### 2019-06-26 paper/paper.tex
paper draft

%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
\documentclass[sigconf,review,anonymous]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em
      b}\kern-0.8em\TeX}}}

\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
\usepackage[caption=false, font=footnotesize]{subfig}
%%%%%%%%%%%%% code listing
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=Java,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=false,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  %breaklines=true,            % Zeilen werden Umgebrochen
  keywords=[2]{class, incorporateUserFeedback, testPushPop, testPopPush},
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

\input{macros}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
%% \acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
%%   Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
%% \acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%   June 03--05, 2018, Woodstock, NY}
%% \acmPrice{15.00}
%% \acmISBN{978-1-4503-9999-9/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page
%% headers.
\title{Synthesis of Intrusion Detection Rules \\ for Increased Network Security}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ben Trovato}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin}
  \state{Ohio}
  \postcode{43017-6221}
}

\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla}
  \country{Iceland}}
\email{larst@affiliation.org}

\affiliation{%
  \institution{Inria Paris-Rocquencourt}
  \city{Rocquencourt}
  \country{France}
}

\author{Aparna Patel}
\affiliation{%
 \institution{Rajiv Gandhi University}
 \streetaddress{Rono-Hills}
 \city{Doimukh}
 \state{Arunachal Pradesh}
 \country{India}}

\author{Huifen Chan}
\affiliation{%
  \institution{Tsinghua University}
  \streetaddress{30 Shuangqing Rd}
  \city{Haidian Qu}
  \state{Beijing Shi}
  \country{China}}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \streetaddress{8600 Datapoint Drive}
  \city{San Antonio}
  \state{Texas}
  \postcode{78229}}
\email{cpalmer@prl.com}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
\Fix{elaborate}
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{datasets, neural networks, gaze detection, text tagging}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
%% \begin{teaserfigure}
%%   \includegraphics[width=\textwidth]{sampleteaser}
%%   \caption{Seattle Mariners at Spring Training, 2010.}
%%   \Description{Enjoying the baseball game from the third-base
%%   seats. Ichiro Suzuki preparing to bat.}
%%   \label{fig:teaser}
%% \end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

Network Intrusion Detection Systems (\nids{}) are an important
instrument that system administrators use to secure networks. These
systems monitor the network traffic for malicious behavior and act
accordingly by blocking messages or alerting humans about suspicious
events. \nids{} are typically placed behind a firewall, vetting the
traffic that the firewall did not block. These systems are very
popular today, with various public (\eg{}, Snort~\Fix{cite} and
Suricata~\Fix{cite}) and private implementations (\eg{}, \Fix{...} and
\Fix{...}). \Fix{...elaborate...}

\sloppy \nids{} can be categorized in two groups: 1)~rule-based and
2)~anomaly-based. A rule-based \nids{} checks if the network traffic
matches a given set of rules. A rule describes a potentially
mallicious message pattern. Figure~\Fix{small example} shows an
example rule of Suricata~\cite{suricata}, a popular open-source
\nids{}, owned and supported by the Open Information Security
Foundation (OISF)~\cite{oisf}. \Fix{quickly explain example}
Rule-based \nids{} focuses on known attacks whereas anomaly-based
\nids{} focuses on unanticipated issues. The hypothesis is that attack
manifestations produces traffic that deviates from regular benign
traffic that has been observed by the network. Anomaly-based \nids{}
have shown useful specially for detecting denial of service attacks
when determining the right threshold for the number of accesses that
triggers an alarm is challenging. Conceptually, these two approaches
are complementary. This paper focuses on rule-based \nids{}, which are
very popular both in academia and industry.

%Despite the great success of rule-based \nids{},

Mantaining \nids{} rules can be a daunting activity. Although a set of
rules are typically provided by the \nids{} (\eg{}, \suri{} provides
\numrulessuri{} rules), 1) enabling all possible rules is not a
practical option and 2) rules not in the original rule set may need to
be added. The first statement is consequential of the inherent
physical limitation of rule-based \nids{}. The cost of checking rules
is proportional to the traffic load and the number of enabled
rules. If system capacity is insufficient to support the load, the
system can potentially miss packets and attacks\Fix{I guess we have
  some reference for this.}. To cope with that issue, system
administrators restrict the attack surface before applying
\nids{}---they block certain protocols on the firewall and configure
the \nids{} to use a selection of rules that apply to the authorized
traffic. The second problem is associated with the observation that
\numundetected{} of the attacks produced by
Metasploit~\cite{metasploit}, a popular open-source penetration
testing tool, went undetected by the \suri{} \nids{}. For example,
\Fix{...}. (Section~\ref{sec:suri-metas-coverage} details the coverage
of \suri{} against the exploits produced by \metas.) Unfortunately,
manually creating those rules can be time-consuming and
error-prone. Figure~\Fix{...} shows an example rule illustrating the
complexity of the Suricata rule language\footnote{Suricata extends the
  language used by Snort~\cite{snort}.}. Collectively, these issues
indicate that synthesizing \nids{} rules is an important aid for
system administrators.

This paper makes the following contributions. \Fix{...}

\section{\suri{} coverage on the \metas{} dataset}
\label{sec:suri-metas-coverage}

\section{Approach}

\Fix{...}

\section{Evaluation}

\Fix{...}

\section{Threats to Validity}

External Validity. \Fix{elaborate ...how popular are Suricata/Metasploit?}

\section{Related Work}

\Fix{...}

%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
%% \begin{acks}
%% To Robert, for the bagels and explaining CMYK and color spaces.
%% \end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\balance \bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2019-06-27 paper/paper.tex
quick pass on text

@@
 Network Intrusion Detection Systems (\nids{}) are an important
-instrument that system administrators use to secure networks. These
-systems monitor the network traffic for malicious behavior and act
-accordingly by blocking messages or alerting humans about suspicious
-events. \nids{} are typically placed behind a firewall, vetting the
-traffic that the firewall did not block. These systems are very
-popular today, with various public (\eg{}, Snort~\Fix{cite} and
-Suricata~\Fix{cite}) and private implementations (\eg{}, \Fix{...} and
-\Fix{...}). \Fix{...elaborate...}
+instrument that system administrators have to secure networks. These
+software systems monitor the network traffic for malicious behavior
+and act accordingly by blocking messages or alerting humans about
+suspicious events. \nids{} are typically placed behind a firewall,
+vetting the traffic that the firewall did not block. \nids{} are
+frequently adopted today. Various open-source (\eg{}, Snort~\Fix{cite}
+and Suricata~\Fix{cite}) and commercial implementations (\eg{},
+\Fix{...}  and \Fix{...}) exist. \Fix{...elaborate...}
 
@@
 example rule of Suricata~\cite{suricata}, a popular open-source
-\nids{}, owned and supported by the Open Information Security
-Foundation (OISF)~\cite{oisf}. \Fix{quickly explain example}
-Rule-based \nids{} focuses on known attacks whereas anomaly-based
-\nids{} focuses on unanticipated issues. The hypothesis is that attack
-manifestations produces traffic that deviates from regular benign
-traffic that has been observed by the network. Anomaly-based \nids{}
-have shown useful specially for detecting denial of service attacks
-when determining the right threshold for the number of accesses that
-triggers an alarm is challenging. Conceptually, these two approaches
+\nids{}, maintained by the Open Information Security Foundation
+(OISF)~\cite{oisf}. \Fix{quickly explain example} Rule-based \nids{}
+focuses on known attacks whereas anomaly-based \nids{} focuses on
+unanticipated issues. The hypothesis of anomaly-based \nids\ is that
+attack manifestations produces traffic that deviates from regular
+benign traffic that has been already observed by the network and that
+deviation is noticeable. Anomaly-based \nids{}, specially those that
+use machine learning, have shown useful specially for detecting denial
+of service attacks~\Fix{cite cite}. In that scenario, determining the
+right threshold on the rate/volume of network accesses that should
+trigger an alarm is challenging. Conceptually, these two approaches
 are complementary. This paper focuses on rule-based \nids{}, which are
-very popular both in academia and industry.
+already very popular both in industry and academia. Our analysis
+focuses on Suricata, but it applies to any other rule-based \nids{}.
 
@@
 
-Mantaining \nids{} rules can be a daunting activity. Although a set of
-rules are typically provided by the \nids{} (\eg{}, \suri{} provides
-\numrulessuri{} rules), 1) enabling all possible rules is not a
-practical option and 2) rules not in the original rule set may need to
-be added. The first statement is consequential of the inherent
-physical limitation of rule-based \nids{}. The cost of checking rules
-is proportional to the traffic load and the number of enabled
-rules. If system capacity is insufficient to support the load, the
-system can potentially miss packets and attacks\Fix{I guess we have
-  some reference for this.}. To cope with that issue, system
-administrators restrict the attack surface before applying
-\nids{}---they block certain protocols on the firewall and configure
-the \nids{} to use a selection of rules that apply to the authorized
-traffic. The second problem is associated with the observation that
-\numundetected{} of the attacks produced by
-Metasploit~\cite{metasploit}, a popular open-source penetration
-testing tool, went undetected by the \suri{} \nids{}. For example,
-\Fix{...}. (Section~\ref{sec:suri-metas-coverage} details the coverage
-of \suri{} against the exploits produced by \metas.) Unfortunately,
-manually creating those rules can be time-consuming and
-error-prone. Figure~\Fix{...} shows an example rule illustrating the
-complexity of the Suricata rule language\footnote{Suricata extends the
-  language used by Snort~\cite{snort}.}. Collectively, these issues
-indicate that synthesizing \nids{} rules is an important aid for
-system administrators.
+Mantaining \nids{} rules can be a daunting activity. Although the
+\nids{} typically provides a set of pre-defined rules, such ruleset is
+neither necessary nor sufficient. More precisely:
+
+\begin{enumerate}
+\item enabling all possible rules is not practical;
+\item rules not in the original rule set may need to be added in a
+  given setup.
+\end{enumerate}
+
+The first statement is consequential of the inherent physical
+limitation of rule-based \nids{}. The cost of checking rules is
+proportional to the traffic load and the number of enabled rules. If
+system capacity is insufficient to support the load, the system can
+potentially miss packets and attacks\Fix{I guess we have some
+  reference for this.}. To cope with that issue, system administrators
+restrict the attack surface before applying \nids{}---they block
+certain protocols on the firewall and configure the \nids{} to use a
+selection of rules that apply to the authorized traffic. The second
+statement is supported by the observation that \numundetected{} of the
+attacks produced by Metasploit~\cite{metasploit}, a popular
+open-source penetration testing tool, went undetected by \suri{}. For
+example, \Fix{...}. (Section~\ref{sec:suri-metas-coverage} provides
+more details.) Unfortunately, manually creating those rules from
+scratch can be time-consuming and error-prone. Figure~\Fix{...}  shows
+an example rule illustrating the complexity of the Suricata rule
+language\footnote{Suricata extends the language used by
+  Snort~\cite{snort}.}.  In addition, looking for rules from untrusted
+sources (say, the internet) is unreliable. Collectively, these issues
+suggest that synthesizing \nids{} rules is an important aid for system
+administrators.
+
+\Fix{why is that challenging (worth of research)?}
 

### 2019-06-27 paper/paper.tex
minor changes in text

@@
 use machine learning, have shown useful specially for detecting denial
-of service attacks~\Fix{cite cite}. In that scenario, determining the
+of service attacks~\cite{tutorial-ml-for-anomaly-detection-dec-2018, cordy-etal-issta19}. In that scenario, determining the
 right threshold on the rate/volume of network accesses that should
@@
 proportional to the traffic load and the number of enabled rules. If
-system capacity is insufficient to support the load, the system can
-potentially miss packets and attacks\Fix{I guess we have some
+system capacity is insufficient to support the load, the intrusion
+detector can miss packets and attacks\Fix{I guess we have some
   reference for this.}. To cope with that issue, system administrators
-restrict the attack surface before applying \nids{}---they block
-certain protocols on the firewall and configure the \nids{} to use a
-selection of rules that apply to the authorized traffic. The second
-statement is supported by the observation that \numundetected{} of the
-attacks produced by Metasploit~\cite{metasploit}, a popular
-open-source penetration testing tool, went undetected by \suri{}. For
-example, \Fix{...}. (Section~\ref{sec:suri-metas-coverage} provides
-more details.) Unfortunately, manually creating those rules from
-scratch can be time-consuming and error-prone. Figure~\Fix{...}  shows
-an example rule illustrating the complexity of the Suricata rule
+block certain protocols on the firewall and configure the \nids{} to
+use a selection of rules that apply to the authorized traffic.  The
+task of selecting rules that are applicable can be non-trivial as it
+depends on the set of firewall rules enabled on the network.
+
+The second statement is supported by the observation that
+\numundetected{} of the attacks produced by
+Metasploit~\cite{metasploit}, a popular open-source penetration
+testing tool, are undetected by \suri{}. For example,
+\Fix{...}. (Section~\ref{sec:suri-metas-coverage} provides more
+details.) Unfortunately, manually creating those rules from scratch
+can be time-consuming and error-prone. Figure~\Fix{...}  shows an
+example rule illustrating the complexity of the Suricata rule
 language\footnote{Suricata extends the language used by

### 2019-09-07 paper/paper.tex
revising manuscript. Lucas/Guilherme, please look at Fix(es) addressed to you

@@
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-\documentclass[sigconf,review,anonymous]{acmart}
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em
-      b}\kern-0.8em\TeX}}}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first
+% footnote. If that is unneeded, please comment it out.
 
 \usepackage{booktabs}
+\usepackage{url}
 \usepackage{multirow}
@@
   keywordstyle=\scriptsize\ttfamily\bfseries,
-  language=Java,             % choose the language of the code
+  language=C,             % choose the language of the code
   frame=single,              % adds a frame around the code
@@
   %breaklines=true,            % Zeilen werden Umgebrochen
-  keywords=[2]{class, incorporateUserFeedback, testPushPop, testPopPush},
+  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
 }
@@
 
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
-\acmDOI{10.1145/1122445.1122456}
+%% \usepackage{cite}
+%% \usepackage{amsmath,amssymb,amsfonts}
+%% \usepackage{algorithmic}
+%% \usepackage{graphicx}
+%% \usepackage{textcomp}
+%% \usepackage{xcolor}
 
-%% These commands are for a PROCEEDINGS abstract or paper.
-%% \acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
-%%   Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
-%% \acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%   June 03--05, 2018, Woodstock, NY}
-%% \acmPrice{15.00}
-%% \acmISBN{978-1-4503-9999-9/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-%%
-%% end of the preamble, start of the body of the document source.
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 \begin{document}
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page
-%% headers.
-\title{Synthesis of Intrusion Detection Rules \\ for Increased Network Security}
+\title{Search-based Synthesis of Network Intrusion Detection Rules}
 
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-\author{Ben Trovato}
-\authornote{Both authors contributed equally to this research.}
-\email{trovato@corporation.com}
-\orcid{1234-5678-9012}
-\author{G.K.M. Tobin}
-\authornotemark[1]
-\email{webmaster@marysville-ohio.com}
-\affiliation{%
-  \institution{Institute for Clarity in Documentation}
-  \streetaddress{P.O. Box 1212}
-  \city{Dublin}
-  \state{Ohio}
-  \postcode{43017-6221}
-}
+%%%%%%%%%%% Anonymized
 
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{larst@affiliation.org}
+%% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address}
+%% \and
+%% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address}
+%% \and
+%% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address}
+%% \and
+%% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address}
+%% \and
+%% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address}
+%% \and
+%% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address}
+%% }
 
-\affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
-}
+\maketitle
 
-\author{Aparna Patel}
-\affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh}
- \state{Arunachal Pradesh}
- \country{India}}
-
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu}
-  \state{Beijing Shi}
-  \country{China}}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas}
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
 \begin{abstract}
-\Fix{elaborate}
+This document is a model and instructions for \LaTeX.
+This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
+or Math in Paper Title or Abstract.
 \end{abstract}
 
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{datasets, neural networks, gaze detection, text tagging}
-
-%% A "teaser" image appears between the author and affiliation
-%% information and the body of the document, and typically spans the
-%% page.
-%% \begin{teaserfigure}
-%%   \includegraphics[width=\textwidth]{sampleteaser}
-%%   \caption{Seattle Mariners at Spring Training, 2010.}
-%%   \Description{Enjoying the baseball game from the third-base
-%%   seats. Ichiro Suzuki preparing to bat.}
-%%   \label{fig:teaser}
-%% \end{teaserfigure}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-\maketitle
+\begin{IEEEkeywords}
+component, formatting, style, styling, insert
+\end{IEEEkeywords}
 
@@
 
-Network Intrusion Detection Systems (\nids{}) are an important
-instrument that system administrators have to secure networks. These
-software systems monitor the network traffic for malicious behavior
-and act accordingly by blocking messages or alerting humans about
-suspicious events. \nids{} are typically placed behind a firewall,
-vetting the traffic that the firewall did not block. \nids{} are
-frequently adopted today. Various open-source (\eg{}, Snort~\Fix{cite}
-and Suricata~\Fix{cite}) and commercial implementations (\eg{},
-\Fix{...}  and \Fix{...}) exist. \Fix{...elaborate...}
+Network Intrusion Detection Systems (\nids{}) are software systems
+that monitor the network traffic for malicious behavior and act
+accordingly by blocking messages or alerting humans about suspicious
+events~\cite{Mitchell:2014:SID:2597757.2542049}. \nids{} are typically
+placed behind a firewall, vetting the traffic that the firewall did
+not block. Various open-source (\eg{}, Snort~\cite{snort} and
+Suricata~\cite{suricata}) and commercial implementations (\eg{},
+SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar}) exist
+today. These systems are very popular in industry to secure local
+computer networks given the amount of potential malicious traffic that
+exist on the Internet today.
 
-\sloppy \nids{} can be categorized in two groups: 1)~rule-based and
-2)~anomaly-based. A rule-based \nids{} checks if the network traffic
-matches a given set of rules. A rule describes a potentially
-mallicious message pattern. Figure~\Fix{small example} shows an
-example rule of Suricata~\cite{suricata}, a popular open-source
-\nids{}, maintained by the Open Information Security Foundation
-(OISF)~\cite{oisf}. \Fix{quickly explain example} Rule-based \nids{}
-focuses on known attacks whereas anomaly-based \nids{} focuses on
-unanticipated issues. The hypothesis of anomaly-based \nids\ is that
-attack manifestations produces traffic that deviates from regular
-benign traffic that has been already observed by the network and that
-deviation is noticeable. Anomaly-based \nids{}, specially those that
-use machine learning, have shown useful specially for detecting denial
-of service attacks~\cite{tutorial-ml-for-anomaly-detection-dec-2018, cordy-etal-issta19}. In that scenario, determining the
-right threshold on the rate/volume of network accesses that should
-trigger an alarm is challenging. Conceptually, these two approaches
-are complementary. This paper focuses on rule-based \nids{}, which are
-already very popular both in industry and academia. Our analysis
-focuses on Suricata, but it applies to any other rule-based \nids{}.
+%% The
+%% hypothesis of anomaly-based \nids\ is that attack manifestations
+%% produces traffic that deviates from regular benign traffic that has
+%% been already observed by the network and that deviation is
+%% noticeable.
+
+\sloppy \nids{} are typically categorized in two
+groups~\cite{kumar2007survey}: 1)~rule-based and 2)~anomaly-based. A
+rule-based \nids{} (\aka\ signature-based NIDS) checks if the network
+traffic matches a fixed set of
+rules. Figure~\ref{fig:synflood-example} shows an example rule of
+Suricata~\cite{suricata}, a popular open-source \nids{}. This rule
+prescribes a method to capture SYN flood
+attacks~\cite{Douligeris:2004:DAD:987153.987158} that can result in
+denial of service of a server. Rule-based \nids{} focuses on known
+attacks whereas anomaly-based \nids{} focuses on suspicious
+unanticipated network traffic, referred to as anomalies. Anomaly-based
+\nids{}, specially those that use machine learning, have shown useful
+specially for detecting denial of service
+attacks~\cite{cordy-etal-issta19}. In that scenario, determining the right
+threshold on the rate/volume of network accesses that should trigger
+an alarm is challenging.
+
+Rule-based and Anomaly-based NIDS are complementary. This paper
+proposes a synthesis technique, dubbed \tname{}, to automatically
+create rules from positive and negative examples (\ie{}, benign and
+malicious traffic) provided as input. The main scenario of application
+is one where anomaly-based NIDS identify abnormal traffic and signals
+that to \tname{} to craft rules that can be used by rule-based
+NIDS. It is worth noting that attackers are very productive in
+creating ways to circumvent existing protections and manual creation
+of rules is tedious and time-consuming.\Mar{Lucas/Guilherme: (IMP) supporting data?}
+Consequently, synthesis of rules can circumvent those issues.
+
+\Mar{--------------------------------- parei aqui}
+
+focuses on rule-based \nids{}, which are already very
+popular both in industry and academia. Our analysis focuses on
+Suricata, but it applies to any other rule-based \nids{}.
 
@@
 
-\section{\suri{} coverage on the \metas{} dataset}
+\section{Illustrative Example}
 \label{sec:suri-metas-coverage}
+
+Figure~\ref{fig:synflood-example} shows an
+example rule of Suricata~\cite{suricata}, a popular open-source
+\nids{}, maintained by the Open Information Security Foundation
+(OISF)~\cite{oisf}. This example shows how Suricata can be used to
+protect against SYN flood attacks. A TCP connection is typically
+established with a 3-way handshake. First, the client sends a SYN
+packet to the server, requesting a connection. Second, the server
+responds with SYN-ACK packet to the client. Third, the client responds
+with an ACK message and the connection is established. Aware of the
+protocol, an attacker sends multiple SYN packets to different ports of
+a server, often using fake IP addresses. Without proper protection,
+the server accepts the connection requests and eventually legitimate
+connection requests cannot be satisfied due to resource exhaustion.
+
+\begin{figure}[t]
+  \lstinputlisting[language=C,numbers=none]{synflood.suricata}
+  \caption{Suricata rule for SYN Flood Attacks.\Mar{colocar
+      propriedades em negrito}}
+  \label{fig:synflood-example}
+\end{figure}
+
+\Mar{Lucas/Guilherme (CRI): por favor, explique o formato de uma regra do
+  Suricata e, em seguida, explique esta regra, detalhando o
+  significado de cada propriedade/campo.}
 
@@
 
-\Fix{...}
+\Mar{Lucas/Guilherme (CRI): vcs. precisam dar pontape inicial (bom
+  exercicio) propondo perguntas de pesquisa. Eu faco ajustes em seguida.}
 
@@
 
-%% The acknowledgments section is defined using the "acks" environment
-%% (and NOT an unnumbered section). This ensures the proper
-%% identification of the section in the article metadata, and the
-%% consistent spelling of the heading.
-%% \begin{acks}
-%% To Robert, for the bagels and explaining CMYK and color spaces.
-%% \end{acks}
-
-%%
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
-\balance \bibliographystyle{ACM-Reference-Format}
+\balance
+%\vspace*{-2\baselineskip}
+\bibliographystyle{IEEEtran}
 \bibliography{references}
@@
 \end{document}
-\endinput
-%%
-%% End of file `sample-sigconf.tex'.
+
+%%  LocalWords:  Suricata SolarWinds QRadar NIDS OISF TCP ACK

### 2019-09-07 paper/paper.tex
minor comments

@@
 \usepackage{booktabs}
+\usepackage{cite}
 \usepackage{url}
@@
 \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+...
 \end{abstract}
@@
 \begin{IEEEkeywords}
-component, formatting, style, styling, insert
+NIDS, synthesis, search
 \end{IEEEkeywords}
@@
 
-%% The
-%% hypothesis of anomaly-based \nids\ is that attack manifestations
-%% produces traffic that deviates from regular benign traffic that has
-%% been already observed by the network and that deviation is
-%% noticeable.
-
 \sloppy \nids{} are typically categorized in two
@@
 Suricata~\cite{suricata}, a popular open-source \nids{}. This rule
-prescribes a method to capture SYN flood
-attacks~\cite{Douligeris:2004:DAD:987153.987158} that can result in
-denial of service of a server. Rule-based \nids{} focuses on known
-attacks whereas anomaly-based \nids{} focuses on suspicious
-unanticipated network traffic, referred to as anomalies. Anomaly-based
-\nids{}, specially those that use machine learning, have shown useful
-specially for detecting denial of service
-attacks~\cite{cordy-etal-issta19}. In that scenario, determining the right
-threshold on the rate/volume of network accesses that should trigger
-an alarm is challenging.
+prescribes a method to capture SYN flood attacks to a
+server~\cite{Douligeris:2004:DAD:987153.987158} that can result in
+denial of service. Rule-based \nids{} focuses on known attacks whereas
+anomaly-based \nids{} focuses on unanticipated attacks that could be
+observed through suspicious network
+traffic~\cite{kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}.
 
 Rule-based and Anomaly-based NIDS are complementary. This paper
-proposes a synthesis technique, dubbed \tname{}, to automatically
-create rules from positive and negative examples (\ie{}, benign and
-malicious traffic) provided as input. The main scenario of application
-is one where anomaly-based NIDS identify abnormal traffic and signals
-that to \tname{} to craft rules that can be used by rule-based
-NIDS. It is worth noting that attackers are very productive in
-creating ways to circumvent existing protections and manual creation
-of rules is tedious and time-consuming.\Mar{Lucas/Guilherme: (IMP) supporting data?}
-Consequently, synthesis of rules can circumvent those issues.
+proposes a technique, dubbed \tname{}, to automatically synthesize
+rules from positive and negative examples (\ie{}, benign and malicious
+traffic). The main scenario of application is one where an
+anomaly-based NIDS identifies abnormal traffic and signals that
+traffic to \tname{} to craft rules that can be used by rule-based
+NIDS. The core motivation is that attackers are productive in creating
+new ways to circumvent existing protections and manual creation of
+rules is tedious and time-consuming.\Mar{Lucas/Guilherme: (IMP) do we
+  have data to support this?} For those reasons, synthesizing rules to
+circumvent those issues is an important problem.
 
 \Mar{--------------------------------- parei aqui}
-
-focuses on rule-based \nids{}, which are already very
-popular both in industry and academia. Our analysis focuses on
-Suricata, but it applies to any other rule-based \nids{}.
-
-%Despite the great success of rule-based \nids{},
-
-Mantaining \nids{} rules can be a daunting activity. Although the
-\nids{} typically provides a set of pre-defined rules, such ruleset is
-neither necessary nor sufficient. More precisely:
-
-\begin{enumerate}
-\item enabling all possible rules is not practical;
-\item rules not in the original rule set may need to be added in a
-  given setup.
-\end{enumerate}
-
-The first statement is consequential of the inherent physical
-limitation of rule-based \nids{}. The cost of checking rules is
-proportional to the traffic load and the number of enabled rules. If
-system capacity is insufficient to support the load, the intrusion
-detector can miss packets and attacks\Fix{I guess we have some
-  reference for this.}. To cope with that issue, system administrators
-block certain protocols on the firewall and configure the \nids{} to
-use a selection of rules that apply to the authorized traffic.  The
-task of selecting rules that are applicable can be non-trivial as it
-depends on the set of firewall rules enabled on the network.
-
-The second statement is supported by the observation that
-\numundetected{} of the attacks produced by
-Metasploit~\cite{metasploit}, a popular open-source penetration
-testing tool, are undetected by \suri{}. For example,
-\Fix{...}. (Section~\ref{sec:suri-metas-coverage} provides more
-details.) Unfortunately, manually creating those rules from scratch
-can be time-consuming and error-prone. Figure~\Fix{...}  shows an
-example rule illustrating the complexity of the Suricata rule
-language\footnote{Suricata extends the language used by
-  Snort~\cite{snort}.}.  In addition, looking for rules from untrusted
-sources (say, the internet) is unreliable. Collectively, these issues
-suggest that synthesizing \nids{} rules is an important aid for system
-administrators.
-
-\Fix{why is that challenging (worth of research)?}
 

### 2019-09-07 paper/paper.tex
pass on text

@@
 
+This section 1) presents an example of a network attack, 2) shows an
+example rule of Suricata to detect this attack and alert system
+administrators about it, and 3) illustrates how \tname{} performs to
+synthesize the corresponding rule.
+
+Let us recall that the 3-way handshake protocol in TCP/IP to establish
+a TCP connection. First, a client sends a SYN packet to the server,
+requesting a connection. Second, the server responds with SYN-ACK
+packet to the client. Third, the client responds with an ACK message
+and the connection is established.
+
 Figure~\ref{fig:synflood-example} shows an
@@
 (OISF)~\cite{oisf}. This example shows how Suricata can be used to
-protect against SYN flood attacks. A TCP connection is typically
-established with a 3-way handshake. First, the client sends a SYN
-packet to the server, requesting a connection. Second, the server
-responds with SYN-ACK packet to the client. Third, the client responds
-with an ACK message and the connection is established. Aware of the
+protect against SYN flood attacks.\Fix{explain the structure of the rule.}
+
+Aware of the
 protocol, an attacker sends multiple SYN packets to different ports of

### 2019-09-07 paper/paper.tex
adding related presentations

@@
 
-\Mar{--------------------------------- parei aqui}
+\Fix{summarize how the technique works}
+
+\Fix{summarize results}
 
@@
 This section 1) presents an example of a network attack, 2) shows an
-example rule of Suricata to detect this attack and alert system
-administrators about it, and 3) illustrates how \tname{} performs to
-synthesize the corresponding rule.
+example rule of Suricata to detect this attack and alert sys admins
+about it, and 3) illustrates how \tname{} performs to synthesize the
+corresponding rule.
 

### 2019-09-07 paper/paper.tex
pass on example

@@
 
+\Mar{Lucas/Guilherme, verifique se isto esta correto}
+Figure~\ref{fig:synflood-example} shows an example rule of
+Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
+the Open Information Security Foundation (OISF)~\cite{oisf}. This
+example shows how Suricata can be used to protect organizations
+against one kind of attack. The rule is divided in three
+parts~\cite{suri-rule-format}--action, header, and rule options. The
+action part always appears as the first word in the rule. In this
+case, \CodeIn{alert}. An action denotes the task that needs to be
+executed if the rule pattern is satisfied. So, in this case, a message
+will be sent to sys admins if the rule pattern is satisfied. The
+header comes after the action in the rule description. It restricts
+the information flow covered by the rule. For this rule, the header is
+\CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any} indicating that
+the rule inspects \CodeIn{tcp} traffic flowing from any port in the
+home network to any other address outside the home network. The rule
+options come after the header. It is a semi-colon-separated sequence
+of key-value pairs characterizing the pattern of attack. For example,
+\Mar{Lucas/Guilherme (IMP): explique o que cada propriedade faz...}
+
+\Mar{------------------------------ parei aqui}
+
 Let us recall that the 3-way handshake protocol in TCP/IP to establish
@@
 and the connection is established.
-
-Figure~\ref{fig:synflood-example} shows an
-example rule of Suricata~\cite{suricata}, a popular open-source
-\nids{}, maintained by the Open Information Security Foundation
-(OISF)~\cite{oisf}. This example shows how Suricata can be used to
-protect against SYN flood attacks.\Fix{explain the structure of the rule.}
 
@@
 
-\Mar{Lucas/Guilherme (CRI): por favor, explique o formato de uma regra do
-  Suricata e, em seguida, explique esta regra, detalhando o
-  significado de cada propriedade/campo.}
-
 \section{Approach}
@@
 
-\Mar{Lucas/Guilherme (CRI): vcs. precisam dar pontape inicial (bom
-  exercicio) propondo perguntas de pesquisa. Eu faco ajustes em seguida.}
+\Mar{Lucas/Guilherme (CRI): precisamos dar pontape inicial propondo
+  perguntas de pesquisa. Eu faco ajustes em seguida.}
 

### 2019-09-08 paper/paper.tex
another pass on example

@@
 
-This section 1) presents an example of a network attack, 2) shows an
-example rule of Suricata to detect this attack and alert sys admins
-about it, and 3) illustrates how \tname{} performs to synthesize the
-corresponding rule.
+This section 1) shows an example rule of Suricata and briefly explains
+the format of Suricata rules, 2) presents an example of
+a network attack, and 3) illustrates how \tname{} performs to
+synthesize the rule that would capture the attack.
+
+\subsection{Suricata Rules}
 
@@
 Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
-the Open Information Security Foundation (OISF)~\cite{oisf}. This
-example shows how Suricata can be used to protect organizations
-against one kind of attack. The rule is divided in three
-parts~\cite{suri-rule-format}--action, header, and rule options. The
-action part always appears as the first word in the rule. In this
+the Open Information Security Foundation (OISF)~\cite{oisf}. A
+Suricata rule is divided in three
+parts---action, header, and rule options~\cite{suri-rule-format}. The
+action part appears as the first word in the rule description. In this
 case, \CodeIn{alert}. An action denotes the task that needs to be
-executed if the rule pattern is satisfied. So, in this case, a message
-will be sent to sys admins if the rule pattern is satisfied. The
+executed if the rule pattern is satisfied. In this case, a message
+will be sent to sys admins if the rule pattern is satisfied. Other
+options of actions include pass, drop, and reject. The
 header comes after the action in the rule description. It restricts
 the information flow covered by the rule. For this rule, the header is
-\CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any} indicating that
-the rule inspects \CodeIn{tcp} traffic flowing from any port in the
-home network to any other address outside the home network. The rule
+\CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any}, which instructs Suricata
+to inspect \CodeIn{tcp} traffic flowing from any port in the
+home network to any other address outside the home network. The
+variables \CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are configurable. The rule
 options come after the header. It is a semi-colon-separated sequence
-of key-value pairs characterizing the pattern of attack. For example,
+of key-value pairs characterizing the attack pattern. For example,
 \Mar{Lucas/Guilherme (IMP): explique o que cada propriedade faz...}
 
-\Mar{------------------------------ parei aqui}
+\subsection{SYN Flood Attack}
 
-Let us recall that the 3-way handshake protocol in TCP/IP to establish
-a TCP connection. First, a client sends a SYN packet to the server,
+The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
+to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
+works as follows. First, a client sends a SYN packet to the server,
 requesting a connection. Second, the server responds with SYN-ACK
@@
 \end{figure}
+
+\subsection{\tname\ on SYN Flood Attack}
+
+\Mar{Guilherme/Lucas (CRI): we need to show how \tname{} works on this
+  example. What is the input? How long does it take to generate the
+  rule? Is the rule generate exactly the same?}
 

### 2019-09-14 paper/paper.tex
pass on technique

@@
 the Open Information Security Foundation (OISF)~\cite{oisf}. A
-Suricata rule is divided in three
-parts---action, header, and rule options~\cite{suri-rule-format}. The
-action part appears as the first word in the rule description. In this
-case, \CodeIn{alert}. An action denotes the task that needs to be
-executed if the rule pattern is satisfied. In this case, a message
-will be sent to sys admins if the rule pattern is satisfied. Other
-options of actions include pass, drop, and reject. The
-header comes after the action in the rule description. It restricts
-the information flow covered by the rule. For this rule, the header is
-\CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any}, which instructs Suricata
-to inspect \CodeIn{tcp} traffic flowing from any port in the
-home network to any other address outside the home network. The
-variables \CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are configurable. The rule
-options come after the header. It is a semi-colon-separated sequence
-of key-value pairs characterizing the attack pattern. For example,
-\Mar{Lucas/Guilherme (IMP): explique o que cada propriedade faz...}
+Suricata rule is divided in three parts---action, header, and rule
+options~\cite{suri-rule-format}. The action part appears as the first
+word in the rule description. In this case, \CodeIn{alert}. An action
+denotes the task that needs to be executed if the rule pattern is
+satisfied. In this case, a message will be sent to sys admins if the
+rule pattern is satisfied. Other options of actions include pass,
+drop, and reject. The header comes after the action in the rule
+description. It restricts the information flow covered by the
+rule. For this rule, the header is \CodeIn{tcp \$HOME\_NET any ->
+  \$EXTERNAL\_NET any}, which instructs Suricata to inspect
+\CodeIn{tcp} traffic flowing from any port in the home network to any
+other address outside the home network. The variables
+\CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are
+configurable. The rule options come after the header. It is a
+semi-colon-separated sequence of properties\Mar{should we call it an
+  option or property?}, where each property is a key-value pair
+describing a distinct characteristic of the attack.
 
@@
 works as follows. First, a client sends a SYN packet to the server,
-requesting a connection. Second, the server responds with SYN-ACK
+requesting a connection. Second, the server responds with a SYN-ACK
 packet to the client. Third, the client responds with an ACK message
-and the connection is established.
+and the connection is established. Aware of the protocol, an attacker
+sends multiple SYN packets to different ports of a server, often using
+fake IP addresses. Without proper protection, the server accepts the
+connection requests and eventually legitimate connection requests
+cannot be satisfied due to resource exhaustion.
 
-Aware of the
-protocol, an attacker sends multiple SYN packets to different ports of
-a server, often using fake IP addresses. Without proper protection,
-the server accepts the connection requests and eventually legitimate
-connection requests cannot be satisfied due to resource exhaustion.
+Figure~\ref{fig:synflood-example} shows a Suricata rule to capture
+this attack. The critical parts are the option \CodeIn{flags: S,12},
+which identifies a SYN packet and the option \CodeIn{threshold: type
+  both, track by\_dst, count 5000, seconds 5}, indicating that a high
+volume of such packets should be requested in a short period of time.
 
@@
 
-\Fix{...}
+\tname{} is a search-based algorithm that uses positive and negative
+examples to synthesize signature-based rules, such as those from Snort
+and Suricata. \tname{} synthesizes rules in two steps. First, it
+produces an overspecified rule to capture the negative example
+provided as input. Then, it uses positive examples to eliminate
+unecessary options from the rule. The goal is to create a rule that
+captures only the negative traffic.
+
+\subsection{Problem Characterization}
+
+\subsection{Fitness Functions}
+
+\subsection{Search}
 

### 2019-09-14 paper/paper.tex
pass on text

@@
 \subsection{Suricata Rules}
+\label{sec:example-suricata-rules}
 
@@
 \tname{} is a search-based algorithm that uses positive and negative
-examples to synthesize signature-based rules, such as those from Snort
-and Suricata. \tname{} synthesizes rules in two steps. First, it
-produces an overspecified rule to capture the negative example
-provided as input. Then, it uses positive examples to eliminate
-unecessary options from the rule. The goal is to create a rule that
-captures only the negative traffic.
+examples to synthesize signature-based NIDS rules, such as those from
+Snort and Suricata. \tname{} synthesizes rules in two steps. First, it
+produces a potentially overspecified rule that captures the negative
+example provided as input. Then, it uses positive examples to
+eliminate unecessary options from the rule. The goal is to create a
+rule that captures only the negative traffic.
 
 \subsection{Problem Characterization}
+
+\tname{} uses an abstract representations of a Suricata rule that
+reflects the concrete representation of the rule. The concrete rule
+representation is the one actually used by Suricata whereas the
+abstract representation is the one \tname{} uses to run the
+search. The abstract representation encodes the rule options segment
+(see Section~\ref{sec:example-suricata-rules}) of a Suricata rule as a
+vector of Option objects. For example, \tname{} models the option
+\CodeIn{flags} from the example on Figure~\ref{fig:synflood-example},
+with a data structure storing two fields, one of character type and
+another of integer type. This enables \tname{} to mutate each element
+separately.\Mar{Lucas, por favor, confirme.}
 
@@
 \subsection{Search}
+
+\tname{} is a single-individual hill-climbing like search.
 

### 2019-09-14 paper/paper.tex
pass

--- 2019-09-14 paper/paper.tex
@@
 
-\tname{} is a search-based algorithm that uses positive and negative
-examples to synthesize signature-based rules, such as those from Snort
-and Suricata. \tname{} synthesizes rules in two steps. First, it
-produces an overspecified rule to capture the negative example
-provided as input. Then, it uses positive examples to eliminate
-unecessary options from the rule. The goal is to create a rule that
-captures only the negative traffic.
+\tname{} is a search-based approach that uses positive and negative
+network traffic examples to synthesize signature-based NIDS rules,
+such as those from Snort and Suricata. \tname{} synthesizes rules in
+two steps. First, it produces a potentially overspecified rule that
+captures the negative example that is provided as input. Then, it uses
+positive examples to eliminate unecessary options from the rule. The
+goal is to create a rule that captures only the negative traffic.
 
@@
 
+Search-based algorithms need to characterize the problem (\ie{}, a
+solution candidate) in some way. In this case, the problem is the
+rule.  \tname\ encodes the rule options segment (see
+Section~\ref{sec:example-suricata-rules}) of a Suricata rule as a
+vector of Option objects. For example, \tname{} models the option
+\CodeIn{flags} from the example on Figure~\ref{fig:synflood-example}
+with a data structure storing two fields, one of character type and
+another of integer type. This enables \tname{} to mutate each element
+separately.\Mar{Lucas, por favor, confirme.}
+
+%% \tname{}
+%% uses an abstract representation of a Suricata rule that reflects the
+%% concrete representation of the rule. The concrete rule representation
+%% is the one actually used by Suricata whereas the abstract
+%% representation is the one \tname{} uses to run the search.
+
 \subsection{Fitness Functions}
 
+The fitness value of a candidate solution $c$, modelling the rule
+options $o_1, \dots, o_n$ is given by $f(c)=1-\Sigma{d(o_i,m)}/n$,
+where $i$ ranges from 1 to $n$ and $m$ denotes the negative
+traffic. The higher the fitness value the better, \ie{}, the fitter
+the candidate solution $c$ will be.  The value of the expression
+$d(o_i,m)$, associated with option $o_i$, ranges over
+0-1. Consequently, the value of the expression $\Sigma{d(o_i,m)}/n$
+itself is in the 0-1 range, as there are $n$ options, and the range of
+the function is 0-1.
+
+As usual, we used customized distance functions to compute $d(o_i,
+m)$. For example, let us consider the \CodeIn{flags} option, discussed
+on Figure~\ref{fig:synflood-example}. Recall that this option
+identifies the SYN packet in the tcp protocol and that its encoding in
+\tname{} stores two attributes---one categorical and one
+numerical. The value of $d(o_i, m)$, in this case, is obtained by
+computing the normalized Euclidian distance in the two-dimensional
+space. Considering that each of the two dimensions has cardinality of
+ten, the distance of a candidate solution holding the option
+\CodeIn{flags(S,13)} to the solution will be 0.22
+(=$1/\sqrt{20}$). Note that the smaller the distance the better.
+
+\tname{} uses a modified version of Suricata to measure distance. It
+updates the rule set that the modified version of Suricata analyses,
+injects the traffic on the network, and monitors the execution. Each
+rule option has an associated \CodeIn{eval} function in code. \tname{}
+monitors the execution of these functions to identify how far it is
+from returning true, which indicates it is closer to trigger the
+action associated with the rule. For the example above,
+\Fix{Lucas/Guilherme, it would be good to show how this looks like in
+  code}
+
 \subsection{Search}
+
+\tname{} is a single-individual hill-climbing like search.
 

### 2019-09-15 paper/paper.tex
pass

@@
 \usepackage{multirow}
+\usepackage{pgfplots}
 \usepackage{tikz}
@@
 
-Search-based algorithms need to characterize the problem (\ie{}, a
-solution candidate) in some way. In this case, the problem is the
-rule.  \tname\ encodes the rule options segment (see
-Section~\ref{sec:example-suricata-rules}) of a Suricata rule as a
-vector of Option objects. For example, \tname{} models the option
-\CodeIn{flags} from the example on Figure~\ref{fig:synflood-example}
-with a data structure storing two fields, one of character type and
-another of integer type. This enables \tname{} to mutate each element
-separately.\Mar{Lucas, por favor, confirme.}
+One important step in search-based optimization is to characterize a
+candidate solution to the target problem. Candidate solutions are the
+objects that the search tries to optimize. In this case, a rule is a
+candidate solution.  \tname\ encodes the rule options segment of a
+Suricata rule (see Section~\ref{sec:example-suricata-rules}) as a
+vector of Option objects. Some options are discarded as they only
+serve for documentation. For example, the options \CodeIn{msg},
+\CodeIn{classtype}, and \CodeIn{miscactivity} have no influence in
+triggering the action of rules. The option \CodeIn{flags}, appearing
+in the example on Figure~\ref{fig:synflood-example}, is relevant to
+trigger the action. \tname{} models that option with a data structure
+storing two fields, one of character type and another of integer
+type. This enables \tname{} to mutate each element
+separately.
+
+Options can be specific to a given protocol or general and they can be
+influential on triggering an action or
+not. Figure~\ref{fig:distribution-rules-protocol} shows the
+distributions of these options per protocol and their relevance for
+triggering rule actions.
+
+\pgfplotsset{width=6cm,compat=1.8}
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+        ybar stacked,
+        enlargelimits=0.15,
+        legend style={at={(0.5,-0.15)},
+          anchor=north,legend columns=-1},
+        ylabel={\#rules},
+        symbolic x coords={gen, http, icmp, udp, tcp},
+        xtick=data,
+        nodes near coords,
+        nodes near coords align={vertical},
+      ]
+      \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
+      \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
+      \legend{relevant, irrelevant}
+    \end{axis}
+  \end{tikzpicture}
+  \caption{\label{fig:distribution-rules-protocol}Distribution of
+    rules per protocol and relevance for triggering action.}
+\end{figure}
+
+To bootstrap the search, \tname{} needs to select an initial
+solution. When creating such candidate solution, \tname{} first looks
+for the protocol related to the malicious input traffic. Then, it
+selects the relevant options for that protocol and initializes the
+option vector with default values. \Mar{Suricata define valores
+  default?  Vcs. definiram?  Isto foi baseado em popularidade ou o
+  que?}
 

### 2019-09-16 paper/paper.tex
pass on technique

@@
 captures the negative example that is provided as input. Then, it uses
-positive examples to eliminate unecessary options from the rule. The
-goal is to create a rule that captures only the negative traffic.
+positive examples to eliminate unnecessary options from the rule. The
+goal of \tname\ is to create a rule that captures only the negative traffic.
 
@@
 candidate solution to the target problem. Candidate solutions are the
-objects that the search tries to optimize. In this case, a rule is a
-candidate solution.  \tname\ encodes the rule options segment of a
-Suricata rule (see Section~\ref{sec:example-suricata-rules}) as a
-vector of Option objects. Some options are discarded as they only
-serve for documentation. For example, the options \CodeIn{msg},
-\CodeIn{classtype}, and \CodeIn{miscactivity} have no influence in
-triggering the action of rules. The option \CodeIn{flags}, appearing
-in the example on Figure~\ref{fig:synflood-example}, is relevant to
-trigger the action. \tname{} models that option with a data structure
-storing two fields, one of character type and another of integer
-type. This enables \tname{} to mutate each element
-separately.
+objects that the search optimize towards solutions. In this case, a a
+candidate solution is an encoding of a Suricata rule.  \tname\ encodes
+the rule options segment of a Suricata rule (see
+Section~\ref{sec:example-suricata-rules}) as a vector of Option
+objects. Some options are discarded as they only serve for
+documentation. For example, the options \CodeIn{msg},
+\CodeIn{classtype}, and \CodeIn{misc-activity}, appearing in the
+example on Figure~\ref{fig:synflood-example}, have no influence in
+triggering the actions associated with rules. The option
+\CodeIn{flags}, however, is relevant to trigger the action. \tname{}
+models that option with a data structure storing two fields, one of
+character type and another of integer type. This enables \tname{} to
+mutate each element separately. Similar encodings are used in modeling
+the other relevant options. We found what is (ir)relevant by
+inspecting Suricata code for the presence of \CodeIn{eval} functions,
+which are invoked from a main look to process every Suricata rule. The
+presence of such a function for a given option indicates that it is
+relevant for triggering some action of some rule.
 
@@
 
-\pgfplotsset{width=6cm,compat=1.8}
+\pgfplotsset{width=5.5cm,compat=1.8}
 \begin{figure}[h!]
@@
 
+\subsection{Initial State}
+
 To bootstrap the search, \tname{} needs to select an initial
@@
 
-The fitness value of a candidate solution $c$, modelling the rule
+The fitness value of a candidate solution $c$, modeling the rule
 options $o_1, \dots, o_n$ is given by $f(c)=1-\Sigma{d(o_i,m)}/n$,
@@
 numerical. The value of $d(o_i, m)$, in this case, is obtained by
-computing the normalized Euclidian distance in the two-dimensional
+computing the normalized Euclidean distance in the two-dimensional
 space. Considering that each of the two dimensions has cardinality of
@@
 
-%%  LocalWords:  Suricata SolarWinds QRadar NIDS OISF TCP ACK
+%%  LocalWords:  Suricata SolarWinds QRadar NIDS OISF TCP ACK msg tcp
+%%  LocalWords:  encodings eval cardinality

### 2019-09-23 paper/paper.tex
pass on intro

@@
 Suricata~\cite{suricata}, a popular open-source \nids{}. This rule
-prescribes a method to capture SYN flood attacks to a
-server~\cite{Douligeris:2004:DAD:987153.987158} that can result in
-denial of service. Rule-based \nids{} focuses on known attacks whereas
+prescribes a method to capture a denial-of-service attack to a
+server~\cite{Douligeris:2004:DAD:987153.987158}. Rule-based \nids{} focuses on known attacks whereas
 anomaly-based \nids{} focuses on unanticipated attacks that could be
@@
 
-Rule-based and Anomaly-based NIDS are complementary. This paper
-proposes a technique, dubbed \tname{}, to automatically synthesize
-rules from positive and negative examples (\ie{}, benign and malicious
-traffic). The main scenario of application is one where an
-anomaly-based NIDS identifies abnormal traffic and signals that
+These intrusion detection mechanisms are complementary. Rule-based
+NIDS are very popular in industry (it is clear to system
+administrators what rules will do), but attacker are creating in
+evading the current set of rules, and writing these rules is tedious
+and error-prone. Anomaly-based NIDS are not restricted to a fixed set
+of rules; they learn common behaviors and dynamically detect uncommon
+behavior. The often use some machine learning algorithm to model
+sequence of events~\cite{7579764}. Unfortunately, anomaly-based NIDS
+can raise false alarms as behavior models are fundamentally
+imprecise. To sum up, both approaches have positive and negative
+aspects.
+
+%% This paper focuses on the
+%% problem of discovering new rules by using anomaly-based NIDS.
+
+This paper proposes a technique, dubbed \tname{}, to automatically
+synthesize rules from positive and negative examples (\ie{}, benign
+and malicious traffic). The main scenario of application is one where
+an anomaly-based NIDS identifies abnormal traffic and signals that
 traffic to \tname{} to craft rules that can be used by rule-based
-NIDS. The core motivation is that attackers are productive in creating
-new ways to circumvent existing protections and manual creation of
-rules is tedious and time-consuming.\Mar{Lucas/Guilherme: (IMP) do we
-  have data to support this?} For those reasons, synthesizing rules to
-circumvent those issues is an important problem.
+NIDS. Those rules can be distributed to rule databases. The core
+motivation is that 1) attackers are productive in creating new ways to
+circumvent existing protections and 2) manual creation of rules is
+tedious and time-consuming.  Note that anomaly-based NIDS are evolving
+pretty quick with the advances in machine learning, but rule-based
+NIDS are still extremely popular. Our goal is to harden rule-based
+NIDS by automatically creating rules from suspicious behavior detected
+by anomaly-based NIDS.
 
@@
 \label{sec:suri-metas-coverage}
+
+\begin{figure}[t]
+  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
+  \caption{Suricata rule for SYN Flood Attacks.}
+  \label{fig:synflood-example}
+\end{figure}
 
@@
 volume of such packets should be requested in a short period of time.
-
-\begin{figure}[t]
-  \lstinputlisting[language=C,numbers=none]{synflood.suricata}
-  \caption{Suricata rule for SYN Flood Attacks.\Mar{colocar
-      propriedades em negrito}}
-  \label{fig:synflood-example}
-\end{figure}
 

### 2019-09-23 paper/paper.tex
pass on intro

@@
 prescribes a method to capture a denial-of-service attack to a
-server~\cite{Douligeris:2004:DAD:987153.987158}. Rule-based \nids{} focuses on known attacks whereas
-anomaly-based \nids{} focuses on unanticipated attacks that could be
+server~\cite{Douligeris:2004:DAD:987153.987158}. Rule-based \nids{} focus on known attacks whereas
+anomaly-based \nids{} focus on unanticipated attacks that could be
 observed through suspicious network
@@
 These intrusion detection mechanisms are complementary. Rule-based
-NIDS are very popular in industry (it is clear to system
-administrators what rules will do), but attacker are creating in
-evading the current set of rules, and writing these rules is tedious
-and error-prone. Anomaly-based NIDS are not restricted to a fixed set
-of rules; they learn common behaviors and dynamically detect uncommon
-behavior. The often use some machine learning algorithm to model
-sequence of events~\cite{7579764}. Unfortunately, anomaly-based NIDS
-can raise false alarms as behavior models are fundamentally
-imprecise. To sum up, both approaches have positive and negative
-aspects.
+NIDS are restricted to a fixed set of rules and not all rules are
+simulneteously enabled (for efficiency). Attackers constantly create
+new attacks, and manually writing rules is tedious and
+error-prone. However, the semantics of the specification language of
+rules is clear making it easy for system administrator to understand
+them. As such, rule-based systems are very popular in industry.
+Anomaly-based NIDS are not restricted to a fixed set of rules; they
+learn common behaviors and dynamically detect uncommon behavior. They
+typically use machine learning algorithm to model event
+sequences~\cite{7579764}. Unfortunately, anomaly-based NIDS can raise
+false alarms as behavior models are fundamentally imprecise. To sum
+up, both approaches have positive and negative aspects.
 

### 2019-09-23 paper/paper.tex
pass on section II

@@
 
-This section 1) shows an example rule of Suricata and briefly explains
-the format of Suricata rules, 2) presents an example of
-a network attack, and 3) illustrates how \tname{} performs to
-synthesize the rule that would capture the attack.
+This section 1)~shows an example rule of Suricata, 2)~briefly explains
+the format of Suricata rules, 3) presents an example of a network
+attack, and 4) illustrates how \tname{} performs to synthesize the
+rule that would capture the attack.
 
@@
 satisfied. In this case, a message will be sent to sys admins if the
-rule pattern is satisfied. Other options of actions include pass,
-drop, and reject. The header comes after the action in the rule
-description. It restricts the information flow covered by the
-rule. For this rule, the header is \CodeIn{tcp \$HOME\_NET any ->
-  \$EXTERNAL\_NET any}, which instructs Suricata to inspect
-\CodeIn{tcp} traffic flowing from any port in the home network to any
-other address outside the home network. The variables
-\CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are
-configurable. The rule options come after the header. It is a
-semi-colon-separated sequence of properties\Mar{should we call it an
-  option or property?}, where each property is a key-value pair
-describing a distinct characteristic of the attack.
+rule pattern is satisfied. Other actions include pass, drop, and
+reject. The header comes after the action in the rule description. It
+restricts the information flow covered by the rule. For this rule, the
+header is \CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any}, which
+instructs Suricata to inspect \CodeIn{tcp} traffic flowing from any
+port in the home network to any other address outside the home
+network. The variables \CodeIn{\$HOME\_NET} and
+\CodeIn{\$EXTERNAL\_NET} are configurable. The rule options come after
+the header. It is a semi-colon-separated sequence of options, where
+each property is a key-value pair describing a distinct characteristic
+of the attack.
 
@@
 sends multiple SYN packets to different ports of a server, often using
-fake IP addresses. Without proper protection, the server accepts the
+fake IP addresses.\Mar{Are you saying that there is no timer that the
+  server starts when it sends a SYN-ACK message?} Without proper protection, the server accepts the
 connection requests and eventually legitimate connection requests
@@
   both, track by\_dst, count 5000, seconds 5}, indicating that a high
-volume of such packets should be requested in a short period of time.
+volume of such packets should be requested in a short period of
+time.\Mar{please, explain the other fields}
 
@@
   example. What is the input? How long does it take to generate the
-  rule? Is the rule generate exactly the same?}
+  rule? Is the rule generated exactly the same? Please explain...}
 

### 2019-09-23 paper/paper.tex
pass on approach

@@
 
+\begin{figure*}[t]
+\centering
+\includegraphics[trim=60 400 65 0,clip,width=0.7\textwidth]{figs/nids-workflow}
+\caption{\tname workflow. The first step initializes the encoding of
+  randomly initializes the rule encoding. Only options associated with
+  the protocol usd in the negative traffic are included in the
+  encoding. This step mutates the values associated with rule options
+  until \suri\ is able to capture the attack.  Note that any subset of
+  options in the rule reported in this step captures the negative
+  traffic. The second step is initialized with a configuration that
+  captures the negative traffic but could also capture benign
+  traffic. It discards options from the rule encoding until no
+  positive traffic is captured.}
+\label{fig:overview}
+\end{figure*}
+
 \tname{} is a search-based approach that uses positive and negative
 network traffic examples to synthesize signature-based NIDS rules,
-such as those from Snort and Suricata. \tname{} synthesizes rules in
-two steps. First, it produces a potentially overspecified rule that
-captures the negative example that is provided as input. Then, it uses
-positive examples to eliminate unnecessary options from the rule. The
-goal of \tname\ is to create a rule that captures only the negative traffic.
+such as those from Snort and Suricata.
+
+Figure~\ref{fig:overview} shows the workflow of \tname{}.  The
+technique synthesizes rules in two steps. First, it produces a
+potentially overspecified rule that captures the negative example that
+is provided as input. Then, it uses positive examples to eliminate
+unnecessary options from the rule. The goal of \tname\ is to create a
+rule that only captures the negative traffic.
+
+\Fix{describes negative and positive traffic}
 

### 2019-09-23 paper/paper.tex
pass on section II.B

@@
 sends multiple SYN packets to different ports of a server, often using
-fake IP addresses.\Mar{Are you saying that there is no timer that the
+fake IP addresses.
+\Luc{Then, after the server responds with the usual SYN-ACK packet, the client sends another SYN to avoid the connection time out.} \Mar{Are you saying that there is no timer that the
   server starts when it sends a SYN-ACK message?} Without proper protection, the server accepts the

### 2019-09-23 simulated-annealing/sa.py
starting to handle flood attacks

@@ -3,8 +3,9 @@ import time
 import copy
 
 keys_by_proto = {}
-keys_by_proto["icmp"] = {"itype":255, "icode":255, "icmp_seq":65525, "icmp_id":65535, "dsize":1480}
+keys_by_proto["icmp"] = {"dsize":1480, "itype":255, "icode":255, "icmp_seq":65525}#, "icmp_id":65535}
 # keys_by_proto["tcp"] = ["flow", "flags", "flowbits", "byte_test", "threshold", "seq", "ack", "window"]
+threshold = {"type":["limit", "threshold", "both"], "track":["by_src", "by_dst"], "count": 65535, "seconds": 1}
 
 default_rule_action = "alert"
 default_rule_header = "any any -> any any"
@@ -31,6 +32,10 @@ def getSuricataPid():
 def getLocalIp():
     return subprocess.Popen(["sh", "getLocalIp.sh"], stdout=subprocess.PIPE, encoding="utf-8").communicate()[0].rstrip()
 
+def sendFlood(local_ip):
+    subprocess.Popen(["sh", "sendFlood.sh", str(local_ip)], stdout=subprocess.DEVNULL).wait()
+    time.sleep(0.1)
+
 def sendPacket(local_ip):
     subprocess.Popen(["sh", "sendPacket.sh", str(local_ip)], stdout=subprocess.DEVNULL).wait()
 
@@ -54,23 +59,34 @@ def evalRule(rule):
     subprocess.Popen(["sudo", "rm", "../suricata/rulesFitness.txt"], stdout=subprocess.DEVNULL).wait()
     writeRuleOnFile(rule)
     reloadSuricataRules(getSuricataPid())    
-    sendPacket(getLocalIp())    
-
+    #sendPacket(getLocalIp())    
+    sendFlood(getLocalIp())
     fitnessFile_path = "../suricata/rulesFitness.txt"
 
     try:
         fitnessFile = open(fitnessFile_path, "r")
     except IOError:
-        return 0
+        return 0, 0
     
     lines = fitnessFile.readlines()
-    last_line = lines[len(lines)-1].rstrip('\n')
-    print(last_line)
-    keywords = last_line.split("-")
+    fitness = 0
+    matches = 0
+    for line in lines:
+        line = line.rstrip('\n')
+        keywords = line.split("-")
+        if rule.threshold == {}:
+            for key in keywords:
+                if "threshold" in key:
+                    keywords = keywords[:-1]
+                    break
[…]

### 2019-09-23 suricata/src/detect-engine-alert.c

@@ -245,6 +245,8 @@ void PacketAlertFinalize(DetectEngineCtx *de_ctx, DetectEngineThreadCtx *det_ctx
         const Signature *s = de_ctx->sig_array[p->alerts.alerts[i].num];
 
         int res = PacketAlertHandle(de_ctx, det_ctx, s, p, &p->alerts.alerts[i]);
+        int fitness = (res == 1) ? 1 : 0;
+        logFitness("threshold", s->id, fitness);
         if (res > 0) {
             /* Now, if we have an alert, we have to check if we want
              * to tag this session or src/dst host */

### 2019-09-23 suricata/src/detect.c

@@ -680,7 +680,7 @@ static inline int DetectRunInspectRulePacketMatches(
 {
     FILE *fp = fopen("./rulesFitness.txt", "a+");
     
-    fprintf(fp, "SID: %d", s->id);
+    fprintf(fp, "\nSID: %d", s->id);
 
     fclose(fp);
 
@@ -719,7 +719,7 @@ static inline int DetectRunInspectRulePacketMatches(
                 if (smd->is_last) {
                     printf("Matches: %d/%d - SID: %d %s\n\n", matches, sig_cnt, s->id, s->msg);
                     FILE *fp = fopen("./rulesFitness.txt", "a");
-                    fprintf(fp, "\n");
+                    //fprintf(fp, "\n");
 
                     fclose(fp);
 

### 2019-09-23 suricata/src/pesquisa.c

@@ -1,4 +1,5 @@
 #include "pesquisa.h"
+#include "string.h"
 
 void logFitness(void *key, int signatureID, float fitness){
     FILE *fp = fopen("./rulesFitness.txt", "a");
@@ -6,8 +7,11 @@ void logFitness(void *key, int signatureID, float fitness){
     char *keyword = (char *) key;
     //fitness = abs(fitness);
 
+    if(strcmp(keyword, "threshold") == 0){
+        fseek(fp, -2, SEEK_END);
+    }
+
     fprintf(fp, " - %s: %f", keyword, fitness);
     fclose(fp);
-    //if (fitness == 0)
     printf("SID: %d - %s keyword fitness: %f\n", signatureID, keyword, fitness);
 }

### 2019-09-24 paper/paper.tex
pass on workflow

--- 2019-09-23 paper/paper.tex
@@
 sends multiple SYN packets to different ports of a server, often using
-fake IP addresses.\Mar{Are you saying that there is no timer that the
-  server starts when it sends a SYN-ACK message?} Without proper protection, the server accepts the
-connection requests and eventually legitimate connection requests
-cannot be satisfied due to resource exhaustion.
+fake IP addresses. Then, after the server responds with a SYN-ACK
+packet, the client sends another SYN to avoid the connection to time
+out. Without proper protection, the server accepts the connection
+requests and eventually legitimate connection requests cannot be
+satisfied due to resource exhaustion.
 
@@
 
+\begin{figure*}[t]
+\centering
+\includegraphics[trim=60 400 65 0,clip,width=0.7\textwidth]{figs/nids-workflow}
+\caption{\tname\ workflow.}
+\label{fig:overview}
+\end{figure*}
+
 \tname{} is a search-based approach that uses positive and negative
-network traffic examples to synthesize signature-based NIDS rules,
-such as those from Snort and Suricata. \tname{} synthesizes rules in
-two steps. First, it produces a potentially overspecified rule that
-captures the negative example that is provided as input. Then, it uses
-positive examples to eliminate unnecessary options from the rule. The
-goal of \tname\ is to create a rule that captures only the negative traffic.
+network examples to synthesize signature-based NIDS rules, such as
+those from Snort and Suricata.
+
+The technique synthesizes rules in two steps. First, it produces a
+potentially overspecified rule that captures the negative example that
+is provided as input. Then, it uses positive examples to eliminate
+unnecessary options from the rule. The goal of \tname\ is to create a
+rule that only captures the negative traffic.
+
+Figure~\ref{fig:overview} shows the workflow of \tname{} as a pipeline
+of two components, each implementing one of the steps mentioned above.
+The goal of the first component is to produce a rule that captures the
+negative traffic.  First, it initializes the rule encoding with random
+values. Only options associated with the used protocol are included in
+the encoding. Then, \tname\ optimizes the rule options until \suri\ is
+able to capture the attack.  Note that, as all options need to be
+satisfied for the traffic to be captured, at this point, any subset of
+options would capture the negative traffic. The second component of
+the pipeline is initialized with the rule produced in the previous
+step. That rule captures the negative traffic but could also capture
+benign traffic. This component systematically discards options from
+the rule encoding until no positive traffic is captured.
+
+\Fix{describes negative and positive traffic}
 

### 2019-09-24 paper/paper.tex
pass on technique

@@
 NIDS are restricted to a fixed set of rules and not all rules are
-simulneteously enabled (for efficiency). Attackers constantly create
+simultaneously enabled (for efficiency). Attackers constantly create
 new attacks, and manually writing rules is tedious and
@@
 
-This paper makes the following contributions. \Fix{...}
+This paper makes the following contributions.
 
@@
 to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
-works as follows. First, a client sends a SYN packet to the server,
-requesting a connection. Second, the server responds with a SYN-ACK
-packet to the client. Third, the client responds with an ACK message
-and the connection is established. Aware of the protocol, an attacker
-sends multiple SYN packets to different ports of a server, often using
-fake IP addresses. Then, after the server responds with a SYN-ACK
-packet, the client sends another SYN to avoid the connection to time
-out. Without proper protection, the server accepts the connection
-requests and eventually legitimate connection requests cannot be
-satisfied due to resource exhaustion.
+works as follows in normal circumstances. First, a client sends a SYN
+packet to the server, requesting a connection. Second, the server
+responds with a SYN-ACK packet to the client. Third, the client
+responds with an ACK message and the connection is established. Aware
+of the protocol, an attacker sends multiple SYN packets to different
+ports of a server, often using fake IP addresses. Then, after the
+server responds with a SYN-ACK packet, the client keeps sending other
+SYN to avoid the connection to time out. Without proper protection,
+the server accepts these malicious requests and eventually legitimate
+requests cannot be satisfied due to resource exhaustion.
 
@@
 \centering
-\includegraphics[trim=60 400 65 0,clip,width=0.7\textwidth]{figs/nids-workflow}
+\includegraphics[trim=10 360 182 130,clip,width=0.7\textwidth]{figs/nids-workflow}
 \caption{\tname\ workflow.}
@@
 The technique synthesizes rules in two steps. First, it produces a
-potentially overspecified rule that captures the negative example that
-is provided as input. Then, it uses positive examples to eliminate
+potentially overspecified rule that captures the negative example
+provided on input. Then, it uses positive examples to eliminate
 unnecessary options from the rule. The goal of \tname\ is to create a
@@
 Figure~\ref{fig:overview} shows the workflow of \tname{} as a pipeline
-of two components, each implementing one of the steps mentioned above.
-The goal of the first component is to produce a rule that captures the
-negative traffic.  First, it initializes the rule encoding with random
-values. Only options associated with the used protocol are included in
-the encoding. Then, \tname\ optimizes the rule options until \suri\ is
-able to capture the attack.  Note that, as all options need to be
-satisfied for the traffic to be captured, at this point, any subset of
-options would capture the negative traffic. The second component of
-the pipeline is initialized with the rule produced in the previous
-step. That rule captures the negative traffic but could also capture
-benign traffic. This component systematically discards options from
-the rule encoding until no positive traffic is captured.
+of two components, appearing numbered in the figure. Each component
+implements one of the steps mentioned above. The goal of the first
+component is to produce a rule that captures the negative traffic.
+First, it initializes the rule with random values. Only options
+associated with the used protocol are included in the rule
+representation. Then, \tname\ optimizes the rule options until
+\suri\ is able to capture the attack associated with the negative
+traffic.  \tname\ unsuccessfully terminates at this point if it cannot
+capture the attack. The second component takes as input the rule
+produced by the first component and tries to minimize that rule. Note
+that any subset of options would capture the negative traffic---as all
+options in the rule need to be satisfied for the negative traffic to
+be captured---but it can also capture positive traffic. This component
+systematically discards options from the rule encoding until no
+positive traffic is captured.
 
-\Fix{describes negative and positive traffic}
+\subsection{Positive Traffic}
+
+Note from the figure that the negative traffic is an input to the
+technique. The positive traffic, in contrast, is not. \tname{} uses
+public databases storing different kinds of positive traffic for all
+protocols covered by Suricata. \Fix{what databases?}
 

### 2019-09-24 paper/paper.tex
pass on evaluation

@@
 \section{Evaluation}
+\Luc{Atualmente estamos considerando os seguintes ataques:
+\subsection{Ping Scan}
+Utilizado para descobrir hosts ativos na rede - pacote individual - protocolo ICMP
+\subsection{Port Scan TCP}
+Utilizado para descobrir portas abertas em um host - pacote individual - protocolo TCP
+\subsection{Port Scan UDP}
+Utilizado para descobrir portas abertas em um host - pacote individual - protocolo UDP
+\subsection{SYN Flood}
+DoS TCP
+\subsection{Black Nurse}
+DoS ICMP
+\subsection{Ping Flood}
+DoS ICMP
+\subsection{GET/POST Flood}
+DoS HTTP
+\
+}
 

### 2019-09-25 paper/paper.tex
pass on text

--- 2019-09-24 paper/paper.tex
@@
 
-\subsection{Positive Traffic}
-
-Note from the figure that the negative traffic is an input to the
-technique. The positive traffic, in contrast, is not. \tname{} uses
-public databases storing different kinds of positive traffic for all
-protocols covered by Suricata. \Fix{what databases?}
-
 \subsection{Problem Characterization}
@@
 
+\subsection{Limitations}
+
+\Fix{...}
+
 \section{Evaluation}
 
-\Mar{Lucas/Guilherme (CRI): precisamos dar pontape inicial propondo
-  perguntas de pesquisa. Eu faco ajustes em seguida.}
+%% \renewcommand{\labelitemi}{}
+%% \setitemize[0]{leftmargin=20pt,itemindent=-20pt}
+%% \begin{itemize}
+%% \end{itemize}
+
+%% \renewcommand{\labelitemi}{}
+%% \setitemize[0]{leftmargin=0pt,itemindent=0pt}
+%% \begin{itemize}
+%%   \item{\textbf{RQ1.}~How precise is \tname?}
+%%   \item{\textbf{RQ2.}~How efficient is \tname?}
+%% \end{itemize}
+
+This section reports precision and efficiency of \tname.
+
+\subsection{Objects of Analysis}
+
+We evaluated \tname{} on a set of attacks covering each of the four
+protocols supported by Suricata and attacks with different
+characteristics. Table~\ref{table:attacks} shows the attacks we
+analyzed and the rule of Suricata that captures the attack. Column
+``Name'' shows the name of the attack, column ``\#'' indicates whether
+single or multiple messages are required to manifest the attack,
+column ``Protocol'' shows the protocol name, column ``Description''
+summarizes the effect of the attack, if successfull, and column
+``Golden Rule'' shows the relevant parts in the options section of the
+Suricata rule that captures the attack.
+
+\begin{table*}[t!]
+  \caption{\label{table:attacks}Characterization of attacks analyzed
+    and Golden Rule of Suricata (reference) to capture it.}  
+  \centering
+  \begin{tabular}{lllll}
+    \toprule
+    \multicolumn{1}{c}{Name} & \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Protocol} & \multicolumn{1}{c}{Description} & \multicolumn{1}{c}{Golden Rule} \\
+    \midrule
+    GET/POST Flood & multiple & HTTP & Denial of Service & - \\      
+    Ping Scan & single & ICMP & Discovers active hosts in a network &  - \\
+    Black Nurse & multiple & ICMP & Denial of Service & - \\
+    Ping Flood  & multiple & ICMP & Denial of Service & - \\  
+    Port Scan TCP & single & TCP & Discovers open ports in a host & -  \\
+    SYN Flood & multiple & TCP & Denial of Service & - \\
+    Port Scan UDP & single & UDP & Discovers open ports in a host & - \\        
+    \bottomrule
+  \end{tabular}
+\end{table*}
+
+\subsection{Setup}
+
+Note from the figure that the negative traffic is an input to the
+technique. The positive traffic, in contrast, is not. \tname{} uses
+public databases storing different kinds of positive traffic for all
+protocols covered by Suricata. \Fix{what databases?}
+
+\Fix{any other detail we should consider?}
+
+\subsection{Precision}
+
+\Fix{...}
+
+\subsection{Efficiency}
+
+\Fix{...}
 

### 2019-09-25 paper/paper.tex
pass

@@
 computer networks given the amount of potential malicious traffic that
-exist on the Internet today.
+exist on the Internet.
 
 \sloppy \nids{} are typically categorized in two
-groups~\cite{kumar2007survey}: 1)~rule-based and 2)~anomaly-based. A
-rule-based \nids{} (\aka\ signature-based NIDS) checks if the network
-traffic matches a fixed set of
+groups~\cite{kumar2007survey}: rule-based and anomaly-based \nids. A
+rule-based intrusion detector\footnote{\aka\ signature-based intrusion
+  detector.} checks if the network traffic matches a fixed set of
 rules. Figure~\ref{fig:synflood-example} shows an example rule of
 Suricata~\cite{suricata}, a popular open-source \nids{}. This rule
-prescribes a method to capture a denial-of-service attack to a
-server~\cite{Douligeris:2004:DAD:987153.987158}. Rule-based \nids{} focus on known attacks whereas
-anomaly-based \nids{} focus on unanticipated attacks that could be
-observed through suspicious network
+prescribes a method to capture a denial-of-service
+attack~\cite{understanding-dos} to a server by matching specific
+conditions about the network traffic. Relevant conditions appear in
+bold. Rule-based \nids{} focus on known attacks whereas anomaly-based
+\nids{} focus on unknown attacks that could be observed by first
+learning the regular traffic and then capturing suspicious
 traffic~\cite{kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}.
 
-These intrusion detection mechanisms are complementary. Rule-based
-NIDS are restricted to a fixed set of rules and not all rules are
-simultaneously enabled (for efficiency). Attackers constantly create
-new attacks, and manually writing rules is tedious and
-error-prone. However, the semantics of the specification language of
-rules is clear making it easy for system administrator to understand
-them. As such, rule-based systems are very popular in industry.
-Anomaly-based NIDS are not restricted to a fixed set of rules; they
-learn common behaviors and dynamically detect uncommon behavior. They
-typically use machine learning algorithm to model event
-sequences~\cite{7579764}. Unfortunately, anomaly-based NIDS can raise
-false alarms as behavior models are fundamentally imprecise. To sum
-up, both approaches have positive and negative aspects.
+These two kinds of intrusion detection mechanisms are
+complementary. Conceptually, rule-based NIDS could miss attacks (as
+rulesets can become outdated) whereas anomaly-based NIDS can raise
+false alarms (as the detection algorithms use are approximate).
+Rule-based NIDS are restricted to a fixed set of rules and attackers
+constantly create new strategies. Furthermore, manually writing rules
+to capture those attacks is tedious and error-prone. As consequence,
+IT-security companies advertize paid rulesets for consumers to
+buy~\cite{proofpoint-etpro,snort-rule-subscriptions}.  Despite this
+fundamental issue, rule-based systems are tremendously popular in
+industry today~\Fix{cite}. Anomaly-based NIDS, in contrast, are not
+restricted to a fixed set of rules. They learn usage behavior from
+positive traffic and, based on that, alert uncommon
+behavior~\cite{7579764}. Unfortunately, anomaly-based NIDS are
+fundamentally imprecise. To sum up, both approaches have positive and
+negative aspects.
 
@@
 
-This paper proposes a technique, dubbed \tname{}, to automatically
-synthesize rules from positive and negative examples (\ie{}, benign
-and malicious traffic). The main scenario of application is one where
-an anomaly-based NIDS identifies abnormal traffic and signals that
+This paper proposes a technique, dubbed \tname{}, to close the gap
+between these two strategies. \tname{} automatically synthesizes rules
+from positive and negative examples (\ie{}, benign and malicious
+traffic). A scenario of application of \tname{} is one where an
+anomaly-based NIDS identifies abnormal traffic and signals that
 traffic to \tname{} to craft rules that can be used by rule-based
-NIDS. Those rules can be distributed to rule databases. The core
-motivation is that 1) attackers are productive in creating new ways to
-circumvent existing protections and 2) manual creation of rules is
-tedious and time-consuming.  Note that anomaly-based NIDS are evolving
-pretty quick with the advances in machine learning, but rule-based
-NIDS are still extremely popular. Our goal is to harden rule-based
-NIDS by automatically creating rules from suspicious behavior detected
-by anomaly-based NIDS.
+NIDS. Note that anomaly-based NIDS are evolving pretty quick with the
+advances in machine learning, but rule-based NIDS are still extremely
+popular.  \tname{} could also leverage existing databases of malicious
+traffic to synthesize rules.  Regardless of how the negative traffic
+is produced (out of scope of this paper), those rules can be
+distributed to rule databases for free. Our goal is to harden the
+protection of NIDS by automatically synthesizing rules from negative
+examples.
+
+%%  The
+%% core motivation is that 1) attackers are productive in creating new
+%% ways to circumvent existing protections and 2) manual creation of
+%% rules is tedious and time-consuming.
 
@@
 candidate solution to the target problem. Candidate solutions are the
-objects that the search optimize towards solutions. In this case, a a
+objects that the search optimize towards solutions. In this case, a 
 candidate solution is an encoding of a Suricata rule.  \tname\ encodes
@@
 example on Figure~\ref{fig:synflood-example}, have no influence in
-triggering the actions associated with rules. The option
+triggering the actions associated with the rule. The option
 \CodeIn{flags}, however, is relevant to trigger the action. \tname{}
@@
 mutate each element separately. Similar encodings are used in modeling
-the other relevant options. We found what is (ir)relevant by
+the other relevant options. We found what options are (ir)relevant by
 inspecting Suricata code for the presence of \CodeIn{eval} functions,
-which are invoked from a main look to process every Suricata rule. The
+which are invoked from a main loop to process each action rule. The
 presence of such a function for a given option indicates that it is
@@
 
-Options can be specific to a given protocol or general and they can be
-influential on triggering an action or
-not. Figure~\ref{fig:distribution-rules-protocol} shows the
+Options can be specific to a given protocol or general and they can
+influence or not triggering an
+action. Figure~\ref{fig:distribution-rules-protocol} shows the
 distributions of these options per protocol and their relevance for

### 2019-10-02 paper/paper.tex
pass on evaluation

@@
 
+\Gui{We are using the bigflows.pcap provided by tcpreplay on http://tcpreplay.appneta.com/wiki/captures.html "This is a capture of real network traffic on a busy private network’s access point to the Internet"
+We also used normal pcaps from Stratosphere Lab on https://www.stratosphereips.org/datasets-normal to verify if Suricata's ET rules weren't catching positive flow. }
+
 \Fix{any other detail we should consider?}

### 2019-10-02 paper/paper.tex
minor comments

--- 2019-09-25 paper/paper.tex
@@
   \caption{\label{table:attacks}Characterization of attacks analyzed
-    and Golden Rule of Suricata (reference) to capture it.}  
+    and Golden Rule of Suricata (reference) to capture it.\Mar{please,
+  show me the Golden Rule, \ie{} the part that is relevant of each rule (for
+  capturing the negative traffic precisely)}}  
   \centering
@@
 
-Note from the figure that the negative traffic is an input to the
-technique. The positive traffic, in contrast, is not. \tname{} uses
-public databases storing different kinds of positive traffic for all
-protocols covered by Suricata. \Fix{what databases?}
+Note from Figure~\ref{fig:overview} that the negative traffic is an
+input to the technique. The positive (non malicious) traffic, in
+contrast, is not. \tname{} uses public databases storing different
+kinds of positive traffic for all protocols covered by Suricata. We
+used two datasets of positive traffic. The Bigflows.pcap dataset is
+part of the TcpReplay~\cite{tcpreplay} open source project. This
+dataset includes real network traffic on a busy private network access
+point to the Internet. According to the TcpReplay web site ``This
+capture is much larger and has a smaller average packet size than the
+previous capture (smallFlows.pcap). It also has many more flows and
+different applications.''. In addition to the Bigflows.pcap dataset,
+we also used datasets of normal traffic made available by Stratosphere
+Lab~\cite{stratosphere-normal}.
+
+%% \Gui{We are using the bigflows.pcap provided by tcpreplay on http://tcpreplay.appneta.com/wiki/captures.html "This is a capture of real network traffic on a busy private network’s access point to the Internet"
+%%  }
 

### 2019-10-03 paper/paper.tex
minor comments

@@
 
+\Luc{entrada: conjunto de 20 pacotes SYN que representam o flood.
+tempo para gerar a regra: ~1m30s
+no caso do syn flood, geramos a mesma regra.
+}
+
 \section{Approach}
@@
 from returning true, which indicates it is closer to trigger the
-action associated with the rule. For the example above,
+action associated with the rule.
+\Luc{Para opções cujos valores são numéricos, o suricata apenas verifica se o valor da opção no pacote é igual/maior/menor ao valor especificado na regra, retornando um booleano indicando isso.
+Alteramos o codigo fonte para obter a diferença entre o valor da opção no pacote e o valor da opção na regra.}
 \Fix{Lucas/Guilherme, it would be good to show how this looks like in

### 2019-10-03 paper/paper.tex
minor comments

--- 2019-10-02 paper/paper.tex
@@
 from returning true, which indicates it is closer to trigger the
-action associated with the rule. For the example above,
+action associated with the rule. 
+\Luc{Originalmente, o suricata interrompe a checagem de uma regra quando alguma das opções não dá match com o pacote analisado.
+O código fonte foi alterado para que as regras sempre tenham todas as suas opções verificadas.
+A função de verificação de uma opção retorna um booleano indicando se o valor da opção na regra corresponde ou não ao valor da opção no pacote.
+Para opções cujos valores são numéricos, alteramos o código fonte para obter a diferença entre os valores da opção na regra e no pacote.}
 \Fix{Lucas/Guilherme, it would be good to show how this looks like in

### 2019-10-04 paper/paper.tex
pass on example and approach

@@
 
-\Luc{entrada: conjunto de 20 pacotes SYN que representam o flood.
-tempo para gerar a regra: ~1m30s
-no caso do syn flood, geramos a mesma regra.
+\Luc{entrada: utilizamos a ferramenta hping3 para executar o ataque em um alvo arbitrario dentro da rede (n ha necessidade de haver um alvo real para executar o hping3) e capturamos o trafego gerado com o wireshark.
+o comando utilizado foi:\\*
+\texttt{\# hping3 -d 80 -w 64 -S -p 80 --flood --rand-source 192.138.1.115} \\*
+que envia pacotes de 120 bytes (-d 120), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80).
+A flag --flood indica q o hping3 vai enviar os pacotes o mais rápido possível.
+a flag --rand-source indica que cada pacote vai ter um ip de origem aleatorio.
+192.138.1.115 é o alvo.\\*
+O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
+regra inicial: -\\*
+regra intermediaria: -\\*
+regra final: -\\*
+iterações para gerar a regra: -
+
 }
@@
 A função de verificação de uma opção retorna um booleano indicando se o valor da opção na regra corresponde ou não ao valor da opção no pacote.
-Para opções cujos valores são numéricos, alteramos o código fonte para obter a diferença entre os valores da opção na regra e no pacote.}
+Para opções cujos valores são numéricos, alteramos o código fonte para obter a diferença entre os valores da opção na regra e no pacote.
+\\*
+Exemplo trecho de código original opção "icode":
+
+\begin{figure}[h]
+  \lstinputlisting[language=C,numbers=none,keywords={}]{original.code}
+  \caption{Original source code.}
+  \label{fig:original-code}
+\end{figure}
+
+Trecho modificado:
+
+\begin{figure}[h]
+  \lstinputlisting[language=C,numbers=none,keywords={}]{modified.code}
+  \caption{Modified source code.}
+  \label{fig:modified-code}
+\end{figure}
+
+}
+
 \Fix{Lucas/Guilherme, it would be good to show how this looks like in

### 2019-10-04 paper/paper.tex
pass on table 1

@@
   \centering
-  \begin{tabular}{lllll}
+  \begin{tabular}{llllllllllll}
     \toprule
@@
     \midrule     
-    GET/POST Flood & multiple & HTTP & Denial of Service & - \\      
-    Ping Scan & single & ICMP & Discovers active hosts in a network &  - \\
-    Black Nurse & multiple & ICMP & Denial of Service & - \\
-    Ping Flood  & multiple & ICMP & Denial of Service & - \\  
-    Port Scan TCP & single & TCP & Discovers open ports in a host & -  \\
-    SYN Flood & multiple & TCP & Denial of Service & - \\
-    Port Scan UDP & single & UDP & Discovers open ports in a host & - \\        
+    Ping Scan & single & ICMP & Discovers active hosts in a network &  dsize:0; itype:8; \\
+    Black Nurse & multiple & ICMP & Denial of Service & itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;\\
+    Ping Flood  & multiple & ICMP & Denial of Service & itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;\\  
+    Port Scan TCP & single & TCP & Discovers open ports in a host & flags: A; ack: 0;  \\
+    SYN Flood & multiple & TCP & Denial of Service & threshold: type both, track by\_dst, count 5000, seconds 5;\\
     \bottomrule

### 2019-10-04 paper/paper.tex
pass on table 1

@@
 O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
+
+\begin{table*}[t!]
+  \caption{\label{table:rules}Rule evolution}  
+  \centering
+  \begin{tabular}{llllllllllll}
+    \toprule
+    \multicolumn{1}{c}{Name} & \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Protocol} & \multicolumn{1}{c}{Description} & \multicolumn{1}{c}{Golden Rule} \\
+    \midrule     
+    Ping Scan & single & ICMP & Discovers active hosts in a network &  dsize:0; itype:8; \\
+    Black Nurse & multiple & ICMP & Denial of Service & itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;\\
+    Ping Flood  & multiple & ICMP & Denial of Service & itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;\\  
+    Port Scan TCP & single & TCP & Discovers open ports in a host & flags: A; ack: 0;  \\
+    SYN Flood & multiple & TCP & Denial of Service & threshold: type both, track by\_dst, count 5000, seconds 5;\\
+    \bottomrule
+  \end{tabular}
+\end{table*}
+
 regra inicial: -\\*
@@
   \centering
-  \begin{tabular}{llllllllllll}
+  \begin{tabular}{lllll}
     \toprule

### 2019-10-04 paper/paper.tex
fix pass on table 1
pass on table I

--- 2019-10-04 paper/paper.tex
@@
     \midrule     
-    GET/POST Flood & multiple & HTTP & Denial of Service & - \\      
-    Ping Scan & single & ICMP & Discovers active hosts in a network &  - \\
-    Black Nurse & multiple & ICMP & Denial of Service & - \\
-    Ping Flood  & multiple & ICMP & Denial of Service & - \\  
-    Port Scan TCP & single & TCP & Discovers open ports in a host & -  \\
-    SYN Flood & multiple & TCP & Denial of Service & - \\
-    Port Scan UDP & single & UDP & Discovers open ports in a host & - \\        
+    Ping Scan & single & ICMP & Discovers active hosts in a network &  dsize:0; itype:8; \\
+    Black Nurse & multiple & ICMP & Denial of Service & itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;\\
+    Ping Flood  & multiple & ICMP & Denial of Service & itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;\\  
+    Port Scan TCP & single & TCP & Discovers open ports in a host & flags: A; ack: 0;dsize: 0; \\
+    SYN Flood & multiple & TCP & Denial of Service & flags: S,12; threshold: type both, track by\_dst, count 5000, seconds 5;\\
+    UDP Flood & multiple & UDP & Denial of Service & fragbits:M; threshold: type both, track by_dst, count 5000, seconds 5; \\
     \bottomrule

### 2019-10-04 paper/paper.tex
fix escaped compilation error

--- 2019-10-04 paper/paper.tex
@@
     Ping Flood  & multiple & ICMP & Denial of Service & itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;\\  
-    Port Scan TCP & single & TCP & Discovers open ports in a host & flags: A; ack: 0;  \\
-    SYN Flood & multiple & TCP & Denial of Service & threshold: type both, track by\_dst, count 5000, seconds 5;\\
+    Port Scan TCP & single & TCP & Discovers open ports in a host & flags: A; ack: 0;dsize: 0; \\
+    SYN Flood & multiple & TCP & Denial of Service & flags: S,12; threshold: type both, track by\_dst, count 5000, seconds 5;\\
+    UDP Flood & multiple & UDP & Denial of Service & fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5; \\
     \bottomrule

### 2019-10-05 paper/paper.tex
pass on example

@@
 \texttt{\# hping3 -d 80 -w 64 -S -p 80 --flood --rand-source 192.138.1.115} \\*
-que envia pacotes de 120 bytes (-d 120), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80).
+que envia pacotes de 80 bytes (-d 80), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80).
 A flag --flood indica q o hping3 vai enviar os pacotes o mais rápido possível.
@@
 O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
-regra inicial: -\\*
-regra intermediaria: -\\*
-regra final: -\\*
-iterações para gerar a regra: -
+O wireshark armazena os pacotes no formato pcap, utilizamos a ferramenta tcpreplay para reproduzir o trafego com o comando:\\*
+\texttt{\# tcpreplay --pps 200 --intf1=wlp2s0 Datasets/syn-flood-sample.pcap}\\*
+Nesse caso, estamos enviando os pacotes armazenados em \texttt{Datasets/syn-flood-sample.pcap} a uma taxa de 200 pacotes por segundo (\texttt{-pps 200}) utilizando a interface wlp2s0 (\texttt{--intf1 wlp2s0})\\*
+Há diferenças entre os pacotes na opção ack.\\* 
+Por que 20?
+Apenas para gerar a regra mais rapidamente, poderia ser qualquer valor maior ou menor.
+No final, o parâmetro \texttt{count} da opção \texttt{threshold} vai ser alterado pelo adm da rede.\\*
+Intervalo de reprodução: 0.005 segundo
+
+\begin{table}[h]
+  \caption{\label{table:rules}Rule evolution}  
+  \centering
+  \begin{tabular}{lllll}
+    \toprule
+    \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
+    \midrule     
+    0 & (ack:0; seq:0; window:0; flags:F;)\\
+    5 & (window:0; flags:F;)\\
+    68 & (window:64; flags:F;)\\
+    73 & (window:64; flags:S,12;)\\
+    94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
+    \bottomrule
+  \end{tabular}
+\end{table}
 

### 2019-10-16 paper/paper.tex
pass on intro

@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first
-% footnote. If that is unneeded, please comment it out.
-
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
+%
+\documentclass[runningheads]{llncs}
+%
+\usepackage{graphicx}
 \usepackage{booktabs}
@@
 
-%% \usepackage{cite}
-%% \usepackage{amsmath,amssymb,amsfonts}
-%% \usepackage{algorithmic}
-%% \usepackage{graphicx}
-%% \usepackage{textcomp}
-%% \usepackage{xcolor}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 \begin{document}
-
-\title{Search-based Synthesis of Network Intrusion Detection Rules}
-
-%%%%%%%%%%% Anonymized
-
-%% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address}
-%% \and
-%% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address}
-%% \and
-%% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address}
-%% \and
-%% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address}
-%% \and
-%% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address}
-%% \and
-%% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-%% \textit{name of organization (of Aff.)}\\
-%% City, Country \\
-%% email address}
-%% }
-
-\maketitle
-
+%
+\title{Synthesis of Network Intrusion Detection Rules}
+%
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{Lucas Alcantara\inst{1}\Comment{\orcidID{0000-1111-2222-3333}} \and
+Guilherme Padilha\inst{1}\Comment{\orcidID{1111-2222-3333-4444}} \and
+Marcelo d'Amorim\inst{1}\Comment{\orcidID{0000-0002-1323-8769}}}
+%
+\authorrunning{Alcantara et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Federal University of Pernambuco (UFPE), Recife, Brazil\\
+\email{\{lama,ghps,damorim\}@cin.ufpe.br}}
+%
+\maketitle              % typeset the header of the contribution
+%
 \begin{abstract}
 ...
+\keywords{NIDS, synthesis, search}
 \end{abstract}
-
-\begin{IEEEkeywords}
-NIDS, synthesis, search
-\end{IEEEkeywords}
-
+%
+%
+%
 \section{Introduction}
@@
 not block. Various open-source (\eg{}, Snort~\cite{snort} and
-Suricata~\cite{suricata}) and commercial implementations (\eg{},
+Suricata~\cite{suricata}) and commercial NIDS implementations (\eg{},
 SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar}) exist
@@
 attack~\cite{understanding-dos} to a server by matching specific
-conditions about the network traffic. Relevant conditions appear in
-bold. Rule-based \nids{} focus on known attacks whereas anomaly-based
-\nids{} focus on unknown attacks that could be observed by first
-learning the regular traffic and then capturing suspicious
-traffic~\cite{kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}.
+conditions about the network traffic. Relevant properties about the
+traffic of interest appear in bold (see
+Section~\ref{sec:suri-metas-coverage}). Rule-based \nids{} focus on
+known attacks whereas anomaly-based \nids{} focus on unknown attacks
+that could be observed by first learning the regular traffic and then
+acting
+properly~\cite{kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}. These
+two kinds of intrusion detection mechanisms are
+complementary. Rule-based NIDS could miss attacks (as rulesets can
+become outdated) whereas anomaly-based NIDS can raise false alarms (as
+detection algorithms are approximate).
 
-These two kinds of intrusion detection mechanisms are
-complementary. Conceptually, rule-based NIDS could miss attacks (as
-rulesets can become outdated) whereas anomaly-based NIDS can raise
-false alarms (as the detection algorithms use are approximate).
-Rule-based NIDS are restricted to a fixed set of rules and attackers
-constantly create new strategies. Furthermore, manually writing rules
-to capture those attacks is tedious and error-prone. As consequence,
-IT-security companies advertize paid rulesets for consumers to
-buy~\cite{proofpoint-etpro,snort-rule-subscriptions}.  Despite this
+Rule-based NIDS are restricted to a fixed set of rules defined by the
+network system administrator. Manually writing these rules is tedious,
+error-prone, but important as attackers constantly create new
+strategies to circumvent existing rules. IT-security companies
+capitalize on this phenomena and put rulesets on the market for
+sale~\cite{proofpoint-etpro,snort-rule-subscriptions}.  Despite this
 fundamental issue, rule-based systems are tremendously popular in
-industry today~\Fix{cite}. Anomaly-based NIDS, in contrast, are not
-restricted to a fixed set of rules. They learn usage behavior from
-positive traffic and, based on that, alert uncommon
-behavior~\cite{7579764}. Unfortunately, anomaly-based NIDS are
-fundamentally imprecise. To sum up, both approaches have positive and
-negative aspects.
+industry today. Anomaly-based NIDS are not restricted to a fixed set
+of rules. They learn usage behavior from positive traffic and, based
+on that, alert uncommon behavior~\cite{7579764}. Unfortunately,
+anomaly-based NIDS are fundamentally imprecise. To sum up, rule-based
+NIDS are precise but incomplete whereas anomaly-based NIDS are (more)
+complete but imprecise.
 
@@
 
-This paper proposes a technique, dubbed \tname{}, to close the gap
-between these two strategies. \tname{} automatically synthesizes rules
-from positive and negative examples (\ie{}, benign and malicious
-traffic). A scenario of application of \tname{} is one where an
-anomaly-based NIDS identifies abnormal traffic and signals that
-traffic to \tname{} to craft rules that can be used by rule-based
-NIDS. Note that anomaly-based NIDS are evolving pretty quick with the
-advances in machine learning, but rule-based NIDS are still extremely
-popular.  \tname{} could also leverage existing databases of malicious
-traffic to synthesize rules.  Regardless of how the negative traffic
-is produced (out of scope of this paper), those rules can be
-distributed to rule databases for free. Our goal is to harden the
-protection of NIDS by automatically synthesizing rules from negative
-examples.
+This paper proposes \tname{}, a technique that uses machine
+intelligence to automatically synthesize rules for rule-based
+\nids. Our general goal is to harden the protection of NIDS by
+improving the creation process of rules. One scenario of application
+of \tname{} is one where an anomaly-based NIDS identifies abnormal
+traffic and signals that traffic to \tname{} to create and distribute
+the rules.
 
+%% Note that anomaly-based NIDS are evolving pretty quick with the
+%% advances in machine learning, but rule-based NIDS are still extremely
+%% popular.  \tname{} could also leverage existing databases of malicious
+%% traffic to synthesize rules.  Regardless of how the negative traffic
+%% is produced (out of scope of this paper), those rules can be
+%% distributed to rule databases for free.
 %%  The
@@
 %% rules is tedious and time-consuming.
+
+\tname{} synthesizes rules from positive and negative examples, \ie{},
+malicious and benign traffic.
 
@@
 \centering
-\includegraphics[trim=10 360 182 130,clip,width=0.7\textwidth]{figs/nids-workflow}
+\includegraphics[trim=10 360 182 130,clip,width=0.95\textwidth]{figs/nids-workflow}
 \caption{\tname\ workflow.}
@@
 \Fix{...}
-
-\balance
-%\vspace*{-2\baselineskip}
-\bibliographystyle{IEEEtran}
+%
+% ---- Bibliography ----
+%
+% BibTeX users should specify bibliography style 'splncs04'.
+% References will then be sorted and formatted in the correct style.
+%
+\bibliographystyle{splncs04}
 \bibliography{references}
+%
 
 \end{document}
-
-%%  LocalWords:  Suricata SolarWinds QRadar NIDS OISF TCP ACK msg tcp
-%%  LocalWords:  encodings eval cardinality

### 2019-10-17 paper/paper.tex
pass up to section 2.1

@@
 
-\Mar{Lucas/Guilherme, verifique se isto esta correto}
 Figure~\ref{fig:synflood-example} shows an example rule of
@@
 options~\cite{suri-rule-format}. The action part appears as the first
-word in the rule description. In this case, \CodeIn{alert}. An action
-denotes the task that needs to be executed if the rule pattern is
-satisfied. In this case, a message will be sent to sys admins if the
-rule pattern is satisfied. Other actions include pass, drop, and
-reject. The header comes after the action in the rule description. It
-restricts the information flow covered by the rule. For this rule, the
-header is \CodeIn{tcp \$HOME\_NET any -> \$EXTERNAL\_NET any}, which
-instructs Suricata to inspect \CodeIn{tcp} traffic flowing from any
-port in the home network to any other address outside the home
-network. The variables \CodeIn{\$HOME\_NET} and
+word in the rule description. An action denotes the task that needs to
+be executed if the rule pattern is satisfied. In this example, a
+message will be sent to system administrators if the rule pattern is
+satisfied. The header comes after the action in the rule
+description. It restricts the information flow covered by the
+rule. For this rule, the header is \CodeIn{tcp \$HOME\_NET any ->
+  \$EXTERNAL\_NET any}. It instructs Suricata to inspect \CodeIn{tcp}
+traffic flowing from any port in the home network to any other address
+outside the home network. The variables \CodeIn{\$HOME\_NET} and
 \CodeIn{\$EXTERNAL\_NET} are configurable. The rule options come after
-the header. It is a semi-colon-separated sequence of options, where
-each property is a key-value pair describing a distinct characteristic
-of the attack.
+the header. It is a semi-colon-separated sequence of key-value
+pairs. Some options serve to document the analyzed traffic (\eg{},
+``msg'', ``classtype'', and ``sid'') whereas the remaining properties
+describe characteristics of the attack (\eg, ``flags'', ``threshold'',
+and ``content''). The main challenge in synthesizing rules is to
+discover the set of options that make a rule to capture the malicious
+traffic, but ignore the benign traffic.
 

### 2019-10-19 paper/paper.tex
pass on abstract

@@
 %
-\title{Synthesis of Network Intrusion Detection Rules}
+\title{Synthesis of Rules for Network Intrusion Detection}
 %
@@
 \begin{abstract}
-...
+Network Intrusion Detection Systems (\nids{}) are a popular defense
+mechanism against network attacks today. These systems monitor the
+network traffic and flag suspicious network behavior. Signature-based
+\nids\ do that by checking the network traffic against a pre-defined
+set of rules, which can become obsolete due to the adaptation of
+attackers to existing rules. This paper proposes \tname{}, a technique
+to synthesize rules for \nids; \tname{} uses previously-classified
+positive and negative examples (\ie{}, malicious and benign traffic)
+to identify the constraints that distiguish positive and negative
+examples. In a nutshell, \tname{} bootstraps the search with a
+candidate solution that uses data from the payload of the malicious
+message, optimizes the candidate solution to capture the positive
+example and, and then optimizes the candidate solution to avoid
+capturing the negative traffic. Results indicate that \Fix{...}
+
 \keywords{NIDS, synthesis, search}

### 2019-10-19 paper/paper.tex
pass on abstract

--- 2019-10-17 paper/paper.tex
@@
 \begin{abstract}
-...
+Network Intrusion Detection Systems (\nids{}) are a popular mechanism
+used by system administrators to defend against network attacks. These
+systems monitor the network traffic and flag suspicious network
+behavior. Signature-based \nids\ do that by checking the network
+traffic against a pre-defined set of rules, which can become obsolete
+as attackers learn new strategies to circumvent existing defenses.
+This paper proposes \tname{}, a technique that automatically
+synthesizes \nids\ rules from positive and negative examples (\ie{},
+malicious and benign traffic). \tname{} formulates synthesis as an
+optimization problem whose candidate solutions are rules that
+maximizes the captured the positive traffic and minimizes the captured
+negative traffic. \tname{} bootstraps the search with candidate
+solutions that use data from the payload of the malicious message and
+produces optimal rule candidates on output. We evaluated \tname{} on a
+diverse set of attacks. Results indicate that \Fix{...}
+
+%% optimizes the
+%% candidate solution to capture the positive example and, and then
+%% optimizes the candidate solution to avoid capturing the negative
+%% traffic.
+
 \keywords{NIDS, synthesis, search}

### 2019-10-20 paper/paper.tex
pass on technique

@@
 %
-\author{Lucas Alcantara\inst{1}\Comment{\orcidID{0000-1111-2222-3333}} \and
-Guilherme Padilha\inst{1}\Comment{\orcidID{1111-2222-3333-4444}} \and
-Marcelo d'Amorim\inst{1}\Comment{\orcidID{0000-0002-1323-8769}}}
+\author{Lucas Alcantara\Comment{\inst{1}}\Comment{\orcidID{0000-1111-2222-3333}} \and
+Guilherme Padilha\Comment{\inst{1}}\Comment{\orcidID{1111-2222-3333-4444}} \and
+Marcelo d'Amorim\Comment{\orcidID{0000-0002-1323-8769}}}
 %
@@
 server responds with a SYN-ACK packet, the client keeps sending other
-SYN to avoid the connection to time out. Without proper protection,
-the server accepts these malicious requests and eventually legitimate
-requests cannot be satisfied due to resource exhaustion.
+SYN packets to avoid the connection to time out. Without proper
+protection, the server accepts these malicious requests and eventually
+legitimate requests cannot be satisfied due to resource exhaustion.
+Figure~\ref{fig:synflood-example} shows a Suricata rule to capture
+this attack. The critical parts appear in bold---the option
+\CodeIn{flags: S,12}, which identifies a SYN packet and the option
+\CodeIn{threshold: type both, track by\_dst, count 5000, seconds 5},
+indicating that a high volume of such packets should be requested in a
+short period of time.
 
-Figure~\ref{fig:synflood-example} shows a Suricata rule to capture
-this attack. The critical parts are the option \CodeIn{flags: S,12},
-which identifies a SYN packet and the option \CodeIn{threshold: type
-  both, track by\_dst, count 5000, seconds 5}, indicating that a high
-volume of such packets should be requested in a short period of
-time.\Mar{please, explain the other fields}
+%\Mar{please, explain the other fields}
 
@@
 
-\Luc{entrada: utilizamos a ferramenta hping3 para executar o ataque em um alvo arbitrario dentro da rede (n ha necessidade de haver um alvo real para executar o hping3) e capturamos o trafego gerado com o wireshark.
-o comando utilizado foi:\\*
-\texttt{\# hping3 -d 80 -w 64 -S -p 80 --flood --rand-source 192.138.1.115} \\*
-que envia pacotes de 80 bytes (-d 80), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80).
-A flag --flood indica q o hping3 vai enviar os pacotes o mais rápido possível.
-a flag --rand-source indica que cada pacote vai ter um ip de origem aleatorio.
-192.138.1.115 é o alvo.\\*
-O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
-O wireshark armazena os pacotes no formato pcap, utilizamos a ferramenta tcpreplay para reproduzir o trafego com o comando:\\*
-\texttt{\# tcpreplay --pps 200 --intf1=wlp2s0 Datasets/syn-flood-sample.pcap}\\*
-Nesse caso, estamos enviando os pacotes armazenados em \texttt{Datasets/syn-flood-sample.pcap} a uma taxa de 200 pacotes por segundo (\texttt{-pps 200}) utilizando a interface wlp2s0 (\texttt{--intf1 wlp2s0})\\*
-Há diferenças entre os pacotes na opção ack.\\* 
-Por que 20?
-Apenas para gerar a regra mais rapidamente, poderia ser qualquer valor maior ou menor.
-No final, o parâmetro \texttt{count} da opção \texttt{threshold} vai ser alterado pelo adm da rede.\\*
-Intervalo de reprodução: 0.005 segundo
-
-\begin{table}[h]
-  \caption{\label{table:rules}Rule evolution}  
-  \centering
-  \begin{tabular}{lllll}
-    \toprule
-    \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
-    \midrule     
-    0 & (ack:0; seq:0; window:0; flags:F;)\\
-    5 & (window:0; flags:F;)\\
-    68 & (window:64; flags:F;)\\
-    73 & (window:64; flags:S,12;)\\
-    94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-}
-
-\section{Approach}
+\section{Technique}
 
@@
 \tname{} is a search-based approach that uses positive and negative
-network examples to synthesize signature-based NIDS rules, such as
-those from Snort and Suricata.
+examples to synthesize signature-based NIDS rules, such as those from
+Snort and Suricata. The technique synthesizes rules in two
+steps. First, \tname{} produces a potentially overspecified rule that
+captures the malicious traffic provided on input. The rule obtained in
+this stage can also capture benign traffic. In this step,
+\tname\ \emph{maximizes} the number of properties in the rule that
+captures the positive traffic. It extracts the tokens from the payload
+of the positive traffic (\ie{}, malicious traffic) and add them as
+content properties (see Section \Fix{xxx}). Furthermore, it searches
+for additional properties satisfied by the positive traffic. Any
+subset of properties found in this stage, by construction, captures
+the positive traffic. In the second step, \tname{} uses negative
+examples to \emph{minimize} the rule, discarding unnecessary
+options. The goal of \tname\ is to create a rule that only captures
+the positive traffic. The search is optimal (\ie{}, we reach this
+goal) if no benign message is captured during the minimization step.
 
-The technique synthesizes rules in two steps. First, it produces a
-potentially overspecified rule that captures the negative example
-provided on input. Then, it uses positive examples to eliminate
-unnecessary options from the rule. The goal of \tname\ is to create a
-rule that only captures the negative traffic.
+\Fix{...............estou aqui}
 
@@
 
+\Luc{entrada: utilizamos a ferramenta hping3 para executar o ataque em um alvo arbitrario dentro da rede (n ha necessidade de haver um alvo real para executar o hping3) e capturamos o trafego gerado com o wireshark.
+o comando utilizado foi:\\*
+\texttt{\# hping3 -d 80 -w 64 -S -p 80 --flood --rand-source 192.138.1.115} \\*
+que envia pacotes de 80 bytes (-d 80), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80).
+A flag --flood indica q o hping3 vai enviar os pacotes o mais rápido possível.
+a flag --rand-source indica que cada pacote vai ter um ip de origem aleatorio.
+192.138.1.115 é o alvo.\\*
+O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
+O wireshark armazena os pacotes no formato pcap, utilizamos a ferramenta tcpreplay para reproduzir o trafego com o comando:\\*
+\texttt{\# tcpreplay --pps 200 --intf1=wlp2s0 Datasets/syn-flood-sample.pcap}\\*
+Nesse caso, estamos enviando os pacotes armazenados em \texttt{Datasets/syn-flood-sample.pcap} a uma taxa de 200 pacotes por segundo (\texttt{-pps 200}) utilizando a interface wlp2s0 (\texttt{--intf1 wlp2s0})\\*
+Há diferenças entre os pacotes na opção ack.\\* 
+Por que 20?
+Apenas para gerar a regra mais rapidamente, poderia ser qualquer valor maior ou menor.
+No final, o parâmetro \texttt{count} da opção \texttt{threshold} vai ser alterado pelo adm da rede.\\*
+Intervalo de reprodução: 0.005 segundo
+
+\begin{table}[h]
+  \caption{\label{table:rules}Rule evolution}  
+  \centering
+  \begin{tabular}{lllll}
+    \toprule
+    \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
+    \midrule     
+    0 & (ack:0; seq:0; window:0; flags:F;)\\
+    5 & (window:0; flags:F;)\\
+    68 & (window:64; flags:F;)\\
+    73 & (window:64; flags:S,12;)\\
+    94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+}
+
 \Fix{any other detail we should consider?}

### 2019-10-21 paper/paper.tex
pass on technique

@@
 %
-\title{Synthesis of Rules for Network Intrusion Detection}
+\title{Synthesis of Rules for Network Intrusion Detectors}
 %
@@
 
-\section{Illustrative Example}
-\label{sec:suri-metas-coverage}
+\section{Background}
 
-\begin{figure}[t]
-  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
-  \caption{Suricata rule for SYN Flood Attacks.}
-  \label{fig:synflood-example}
-\end{figure}
-
-This section 1)~shows an example rule of Suricata, 2)~briefly explains
-the format of Suricata rules, 3) presents an example of a network
-attack, and 4) illustrates how \tname{} performs to synthesize the
-rule that would capture the attack.
+This paper focuses on signature-based \nids, Suricata, in
+particular. But the principles we use are general to any other
+signature-based \nids~(\eg{}, Snort~\cite{snort}).
 
@@
 
+\Fix{tabulate some important rules that we support}
+
+\section{Illustrative Examples}
+\label{sec:suri-metas-coverage}
+
+\begin{figure}[t]
+  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
+  \caption{Suricata rule for SYN Flood Attacks.}
+  \label{fig:synflood-example}
+\end{figure}
+
+This section briefly presents examples of attacks and rules that are
+synthesized with \tname{} to capture those attacks.
+
 \subsection{SYN Flood Attack}
@@
 
-\subsection{\tname\ on SYN Flood Attack}
+\subsubsection{\tname\ on SYN Flood Attack}
 
@@
 
+\Fix{repeat this for two other attacks. one non-dos/flood non-content}
+
 \section{Technique}
+
+\tname{} is a technique that uses positive and negative examples to
+synthesize signature-based NIDS rules, such as those from Snort and
+Suricata. The goal of \tname\ is to create a rule that \emph{only}
+captures the positive traffic, \ie{}, the malicious sequence of
+messages.
+
+\vspace{1ex}
+\noindent\textbf{Overview.}~\tname{} synthesizes rules in two
+steps. First, it produces a potentially overspecified rule that
+captures the malicious traffic provided on input. In this step,
+\tname\ \emph{maximizes} the number of options in the rule that
+captures the positive traffic. It extracts tokens from the malicious
+message and use them to define the initial rule. Furthermore, it
+searches for additional rule option (\eg, \Fix{xxx}) that are not
+found in the message data but are satisfied by the traffic. We say
+that this rule overfits the problem as it clearly satisfy the success
+criterion of only capturing the positive traffic, however it is unable
+to generalize to other attacks of the same kind with different
+data. In the second step, \tname{} uses negative examples to
+\emph{minimize} the rule, discarding unnecessary options. Note that
+any subset of options in the rule obtained in this stage captures the
+positive traffic by construction. \Fix{...elaborate...}
+
+\subsection{Step 1: Rule Maximization}
+\pgfplotsset{width=5.5cm,compat=1.8}
+
+The goal of the first step is to synthesize a rule with the maximum
+number of options that satisfy the malicious traffic. A rule is
+maximal if no other option can be added without making it miss the
+traffic.
+
+\begin{wrapfigure}[14]{r}{0.5\textwidth}
+  \centering
+  \vspace{-5ex}  
+  \begin{tikzpicture}
+    \begin{axis}[
+        ybar stacked,
+        enlargelimits=0.15,
+        legend style={at={(0.5,-0.15)},
+          anchor=north,legend columns=-1},
+        ylabel={\#rules},
+        symbolic x coords={gen, http, icmp, udp, tcp},
+        xtick=data,
+        nodes near coords,
+        nodes near coords align={vertical},
+      ]
+      \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
+      \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
+      \legend{relevant, irrelevant}
+    \end{axis}
+  \end{tikzpicture}
+  \caption{\label{fig:distribution-rules-protocol}Distribution of
+    rules per protocol and relevance for triggering action.}
+\end{wrapfigure}
+It is important to note that options can be specific to a given
+protocol or general and they can influence or not triggering an
+action. Figure~\ref{fig:distribution-rules-protocol} shows the
+distributions of these options per protocol and their relevance for
+triggering rule actions. This plot indicates that the problem search
+space is defined based on the protocol of the malicious traffic. For
+instance, if \tname{} detects that the attack involves \CodeIn{http}
+traffic, then 34 different rule options will be considered in the
+creation of the rule.
+
+\Fix{describe extraction...}
+
+\subsection{Step 2: Rule Minimization}
+
+Any rule created from a subset of rule options still captures the
+traffic as every rule option we infer in this step is independent and
+satisfied by the traffic.
+
+\Fix{...............estou aqui}
 
@@
 \end{figure*}
-
-\tname{} is a search-based approach that uses positive and negative
-examples to synthesize signature-based NIDS rules, such as those from
-Snort and Suricata. The technique synthesizes rules in two
-steps. First, \tname{} produces a potentially overspecified rule that
-captures the malicious traffic provided on input. The rule obtained in
-this stage can also capture benign traffic. In this step,
-\tname\ \emph{maximizes} the number of properties in the rule that
-captures the positive traffic. It extracts the tokens from the payload
-of the positive traffic (\ie{}, malicious traffic) and add them as
-content properties (see Section \Fix{xxx}). Furthermore, it searches
-for additional properties satisfied by the positive traffic. Any
-subset of properties found in this stage, by construction, captures
-the positive traffic. In the second step, \tname{} uses negative
-examples to \emph{minimize} the rule, discarding unnecessary
-options. The goal of \tname\ is to create a rule that only captures
-the positive traffic. The search is optimal (\ie{}, we reach this
-goal) if no benign message is captured during the minimization step.
-
-\Fix{...............estou aqui}
 
@@
 relevant for triggering some action of some rule.
-
-Options can be specific to a given protocol or general and they can
-influence or not triggering an
-action. Figure~\ref{fig:distribution-rules-protocol} shows the
-distributions of these options per protocol and their relevance for
-triggering rule actions.
-
-\pgfplotsset{width=5.5cm,compat=1.8}
-\begin{figure}[h!]
-  \centering
-  \begin{tikzpicture}
-    \begin{axis}[
-        ybar stacked,
-        enlargelimits=0.15,
-        legend style={at={(0.5,-0.15)},
-          anchor=north,legend columns=-1},
-        ylabel={\#rules},
-        symbolic x coords={gen, http, icmp, udp, tcp},
-        xtick=data,
-        nodes near coords,
-        nodes near coords align={vertical},
-      ]
-      \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
-      \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
-      \legend{relevant, irrelevant}
-    \end{axis}
-  \end{tikzpicture}
-  \caption{\label{fig:distribution-rules-protocol}Distribution of
-    rules per protocol and relevance for triggering action.}
-\end{figure}
 

### 2019-10-22 paper/paper.tex
pass on technique

@@
         nodes near coords,
+        every node near coord/.append style={font=\scriptsize},        
         nodes near coords align={vertical},
@@
 protocol or general and they can influence or not triggering an
-action. Figure~\ref{fig:distribution-rules-protocol} shows the
+action. The purpose of rule options \CodeIn{msg}, for instance, is to
+print on the output an informative message indicating what kind of
+intrusion has been
+observed. Figure~\ref{fig:distribution-rules-protocol} shows the
 distributions of these options per protocol and their relevance for
@@
 
-\Fix{describe extraction...}
+To infer options, \tname{} parses the protocol message, looking for
+fields in the message associated with options. For example, fields
+\Fix{...and...}  can be used to determine the option
+\Fix{...}. Unfortunately, inferring options at the field granularity
+level is insufficient to capture content options (see
+Section~\ref{sec:example-suricata-rules}), which refer to parts of the
+payload of the message and cannot be directly extracted directly from
+the fields. Handling this case is critically important as most rules
+contain content options. Figure~\ref{fig:distribution-contents} shows
+the histogram of number of content options per rule (up to 10) for the
+Suricata public rule database \Fix{cite}. Rules without content
+options is an exception and half of the rules contain at least two
+content options.
+\begin{wrapfigure}[12]{r}{0.5\textwidth}
+  \centering
+  \vspace{-5ex}  
+  \begin{tikzpicture}
+    \begin{axis}[
+        bar width=8pt,
+        scaled ticks=false,
+        tick label style={/pgf/number format/fixed},
+        ybar stacked,
+        enlargelimits=0.1,
+        legend style={at={(0.5,-0.15)},
+          anchor=north,legend columns=-1},
+        symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+        xtick=data,
+        nodes near coords,
+        every node near coord/.append style={font=\tiny},
+        nodes near coords align={vertical},
+      ]
+      \addplot coordinates {(0,25) (1,10222) (2,4956) (3,5589)
+        (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{\label{fig:distribution-contents}Histogram of number of
+    content options per rule (up to 10).}
+\end{wrapfigure}
+To find the values associated with these options, \tname{} splits the
+payload of the malicious message in tokens using natural language
+delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
+spaces. The rationale for this decision is that we observed, by
+inspecting existing rules, that content options typically refer to
+sequence of characters in the payload of messages separated by these
+delimiters. \Fix{describe extraction of options not in the
+  message. give examples! search only makes sense if this exists!}
 
@@
 
-\begin{figure*}[t]
-\centering
-\includegraphics[trim=10 360 182 130,clip,width=0.95\textwidth]{figs/nids-workflow}
-\caption{\tname\ workflow.}
-\label{fig:overview}
-\end{figure*}
+%% \begin{figure*}[t]
+%% \centering
+%% \includegraphics[trim=10 360 182 130,clip,width=0.95\textwidth]{figs/nids-workflow}
+%% \caption{\tname\ workflow.}
+%% \label{fig:overview}
+%% \end{figure*}
 

### 2019-10-22 paper/paper.tex
pass

--- 2019-10-21 paper/paper.tex
@@
 \subsection{Step 1: Rule Maximization}
-\pgfplotsset{width=5.5cm,compat=1.8}
+\pgfplotsset{width=6cm,compat=1.8}
+\pgfplotsset{every tick label/.append style={font=\tiny}}
 
@@
 number of options that satisfy the malicious traffic. A rule is
-maximal if no other option can be added without making it miss the
+maximal if no other option can be added without missing the malicious
 traffic.
 
-\begin{wrapfigure}[14]{r}{0.5\textwidth}
+\begin{wrapfigure}[16]{r}{0.5\textwidth}
   \centering
   \vspace{-5ex}
+  \scalebox{1.05}{
     \begin{tikzpicture}
@@
             anchor=north,legend columns=-1},
-        ylabel={\#rules},
+%          ylabel={\#rules},
           symbolic x coords={gen, http, icmp, udp, tcp},
@@
           nodes near coords,
+          every node near coord/.append style={font=\scriptsize},        
           nodes near coords align={vertical},
@@
     \end{tikzpicture}
+  }
   \caption{\label{fig:distribution-rules-protocol}Distribution of
-    rules per protocol and relevance for triggering action.}
+    number of rules per protocol and relevance for triggering action.}
 \end{wrapfigure}
@@
 protocol or general and they can influence or not triggering an
-action. Figure~\ref{fig:distribution-rules-protocol} shows the
+action. The purpose of rule options \CodeIn{msg}, for instance, is to
+print on the output an informative message indicating what kind of
+intrusion has been
+observed. Figure~\ref{fig:distribution-rules-protocol} shows the
 distributions of these options per protocol and their relevance for
-triggering rule actions. This plot indicates that the problem search
+triggering rule actions. We obtained this information
+by\Fix{...explicar como...} \Fix{eh um pouco suspeito ter tantas regras
+  irrelevantes...}This plot indicates that the problem search
 space is defined based on the protocol of the malicious traffic. For
-instance, if \tname{} detects that the attack involves \CodeIn{http}
+instance, if \tname{} detects by analyzing the format of network packets that the attack involves \CodeIn{http}
 traffic, then 34 different rule options will be considered in the
@@
 
-\Fix{describe extraction...}
+To infer options, \tname{} parses the protocol message, looking for
+fields in the message associated with options. For example, fields
+\Fix{...and...}  can be used to determine the option
+\Fix{...}. Unfortunately, inferring options at the field granularity
+level is insufficient to capture \CodeIn{content} options (see
+Section~\ref{sec:example-suricata-rules}), which refer to parts of the
+payload of the message and cannot be directly extracted directly from
+the fields. Handling this case is critically important as most rules
+contain \CodeIn{content} options. Figure~\ref{fig:distribution-contents} shows
+the histogram of number of these options per rule (up to 10) for the
+Suricata public rule database \Fix{cite}. Rules without \CodeIn{content}
+options is an exception and half of the rules contain at least two
+\CodeIn{content} options.
+\begin{wrapfigure}[14]{r}{0.56\textwidth}
+  \centering
+  \vspace{-3ex}
+  \scalebox{1.2}{  
+    \begin{tikzpicture}
+      \begin{axis}[
+          bar width=6pt,
+          scaled ticks=false,
+          tick label style={/pgf/number format/fixed},
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          xtick=data,
+          nodes near coords,
+          every node near coord/.append style={font=\tiny},
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(0,25) (1,10222) (2,4956) (3,5589)
+          (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \caption{\label{fig:distribution-contents}Histogram of number of
+    \CodeIn{content} options per rule (up to 10).}
+\end{wrapfigure}
+To find the values associated with these options, \tname{} splits the
+payload of the malicious message in tokens using natural language
+delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
+spaces. The rationale for this decision is that we observed, by
+inspecting existing rules, that \CodeIn{content} options typically refer to
+sequence of characters in the payload of messages separated by these
+delimiters. \Fix{describe extraction of options not in the
+  message. give examples! search only makes sense if this exists!}
 
@@
 
-Any rule created from a subset of rule options still captures the
-traffic as every rule option we infer in this step is independent and
-satisfied by the traffic.
+The rule reported by the first step satisfies the criterion to only
+capture positive traffic, however it is ``too'' specific---it is
+unable to capture different manifestations of the same attack as the
+rule would unlikely match exactly the options we found for the input
+message. Some options added to the rule during the first step are
+irrelevant for determining the attack and should be removed.
+
+The goal of this step is to create a \emph{minimal} rule that captures
+\emph{only} the positive malicious traffic. A rule is minimal if
+discarding any option results in capturing negative (\ie{}, benign)
+traffic. Recall that the previous maximization step produces a rule
+that captures the positive traffic and all options are satisifed by
+the positive traffic. Consequently, any subset of these options also
+captures the positive traffic. The minimization procedure we propose
+in this step builds on that observation. It iteratively discards
+options from the candidate rule until no other options can be
+removed. The procedure uses negative traffic to guide the
+search. Public databases of negative traffic with \Fix{thousands} of
+messages exist \Fix{cite cite} and can be used to increase the
+accuracy of the minimization.
 

### 2019-10-23 paper/paper.tex
supported options table

@@
 
-\Fix{tabulate some important rules that we support}
+\begin{table}[h]
+  \caption{\label{table:rules}Supported options}  
+  \centering
+  \begin{tabular}{lllll}
+    \toprule
+    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Option} &  \multicolumn{1}{c}{Description}\\
+    \midrule     
+    1 & dsize & matches on the size of the packet payload\\
+    2 & itype & matches on a specific ICMP type\\
+    3 & icode & matches on a specific ICMP code\\
+    4 & icmp\_seq & checks a ICMP sequence number\\
+    5 & icmp\_id & matches on specific ICMP id-values\\
+    6 & window & check for a specific TCP window size\\
+    7 & flags & matches on TCP flags\\
+    8 & fragbits & checks if the fragmentation or reserved bits are set in the IP header\\
+    9 & threshold & used to control the rule’s alert frequency\\
+    10 & content & matches on bytes of the payload\\
+    \bottomrule
+  \end{tabular}
+\end{table}
 

### 2019-10-23 paper/paper.tex
pass

--- 2019-10-22 paper/paper.tex
@@
 
-This paper focuses on signature-based \nids, Suricata, in
-particular. But the principles we use are general to any other
+This paper focuses on signature-based \nids. Although we focused on
+the Suricata \nids, the principles we used are general to any other
 signature-based \nids~(\eg{}, Snort~\cite{snort}).
@@
 the header. It is a semi-colon-separated sequence of key-value
-pairs. Some options serve to document the analyzed traffic (\eg{},
-``msg'', ``classtype'', and ``sid'') whereas the remaining properties
-describe characteristics of the attack (\eg, ``flags'', ``threshold'',
-and ``content''). The main challenge in synthesizing rules is to
-discover the set of options that make a rule to capture the malicious
-traffic, but ignore the benign traffic.
+pairs. Options serve to 1) document the analyzed traffic (\eg{},
+``msg'', ``classtype'', and ``sid'') and to 2) describe
+characteristics of the attack (\eg, ``flags'', ``threshold'', and
+``content''). \tname{} produces options of the second kind, which are
+\emph{relevant} to capturing the attack.
 
-\Fix{tabulate some important rules that we support}
+Suricata supports a total of \totoptions\ options covering
+\numproto\ distinct protocols. Of these total of options,
+\totoptionsrelevant\ are relevant to capture traffic (see
+Figure~\ref{fig:distribution-rules-protocol}). Table~\ref{table:rules}
+shows a sample of the relevant options of Suricata. Column ``\#''
+shows the option number, column ``Name'' shows the name of the
+option, column ``\#~Params'' shows the number of additional parameters
+associated with a given option, and column ``Description'' presents a
+short description of the option purpose.
+
+\begin{table}[h]
+  \caption{\label{table:rules}Example options.}  
+  \centering
+  \begin{tabular}{p{0.5cm}lrp{8cm}}
+    \toprule
+    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &
+    \multicolumn{1}{c}{\# Params} &  \multicolumn{1}{c}{Description}\\
+    \midrule     
+    1 & dsize & - & matches on the size of the packet payload\\
+    2 & itype & - & matches on a specific ICMP type\\
+    3 & icode & - & matches on a specific ICMP code\\
+    4 & icmp\_seq & - & checks a ICMP sequence number\\
+    5 & icmp\_id & - & matches on specific ICMP id-values\\
+    6 & window & - & checks the size of the TCP window\\
+    7 & flags & - & matches on TCP flags\\
+    8 & fragbits & - & checks if the fragmentation or reserved bits are set in the IP header\\
+    9 & threshold & - & controls the rule’s alert frequency\\
+    10 & content & - & matches on bytes of the payload\\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\subsection{Rules and Packets}
+
+\Fix{Illustrate how information is extracted from packets}
 
@@
 
-\subsection{SYN Flood Attack}
+\subsection{Active Reconnaissance}
 
-The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
+Active reconnaissance is a method used by malicious individuals to
+gather information about vulnerabilities from a computing system. The
+most common approach is to ``scan'' open ports to discover
+vulnerabilities. Active reconnaissance is a preliminary step towards
+an attack that actually exploits the system. A Ping Scan is a very
+common type of port scan based on the ICMP
+protocol. \Fix{...explain this attack...}
+
+\Fix{...explain how \tname\ works on it...}
+
+\subsection{Denial-of-Service}
+
+The SYN Flood attack is a denial-of-service attack that exploits a vulnerability in the TCP/IP handshake
 to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
@@
 legitimate requests cannot be satisfied due to resource exhaustion.
-Figure~\ref{fig:synflood-example} shows a Suricata rule to capture
-this attack. The critical parts appear in bold---the option
-\CodeIn{flags: S,12}, which identifies a SYN packet and the option
-\CodeIn{threshold: type both, track by\_dst, count 5000, seconds 5},
-indicating that a high volume of such packets should be requested in a
-short period of time.
+Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
+this attack. This rule was obtained from the \Fix{...what source...}
+The relevant options in the rule appear in bold---the option
+\CodeIn{flags: S,12}, identifies a SYN packet in a TCP packet and the
+option \CodeIn{threshold: type both, track by\_dst, count 5000,
+  seconds 5} indicates that a high volume of such packets should be
+requested in a short period of time.
 
-%\Mar{please, explain the other fields}
+%% Explain how \tname\ behaves on this attack
 
-\subsubsection{\tname\ on SYN Flood Attack}
+The main challenge in synthesizing rules is to discover the set of
+options that makes a rule to capture the malicious traffic, but
+ignores the benign traffic.
+
+\tname{} synthesizes rules with the action header fixed to
+\CodeIn{alert} and with the header \CodeIn{<proto> any any -> any
+  any}, indicating that it infers the target protocol and uses the
+most general flow description possible, \ie{}, the rule applies to the
+traffic flowing from/to any address and port. The rule that \tname{}
+synthesizes for the SYN Flood denial-of-service attack is as follows.
+
+\begin{figure}[h!]
+  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
+%  \caption{Suricata rule for SYN Flood Attacks.}
+%  \label{fig:synflood-example-synt}
+\end{figure}
 
@@
 
-\Fix{repeat this for two other attacks. one non-dos/flood non-content}
+\begin{table}[h]
+  \caption{\label{table:rules}Rule evolution}  
+  \centering
+  \begin{tabular}{lllll}
+    \toprule
+    \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
+    \midrule     
+    0 & (ack:0; seq:0; window:0; flags:F;)\\
+    5 & (window:0; flags:F;)\\
+    68 & (window:64; flags:F;)\\
+    73 & (window:64; flags:S,12;)\\
+    94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
 
-\subsection{Problem Characterization}
-
-One important step in search-based optimization is to characterize a
-candidate solution to the target problem. Candidate solutions are the
-objects that the search optimize towards solutions. In this case, a 
-candidate solution is an encoding of a Suricata rule.  \tname\ encodes
-the rule options segment of a Suricata rule (see
-Section~\ref{sec:example-suricata-rules}) as a vector of Option
-objects. Some options are discarded as they only serve for
-documentation. For example, the options \CodeIn{msg},
-\CodeIn{classtype}, and \CodeIn{misc-activity}, appearing in the
-example on Figure~\ref{fig:synflood-example}, have no influence in
-triggering the actions associated with the rule. The option
-\CodeIn{flags}, however, is relevant to trigger the action. \tname{}
-models that option with a data structure storing two fields, one of
-character type and another of integer type. This enables \tname{} to
-mutate each element separately. Similar encodings are used in modeling
-the other relevant options. We found what options are (ir)relevant by
-inspecting Suricata code for the presence of \CodeIn{eval} functions,
-which are invoked from a main loop to process each action rule. The
-presence of such a function for a given option indicates that it is
-relevant for triggering some action of some rule.
-
-\subsection{Initial State}
-
-To bootstrap the search, \tname{} needs to select an initial
-solution. When creating such candidate solution, \tname{} first looks
-for the protocol related to the malicious input traffic. Then, it
-selects the relevant options for that protocol and initializes the
-option vector with default values. \Mar{Suricata define valores
-  default?  Vcs. definiram?  Isto foi baseado em popularidade ou o
-  que?}
-
-%% \tname{}
-%% uses an abstract representation of a Suricata rule that reflects the
-%% concrete representation of the rule. The concrete rule representation
-%% is the one actually used by Suricata whereas the abstract
-%% representation is the one \tname{} uses to run the search.
-
-\subsection{Fitness Functions}
-
-The fitness value of a candidate solution $c$, modeling the rule
-options $o_1, \dots, o_n$ is given by $f(c)=1-\Sigma{d(o_i,m)}/n$,
-where $i$ ranges from 1 to $n$ and $m$ denotes the negative
-traffic. The higher the fitness value the better, \ie{}, the fitter
-the candidate solution $c$ will be.  The value of the expression
-$d(o_i,m)$, associated with option $o_i$, ranges over
-0-1. Consequently, the value of the expression $\Sigma{d(o_i,m)}/n$
-itself is in the 0-1 range, as there are $n$ options, and the range of
-the function is 0-1.
-
-As usual, we used customized distance functions to compute $d(o_i,
-m)$. For example, let us consider the \CodeIn{flags} option, discussed
-on Figure~\ref{fig:synflood-example}. Recall that this option
-identifies the SYN packet in the tcp protocol and that its encoding in
-\tname{} stores two attributes---one categorical and one
-numerical. The value of $d(o_i, m)$, in this case, is obtained by
-computing the normalized Euclidean distance in the two-dimensional
-space. Considering that each of the two dimensions has cardinality of
-ten, the distance of a candidate solution holding the option
-\CodeIn{flags(S,13)} to the solution will be 0.22
-(=$1/\sqrt{20}$). Note that the smaller the distance the better.
-
-\tname{} uses a modified version of Suricata to measure distance. It
-updates the rule set that the modified version of Suricata analyses,
-injects the traffic on the network, and monitors the execution. Each
-rule option has an associated \CodeIn{eval} function in code. \tname{}
-monitors the execution of these functions to identify how far it is
-from returning true, which indicates it is closer to trigger the
-action associated with the rule. 
-\Luc{Originalmente, o suricata interrompe a checagem de uma regra quando alguma das opções não dá match com o pacote analisado.
-O código fonte foi alterado para que as regras sempre tenham todas as suas opções verificadas.
-A função de verificação de uma opção retorna um booleano indicando se o valor da opção na regra corresponde ou não ao valor da opção no pacote.
-Para opções cujos valores são numéricos, alteramos o código fonte para obter a diferença entre os valores da opção na regra e no pacote.
-\\*
-Exemplo trecho de código original opção "icode":
-
-\begin{figure}[h]
-  \lstinputlisting[language=C,numbers=none,keywords={}]{original.code}
-  \caption{Original source code.}
-  \label{fig:original-code}
-\end{figure}
-
-Trecho modificado:
-
-\begin{figure}[h]
-  \lstinputlisting[language=C,numbers=none,keywords={}]{modified.code}
-  \caption{Modified source code.}
-  \label{fig:modified-code}
-\end{figure}
-
-}
-
-\Fix{Lucas/Guilherme, it would be good to show how this looks like in
-  code}
-
-\subsection{Search}
-
-\tname{} is a single-individual hill-climbing like search.
-
 \subsection{Limitations}
@@
 
-\begin{table}[h]
-  \caption{\label{table:rules}Rule evolution}  
-  \centering
-  \begin{tabular}{lllll}
-    \toprule
-    \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
-    \midrule     
-    0 & (ack:0; seq:0; window:0; flags:F;)\\
-    5 & (window:0; flags:F;)\\
-    68 & (window:64; flags:F;)\\
-    73 & (window:64; flags:S,12;)\\
-    94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
 }

### 2019-10-23 paper/paper.tex
pass on intro+background

@@
   \centering
-  \begin{tabular}{p{0.5cm}lrp{8cm}}
+  \begin{tabular}{p{0.35cm}ll}
     \toprule
-    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &
-    \multicolumn{1}{c}{\# Params} &  \multicolumn{1}{c}{Description}\\
+    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &  \multicolumn{1}{c}{Description}\\
     \midrule     
-    1 & dsize & - & matches on the size of the packet payload\\
-    2 & itype & - & matches on a specific ICMP type\\
-    3 & icode & - & matches on a specific ICMP code\\
-    4 & icmp\_seq & - & checks a ICMP sequence number\\
-    5 & icmp\_id & - & matches on specific ICMP id-values\\
-    6 & window & - & checks the size of the TCP window\\
-    7 & flags & - & matches on TCP flags\\
-    8 & fragbits & - & checks if the fragmentation or reserved bits are set in the IP header\\
-    9 & threshold & - & controls the rule’s alert frequency\\
-    10 & content & - & matches on bytes of the payload\\
+    1 & dsize & matches on the size of the packet payload\\
+    2 & itype &  matches on a specific ICMP type\\
+    3 & icode & matches on a specific ICMP code\\
+    4 & icmp\_seq  & checks a ICMP sequence number\\
+    5 & icmp\_id & matches on specific ICMP id-values\\
+    6 & window & checks the size of the TCP window\\
+    7 & flags & matches on TCP flags\\
+    8 & fragbits & checks if the fragmentation or reserved bits are set in the IP header\\
+    9 & threshold & controls the rule’s alert frequency\\
+    10 & content & matches on bytes of the payload\\
     \bottomrule
@@
 synthesized with \tname{} to capture those attacks.
-
-\subsection{Active Reconnaissance}
-
-Active reconnaissance is a method used by malicious individuals to
-gather information about vulnerabilities from a computing system. The
-most common approach is to ``scan'' open ports to discover
-vulnerabilities. Active reconnaissance is a preliminary step towards
-an attack that actually exploits the system. A Ping Scan is a very
-common type of port scan based on the ICMP
-protocol. \Fix{...explain this attack...}
-
-\Fix{...explain how \tname\ works on it...}
 
@@
 
-The main challenge in synthesizing rules is to discover the set of
-options that makes a rule to capture the malicious traffic, but
-ignores the benign traffic.
+In the following, we illustrate how \tname{} proceeds for obtaining
+the rule for the SYN Flood attack. The main challenge in synthesizing
+rules is to discover the set of options that enables the \nids\ to
+capture the malicious traffic, but ignore the benign traffic.  The
+other parts of the rule (\ie{}, the action and the header) are either
+inferred or left constant. \tname{} sets the action header to
+\CodeIn{alert} and sets the header to \CodeIn{<proto> any any -> any
+  any}, indicating that it infers the target protocol, and it uses the
+most general flow description possible, \ie{}, analyzing the traffic
+flowing from/to any address and port.  We conjecture that the system
+administrator would know how to modify the constant parts.
 
-\tname{} synthesizes rules with the action header fixed to
-\CodeIn{alert} and with the header \CodeIn{<proto> any any -> any
-  any}, indicating that it infers the target protocol and uses the
-most general flow description possible, \ie{}, the rule applies to the
-traffic flowing from/to any address and port. The rule that \tname{}
-synthesizes for the SYN Flood denial-of-service attack is as follows.
+\tname{} starts the search for rule options for the SYN Flood attack
+from the initial set of options appearing below.
 
+\Fix{atualize isto para o valor correto, por favor. $\rightarrow$}
 \begin{figure}[h!]
+  \vspace{-3ex}
   \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
@@
   %  \label{fig:synflood-example-synt}
+  \vspace{-3ex}  
 \end{figure}
 
-\Mar{Guilherme/Lucas (CRI): we need to show how \tname{} works on this
-  example. What is the input? How long does it take to generate the
-  rule? Is the rule generated exactly the same? Please explain...}
+This set is based on an analysis of the packets from the positive
+traffic. Consequently, a rule that uses this set of options would
+capture the positive traffic by construction. Unfortunately, such rule
+would be overspecified, \ie{}, the rule would unlikely capture the
+same attacks as the traffic data would not be the same and \tname{}
+includes in the set all possible options that can be inferred from the
+data. To address that, \tname{} minimizes the rule with the goal of
+finding the minimal set of rules that only captures the positive
+traffic, hence mitigating the overfitting problem. \tname{}
+potentially finds several minimal rules. The list below shows the
+options for the five shortest rules it produces for this attack.
 
-\begin{table}[h]
-  \caption{\label{table:rules}Rule evolution}  
-  \centering
-  \begin{tabular}{lllll}
-    \toprule
-    \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
-    \midrule     
-    0 & (ack:0; seq:0; window:0; flags:F;)\\
-    5 & (window:0; flags:F;)\\
-    68 & (window:64; flags:F;)\\
-    73 & (window:64; flags:S,12;)\\
-    94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
-    \bottomrule
-  \end{tabular}
-\end{table}
+\Fix{...elaborate...}
+
+\subsection{Active Reconnaissance}
+
+Active reconnaissance is a method used by malicious individuals to
+gather information about vulnerabilities from a computing system. The
+most common approach is to ``scan'' open ports to discover
+vulnerabilities. Active reconnaissance is a preliminary step towards
+an attack that actually exploits the system. A Ping Scan is a very
+common type of port scan based on the ICMP
+protocol. \Fix{...explain this attack...}
+
+\Fix{...explain how \tname\ works on it...}
+
+\subsection{\Fix{...some other with contents...}}
+
+\Fix{...elaborate...}
+
+%% \begin{table}[h]
+%%   \caption{\label{table:rules}Rule evolution}  
+%%   \centering
+%%   \begin{tabular}{lllll}
+%%     \toprule
+%%     \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
+%%     \midrule     
+%%     0 & (ack:0; seq:0; window:0; flags:F;)\\
+%%     5 & (window:0; flags:F;)\\
+%%     68 & (window:64; flags:F;)\\
+%%     73 & (window:64; flags:S,12;)\\
+%%     94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
+%%     \bottomrule
+%%   \end{tabular}
+%% \end{table}
 
 \section{Technique}
+
+\Fix{---------------------------------------------------
+  estou aqui}
 

### 2019-10-23 paper/paper.tex
pass on examples and technique
pass

@@
 Suricata supports a total of \totoptions\ options covering
-\numproto\ distinct protocols. Of these total of options,
-\totoptionsrelevant\ are relevant to capture traffic (see
-Figure~\ref{fig:distribution-rules-protocol}). Table~\ref{table:rules}
+\numproto\ distinct protocols. Of these \totoptions,
+\totoptionsrelevant\ options are relevant to capture traffic
+\begin{wrapfigure}[17]{r}{0.52\textwidth}
+  \centering
+  \vspace{-4ex}
+  \scalebox{.75}{
+    \begin{tikzpicture}
+      \begin{axis}[
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+%          ylabel={\#rules},
+          symbolic x coords={gen, http, icmp, udp, tcp},
+          xtick=data,
+          nodes near coords,
+          every node near coord/.append style={font=\scriptsize},        
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
+        \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
+        \legend{relevant, irrelevant}
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \caption{\label{fig:distribution-rules-protocol}Distribution of
+    number of options per protocol and relevance for triggering action.}
+\end{wrapfigure}
+Options can be specific
+to a given protocol or general and they can influence or not
+triggering an action. The purpose of rule option \CodeIn{msg}, for
+instance, is to print on the output an informative message indicating
+what kind of intrusion has been
+observed. Figure~\ref{fig:distribution-rules-protocol} shows the
+distributions of these options per protocol and their relevance for
+triggering rule actions. We obtained this information
+by\Fix{...explicar como...} \Fix{eh um pouco suspeito ter tantas
+  regras irrelevantes...}
+
+%% This plot indicates that the problem search
+%% space is defined based on the protocol of the malicious traffic. For
+%% instance, if \tname{} detects by analyzing the format of network
+%% packets that the attack involves \CodeIn{http} traffic, then 34
+%% different rule options will be considered in the creation of the rule.
+
+Table~\ref{table:rules}
 shows a sample of the relevant options of Suricata. Column ``\#''
@@
 associated with a given option, and column ``Description'' presents a
-short description of the option purpose.
+short description of the option purpose. \Fix{...explain content...}
 
@@
 
-\begin{figure}[t]
-  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
-  \caption{Suricata rule for SYN Flood Attacks.}
-  \label{fig:synflood-example}
-\end{figure}
-
 This section briefly presents examples of attacks and rules that are
@@
 legitimate requests cannot be satisfied due to resource exhaustion.
+
+\begin{figure}[h!]
+  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
+  \caption{Suricata rule for SYN Flood Attacks.}
+  \label{fig:synflood-example}
+\end{figure}
+
 Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
@@
 
-\Fix{---------------------------------------------------
-  estou aqui}
-
 \tname{} is a technique that uses positive and negative examples to
@@
 
+\begin{figure*}[t]
+\centering
+\includegraphics[trim=0 300 350 0,clip,width=0.95\textwidth]{figs/nids-workflow}
+\caption{\tname\ workflow.}
+\label{fig:overview}
+\end{figure*}
+
 \vspace{1ex}
-\noindent\textbf{Overview.}~\tname{} synthesizes rules in two
-steps. First, it produces a potentially overspecified rule that
-captures the malicious traffic provided on input. In this step,
-\tname\ \emph{maximizes} the number of options in the rule that
-captures the positive traffic. It extracts tokens from the malicious
-message and use them to define the initial rule. Furthermore, it
-searches for additional rule option (\eg, \Fix{xxx}) that are not
-found in the message data but are satisfied by the traffic. We say
-that this rule overfits the problem as it clearly satisfy the success
-criterion of only capturing the positive traffic, however it is unable
-to generalize to other attacks of the same kind with different
-data. In the second step, \tname{} uses negative examples to
-\emph{minimize} the rule, discarding unnecessary options. Note that
-any subset of options in the rule obtained in this stage captures the
-positive traffic by construction. \Fix{...elaborate...}
+\noindent\textbf{Overview.}~ The input of the technique is the
+positive traffic whereas the output is a list of minimized rules
+capturing only the positive, malicious,
+traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
+pipeline of two components, which appear numbered in the figure.
+\tname{} synthesizes rules in two steps. First, it produces a
+potentially overspecified rule that captures the malicious traffic
+provided on input. In this step, \tname\ finds options by
+\emph{reverse engineering} the positive traffic. It extracts tokens
+from the malicious message and uses them to define the initial
+rule. We say that this rule overfits the problem as it satisfies the
+success criterion of only capturing the positive traffic, however it
+is unlikely that it will be able to generalize to other attacks (say,
+attacks of the same kind, but with different data). The reason is that
+many options added to the rule at the first step are not important for
+discriminating the attack. In the second step, \tname{} uses negative
+examples to \emph{minimize} the rule, discarding unnecessary
+options. Note that any subset of the original set of options captures
+the positive traffic. The minimization step uses a population-based
+gradient descent-like search to generate minimal solutions.
 
-\subsection{Step 1: Rule Maximization}
+%% The goal of the first
+%% component is to produce a rule that captures the negative traffic.
+%% First, it initializes the rule with random values. Only options
+%% associated with the used protocol are included in the rule
+%% representation. Then, \tname\ optimizes the rule options until
+%% \suri\ is able to capture the attack associated with the negative
+%% traffic.  \tname\ unsuccessfully terminates at this point if it cannot
+%% capture the attack. The second component takes as input the rule
+%% produced by the first component and tries to minimize that rule. Note
+%% that any subset of options would capture the negative traffic---as all
+%% options in the rule need to be satisfied for the negative traffic to
+%% be captured---but it can also capture positive traffic. This component
+%% systematically discards options from the rule encoding until no
+%% positive traffic is captured.
+
+\subsection{Step 1: Reverse Engineering}
 \pgfplotsset{width=6cm,compat=1.8}
@@
 
-The goal of the first step is to synthesize a rule with the maximum
-number of options that satisfy the malicious traffic. A rule is
-maximal if no other option can be added without missing the malicious
-traffic.
+\Fix{--------------------------- estou aqui}
 
-\begin{wrapfigure}[16]{r}{0.5\textwidth}
-  \centering
-  \vspace{-5ex}
-  \scalebox{1.05}{
-    \begin{tikzpicture}
-      \begin{axis}[
-          ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
-%          ylabel={\#rules},
-          symbolic x coords={gen, http, icmp, udp, tcp},
-          xtick=data,
-          nodes near coords,
-          every node near coord/.append style={font=\scriptsize},        
-          nodes near coords align={vertical},
-        ]
-        \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
-        \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
-        \legend{relevant, irrelevant}
-      \end{axis}
-    \end{tikzpicture}
-  }
-  \caption{\label{fig:distribution-rules-protocol}Distribution of
-    number of rules per protocol and relevance for triggering action.}
-\end{wrapfigure}
-It is important to note that options can be specific to a given
-protocol or general and they can influence or not triggering an
-action. The purpose of rule options \CodeIn{msg}, for instance, is to
-print on the output an informative message indicating what kind of
-intrusion has been
-observed. Figure~\ref{fig:distribution-rules-protocol} shows the
-distributions of these options per protocol and their relevance for
-triggering rule actions. We obtained this information
-by\Fix{...explicar como...} \Fix{eh um pouco suspeito ter tantas regras
-  irrelevantes...}This plot indicates that the problem search
-space is defined based on the protocol of the malicious traffic. For
-instance, if \tname{} detects by analyzing the format of network packets that the attack involves \CodeIn{http}
-traffic, then 34 different rule options will be considered in the
-creation of the rule.
-
-To infer options, \tname{} parses the protocol message, looking for
+The goal of the first step is to extract options from the positive
+traffic. To infer options, \tname{} parses the protocol message, looking for
 fields in the message associated with options. For example, fields
@@
 accuracy of the minimization.
-
-\Fix{...............estou aqui}
-
-%% \begin{figure*}[t]
-%% \centering
-%% \includegraphics[trim=10 360 182 130,clip,width=0.95\textwidth]{figs/nids-workflow}
-%% \caption{\tname\ workflow.}
-%% \label{fig:overview}
-%% \end{figure*}
-
-Figure~\ref{fig:overview} shows the workflow of \tname{} as a pipeline
-of two components, appearing numbered in the figure. Each component
-implements one of the steps mentioned above. The goal of the first
-component is to produce a rule that captures the negative traffic.
-First, it initializes the rule with random values. Only options
-associated with the used protocol are included in the rule
-representation. Then, \tname\ optimizes the rule options until
-\suri\ is able to capture the attack associated with the negative
-traffic.  \tname\ unsuccessfully terminates at this point if it cannot
-capture the attack. The second component takes as input the rule
-produced by the first component and tries to minimize that rule. Note
-that any subset of options would capture the negative traffic---as all
-options in the rule need to be satisfied for the negative traffic to
-be captured---but it can also capture positive traffic. This component
-systematically discards options from the rule encoding until no
-positive traffic is captured.
 

### 2019-10-24 paper/paper.tex
updating background and technique

@@
 
-Suricata supports a total of \totoptions\ options covering
-\numproto\ distinct protocols. Of these \totoptions,
-\totoptionsrelevant\ options are relevant to capture traffic
-\begin{wrapfigure}[17]{r}{0.52\textwidth}
+Suricata supports a total of \totoptions\ distinct options covering
+\numproto\ distinct protocols. A total of 9 of these options are for
+documentation.
+\begin{wrapfigure}[16]{r}{0.5\textwidth}
   \centering
-  \vspace{-4ex}
-  \scalebox{.75}{
+  \vspace{-5ex}
+  \scalebox{0.7}{
     \begin{tikzpicture}
@@
           %          ylabel={\#rules},
-          symbolic x coords={gen, http, icmp, udp, tcp},
+          symbolic x coords={dns, ftp, http, icmp, ip, smb, smtp, ssh,
+          tcp, tls, udp},
           xtick=data,
+          xticklabel style={rotate=45},          
           nodes near coords,
@@
         ]
-        \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
-        \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
-        \legend{relevant, irrelevant}
+        \addplot coordinates {(dns,10) (ftp,11) (http,20) (icmp,18)(ip,16) (smb,13) (smtp,11) (ssh, 10) (tcp, 32) (tls,18) (udp, 18)};
+%        \legend{relevant, irrelevant}
       \end{axis}
@@
   }
-  \caption{\label{fig:distribution-rules-protocol}Distribution of
-    number of options per protocol and relevance for triggering action.}
+  \caption{\label{fig:distribution-rules-protocol}Number of relevant options per protocol.}
 \end{wrapfigure}
-Options can be specific
-to a given protocol or general and they can influence or not
-triggering an action. The purpose of rule option \CodeIn{msg}, for
-instance, is to print on the output an informative message indicating
-what kind of intrusion has been
-observed. Figure~\ref{fig:distribution-rules-protocol} shows the
-distributions of these options per protocol and their relevance for
-triggering rule actions. We obtained this information
-by\Fix{...explicar como...} \Fix{eh um pouco suspeito ter tantas
-  regras irrelevantes...}
-
-%% This plot indicates that the problem search
-%% space is defined based on the protocol of the malicious traffic. For
-%% instance, if \tname{} detects by analyzing the format of network
-%% packets that the attack involves \CodeIn{http} traffic, then 34
-%% different rule options will be considered in the creation of the rule.
-
-Table~\ref{table:rules}
-shows a sample of the relevant options of Suricata. Column ``\#''
-shows the option number, column ``Name'' shows the name of the
-option, column ``\#~Params'' shows the number of additional parameters
-associated with a given option, and column ``Description'' presents a
-short description of the option purpose. \Fix{...explain content...}
+The purpose of rule option \CodeIn{msg}, for instance, is to print on
+the output an informative message indicating what kind of intrusion
+has been observed. Figure~\ref{fig:distribution-rules-protocol} shows
+the distributions of these options per protocol. Note that the sum of
+the number of protocol options is higher than \totoptions\ as some
+options can be used by multiple protocols. Table~\ref{table:rules}
+shows a sample of the options of Suricata. Column ``\#'' shows the
+option number, column ``Name'' shows the name of the option, column
+``\#~Params'' shows the number of additional parameters associated
+with a given option, and column ``Description'' presents a short
+description of the option purpose. \Fix{...explain content...}
 
@@
 \Fix{atualize isto para o valor correto, por favor. $\rightarrow$}
+
 \begin{figure}[h!]

### 2019-10-24 paper/paper.tex
pass

@@
 documentation.
-\begin{wrapfigure}[16]{r}{0.5\textwidth}
+\begin{wrapfigure}[15]{r}{0.5\textwidth}
   \centering
@@
   }
-  \caption{\label{fig:distribution-rules-protocol}Number of relevant options per protocol.}
+  \vspace{-1ex}
+  \caption{\label{fig:distribution-rules-protocol}Number of relevant
+    options per protocol supported by Suricata.}
 \end{wrapfigure}
-The purpose of rule option \CodeIn{msg}, for instance, is to print on
+\Comment{For instance, the purpose of rule option \CodeIn{msg} is to print on
 the output an informative message indicating what kind of intrusion
-has been observed. Figure~\ref{fig:distribution-rules-protocol} shows
-the distributions of these options per protocol. Note that the sum of
+has been observed.} Figure~\ref{fig:distribution-rules-protocol} shows
+the distributions of these options per protocol. The sum of
 the number of protocol options is higher than \totoptions\ as some
-options can be used by multiple protocols. Table~\ref{table:rules}
-shows a sample of the options of Suricata. Column ``\#'' shows the
-option number, column ``Name'' shows the name of the option, column
-``\#~Params'' shows the number of additional parameters associated
-with a given option, and column ``Description'' presents a short
-description of the option purpose. \Fix{...explain content...}
+options is shared across different protocols.
+
+Table~\ref{table:rules} shows a sample of the options of
+Suricata. Column ``\#'' shows the option number, column ``Name'' shows
+the name of the option, column ``\#~Params'' shows the number of
+additional parameters associated with a given option, and column
+``Description'' presents a short description of the option purpose. We
+found that the option \CodeIn{content} is particularly prevalent in
+rules from public databases (see
+Figure~\ref{fig:distribution-contents}). This option takes a sequence
+of bytes as argument (\eg{}, \CodeIn{content:''SELECT ''}) and checks
+if that sequence is present in the payload of the message. Detailed
+description of the Suricata rules can be found
+elsewhere~\cite{suri-rule-format}.
 
 \begin{table}[h]
-  \caption{\label{table:rules}Example options.}  
+  \caption{\label{table:rules}Example options of Suricata.}  
   \centering
@@
     9 & threshold & controls the rule’s alert frequency\\
-    10 & content & matches on bytes of the payload\\
+    10 & content & checks if argument is present on the payload\\
     \bottomrule
@@
 \subsection{Rules and Packets}
+\label{sec:rules-and-packets}
 
@@
 \Fix{atualize isto para o valor correto, por favor. $\rightarrow$}
-
-\begin{figure}[h!]
+\begin{figure}[h]
   \vspace{-3ex}
@@
 capture the positive traffic by construction. Unfortunately, such rule
-would be overspecified, \ie{}, the rule would unlikely capture the
-same attacks as the traffic data would not be the same and \tname{}
-includes in the set all possible options that can be inferred from the
-data. To address that, \tname{} minimizes the rule with the goal of
-finding the minimal set of rules that only captures the positive
-traffic, hence mitigating the overfitting problem. \tname{}
-potentially finds several minimal rules. The list below shows the
-options for the five shortest rules it produces for this attack.
+would be overspecified. It would unlikely capture different
+manifestations of the same attack. Informally, the rule includes ``too
+many'' characteristics of the malicious traffic used to define it, but
+new attacks may not use the same data. To address that, \tname{}
+minimizes the rule with the goal of finding the minimal set of rules
+that only captures the positive traffic, hence mitigating the
+overfitting problem. \tname{} potentially finds several minimal
+rules. The list below shows the options for the five shortest rules it
+produces for this attack.
 
@@
 \subsection{\Fix{...some other with contents...}}
+\label{sec:content-example}
 
@@
 success criterion of only capturing the positive traffic, however it
-is unlikely that it will be able to generalize to other attacks (say,
-attacks of the same kind, but with different data). The reason is that
-many options added to the rule at the first step are not important for
+is unlikely that it will be able to generalize to other attacks,
+\ie{}, attacks of the same kind, but with different data. The reason
+is that many options added to the rule are unimportant for
 discriminating the attack. In the second step, \tname{} uses negative
@@
 
-\Fix{--------------------------- estou aqui}
-
-The goal of the first step is to extract options from the positive
-traffic. To infer options, \tname{} parses the protocol message, looking for
-fields in the message associated with options. For example, fields
-\Fix{...and...}  can be used to determine the option
-\Fix{...}. Unfortunately, inferring options at the field granularity
-level is insufficient to capture \CodeIn{content} options (see
-Section~\ref{sec:example-suricata-rules}), which refer to parts of the
-payload of the message and cannot be directly extracted directly from
-the fields. Handling this case is critically important as most rules
-contain \CodeIn{content} options. Figure~\ref{fig:distribution-contents} shows
-the histogram of number of these options per rule (up to 10) for the
-Suricata public rule database \Fix{cite}. Rules without \CodeIn{content}
-options is an exception and half of the rules contain at least two
-\CodeIn{content} options.
 \begin{wrapfigure}[14]{r}{0.56\textwidth}
   \centering
-  \vspace{-3ex}
+  \vspace{-8ex}
   \scalebox{1.2}{  
@@
 \end{wrapfigure}
-To find the values associated with these options, \tname{} splits the
-payload of the malicious message in tokens using natural language
-delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
-spaces. The rationale for this decision is that we observed, by
-inspecting existing rules, that \CodeIn{content} options typically refer to
-sequence of characters in the payload of messages separated by these
-delimiters. \Fix{describe extraction of options not in the
-  message. give examples! search only makes sense if this exists!}
+The goal of the first step is to extract options from the positive
+traffic. For that, \tname{} parses the protocol message, looking for
+fields in the message associated with options (see
+Section~\ref{sec:rules-and-packets}). For example, fields
+\Fix{...and...}  can be used to determine the option
+\Fix{zzz}. Unfortunately, inferring options at the field granularity
+level is insufficient to capture \CodeIn{content} options (see
+Section~\ref{sec:example-suricata-rules}), which refer to parts of the
+payload of the message and cannot be directly extracted directly from
+the fields. Handling this case is critically important as most rules
+contain \CodeIn{content}
+options. Figure~\ref{fig:distribution-contents} shows the histogram of
+number of these options per rule (up to 10) for the Suricata public
+rule database \Fix{cite}. Rules without \CodeIn{content} options is an
+exception and half of the rules contain at least two \CodeIn{content}
+options.  To find the values associated with these options, \tname{}
+splits the payload of the malicious message in tokens using natural
+language delimiters, such as $\backslash$t, $\backslash$n,
+$\backslash$r, and spaces. The rationale for this decision is that we
+observed, by inspecting existing rules, that \CodeIn{content} options
+typically refer to sequence of characters in the payload of messages
+separated by these delimiters. This tokenization step can generate
+dozens of options. For the example from
+Section~\ref{sec:content-example}, \tname{} produces a total of
+\Fix{xx} content options.
 
 \subsection{Step 2: Rule Minimization}
+
+\Fix{--------------------------- estou aqui}
 

### 2019-10-24 paper/paper.tex
pass on multiple parts

@@
 as attackers learn new strategies to circumvent existing defenses.
-This paper proposes \tname{}, a technique that automatically
+This paper proposes \tname{}\footnote{Abbreviation of \textbf{Sy}nthesis of
+  Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a technique that automatically
 synthesizes \nids\ rules from positive and negative examples (\ie{},
@@
 \CodeIn{alert} and sets the header to \CodeIn{<proto> any any -> any
-  any}, indicating that it infers the target protocol, and it uses the
-most general flow description possible, \ie{}, analyzing the traffic
-flowing from/to any address and port.  We conjecture that the system
-administrator would know how to modify the constant parts.
+  any}, indicating that it infers the target protocol and uses the
+most general flow description possible, \ie{}, it instructs the
+\nids\ to analyze the traffic flowing from/to any address and port.
+We understand that this is domain-specific information and conjecture
+that the system administrator would know how to modify the constant
+parts.
 
-\tname{} starts the search for rule options for the SYN Flood attack
-from the initial set of options appearing below.
+\tname{} proceeds as follows to synthesize rules for the SYN Flood
+attack. First, it analyzes the positive traffic and extracts options
+expressed in that traffic. For this example, it finds the following
+set of options:
 
@@
 
-This set is based on an analysis of the packets from the positive
-traffic. Consequently, a rule that uses this set of options would
-capture the positive traffic by construction. Unfortunately, such rule
-would be overspecified. It would unlikely capture different
-manifestations of the same attack. Informally, the rule includes ``too
-many'' characteristics of the malicious traffic used to define it, but
-new attacks may not use the same data. To address that, \tname{}
-minimizes the rule with the goal of finding the minimal set of rules
-that only captures the positive traffic, hence mitigating the
-overfitting problem. \tname{} potentially finds several minimal
-rules. The list below shows the options for the five shortest rules it
-produces for this attack.
+Note that a rule that uses this set of options would capture the
+positive traffic as these options originated from the positive
+traffic. Unfortunately, such rule would be overspecified. It would
+unlikely capture different manifestations of the same
+attack. Informally, the rule includes ``too many'' characteristics of
+the malicious traffic used to define it. To address this overfitting
+issue, \tname{} minimizes the rule preserving the invariant that
+plausible rules should only capture the positive traffic. \tname{}
+potentially finds several minimal rules. The list below shows the
+options for the five shortest rules it produces for this attack.
 

### 2019-10-24 paper/paper.tex
pass on minimization procedure

@@
 %
+%\usepackage[margin=1in]{geometry}% http://ctan.org/pkg/geometry
+\usepackage{lipsum}% http://ctan.org/pkg/lipsum
+\usepackage{algorithm}% http://ctan.org/pkg/algorithm
+\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
 \usepackage{graphicx}
+\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption
 \usepackage{booktabs}
@@
 %
-\author{Lucas Alcantara\Comment{\inst{1}}\Comment{\orcidID{0000-1111-2222-3333}} \and
-Guilherme Padilha\Comment{\inst{1}}\Comment{\orcidID{1111-2222-3333-4444}} \and
-Marcelo d'Amorim\Comment{\orcidID{0000-0002-1323-8769}}}
+\author{Lucas Alcantara\MyComment{\inst{1}}\MyComment{\orcidID{0000-1111-2222-3333}} \and
+Guilherme Padilha\MyComment{\inst{1}}\MyComment{\orcidID{1111-2222-3333-4444}} \and
+Marcelo d'Amorim\MyComment{\orcidID{0000-0002-1323-8769}}}
 %
@@
 \end{wrapfigure}
-\Comment{For instance, the purpose of rule option \CodeIn{msg} is to print on
+\MyComment{For instance, the purpose of rule option \CodeIn{msg} is to print on
 the output an informative message indicating what kind of intrusion
@@
 
-\Fix{--------------------------- estou aqui}
+The rule reported in the first step satisfies the criterion to only
+capture positive traffic. Unfortunately, the rule produced in the
+first step is overconstrained---it is unable to capture different
+manifestations of the same attack as the rule matches all options that
+\tname\ found for the input message.
 
-The rule reported by the first step satisfies the criterion to only
-capture positive traffic, however it is ``too'' specific---it is
-unable to capture different manifestations of the same attack as the
-rule would unlikely match exactly the options we found for the input
-message. Some options added to the rule during the first step are
-irrelevant for determining the attack and should be removed.
+%%  Some options added to the rule
+%% during the first step are irrelevant for determining the attack and
+%% should be removed to obtain more general rules.
 
-The goal of this step is to create a \emph{minimal} rule that captures
-\emph{only} the positive malicious traffic. A rule is minimal if
-discarding any option results in capturing negative (\ie{}, benign)
-traffic. Recall that the previous maximization step produces a rule
-that captures the positive traffic and all options are satisifed by
-the positive traffic. Consequently, any subset of these options also
-captures the positive traffic. The minimization procedure we propose
-in this step builds on that observation. It iteratively discards
-options from the candidate rule until no other options can be
-removed. The procedure uses negative traffic to guide the
-search. Public databases of negative traffic with \Fix{thousands} of
-messages exist \Fix{cite cite} and can be used to increase the
-accuracy of the minimization.
+\tname{} leverages public datasets\Fix{cite cite} of negative traffic
+to guide a search for more general rules, hence addressing the
+overffiting issue. The goal of this step is to create rules that
+captures traffic similar to the input malicious traffic. To achieve
+that goal, \tname{} \emph{minimizes} the rules produced in the
+previous step, preserving the invariant of \emph{only} capturing the
+positive malicious traffic. A rule is minimal if discarding any option
+results in capturing negative (\ie{}, benign) traffic.  The
+minimization procedure we propose in this step builds on the
+observation that all options from the rule produced in the previous
+step are satisifed by the positive traffic. Consequently, any subset
+of these options also captures the positive traffic. The procedure
+iteratively discards options from the candidate rule until it detects
+that the rule is minimal, \ie{}, it finds that no other options can be
+removed. The procedure uses negative traffic to guide the search.
+
+\begin{algorithm}[t!]
+\textbf{INPUT:} overconstrained rule $or$ \\
+\textbf{OUTPUT:} Set of minimal rules $R$\\
+\vspace{-3ex}
+%\noindent\begin{minipage}{\textwidth}
+\captionof{algorithm}{Minimization algorithm}\label{algo1}
+\begin{algorithmic}[1]
+%  \Procedure{MinRule}{$or$}\Comment{The g.c.d. of a and b}
+    \State $r\gets a\bmod b$
+    \While{$r\not=0$}\Comment{We have the answer if r is 0}
+      \State $a\gets b$
+      \State $b\gets r$
+      \State $r\gets a\bmod b$
+    \EndWhile
+    \State \textbf{return} $b$\Comment{The gcd is b}
+%  \EndProcedure
+\end{algorithmic}
+%\end{minipage}%
+\end{algorithm}
+
+%% Public databases of negative traffic with
+%% \Fix{thousands} of messages exist  and can be used to
+%% increase the accuracy of the minimization.
 

### 2019-10-24 paper/paper.tex
pass on rules and packets

@@
 
-\Fix{Illustrate how information is extracted from packets}
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.3]{figs/ICMP-packet-structure.png}
+\caption{ICMP Packet Structure}
+\label{fig:overview}
+\end{figure*}
+
+\Luc{Os campos Flags e Time to live no header IP são mapeados para as opções fragbits e ttl, respectivamente. Os campos Type e Code no header ICMP são mapeados para as opções itype e icode. \\
+Os campos, Time to Live, Type e Code recebem diretamente o valor que está no pacote. Por exemplo, se o campo Type no pacote possuir valor 8, na regra ele vai aparecer como itype:8.\\
+O campo Flags é intepretado como um vetor de bits, onde cada posição representa uma flag.
+Na regra, o valor do campo Flags é traduzido em caracteres que representam cada flag presente no pacote.
+Por exemplo: se o campo Flags possuir o valor 1010, na regra aparecerá fragbits:RM, pois os caracteres R e M indicam a presença do quarto e segundo bits.}
 
@@
 Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
-this attack. This rule was obtained from the \Fix{...what source...}
+this attack. This rule was obtained from the \Luc{Emerging Threats Open Ruleset}.
 The relevant options in the rule appear in bold---the option
@@
 
-\Fix{atualize isto para o valor correto, por favor. $\rightarrow$}
 \begin{figure}[h]
@@
 options. Figure~\ref{fig:distribution-contents} shows the histogram of
-number of these options per rule (up to 10) for the Suricata public
-rule database \Fix{cite}. Rules without \CodeIn{content} options is an
+number of these options per rule (up to 10) for the \Luc{Emerging Threats Open Ruleset}. Rules without \CodeIn{content} options is an
 exception and half of the rules contain at least two \CodeIn{content}

### 2019-10-24 paper/paper.tex
another pass

--- 2019-10-24 paper/paper.tex
@@
 
-\keywords{NIDS, synthesis, search}
+\keywords{\nids, synthesis, search}
 \end{abstract}
@@
 not block. Various open-source (\eg{}, Snort~\cite{snort} and
-Suricata~\cite{suricata}) and commercial NIDS implementations (\eg{},
+Suricata~\cite{suricata}) and commercial \nids implementations (\eg{},
 SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar}) exist
@@
 two kinds of intrusion detection mechanisms are
-complementary. Rule-based NIDS could miss attacks (as rulesets can
-become outdated) whereas anomaly-based NIDS can raise false alarms (as
+complementary. Rule-based \nids could miss attacks (as rulesets can
+become outdated) whereas anomaly-based \nids can raise false alarms (as
 detection algorithms are approximate).
 
-Rule-based NIDS are restricted to a fixed set of rules defined by the
+Rule-based \nids are restricted to a fixed set of rules defined by the
 network system administrator. Manually writing these rules is tedious,
@@
 fundamental issue, rule-based systems are tremendously popular in
-industry today. Anomaly-based NIDS are not restricted to a fixed set
+industry today. Anomaly-based \nids are not restricted to a fixed set
 of rules. They learn usage behavior from positive traffic and, based
 on that, alert uncommon behavior~\cite{7579764}. Unfortunately,
-anomaly-based NIDS are fundamentally imprecise. To sum up, rule-based
-NIDS are precise but incomplete whereas anomaly-based NIDS are (more)
+anomaly-based \nids are fundamentally imprecise. To sum up, rule-based
+\nids are precise but incomplete whereas anomaly-based \nids are (more)
 complete but imprecise.
@@
 %% This paper focuses on the
-%% problem of discovering new rules by using anomaly-based NIDS.
+%% problem of discovering new rules by using anomaly-based \nids.
 
@@
 intelligence to automatically synthesize rules for rule-based
-\nids. Our general goal is to harden the protection of NIDS by
+\nids. Our general goal is to harden the protection of \nids by
 improving the creation process of rules. One scenario of application
-of \tname{} is one where an anomaly-based NIDS identifies abnormal
+of \tname{} is one where an anomaly-based \nids identifies abnormal
 traffic and signals that traffic to \tname{} to create and distribute
@@
 
-%% Note that anomaly-based NIDS are evolving pretty quick with the
-%% advances in machine learning, but rule-based NIDS are still extremely
+%% Note that anomaly-based \nids are evolving pretty quick with the
+%% advances in machine learning, but rule-based \nids are still extremely
 %% popular.  \tname{} could also leverage existing databases of malicious
@@
 
-Suricata supports a total of \totoptions\ distinct options covering
-\numproto\ distinct protocols. A total of 9 of these options are for
-documentation.
-\begin{wrapfigure}[15]{r}{0.5\textwidth}
+\begin{wrapfigure}[14]{r}{0.5\textwidth}
   \centering
-  \vspace{-5ex}
+  \vspace{-7ex}
   \scalebox{0.7}{
@@
   }
-  \vspace{-1ex}
+  \vspace{-2ex}
   \caption{\label{fig:distribution-rules-protocol}Number of relevant
@@
 \end{wrapfigure}
-\MyComment{For instance, the purpose of rule option \CodeIn{msg} is to print on
+Suricata supports a total of \totoptions\ distinct options covering
+\numproto\ distinct protocols. A total of 9 of these options are for
+documentation.\MyComment{For instance, the purpose of rule option \CodeIn{msg} is to print on
 the output an informative message indicating what kind of intrusion
@@
 Suricata. Column ``\#'' shows the option number, column ``Name'' shows
-the name of the option, column ``\#~Params'' shows the number of
-additional parameters associated with a given option, and column
+the name of the option, and column
 ``Description'' presents a short description of the option purpose. We
@@
 
-\begin{table}[h]
+\begin{table}[t!]
   \caption{\label{table:rules}Example options of Suricata.}  
@@
 
-\Fix{Illustrate how information is extracted from packets}
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.3]{figs/ICMP-packet-structure.png}
+\caption{Layout of an ICMP packet.}
+\label{fig:icmp-packet-layout}
+\end{figure*}
+
+There exists a correspondence between network packets and most rule
+options. Let's consider the ICMP protocol as an
+example. Figure~\ref{fig:icmp-packet-layout} shows the layout of an
+ICMP packet. The packet fields ``Flags'' and ``Time to Live'',
+declared in the IP header, are mapped to the options \CodeIn{flagbits}
+and \CodeIn{ttl}, respectively. Likewise, the fields ``Type'' and
+``Code'', declared in the ICMP header, are mapped to the options
+\CodeIn{itype} and \CodeIn{icode}. (Table~\ref{table:rules} lists some
+example rule options.) The values of the options ``ttl'', ``itype'',
+and ``icode'' are the same as the value of the corresponding fields in
+the ICMP packet. For the field ``Flags'' the value of the
+corresponding option is encoded in its bitvector. For example, if the
+``Flags'' field stores the value 1010, the corresponding rule option
+will be \CodeIn{flagbits:RM}, as each position of the vector encodes a
+different character in the option \CodeIn{flagbits}.
+
+The value of most options can be obtained through transformation
+functions, as those described above. However, this is not the case for
+the option \CodeIn{threshold}, which is typically used in
+denial-of-service attacks. Conceptually, that option operates over
+sequence of messages as opposed to a single message. More precisely,
+to trigger an action on a rule containing that option the \nids\ tool
+needs to monitor several messages.
+
+%% \Luc{
+%%   Os campos Flags e Time to live no header IP são mapeados para as opções fragbits e ttl, respectivamente. Os campos Type e Code no header ICMP são mapeados para as opções itype e icode. \\
+%% Os campos, Time to Live, Type e Code recebem diretamente o valor que está no pacote. Por exemplo, se o campo Type no pacote possuir valor 8, na regra ele vai aparecer como itype:8.\\
+%% O campo Flags é intepretado como um vetor de bits, onde cada posição representa uma flag.
+Na regra, o valor do campo Flags é traduzido em caracteres que representam cada flag presente no pacote.
+Por exemplo: se o campo Flags possuir o valor 1010, na regra aparecerá fragbits:RM, pois os caracteres R e M indicam a presença do quarto e segundo bits.}
 
@@
 
-This section briefly presents examples of attacks and rules that are
-synthesized with \tname{} to capture those attacks.
+This section briefly presents examples of attacks and illustrates the
+rules that \tname{} synthesizes to capture those attacks.
 
@@
 Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
-this attack. This rule was obtained from the \Fix{...what source...}
-The relevant options in the rule appear in bold---the option
-\CodeIn{flags: S,12}, identifies a SYN packet in a TCP packet and the
-option \CodeIn{threshold: type both, track by\_dst, count 5000,
-  seconds 5} indicates that a high volume of such packets should be
-requested in a short period of time.
+this attack. This rule was obtained from the Emerging Threats Open
+Ruleset~\cite{emerging-threats-open}. The relevant options in the
+rule appear in bold---the option \CodeIn{flags: S,12}, identifies a
+SYN packet in a TCP packet and the option \CodeIn{threshold: type
+  both, track by\_dst, count 5000, seconds 5} indicates that a high
+volume of such packets should be requested in a short period of time.
 
@@
 \tname{} is a technique that uses positive and negative examples to
-synthesize signature-based NIDS rules, such as those from Snort and
+synthesize signature-based \nids rules, such as those from Snort and
 Suricata. The goal of \tname\ is to create a rule that \emph{only}
@@
 \pgfplotsset{every tick label/.append style={font=\tiny}}
-
-\begin{wrapfigure}[14]{r}{0.56\textwidth}
+The goal of the first step is to extract options from the positive
+traffic. For that, \tname{} parses the protocol message, looking for
+fields in the message associated with options (see
+Section~\ref{sec:rules-and-packets}).
+Unfortunately, inferring options at the field granularity
+level is insufficient to capture \CodeIn{content} options (see
+Section~\ref{sec:example-suricata-rules}), which refer to parts of the
+payload of the message and cannot be directly extracted directly from
+the fields. Handling this case is critically important as most rules
+contain \CodeIn{content}
+options. Figure~\ref{fig:distribution-contents} shows the histogram of
+number of these options per rule (up to 10) obtained from a public
+ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
+rules. Rules without \CodeIn{content} options is an exception and half
+of the rules contain at least two \CodeIn{content} options.  To find
+\begin{wrapfigure}[15]{r}{0.56\textwidth}
   \centering
-  \vspace{-8ex}
+  \vspace{-4ex}
   \scalebox{1.2}{  
@@
 \end{wrapfigure}
-The goal of the first step is to extract options from the positive
-traffic. For that, \tname{} parses the protocol message, looking for
-fields in the message associated with options (see
-Section~\ref{sec:rules-and-packets}). For example, fields
-\Fix{...and...}  can be used to determine the option
-\Fix{zzz}. Unfortunately, inferring options at the field granularity
-level is insufficient to capture \CodeIn{content} options (see
-Section~\ref{sec:example-suricata-rules}), which refer to parts of the
-payload of the message and cannot be directly extracted directly from
-the fields. Handling this case is critically important as most rules
-contain \CodeIn{content}
-options. Figure~\ref{fig:distribution-contents} shows the histogram of
-number of these options per rule (up to 10) for the Suricata public
-rule database \Fix{cite}. Rules without \CodeIn{content} options is an
-exception and half of the rules contain at least two \CodeIn{content}
-options.  To find the values associated with these options, \tname{}
-splits the payload of the malicious message in tokens using natural
-language delimiters, such as $\backslash$t, $\backslash$n,
-$\backslash$r, and spaces. The rationale for this decision is that we
-observed, by inspecting existing rules, that \CodeIn{content} options
-typically refer to sequence of characters in the payload of messages
-separated by these delimiters. This tokenization step can generate
-dozens of options. For the example from
-Section~\ref{sec:content-example}, \tname{} produces a total of
-\Fix{xx} content options.
+the values associated with these options, \tname{} splits the payload
+of the malicious message in tokens using natural language delimiters,
+such as $\backslash$t, $\backslash$n, $\backslash$r, and spaces. The
+rationale for this decision is that we observed, by inspecting
+existing rules, that \CodeIn{content} options typically refer to
+sequence of characters in the payload of messages separated by these
+delimiters. This tokenization step can generate dozens of options. For
+the example from Section~\ref{sec:content-example}, \tname{} produces
+a total of \Fix{xx} content options.
 

### 2019-10-24 paper/paper.tex
pass on recon
pass on pseudo-code

@@
 produces optimal rule candidates on output. We evaluated \tname{} on a
-diverse set of attacks. Results indicate that \Fix{...}
+diverse set of attacks. Results indica te that \Fix{...}
 
@@
 \tname{} is a technique that uses positive and negative examples to
-synthesize signature-based \nids rules, such as those from Snort and
+synthesize signature-based \nids\ rules, such as those from Snort and
 Suricata. The goal of \tname\ is to create a rule that \emph{only}
@@
 
-\tname{} leverages public datasets\Fix{cite cite} of negative traffic
-to guide a search for more general rules, hence addressing the
+\tname{} leverages public datasets\cite{tcpreplay} of negative traffic
+to guide the search for more general rules, hence addressing the
 overffiting issue. The goal of this step is to create rules that
@@
 
+\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%
+\algnewcommand\algorithmicforeach{\textbf{for each}}
+\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
+\algnewcommand\And{\textbf{and} }
 \begin{algorithm}[t!]
 \textbf{INPUT:} overconstrained rule $or$ \\
-\textbf{OUTPUT:} Set of minimal rules $R$\\
+\textbf{OUTPUT:} Set of minimal rules $\mathcal R$\\
 \vspace{-3ex}
-%\noindent\begin{minipage}{\textwidth}
 \captionof{algorithm}{Minimization algorithm}\label{algo1}
 \begin{algorithmic}[1]
-%  \Procedure{MinRule}{$or$}\Comment{The g.c.d. of a and b}
-    \State $r\gets a\bmod b$
-    \While{$r\not=0$}\Comment{We have the answer if r is 0}
-      \State $a\gets b$
-      \State $b\gets r$
-      \State $r\gets a\bmod b$
-    \EndWhile
-    \State \textbf{return} $b$\Comment{The gcd is b}
-%  \EndProcedure
+  \State $\mathcal R \gets \{\mathit{or}\}$\Comment{initial population
+    of candidate solutions contains only the input rule.}\label{line-init}
+  \Do\label{start-do-while}
+      \State $\mathcal R' \gets \emptyset$\Comment{starting new
+        iteration. new generation should be empty.}\label{init-rprime}
+      \ForEach {$ro \in \mathcal R $}\Comment{analyze each candidate individual/rule separately.}\label{start-foreach}
+           \For {$i \gets 1$ to $N$}\Comment{mutate each rule $N$ times.}
+               \State $r \gets copy(ro)$      
+               \State $j\gets \mathit{rand}(0,\mathit{len}(r))$
+               \State $r[j]\gets 0$\Comment{discard option $j$ of rule
+                 $r$, acopy of the original rule $\mathit{ro}$}
+               \If{$\mathit{isnew}(r)$ \And $\mathit{isfit}(r)$}
+                   \State $\mathcal R' \cup\gets \{r\}$\Comment{$r$
+                     can lead to a plausible candidate solution.}\label{line-progress}
+               \EndIf
+           \EndFor
+       \EndFor\label{end-foreach}
+       \State $\mathcal R \gets R'$\Comment{update new generation}
+  \doWhile{$\mathcal R'\not= \{\}$}\Comment{repeat until reaching a fixpoint}\label{end-do-while}
 \end{algorithmic}
-%\end{minipage}%
 \end{algorithm}
+
+Algorithm~\ref{algo1} shows the pseudocode of the \tname{}
+minimization algorithm. The algorithm takes as input the
+overconstrained rule produced by the reverse engineering step and
+produces a set $\mathcal R$ of rules that are minimal and only
+captures the positive traffic. Line~\ref{line-init} initializes the
+set $\mathcal{R}$ of candidate solutions with the input rule $or$. We
+encode a rule as a bitvector where each index in the vector represents
+one option; the value assigned at a given position indicates whether
+or not the option is present in the rule. Each iteration of the outer
+loop (lines~\ref{start-do-while}-\ref{end-do-while}) represents one
+generation of the population of individuals produced by the
+search. The algorithm terminates when no rule in the current
+generation of individuals could be further reduced without violating
+the property that a candidate solution should only capture positive
+traffic. This means that execution has not reached
+line~\ref{line-progress} as to augment the set $\mathcal{R'}$ and that
+set will be empty at the end of one iteration of the outer
+loop. Line~\ref{init-rprime} initializes the new generation of
+candidate rules. The loop from
+lines~\ref{start-foreach}--\ref{end-foreach} iterates through each
+candidate rule in $\mathcal R$. Note that these rules have been found
+in the previous generation. The algorithm mutates each of these rules
+for $N$ times, discarding one option of the rule each time. The call
+to $\mathit{isNew}$ checks if $r$ has not been previously generated
+through another path whereas the call to $\mathit{isfit}$ checks if
+the rule only captures positive traffic. Function $\mathit{copy}$
+creates a copy of the bitvector of a rule, function $\mathit{len}$
+returns the length of the vector, and function $\mathit{rand}$
+randomly chooses one index of the vector.
 
@@
 %% increase the accuracy of the minimization.
-
-\subsection{Limitations}
-
-\Fix{...}
+%\subsection{Limitations}
+%\Fix{...}
 

### 2019-10-24 paper/paper.tex
acks

@@
 common type of port scan based on the ICMP
-protocol. \Fix{...explain this attack...}
+protocol.
+\Luc{The attacker sends a set of ICMP Echo requests to a network of machines and sees which one responds to the request.
+This allow the attacker to know which machines are active in the network, then it can choose one of them to attack.}
 
@@
 \Fix{...}
+
+\section*{Acknowledgments}
+This work is partially supported by RNP grant number 002951.
+
 %

### 2019-10-24 paper/paper.tex
pass

@@
 iteratively discards options from the candidate rule until it detects
-that the rule is minimal, \ie{}, it finds that no other options can be
+that the rule is minimal, \ie{}, it finds that no other option can be
 removed. The procedure uses negative traffic to guide the search.
@@
       \ForEach {$ro \in \mathcal R $}\Comment{analyze each candidate individual/rule separately.}\label{start-foreach}
-           \For {$i \gets 1$ to $N$}\Comment{mutate each rule $N$ times.}
+           \For {$i \gets 1$ to $N$}\Comment{mutate each rule $N$ times.}\label{start-for}
                \State $r \gets copy(ro)$      
@@
                \State $r[j]\gets 0$\Comment{discard option $j$ of rule
-                 $r$, acopy of the original rule $\mathit{ro}$}
+                 $r$, a copy of the original rule $\mathit{ro}$}
                \If{$\mathit{isnew}(r)$ \And $\mathit{isfit}(r)$}
@@
                \EndIf
-           \EndFor
+           \EndFor\label{end-for}
        \EndFor\label{end-foreach}
@@
 
+\begin{theorem}
+  The minimization algorithm computes rules that capture only
+  the positive traffic.
+\end{theorem}
+~
+\begin{proof}
+  Every rule created during the search captures the positive traffic
+  as every rule contains a proper subset of the option from the
+  original rule and all those options are satisfied by the positive
+  traffic. By definition of function $\mathit{isfit}$, these rules
+  also capture \emph{only} the positive traffic. If some negative
+  traffic was captured, a rule would not be added to $\mathcal
+  R'$.\qed
+\end{proof}
+
+\begin{proposition}
+  The minimization algorithm may not produce minimal rules.
+\end{proposition}
+
+The output of the minimization algorithm may contain rules that are
+not minimal given the non-determinism of the algorithm. It is possible
+that options that could be discarded are not selected in the loop from
+lines~\ref{start-for}--\ref{end-for}. However, we found empirically
+that a small value of $N$ suffices to produce minimal rules.
+
 %% Public databases of negative traffic with
@@
 %\Fix{...}
+
+\vspace{1ex}
+\noindent\textbf{Limitation.} It is worth noting that although
+\tname{} is accurate for generating rules that discrimate positive and
+negative traffic, the quality of these rules is limited by the quality
+of the training data it uses. For the positive traffic, \tname{} only
+requires one sample, but can benefit of several samples. For the
+negative traffic, \tname{} can use public datasets available online.
 

### 2019-10-25 paper/paper.tex
pass on recon

@@
 protocol.
-\Luc{The attacker sends a set of ICMP Echo requests to a network of machines and sees which one responds to the request.
+\Luc{To scan a network, an attacker sends a set of ICMP Echo requests addresed to a range of IP addresses in the network and sees which one responds to the request.
 This allow the attacker to know which machines are active in the network, then it can choose one of them to attack.}
 
-\Fix{...explain how \tname\ works on it...}
+\begin{figure}[h!]
+  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
+  \caption{Suricata rule for Ping Scan.}
+  \label{fig:synflood-example}
+\end{figure}
+
+\Luc{Figure 4 shows a Suricata rule that capture a ping scan performed by the NMAP Tool.
+The relevant options are in bold---the option \CodeIn{dsize: 0}, indicates that the packet has no bytes in the payload.
+\CodeIn{itype: 8} identifies the packet as a ICMP Echo request.}
+
+\Luc{To synthesize rules for a Ping Scan, \tname{} analyzes the positive traffic and extract the following set of options:}
+
+\begin{figure}[h]
+\vspace{-3ex} \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{pingscan.suricata.synth}
+%  \caption{Suricata rule for SYN Flood Attacks.}
+  %  \label{fig:synflood-example-synt}
+\vspace{-3ex}  
+\end{figure}
+
+\Luc{Again, a rule with this options would capture the positive traffic, but it would be overspecified.
+So, \tname{} tries to minimize the rule...
+After the minimization, it produces the following rule:}
+\Luc{}
 

### 2019-10-25 paper/paper.tex
another pass

@@
 \subsection{Step 2: Rule Minimization}
+\label{sec:minimization}
 
@@
 
+\vspace{1ex}
+\noindent\textbf{Implementation details.} \Fix{discuss function isFit}
+
 \section{Evaluation}
 
-%% \renewcommand{\labelitemi}{}
-%% \setitemize[0]{leftmargin=20pt,itemindent=-20pt}
-%% \begin{itemize}
-%% \end{itemize}
-
-%% \renewcommand{\labelitemi}{}
-%% \setitemize[0]{leftmargin=0pt,itemindent=0pt}
-%% \begin{itemize}
-%%   \item{\textbf{RQ1.}~How precise is \tname?}
-%%   \item{\textbf{RQ2.}~How efficient is \tname?}
-%% \end{itemize}
-
-This section reports precision and efficiency of \tname.
+This section evaluates the precision and efficiency of \tname{}.
 
@@
 
-We evaluated \tname{} on a set of attacks covering each of the four
-protocols supported by Suricata and attacks with different
-characteristics. Table~\ref{table:attacks} shows the attacks we
-analyzed and the rule of Suricata that captures the attack. Column
-``Name'' shows the name of the attack, column ``\#'' indicates whether
-single or multiple messages are required to manifest the attack,
-column ``Protocol'' shows the protocol name, column ``Description''
-summarizes the effect of the attack, if successfull, and column
-``Golden Rule'' shows the relevant parts in the options section of the
-Suricata rule that captures the attack.
+Table~\ref{table:attacks} shows the attacks we analyzed and the
+options of the rule from the Emerging Threats Open
+Ruleset~\cite{emerging-threats-open} that captures the attack. Column
+``Name'' shows the name of the attack, column ``Protocol'' shows the
+protocol name, and column ``Ground Truth'' shows the options of the
+rule that captures the attack.
 
-\begin{table*}[t!]
-  \caption{\label{table:attacks}Characterization of attacks analyzed
-    and Golden Rule of Suricata (reference) to capture it.\Mar{please,
-  show me the Golden Rule, \ie{} the part that is relevant of each rule (for
-  capturing the negative traffic precisely)}}  
+\begin{table*}[h]
+  \caption{\label{table:attacks}List of attacks.\Mar{Guilherme, por
+      favor, adicione mais ataques aqui...}}
   \centering
-  \begin{tabular}{lllll}
+  \begin{tabular}{lll}
     \toprule
-    \multicolumn{1}{c}{Name} & \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Protocol} & \multicolumn{1}{c}{Description} & \multicolumn{1}{c}{Golden Rule} \\
+    \multicolumn{1}{c}{Name} &
+    \multicolumn{1}{c}{Protocol} &
+    \multicolumn{1}{c}{Groud Truth} \\
     \midrule     
-    Ping Scan & single & ICMP & Discovers active hosts in a network &  dsize:0; itype:8; \\
-    Black Nurse & multiple & ICMP & Denial of Service & itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;\\
-    Ping Flood  & multiple & ICMP & Denial of Service & itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;\\  
-    Port Scan TCP & single & TCP & Discovers open ports in a host & flags: A; ack: 0;dsize: 0; \\
-    SYN Flood & multiple & TCP & Denial of Service & flags: S,12; threshold: type both, track by\_dst, count 5000, seconds 5;\\
-    UDP Flood & multiple & UDP & Denial of Service & fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5; \\
+    Ping Scan \MyComment{& single} & ICMP \MyComment{& Discovers
+      active hosts in a network} & \scriptsize{dsize:0; itype:8;} \\
+    Black Nurse \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;}\\
+    Ping Flood  \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;}\\  
+    Port Scan \MyComment{& single} & TCP \MyComment{& Discovers open ports in a host} & \scriptsize{flags: A; ack: 0;dsize: 0;} \\
+    SYN Flood \MyComment{& multiple} & TCP \MyComment{& Denial of Service} & \scriptsize{flags: S,12; threshold: type both, track by\_dst, count 5000, seconds 5;}\\
+    UDP Flood \MyComment{& multiple} & UDP \MyComment{& Denial of Service} & \scriptsize{fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5;} \\
     \bottomrule
@@
 
-\subsection{Setup}
+\subsection{Dataset of Benign Traffic}
 
-Note from Figure~\ref{fig:overview} that the negative traffic is an
-input to the technique. The positive (non malicious) traffic, in
-contrast, is not. \tname{} uses public databases storing different
-kinds of positive traffic for all protocols covered by Suricata. We
-used two datasets of positive traffic. The Bigflows.pcap dataset is
-part of the TcpReplay~\cite{tcpreplay} open source project. This
-dataset includes real network traffic on a busy private network access
-point to the Internet. According to the TcpReplay web site ``This
-capture is much larger and has a smaller average packet size than the
-previous capture (smallFlows.pcap). It also has many more flows and
-different applications.''. In addition to the Bigflows.pcap dataset,
-we also used datasets of normal traffic made available by Stratosphere
+Note from Figure~\ref{fig:overview} that the positive traffic is an
+input to the technique. The negative non-malicious traffic, however,
+is not. \tname{} uses public databases storing different kinds of
+positive traffic for all protocols covered by Suricata. We used two
+datasets of positive traffic. The Bigflows.pcap dataset is part of the
+TcpReplay~\cite{tcpreplay} open source project. This dataset includes
+real network traffic on a busy private network access point to the
+Internet. According to the TcpReplay web site ``This capture is much
+larger and has a smaller average packet size than the previous capture
+(smallFlows.pcap). It also has many more flows and different
+applications.''. In addition to the Bigflows.pcap dataset, we also
+used datasets of normal traffic made available by Stratosphere
 Lab~\cite{stratosphere-normal}.
+
+\subsection{Methodology}
+
+\tname{} expects (a) message(s) reproducing an attack on input. As
+previously discussed, those messages could be directly provided by an
+anomaly-based \nids\ tool. To evaluate \tname{}, however, we created
+those messages ourselves. For all attacks listed on
+Table~\ref{table:attacks}, we found reproduction scripts available on
+the web. With the script to reproduce the attack, we proceeded as
+follows: 1)~we instantiated the script on a given target machine,
+2)~we executed the script, 3)~we monitored the traffic with the
+wireshark network monitor~\cite{wireshark-net-monitor}, and 4) we
+simulated the delivery of the corresponding network messages by
+directly executing Suricata on the ``pcap'' files that wireshark
+produced in the previous step. It is worth noting that Suricata is
+invoked multiple times in the process of minimizing the rule (see
+Section~\ref{sec:minimization}) and that processing pcap files is much
+more efficient compared to sending and processing network messages. To
+illustrate, the SYN Flood attack was reproduced with the hping3
+tool~\cite{hping3} using the following command \CodeIn{hping3 -d 80 -w
+  64 -S -p 80 --flood --rand-source <IP>}.
 
@@
 
-\Luc{entrada: utilizamos a ferramenta hping3 para executar o ataque em um alvo arbitrario dentro da rede (n ha necessidade de haver um alvo real para executar o hping3) e capturamos o trafego gerado com o wireshark.
+%% \Luc{entrada: utilizamos a ferramenta hping3 para executar o ataque em um alvo arbitrario dentro da rede (n ha necessidade de haver um alvo real para executar o hping3) e capturamos o trafego gerado com o wireshark.
 o comando utilizado foi:\\*
-\texttt{\# hping3 -d 80 -w 64 -S -p 80 --flood --rand-source 192.138.1.115} \\*
-que envia pacotes de 80 bytes (-d 80), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80).
+%% \texttt{\# hping3 -d 80 -w 64 -S -p 80 --flood --rand-source 192.138.1.115} \\*
+%% que envia pacotes de 80 bytes (-d 80), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80).
 A flag --flood indica q o hping3 vai enviar os pacotes o mais rápido possível.
@@
 192.138.1.115 é o alvo.\\*
-O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
-O wireshark armazena os pacotes no formato pcap, utilizamos a ferramenta tcpreplay para reproduzir o trafego com o comando:\\*
-\texttt{\# tcpreplay --pps 200 --intf1=wlp2s0 Datasets/syn-flood-sample.pcap}\\*
-Nesse caso, estamos enviando os pacotes armazenados em \texttt{Datasets/syn-flood-sample.pcap} a uma taxa de 200 pacotes por segundo (\texttt{-pps 200}) utilizando a interface wlp2s0 (\texttt{--intf1 wlp2s0})\\*
-Há diferenças entre os pacotes na opção ack.\\* 
-Por que 20?
+%% O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
+%% O wireshark armazena os pacotes no formato pcap, utilizamos a ferramenta tcpreplay para reproduzir o trafego com o comando:\\*
+%% \texttt{\# tcpreplay --pps 200 --intf1=wlp2s0 Datasets/syn-flood-sample.pcap}\\*
+%% Nesse caso, estamos enviando os pacotes armazenados em \texttt{Datasets/syn-flood-sample.pcap} a uma taxa de 200 pacotes por segundo (\texttt{-pps 200}) utilizando a interface wlp2s0 (\texttt{--intf1 wlp2s0})\\*
+%% Há diferenças entre os pacotes na opção ack.\\* 
+%% Por que 20?
 Apenas para gerar a regra mais rapidamente, poderia ser qualquer valor maior ou menor.
 No final, o parâmetro \texttt{count} da opção \texttt{threshold} vai ser alterado pelo adm da rede.\\*
-Intervalo de reprodução: 0.005 segundo
+%% Intervalo de reprodução: 0.005 segundo
 
-}
+%% }
 
-\Fix{any other detail we should consider?}
+%% \Fix{any other detail we should consider?}
 
@@
 
-\Fix{...}
-
 \subsection{Efficiency}
-
-\Fix{...}
 

### 2019-10-25 paper/paper.tex
other updates

@@
 \end{theorem}
-~
 \begin{proof}
@@
 
-\vspace{1ex}
+\vspace{2ex}
 \noindent\textbf{Limitation.} It is worth noting that although
@@
 
-\vspace{1ex}
-\noindent\textbf{Implementation details.} \Fix{discuss function isFit}
+\vspace{2ex}
+\noindent\textbf{Implementation details.} We implemented function
+\emph{isFit} by 1)~updating Suricata to only use the informed rule,
+2)~replaying the negative traffic (see
+Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
+check if any benign message was captured by the rule. It is worth
+noting that the benign traffic is stored in ``pcap'' files~\cite{pcap}
+and Suricata can process those files very efficiently compared to
+analyzing network messages directly.
 
@@
 \subsection{Dataset of Benign Traffic}
+\label{sec:dataset-benign}
 
@@
 
+\Fix{}
+
 \subsection{Efficiency}
 
-\section{Threats to Validity}
+\subsection{Threats to Validity}
 
@@
 \section{Related Work}
+
+\Fix{...}
+
+\section{Conclusions}
 

### 2020-01-03 paper/paper.tex
ISSTA format + pass on text

@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-%\usepackage[margin=1in]{geometry}% http://ctan.org/pkg/geometry
+%%
+%% This is file `sample-sigconf.tex',
+%% generated with the docstrip utility.
+%%
+%% The original source files were:
+%%
+%% samples.dtx  (with options: `sigconf')
+%% 
+%% IMPORTANT NOTICE:
+%% 
+%% For the copyright see the source file.
+%% 
+%% Any modified versions of this file must be renamed
+%% with new filenames distinct from sample-sigconf.tex.
+%% 
+%% For distribution of the original source see the terms
+%% for copying and modification in the file samples.dtx.
+%% 
+%% This generated file may be distributed as long as the
+%% original source files, as listed above, are part of the
+%% same distribution. (The sources need not necessarily be
+%% in the same archive or directory.)
+%%
+%% The first command in your LaTeX source must be the \documentclass command.
+\documentclass[sigconf,anonymous]{acmart}
+
+\usepackage{libertine}
 \usepackage{lipsum}% http://ctan.org/pkg/lipsum
@@
 \usepackage{booktabs}
-\usepackage{cite}
+%% \usepackage{cite}
 \usepackage{url}
@@
 \usepackage[caption=false, font=footnotesize]{subfig}
-%%%%%%%%%%%%% code listing
 \renewcommand{\ttdefault}{pcr}
@@
 
+%% 
 \input{macros}
 
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
+%%
+%% \BibTeX command to typeset BibTeX logo in the docs
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
 
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmcopyright}
+\copyrightyear{2018}
+\acmYear{2018}
+\acmDOI{10.1145/1122445.1122456}
+
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
+  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
+\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+  June 03--05, 2018, Woodstock, NY}
+\acmPrice{15.00}
+\acmISBN{978-1-4503-XXXX-X/18/06}
+
+%%
+%% Submission ID.
+%% Use this when submitting an article to a sponsored event. You'll
+%% receive a unique submission ID from the organizers
+%% of the event, and this ID should be used as the parameter to this command.
+%%\acmSubmissionID{123-A56-BU3}
+
+%%
+%% The majority of ACM publications use numbered citations and
+%% references.  The command \citestyle{authoryear} switches to the
+%% "author year" style.
+%%
+%% If you are preparing content for an event
+%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+%% citations and references.
+%% Uncommenting
+%% the next command will enable that style.
+%%\citestyle{acmauthoryear}
+
+%%
+%% end of the preamble, start of the body of the document source.
 \begin{document}
-%
-\title{Synthesis of Rules for Network Intrusion Detectors}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{Lucas Alcantara\MyComment{\inst{1}}\MyComment{\orcidID{0000-1111-2222-3333}} \and
-Guilherme Padilha\MyComment{\inst{1}}\MyComment{\orcidID{1111-2222-3333-4444}} \and
-Marcelo d'Amorim\MyComment{\orcidID{0000-0002-1323-8769}}}
-%
-\authorrunning{Alcantara et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Federal University of Pernambuco (UFPE), Recife, Brazil\\
-\email{\{lama,ghps,damorim\}@cin.ufpe.br}}
-%
-\maketitle              % typeset the header of the contribution
-%
+
+%%
+%% The "title" command has an optional parameter,
+%% allowing the author to define a "short title" to be used in page headers.
+\title{Learning to Synthesize Rules for Network Intrusion Detectors}
+
+%%
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+%% Of note is the shared affiliation of the first two authors, and the
+%% "authornote" and "authornotemark" commands
+%% used to denote shared contribution to the research.
+\author{Ben Trovato}
+\authornote{Both authors contributed equally to this research.}
+\email{trovato@corporation.com}
+\orcid{1234-5678-9012}
+\author{G.K.M. Tobin}
+\authornotemark[1]
+\email{webmaster@marysville-ohio.com}
+\affiliation{%
+  \institution{Institute for Clarity in Documentation}
+  \streetaddress{P.O. Box 1212}
+  \city{Dublin}
+  \state{Ohio}
+  \postcode{43017-6221}
+}
+
+\author{Lars Th{\o}rv{\"a}ld}
+\affiliation{%
+  \institution{The Th{\o}rv{\"a}ld Group}
+  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
+  \city{Hekla}
+  \country{Iceland}}
+\email{larst@affiliation.org}
+
+\author{Valerie B\'eranger}
+\affiliation{%
+  \institution{Inria Paris-Rocquencourt}
+  \city{Rocquencourt}
+  \country{France}
+}
+
+\author{Aparna Patel}
+\affiliation{%
+ \institution{Rajiv Gandhi University}
+ \streetaddress{Rono-Hills}
+ \city{Doimukh}
+ \state{Arunachal Pradesh}
+ \country{India}}
+
+\author{Huifen Chan}
+\affiliation{%
+  \institution{Tsinghua University}
+  \streetaddress{30 Shuangqing Rd}
+  \city{Haidian Qu}
+  \state{Beijing Shi}
+  \country{China}}
+
+\author{Charles Palmer}
+\affiliation{%
+  \institution{Palmer Research Laboratories}
+  \streetaddress{8600 Datapoint Drive}
+  \city{San Antonio}
+  \state{Texas}
+  \postcode{78229}}
+\email{cpalmer@prl.com}
+
+\author{John Smith}
+\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
+\email{jsmith@affiliation.org}
+
+\author{Julius P. Kumquat}
+\affiliation{\institution{The Kumquat Consortium}}
+\email{jpkumquat@consortium.net}
+
+%%
+%% By default, the full list of authors will be used in the page
+%% headers. Often, this list is too long, and will overlap
+%% other information printed in the page headers. This command allows
+%% the author to define a more concise list
+%% of authors' names for this purpose.
+\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
+
+%%
+%% The abstract is a short summary of the work to be presented in the
+%% article.
 \begin{abstract}
@@
 as attackers learn new strategies to circumvent existing defenses.
-This paper proposes \tname{}\footnote{Abbreviation of \textbf{Sy}nthesis of
-  Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a technique that automatically
-synthesizes \nids\ rules from positive and negative examples (\ie{},
-malicious and benign traffic). \tname{} formulates synthesis as an
-optimization problem whose candidate solutions are rules that
-maximizes the captured the positive traffic and minimizes the captured
-negative traffic. \tname{} bootstraps the search with candidate
-solutions that use data from the payload of the malicious message and
-produces optimal rule candidates on output. We evaluated \tname{} on a
-diverse set of attacks. Results indica te that \Fix{...}
+This paper proposes \tname{}, a technique that automatically
+synthesizes \nids\ rules from positive and negative examples, \ie{},
+malicious and benign traffic. \tname{} formulates synthesis as an
+optimization problem whose candidate solutions are rules that maximize
+the capture of positive traffic and minimize the capture of negative
+traffic. \tname{} bootstraps the search with candidate solutions that
+use data from the payload of the malicious message and produces
+optimal rule candidates on output. We evaluated \tname{} on a diverse
+set of attacks. Results indicate that \Fix{...}
+\end{abstract}
 
-%% optimizes the
-%% candidate solution to capture the positive example and, and then
-%% optimizes the candidate solution to avoid capturing the negative
-%% traffic.
+%%
+%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+%% Please copy and paste the code instead of the example below.
+%%
+\begin{CCSXML}
+<ccs2012>
+ <concept>
+  <concept_id>10010520.10010553.10010562</concept_id>
+  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+  <concept_significance>500</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010575.10010755</concept_id>
+  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+  <concept_significance>300</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010553.10010554</concept_id>
+  <concept_desc>Computer systems organization~Robotics</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10003033.10003083.10003095</concept_id>
+  <concept_desc>Networks~Network reliability</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+</ccs2012>
+\end{CCSXML}
 
+\ccsdesc[500]{Computer systems organization~Embedded systems}
+\ccsdesc[300]{Computer systems organization~Redundancy}
+\ccsdesc{Computer systems organization~Robotics}
+\ccsdesc[100]{Networks~Network reliability}
+
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
 \keywords{\nids, synthesis, search}
-\end{abstract}
-%
-%
-%
+
+%% This command processes the author and affiliation and title
+%% information and builds the first part of the formatted document.
+\maketitle
+
 \section{Introduction}
+\label{sec:intro}
 
@@
 not block. Various open-source (\eg{}, Snort~\cite{snort} and
-Suricata~\cite{suricata}) and commercial \nids implementations (\eg{},
+Suricata~\cite{suricata}) and commercial \nids\ implementations (\eg{},
 SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar}) exist
@@
 
-\sloppy \nids{} are typically categorized in two
-groups~\cite{kumar2007survey}: rule-based and anomaly-based \nids. A
-rule-based intrusion detector\footnote{\aka\ signature-based intrusion
-  detector.} checks if the network traffic matches a fixed set of
-rules. Figure~\ref{fig:synflood-example} shows an example rule of
-Suricata~\cite{suricata}, a popular open-source \nids{}. This rule
-prescribes a method to capture a denial-of-service
-attack~\cite{understanding-dos} to a server by matching specific
-conditions about the network traffic. Relevant properties about the
-traffic of interest appear in bold (see
-Section~\ref{sec:suri-metas-coverage}). Rule-based \nids{} focus on
-known attacks whereas anomaly-based \nids{} focus on unknown attacks
-that could be observed by first learning the regular traffic and then
-acting
-properly~\cite{kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}. These
-two kinds of intrusion detection mechanisms are
-complementary. Rule-based \nids could miss attacks (as rulesets can
-become outdated) whereas anomaly-based \nids can raise false alarms (as
-detection algorithms are approximate).
+This paper focuses on rule-based \nids{}, which is a very popular kind
+of NIDS used in industry (see Section~\ref{sec:background}). A
+\emph{rule-based intrusion detector}\footnote{\aka\ signature-based
+  intrusion detector.} checks if the network traffic matches a fixed
+set of rules. Figure~\ref{fig:synflood-example} shows an example rule
+of Suricata~\cite{suricata}, a popular open-source \nids{}. This
+particular rule prescribes a method to capture a denial-of-service
+attack to a server by matching specific conditions about the network
+traffic~\cite{understanding-dos}. Relevant properties about the
+traffic of interest appear in bold in this rule (see
+Section~\ref{sec:suri-metas-coverage}). Deployments of rule-based
+\nids\ are restricted to a fixed set of rules defined by the network
+system administrator. Updating these rules is tedious and error-prone,
+but it is important as attackers constantly create new strategies to
+circumvent existing rules. IT-security companies capitalize on this
+phenomena and offer rulesets on the
+market~\cite{proofpoint-etpro,snort-rule-subscriptions}.
 
-Rule-based \nids are restricted to a fixed set of rules defined by the
-network system administrator. Manually writing these rules is tedious,
-error-prone, but important as attackers constantly create new
-strategies to circumvent existing rules. IT-security companies
-capitalize on this phenomena and put rulesets on the market for
-sale~\cite{proofpoint-etpro,snort-rule-subscriptions}.  Despite this
-fundamental issue, rule-based systems are tremendously popular in
-industry today. Anomaly-based \nids are not restricted to a fixed set
-of rules. They learn usage behavior from positive traffic and, based
-on that, alert uncommon behavior~\cite{7579764}. Unfortunately,
-anomaly-based \nids are fundamentally imprecise. To sum up, rule-based
-\nids are precise but incomplete whereas anomaly-based \nids are (more)
-complete but imprecise.
-
-%% This paper focuses on the
-%% problem of discovering new rules by using anomaly-based \nids.
-
-This paper proposes \tname{}, a technique that uses machine
-intelligence to automatically synthesize rules for rule-based
-\nids. Our general goal is to harden the protection of \nids by
-improving the creation process of rules. One scenario of application
-of \tname{} is one where an anomaly-based \nids identifies abnormal
-traffic and signals that traffic to \tname{} to create and distribute
-the rules.
+This paper proposes \tname{}\footnote{Abbreviation for
+  \textbf{Sy}nthesis of Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a
+technique that uses machine intelligence to automatically synthesize
+rules for rule-based \nids. Our goal is to facilitate the creation
+process of rules and, consequently, harden the protection of \nids\ to
+network attacks. \tname{} synthesizes rules from malicious and benign
+traffic and from examples of correct rules.
 
@@
 
-\tname{} synthesizes rules from positive and negative examples, \ie{},
-malicious and benign traffic.
-
 \Fix{summarize how the technique works}
@@
 
-\section{Background}
+\section{Network Intrusion Detection Systems (\nids)}
+\label{sec:background}
 
-This paper focuses on signature-based \nids. Although we focused on
-the Suricata \nids, the principles we used are general to any other
-signature-based \nids~(\eg{}, Snort~\cite{snort}).
+\subsection{Rule-based and Anomaly-based \nids}
+
+\sloppy \nids{} are typically categorized in two
+groups~\cite{kumar2007survey}: rule-based (see
+Section~\ref{sec:intro}) and anomaly-based \nids. An
+\emph{anomaly-based intrusion detector} learn usage behavior from
+benign traffic and, based on that, alert uncommon
+behavior~\cite{7579764,kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}. More
+precisely, an anomaly-based \nids{} flags an attack when the network
+traffic manifests different characteristics compared to those of the
+benign traffic.
+
+Rule-based \nids\ can miss attacks as rulesets can become outdated
+whereas anomaly-based \nids\ can report false alarms as learning
+algorithms that generalize observations from bening traffic are
+approximate. To sum up, rule-based \nids{} and anomaly-based \nids{}
+are complementary---a rule-based \nids{} focuses on known attacks
+whereas anomaly-based \nids{} focuses on unknown potential attacks.
+This paper focuses on rule-based \nids\ for its tremendous popularity
+in industry. Although we focused on the Suricata \nids, the principles
+we used are general to any other signature-based \nids~(\eg{},
+Snort~\cite{snort}).
+
+\Mar{estou aqui-----------------}
 
@@
 
-\begin{wrapfigure}[14]{r}{0.5\textwidth}
+%\begin{wrapfigure}[14]{r}{0.5\textwidth}
+\begin{figure}[t!]
   \centering
-  \vspace{-7ex}
-  \scalebox{0.7}{
+%  \vspace{-7ex}
+  \scalebox{0.85}{
     \begin{tikzpicture}
@@
     options per protocol supported by Suricata.}
-\end{wrapfigure}
+%\end{wrapfigure}
+\end{figure}
+
 Suricata supports a total of \totoptions\ distinct options covering
@@
   \centering
-  \begin{tabular}{p{0.35cm}ll}
+  \begin{tabular}{clp{5.5cm}}
     \toprule
@@
 \subsection{Step 1: Reverse Engineering}
+
 \pgfplotsset{width=6cm,compat=1.8}
 \pgfplotsset{every tick label/.append style={font=\tiny}}
-The goal of the first step is to extract options from the positive
-traffic. For that, \tname{} parses the protocol message, looking for
-fields in the message associated with options (see
-Section~\ref{sec:rules-and-packets}).
-Unfortunately, inferring options at the field granularity
-level is insufficient to capture \CodeIn{content} options (see
-Section~\ref{sec:example-suricata-rules}), which refer to parts of the
-payload of the message and cannot be directly extracted directly from
-the fields. Handling this case is critically important as most rules
-contain \CodeIn{content}
-options. Figure~\ref{fig:distribution-contents} shows the histogram of
-number of these options per rule (up to 10) obtained from a public
-ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
-rules. Rules without \CodeIn{content} options is an exception and half
-of the rules contain at least two \CodeIn{content} options.  To find
-\begin{wrapfigure}[15]{r}{0.56\textwidth}
+
+%\begin{wrapfigure}[15]{r}{0.56\textwidth}
+\begin{figure}[t!]
   \centering
@@
     \CodeIn{content} options per rule (up to 10).}
-\end{wrapfigure}
+%\end{wrapfigure}
+\end{figure}
+
+The goal of the first step is to extract options from the positive
+traffic. For that, \tname{} parses the protocol message, looking for
+fields in the message associated with options (see
+Section~\ref{sec:rules-and-packets}).
+Unfortunately, inferring options at the field granularity
+level is insufficient to capture \CodeIn{content} options (see
+Section~\ref{sec:example-suricata-rules}), which refer to parts of the
+payload of the message and cannot be directly extracted directly from
+the fields. Handling this case is critically important as most rules
+contain \CodeIn{content}
+options. Figure~\ref{fig:distribution-contents} shows the histogram of
+number of these options per rule (up to 10) obtained from a public
+ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
+rules. Rules without \CodeIn{content} options is an exception and half
+of the rules contain at least two \CodeIn{content} options.  To find
 the values associated with these options, \tname{} splits the payload
@@
 \begin{algorithm}[t!]
+  \noindent\hspace{-31ex}
   \textbf{INPUT:} overconstrained rule $or$ \\
+  \noindent\hspace{-30ex} 
   \textbf{OUTPUT:} Set of minimal rules $\mathcal R$\\
-\vspace{-3ex}
+%\vspace{-3ex}
 \captionof{algorithm}{Minimization algorithm}\label{algo1}
@@
   \Do\label{start-do-while}
-      \State $\mathcal R' \gets \emptyset$\Comment{starting new
+      \State $\mathcal R' \gets \emptyset$ \Comment{starting new
         iteration. new generation should be empty.}\label{init-rprime}
@@
                  $r$, a copy of the original rule $\mathit{ro}$}
-               \If{$\mathit{isnew}(r)$ \And $\mathit{isfit}(r)$}
+               \If{$\mathit{isnew(r)} \And \mathit{isfit}(r)$} 
                    \State $\mathcal R' \cup\gets \{r\}$\Comment{$r$
@@
 
-\section*{Acknowledgments}
-This work is partially supported by RNP grant number 002951.
+%% \section*{Acknowledgments}
+%% This work is partially supported by RNP grant number 002951.
 
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-\bibliographystyle{splncs04}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{references}
-%
 
 \end{document}
+\endinput
+%%
+%% End of file `sample-sigconf.tex'.

### 2020-01-04 paper/paper.tex
another pass

@@
 
-\Mar{estou aqui-----------------}
-
 \subsection{Suricata Rules}
@@
 the header. It is a semi-colon-separated sequence of key-value
-pairs. Options serve to 1) document the analyzed traffic (\eg{},
-``msg'', ``classtype'', and ``sid'') and to 2) describe
-characteristics of the attack (\eg, ``flags'', ``threshold'', and
-``content''). \tname{} produces options of the second kind, which are
-\emph{relevant} to capturing the attack.
+pairs. Rule options serve to document the analyzed traffic (\eg{},
+``msg'', ``classtype'', and ``sid'') and to characterize the attack
+pattern (\eg, ``flags'', ``threshold'', and ``content''). \tname{}
+produces options of the second kind, which are \emph{relevant} to
+capturing the attack.
+
+\begin{table}[t!]
+  \caption{\label{table:rules}Example options of Suricata.}  
+  \centering
+  \begin{tabular}{clp{5.5cm}}
+    \toprule
+    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &  \multicolumn{1}{c}{Description}\\
+    \midrule     
+    1 & dsize & matches on the size of the packet payload\\
+    2 & itype &  matches on a specific ICMP type\\
+    3 & icode & matches on a specific ICMP code\\
+    4 & icmp\_seq  & checks a ICMP sequence number\\
+    5 & icmp\_id & matches on specific ICMP id-values\\
+    6 & window & checks the size of the TCP window\\
+    7 & flags & matches on TCP flags\\
+    8 & fragbits & checks if the fragmentation or reserved bits are set in the IP header\\
+    9 & threshold & controls the rule’s alert frequency\\
+    10 & content & checks if argument is present on the payload\\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+Table~\ref{table:rules} shows a sample of the options of
+Suricata. Column ``\#'' shows the option number, column ``Name'' shows
+the name of the option, and column ``Description'' presents a short
+description of the option purpose. We found that the option
+\CodeIn{content} is particularly prevalent in rules from public
+databases\MyComment{ (see Figure~\ref{fig:distribution-contents})}. This
+option takes a sequence of bytes as argument (\eg{},
+\CodeIn{content:''SELECT ''}) and checks if that sequence is present
+in the payload of the message. Detailed description of the Suricata
+rules can be found elsewhere~\cite{suri-rule-format}. It is worth
+noting that Suricata supports a total of \totoptions\ distinct options
+covering \numproto\ distinct protocols. A total of 9 of these options
+are for documentation.\MyComment{For instance, the purpose of rule
+  option \CodeIn{msg} is to print on the output an informative message
+  indicating what kind of intrusion has been observed.}
+Figure~\ref{fig:distribution-rules-protocol} shows the distributions
+of these options per protocol. The sum of the number of protocol
+options is higher than \totoptions\ as some options are shared across
+different protocols. \tname{} does not make distinction among protocols.
 
@@
 
-Suricata supports a total of \totoptions\ distinct options covering
-\numproto\ distinct protocols. A total of 9 of these options are for
-documentation.\MyComment{For instance, the purpose of rule option \CodeIn{msg} is to print on
-the output an informative message indicating what kind of intrusion
-has been observed.} Figure~\ref{fig:distribution-rules-protocol} shows
-the distributions of these options per protocol. The sum of
-the number of protocol options is higher than \totoptions\ as some
-options is shared across different protocols.
-
-Table~\ref{table:rules} shows a sample of the options of
-Suricata. Column ``\#'' shows the option number, column ``Name'' shows
-the name of the option, and column
-``Description'' presents a short description of the option purpose. We
-found that the option \CodeIn{content} is particularly prevalent in
-rules from public databases (see
-Figure~\ref{fig:distribution-contents}). This option takes a sequence
-of bytes as argument (\eg{}, \CodeIn{content:''SELECT ''}) and checks
-if that sequence is present in the payload of the message. Detailed
-description of the Suricata rules can be found
-elsewhere~\cite{suri-rule-format}.
-
-\begin{table}[t!]
-  \caption{\label{table:rules}Example options of Suricata.}  
-  \centering
-  \begin{tabular}{clp{5.5cm}}
-    \toprule
-    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &  \multicolumn{1}{c}{Description}\\
-    \midrule     
-    1 & dsize & matches on the size of the packet payload\\
-    2 & itype &  matches on a specific ICMP type\\
-    3 & icode & matches on a specific ICMP code\\
-    4 & icmp\_seq  & checks a ICMP sequence number\\
-    5 & icmp\_id & matches on specific ICMP id-values\\
-    6 & window & checks the size of the TCP window\\
-    7 & flags & matches on TCP flags\\
-    8 & fragbits & checks if the fragmentation or reserved bits are set in the IP header\\
-    9 & threshold & controls the rule’s alert frequency\\
-    10 & content & checks if argument is present on the payload\\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
 \subsection{Rules and Packets}
@@
 
-\begin{figure*}[h!]
+\begin{figure}[h!]
 \centering
-\includegraphics[scale=0.3]{figs/ICMP-packet-structure.png}
+\includegraphics[scale=0.28]{figs/ICMP-packet-structure.png}
 \caption{Layout of an ICMP packet.}
 \label{fig:icmp-packet-layout}
-\end{figure*}
+\end{figure}
 
-There exists a correspondence between network packets and most rule
-options. Let's consider the ICMP protocol as an
+Rule options are often associated with specific fields of network
+packets. Let's consider the ICMP protocol as an
 example. Figure~\ref{fig:icmp-packet-layout} shows the layout of an
@@
 the ICMP packet. For the field ``Flags'' the value of the
-corresponding option is encoded in its bitvector. For example, if the
+corresponding option is encoded in the bitvector. For example, if the
 ``Flags'' field stores the value 1010, the corresponding rule option
 will be \CodeIn{flagbits:RM}, as each position of the vector encodes a
-different character in the option \CodeIn{flagbits}.
-
-The value of most options can be obtained through transformation
-functions, as those described above. However, this is not the case for
-the option \CodeIn{threshold}, which is typically used in
-denial-of-service attacks. Conceptually, that option operates over
-sequence of messages as opposed to a single message. More precisely,
-to trigger an action on a rule containing that option the \nids\ tool
-needs to monitor several messages.
+different character in the \CodeIn{flagbits} rule option. The value of
+several options can be determined through transformation functions, as
+those described above. One expection to this is the option
+\CodeIn{threshold}, which is typically used in denial-of-service
+attacks. That option operates over sequence of messages as opposed to
+a single message. Consequently, it is not possible to determined its
+values from a single message. More precisely, to trigger an action on
+a rule containing that option a \nids\ tool needs to monitor several
+messages.
 
@@
 \label{sec:suri-metas-coverage}
+
+\Mar{----------------- estou aqui}
 
@@
 
-%% Explain how \tname\ behaves on this attack
-
 In the following, we illustrate how \tname{} proceeds for obtaining
-the rule for the SYN Flood attack. The main challenge in synthesizing
-rules is to discover the set of options that enables the \nids\ to
-capture the malicious traffic, but ignore the benign traffic.  The
+the rule for the SYN Flood attack. One challenge in synthesizing
+rules is to discover the set of options in a rule that makes the \nids\ to
+capture the malicious traffic, but allow the benign traffic to go through.  The
 other parts of the rule (\ie{}, the action and the header) are either
@@
 We understand that this is domain-specific information and conjecture
-that the system administrator would know how to modify the constant
-parts.
+that the system administrator would know how to customize.
 
-\tname{} proceeds as follows to synthesize rules for the SYN Flood
+\tname{} proceeds as follows to synthesize rules for this
 attack. First, it analyzes the positive traffic and extracts options
@@
 
-Note that a rule that uses this set of options would capture the
-positive traffic as these options originated from the positive
-traffic. Unfortunately, such rule would be overspecified. It would
+\Mar{<~ Lucas/Guilheme, logo acima a gente falou que threshold nao
+  poderia ser inferido a partir de uma unica mensagem e aqui a gente
+  fala isto. Vcs. podem explicar o que realmente acontece?}
+
+Note that, by construction, a rule that uses this set of options
+captures the positive traffic as these options originated from the
+positive traffic. Unfortunately, this rule is overspecified. It would
 unlikely capture different manifestations of the same
@@
 
-\Fix{...elaborate...}
+\Mar{<~ Lucas/Guilheme, vcs. podem atualizar esta informacao?}
 
 \subsection{Active Reconnaissance}
+
+\Mar{------------------------------- estou aqui}
 

### 2020-01-06 paper/paper.tex
minor change
pass on comments

@@
 
-\Mar{----------------- estou aqui}
-
 This section briefly presents examples of attacks and illustrates the
@@
   fala isto. Vcs. podem explicar o que realmente acontece?}
+\Luc{O input considerado para ataques de flood eh um conjunto de pacotes e não apenas um pacote unico, nesse caso um conjunto com 20 pacotes.
+A opção de threshold é incluida na regra quando o input contem mais de um pacote, o parametro count recebe o valor da quantidade de pacotes do conjunto de entrada.}
 
@@
 \Mar{<~ Lucas/Guilheme, vcs. podem atualizar esta informacao?}
+\Gui{}
 

### 2020-01-06 paper/paper.tex
pass

--- 2020-01-04 paper/paper.tex
@@
 In the following, we illustrate how \tname{} proceeds for obtaining
-the rule for the SYN Flood attack. One challenge in synthesizing
-rules is to discover the set of options in a rule that makes the \nids\ to
-capture the malicious traffic, but allow the benign traffic to go through.  The
-other parts of the rule (\ie{}, the action and the header) are either
-inferred or left constant. \tname{} sets the action header to
-\CodeIn{alert} and sets the header to \CodeIn{<proto> any any -> any
-  any}, indicating that it infers the target protocol and uses the
-most general flow description possible, \ie{}, it instructs the
-\nids\ to analyze the traffic flowing from/to any address and port.
-We understand that this is domain-specific information and conjecture
-that the system administrator would know how to customize.
+the rule for the SYN Flood attack. The challenge in synthesizing rules
+is to discover the set of options in a rule that makes the \nids\ to
+capture the malicious traffic, but allow the benign traffic to go
+through.  The other parts of the rule (\ie{}, the action and the
+header) are either inferred or left constant. For example, in this
+case, \tname{} sets the action header to \CodeIn{alert} and sets the
+header to \CodeIn{<proto> any any -> any any}, indicating that it
+infers the target protocol and uses the most general flow description
+possible, \ie{}, it instructs the \nids\ to analyze the traffic
+flowing from/to any address and port.  We understand that this is
+domain-specific information and assume that the system administrator
+would know how to customize it.
 
 \tname{} proceeds as follows to synthesize rules for this
-attack. First, it analyzes the positive traffic and extracts options
-expressed in that traffic. For this example, it finds the following
-set of options:
+attack. First, it analyzes the positive (\ie, malicious) traffic and
+extracts options expressed in that traffic. In this case, the
+malicious traffic is characterized as a sequence of 20
+messages. \tname{} finds the following set of options at this stage:
 
@@
   \vspace{-3ex}
-  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
+  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window}]{synflood.suricata.synth}
 %  \caption{Suricata rule for SYN Flood Attacks.}
@@
 
-\Mar{<~ Lucas/Guilheme, logo acima a gente falou que threshold nao
-  poderia ser inferido a partir de uma unica mensagem e aqui a gente
-  fala isto. Vcs. podem explicar o que realmente acontece?}
-
-Note that, by construction, a rule that uses this set of options
-captures the positive traffic as these options originated from the
-positive traffic. Unfortunately, this rule is overspecified. It would
-unlikely capture different manifestations of the same
-attack. Informally, the rule includes ``too many'' characteristics of
-the malicious traffic used to define it. To address this overfitting
-issue, \tname{} minimizes the rule preserving the invariant that
-plausible rules should only capture the positive traffic. \tname{}
-potentially finds several minimal rules. The list below shows the
-options for the five shortest rules it produces for this attack.
-
-\Mar{<~ Lucas/Guilheme, vcs. podem atualizar esta informacao?}
+%% \Mar{<~ Lucas/Guilheme, logo acima a gente falou que threshold nao
+%%   poderia ser inferido a partir de uma unica mensagem e aqui a gente
+%%   fala isto. Vcs. podem explicar o que realmente acontece?}
+%% \Luc{O input considerado para ataques de flood eh um conjunto de pacotes e não apenas um pacote unico, nesse caso um conjunto com 20 pacotes.
+A opção de threshold é incluida na regra quando o input contem mais de um pacote, o parametro count recebe o valor da quantidade de pacotes do conjunto de entrada.}
+
+\tname{} only reports a rule if the \nids\ (\ie{}, Suricata)
+instantiated with that rule is able to capture the traffic passed on
+input. Consequently, by construction, a rule that uses this set of
+options captures the malicious traffic. Unfortunately, this rule is
+potentially overspecified. Informally, the rule includes ``too many''
+characteristics of the malicious traffic used to define it and could
+potentially result in false negatives.
+
+Second, \tname{} uses negative (\ie{}, benign) traffic to address the
+overfitting issue. \tname{} searches for alternative rules that
+preserves the invariant that a rule should only capture the positive
+traffic. It uses the rule obtained in the previous step to bootstrap
+the search.
+
+Finally, \tname{} ranks the inferred rules according to heuristics
+extracted from a database of existing \tname{} rules. One example
+heuristic is \Fix{...}. Instead of traffic, this step looks for the
+structure of existing rules. The list below shows the rule options for
+the five shortest rules \tname{} produces for this attack.
+
+\Mar{<~ Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
+  Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
+  respondam isto logo.}  \Gui{}
 

### 2020-01-06 paper/paper.tex
pass intro
quick pass

@@
 
+\Mar{@Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
+  Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
+  respondam isto logo.}  \Gui{}
+
 \tname{} only reports a rule if the \nids\ (\ie{}, Suricata)
@@
 extracted from a database of existing \tname{} rules. One example
-heuristic is \Fix{...}. Instead of traffic, this step looks for the
+heuristic is \Fix{...informa um exemplo concreto que
+  vc. usaram...}. Instead of traffic, this step looks for the
 structure of existing rules. The list below shows the rule options for
-the five shortest rules \tname{} produces for this attack.
-
-\Mar{<~ Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
-  Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
-  respondam isto logo.}  \Gui{}
-
-\subsection{Active Reconnaissance}
+the top-5 rules \tname{} produces for this attack.
+
+\Mar{listem 5 primeira regras neste caso}
+
+\subsection{Active Recnnaissance}
+
+Active reconnaissance is a method used by malicious individuals to
+collection information about a computing system to determine potential
+vulnerabilities. Active reconnaissance is a preliminary step towards
+an actual attack. The most popular technique is to ``scan'' open ports
+in a host. A Ping Scan is a very common type of port scan based and it
+is based on the ICMP protocol. \Mar{@Lucas, acima falamos de
+  porta. Abaixo nao falamos de porta, apenas de maquina. Vc. pode ser
+  mais especifico? ->}To scan a network, an attacker sends a several
+ICMP Echo requests addressed to a range of IP addresses in the network
+and checks which ones respond to the requests. This allow the attacker
+to know which machines are active in the network, then it can choose
+one of them to attack.
 
 \Mar{------------------------------- estou aqui}
-
-Active reconnaissance is a method used by malicious individuals to
-gather information about vulnerabilities from a computing system. The
-most common approach is to ``scan'' open ports to discover
-vulnerabilities. Active reconnaissance is a preliminary step towards
-an attack that actually exploits the system. A Ping Scan is a very
-common type of port scan based on the ICMP
-protocol.
-\Luc{To scan a network, an attacker sends a set of ICMP Echo requests addresed to a range of IP addresses in the network and sees which one responds to the request.
-This allow the attacker to know which machines are active in the network, then it can choose one of them to attack.}
 

### 2020-01-06 paper/paper.tex
pass

@@
 
-\subsection{Active Recnnaissance}
+\subsection{Active Reconnaissance}
 
@@
 
-\Mar{------------------------------- estou aqui}
-
 \begin{figure}[h!]
@@
 
-\Luc{Figure 4 shows a Suricata rule that capture a ping scan performed by the NMAP Tool.
-The relevant options are in bold---the option \CodeIn{dsize: 0}, indicates that the packet has no bytes in the payload.
-\CodeIn{itype: 8} identifies the packet as a ICMP Echo request.}
-
-\Luc{To synthesize rules for a Ping Scan, \tname{} analyzes the positive traffic and extract the following set of options:}
+Figure 4 shows a Suricata rule that captures a ping scan implemented
+with the NMAP Tool~\cite{netmap}. The relevant options are in
+bold---the option \CodeIn{dsize: 0} indicates that the packet has no
+bytes in the payload and the option \CodeIn{itype: 8} indicates that
+the packet to be captured are ICMP Echo requests.
+
+Again, to synthesize rules for a Ping Scan, \tname{} analyzes the
+positive traffic and extracts the following rule options:
 
 \begin{figure}[h]
-\vspace{-3ex} \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{pingscan.suricata.synth}
-%  \caption{Suricata rule for SYN Flood Attacks.}
-  %  \label{fig:synflood-example-synt}
+  \vspace{-2ex}
+  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype}]{pingscan.suricata.synth}
   \vspace{-3ex}  
@@
 
-\Luc{Again, a rule with this options would capture the positive traffic, but it would be overspecified.
-So, \tname{} tries to minimize the rule...
-After the minimization, it produces the following rule:}
-\Luc{}
-
-\subsection{\Fix{...some other with contents...}}
+Again, note that rule \CodeIn{icode} was not present in the original
+rule.  \tname{} produces the following five rules at the top of the
+ranking after minimization and ranking.
+
+\Fix{...show rules...}
+
+\subsection{\Fix{Name of attack ***with content***}}
 \label{sec:content-example}
 
-\Fix{...elaborate...}
+\Fix{...please elaborate. We need on example that shows that the
+  search can be non-trivial. In the examples above, we had to discard
+  only one rule option!}
 
@@
 \section{Technique}
+
+\Mar{----------------------------------- estou aqui}
 

### 2020-01-06 paper/paper.tex
minor

@@
 Figure 4 shows a Suricata rule that captures a ping scan implemented
-with the NMAP Tool~\cite{netmap}. The relevant options are in
-bold---the option \CodeIn{dsize: 0} indicates that the packet has no
-bytes in the payload and the option \CodeIn{itype: 8} indicates that
-the packet to be captured are ICMP Echo requests.
+with NMAP~\cite{netmap}, a network discovery and security auditing
+open-source tool. The relevant options are in bold---the option
+\CodeIn{dsize: 0} indicates that the packet has no bytes in the
+payload and the option \CodeIn{itype: 8} indicates that the packet to
+be captured are ICMP Echo requests.
 
@@
 
+\tname{} is a technique that uses (1) benign and malicious network
+traffic and (2) existing signature-based \nids\ rules to synthesize
+new rules. The goal of \tname\ is to create plausible rules that
+\emph{only} captures the positive traffic, \ie{}, the malicious
+messages.
+
 \Mar{----------------------------------- estou aqui}
-
-\tname{} is a technique that uses positive and negative examples to
-synthesize signature-based \nids\ rules, such as those from Snort and
-Suricata. The goal of \tname\ is to create a rule that \emph{only}
-captures the positive traffic, \ie{}, the malicious sequence of
-messages.
 

### 2020-01-06 paper/paper.tex
pass on workflow (input and output)
pass

@@
 
-\tname{} is a technique that uses (1) benign and malicious network
-traffic and (2) existing signature-based \nids\ rules to synthesize
-new rules. The goal of \tname\ is to create plausible rules that
-\emph{only} captures the positive traffic, \ie{}, the malicious
-messages.
-
-\Mar{----------------------------------- estou aqui}
+\tname{} is a technique to synthesize signature-based rules for
+\nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
+goal of \tname\ is to create rules that captures \emph{only} the
+malicious messages.
 
@@
 \centering
-\includegraphics[trim=0 300 350 0,clip,width=0.95\textwidth]{figs/nids-workflow}
+\includegraphics[trim=0 350 100 0,clip,width=0.65\textwidth]{figs/nids-workflow}
 \caption{\tname\ workflow.}
@@
 \vspace{1ex}
-\noindent\textbf{Overview.}~ The input of the technique is the
-positive traffic whereas the output is a list of minimized rules
-capturing only the positive, malicious,
-traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
-pipeline of two components, which appear numbered in the figure.
-\tname{} synthesizes rules in two steps. First, it produces a
+\noindent\textbf{Overview.}~To create rules, \tname\ leverages (1)
+benign and malicious network traffic and (2) rules from public
+rulesets. Figure~\ref{fig:overview} shows the workflow of \tname{} as
+a pipeline of three components, which appear numbered in the figure.
+The inputs of the technique appear at the bottom of the figure whereas
+the output appear at the right. \tname\ takes on input malicious and
+benign traffic and a set of rules from public rulesets and produces on
+output a list of plausible rules to isolate the malicious traffic.
+
+\Mar{------------------ estou aqui}
+
+First, it produces a
 potentially overspecified rule that captures the malicious traffic
@@
   \centering
-  \vspace{-4ex}
+%  \vspace{-4ex}
   \scalebox{1.2}{  

### 2020-01-06 paper/paper.tex
comment on 3.1

@@
   Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
-  respondam isto logo.}  \Gui{}
+  respondam isto logo.}  \Gui{O window:64 seria removido sim. Na documentação do Snort tem isso:
+
+The reserved bits can be used to detect unusual behavior, such as IP stack fingerprinting attempts or other suspicious activity.
+
+O que realmente caracteriza o ataque eh apenas a flag S e o numero elevado de pacotes, os bits reservados 12 auxiliam na deteccao desse ataque.
+
+Alem disso, tem mais:
+
+ The reserved bits '1' and '2' have been replaced with 'C' and 'E', respectively, to match RFC 3168
+
+ O 12 ainda funcionam no programa, mas nao sao mais utilizados.}
 

### 2020-01-06 paper/paper.tex
pass on comments

--- 2020-01-06 paper/paper.tex
@@
   Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
-  respondam isto logo.}  \Gui{}
+  respondam isto logo.}  \Gui{O window:64 seria removido sim.}
+  \Luc{O segundo campo da opção flags define flags que serão ignoradas.
+Nesse caso a regra procura por pacotes onde apenas a flag S está 'setada', independentemente do valor das flags 1 e 2.
+Portanto a regra gerada pode capturar coisas diferentes da regra original.
+A inferencia do "12" n pode ser feita apenas a partir do pacote de entrada, precisariamos tratar isso.}
 
@@
   porta. Abaixo nao falamos de porta, apenas de maquina. Vc. pode ser
-  mais especifico? ->}To scan a network, an attacker sends a several
+  mais especifico? ->}
+  \Luc{De fato estamos falando de duas coisas diferentes, mas ambas são exemplos de 'active reconnaissance'.
+Primeiramente citamos o port scanning, q tem o objetivo de encontrar portas abertas em uma maquina especifica.
+Depois falamos do ping scan, q tem o objetivo de encontrar ip's ativos numa rede.} To scan a network, an attacker sends a several
 ICMP Echo requests addressed to a range of IP addresses in the network

### 2020-01-06 paper/paper.tex
pass on section 4

@@
 
-\Mar{------------------ estou aqui}
-
-First, it produces a
-potentially overspecified rule that captures the malicious traffic
-provided on input. In this step, \tname\ finds options by
-\emph{reverse engineering} the positive traffic. It extracts tokens
-from the malicious message and uses them to define the initial
-rule. We say that this rule overfits the problem as it satisfies the
-success criterion of only capturing the positive traffic, however it
-is unlikely that it will be able to generalize to other attacks,
-\ie{}, attacks of the same kind, but with different data. The reason
-is that many options added to the rule are unimportant for
-discriminating the attack. In the second step, \tname{} uses negative
-examples to \emph{minimize} the rule, discarding unnecessary
-options. Note that any subset of the original set of options captures
-the positive traffic. The minimization step uses a population-based
-gradient descent-like search to generate minimal solutions.
+\tname\ works as follows. First, it produces a potentially
+overspecified rule that captures the malicious traffic, provided on
+input in the form of pcap files~\cite{pcap}. In this step,
+\tname\ detects rule options in correspondence with the input messages
+(see Section~\ref{sec:rules-and-packets}). It is worth noting that the
+rule produced in this step can include more options than the golden
+rule (see Section~\ref{sec:suri-metas-coverage}). Consequently, the
+use of that rule could lead, in principle, to false negatives as not
+all manifestations of the same attack are identical, \ie{}, the
+content of messages for two instantiations of a given attack may
+differ. Second, \tname{} generates plausible rules similar to the one
+produced in the previous step. It is worth noting that any subset of
+the original set of options captures the malicious traffic. As the
+generation procedure discards options from the initial rule, it
+satisfies the invariant of capturing malicious traffic by
+construction. However, the goal of \tname\ is to produce rules that
+capture \emph{only} malicious traffic. To achieve that goal, \tname{}
+uses benign traffic from public databases to check that the
+\nids\ instantiated with a given rule does not capture benign
+traffic. The minimization step uses a population-based gradient
+descent-like search to find all plausible rules, which are derived
+from the rule obtained in the previous step. Third,
+\Fix{...elaborate...}
 

### 2020-01-07 paper/paper.tex
comments

@@
 
-%% \Mar{<~ Lucas/Guilheme, logo acima a gente falou que threshold nao
-%%   poderia ser inferido a partir de uma unica mensagem e aqui a gente
-%%   fala isto. Vcs. podem explicar o que realmente acontece?}
-%% \Luc{O input considerado para ataques de flood eh um conjunto de pacotes e não apenas um pacote unico, nesse caso um conjunto com 20 pacotes.
-A opção de threshold é incluida na regra quando o input contem mais de um pacote, o parametro count recebe o valor da quantidade de pacotes do conjunto de entrada.}
-
 \Mar{@Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
   Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
-  respondam isto logo.}  \Gui{O window:64 seria removido sim.}
-  \Luc{O segundo campo da opção flags define flags que serão ignoradas.
-Nesse caso a regra procura por pacotes onde apenas a flag S está 'setada', independentemente do valor das flags 1 e 2.
-Portanto a regra gerada pode capturar coisas diferentes da regra original.
-A inferencia do "12" n pode ser feita apenas a partir do pacote de entrada, precisariamos tratar isso.}
+  respondam isto logo.}  \Gui{O window:64 seria removido
+  sim.}\Mar{ok. vou explicar isto.}
+
+  \Luc{O segundo campo da opção flags define flags que serão
+    ignoradas. Nesse caso a regra procura por pacotes onde apenas a
+    flag S está 'setada', independentemente do valor das flags 1 e
+    2. Portanto a regra gerada pode capturar coisas diferentes da
+    regra original. A inferencia do "12" n pode ser feita apenas a
+    partir do pacote de entrada, precisariamos tratar
+    isso.}\Mar{vc. pode explicar isto **mais detalhadamente** E (1)
+    explicar que isto eh uma limitacao e (2) indicar a possivel consequencia
+    disto?}
 

### 2020-01-07 paper/paper.tex
pass on comments

@@
     partir do pacote de entrada, precisariamos tratar
-    isso.}\Mar{vc. pode explicar isto **mais detalhadamente** E (1)
-    explicar que isto eh uma limitacao e (2) indicar a possivel consequencia
-    disto?}
+    isso.}\Mar{vc.
+pode explicar isto **mais detalhadamente** \Luc{a opção flags pode ter dois campos separados por virgula (ex: flags:S,12).
+o primeiro campo indica quais flags devem estar ativas em um pacote para que o suricata gere um alerta.
+se uma ou mais flags forem irrelevantes para a detecção de um determinado ataque (ou seja, se n importa se a flag está ou não ativa), elas serão adicionadas ao segundo campo.} E (1)
+    explicar que isto eh uma limitacao \Luc{n entendi exatamente oq explicar aqui.} e (2) indicar a possivel consequencia
+    disto?
+\Luc{a consequencia sao possiveis falsos negativos.
+ao não incluir flags no segundo campo podemos descartar pacotes por conta da presença de flags q n são relevantes (ou seja, vamos levar em conta flags q n deveriam interferir na detecção).}}
 

### 2020-01-07 paper/paper.tex
pass on section 3.1

--- 2020-01-07 paper/paper.tex
@@
 \begin{figure}[h]
-  \vspace{-3ex}
+  \vspace{-2ex}
   \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window}]{synflood.suricata.synth}
@@
   %  \label{fig:synflood-example-synt}
-  \vspace{-3ex}  
+  \vspace{-2ex}  
 \end{figure}
 
-\Mar{@Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
-  Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
-  respondam isto logo.}  \Gui{O window:64 seria removido
-  sim.}\Mar{ok. vou explicar isto.}
-
-  \Luc{O segundo campo da opção flags define flags que serão
-    ignoradas. Nesse caso a regra procura por pacotes onde apenas a
-    flag S está 'setada', independentemente do valor das flags 1 e
-    2. Portanto a regra gerada pode capturar coisas diferentes da
-    regra original. A inferencia do "12" n pode ser feita apenas a
-    partir do pacote de entrada, precisariamos tratar
-    isso.}\Mar{vc. pode explicar isto **mais detalhadamente** E (1)
-    explicar que isto eh uma limitacao e (2) indicar a possivel consequencia
-    disto?}
-
-\tname{} only reports a rule if the \nids\ (\ie{}, Suricata)
+%% \Mar{@Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
+%%   Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
+%%   respondam isto logo.}  \Gui{O window:64 seria removido
+%%   sim.}\Mar{ok. vou explicar isto.}
+  %% \Luc{O segundo campo da opção flags define flags que serão
+  %%   ignoradas. Nesse caso a regra procura por pacotes onde apenas a
+  %%   flag S está 'setada', independentemente do valor das flags 1 e
+  %%   2. Portanto a regra gerada pode capturar coisas diferentes da
+  %%   regra original. A inferencia do "12" n pode ser feita apenas a
+  %%   partir do pacote de entrada, precisariamos tratar
+  %%   isso.}\Mar{vc.
+pode explicar isto **mais detalhadamente** \Luc{a opção flags pode ter dois campos separados por virgula (ex: flags:S,12).
+o primeiro campo indica quais flags devem estar ativas em um pacote para que o suricata gere um alerta.
+se uma ou mais flags forem irrelevantes para a detecção de um determinado ataque (ou seja, se n importa se a flag está ou não ativa), elas serão adicionadas ao segundo campo.} E (1)
+  %%   explicar que isto eh uma limitacao \Luc{n entendi exatamente oq explicar aqui.} e (2) indicar a possivel consequencia
+  %%   disto?
+\Luc{a consequencia sao possiveis falsos negativos.
+ao não incluir flags no segundo campo podemos descartar pacotes por conta da presença de flags q n são relevantes (ou seja, vamos levar em conta flags q n deveriam interferir na detecção).}}
+
+\noindent
+\tname{} only reports a rule if the \nids\ (\eg{}, Suricata)
 instantiated with that rule is able to capture the traffic passed on
@@
 characteristics of the malicious traffic used to define it and could
-potentially result in false negatives.
+potentially result in false negatives. In this particular case, note
+that the option \CodeIn{windows: 64} is not included in the golden
+rule.
 
@@
 overfitting issue. \tname{} searches for alternative rules that
-preserves the invariant that a rule should only capture the positive
-traffic. It uses the rule obtained in the previous step to bootstrap
-the search.
+preserves the invariant that a rule should capture the positive
+traffic but not capture the negative traffic. It uses the rule
+obtained in the previous step to bootstrap the search and generates
+various rules containing subsets of these three options.
 
@@
 \Mar{listem 5 primeira regras neste caso}
+
+\noindent
+\textbf{Differences.}~Note that there are differences between the
+options produced in the first step of \tname{} and the options from
+the golden rule, as per Figure~\ref{fig:synflood-example}. The option
+\CodeIn{flags} does not mention parameters \CodeIn{1 2}, there is an
+extra option \CodeIn{window}, and there are differences in the option
+\CodeIn{threshold}. Considering the option \CodeIn{flags}, the first
+parameter of that option indicates what flags in the packet must be
+set while the second parameter indicates non-mandatory flags (\ie{},
+``don't cares''). The option \CodeIn{flags: S} indicates that the flag
+``S'' must be set and all flags not mentioned (``1'' and ``2'', in our
+case) must be unset for the \nids\ to capture the
+traffic. Consequently, that choice of option alone could result in
+false negatives and false positives. An attack that satisfies all
+other conditions in a rule and have flags ``1'' or ``2'' set would be
+captured by the original rule but not the inferred rule---a case of
+false negative. Likewise, an attack that satisfies all conditions and
+have flags ``1'' and ``2'' unset would be captured by the inferred
+rule but not the original rule---a case of false positive. The set of
+positive examples is to blame in this case. To mitigate this issue,
+the input messages should have covered cases with options ``1'' or
+``2'' set. Considering option \CodeIn{windows: 64}, \tname{} is able
+to generate rules without that option, as discussed above. Considering
+option \CodeIn{threshold}, the values used in the golden rule are
+domain-specific. \tname{} uses default values for these options. We
+conjecture that the system administrator would be able to configure
+this option.
 
@@
   \caption{Suricata rule for Ping Scan.}
-  \label{fig:synflood-example}
+  \label{fig:pingscan-example}
 \end{figure}

### 2020-01-07 paper/paper.tex
pass on examples

@@
 extracted from a database of existing \tname{} rules. One example
-heuristic is \Fix{...informa um exemplo concreto que
+heuristic is \Luc{Ataque "JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit"
+. O payload do pacote é o seguinte:
+\begin{figure}[H]
+  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-rule.suricata}
+  \label{fig:rule-example}
+\end{figure}
+
+A partir disso produzimos uma regra com os seguintes contents:
+
+\begin{figure}[H]
+  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-contents-splitted.suricata}
+  \label{fig:rule-example}
+\end{figure}
+
+Depois da etapa da minimização, as top 5 regras são:
+
+\begin{figure}[H]
+  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-top5-rules.suricata}
+  \label{fig:rule-example}
+\end{figure}
+
+A regra de ouro é:
+
+\begin{figure}[H]
+  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-golden-rule.suricata}
+  \label{fig:rule-example}
+\end{figure}
+
+}
+
+\Fix{...informa um exemplo concreto que
   vc. usaram...}. Instead of traffic, this step looks for the

### 2020-01-07 paper/paper.tex
changes

--- 2020-01-07 paper/paper.tex
@@
 \noindent
-\tname{} only reports a rule if the \nids\ (\eg{}, Suricata)
-instantiated with that rule is able to capture the traffic passed on
-input. Consequently, by construction, a rule that uses this set of
-options captures the malicious traffic. Unfortunately, this rule is
-potentially overspecified. Informally, the rule includes ``too many''
-characteristics of the malicious traffic used to define it and could
-potentially result in false negatives. In this particular case, note
-that the option \CodeIn{windows: 64} is not included in the golden
-rule.
-
-Second, \tname{} uses negative (\ie{}, benign) traffic to address the
-overfitting issue. \tname{} searches for alternative rules that
-preserves the invariant that a rule should capture the positive
-traffic but not capture the negative traffic. It uses the rule
-obtained in the previous step to bootstrap the search and generates
-various rules containing subsets of these three options.
+\tname{} only reports a rule if the \nids\ instantiated with that rule
+is able to capture the traffic passed on input. Consequently, by
+construction, a rule that uses this set of options captures the
+malicious traffic. Unfortunately, this rule is potentially
+overspecified. Informally, the rule includes ``too specific''
+characteristics of the malicious traffic and could potentially result
+in false negatives. For example, in this particular case, note that
+the option \CodeIn{windows: 64} is not included in the golden rule
+from Figure~\ref{fig:synflood-example}.
+
+Second, \tname{} uses negative (\ie{}, benign) traffic to mitigate the
+overfitting issue. In this step, \tname{} searches for alternative
+rules that preserves the invarianxt that a rule should capture the
+positive traffic but not capture the negative traffic. It uses the
+rule obtained in the previous step, containing three options, to
+bootstrap the search for rules. The search generates rules containing
+subsets of the options from the seed rule.
+
+%% \Luc{Ataque "JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit"
+%% . O payload do pacote é o seguinte:
+%% \begin{figure}[H]
+%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-rule.suricata}
+%%   \label{fig:rule-example}
+%% \end{figure}
+
+%% A partir disso produzimos uma regra com os seguintes contents:
+
+%% \begin{figure}[H]
+%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-contents-splitted.suricata}
+%%   \label{fig:rule-example}
+%% \end{figure}
+
+%% Depois da etapa da minimização, as top 5 regras são:
+
+%% \begin{figure}[H]
+%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-top5-rules.suricata}
+%%   \label{fig:rule-example}
+%% \end{figure}
+
+%% A regra de ouro é:
+
+%% \begin{figure}[H]
+%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-golden-rule.suricata}
+%%   \label{fig:rule-example}
+%% \end{figure}
+
+%% }
 
 Finally, \tname{} ranks the inferred rules according to heuristics
-extracted from a database of existing \tname{} rules. One example
-heuristic is \Fix{...informa um exemplo concreto que
-  vc. usaram...}. Instead of traffic, this step looks for the
-structure of existing rules. The list below shows the rule options for
-the top-5 rules \tname{} produces for this attack.
+extracted from a database of existing rules. One example heuristic
+function we used to rank the generated rules was \Fix{...informa um
+  exemplo concreto que vc. usaram...}. The list below shows the rule
+options for the top-5 rules \tname{} produces for this attack.
 
 \Mar{listem 5 primeira regras neste caso}
+
+To sum up, \tname{} uses the messages that manifest an attack in the
+first step (to create an overspecific seed rule), benign traffic in
+the second step (to create a set of plausible rules from the seed
+rule), and existing rules from popular sources in the third step (to
+rank the generated rules).
 
@@
 
+%% \Mar{@Lucas, acima falamos de porta. Abaixo nao falamos de
+%%   porta, apenas de maquina. Vc. pode ser mais especifico? ->} \Luc{De
+%%   fato estamos falando de duas coisas diferentes, mas ambas são
+%%   exemplos de 'active reconnaissance'. Primeiramente citamos o port
+%%   scanning, q tem o objetivo de encontrar portas abertas em uma
+%%   maquina especifica. Depois falamos do ping scan, q tem o objetivo de
+%%   encontrar ip's ativos numa rede.}
+
 Active reconnaissance is a method used by malicious individuals to
-collection information about a computing system to determine potential
+collect information about a computing system to determine potential
 vulnerabilities. Active reconnaissance is a preliminary step towards
-an actual attack. The most popular technique is to ``scan'' open ports
-in a host. A Ping Scan is a very common type of port scan based and it
-is based on the ICMP protocol. \Mar{@Lucas, acima falamos de
-  porta. Abaixo nao falamos de porta, apenas de maquina. Vc. pode ser
-  mais especifico? ->}
-  \Luc{De fato estamos falando de duas coisas diferentes, mas ambas são exemplos de 'active reconnaissance'.
-Primeiramente citamos o port scanning, q tem o objetivo de encontrar portas abertas em uma maquina especifica.
-Depois falamos do ping scan, q tem o objetivo de encontrar ip's ativos numa rede.} To scan a network, an attacker sends a several
-ICMP Echo requests addressed to a range of IP addresses in the network
-and checks which ones respond to the requests. This allow the attacker
-to know which machines are active in the network, then it can choose
-one of them to attack.
+an actual attack. For example, Ping Scan is a very common type of
+attack that scans the network for IPs of accessible hosts. An attacker
+sends several ICMP Echo requests to a range of IP addresses in the
+network and checks which ones respond to the requests. This allow the
+attacker to know which machines are active in the network. Then (s)he
+can choose one of them to attack.
 
@@
 
-Figure 4 shows a Suricata rule that captures a ping scan implemented
-with NMAP~\cite{netmap}, a network discovery and security auditing
-open-source tool. The relevant options are in bold---the option
-\CodeIn{dsize: 0} indicates that the packet has no bytes in the
-payload and the option \CodeIn{itype: 8} indicates that the packet to
-be captured are ICMP Echo requests.
+Figure 4 shows a Suricata rule that captures a reproduction of a Ping
+Scan attack available from NMAP~\cite{netmap}, a network discovery and
+security auditing open-source tool. The relevant options are in
+bold---the option \CodeIn{dsize: 0} indicates that the packet has no
+bytes in the payload and the option \CodeIn{itype: 8} indicates that
+the packet to be captured are ICMP Echo requests.
 
 Again, to synthesize rules for a Ping Scan, \tname{} analyzes the
-positive traffic and extracts the following rule options:
+positive malicious traffic and extracts the following rule options:
 
@@
 
-\Fix{...show rules...}
+\Fix{...listem 5 primeiras regras neste caso...}
 
@@
 
-\tname{} is a technique to synthesize signature-based rules for
+\tname{} is a technique to synthesize rules for signature-based
 \nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
-goal of \tname\ is to create rules that captures \emph{only} the
-malicious messages.
+goal of \tname\ is to create rules that capture \emph{only} the
+malicious traffic.
 
@@
 The inputs of the technique appear at the bottom of the figure whereas
-the output appear at the right. \tname\ takes on input malicious and
+the output appears at the right. \tname\ takes on input malicious and
 benign traffic and a set of rules from public rulesets and produces on

### 2020-01-07 paper/paper.tex
pass on workflow

@@
 use of that rule could lead, in principle, to false negatives as not
-all manifestations of the same attack are identical, \ie{}, the
-content of messages for two instantiations of a given attack may
-differ. Second, \tname{} generates plausible rules similar to the one
-produced in the previous step. It is worth noting that any subset of
-the original set of options captures the malicious traffic. As the
-generation procedure discards options from the initial rule, it
-satisfies the invariant of capturing malicious traffic by
-construction. However, the goal of \tname\ is to produce rules that
-capture \emph{only} malicious traffic. To achieve that goal, \tname{}
-uses benign traffic from public databases to check that the
-\nids\ instantiated with a given rule does not capture benign
-traffic. The minimization step uses a population-based gradient
-descent-like search to find all plausible rules, which are derived
-from the rule obtained in the previous step. Third,
-\Fix{...elaborate...}
+all manifestations of the same attack are identical. Second, \tname{}
+generates plausible rules similar to the one produced in the previous
+step. It is worth noting that any subset of the original set of
+options captures the malicious traffic. As the generation procedure
+discards options from the initial rule, it satisfies the invariant of
+capturing malicious traffic by construction. However, the goal of
+\tname\ is to produce rules that capture \emph{only} malicious
+traffic. To achieve that goal, \tname{} uses benign traffic from
+public databases to check that the \nids\ instantiated with a given
+rule does not capture benign traffic. The minimization step uses a
+population-based gradient descent-like search to find all plausible
+rules, which are derived from the rule obtained in the previous
+step. Third, uses several heuristic functions derived from a dataset
+of rules to rank the rules generated in the previous step.
 

### 2020-01-07 paper/paper.tex
pass on technique

@@
 
+The goal of the first step is to extract options from the positive
+traffic. For that, \tname{} parses the protocol message, looking for
+fields in the message associated with options (see
+Section~\ref{sec:rules-and-packets}).  Unfortunately, inferring
+options at the field granularity level is insufficient to capture
+\CodeIn{content} options (see
+Section~\ref{sec:example-suricata-rules}), which refer to parts of the
+payload of the message and cannot be extracted directly from the
+fields. Handling this case is critically important as \Fix{xx}\% of the
+rules contain \CodeIn{content} options as the one from
+Section~\ref{sec:content-example}.
+
 \pgfplotsset{width=6cm,compat=1.8}
@@
 
-The goal of the first step is to extract options from the positive
-traffic. For that, \tname{} parses the protocol message, looking for
-fields in the message associated with options (see
-Section~\ref{sec:rules-and-packets}).
-Unfortunately, inferring options at the field granularity
-level is insufficient to capture \CodeIn{content} options (see
-Section~\ref{sec:example-suricata-rules}), which refer to parts of the
-payload of the message and cannot be directly extracted directly from
-the fields. Handling this case is critically important as most rules
-contain \CodeIn{content}
-options. Figure~\ref{fig:distribution-contents} shows the histogram of
+Figure~\ref{fig:distribution-contents} shows the histogram of the
 number of these options per rule (up to 10) obtained from a public
@@
 of the malicious message in tokens using natural language delimiters,
-such as $\backslash$t, $\backslash$n, $\backslash$r, and spaces. The
-rationale for this decision is that we observed, by inspecting
-existing rules, that \CodeIn{content} options typically refer to
-sequence of characters in the payload of messages separated by these
-delimiters. This tokenization step can generate dozens of options. For
-the example from Section~\ref{sec:content-example}, \tname{} produces
-a total of \Fix{xx} content options.
+such as $\backslash$t, $\backslash$n, $\backslash$r, and spaces. We
+observed, by inspecting existing rules, that \CodeIn{content} options
+typically refer to sequence of characters in the payload of messages
+separated by these delimiters. This tokenization step can generate
+dozens of options. For the example from
+Section~\ref{sec:content-example}, \tname{} produces a total of
+\Fix{xx} content options.
+
+\Mar{pessoal, isto explica como inferimos opcoes de content, mas nao
+  explica como inferimos as outras opcoes. Em particular, as opcoes
+  dos primeiros dois exemplos da Secao 3.}
+
+\Mar{-------------------------------- estou aqui}
 

### 2020-01-07 paper/paper.tex
pass on examples

@@
 extracted from a database of existing rules. One example heuristic
-function we used to rank the generated rules was \Fix{...informa um
+function we used to rank the generated rules was
+\Luc{
+%\begin{wrapfigure}[14]{r}{0.5\textwidth}
+\begin{figure}[H]
+  \centering
+%  \vspace{-7ex}
+  \scalebox{0.85}{
+    \begin{tikzpicture}
+      \begin{axis}[
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          %          ylabel={\#rules},
+          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          xtick=data,
+          xticklabel style={rotate=45},          
+          nodes near coords,
+          every node near coord/.append style={font=\scriptsize},        
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
+%        \legend{relevant, irrelevant}
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \vspace{-2ex}
+  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
+%\end{wrapfigure}
+\end{figure}
+
+}
+
+ \Fix{...informa um
   exemplo concreto que vc. usaram...}. The list below shows the rule

### 2020-01-08 paper/paper.tex
pass on examples
comments
pass on 3 and 4.1

@@
 \subsection{Denial-of-Service}
+\label{sec:dos}
 
@@
 function we used to rank the generated rules was
-\Luc{
-%\begin{wrapfigure}[14]{r}{0.5\textwidth}
-\begin{figure}[H]
-  \centering
-%  \vspace{-7ex}
-  \scalebox{0.85}{
-    \begin{tikzpicture}
-      \begin{axis}[
-          ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
-          %          ylabel={\#rules},
-          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
-          xtick=data,
-          xticklabel style={rotate=45},          
-          nodes near coords,
-          every node near coord/.append style={font=\scriptsize},        
-          nodes near coords align={vertical},
-        ]
-        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
-%        \legend{relevant, irrelevant}
-      \end{axis}
-    \end{tikzpicture}
-  }
-  \vspace{-2ex}
-  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
-%\end{wrapfigure}
-\end{figure}
-
-}
-
- \Fix{...informa um
-  exemplo concreto que vc. usaram...}. The list below shows the rule
-options for the top-5 rules \tname{} produces for this attack.
+\Fix{...elaborate...}. The list below shows the rule options for the
+top-5 rules \tname{} produces for this attack.
 
@@
 \subsection{Active Reconnaissance}
+\label{sec:active-recon}
 
@@
 \begin{figure}[h!]
+  \vspace{1ex}
   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
@@
   \label{fig:pingscan-example}
+  \vspace{1ex}  
 \end{figure}
@@
 \nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
-goal of \tname\ is to create rules that capture \emph{only} the
+goal is to create rules that capture \emph{only} the
 malicious traffic.
 
-\begin{figure*}[t]
+\begin{figure*}[t!]
 \centering
@@
 benign and malicious network traffic and (2) rules from public
-rulesets. Figure~\ref{fig:overview} shows the workflow of \tname{} as
-a pipeline of three components, which appear numbered in the figure.
+rulesets. \tname\ takes on input malicious and benign traffic and a
+set of rules from public rulesets and produces on output a list of
+plausible rules to isolate the malicious
+traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
+pipeline of three components, which appear numbered in the figure.
 The inputs of the technique appear at the bottom of the figure whereas
-the output appears at the right. \tname\ takes on input malicious and
-benign traffic and a set of rules from public rulesets and produces on
-output a list of plausible rules to isolate the malicious traffic.
+the output appears at the right.
 
@@
 The goal of the first step is to extract options from the positive
-traffic. For that, \tname{} parses the protocol message, looking for
-fields in the message associated with options (see
-Section~\ref{sec:rules-and-packets}).  Unfortunately, inferring
-options at the field granularity level is insufficient to capture
-\CodeIn{content} options (see
-Section~\ref{sec:example-suricata-rules}), which refer to parts of the
+traffic, which is characterized with pcap files~\cite{pcap}, a popular
+format to describe network packets.
+
+For that, \tname{} parses the message looking for options associated
+with the data. As described in Section~\ref{sec:rules-and-packets},
+there is a correspondence between packet fields and rule
+options. \tname{} leverages that mapping to infer options. The
+examples of sections~\ref{sec:dos} and~\ref{sec:active-recon} show
+cases where \tname{} infers options with this mapping. Unfortunately,
+inferring options at the field granularity level is insufficient to
+capture \CodeIn{content} options as the described on
+Section~\ref{sec:content-example}. These options refer to parts of the
 payload of the message and cannot be extracted directly from the
-fields. Handling this case is critically important as \Fix{xx}\% of the
-rules contain \CodeIn{content} options as the one from
-Section~\ref{sec:content-example}.
+fields. Handling this case is critically important as \Fix{xx}\% of
+the rules contain \CodeIn{content} options.
 
@@
 \pgfplotsset{every tick label/.append style={font=\tiny}}
-
-%\begin{wrapfigure}[15]{r}{0.56\textwidth}
 \begin{figure}[t!]
   \centering
-%  \vspace{-4ex}
   \scalebox{1.2}{  
@@
             anchor=north,legend columns=-1},
-          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          symbolic x coords={1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
           xtick=data,
@@
         ]
-        \addplot coordinates {(0,25) (1,10222) (2,4956) (3,5589)
+        \addplot coordinates {(1,10222) (2,4956) (3,5589)
           (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
@@
   \caption{\label{fig:distribution-contents}Histogram of number of
-    \CodeIn{content} options per rule (up to 10).}
-%\end{wrapfigure}
+    \CodeIn{content} options per rule.}
 \end{figure}
@@
 Figure~\ref{fig:distribution-contents} shows the histogram of the
-number of these options per rule (up to 10) obtained from a public
+number of these options per rule obtained from a public
 ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
-rules. Rules without \CodeIn{content} options is an exception and half
-of the rules contain at least two \CodeIn{content} options.  To find
-the values associated with these options, \tname{} splits the payload
-of the malicious message in tokens using natural language delimiters,
-such as $\backslash$t, $\backslash$n, $\backslash$r, and spaces. We
-observed, by inspecting existing rules, that \CodeIn{content} options
-typically refer to sequence of characters in the payload of messages
-separated by these delimiters. This tokenization step can generate
-dozens of options. For the example from
-Section~\ref{sec:content-example}, \tname{} produces a total of
-\Fix{xx} content options.
-
-\Mar{pessoal, isto explica como inferimos opcoes de content, mas nao
-  explica como inferimos as outras opcoes. Em particular, as opcoes
-  dos primeiros dois exemplos da Secao 3.}
-
-\Mar{-------------------------------- estou aqui}
+rules. We limited the number of content options per rule to 10 for
+space. Rules without \CodeIn{content} options is an exception
+(\Fix{xx}\%) and half of the rules contain at least two
+\CodeIn{content} options.  To find the values associated with these
+options, \tname{} splits the payload of the malicious message in
+tokens using natural language delimiters, such as $\backslash$t,
+$\backslash$n, $\backslash$r, and spaces. We observed, by inspecting
+existing rules, that \CodeIn{content} options typically refer to
+sequence of characters in the payload of messages separated by these
+delimiters. This tokenization step can generate dozens of options.
+\tname{} produced a total of \Fix{xx} content options for the example
+Section~\ref{sec:content-example}.
 
@@
 \label{sec:minimization}
+
+\Mar{-------------------------- estou aqui}
 
@@
 
+\subsection{Step 3: Ranking}
+
+\Fix{...elaborate...}
+
+\begin{figure}[H]
+  \centering
+  \scalebox{0.85}{
+    \begin{tikzpicture}
+      \begin{axis}[
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          %          ylabel={\#rules},
+          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          xtick=data,
+          xticklabel style={rotate=45},          
+          nodes near coords,
+          every node near coord/.append style={font=\scriptsize},        
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \vspace{-2ex}
+  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
+\end{figure}
+
 \section{Evaluation}

### 2020-01-08 paper/paper.tex
pass on step 2

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf,anonymous]{acmart}
+\documentclass[sigconf,anonymous,pre-print]{acmart}
+
+\pagenumbering{arabic} % Remove for camera-ready
 
@@
 
+\vspace{1ex}
 \noindent
@@
 
+\begin{figure*}[ht!]
+  \centering
+  \includegraphics[trim=0 350 100 0,clip,width=0.65\textwidth]{figs/nids-workflow}
+  \caption{\tname\ workflow.}
+  \label{fig:overview}
+  \vspace{-1ex}
+\end{figure*}
+
 Figure 4 shows a Suricata rule that captures a reproduction of a Ping
@@
 malicious traffic.
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[trim=0 350 100 0,clip,width=0.65\textwidth]{figs/nids-workflow}
-\caption{\tname\ workflow.}
-\label{fig:overview}
-\end{figure*}
 
@@
 
-\Mar{-------------------------- estou aqui}
-
 The rule reported in the first step satisfies the criterion to only
-capture positive traffic. Unfortunately, the rule produced in the
-first step is overconstrained---it is unable to capture different
-manifestations of the same attack as the rule matches all options that
-\tname\ found for the input message.
+capture positive traffic. Unfortunately, that rule is overconstrained.
+The rule can include options that are irrelevant for the input
+message. Consequently, the \nids\ may be unable to capture a different
+manifestation of the same attack.
 
@@
 
-\tname{} leverages public datasets\cite{tcpreplay} of negative traffic
+\tname{} leverages public datasets~\cite{tcpreplay} of benign traffic
 to guide the search for more general rules, hence addressing the
-overffiting issue. The goal of this step is to create rules that
-captures traffic similar to the input malicious traffic. To achieve
-that goal, \tname{} \emph{minimizes} the rules produced in the
-previous step, preserving the invariant of \emph{only} capturing the
-positive malicious traffic. A rule is minimal if discarding any option
-results in capturing negative (\ie{}, benign) traffic.  The
-minimization procedure we propose in this step builds on the
-observation that all options from the rule produced in the previous
-step are satisifed by the positive traffic. Consequently, any subset
-of these options also captures the positive traffic. The procedure
-iteratively discards options from the candidate rule until it detects
-that the rule is minimal, \ie{}, it finds that no other option can be
-removed. The procedure uses negative traffic to guide the search.
+overffiting issue. To that end, \tname{} \emph{minimizes} the rules
+produced in the previous step, preserving the invariant of \emph{only}
+capturing the positive malicious traffic. We say that a rule is
+minimal if discarding any option results in capturing negative (\ie{},
+benign) traffic.  The minimization procedure we propose in this step
+builds on the observation that all options from the rule produced in
+the previous step are satisifed by the positive traffic. Consequently,
+any subset of these options also captures the positive traffic. The
+procedure iteratively discards options from the candidate rule until
+it detects that the rule is minimal, \ie{}, it finds that no other
+option can be removed. The procedure uses negative traffic to guide
+the search.
 
@@
 \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
-\algnewcommand\And{\textbf{and} }
+\algnewcommand\And{\textbf{and} }%
 \begin{algorithm}[t!]
-  \noindent\hspace{-31ex}
-  \textbf{INPUT:} overconstrained rule $or$ \\
-  \noindent\hspace{-30ex} 
-  \textbf{OUTPUT:} Set of minimal rules $\mathcal R$\\
+  \noindent\hspace{-25ex}
+  \textbf{INPUT:} (overconstrained) seed rule $sr$ \\
+  \noindent\hspace{-27.5ex} 
+  \textbf{OUTPUT:} Set of minimized rules $\mathcal O$\\
 %\vspace{-3ex}
@@
 \begin{algorithmic}[1]
-  \State $\mathcal R \gets \{\mathit{or}\}$\Comment{initial population
-    of candidate solutions contains only the input rule.}\label{line-init}
+  \State $\mathcal R \gets \{\mathit{sr}\}$\Comment{initial population
+    of candidate solutions contains only the input
+    rule.}\label{line-init}
+  \State $\mathcal O \gets \emptyset$
   \Do\label{start-do-while}
-      \State $\mathcal R' \gets \emptyset$ \Comment{starting new
-        iteration. new generation should be empty.}\label{init-rprime}
-      \ForEach {$ro \in \mathcal R $}\Comment{analyze each candidate individual/rule separately.}\label{start-foreach}
+      \State $\mathcal R' \gets \emptyset$ \Comment{new iteration. reset.}\label{init-rprime}
+      \ForEach {$ro \in \mathcal R $}\Comment{analyze each candidate rule.}\label{start-foreach}
            \For {$i \gets 1$ to $N$}\Comment{mutate each rule $N$ times.}\label{start-for}
@@
                \State $j\gets \mathit{rand}(0,\mathit{len}(r))$
-               \State $r[j]\gets 0$\Comment{discard option $j$ of rule
-                 $r$, a copy of the original rule $\mathit{ro}$}
-               \If{$\mathit{isnew(r)} \And \mathit{isfit}(r)$} 
-                   \State $\mathcal R' \cup\gets \{r\}$\Comment{$r$
-                     can lead to a plausible candidate solution.}\label{line-progress}
+               \State $r[j]\gets 0$\Comment{discard option $j$ in $r$}
+               \If{$\mathit{isNew(r)} \And \mathit{isPlausible}(r)$} 
+                   \State $\mathcal R' \gets \mathcal R' \cup \{r\}$\label{line-progress}
                \EndIf
@@
        \EndFor\label{end-foreach}
-       \State $\mathcal R \gets R'$\Comment{update new generation}
-  \doWhile{$\mathcal R'\not= \{\}$}\Comment{repeat until reaching a fixpoint}\label{end-do-while}
+       \State $\mathcal R \gets \mathcal R'$\Comment{update new generation}
+       \State $\mathcal O \gets \mathcal O \cup \mathcal
+       R$\Comment{remember all rules}
+  \doWhile{$\mathcal R\not= \{\}$}\Comment{repeat until reaching a fixpoint}\label{end-do-while}
 \end{algorithmic}
@@
 overconstrained rule produced by the reverse engineering step and
-produces a set $\mathcal R$ of rules that are minimal and only
-captures the positive traffic. Line~\ref{line-init} initializes the
-set $\mathcal{R}$ of candidate solutions with the input rule $or$. We
-encode a rule as a bitvector where each index in the vector represents
-one option; the value assigned at a given position indicates whether
-or not the option is present in the rule. Each iteration of the outer
-loop (lines~\ref{start-do-while}-\ref{end-do-while}) represents one
+produces a set $\mathcal R$ of minimized rules. Line~\ref{line-init}
+initializes the set $\mathcal{R}$ of candidate solutions with the seed
+rule $sr$. We encode a rule as a bitvector where each index in the
+vector represents one option; the value assigned at a given position
+indicates whether or not the option is present in the rule. Each
+iteration of the outer loop
+(lines~\ref{start-do-while}-\ref{end-do-while}) represents one
 generation of the population of individuals produced by the
@@
 the property that a candidate solution should only capture positive
-traffic. This means that execution has not reached
-line~\ref{line-progress} as to augment the set $\mathcal{R'}$ and that
-set will be empty at the end of one iteration of the outer
+traffic. This means that execution of one iteration has not reached
+the statement declared at line~\ref{line-progress}, which adds an
+element to the set $\mathcal{R'}$. Consequently, that set will be
+empty at the end of one iteration of the outer
 loop. Line~\ref{init-rprime} initializes the new generation of
-candidate rules. The loop from
-lines~\ref{start-foreach}--\ref{end-foreach} iterates through each
+candidate rules. The inner loop
+(lines~\ref{start-foreach}--\ref{end-foreach}) iterates through each
 candidate rule in $\mathcal R$. Note that these rules have been found
@@
 to $\mathit{isNew}$ checks if $r$ has not been previously generated
-through another path whereas the call to $\mathit{isfit}$ checks if
-the rule only captures positive traffic. Function $\mathit{copy}$
+through another path whereas the call to $\mathit{isPlausible}$ checks
+if the rule only captures positive traffic. Function $\mathit{copy}$
 creates a copy of the bitvector of a rule, function $\mathit{len}$
@@
 randomly chooses one index of the vector.
+
+Note that no rules are produced in the last iteration of the
+algorithm. Minimal rules are produced in the second-to-last
+iteration. In the other iterations, intermediate rules are
+produced. Conceptually, the search induces a search tree with the seed
+rule $sr$ as the root, minimal rules as child nodes, and intermediate
+rules as internal nodes. The edges of the tree denote the ancestor
+relation between two rules. Note also that the algorithm reports every
+rule in the tree instead of only minimal rules. The reason is that the
+correct rule may be non-minimal. In summary, this algorithm reports
+many rules that are plausible by using benign traffic to minimize an
+overconstrained rule that is focused on the malicious traffic. In the
+following step, \tname{} ranks the rules this algorithm reports by
+their similarity with previous rules.
 
@@
   traffic was captured, a rule would not be added to $\mathcal
-  R'$.\qed
+  R'$.
 \end{proof}
@@
 
+\begin{proof}
 The output of the minimization algorithm may contain rules that are
-not minimal given the non-determinism of the algorithm. It is possible
-that options that could be discarded are not selected in the loop from
-lines~\ref{start-for}--\ref{end-for}. However, we found empirically
-that a small value of $N$ suffices to produce minimal rules.
+not minimal given the non-determinism of the algorithm. In principle,
+it is possible that options that could be discarded are not selected
+in the loop from lines~\ref{start-for}--\ref{end-for}. However, we
+found empirically that a small value of $N$ suffices to discard those
+options from a rule.
+\end{proof}
 
@@
 of the training data it uses. For the positive traffic, \tname{} only
-requires one sample, but can benefit of several samples. For the
-negative traffic, \tname{} can use public datasets available online.
+requires one sample, but can benefit of several samples as the example
+from Section~\ref{sec:dos} illustrated. The same applies for the
+negative traffic. In this case, however, \tname{} can use public
+datasets available online.
 
@@
 \noindent\textbf{Implementation details.} We implemented function
-\emph{isFit} by 1)~updating Suricata to only use the informed rule,
-2)~replaying the negative traffic (see
-Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
-check if any benign message was captured by the rule. It is worth
-noting that the benign traffic is stored in ``pcap'' files~\cite{pcap}
-and Suricata can process those files very efficiently compared to
-analyzing network messages directly.
+\emph{isPlausible} by 1)~updating Suricata to only use the rule that
+is passed as parameter to this function, 2)~replaying the negative
+traffic (see Section~\ref{sec:dataset-benign}), and 3)~reading the
+Suricata logs to check if any benign message was captured by the
+rule. It is worth noting that the benign traffic is stored in ``pcap''
+files~\cite{pcap} and Suricata can process those files very
+efficiently compared to analyzing network messages directly.
 
 \subsection{Step 3: Ranking}
+
+\Mar{------------------------------------ estou aqui}
 

### 2020-01-08 paper/paper.tex
adjusting style as per website recommendation

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf,anonymous,pre-print]{acmart}
+\documentclass[sigconf,review, anonymous]{acmart}
+\acmConference[ISSTA 2020]{ACM SIGSOFT International Symposium on Software Testing and Analysis}{18–22 July, 2020}{Los Angeles, US}
 
@@
 %% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
-\acmDOI{10.1145/1122445.1122456}
+%% \setcopyright{acmcopyright}
+%% \copyrightyear{2018}
+%% \acmYear{2018}
+%% \acmDOI{10.1145/1122445.1122456}
 
 %% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
-  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
-\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-  June 03--05, 2018, Woodstock, NY}
-\acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
+%% \acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
+%%   Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
+%% \acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+%%   June 03--05, 2018, Woodstock, NY}
+%% \acmPrice{15.00}
+%% \acmISBN{978-1-4503-XXXX-X/18/06}
 

### 2020-01-09 paper/paper.tex
section 3.3

@@
 
-\subsection{\Fix{Name of attack ***with content***}}
+\subsection{\Fix{JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit}}
 \label{sec:content-example}
+\Luc{JBoss é um servidor de aplicação Java muito popular, de codigo aberto.
+Esse ataque tem como alvo esses servidores e permite que seja feita execução remota de comandos.
+Os pacotes que estamos utilizando para testar esse ataque foram gerados pelo Nikto, mas o Metasploit tambem possui um modulo para ele.
+A regra de ouro para esse ataque é:
+
+\begin{figure}[H]
+  \lstinputlisting[language=C,numbers=none,keywords={content}]{adaptor-golden-rule.suricata}
+  \label{fig:rule-example}
+\end{figure}
+
+A opção flow:established,to\_server; define que o match só será feito em pacotes que fazem parte de uma conexão já estabelecida entre um cliente e um servidor.
+Além disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.
+Nos contents os modificadores nocase e http\_uri indicam que a detecção da string não será case sensitive e que essa string deverá se encontrar no campo de request uri.}
+
+\Luc{... sobre comportamento do syrius ...}
 

### 2020-01-09 paper/paper.tex
fix histogram

@@
 payload of the message and cannot be extracted directly from the
-fields. Handling this case is critically important as \Fix{xx}\% of
+fields. Handling this case is critically important as 93\% of
 the rules contain \CodeIn{content} options.
@@
             anchor=north,legend columns=-1},
-          symbolic x coords={1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
           xtick=data,
@@
         ]
-        \addplot coordinates {(1,10222) (2,4956) (3,5589)
-          (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
+        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
       \end{axis}
@@
 space. Rules without \CodeIn{content} options is an exception
-(\Fix{xx}\%) and half of the rules contain at least two
+(7\%) and half of the rules contain at least two
 \CodeIn{content} options.  To find the values associated with these
@@
 
-\begin{figure}[H]
-  \centering
-  \scalebox{0.85}{
-    \begin{tikzpicture}
-      \begin{axis}[
-          ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
-          %          ylabel={\#rules},
-          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
-          xtick=data,
-          xticklabel style={rotate=45},          
-          nodes near coords,
-          every node near coord/.append style={font=\scriptsize},        
-          nodes near coords align={vertical},
-        ]
-        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
-      \end{axis}
-    \end{tikzpicture}
-  }
-  \vspace{-2ex}
-  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
-\end{figure}
+%%\begin{figure}[H]
+%%  \centering
+%%  \scalebox{1.2}{
+%%    \begin{tikzpicture}
+%%      \begin{axis}[
+%%          ybar stacked,
+%%          enlargelimits=0.15,
+%%          legend style={at={(0.5,-0.15)},
+%%            anchor=north,legend columns=-1},
+%%          %          ylabel={\#rules},
+%%          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+%%          xtick=data,
+%%          xticklabel style={rotate=45},          
+%%          nodes near coords,
+%%          every node near coord/.append style={font=\scriptsize},        
+%%          nodes near coords align={vertical},
+%%        ]
+%%        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
+%%      \end{axis}
+%%    \end{tikzpicture}
+%%  }
+%%  \vspace{-2ex}
+%%  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
+%%\end{figure}
 

### 2020-01-10 paper/paper.tex
pass on 4.3

@@
 
+\Luc{Componente 1: Tamanho da regra.
+\begin{figure}[H]
+  \centering
+  \scalebox{1.2}{
+    \begin{tikzpicture}
+      \begin{axis}[
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          %          ylabel={\#rules},
+          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          xtick=data,
+          xticklabel style={rotate=45},          
+          nodes near coords,
+          every node near coord/.append style={font=\scriptsize},        
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(0,11) (1,1301) (2,4910) (3,3662) (4,3250) (5,5353) (6,2911) (7, 1337) (8, 1246) (9,763) (10, 495)};
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \vspace{-2ex}
+  \caption{\label{fig:distribution-rule-size}Number of rules per number of options.}
+\end{figure}
+
+Regras com um tamanho mais comum pontuam mais.
+Nesse caso, a pontuação dada por esse componente será: \[\frac{rules\_per\_size[len(rule)]}{max(rules\_per\_size)}\]
+
+Componente 2: Quantidade de contents na regra (histograma da figura 6)
+Regras com uma quantidade mais comum de contents pontuam mais.
+Assim como no caso anterior, a pontuação dada por esse componente será: \[\frac{rules\_per\_number\_of\_contents[len(rule)]}{max(rules\_per_number\_of\_contents)}\]
+
+Componente 3: Raridade dos contents \Fix{adicionar histograma}
+
+Contents mais comuns recebem maior pontuação.
+Nesse componente, para cada content da regra nós verificamos quão comum ele é. A pontuação é dada por:
+
+\[\frac{\sum_{i=1}^{N}\frac{content\_frequency[content_i]}{max(content\_frequency)}}{N}\]
+
+Onde 'N' é a quantidade de contents na regra.
+
+Componente 4: Frequencia dos modificadores de content
+
+Esse componente pontua um content de acordo com o campo ao qual ele pertence no pacote.
+Contents que pertencem a campos mais comuns (campos que são mais referenciados nas regras) recebem maior pontuação.
+Modificadores diferentes podem acessar campos iguais em um pacote.
+Por exemplo os modificadores http\_method e http\_header, o http\_method acessa apenas o metodo do pacote http, mas o http\_header acessa todo o cabeçalho (incluindo o proprio metodo).
+Com isso, a string "GET" receberia a pontuação correspondente a esses dois modificadores.
+\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{modifier\_frequency[modifier_{ij}]}{max(modifier\_frequency)}}{N}\]
+
+N é o numero de contents e M é a quantidade de modificadores de um content.
+
+\begin{figure}[H]
+  \centering
+  \scalebox{1.2}{
+    \begin{tikzpicture}
+      \begin{axis}[
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          %          ylabel={\#rules},
+          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection},
+          xtick=data,
+          xticklabel style={rotate=45},          
+          nodes near coords,
+          every node near coord/.append style={font=\scriptsize},        
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29)};
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \vspace{-2ex}
+  \caption{\label{fig:distribution-content_modifiers} Content modifiers frequency.}
+\end{figure}
+
+}
 \Fix{...elaborate...}
-
-%%\begin{figure}[H]
-%%  \centering
-%%  \scalebox{1.2}{
-%%    \begin{tikzpicture}
-%%      \begin{axis}[
-%%          ybar stacked,
-%%          enlargelimits=0.15,
-%%          legend style={at={(0.5,-0.15)},
-%%            anchor=north,legend columns=-1},
-%%          %          ylabel={\#rules},
-%%          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
-%%          xtick=data,
-%%          xticklabel style={rotate=45},          
-%%          nodes near coords,
-%%          every node near coord/.append style={font=\scriptsize},        
-%%          nodes near coords align={vertical},
-%%        ]
-%%        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
-%%      \end{axis}
-%%    \end{tikzpicture}
-%%  }
-%%  \vspace{-2ex}
-%%  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
-%%\end{figure}
 

### 2020-01-11 paper/paper.tex
pass

@@
 \label{sec:content-example}
+
+\Mar{Note que a gente organizou estas secoes por tipo de ataque. Este
+  titulo eh muito especifico.}
 \Luc{JBoss é um servidor de aplicação Java muito popular, de codigo aberto.
@@
   original rule and all those options are satisfied by the positive
-  traffic. By definition of function $\mathit{isfit}$, these rules
+  traffic. By definition of function $\mathit{isPlausible}$, these rules
   also capture \emph{only} the positive traffic. If some negative
@@
 
-\Mar{------------------------------------ estou aqui}
-
-\Luc{Componente 1: Tamanho da regra.
+\tname{} uses a composition of heurisitc functions to sort the rules
+it produces. Each function uses a different criterion to estimate how
+similar a rule created in the previous step is to existing Suricata
+rules~\cite{emerging-threats-open}. We describe in the following each
+of these heuristic functions. Each function $f_i$ ranges in the 0-1
+domain. \tname{} uses the weighted average of these functions to sort
+rules: $\sum{f_i}/|f_i|$.
+
+\subsubsection{Size of a rule}
+\tname{} uses the size of existing rules as a heuristic function. For
+example, we noted that most rules in the dataset have 2 to 5 options,
+so \tname{} values rules in this size range higher compared to other
+sizes. Figure~\ref{fig:distribution-rule-size} shows the histogram of
+rule sizes. The x-axis shows the number of options (\ie{}, size)
+included in a rule and the y-axis shows the number of rules that have
+that number of options. \tname{} uses this information to rank rules.
+
 \begin{figure}[H]
@@
           ybar stacked,
-          enlargelimits=0.15,
+          enlargelimits=0.09,
           legend style={at={(0.5,-0.15)},
@@
   \vspace{-2ex}
-  \caption{\label{fig:distribution-rule-size}Number of rules per number of options.}
+  \caption{\label{fig:distribution-rule-size}Distribution of rule
+    size.\Mar{how is it possible a rule with 0 options?}}
 \end{figure}
+
+\Mar{----------------------------------------- estou aqui}
 
@@
 
-}
-\Fix{...elaborate...}
-
 \section{Evaluation}

### 2020-01-11 paper/paper.tex
pass on step 3

@@
 it produces. Each function uses a different criterion to estimate how
-similar a rule created in the previous step is to existing Suricata
+similar a rule, created in the previous step, is to existing Suricata
 rules~\cite{emerging-threats-open}. We describe in the following each
@@
 \subsubsection{Size of a rule}
-\tname{} uses the size of existing rules as a heuristic function. For
-example, we noted that most rules in the dataset have 2 to 5 options,
-so \tname{} values rules in this size range higher compared to other
-sizes. Figure~\ref{fig:distribution-rule-size} shows the histogram of
-rule sizes. The x-axis shows the number of options (\ie{}, size)
-included in a rule and the y-axis shows the number of rules that have
-that number of options. \tname{} uses this information to rank rules.
+\tname{} uses the size of existing rules as one element to measure
+similarity. Informally, rules with popular sizes score high; inpopular
+rules score low.
+
+Figure~\ref{fig:distribution-rule-size} shows the histogram of rule
+sizes for our rule dataset, including a total of
+\numrulessuri\ rules. The x-axis shows the number of options included
+in a rule and the y-axis shows the number of rules with the
+corresponding number of options. For example, we noted that most rules
+have 2 to 6 options. Based on that, \tname{} values rules in that size
+range higher compared to other sizes.
+
+%\tname{} uses this information to rank rules.
 
@@
 
-\Mar{----------------------------------------- estou aqui}
-
-Regras com um tamanho mais comum pontuam mais.
-Nesse caso, a pontuação dada por esse componente será: \[\frac{rules\_per\_size[len(rule)]}{max(rules\_per\_size)}\]
+More precisely, we calculate the contribution of size of a given rule
+$r$ with the following formula:
+
+\[\frac{\mathit{rules\_per\_size}[\mathit{len(r)}]}{\mathit{num\_rules}}\]
+
+\noindent
+, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
+$\mathit{rules\_per\_size}$ encodes the size distribution--it is a
+dictionary mapping the number of rules associated with a given size,
+and $\mathit{num\_rules}$ returns the number of elements in the
+dictionary. Consequently, the contribution of size is proportion;al to
+its popularity. For example, the contribution of this heuristic for a
+rule with 3 options would be 0.145 (=3,662/25,239) whereas the
+contribution of a rule with 9 options would be 0.03 (=763/25,239).
+
+\Mar{--------------------------- estou aqui}
 

### 2020-01-11 paper/paper.tex
minor fixes to 4.3.1

--- 2020-01-11 paper/paper.tex
@@
 it produces. Each function uses a different criterion to estimate how
-similar a rule created in the previous step is to existing Suricata
-rules~\cite{emerging-threats-open}. We describe in the following each
-of these heuristic functions. Each function $f_i$ ranges in the 0-1
+similar a rule, created in the previous step, is to rules from a
+public database~\cite{emerging-threats-open} including a total of
+\numrulessuri\ rules. We describe in the following each of these
+heuristic functions. Each function $f_i$ ranges in the 0-1
 domain. \tname{} uses the weighted average of these functions to sort
@@
 \subsubsection{Size of a rule}
-\tname{} uses the size of existing rules as a heuristic function. For
-example, we noted that most rules in the dataset have 2 to 5 options,
-so \tname{} values rules in this size range higher compared to other
-sizes. Figure~\ref{fig:distribution-rule-size} shows the histogram of
-rule sizes. The x-axis shows the number of options (\ie{}, size)
-included in a rule and the y-axis shows the number of rules that have
-that number of options. \tname{} uses this information to rank rules.
+One element that \tname{} uses to measure similarity of synthesized to
+human-created rules is the size of the synthetic rule. Informally,
+rules with popular sizes score high whereas inpopular rules score low.
+
+Figure~\ref{fig:distribution-rule-size} shows the histogram of rule
+sizes for our rule dataset. The x-axis shows the number of options
+included in a rule (up to 10) while the y-axis shows the number of
+rules with the corresponding number of options. Note that there are 11
+rules with no options. These rules only use the header to signal a
+traffic as malicious. Also note that most rules have 2 to 6
+options. \tname{} values rules in that size range higher compared to
+other sizes.
+
+%\tname{} uses this information to rank rules.
 
@@
   \caption{\label{fig:distribution-rule-size}Distribution of rule
-    size.\Mar{how is it possible a rule with 0 options?}}
+    size.}
 \end{figure}
 
-\Mar{----------------------------------------- estou aqui}
-
-Regras com um tamanho mais comum pontuam mais.
-Nesse caso, a pontuação dada por esse componente será: \[\frac{rules\_per\_size[len(rule)]}{max(rules\_per\_size)}\]
+More precisely, we calculate the contribution of size of a given rule
+$r$ with the following formula:
+
+\[\frac{\mathit{rules\_per\_size}[\mathit{len(r)}]}{\mathit{max(rules\_per\_size})}\]
+
+\noindent
+, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
+$\mathit{rules\_per\_size}$ encodes distribution of sizes as a
+dictionary mapping sizes to number of associated rules, and
+$\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
+for any dictionary entry. Consequently, the contribution of size is
+proportional to its popularity. For example, the contribution of this
+heuristic for a rule with 3 options would be 0.684 (=3,662/5,353)
+whereas the contribution of a rule with 9 options would be 14.25
+(=763/5,353). Note that the measurement is normalized in the 0-1
+range.
+
+\Mar{--------------------------- estou aqui}
 

### 2020-01-11 paper/paper.tex
pass

@@
 
-\pgfplotsset{width=6cm,compat=1.8}
-\pgfplotsset{every tick label/.append style={font=\tiny}}
-\begin{figure}[t!]
-  \centering
-  \scalebox{1.2}{  
-    \begin{tikzpicture}
-      \begin{axis}[
-          bar width=6pt,
-          scaled ticks=false,
-          tick label style={/pgf/number format/fixed},
-          ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
-          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
-          xtick=data,
-          nodes near coords,
-          every node near coord/.append style={font=\tiny},
-          nodes near coords align={vertical},
-        ]
-        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
-      \end{axis}
-    \end{tikzpicture}
-  }
-  \caption{\label{fig:distribution-contents}Histogram of number of
-    \CodeIn{content} options per rule.}
-\end{figure}
-
 Figure~\ref{fig:distribution-contents} shows the histogram of the
@@
 
+\pgfplotsset{width=6cm,compat=1.8}
+\pgfplotsset{every tick label/.append style={font=\tiny}}
+
 \begin{figure}[H]
@@
       \begin{axis}[
+          bar width=6pt,
+          scaled ticks=false,
+          tick label style={/pgf/number format/fixed},
           ybar stacked,
-          enlargelimits=0.09,
+          enlargelimits=0.15,
           legend style={at={(0.5,-0.15)},
             anchor=north,legend columns=-1},
-          %          ylabel={\#rules},
           symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
           xtick=data,
-          xticklabel style={rotate=45},          
           nodes near coords,
-          every node near coord/.append style={font=\scriptsize},        
+          every node near coord/.append style={font=\tiny},
           nodes near coords align={vertical},
@@
   \vspace{-2ex}
-  \caption{\label{fig:distribution-rule-size}Distribution of rule
-    size.}
+  \caption{\label{fig:distribution-rule-size}Distribution of rule sizes.}
 \end{figure}
@@
 
+\subsubsection{Number of content options.} We empirically observed that content options are very
+common. As consequence, we defined a heuristic function that takes the
+number of content options into account. We calculate the contribution
+of number of content options with the following formula:
+
+\begin{figure}[t!]
+  \centering
+  \scalebox{1.2}{  
+    \begin{tikzpicture}
+      \begin{axis}[
+          bar width=6pt,
+          scaled ticks=false,
+          tick label style={/pgf/number format/fixed},
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          xtick=data,
+          nodes near coords,
+          every node near coord/.append style={font=\tiny},
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \caption{\label{fig:distribution-contents}Distribution of number of
+    content options.}
+\end{figure}
+
+\[\frac{\mathit{rules\_per\_number\_of\_contents[len(rule)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
+
+Note that the only difference compared to the formula to compute
+contribution of size is the distribution used, which is given by the
+term $\mathit{rules\_per\_number\_of\_contents}$ denoting a dictionary
+of number of content options to number of rules that contain that
+number of content options.
+
 \Mar{--------------------------- estou aqui}
-
-Componente 2: Quantidade de contents na regra (histograma da figura 6)
-Regras com uma quantidade mais comum de contents pontuam mais.
-Assim como no caso anterior, a pontuação dada por esse componente será: \[\frac{rules\_per\_number\_of\_contents[len(rule)]}{max(rules\_per_number\_of\_contents)}\]
 

### 2020-01-11 paper/paper.tex
pass on 4.3.3

@@
 
-\Mar{--------------------------- estou aqui}
-
-Componente 3: Raridade dos contents \Fix{adicionar histograma}
-
-Contents mais comuns recebem maior pontuação.
-Nesse componente, para cada content da regra nós verificamos quão comum ele é. A pontuação é dada por:
-
-\[\frac{\sum_{i=1}^{N}\frac{content\_frequency[content_i]}{max(content\_frequency)}}{N}\]
-
-Onde 'N' é a quantidade de contents na regra.
-
-Componente 4: Frequencia dos modificadores de content
+\subsubsection{Content rarity.} This heuristic weighs higher rules
+that use popular strings in content options. More precisely, for a
+given rule \tname{} takes the arithmetic mean of the popularity scores
+of each content in that rule. More precisely, we calculate the
+contribution of content rarity with the following formula:
+
+\[\frac{\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}}{N}\]
+
+\noindent
+, where $N$ is the number of content options in a given rule. Figure~\Fix{...elaborate...}
+
+\Mar{----------------------------- parei aqui}
+
+\subsubsection{Frequency of content modifiers.}
 

### 2020-01-11 paper/paper.tex
pass on 4.3.3

--- 2020-01-11 paper/paper.tex
@@
 Componente 3: Raridade dos contents \Fix{adicionar histograma}
+\begin{figure}[H]
+  \centering
+  \scalebox{1.2}{  
+    \begin{tikzpicture}
+      \begin{axis}[
+          bar width=6pt,
+          scaled ticks=false,
+          tick label style={/pgf/number format/fixed},
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},
+          xtick=data,
+          xticklabel style={rotate=45},          
+          nodes near coords,
+          every node near coord/.append style={font=\tiny},
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387) (INSERT, 386) (200, 352) (/index.php?, 273)};
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \caption{\label{fig:distribution-contents}-}
+\end{figure}
 

### 2020-01-11 paper/paper.tex

--- 2020-01-11 paper/paper.tex
@@
 
-\subsubsection{Content rarity.} This heuristic weighs higher rules
+\subsubsection{Content rarity.} 
+This heuristic weighs higher rules
 that use popular strings in content options. More precisely, for a
@@
 \Mar{----------------------------- parei aqui}
+
+\begin{figure}[H]
+  \centering
+  \scalebox{1.2}{  
+    \begin{tikzpicture}
+      \begin{axis}[
+          bar width=6pt,
+          scaled ticks=false,
+          tick label style={/pgf/number format/fixed},
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},
+          xtick=data,
+          xticklabel style={rotate=45},          
+          nodes near coords,
+          every node near coord/.append style={font=\tiny},
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387) (INSERT, 386) (200, 352) (/index.php?, 273)};
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \caption{\label{fig:distribution-contents}-}
+\end{figure}
 

### 2020-01-12 paper/paper.tex
pass on 4.3

@@
 public database~\cite{emerging-threats-open} including a total of
-\numrulessuri\ rules. We describe in the following each of these
+\numrulessuri\ rules \Luc{acima na seção 4.1 falamos que o dataset contem 27.8K regras.
+No histograma da fig 7 de fato mostra apenas 25K, mas isso porque decidi colocar nele apenas 11 barras}.
+We describe in the following each of these
 heuristic functions. Each function $f_i$ ranges in the 0-1
-domain. \tname{} uses the weighted average of these functions to sort
+domain.
+\tname{} uses the weighted average \Luc{atualmente eh uma média simples e não ponderada ja q as funções tem o mesmo peso.
+conversamos na ultima reuniao sobre testar diferentes pesos para cada função, mas n sei quão prioritário eh isso comparado as outras atividades que definimos.} of these functions to sort
 rules: $\sum{f_i}/|f_i|$.
@@
 rules with popular sizes score high whereas inpopular rules score low.
-
+\Luc{talvez explicitar que "size of a rule" quer dizer a quantidade de opções que ha nela?}
 Figure~\ref{fig:distribution-rule-size} shows the histogram of rule
@@
 heuristic for a rule with 3 options would be 0.684 (=3,662/5,353)
-whereas the contribution of a rule with 9 options would be 14.25
-(=763/5,353). Note that the measurement is normalized in the 0-1
+whereas the contribution of a rule with 9 options would be 0.028
+(=154/5,353). Note that the measurement is normalized in the 0-1
 range.
@@
 \subsubsection{Number of content options.} We empirically observed that content options are very
-common. As consequence, we defined a heuristic function that takes the
+common. \Luc{Se for preciso temos numeros para fundamentar isso melhor, sao cerca de 70K contents nas 28K regras do dataset} As consequence, we defined a heuristic function that takes the
 number of content options into account. We calculate the contribution
@@
 
-\[\frac{\mathit{rules\_per\_number\_of\_contents[len(rule)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
+\[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
 
@@
 , where $N$ is the number of content options in a given rule. Figure~\Fix{...elaborate...}
+
+\Luc{tudo ok ate aqui}
 

### 2020-01-12 paper/paper.tex
pass

--- 2020-01-11 paper/paper.tex
@@
 \numrulessuri\ rules. We describe in the following each of these
-heuristic functions. Each function $f_i$ ranges in the 0-1
-domain. \tname{} uses the weighted average of these functions to sort
-rules: $\sum{f_i}/|f_i|$.
+heuristic functions. The codomain of each function ranges in the 0-1
+interval. The score of a rule is the average across the values
+produced by each function.
+
+%% \Luc{atualmente eh uma
+%%   média simples e não ponderada ja q as funções tem o mesmo
+%%   peso. conversamos na ultima reuniao sobre testar diferentes pesos
+%%   para cada função, mas n sei quão prioritário eh isso comparado as
+%%   outras atividades que definimos.} $\sum{f_i}/|f_i|$.
 
 \subsubsection{Size of a rule}
-One element that \tname{} uses to measure similarity of synthesized to
-human-created rules is the size of the synthetic rule. Informally,
-rules with popular sizes score high whereas inpopular rules score low.
-
-Figure~\ref{fig:distribution-rule-size} shows the histogram of rule
-sizes for our rule dataset. The x-axis shows the number of options
-included in a rule (up to 10) while the y-axis shows the number of
-rules with the corresponding number of options. Note that there are 11
-rules with no options. These rules only use the header to signal a
-traffic as malicious. Also note that most rules have 2 to 6
+One element that \tname{} uses to measure similarity between a
+synthesized and a human-created rule is the size of the synthetic
+rule, \ie{}, the number of options it contains. Informally, rules with
+popular sizes score high whereas inpopular rules score
+low. Figure~\ref{fig:distribution-rule-size} shows the histogram of
+rule sizes for our rule dataset. The x-axis shows the number of
+options included in a rule (up to 10) while the y-axis shows the
+number of rules with the corresponding number of options. Note that
+there are 11 rules with no options. These rules only use the header to
+signal a traffic as malicious. Also note that most rules have 2 to 6
 options. \tname{} values rules in that size range higher compared to
@@
 proportional to its popularity. For example, the contribution of this
-heuristic for a rule with 3 options would be 0.684 (=3,662/5,353)
-whereas the contribution of a rule with 9 options would be 14.25
-(=763/5,353). Note that the measurement is normalized in the 0-1
+heuristic for a rule with three options would be 0.684 (=3,662/5,353)
+whereas the contribution of a rule with nine options would be 0.028
+(=154/5,353). Note that the measurement is normalized in the 0-1
 range.
@@
 \subsubsection{Number of content options.} We empirically observed that content options are very
-common. As consequence, we defined a heuristic function that takes the
-number of content options into account. We calculate the contribution
-of number of content options with the following formula:
+common. \Luc{Se for preciso temos numeros para fundamentar isso
+  melhor, sao cerca de 70K contents nas 28K regras do
+  dataset}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
+  (\%) das 28K regras tem contents.} As consequence, we defined a
+heuristic function that takes the number of content options into
+account. We calculate the contribution of number of content options
+with the following formula:
 
@@
 
-\[\frac{\mathit{rules\_per\_number\_of\_contents[len(rule)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
-
-Note that the only difference compared to the formula to compute
-contribution of size is the distribution used, which is given by the
-term $\mathit{rules\_per\_number\_of\_contents}$ denoting a dictionary
-of number of content options to number of rules that contain that
-number of content options.
+\[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
+
+\noindent
+Note that the only difference of this formula compared to the formula
+to compute contribution of size is the distribution used, which is
+given by the term $\mathit{rules\_per\_number\_of\_contents}$ denoting
+a dictionary of number of content options to number of rules that
+contain that number of content options.
 
 \subsubsection{Content rarity.} 
-This heuristic weighs higher rules
-that use popular strings in content options. More precisely, for a
-given rule \tname{} takes the arithmetic mean of the popularity scores
-of each content in that rule. More precisely, we calculate the
-contribution of content rarity with the following formula:
+This heuristic weighs higher rules that use popular strings in content
+options. For a given rule \tname{} takes the arithmetic mean of the
+popularity scores of each string used in contents in that rule. More
+precisely, we calculate the contribution of content rarity with the
+following formula:
 
@@
 \noindent
-, where $N$ is the number of content options in a given rule. Figure~\Fix{...elaborate...}
-
-\Mar{----------------------------- parei aqui}
+, where the term $\mathit{content_i}$ denotes the $i$-th xcontent in a
+given rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
+denotes the frequency of the string used in the $i$-th content, the
+term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
+observed across all strings appearing as content parameters, and $N$
+is the number of content options in a given
+rule. Figure~\ref{fig:distribution-strings} shows the distribution of
+the ten most popular strings in our rule dataset. To illustrate this
+function, let us consider a rule with two content options, using the
+strings ``GET'' and ``ID'', respectively. The contribution of that
+rule for the computation of similarity score would be 0.601
+(=(1+.203)/2).
 
@@
   }
-  \caption{\label{fig:distribution-contents}-}
+  \caption{\label{fig:distribution-strings}Distribution of strings
+    used as parameters of content options.}
 \end{figure}
@@
 \subsubsection{Frequency of content modifiers.}
+
+\Mar{----------------------------- parei aqui}
 

### 2020-01-12 simulated-annealing/sa.py
training-test for sa.py

@@ -7,6 +7,7 @@ import random
 import re
 import binascii
 import argparse
+import csv
 from itertools import combinations
 
 open("bad.rules", 'w').close()
@@ -487,7 +488,7 @@ def isEmpty(fpath):
 
 def sendAttackVariation(attack):
     subprocess.Popen(["sh", "sendAttackVariation.sh", attack], stdout=subprocess.DEVNULL).wait()
-    time.sleep(2.0)
+    time.sleep(0.5)
 
 def checkFalseNegative(rules):
     global fitnessFile_path
@@ -507,8 +508,39 @@ def checkFalseNegative(rules):
         fitnessFile = fitnessFile.read()
     
     for i in range(len(rules)):
-        if fitnessFile.count(':'+str(rules[i].sid)+':') >= 1:
-            output[i] = fitnessFile.count(':'+str(rules[i].sid)+':')
+        if fitnessFile.count('1:'+str(rules[i].sid)+':') >= 1:
+            output[i] = fitnessFile.count('1:'+str(rules[i].sid)+':')
+        #else:
+            #print("fitness file count", ':'+str(rules[i].sid)+':', fitnessFile.count(str(rules[i].sid)))
+            #print("False negative rule:", rules[i])
+            #print(fitnessFile)
+
+    return output
+
+def sendTest(attack):
+    subprocess.Popen(["sh", "sendTest.sh", attack], stdout=subprocess.DEVNULL).wait()
+    time.sleep(0.5)
+
+def checkPrecision(rules):
+    global fitnessFile_path
+    global args
+    variation_packets = 2000
+
+    open(fitnessFile_path, 'w').close()
+    writeRuleOnFile(rules)
+    sendTest(args.attack)
+    
+    output = []
+
+    for i in range(len(rules)):
+        output.append(0)
+
+    with open(fitnessFile_path, "r") as fitnessFile:
+        fitnessFile = fitnessFile.read()
+    
+    for i in range(len(rules)):
+        if fitnessFile.count('1:'+str(rules[i].sid)+':') >= 1:
+            output[i] = fitnessFile.count('1:'+str(rules[i].sid)+':')
         #else:
             #print("fitness file count", ':'+str(rules[i].sid)+':', fitnessFile.count(str(rules[i].sid)))
             #print("False negative rule:", rules[i])
@@ -636,6 +668,7 @@ def newRuleFitness(rule):
     global max_fit2
     global max_fit3
     global max_fit4
+
[…]

### 2020-01-12 simulated-annealing/sendTest.sh

@@ -1,0 +1,1 @@
+suricata -c inputs/suricata-$1.yaml -r Datasets/test.pcap

### 2020-01-12 paper/paper.tex
pass on methodology

@@
 
+\Gui{
+O procedimento que já tínhamos representa a etapa de treinamento.
+Toda aquela parte que foi explicada na apresentação para os outros professores e descrito no paper.
+Descreverei novamente pra confirmar, e também foi adicionado uma pequena etapa.
+
+TREINAMENTO
+
+->Syrius recebe como input um pcap contendo o ataque (no caso de ataques com content, usamos ataques do Nikto), um pcap benígno (usamos o bigflows.pcap) e um pcap contendo as variações de ataque (isso pode ser otimizado unificando com o pcap de ataque, mas por agora está assim)
+->É gerado a regra máxima através do tokenizer e engenharia reversa.
+->Syrius remove UM content de uma regra da lista de regras atuais (inicialmente apenas a regra máxima) e checa se a regra criada já existe na lista.
+Se não, a regra é adicionada à lista de regras-não-testadas.
+Syrius faz isso duas vezes por regra, para todas as regras da lista.
+->A lista de regras-não-testadas é inicializada no suricata e então o rodamos com o pcap benígno de treinamento (75porcento do total).
+Verificando no arquivo de log do suricata, caso a regra alerte, ela é removida da lista.
+As regras restantes vão para a lista de regras atuais.
+->Quando não houver alterações, retornamos a lista com todas as regras que deram certo.
+->Opcionalmente, ainda rodamos no final uma variação de ataque (no momento apenas um pacote devido ao número limitado que temos do ataque).
+Ao contrário do método com pacotes benígnos, se a regra NÃO alertar, ela é removida.
+->Calculamos o fitness de cada regra e ordenamos de acordo com o fitness resultante. Esta lista ordenada é o fim do treinamento.
+
+TESTE
+
+->Inicializamos a lista de regras ordenadas no suricata e rodamos o pcap benígno de teste (25porcento restantes) para cálculo do precision.
+O cálculo ficou basicamente, por regra: [100 - (numero-de-alertas/total-de-pacotes)]
+
+->Inicializamos a lista de regras ordenadas no suricata e rodamos o pcap com variações de ataque (caso a etapa opcional ocorrer, não usamos aquele pacote especifico) para cálculo do recall.
+O cálculo ficou basicamente, por regra: [(100/total-de-pacotes) * n-alertas].
+
+->Por final calculamos o F1 Score por regra: [2*((precision*recall)/(precision+recall))]
+
+->O teste retorna um csv contendo a lista de regras ordenadas com seus respectivos precision, recall e f1 score.
+}
+
 %% \Gui{We are using the bigflows.pcap provided by tcpreplay on http://tcpreplay.appneta.com/wiki/captures.html "This is a capture of real network traffic on a busy private network’s access point to the Internet"

### 2020-01-13 paper/paper.tex
minor

@@
 
-\subsubsection{Frequency of content modifiers.}
-
-\Mar{----------------------------- parei aqui}
-
-Esse componente pontua um content de acordo com o campo ao qual ele pertence no pacote.
+\subsubsection{Frequency of content modifiers.} Content options are
+associated with specific packets fields\Mar{or modifiers?}.\Mar{why
+  GET is associated with http\_method and http\_header?}\Mar{are all
+  content associated with one or many fields or positions in a packet?
+  can you give examples other than GET to explain this? if yes,
+  shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
+  is important to justify (to explain rationale)---how is it different
+  from counting rarity?}
+
+%% Esse componente pontua um content de acordo com o campo ao qual ele pertence no pacote.
 Contents que pertencem a campos mais comuns (campos que são mais referenciados nas regras) recebem maior pontuação.
-Modificadores diferentes podem acessar campos iguais em um pacote.
-Por exemplo os modificadores http\_method e http\_header, o http\_method acessa apenas o metodo do pacote http, mas o http\_header acessa todo o cabeçalho (incluindo o proprio metodo).
-Com isso, a string "GET" receberia a pontuação correspondente a esses dois modificadores.
+%% Modificadores diferentes podem acessar campos iguais em um pacote. Por
+%% exemplo os modificadores http\_method e http\_header, o http\_method
+%% acessa apenas o metodo do pacote http, mas o http\_header acessa todo
+%% o cabeçalho (incluindo o proprio metodo). Com isso, a string "GET"
+%% receberia a pontuação correspondente a esses dois modificadores.
+
+More precisely, we calculate the contribution of the frequency of
+content modifiers with the following formula:
+
 \[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{modifier\_frequency[modifier_{ij}]}{max(modifier\_frequency)}}{N}\]
 
-N é o numero de contents e M é a quantidade de modificadores de um content.
+\noindent
+, where $N$ is the number of contents in a given rule and $M$
+is the number of content modifiers\Mar{in a rule? what is a content
+  modifier in a rule?}.
 
@@
   \vspace{-2ex}
-  \caption{\label{fig:distribution-content_modifiers} Content modifiers frequency.}
+  \caption{\label{fig:distribution-content_modifiers}Content modifiers
+    frequency.\Mar{This is awkward. This is the only histogram that
+      only shows 6 values in the x-axis. Aren't there more?}}
 \end{figure}

### 2020-01-13 paper/paper.tex
quick pass on evaluation

@@
 \vspace{2ex}
-\noindent\textbf{Limitation.} It is worth noting that although
+\noindent\textbf{Limitations.} It is worth noting that although
 \tname{} is accurate for generating rules that discrimate positive and
@@
 \subsection{Objects of Analysis}
-
-Table~\ref{table:attacks} shows the attacks we analyzed and the
-options of the rule from the Emerging Threats Open
-Ruleset~\cite{emerging-threats-open} that captures the attack. Column
-``Name'' shows the name of the attack, column ``Protocol'' shows the
-protocol name, and column ``Ground Truth'' shows the options of the
-rule that captures the attack.
+\label{sec:dataset-benign}
+
+Table~\ref{table:attacks} shows the attacks we considered in our
+evaluation. These attacks cover different protocols and features of
+\tname{}. Column ``Name'' shows the name of the attack, column
+``Protocol'' shows the protocol name, and column ``Ground Truth''
+shows the options of the rule that captures the attack.
+
+\vspace{1ex}
+\noindent
+\textbf{Malicious traffic.}~Nikito?\Fix{where do these
+  attacks/messages come from?}
+
+\vspace{1ex}
+\noindent
+\textbf{Benign traffic.}~\tname{} uses public databases storing benign
+traffic for all protocols covered by Suricata (see
+Figure~\ref{fig:distribution-rules-protocol}). We used two datasets of
+positive traffic. The Bigflows.pcap dataset is part of the
+TcpReplay~\cite{tcpreplay} open source project. This dataset includes
+real network traffic on a busy private network access point to the
+Internet. According to the TcpReplay web site ``This capture is much
+larger and has a smaller average packet size than the previous capture
+(smallFlows.pcap). It also has many more flows and different
+applications.''. In addition to the Bigflows.pcap dataset, we also
+used datasets of normal traffic made available by Stratosphere
+Lab~\cite{stratosphere-normal}.
+
+\vspace{1ex}
+\noindent
+\textbf{Ground Truth.}~The rules that detect those attacks were
+obtained from the Emerging Threats Open
+Ruleset~\cite{emerging-threats-open}.
+
+%% Note from Figure~\ref{fig:overview} that the positive traffic is an
+%% input to the technique. The negative non-malicious traffic, however,
+%% is not.
 
@@
 
-\subsection{Dataset of Benign Traffic}
-\label{sec:dataset-benign}
-
-Note from Figure~\ref{fig:overview} that the positive traffic is an
-input to the technique. The negative non-malicious traffic, however,
-is not. \tname{} uses public databases storing different kinds of
-positive traffic for all protocols covered by Suricata. We used two
-datasets of positive traffic. The Bigflows.pcap dataset is part of the
-TcpReplay~\cite{tcpreplay} open source project. This dataset includes
-real network traffic on a busy private network access point to the
-Internet. According to the TcpReplay web site ``This capture is much
-larger and has a smaller average packet size than the previous capture
-(smallFlows.pcap). It also has many more flows and different
-applications.''. In addition to the Bigflows.pcap dataset, we also
-used datasets of normal traffic made available by Stratosphere
-Lab~\cite{stratosphere-normal}.
-
 \subsection{Methodology}
 
-\tname{} expects (a) message(s) reproducing an attack on input. As
-previously discussed, those messages could be directly provided by an
-anomaly-based \nids\ tool. To evaluate \tname{}, however, we created
-those messages ourselves. For all attacks listed on
-Table~\ref{table:attacks}, we found reproduction scripts available on
-the web. With the script to reproduce the attack, we proceeded as
-follows: 1)~we instantiated the script on a given target machine,
-2)~we executed the script, 3)~we monitored the traffic with the
-wireshark network monitor~\cite{wireshark-net-monitor}, and 4) we
-simulated the delivery of the corresponding network messages by
-directly executing Suricata on the ``pcap'' files that wireshark
-produced in the previous step. It is worth noting that Suricata is
-invoked multiple times in the process of minimizing the rule (see
-Section~\ref{sec:minimization}) and that processing pcap files is much
-more efficient compared to sending and processing network messages. To
-illustrate, the SYN Flood attack was reproduced with the hping3
-tool~\cite{hping3} using the following command \CodeIn{hping3 -d 80 -w
-  64 -S -p 80 --flood --rand-source <IP>}.
+\subsubsection{Attack reproduction.} \tname{} requires the attack to
+be replayed in order to check whether or not the rule it generates
+captures the attack.  See function $\mathit{isPlausible}$, used in the
+minimization algorithm from Section~\ref{sec:minimization}. For all
+attacks listed on Table~\ref{table:attacks}, we found scripts
+available on the web to replay the attacks. For example, the SYN Flood
+attack can be replayed with hping3~\cite{hping3}, using the following
+command: \CodeIn{hping3 -d 80 -w 64 -S -p 80 --flood --rand-source
+  <IP>}. With the script to reproduce the attack, we executed the
+following steps to obtain ``pcap'' files, which abstracts network
+messages~\cite{pcap}, to run \tname:~1)~instantiate the script (\eg{},
+define the IP, for the example above), 2)~execute the script,
+3)~monitor the traffic with the wireshark network
+monitor~\cite{wireshark-net-monitor}.
+
+Recall that Suricata is invoked multiple times in the process of
+minimizing the rule (see Section~\ref{sec:minimization}). Suricata
+supports a feature to process ``pcap'' files, enabling very efficient
+performance compared to processing actual network messages.
 

### 2020-01-14 paper/paper.tex
4.3.4

@@
 
-\subsubsection{Frequency of content modifiers.} Content options are
+\subsubsection{Frequency of content modifiers.} 
+\Luc{\section{Modificadores}É possivel deixar que uma regra verifique todo o payload do pacote em busca de um match com um content ou especificar partes do payload onde esse content deve ser procurado.
+Essa definição é feita utilizando os chamados modificadores de content, eles adicionam mais detalhes/informações sobre a detecção de um determinado content.
+Os modificadores que estamos considerando são validos apenas para regras dos protocolos tcp e http.
+
+Exemplo do payload "cru" de um pacote http:\\
+Aqui, todas as informações do pacote estão juntas.
+
+\begin{figure}[h!]
+  \lstinputlisting[language=C,numbers=none,keywords={content}]{http-example.payload}
+\end{figure}
+
+O mesmo payload interpretado pelo Wireshark:\\
+Aqui, o wireshark interpreta as informações dos pacotes e as divide para melhor visualização e entendimento.
+
+\begin{figure}[H]
+  \lstinputlisting[language=C,numbers=none,keywords={content}]{http-example-wireshark.payload}
+\end{figure}
+
+O que o suricata faz eh identificar os campos e seus valores e com isso permite que as regras façam referencia direta a eles.
+Por exemplo, na figura do payload "cru" não conseguimos identificar facilmente oq eh o "Request URI Path:", já na segunda figura isso está explícito.
+
+Eu disse que o "GET" estava associado tambem ao http\_header, mas não está, me enganei.
+Mas sim, pode haver interseção entre os campos que dois ou mais modificadores referenciam.
+O seguinte trecho do payload é acessado pelo modificador http\_header:
+
+\begin{figure}[H]
+  \lstinputlisting[language=C,numbers=none,keywords={content}]{http-header.payload}
+\end{figure}
+
+Entretando, tambem existem os modificadores http\_host, http\_accept\_enc, http\_connection e http\_accept que permitem acessar diretamente cada uma das linhas acima.
+
+\subsection{Pontuação}
+A avaliação que fazemos aqui pontua um dado content de acordo com quantas vezes o campo em que ele se encontra eh referenciado nas regras.
+Por exemplo, vamos supor que o tokenizer produziu a opção content: "/HtmlAdaptor".
+Essa string pertence ao campo "Request URI:".
+Esse campo pode ser acessado pelo modificador http\_uri.
+O tokenizer associa o content a esse modificador.
+Posteriormente, esse content vai receber uma pontuação correspondente a popularidade desse modificador:\\
+\[\frac{10111}{10111*1}\]
+A multiplicação por 1 no denominador eh referente a quantidade de modificadores que estão associadas a esse content.
+No caso da opção content: "keep-alive", essa string pertence ao campo "Connection:".
+Como falado acima, o suricata pode acessar esse campo pelos modificadores http\_header ou http\_connection.
+Portanto, esse content será pontuado de acordo com a popularidade desses dois modificadores:
+\[\frac{4089+29}{10111*2}\]
+}
+
+Content options are
 associated with specific packets fields\Mar{or modifiers?}.\Mar{why
@@
   modifier in a rule?}.
+\Luc{M eh o numero de modificadores associados a cada content.
+Temos content[1], content[2], ..., content[N].
+Cada um desses contents tem modificadores associados content[i][1], content[i][2], ..., content[i][M].}
 

### 2020-01-14 paper/paper.tex
pass

--- 2020-01-13 paper/paper.tex
@@
 
-\subsubsection{Frequency of content modifiers.} Content options are
-associated with specific packets fields\Mar{or modifiers?}.\Mar{why
-  GET is associated with http\_method and http\_header?}\Mar{are all
-  content associated with one or many fields or positions in a packet?
-  can you give examples other than GET to explain this? if yes,
-  shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
-  is important to justify (to explain rationale)---how is it different
-  from counting rarity?}
-
-%% Esse componente pontua um content de acordo com o campo ao qual ele pertence no pacote.
-Contents que pertencem a campos mais comuns (campos que são mais referenciados nas regras) recebem maior pontuação.
-%% Modificadores diferentes podem acessar campos iguais em um pacote. Por
-%% exemplo os modificadores http\_method e http\_header, o http\_method
-%% acessa apenas o metodo do pacote http, mas o http\_header acessa todo
-%% o cabeçalho (incluindo o proprio metodo). Com isso, a string "GET"
-%% receberia a pontuação correspondente a esses dois modificadores.
-
-More precisely, we calculate the contribution of the frequency of
-content modifiers with the following formula:
-
-\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{modifier\_frequency[modifier_{ij}]}{max(modifier\_frequency)}}{N}\]
-
-\noindent
-, where $N$ is the number of contents in a given rule and $M$
-is the number of content modifiers\Mar{in a rule? what is a content
-  modifier in a rule?}.
+\subsubsection{Frequency of content modifiers.} As previously
+mentioned, \percRulesWithContent\ of Suricata rules involve content
+options. Furthermnore, we observed that many of these rules indicate
+what parts of the message the content string should be checked
+against.\Mar{Lucas, (1) vc. tem dados para dar suporte a isto (de http
+  e tcp)? ou seja, quantas options do tipo ``content'' possuem
+  modificadores.}  For example, the creator of the rule may want to
+indicate that a specific string be present in the header of an HTTP
+request message. Content \emph{modifiers} serve this purpose. For
+example, the modifier \CodeIn{http\_header} in the option
+\CodeIn{content: ``keep-alive''; http\_header} states the requirement that
+the string ``keep-alive'' be present in the header of the message. (A
+keep-alive connection attribute indicates that a single TCP connection
+should remain open for multiple HTTP requests/responses.)
+Figure~\Fix{adicione pequena figura...verifique se ainda faz sentido
+  depois do que falei acima...}
+
+Suricata supports content modifiers for tcp and http. Given that
+content options often use modifiers\Mar{assumindo que eh
+  verdade. confirmar estat. solicitada acima...}, \tname\ uses
+associated frequency to discriminate more or less likely rules. For
+example, historical data indicates that content options that use the
+\CodeIn{http\_connection} modifiers are much less likely than content
+options that use the modifier
+\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
+shows the distribution of content modifiers for http. We ommit the
+distribution for tcp for space. Similarly to the the previous
+heuristic functions, this function uses the frequency of content
+modifiers to calibrate the contribution of each rule.
 
@@
   \caption{\label{fig:distribution-content_modifiers}Content modifiers
-    frequency.\Mar{This is awkward. This is the only histogram that
-      only shows 6 values in the x-axis. Aren't there more?}}
+    frequency.\Mar{what about ``none'' (no modifier)?}}
 \end{figure}
+
+%% \section{Modificadores}É possivel deixar que uma regra verifique todo o payload do pacote em busca de um match com um content ou especificar partes do payload onde esse content deve ser procurado.
+Essa definição é feita utilizando os chamados modificadores de content, eles adicionam mais detalhes/informações sobre a detecção de um determinado content.
+Os modificadores que estamos considerando são validos apenas para regras dos protocolos tcp e http.
+
+%% Exemplo do payload "cru" de um pacote http:\\
+%% Aqui, todas as informações do pacote estão juntas.
+
+%% \begin{figure}[h!]
+%%   \lstinputlisting[language=C,numbers=none,keywords={content}]{http-example.payload}
+%% \end{figure}
+
+%% O mesmo payload interpretado pelo Wireshark:\\
+%% Aqui, o wireshark interpreta as informações dos pacotes e as divide para melhor visualização e entendimento.
+
+%% \begin{figure}[H]
+%%   \lstinputlisting[language=C,numbers=none,keywords={content}]{http-example-wireshark.payload}
+%% \end{figure}
+
+%% O que o suricata faz eh identificar os campos e seus valores e com isso permite que as regras façam referencia direta a eles.
+Por exemplo, na figura do payload "cru" não conseguimos identificar facilmente oq eh o "Request URI Path:", já na segunda figura isso está explícito.
+
+%% Eu disse que o "GET" estava associado tambem ao http\_header, mas não está, me enganei.
+Mas sim, pode haver interseção entre os campos que dois ou mais modificadores referenciam.
+O seguinte trecho do payload é acessado pelo modificador http\_header:
+
+%% \begin{figure}[H]
+%%   \lstinputlisting[language=C,numbers=none,keywords={content}]{http-header.payload}
+%% \end{figure}
+
+%% Entretando, tambem existem os modificadores http\_host,
+%% http\_accept\_enc, http\_connection e http\_accept que permitem
+%% acessar diretamente cada uma das linhas acima.
+
+\Mar{----------------------------------------------------------- estou
+aqui}
+
+\subsection{Pontuação}
+A avaliação que fazemos aqui pontua um dado content de acordo com quantas vezes o campo em que ele se encontra eh referenciado nas regras.
+Por exemplo, vamos supor que o tokenizer produziu a opção content: "/HtmlAdaptor".
+Essa string pertence ao campo "Request URI:".
+Esse campo pode ser acessado pelo modificador http\_uri.
+O tokenizer associa o content a esse modificador.
+Posteriormente, esse content vai receber uma pontuação correspondente a popularidade desse modificador:\\
+\[\frac{10111}{10111*1}\]
+A multiplicação por 1 no denominador eh referente a quantidade de modificadores que estão associadas a esse content.
+No caso da opção content: "keep-alive", essa string pertence ao campo "Connection:".
+Como falado acima, o suricata pode acessar esse campo pelos modificadores http\_header ou http\_connection.
+Portanto, esse content será pontuado de acordo com a popularidade desses dois modificadores:
+\[\frac{4089+29}{10111*2}\]
+
+Content options are
+associated with specific packets fields\Mar{or modifiers?}.\Mar{why
+  GET is associated with http\_method and http\_header?}\Mar{are all
+  content associated with one or many fields or positions in a packet?
+  can you give examples other than GET to explain this? if yes,
+  shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
+  is important to justify (to explain rationale)---how is it different
+  from counting rarity?}
+
+%% Esse componente pontua um content de acordo com o campo ao qual ele pertence no pacote.
+Contents que pertencem a campos mais comuns (campos que são mais referenciados nas regras) recebem maior pontuação.
+%% Modificadores diferentes podem acessar campos iguais em um pacote. Por
+%% exemplo os modificadores http\_method e http\_header, o http\_method
+%% acessa apenas o metodo do pacote http, mas o http\_header acessa todo
+%% o cabeçalho (incluindo o proprio metodo). Com isso, a string "GET"
+%% receberia a pontuação correspondente a esses dois modificadores.
+
+More precisely, we calculate the contribution of the frequency of
+content modifiers with the following formula:
+
+\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{modifier\_frequency[modifier_{ij}]}{max(modifier\_frequency)}}{N}\]
+
+\noindent
+, where $N$ is the number of contents in a given rule and $M$
+is the number of content modifiers\Mar{in a rule? what is a content
+  modifier in a rule?}. \Luc{M eh o numero de modificadores associados
+  a cada content. Temos content[1], content[2], ..., content[N]. Cada
+  um desses contents tem modificadores associados content[i][1],
+  content[i][2], ..., content[i][M].}
 

### 2020-01-14 paper/paper.tex
4.3.4 comments

@@
   e tcp)? ou seja, quantas options do tipo ``content'' possuem
-  modificadores.}  For example, the creator of the rule may want to
+  modificadores.} \Luc{conseguirei esses dados} For example, the creator of the rule may want to
 indicate that a specific string be present in the header of an HTTP
@@
 
-Suricata supports content modifiers for tcp and http. Given that
+Suricata supports content modifiers for tcp and http.
+\Luc{Eh possivel utilizar os modificadores em qualquer protocolo.
+Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo.
+O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós.
+Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso.
+Se necessário posso explicar.} Given that
 content options often use modifiers\Mar{assumindo que eh

### 2020-01-15 paper/paper.tex
pass

--- 2020-01-14 paper/paper.tex
@@
 \subsection{\Fix{JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit}}
-\label{sec:content-example}
+\label{sec:jboss}
 
@@
 \section{Technique}
+\label{sec:technique}
 
@@
 Section~\ref{sec:content-example}.
+
+\Mar{Revisar.........................}
+\Fix{
+A avaliação que fazemos aqui pontua um dado content de acordo com
+quantas vezes o campo em que ele se encontra eh referenciado nas
+regras. Por exemplo, vamos supor que o tokenizer produziu a opção
+content: "/HtmlAdaptor". Essa string pertence ao campo "Request
+URI:". Esse campo pode ser acessado pelo modificador http\_uri. O
+tokenizer associa o content a esse modificador.
+}
 
@@
 mentioned, \percRulesWithContent\ of Suricata rules involve content
-options. Furthermnore, we observed that many of these rules indicate
-what parts of the message the content string should be checked
-against.\Mar{Lucas, (1) vc. tem dados para dar suporte a isto (de http
-  e tcp)? ou seja, quantas options do tipo ``content'' possuem
-  modificadores.}  For example, the creator of the rule may want to
-indicate that a specific string be present in the header of an HTTP
-request message. Content \emph{modifiers} serve this purpose. For
-example, the modifier \CodeIn{http\_header} in the option
-\CodeIn{content: ``keep-alive''; http\_header} states the requirement that
-the string ``keep-alive'' be present in the header of the message. (A
-keep-alive connection attribute indicates that a single TCP connection
-should remain open for multiple HTTP requests/responses.)
+options. Furthermore, we observed---by analyzing the dataset of
+existing rules (see Section~\ref{sec:dataset-benign})---that often
+these rules indicate what part of the message the content string
+should be checked against.\Mar{Lucas, (1) vc. tem dados para dar
+  suporte a isto (de http e tcp)? ou seja, quantas options do tipo
+  ``content'' possuem modificadores.} \Luc{conseguirei esses dados}
+For example, the creator of the rule may want to indicate that a
+specific string be present in the header of an HTTP request
+message. Content \emph{modifiers} serve this purpose. For example, the
+modifier \CodeIn{http\_header} in the option \CodeIn{content:
+  ``keep-alive''; http\_header} states the requirement that the string
+``keep-alive'' be present in the header of the message. (A keep-alive
+connection attribute indicates that a single TCP connection should
+remain open for multiple HTTP requests/responses.)
 Figure~\Fix{adicione pequena figura...verifique se ainda faz sentido
@@
 
-Suricata supports content modifiers for tcp and http. Given that
-content options often use modifiers\Mar{assumindo que eh
-  verdade. confirmar estat. solicitada acima...}, \tname\ uses
-associated frequency to discriminate more or less likely rules. For
-example, historical data indicates that content options that use the
-\CodeIn{http\_connection} modifiers are much less likely than content
-options that use the modifier
+%% Suricata supports content modifiers for tcp and http.
+\Luc{Eh possivel utilizar os modificadores em qualquer protocolo.
+Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo.
+O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós.
+Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso.
+Se necessário posso explicar.} Given that
+%% content options often use modifiers\Mar{assumindo que eh
+%%   verdade. confirmar estat. solicitada acima...},
+
+\tname\ uses the frequency of content modifiers to discriminate more
+or less likely rules. For example, historical data indicates that
+content options that use the \CodeIn{http\_connection} modifier are
+much less likely than content options that use the modifier
 \CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
-shows the distribution of content modifiers for http. We ommit the
-distribution for tcp for space. Similarly to the the previous
-heuristic functions, this function uses the frequency of content
-modifiers to calibrate the contribution of each rule.
+shows the histogram of content modifiers for http. Although content
+modifiers are applicable for rules of any protocol in Suricatax,
+currently, this heuristic function is only supported for tcp and http
+in \tname. The rationale is that (1) these are the protocols involving
+most attacks \Mar{Lucas, please confirm.} and (2) the payload of
+messages of different protocols are not as segmented as the case of
+tcp and http, which makes the reference to specific parts of the
+payload less common for rule writers and more challenging for
+\tname\Mar{Lucas, please confirm}.
+
+Similarly to the the previous heuristic functions, this function uses
+the frequency of content modifiers to calibrate the contribution of
+each rule. More precisely, we calculate this heuristic function with
+the following formula:
+
+\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{N}\]
+
+\noindent
+, where $N$ is the number of contents in a given rule and $M$ is the
+number of content modifiers\Mar{isto eh confuso. vc. esta dizendo que
+  o mesmo content pode ter mais de um modificador? ou a string poderia
+  ser acessada por mais de um modificador. se for este o caso,
+  sinceramente, achei isto bem confuso. afinal, a regra jah indica
+  qual o modificador foi escolhido para acessar a string.}. \Luc{M eh
+  o numero de modificadores associados a cada content. Temos
+  content[1], content[2], ..., content[N]. Cada um desses contents tem
+  modificadores associados content[i][1], content[i][2], ...,
+  content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
+  content. a frequencia seria 0, correto?}
 
@@
 
-\Mar{----------------------------------------------------------- estou
-aqui}
-
-\subsection{Pontuação}
-A avaliação que fazemos aqui pontua um dado content de acordo com quantas vezes o campo em que ele se encontra eh referenciado nas regras.
+\Mar{Lucas, use uma regra concreta com a regra da Secao 3.3 (assumindo
+que vai mante-la) e mostre como eh feito o calculo da regra desta
+funcao heuristica para regra inteira.}
+
+Consider a rule containing the option \CodeIn{content:
+  ``/HtmlAdaptor''; http\_uri}, as the one from
+Section~\ref{sec:jboss}. The contribution of this option to the score
+of the corresponding rule would be $\frac{10,111}{10,111*1}$. The
+value of $N$ for this case is 1 as the field of the HTTP message
+\CodeIn{Request URI} holding the string can be accessed only through
+the modifier \CodeIn{http\_uri}.
+
+%% A avaliação que fazemos aqui pontua um dado content de acordo com quantas vezes o campo em que ele se encontra eh referenciado nas regras.
 Por exemplo, vamos supor que o tokenizer produziu a opção content: "/HtmlAdaptor".
@@
 O tokenizer associa o content a esse modificador.
-Posteriormente, esse content vai receber uma pontuação correspondente a popularidade desse modificador:\\
-\[\frac{10111}{10111*1}\]
-A multiplicação por 1 no denominador eh referente a quantidade de modificadores que estão associadas a esse content.
-No caso da opção content: "keep-alive", essa string pertence ao campo "Connection:".
-Como falado acima, o suricata pode acessar esse campo pelos modificadores http\_header ou http\_connection.
-Portanto, esse content será pontuado de acordo com a popularidade desses dois modificadores:
+%% Posteriormente, esse content vai receber uma pontuação
+%% correspondente a popularidade desse modificador:\\
+
+\Mar{----------------------------------------------------------- estou
+aqui}
+
+A multiplicação por 1 no denominador eh referente a quantidade de
+modificadores que estão associadas a esse content.\Mar{nao foi isto
+  que vc. disse acima, Lucas. foi dito que N (=2) era o numero de
+  contents e o exemplo acima tem apenas 1 content.}  No caso da opção
+content: "keep-alive", essa string pertence ao campo
+"Connection:". Como falado acima, o suricata pode acessar esse campo
+pelos modificadores http\_header ou http\_connection. Portanto, esse
+content será pontuado de acordo com a popularidade desses dois
+modificadores:
 \[\frac{4089+29}{10111*2}\]
 
-Content options are
-associated with specific packets fields\Mar{or modifiers?}.\Mar{why
-  GET is associated with http\_method and http\_header?}\Mar{are all
-  content associated with one or many fields or positions in a packet?
-  can you give examples other than GET to explain this? if yes,
-  shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
-  is important to justify (to explain rationale)---how is it different
-  from counting rarity?}
+%% Content options are
+%% associated with specific packets fields\Mar{or modifiers?}.\Mar{why
+%%   GET is associated with http\_method and http\_header?}\Mar{are all
+%%   content associated with one or many fields or positions in a packet?
+%%   can you give examples other than GET to explain this? if yes,
+%%   shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
+%%   is important to justify (to explain rationale)---how is it different
+%%   from counting rarity?}
 
@@
 %% receberia a pontuação correspondente a esses dois modificadores.
-
-More precisely, we calculate the contribution of the frequency of
-content modifiers with the following formula:
-
-\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{modifier\_frequency[modifier_{ij}]}{max(modifier\_frequency)}}{N}\]
-
-\noindent
-, where $N$ is the number of contents in a given rule and $M$
-is the number of content modifiers\Mar{in a rule? what is a content
-  modifier in a rule?}. \Luc{M eh o numero de modificadores associados
-  a cada content. Temos content[1], content[2], ..., content[N]. Cada
-  um desses contents tem modificadores associados content[i][1],
-  content[i][2], ..., content[i][M].}
 

### 2020-01-15 paper/paper.tex
4.3.4 comments

@@
 shows the histogram of content modifiers for http. Although content
-modifiers are applicable for rules of any protocol in Suricatax,
+modifiers are applicable for rules of any protocol in Suricata,
 currently, this heuristic function is only supported for tcp and http
 in \tname. The rationale is that (1) these are the protocols involving
-most attacks \Mar{Lucas, please confirm.} and (2) the payload of
+most attacks \Luc{Ainda n encontrei nenhuma informação concreta afirmando isso.
+N podemos citar o fato de que 80\% das regras do dataset são desses 2 protocolos?}\Mar{Lucas, please confirm.} and (2) the payload of
 messages of different protocols are not as segmented as the case of
@@
 payload less common for rule writers and more challenging for
-\tname\Mar{Lucas, please confirm}.
+\tname\Mar{Lucas, please confirm}\Luc{verificar}.
 
@@
 
-\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{N}\]
+$\sum_{i=1}^{N}\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}$
 
@@
 , where $N$ is the number of contents in a given rule and $M$ is the
-number of content modifiers\Mar{isto eh confuso. vc. esta dizendo que
+number of content modifiers\Luc{$M_i$ eh o numero de modificadores que podem referenciar a posição em q se encontra o $content_i$}\Mar{isto eh confuso. vc. esta dizendo que
   o mesmo content pode ter mais de um modificador? ou a string poderia
@@
   sinceramente, achei isto bem confuso. afinal, a regra jah indica
-  qual o modificador foi escolhido para acessar a string.}. \Luc{M eh
+  qual o modificador foi escolhido para acessar a string.}.\Luc{Cada content em uma regra só pode possuir um modificador para indicar a posição onde ele deve ser procurado.
+Entretanto sim, uma "string poderia
+  ser acessada por mais de um modificador", foi o caso daquele exemplo com o http\_header.} \Luc{M eh
   o numero de modificadores associados a cada content. Temos
@@
   content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
-  content. a frequencia seria 0, correto?}
+  content. a frequencia seria 0, correto?} \Luc{Sim}
 
@@
 que vai mante-la) e mostre como eh feito o calculo da regra desta
-funcao heuristica para regra inteira.}
+funcao heuristica para regra inteira.}\\
+\Luc{-------------}
 
@@
 A multiplicação por 1 no denominador eh referente a quantidade de
-modificadores que estão associadas a esse content.\Mar{nao foi isto
+modificadores que estão associadas a esse content.\Luc{O correto eh ser dividido pela quantidade de modificadores.
+Atualizei a formula para tentar representar isso.
+Considere que no exemplo acima o content pudesse ser acessado por mais de um modificador.
+A pontuação dele então seria $\frac{10,111+x}{10,111*1}$ > 1.
+Para manter a pontuação entre 0-1, o certo nesse caso seria fazer $\frac{10,111+x}{10,111*2}$}\Mar{nao foi isto
   que vc. disse acima, Lucas. foi dito que N (=2) era o numero de

### 2020-01-15 paper/paper.tex
pass

--- 2020-01-15 paper/paper.tex
@@
 
-\begin{figure}[H]
+\begin{figure}[t!]
   \centering
@@
 these rules indicate what part of the message the content string
-should be checked against.\Mar{Lucas, (1) vc. tem dados para dar
-  suporte a isto (de http e tcp)? ou seja, quantas options do tipo
-  ``content'' possuem modificadores.} \Luc{conseguirei esses dados}
-For example, the creator of the rule may want to indicate that a
-specific string be present in the header of an HTTP request
-message. Content \emph{modifiers} serve this purpose. For example, the
-modifier \CodeIn{http\_header} in the option \CodeIn{content:
-  ``keep-alive''; http\_header} states the requirement that the string
-``keep-alive'' be present in the header of the message. (A keep-alive
-connection attribute indicates that a single TCP connection should
-remain open for multiple HTTP requests/responses.)
-Figure~\Fix{adicione pequena figura...verifique se ainda faz sentido
-  depois do que falei acima...}
+should be checked against.\Mar{Lucas, vc. tem dados para dar suporte a
+  isto? ou seja, quantas options do tipo ``content'' possuem
+  modificadores.} \Luc{conseguirei esses dados} For example, the
+creator of the rule may want to indicate that a specific string be
+present in the header of an HTTP request message. Content
+\emph{modifiers} serve this purpose. For example, the modifier
+\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
+  http\_header} states the requirement that the string ``keep-alive''
+be present in the header of the message. (A keep-alive connection
+attribute indicates that a single TCP connection should remain open
+for multiple HTTP requests/responses.)  Figure~\Fix{adicione pequena
+  figura...verifique se ainda faz sentido depois do que falei
+  acima...}
 
@@
 shows the histogram of content modifiers for http. Although content
-modifiers are applicable for rules of any protocol in Suricatax,
-currently, this heuristic function is only supported for tcp and http
-in \tname. The rationale is that (1) these are the protocols involving
-most attacks \Mar{Lucas, please confirm.} and (2) the payload of
-messages of different protocols are not as segmented as the case of
-tcp and http, which makes the reference to specific parts of the
-payload less common for rule writers and more challenging for
-\tname\Mar{Lucas, please confirm}.
+modifiers are applicable to rules of any protocol in Suricata,
+currently, \tname\ only supports tcp and http. These are the protocols
+involving ~80\% of the attacks.
+
+%% and
+%% (2) the payload of messages of different protocols are not as
+%% segmented as the case of tcp and http. Consequently, rule writers do
+%% not refer to specific parts of the payload as often as the task o and
+%% more challenging for
 
@@
 the frequency of content modifiers to calibrate the contribution of
-each rule. More precisely, we calculate this heuristic function with
-the following formula:
-
-\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{N}\]
+each rule. More precisely, we compute this function with the following
+formula:
+
+\[\sum_{i=1}^{N}\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]
 
 \noindent
-, where $N$ is the number of contents in a given rule and $M$ is the
-number of content modifiers\Mar{isto eh confuso. vc. esta dizendo que
-  o mesmo content pode ter mais de um modificador? ou a string poderia
-  ser acessada por mais de um modificador. se for este o caso,
-  sinceramente, achei isto bem confuso. afinal, a regra jah indica
-  qual o modificador foi escolhido para acessar a string.}. \Luc{M eh
-  o numero de modificadores associados a cada content. Temos
-  content[1], content[2], ..., content[N]. Cada um desses contents tem
-  modificadores associados content[i][1], content[i][2], ...,
-  content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
-  content. a frequencia seria 0, correto?}
-
-\begin{figure}[H]
+, where $N$ is the number of contents in a given rule and $M_i$ is the
+number of content modifiers that could access the string associated
+with content $i$. \Mar{Qual o rationale para considerar as varias
+  alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}%% Isto
+                                                                             %% nao
+                                                                             %% intuitivo
+                                                                             %% para
+                                                                             %% mim
+                                                                             %% -Marcelo
+
+%% \Luc{$M_i$ eh o numero de modificadores que podem referenciar a posição em q se encontra o $content_i$}\Mar{isto eh confuso. vc. esta dizendo que
+%%   o mesmo content pode ter mais de um modificador? ou a string poderia
+%%   ser acessada por mais de um modificador. se for este o caso,
+%%   sinceramente, achei isto bem confuso. afinal, a regra jah indica
+%%   qual o modificador foi escolhido para acessar a string.}.\Luc{Cada content em uma regra só pode possuir um modificador para indicar a posição onde ele deve ser procurado.
+Entretanto sim, uma "string poderia
+%%   ser acessada por mais de um modificador", foi o caso daquele exemplo com o http\_header.} \Luc{M eh
+%%   o numero de modificadores associados a cada content. Temos
+%%   content[1], content[2], ..., content[N]. Cada um desses contents tem
+%%   modificadores associados content[i][1], content[i][2], ...,
+%%   content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
+%%   content. a frequencia seria 0, correto?} \Luc{Sim}
+
+\begin{figure}[t!]
   \centering
@@
 
+Consider a rule containing the option \CodeIn{content:
+  ``/HtmlAdaptor''; http\_uri}, as the one from
+Section~\ref{sec:jboss}. The contribution of this option to the score
+of the corresponding rule would be $0.5=\frac{10,111}{10,111*1}$. The
+value of $M_i$ for this case is 1 as the field of the HTTP message
+\CodeIn{Request URI} holding the string can only be accessed through
+the modifier \CodeIn{http\_uri}. For the option \CodeIn{content:
+  ``keep-alive''; <mod>}, either \CodeIn{http\_header} or
+\CodeIn{http\_connection} could substitute the placeholder
+\CodeIn{<mod>} to locate the string ``keep-alive'' stored in the field
+\CodeIn{Connection} in the header of an HTTP request
+message\Mar{Lucas, a imagem mostrando header com campos Request URI e
+  Connection e os conteudos citados acima pode ajudar.}. Consequently,
+for this case, the constribution of content modifiers would be
+$0.2=\frac{4089+29}{10111*2}$. Assuming that both options are
+mentioned in a rule, the aggregated contribution of this heuristic
+function would be \Fix{...}\Mar{Lucas, confirme que nao esta faltando
+  calcular a media (dividir tudo por N).}
+
 %% \section{Modificadores}É possivel deixar que uma regra verifique todo o payload do pacote em busca de um match com um content ou especificar partes do payload onde esse content deve ser procurado.
@@
 
-\Mar{Lucas, use uma regra concreta com a regra da Secao 3.3 (assumindo
-que vai mante-la) e mostre como eh feito o calculo da regra desta
-funcao heuristica para regra inteira.}
-
-Consider a rule containing the option \CodeIn{content:
-  ``/HtmlAdaptor''; http\_uri}, as the one from
-Section~\ref{sec:jboss}. The contribution of this option to the score
-of the corresponding rule would be $\frac{10,111}{10,111*1}$. The
-value of $N$ for this case is 1 as the field of the HTTP message
-\CodeIn{Request URI} holding the string can be accessed only through
-the modifier \CodeIn{http\_uri}.
+%% \Mar{Lucas, use uma regra concreta com a regra da Secao 3.3 (assumindo
+%% que vai mante-la) e mostre como eh feito o calculo da regra desta
+%% funcao heuristica para regra inteira.}\\
+%% \Luc{-------------}
 
@@
 
-\Mar{----------------------------------------------------------- estou
-aqui}
-
-A multiplicação por 1 no denominador eh referente a quantidade de
-modificadores que estão associadas a esse content.\Mar{nao foi isto
-  que vc. disse acima, Lucas. foi dito que N (=2) era o numero de
-  contents e o exemplo acima tem apenas 1 content.}  No caso da opção
-content: "keep-alive", essa string pertence ao campo
-"Connection:". Como falado acima, o suricata pode acessar esse campo
-pelos modificadores http\_header ou http\_connection. Portanto, esse
-content será pontuado de acordo com a popularidade desses dois
-modificadores:
-\[\frac{4089+29}{10111*2}\]
+%% A multiplicação por 1 no denominador eh referente a quantidade de
+%% modificadores que estão associadas a esse content.\Luc{O correto eh ser dividido pela quantidade de modificadores.
+Atualizei a formula para tentar representar isso.
+Considere que no exemplo acima o content pudesse ser acessado por mais de um modificador.
+A pontuação dele então seria $\frac{10,111+x}{10,111*1}$ > 1.
+Para manter a pontuação entre 0-1, o certo nesse caso seria fazer $\frac{10,111+x}{10,111*2}$}\Mar{nao foi isto
+%%   que vc. disse acima, Lucas. foi dito que N (=2) era o numero de
+%%   contents e o exemplo acima tem apenas 1 content.}
 

### 2020-01-15 paper/paper.tex
4.3.4

@@
 
-\[\sum_{i=1}^{N}\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]
+\[\frac{\sum_{i=1}^{N}\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}}{N}\]
 
@@
 with content $i$. \Mar{Qual o rationale para considerar as varias
-  alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}%% Isto
+  alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}0
+  \Luc{Vc está supondo que as regras que geramos incluem os modificadores e que deveriamos pontuar somente de acordo com os que estão presentes?
+Não eh o caso.
+A ideia aqui basicamente eh: se um certo content do payload está numa posição q eh muito verificada nas regras, possivelmente esse content eh mais relevante do q outro q está numa posição q raramente eh verificada.
+Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no dataset.}
+  \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content.
+Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}
+
+                                                                             %% Isto
                                                                              %% nao
@@
           %          ylabel={\#rules},
-          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection},
+          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection, http\_request\_line},
           xtick=data,
@@
         ]
-        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29)};
+        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29) (http\_request\_line,11)};
       \end{axis}
@@
     frequency.\Mar{what about ``none'' (no modifier)?}}
+  \Luc{Uma regra pode sim ter contents sem modificadores, mas isso n vem ao caso aqui.
+Nessa heuristica toda string do payload vai ter associada a ela os modificadores q podem acessar a sua posição.
+Portanto, não vai existir o caso de não haver modificador.
+Atenção: isso não quer dizer q as regras q geramos incluem os modificadores, elas n incluem.}
 \end{figure}
@@
 message\Mar{Lucas, a imagem mostrando header com campos Request URI e
-  Connection e os conteudos citados acima pode ajudar.}. Consequently,
+  Connection e os conteudos citados acima pode ajudar.}\Luc{adicionar}. Consequently,
 for this case, the constribution of content modifiers would be
@@
 function would be \Fix{...}\Mar{Lucas, confirme que nao esta faltando
-  calcular a media (dividir tudo por N).}
+  calcular a media (dividir tudo por N).} \Luc{Sim, tem q dividir. A implementação estava errada, está de acordo com a formula agora}
 

### 2020-01-15 paper/paper.tex
4.3.4 fix

@@
   \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content.
-Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}
+Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores que conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}
 

### 2020-01-15 paper/paper.tex
pass on 4.3.3

--- 2020-01-15 paper/paper.tex
@@
 Similarly to the the previous heuristic functions, this function uses
-the frequency of content modifiers to calibrate the contribution of
-each rule. More precisely, we compute this function with the following
+the frequency of content modifiers to calibrate the score of each
+rule. More precisely, we compute this function with the following
 formula:
 
-\[\sum_{i=1}^{N}\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]
+\[\frac{\sum_{i=1}^{N}g(i)}{N},~g(i)=\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]
 
 \noindent
-, where $N$ is the number of contents in a given rule and $M_i$ is the
-number of content modifiers that could access the string associated
-with content $i$. \Mar{Qual o rationale para considerar as varias
-  alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}%% Isto
-                                                                             %% nao
-                                                                             %% intuitivo
-                                                                             %% para
-                                                                             %% mim
-                                                                             %% -Marcelo
+, where $N$ is the number of contents options, $g(i)$ is the
+contribution of content option $i$ to the score,
+$\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
+of a modifier $j$ to content option $i$,
+$\mathit{max(modifier\_frequency)}$ refers to the modifier with
+maximum popularity (for normalization), and $M_i$ is the number of
+content modifiers that could access the string associated with content
+$i$.
+
+%% \Mar{Qual o rationale para considerar as varias
+%%   alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}0
+%%   \Luc{Vc está supondo que as regras que geramos incluem os modificadores e que deveriamos pontuar somente de acordo com os que estão presentes?
+Não eh o caso.
+A ideia aqui basicamente eh: se um certo content do payload está numa posição q eh muito verificada nas regras, possivelmente esse content eh mais relevante do q outro q está numa posição q raramente eh verificada.
+Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no dataset.}
+%%   \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content.
+Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores que conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}
 
@@
   \caption{\label{fig:distribution-content_modifiers}Content modifiers
-    frequency.\Mar{what about ``none'' (no modifier)?}}
+    frequency.}
 \end{figure}
 
+%\Mar{what about ``none'' (no modifier)?}}
+%  \Luc{Uma regra pode sim ter contents sem modificadores, mas isso n
+    %% vem ao caso aqui. Nessa heuristica toda string do payload vai ter
+    %% associada a ela os modificadores q podem acessar a sua
+    %% posição. Portanto, não vai existir o caso de não haver
+    %% modificador. Atenção: isso não quer dizer q as regras q geramos
+    %% incluem os modificadores, elas n incluem.
+
 Consider a rule containing the option \CodeIn{content:
-  ``/HtmlAdaptor''; http\_uri}, as the one from
-Section~\ref{sec:jboss}. The contribution of this option to the score
-of the corresponding rule would be $0.5=\frac{10,111}{10,111*1}$. The
-value of $M_i$ for this case is 1 as the field of the HTTP message
-\CodeIn{Request URI} holding the string can only be accessed through
-the modifier \CodeIn{http\_uri}. For the option \CodeIn{content:
-  ``keep-alive''; <mod>}, either \CodeIn{http\_header} or
-\CodeIn{http\_connection} could substitute the placeholder
-\CodeIn{<mod>} to locate the string ``keep-alive'' stored in the field
-\CodeIn{Connection} in the header of an HTTP request
-message\Mar{Lucas, a imagem mostrando header com campos Request URI e
-  Connection e os conteudos citados acima pode ajudar.}. Consequently,
-for this case, the constribution of content modifiers would be
-$0.2=\frac{4089+29}{10111*2}$. Assuming that both options are
-mentioned in a rule, the aggregated contribution of this heuristic
-function would be \Fix{...}\Mar{Lucas, confirme que nao esta faltando
-  calcular a media (dividir tudo por N).}
+  ``/HtmlAdaptor''}. The contribution of this option to the score of
+the corresponding rule would be $0.5$
+(=$\frac{10,111}{10,111*1}$). The value of $M_i$ for this case is 1 as
+the field of the HTTP message \CodeIn{Request URI}---holding the
+string ``/HtmlAdaptor''---can be accessed through the modifier
+\CodeIn{http\_uri}. Conceptually, the relatively high heuristic value
+given to this option indicates that this is a reasonable option
+because the string used in the content option (``/HtmlAdaptor'') is
+stored in a specific field of the HTTP message (\CodeIn{Request URI})
+and can be accessible through content modifiers
+(\CodeIn{http\_uri}). For the option \CodeIn{content: ``keep-alive''},
+either \CodeIn{http\_header} or \CodeIn{http\_coxnnection} could be
+used as modifier to locate the string ``keep-alive'' stored in the
+field \CodeIn{Connection} (see Figure~\Fix{xxx}). Consequently, for
+this case, the contribution of modifiers to this content option would
+be $0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
+used in a rule, the aggregated contribution of this heuristic function
+would be 0.35 (=(0.2+0.5)/2).
 

### 2020-01-15 paper/paper.tex
4.3.4

@@
 currently, \tname\ only supports tcp and http. These are the protocols
-involving ~80\% of the attacks.
+involving ~80\% of the attacks. \Luc{80\% das regras do nosso dataset.}
 
@@
 the corresponding rule would be $0.5$
-(=$\frac{10,111}{10,111*1}$). The value of $M_i$ for this case is 1 as
+(=$\frac{10,111+11}{10,111*2}$). The value of $M_i$ for this case is 2 as
 the field of the HTTP message \CodeIn{Request URI}---holding the
-string ``/HtmlAdaptor''---can be accessed through the modifier
-\CodeIn{http\_uri}. Conceptually, the relatively high heuristic value
+string ``/HtmlAdaptor''---can be accessed through the modifiers
+\CodeIn{http\_uri} and \CodeIn{http\_request\_line} \Luc{Inclui esse modificador pq ele tambem pode acessar o campo Request URI.
+Isso eh um problema?
+O http\_uri eh o modificador mais comum, mas o fato de haver o modificador http\_request\_line que tambem accessa a mesma posição faz com que o fitness maximo seja 0.5, pois a ocorrencia de http\_request\_line eh irrelevante no calculo.}.
+Conceptually, the relatively high heuristic value
 given to this option indicates that this is a reasonable option
@@
 and can be accessible through content modifiers
-(\CodeIn{http\_uri}). For the option \CodeIn{content: ``keep-alive''},
-either \CodeIn{http\_header} or \CodeIn{http\_coxnnection} could be
+(\CodeIn{http\_uri} and \CodeIn{http\_request\_line}). For the option \CodeIn{content: ``keep-alive''},
+either \CodeIn{http\_header} or \CodeIn{http\_connection} could be
 used as modifier to locate the string ``keep-alive'' stored in the

### 2020-01-15 paper/paper.tex
another pass on 4.3.4

--- 2020-01-15 paper/paper.tex
@@
 attribute indicates that a single TCP connection should remain open
-for multiple HTTP requests/responses.)  Figure~\Fix{adicione pequena
-  figura...verifique se ainda faz sentido depois do que falei
-  acima...}
+for multiple HTTP requests/responses.)
+Figure \Fix{xxx}
 
@@
 
-\tname\ uses the frequency of content modifiers to discriminate more
-or less likely rules. For example, historical data indicates that
-content options that use the \CodeIn{http\_connection} modifier are
-much less likely than content options that use the modifier
+\tname\ uses the frequency of content modifiers to discriminate rules
+more or less likely to reflect actual rules. For example, historical
+data indicates that content options that use the
+\CodeIn{http\_connection} modifier are much less likely than content
+options that use the modifier
 \CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
 shows the histogram of content modifiers for http. Although content
-modifiers are applicable to rules of any protocol in Suricata,
-currently, \tname\ only supports tcp and http. These are the protocols
-involving ~80\% of the attacks.
+modifiers are applicable to rules of any protocol supported by
+Suricata, currently, \tname\ only supports tcp and http. These are the
+protocols involving ~80\% of the rules from our dataset.
 
@@
 %% incluem os modificadores, elas n incluem.
+%\Luc{Inclui esse modificador pq ele tambem pode acessar o campo Request URI.
+Isso eh um problema?
+O http\_uri eh o modificador mais comum, mas o fato de haver o modificador http\_request\_line que tambem accessa a mesma posição faz com que o fitness maximo seja 0.5, pois a ocorrencia de http\_request\_line eh irrelevante no calculo.}
 
@@
   ``/HtmlAdaptor''}. The contribution of this option to the score of
-the corresponding rule would be $0.5$
-(=$\frac{10,111}{10,111*1}$). The value of $M_i$ for this case is 1 as
-the field of the HTTP message \CodeIn{Request URI}---holding the
-string ``/HtmlAdaptor''---can be accessed through the modifier
-\CodeIn{http\_uri}. Conceptually, the relatively high heuristic value
-given to this option indicates that this is a reasonable option
-because the string used in the content option (``/HtmlAdaptor'') is
-stored in a specific field of the HTTP message (\CodeIn{Request URI})
-and can be accessible through content modifiers
-(\CodeIn{http\_uri}). For the option \CodeIn{content: ``keep-alive''},
-either \CodeIn{http\_header} or \CodeIn{http\_coxnnection} could be
-used as modifier to locate the string ``keep-alive'' stored in the
-field \CodeIn{Connection} (see Figure~\Fix{xxx}). Consequently, for
-this case, the contribution of modifiers to this content option would
-be $0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
+the rule that uses it would be $0.5$
+(=$\frac{10,111+11}{10,111*2}$). The value of $M_i$ for this case is 2
+as the field of the HTTP message \CodeIn{Request URI}---holding the
+string ``/HtmlAdaptor''---can be accessed through the modifiers
+\CodeIn{http\_uri} and \CodeIn{http\_request\_line}. The relatively
+high heuristic value given to this option suggests it is a plausible
+option to be used in rules---the string used in the content option
+(``/HtmlAdaptor'') is stored in a specific field of the HTTP message
+(\CodeIn{Request URI}) and can be accessed by two content modifiers,
+namely (\CodeIn{http\_uri} and \CodeIn{http\_request\_line}). For the
+option \CodeIn{content: ``keep-alive''}, either the modifier
+\CodeIn{http\_header} or the modifier \CodeIn{http\_connection} could
+be used to locate the string ``keep-alive'' stored in the field
+\CodeIn{Connection} (see Figure~\Fix{xxx}). Consequently, for this
+case, the contribution of modifiers to this content option would be
+$0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
 used in a rule, the aggregated contribution of this heuristic function

### 2020-01-17 paper/paper.tex
pass

@@
   \vspace{-2ex}
-  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype}]{pingscan.suricata.synth}
+  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{pingscan.suricata.synth}
   \vspace{-3ex}  
@@
 
+\begin{figure}[h]
+  \vspace{-2ex}
+  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{top-pingscan.rules}
+  \vspace{-3ex}  
+\end{figure}
+
+\Luc{Apenas 3 regras tiveram f1 de 100\%. O total de regras nesse caso foi de 13.}
 \Fix{...listem 5 primeiras regras neste caso...}
 
-\subsection{\Fix{JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit}}
-\label{sec:jboss}
+\subsection{Remote Code Execution}
+\label{sec:rce}
+
+"RCE refers to the process by which an agent can exploit a network vulnerability to run arbitrary code on a targeted machine or system.
+In an RCE attack, hackers intentionally exploit a remote code execution vulnerability to run malware.
+By prompting the targeted device to perform code execution, a hacker can run their own programming in its place.
+This programming can then enable them to gain full access, steal data, carry out a full distributed denial of service (DDoS) attack, destroy files and infrastructure, or engage in illegal activity.
+And as the term remote execution suggests, an RCE cyberattack can take place from any geophysical location."\\ Texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution
+\Luc{---------------- estou aqui}
 
@@
   dataset}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
-  (\%) das 28K regras tem contents.} As consequence, we defined a
+  (\%) das 28K regras tem contents.} \Luc{sao 69417 contents nas 26094 regras q utilizam contents} As consequence, we defined a
 heuristic function that takes the number of content options into
@@
   isto? ou seja, quantas options do tipo ``content'' possuem
-  modificadores.} \Luc{conseguirei esses dados} For example, the
+  modificadores.} \Luc{das 26k regras que utilizam content, 13k delas utilizam modificadores http q sao os q nós consideramos.} For example, the
 creator of the rule may want to indicate that a specific string be
@@
 for multiple HTTP requests/responses.)
-Figure \Fix{xxx}
+Figure~\ref{fig:http-header-example}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.5]{figs/http_header-example.png}
+\caption{\CodeIn{http\_header} example.}
+\label{fig:http-header-example}
+\end{figure}
 
@@
 be used to locate the string ``keep-alive'' stored in the field
-\CodeIn{Connection} (see Figure~\Fix{xxx}). Consequently, for this
+\CodeIn{Connection} (see Figurgure~\ref{fig:http-header-example}). Consequently, for this
 case, the contribution of modifiers to this content option would be
@@
 \noindent
-\textbf{Malicious traffic.}~Nikito?\Fix{where do these
+\textbf{Malicious traffic.}~Nikto?\Fix{where do these
   attacks/messages come from?}
+  \Luc{Nós executamos um servidor http (Apache2) numa maquina e executamos o Nikto em outra maquina atacando o servidor http. Capturamos o trafego q chegava ao servidor e eh ele q estamos utilizando.}
 

### 2020-01-17 paper/paper.tex

@@
 
-"RCE refers to the process by which an agent can exploit a network vulnerability to run arbitrary code on a targeted machine or system.
+\Luc{"RCE refers to the process by which an agent can exploit a network vulnerability to run arbitrary code on a targeted machine or system.
 In an RCE attack, hackers intentionally exploit a remote code execution vulnerability to run malware.
@@
 This programming can then enable them to gain full access, steal data, carry out a full distributed denial of service (DDoS) attack, destroy files and infrastructure, or engage in illegal activity.
-And as the term remote execution suggests, an RCE cyberattack can take place from any geophysical location."\\ Texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution
-\Luc{---------------- estou aqui}
+And as the term remote execution suggests, an RCE cyberattack can take place from any geophysical location."\\ Texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution \\
+---------------- estou aqui}
 

### 2020-01-17 paper/paper.tex
quick pass

--- 2020-01-15 paper/paper.tex
@@
 \label{sec:jboss}
+\label{sec:content-example}
 
@@
 set of rules from public rulesets and produces on output a list of
-plausible rules to isolate the malicious
+plausible rules to isolate the malicious from the benign
 traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
@@
 input in the form of pcap files~\cite{pcap}. In this step,
-\tname\ detects rule options in correspondence with the input messages
-(see Section~\ref{sec:rules-and-packets}). It is worth noting that the
-rule produced in this step can include more options than the golden
-rule (see Section~\ref{sec:suri-metas-coverage}). Consequently, the
-use of that rule could lead, in principle, to false negatives as not
-all manifestations of the same attack are identical. Second, \tname{}
-generates plausible rules similar to the one produced in the previous
-step. It is worth noting that any subset of the original set of
-options captures the malicious traffic. As the generation procedure
-discards options from the initial rule, it satisfies the invariant of
-capturing malicious traffic by construction. However, the goal of
-\tname\ is to produce rules that capture \emph{only} malicious
-traffic. To achieve that goal, \tname{} uses benign traffic from
-public databases to check that the \nids\ instantiated with a given
-rule does not capture benign traffic. The minimization step uses a
-population-based gradient descent-like search to find all plausible
-rules, which are derived from the rule obtained in the previous
-step. Third, uses several heuristic functions derived from a dataset
-of rules to rank the rules generated in the previous step.
+\tname\ detects rule options directly from the input messages (see
+Section~\ref{sec:rules-and-packets}). As only malicious traffic is
+considered in this step, the rule produced is likely overspecified,
+\ie{}, it includes more options than the golden rule (see
+Section~\ref{sec:suri-metas-coverage}). Consequently, the use of that
+rule could lead to false negatives as not all manifestations of the
+same attack are identical. Second, \tname{} generates plausible rules
+that are similar to the seed rule produced in the previous step. It is
+worth noting that any subset of the original set of options captures
+the malicious traffic. \tname{} leverages that observation to search
+for rules containing a proper subset of rules defined in the original
+seed option. As the generation procedure discards options from the
+initial rule, the rules \tname{} creates satisfies the property of
+capturing malicious traffic. However, the goal of \tname\ is to
+produce rules that capture \emph{only} malicious traffic. To achieve
+that goal, \tname{} uses benign traffic to reduce the likelihood the
+\nids\ will report false positives on the synthesized rule. The
+minimization step uses a population-based gradient descent-like search
+to find all plausible rules derived from the seed rule, obtained in
+the previous step. Third, \tname{} uses several heuristic functions
+derived from a dataset of rules to rank the rules generated in the
+previous step.
 
@@
 inferring options at the field granularity level is insufficient to
-capture \CodeIn{content} options as the described on
+capture content options as those described on
 Section~\ref{sec:content-example}. These options refer to parts of the
 payload of the message and cannot be extracted directly from the
-fields. Handling this case is critically important as 93\% of
-the rules contain \CodeIn{content} options.
+fields. Handling this case is critically important as 93\% of the
+rules contain \CodeIn{content} options.
 
@@
 
-\Gui{
-O procedimento que já tínhamos representa a etapa de treinamento.
+%% \Gui{
+%% O procedimento que já tínhamos representa a etapa de treinamento.
 Toda aquela parte que foi explicada na apresentação para os outros professores e descrito no paper.
@@
 
-TREINAMENTO
-
-->Syrius recebe como input um pcap contendo o ataque (no caso de ataques com content, usamos ataques do Nikto), um pcap benígno (usamos o bigflows.pcap) e um pcap contendo as variações de ataque (isso pode ser otimizado unificando com o pcap de ataque, mas por agora está assim)
-->É gerado a regra máxima através do tokenizer e engenharia reversa.
-->Syrius remove UM content de uma regra da lista de regras atuais (inicialmente apenas a regra máxima) e checa se a regra criada já existe na lista.
+%% TREINAMENTO
+
+%% ->Syrius recebe como input um pcap contendo o ataque (no caso de ataques com content, usamos ataques do Nikto), um pcap benígno (usamos o bigflows.pcap) e um pcap contendo as variações de ataque (isso pode ser otimizado unificando com o pcap de ataque, mas por agora está assim)
+%% ->É gerado a regra máxima através do tokenizer e engenharia reversa.
+%% ->Syrius remove UM content de uma regra da lista de regras atuais (inicialmente apenas a regra máxima) e checa se a regra criada já existe na lista.
 Se não, a regra é adicionada à lista de regras-não-testadas.
 Syrius faz isso duas vezes por regra, para todas as regras da lista.
-->A lista de regras-não-testadas é inicializada no suricata e então o rodamos com o pcap benígno de treinamento (75porcento do total).
+%% ->A lista de regras-não-testadas é inicializada no suricata e então o rodamos com o pcap benígno de treinamento (75porcento do total).
 Verificando no arquivo de log do suricata, caso a regra alerte, ela é removida da lista.
 As regras restantes vão para a lista de regras atuais.
-->Quando não houver alterações, retornamos a lista com todas as regras que deram certo.
-->Opcionalmente, ainda rodamos no final uma variação de ataque (no momento apenas um pacote devido ao número limitado que temos do ataque).
+%% ->Quando não houver alterações, retornamos a lista com todas as regras que deram certo.
+%% ->Opcionalmente, ainda rodamos no final uma variação de ataque (no momento apenas um pacote devido ao número limitado que temos do ataque).
 Ao contrário do método com pacotes benígnos, se a regra NÃO alertar, ela é removida.
-->Calculamos o fitness de cada regra e ordenamos de acordo com o fitness resultante. Esta lista ordenada é o fim do treinamento.
-
-TESTE
-
-->Inicializamos a lista de regras ordenadas no suricata e rodamos o pcap benígno de teste (25porcento restantes) para cálculo do precision.
+%% ->Calculamos o fitness de cada regra e ordenamos de acordo com o fitness resultante. Esta lista ordenada é o fim do treinamento.
+
+%% TESTE
+
+%% ->Inicializamos a lista de regras ordenadas no suricata e rodamos o pcap benígno de teste (25porcento restantes) para cálculo do precision.
 O cálculo ficou basicamente, por regra: [100 - (numero-de-alertas/total-de-pacotes)]
 
-->Inicializamos a lista de regras ordenadas no suricata e rodamos o pcap com variações de ataque (caso a etapa opcional ocorrer, não usamos aquele pacote especifico) para cálculo do recall.
+%% ->Inicializamos a lista de regras ordenadas no suricata e rodamos o pcap com variações de ataque (caso a etapa opcional ocorrer, não usamos aquele pacote especifico) para cálculo do recall.
 O cálculo ficou basicamente, por regra: [(100/total-de-pacotes) * n-alertas].
 
-->Por final calculamos o F1 Score por regra: [2*((precision*recall)/(precision+recall))]
-
-->O teste retorna um csv contendo a lista de regras ordenadas com seus respectivos precision, recall e f1 score.
-}
+%% ->Por final calculamos o F1 Score por regra: [2*((precision*recall)/(precision+recall))]
+
+%% ->O teste retorna um csv contendo a lista de regras ordenadas com seus respectivos precision, recall e f1 score.
+%% }
 

### 2020-01-17 paper/paper.tex
merging

@@
 
+\begin{figure}[h]
+  \vspace{-2ex}
+  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{top-pingscan.rules}
+  \vspace{-3ex}  
+\end{figure}
+
+\Luc{Apenas 3 regras tiveram f1 de 100\%. O total de regras nesse caso foi de 13.}
 \Fix{...listem 5 primeiras regras neste caso...}
 
-\subsection{\Fix{JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit}}
+\subsection{Remote Code Execution}
+\label{sec:rce}
 \label{sec:jboss}
 \label{sec:content-example}
+
+\Luc{"RCE refers to the process by which an agent can exploit a network vulnerability to run arbitrary code on a targeted machine or system.
+In an RCE attack, hackers intentionally exploit a remote code execution vulnerability to run malware.
+By prompting the targeted device to perform code execution, a hacker can run their own programming in its place.
+This programming can then enable them to gain full access, steal data, carry out a full distributed denial of service (DDoS) attack, destroy files and infrastructure, or engage in illegal activity.
+And as the term remote execution suggests, an RCE cyberattack can take place from any geophysical location."\\ Texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution \\
 

### 2020-01-17 paper/paper.tex
fix

@@
 This programming can then enable them to gain full access, steal data, carry out a full distributed denial of service (DDoS) attack, destroy files and infrastructure, or engage in illegal activity.
-And as the term remote execution suggests, an RCE cyberattack can take place from any geophysical location."\\ Texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution \\
+And as the term remote execution suggests, an RCE cyberattack can take place from any geophysical location."\\ Texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution} \\
 

### 2020-01-18 paper/paper.tex
pass on eval

@@
 
-This section evaluates the precision and efficiency of \tname{}.
-
-\subsection{Objects of Analysis}
-\label{sec:dataset-benign}
-
-Table~\ref{table:attacks} shows the attacks we considered in our
-evaluation. These attacks cover different protocols and features of
-\tname{}. Column ``Name'' shows the name of the attack, column
-``Protocol'' shows the protocol name, and column ``Ground Truth''
-shows the options of the rule that captures the attack.
-
-\vspace{1ex}
-\noindent
-\textbf{Malicious traffic.}~Nikto?\Fix{where do these
-  attacks/messages come from?}
-  \Luc{Nós executamos um servidor http (Apache2) numa maquina e executamos o Nikto em outra maquina atacando o servidor http. Capturamos o trafego q chegava ao servidor e eh ele q estamos utilizando.}
-
-\vspace{1ex}
-\noindent
-\textbf{Benign traffic.}~\tname{} uses public databases storing benign
-traffic for all protocols covered by Suricata (see
-Figure~\ref{fig:distribution-rules-protocol}). We used two datasets of
-positive traffic. The Bigflows.pcap dataset is part of the
-TcpReplay~\cite{tcpreplay} open source project. This dataset includes
-real network traffic on a busy private network access point to the
-Internet. According to the TcpReplay web site ``This capture is much
-larger and has a smaller average packet size than the previous capture
-(smallFlows.pcap). It also has many more flows and different
-applications.''. In addition to the Bigflows.pcap dataset, we also
-used datasets of normal traffic made available by Stratosphere
-Lab~\cite{stratosphere-normal}.
-
-\vspace{1ex}
-\noindent
-\textbf{Ground Truth.}~The rules that detect those attacks were
-obtained from the Emerging Threats Open
-Ruleset~\cite{emerging-threats-open}.
-
-%% Note from Figure~\ref{fig:overview} that the positive traffic is an
-%% input to the technique. The negative non-malicious traffic, however,
-%% is not.
-
-\begin{table*}[h]
+\begin{table*}[h!]
   \caption{\label{table:attacks}List of attacks.\Mar{Guilherme, por
@@
 
+This section reports on the evaluation of \tname{}.
+
+\subsection{Objects of Analysis}
+\label{sec:dataset-benign}
+
+Table~\ref{table:attacks} shows the attacks we considered in our
+evaluation. These attacks cover different protocols and features of
+\tname{}. Column ``Name'' shows the name of the attack, column
+``Protocol'' shows the protocol name, and column ``Ground Truth''
+shows the options of the rule that captures the attack.
+
+\subsubsection{Malicious traffic.}To reproduce the malicious traffic, we
+configured an HTTP server (Apache2) in one machine and ran
+Nikto2~\cite{nikto} in a different machine. We used this setup to
+replay each of the attacks from Table~\ref{table:attacks}.  On each
+execution, we captured the incoming traffic to the web server with
+Wireshark~\cite{wireshark-net-monitor} and used the corresponding pcap
+files as an input to \tname.\Mar{@Lucas, eu traduzi o que vc. falou
+  para ingles. Mas, aparentemente, isto esta **inconsistente** com o
+  que eh dito na Secao~\ref{attack-reproduction}. Por favor, esclareca!}
+
+\subsubsection{Benign traffic.}\tname{} uses public databases storing benign
+traffic for all protocols covered by Suricata (see
+Figure~\ref{fig:distribution-rules-protocol}). We used two datasets of
+benign traffic. The Bigflows.pcap dataset is part of the
+TcpReplay~\cite{tcpreplay} open source project. This dataset includes
+real network traffic of a busy private network access point to the
+Internet. According to the TcpReplay web site ``This capture is much
+larger and has a smaller average packet size than the previous capture
+(smallFlows.pcap). It also has many more flows and different
+applications.''. In addition to the Bigflows.pcap dataset, we also
+used datasets of normal traffic made available by Stratosphere
+Lab~\cite{stratosphere-normal}.
+
+\subsubsection{Ground Truth.}To evaluate precision of \tname\ we
+need rules that are considered correct to use as reference. The rules
+that detect the attacks from Table~\ref{table:attacks} were obtained
+from the Emerging Threats Open Ruleset~\cite{emerging-threats-open}.
+
+%% Note from Figure~\ref{fig:overview} that the positive traffic is an
+%% input to the technique. The negative non-malicious traffic, however,
+%% is not.
+
 \subsection{Methodology}
 
-\subsubsection{Attack reproduction.} \tname{} requires the attack to
+\subsubsection{Attack reproduction.}\label{attack-reproduction} \tname{} requires the attack to
 be replayed in order to check whether or not the rule it generates
@@
 
-\subsection{Precision}
-
-\Fix{}
-
-\subsection{Efficiency}
+\subsection{Results}
 

### 2020-01-18 paper/paper.tex
pass on 4.3.4

@@
 \subsubsection{Frequency of content modifiers.} As previously
-mentioned, \percRulesWithContent\ of Suricata rules involve content
-options. Furthermore, we observed---by analyzing the dataset of
-existing rules (see Section~\ref{sec:dataset-benign})---that often
-these rules indicate what part of the message the content string
-should be checked against.\Mar{Lucas, vc. tem dados para dar suporte a
-  isto? ou seja, quantas options do tipo ``content'' possuem
-  modificadores.} \Luc{das 26k regras que utilizam content, 13k delas utilizam modificadores http q sao os q nós consideramos.} For example, the
-creator of the rule may want to indicate that a specific string be
-present in the header of an HTTP request message. Content
-\emph{modifiers} serve this purpose. For example, the modifier
+mentioned, \percRulesWithContent\ (=26K/\numrulessuri) of Suricata
+rules involve content options. Furthermore, of the 26K rules that use
+content options, 13K use HTTP modifiers\Mar{@Lucas, pode dar o numero
+  desconsiderando HTTP? Esta parte eh generica.}. Content modifiers
+serve to indicate what part of a network packet the string mentioned
+in the content option should appear. For example, the modifier
 \CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
@@
 for multiple HTTP requests/responses.)
-Figure~\ref{fig:http-header-example}
-
-\begin{figure}[h!]
+Figure~\ref{fig:http-header-example} illustrates possible content
+options used to match against certain strings that appear in an HTTP
+message. A check (resp., cross) mark indicates a match (resp.,
+unmatch).
+
+\begin{figure}[t!]
 \centering
@@
 
+%% we observed---by analyzing
+%% the dataset of existing rules (see
+%% Section~\ref{sec:dataset-benign})---that these rules often indicate
+%% what part of the message the content string should be checked
+%% against. For example,
+%% For example, the
+%% creator of the rule may want to indicate that a specific string be
+%% present in the header of an HTTP request message. Content
+%% \emph{modifiers} serve this purpose.
 %% Suricata supports content modifiers for tcp and http.
@@
 
-\tname\ uses the frequency of content modifiers to discriminate rules
-more or less likely to reflect actual rules. For example, historical
-data indicates that content options that use the
-\CodeIn{http\_connection} modifier are much less likely than content
-options that use the modifier
+\tname{} leverages the fact that these modifiers are commonly used to
+define rules as to better rank the rules it synthesizes. More
+precisely, it builds on the observation that strings acessible in
+positions associated with content modifiers (\eg{}, GET, Host,
+Connection) are more likely to produce useful rules. \tname\ uses the
+frequency of content modifiers to discriminate rules that are more or
+less likely to reflect actual rules. For example, historical data
+indicates that content options that use the \CodeIn{http\_connection}
+modifier are much less likely than content options that use the
+modifier
 \CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}

### 2020-01-18 paper/paper.tex
explain RCE

@@
   \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
-  \caption{Suricata rule for SYN Flood Attacks.}
+  \caption{Suricata rule for SYN Flood attack.}
   \label{fig:synflood-example}
@@
   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
-  \caption{Suricata rule for Ping Scan.}
+  \caption{Suricata rule for Ping Scan attack.}
   \label{fig:pingscan-example}
@@
 
-\subsection{Remote Code Execution}
+\subsection{Remote Code Execution (RCE)}
 \label{sec:rce}
@@
 
-\Luc{"RCE refers to the process by which an agent can exploit a network vulnerability to run arbitrary code on a targeted machine or system.
-In an RCE attack, hackers intentionally exploit a remote code execution vulnerability to run malware.
-By prompting the targeted device to perform code execution, a hacker can run their own programming in its place.
-This programming can then enable them to gain full access, steal data, carry out a full distributed denial of service (DDoS) attack, destroy files and infrastructure, or engage in illegal activity.
-And as the term remote execution suggests, an RCE cyberattack can take place from any geophysical location."\\ Texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution} \\
-
-\Mar{Note que a gente organizou estas secoes por tipo de ataque. Este
-  titulo eh muito especifico.}
-\Luc{JBoss é um servidor de aplicação Java muito popular, de codigo aberto.
-Esse ataque tem como alvo esses servidores e permite que seja feita execução remota de comandos.
-Os pacotes que estamos utilizando para testar esse ataque foram gerados pelo Nikto, mas o Metasploit tambem possui um modulo para ele.
-A regra de ouro para esse ataque é:
+%% Adaptacao do texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution
+
+RCE refers to the family of attacks where hackers exploit a network
+vulnerability to run arbitrary code---that could damage the system or
+steal sensitive information---on a remote machine. JBoss~\cite{jboss}
+is popular open source Java application server. This attacks enables
+remote execution of code on the JBoss application server.
+\Mar{@Lucas, por favor, observe o padrao usado nos exemplos
+  anteriores. eh preciso explicar em alto nivel o ataque antes de
+  explicar a regra!}
+
+%% In an RCE attack, hackers
+%% intentionally exploit a vulnerability in the system to run malware
+%% By prompting the targeted device to perform code execution, a hacker
+%% can run arbritrary code to
+%% their own programming in its
+%% place. This programming can then enable them to gain full access,
+%% steal data, carry out a full distributed denial of service (DDoS)
+%% attack, destroy files and infrastructure, or engage in illegal
+%% activity. And as the term remote execution suggests, an RCE
+%% cyberattack can take place from any geophysical location."\\
 
 \begin{figure}[H]
-  \lstinputlisting[language=C,numbers=none,keywords={content}]{adaptor-golden-rule.suricata}
-  \label{fig:rule-example}
+  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
+  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code Execution.}
 \end{figure}
 
-A opção flow:established,to\_server; define que o match só será feito em pacotes que fazem parte de uma conexão já estabelecida entre um cliente e um servidor.
-Além disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.
-Nos contents os modificadores nocase e http\_uri indicam que a detecção da string não será case sensitive e que essa string deverá se encontrar no campo de request uri.}
-
-\Luc{... sobre comportamento do syrius ...}
-
-\Fix{...please elaborate. We need on example that shows that the
-  search can be non-trivial. In the examples above, we had to discard
-  only one rule option!}
+%A opção  define que o match só será feito em pacotes que fazem parte
+%de uma conexão já estabelecida entre um cliente e um servidor. Além
+%disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.
+
+Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
+JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
+indicates that a rule match will only occur on packets that (1) flow
+on an already-established connection and that (2) flow to the server
+as opposed to the client. For the content options the modifiers
+\CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
+the matching not be case sensitive and that the string should be
+located at the packet field ``Request URI''. \Mar{Lucas, ``nocase''
+  tambem eh considerado modificador? Isto nao eh confuso? Na secao
+  4.3.4 consideramos modificadores como sinonimo para qualificador de
+  campo, nao?}
+
+\tname{} reproduces the malicious JBoss attack using the traffic from
+a Nikto script~\cite{nikto}. However, similar JBoss attack is also
+available from the Metasploit database~\cite{metasploit}. As in the
+other examples, \tname{} uses, as additional inputs, benign traffic
+from \Fix{...} and a set of existing Suricata rules from
+\Fix{...}. \tname{} initially produces one seed rule containing
+\Fix{xx} options, by analyzing the malicious traffic alone. Then, it
+uses that seed rule to derive \Fix{yy} rules, each rule containing
+subsets of the option in the initial rule. Finally, it ranks that
+list. The list below shows the top-5 rules that \tname{}
+produces.\Mar{<- check/revise}
+
+\Fix{list rules}
+
+Note that the golden rule is in position \Fix{...} on this
+case. \Mar{explain syntactical difference, if exist}
 

### 2020-01-19 paper/paper.tex
pass

@@
   4.3.4 consideramos modificadores como sinonimo para qualificador de
-  campo, nao?}
+  campo, nao?}\Luc{Sim, ``nocase'' define que a busca pela string nao deve ser case sensitive. Em 4.3.4 estamos considerando apenas os modificadores que definem posições/campos de pacotes http.}
 
@@
 content options, 13K use HTTP modifiers\Mar{@Lucas, pode dar o numero
-  desconsiderando HTTP? Esta parte eh generica.}. Content modifiers
+  desconsiderando HTTP? Esta parte eh generica.}. \Luc{Posso, mas nessa função nós consideramos apenas os modificadores http.} Content modifiers
 serve to indicate what part of a network packet the string mentioned
@@
       active hosts in a network} & \scriptsize{dsize:0; itype:8;} \\
-    Black Nurse \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;}\\
-    Ping Flood  \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;}\\  
-    Port Scan \MyComment{& single} & TCP \MyComment{& Discovers open ports in a host} & \scriptsize{flags: A; ack: 0;dsize: 0;} \\
+    %%Black Nurse \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;}\\
+    %%Ping Flood  \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;}\\  
+    %%Port Scan \MyComment{& single} & TCP \MyComment{& Discovers open ports in a host} & \scriptsize{flags: A; ack: 0;dsize: 0;} \\
     SYN Flood \MyComment{& multiple} & TCP \MyComment{& Denial of Service} & \scriptsize{flags: S,12; threshold: type both, track by\_dst, count 5000, seconds 5;}\\
-    UDP Flood \MyComment{& multiple} & UDP \MyComment{& Denial of Service} & \scriptsize{fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5;} \\
+    %%UDP Flood \MyComment{& multiple} & UDP \MyComment{& Denial of Service} & \scriptsize{fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5;} \\
+    JBoss JMX Console Beanshell Exploit \MyComment{& single} & HTTP \MyComment{& Remote Code Execution} & \scriptsize{content:"/HtmlAdaptor"; nocase; http\_uri; content:"action=inspect"; nocase; http\_uri; content:"bean"; nocase; http\_uri; content:"name="; http\_uri;} \\
+    WeBid Local File Inclusion \MyComment{& single} & TCP \MyComment{& Local File Inclusion} & \scriptsize{content:"GET "; depth:4; content:"/cron.php?"; http\_uri; nocase; content:"include\_path="; http\_uri; nocase; content:"../";} \\
+    ColdFusion Privilege Escalation \MyComment{& single} & HTTP \MyComment{& Privilege Escalation} & \scriptsize{content:"GET"; http\_method; nocase; content:"/CFIDE/administrator"; http\_uri; nocase;} \\
+    \Luc{Adicionar nome} \MyComment{& single} & HTTP \MyComment{& \Luc{Adicionar}} & \scriptsize{content:".htaccess"; nocase; http\_uri;} \\
     \bottomrule
@@
   que eh dito na Secao~\ref{attack-reproduction}. Por favor, esclareca!}
+  \Luc{Essa descrição é referente aos ataques HTTP/TCP da tabela.
+O hping3 gera apenas o trafego do SYN Flood.
+Ambas as explicações estão corretas.
+Creio q a descrição da seção 5.2.1 tenha q ser movida para cá pois está explicando como capturamos o trafego malicioso.
+Nós executamos o Nikto e o hping3 uma unica vez para capturar com o wireshark o trafego q eles geram.
+O trafego produzido pelo Nikto ja inclui todos os ataques que utilizamos, nós separamos os pacotes de cada um deles manualmente utilizando o wireshark.}
 

### 2020-01-19 paper/paper.tex
minor changes on format tables

@@
 
+    %%Black Nurse \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;}\\
+    %%Ping Flood  \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;}\\  
+    %%Port Scan \MyComment{& single} & TCP \MyComment{& Discovers open ports in a host} & \scriptsize{flags: A; ack: 0;dsize: 0;} \\
+
+    %%UDP Flood \MyComment{& multiple} & UDP \MyComment{& Denial of Service} & \scriptsize{fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5;} \\
+
 \begin{table*}[h!]
-  \caption{\label{table:attacks}List of attacks.\Mar{Guilherme, por
-      favor, adicione mais ataques aqui...}}
+  \small
+  \caption{\label{table:attacks}List of attacks.}
   \centering
-  \begin{tabular}{lll}
+  \begin{tabular}{lllp{11cm}}
     \toprule
     \multicolumn{1}{c}{Name} &
+    \multicolumn{1}{c}{Category} &
     \multicolumn{1}{c}{Protocol} &
@@
     \midrule     
-    Ping Scan \MyComment{& single} & ICMP \MyComment{& Discovers
-      active hosts in a network} & \scriptsize{dsize:0; itype:8;} \\
-    %%Black Nurse \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;}\\
-    %%Ping Flood  \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;}\\  
-    %%Port Scan \MyComment{& single} & TCP \MyComment{& Discovers open ports in a host} & \scriptsize{flags: A; ack: 0;dsize: 0;} \\
-    SYN Flood \MyComment{& multiple} & TCP \MyComment{& Denial of Service} & \scriptsize{flags: S,12; threshold: type both, track by\_dst, count 5000, seconds 5;}\\
-    %%UDP Flood \MyComment{& multiple} & UDP \MyComment{& Denial of Service} & \scriptsize{fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5;} \\
-    JBoss JMX Console Beanshell Exploit \MyComment{& single} & HTTP \MyComment{& Remote Code Execution} & \scriptsize{content:"/HtmlAdaptor"; nocase; http\_uri; content:"action=inspect"; nocase; http\_uri; content:"bean"; nocase; http\_uri; content:"name="; http\_uri;} \\
-    WeBid Local File Inclusion \MyComment{& single} & TCP \MyComment{& Local File Inclusion} & \scriptsize{content:"GET "; depth:4; content:"/cron.php?"; http\_uri; nocase; content:"include\_path="; http\_uri; nocase; content:"../";} \\
-    ColdFusion Privilege Escalation \MyComment{& single} & HTTP \MyComment{& Privilege Escalation} & \scriptsize{content:"GET"; http\_method; nocase; content:"/CFIDE/administrator"; http\_uri; nocase;} \\
-    \Luc{Adicionar nome} \MyComment{& single} & HTTP \MyComment{& \Luc{Adicionar}} & \scriptsize{content:".htaccess"; nocase; http\_uri;} \\
+    Ping Scan & Active Recon. & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
+    SYN Flood & DoS & TCP & \textbf{flags}: S,12; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\
+
+    JBoss JMX\MyComment{ Console Beanshell Exploit} & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
+    WeBid \MyComment{Local File Inclusion} & - & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
+    ColdFusion \MyComment{Privilege Escalation} & - & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
+    - & - & HTTP \MyComment{& \Luc{Adicionar}} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
     \bottomrule
@@
 \end{table*}
+
+\begin{table}[h!]
+  \small
+  \caption{\label{table:results}Results.}
+  \centering
+  \begin{tabular}{lllp{9cm}}
+    \toprule
+    \multicolumn{1}{c}{Name} &
+    \multicolumn{1}{c}{\#~Rules created} &
+    \multicolumn{1}{c}{Rank} \\
+    \midrule
+    Ping Scan & - & - \\
+    SYN Flood & - & - \\
+    JBoss JMX & - & - \\
+    WeBid & - & - \\    
+    ColdFusion & - & - \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
   que eh dito na Secao~\ref{attack-reproduction}. Por favor, esclareca!}
-  \Luc{Essa descrição é referente aos ataques HTTP/TCP da tabela.
-O hping3 gera apenas o trafego do SYN Flood.
-Ambas as explicações estão corretas.
-Creio q a descrição da seção 5.2.1 tenha q ser movida para cá pois está explicando como capturamos o trafego malicioso.
-Nós executamos o Nikto e o hping3 uma unica vez para capturar com o wireshark o trafego q eles geram.
-O trafego produzido pelo Nikto ja inclui todos os ataques que utilizamos, nós separamos os pacotes de cada um deles manualmente utilizando o wireshark.}
+  \Luc{Essa descrição é referente aos ataques HTTP/TCP da tabela. O
+    hping3 gera apenas o trafego do SYN Flood. Ambas as explicações
+    estão corretas. Creio q a descrição da seção 5.2.1 tenha q ser
+    movida para cá pois está explicando como capturamos o trafego
+    malicioso. Nós executamos o Nikto e o hping3 uma unica vez para
+    capturar com o wireshark o trafego q eles geram. O trafego
+    produzido pelo Nikto ja inclui todos os ataques que utilizamos,
+    nós separamos os pacotes de cada um deles manualmente utilizando o
+    wireshark.}\Mar{Fiquei na duvida se o hping soh serve para o SYN
+    Flood e se/porque o Nikto nao eh usado neste caso...Enfim, nao
+    precisa responder por aqui. Mova o texto deixando-o *consistente*.}
 
@@
 
+\Mar{Explique resumidamente o que cada ataque listado na tabela 2 faz...}
+
 \subsection{Threats to Validity}

### 2020-01-19 paper/paper.tex
top 5 synflood rules

@@
 in false negatives. For example, in this particular case, note that
-the option \CodeIn{windows: 64} is not included in the golden rule
+the option \CodeIn{windows: 512} is not included in the golden rule
 from Figure~\ref{fig:synflood-example}.
@@
 
-\Mar{listem 5 primeira regras neste caso}
+\begin{figure}[h]
+  \vspace{-2ex}
+  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window,icmp\_id,icmp\_seq}]{synflood-top5-rules.suricata}
+  \vspace{-3ex}  
+\end{figure}
 

### 2020-01-19 paper/paper.tex
reverting last commit
update table 2 and 3

@@
 
-\begin{figure}[h]
-  \vspace{-2ex}
-  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window,icmp\_id,icmp\_seq}]{synflood-top5-rules.suricata}
-  \vspace{-3ex}  
-\end{figure}
+\Mar{listem 5 primeira regras neste caso}
 
@@
     ColdFusion \MyComment{Privilege Escalation} & - & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
-    - & - & HTTP \MyComment{& \Luc{Adicionar}} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
+    .htaccess access\MyComment{attempted-recon} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
+    .idq access\MyComment{EXPLOIT ISAPI} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
+    iisadmin access\MyComment{web-application-attack} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
+    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
+    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
     \bottomrule
@@
     SYN Flood & - & - \\
-    JBoss JMX & - & - \\
+    JBoss JMX & 283 & 19 \\
     WeBid & - & - \\    
-    ColdFusion & - & - \\
+    ColdFusion & 566 & 37 \\
+    .htaccess access & 572 & 1 \\
+    .idq access & 700 & 136 \\
+    iisadmin access & 502 & 117 \\
+    /system32/ in URI & 1060 & 191 \\
+    Script tag in URI & 1332 & 304 \\
     \bottomrule

### 2020-01-19 paper/paper.tex
minor formatting changes

@@
   \centering
-  \begin{tabular}{lllp{9cm}}
+  \begin{tabular}{lrrp{9cm}}
     \toprule
@@
     iisadmin access & 502 & 117 \\
-    /system32/ in URI & 1060 & 191 \\
-    Script tag in URI & 1332 & 304 \\
+    /system32/ in URI & 1,060 & 191 \\
+    Script tag in URI & 1,332 & 304 \\
     \bottomrule

### 2020-01-20 paper/paper.tex
comment on 4.3 and table 4

@@
 
+\Gui{Fizemos um teste para encontrar melhores pesos para cada ataque.
+Calculamos o fitness de cada regra atribuindo diferentes pesos para cada função heurística.
+Os valores atribuídos foram 0, 0.25, 0.5, 0.75 e 1.
+Depois de calculado, ordenamos todas as listas criadas e pegamos a posição da Golden Rule.
+As melhores combinações de pesos (que conseguiram melhor posição da golden rule) estão na tabela 4, onde na coluna "Weights", há quatro números separados por vírgula, cada um deles representando o peso de uma função (na ordem que é representado no paper.
+O total de regras criadas é igual ao da Tabela 3.}
+
 %% \section{Modificadores}É possivel deixar que uma regra verifique todo o payload do pacote em busca de um match com um content ou especificar partes do payload onde esse content deve ser procurado.
@@
 
+\begin{table}[h!]
+  \small
+  \caption{\label{table:weight}Weight Combinations.}
+  \centering
+  \begin{tabular}{lrrp{15cm}}
+    \toprule
+    \multicolumn{1}{c}{Name} &
+    \multicolumn{1}{c}{Weights} &
+    \multicolumn{1}{c}{Rank} \\
+    \midrule
+    Ping Scan & - & - \\
+    SYN Flood & - & - \\
+    JBoss JMX & 1,0,1,1 & 107 \\
+    WeBid & - & - \\    
+    ColdFusion & 0,1,1,1 & 7 \\
+    .htaccess access & 0,1,1,1 & 11 \\
+    .idq access & 0,1,1,1 & 7 \\
+    iisadmin access & 0,1,1,1 & 2 \\
+    /system32/ in URI & 0,1,1,1 & 11 \\
+    Script tag in URI & 0,1,1,1 & 9 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
 This section reports on the evaluation of \tname{}.

### 2020-02-08 paper/paper.tex
quick pass on abstract and intro

@@
 %%
-%% The first command in your LaTeX source must be the \documentclass command.
+%% The first command in your LaTeX source must be the \documentclass
+%% command.
 \documentclass[sigconf,review, anonymous]{acmart}
-\acmConference[ISSTA 2020]{ACM SIGSOFT International Symposium on Software Testing and Analysis}{18–22 July, 2020}{Los Angeles, US}
+\acmConference[ESEC/FSE 2020]{The 28th ACM Joint European Software Engineering Conference and Symposium on the Foundations of Software Engineering}{8 - 13 November, 2020}{Sacramento, California, United States}
 
@@
 systems monitor the network traffic and flag suspicious network
-behavior. Signature-based \nids\ do that by checking the network
-traffic against a pre-defined set of rules, which can become obsolete
+behavior. Signature-based \nids\ do that by checking the\MyComment{ network}
+traffic against a pre-defined set of rules, which become obsolete
 as attackers learn new strategies to circumvent existing defenses.
+
 This paper proposes \tname{}, a technique that automatically
@@
 malicious and benign traffic. \tname{} formulates synthesis as an
-optimization problem whose candidate solutions are rules that maximize
-the capture of positive traffic and minimize the capture of negative
-traffic. \tname{} bootstraps the search with candidate solutions that
-use data from the payload of the malicious message and produces
-optimal rule candidates on output. We evaluated \tname{} on a diverse
+optimization problem whose candidate solutions are rules. Optimal
+solutions maximize the capture of positive traffic and minimize the
+capture of negative traffic. \tname{} bootstraps the search with a
+candidate solution inferred from the payload of the malicious message.
+It uses that initial overfitting solution to search for alternative
+rules that only capture the malicious traffic. Then, it ranks those
+rules according to heuristics---drawn from a data set of existing
+rules---denoting the likelihood of those rules of being useful.
+
+We evaluated \tname{} on a diverse
 set of attacks. Results indicate that \Fix{...}

### 2020-02-18 simulated-annealing/sa.py
rule options fitness

@@ -9,7 +9,7 @@ import binascii
 import argparse
 import csv
 import math
-from functools import partial
+from functools import partial, reduce
 from itertools import combinations
 
 open("bad.rules", 'w').close()
@@ -426,8 +426,9 @@ for content in sd:
         break    
 
 
-#print("keyword_freq:", keywords_freq)
-#print()
+print("keyword_freq:", keywords_freq)
+#exit()
+
 html_modifiers_freq = {}
 def getHtmlModifiersFreq(keywords_freq):
     global html_modifiers
@@ -465,7 +466,7 @@ if rule_protocol == "http":
     lower_case_pkt_content_modifiers = dict((k, v.lower()) for k,v in pkt_content_modifiers.items())
     print()
 
-max_fitness = [0,0,0,0]
+max_fitness = [0,0,0,0,0]
 
 def ruleSizeFitness(rule):
     global rules_per_size
@@ -558,6 +559,21 @@ def ruleContentsModifiersFitness(rule):
     
     return fit_aux/len(rule_contents)
 
+def ruleOptionsFitness(rule):
+    global keywords_freq
+    count = 0
+    fitness = 0
+
+    for option in rule.options:
+        count += 1
+        fitness += keywords_freq[option]/max(list(keywords_freq.values()))
+    
+    if count > 0:
+        fitness = fitness/count
+
+    return fitness
+
+
 def writeRuleOnFile(rules):
     global ruleFile_path
     open(ruleFile_path, 'w').close()
@@ -565,13 +581,13 @@ def writeRuleOnFile(rules):
     ruleFile.seek(0)
     ruleFile.truncate()
     for rule in rules:
-        ruleFile.write(str(rule) + "\n" + "\n")
+        ruleFile.write(str(rule) + "\n")
     ruleFile.close()
-    time.sleep(0.050)
+    #time.sleep(0.050)
 
 def sendGoodTraffic(attack):
     subprocess.Popen(["sh", "sendGoodTraffic.sh", attack], stdout=subprocess.DEVNULL).wait()
-    time.sleep(0.05)
+    #time.sleep(0.05)
 
 def isEmpty(fpath):
     result=False
@@ -585,12 +601,12 @@ def isEmpty(fpath):
 
 def sendAttackVariation(attack):
     subprocess.Popen(["sh", "sendAttackVariation.sh", attack], stdout=subprocess.DEVNULL).wait()
-    time.sleep(0.5)
+    #time.sleep(0.5)
 
[…]

### 2020-02-18 paper/paper.tex
fix table

--- 2020-01-20 paper/paper.tex
@@
     SYN Flood & - & - \\
-    JBoss JMX & 283 & 19 \\
+    JBoss JMX & 283 & 219 \\
     WeBid & - & - \\    
-    ColdFusion & 566 & 37 \\
-    .htaccess access & 572 & 1 \\
-    .idq access & 700 & 136 \\
-    iisadmin access & 502 & 117 \\
-    /system32/ in URI & 1,060 & 191 \\
-    Script tag in URI & 1,332 & 304 \\
+    ColdFusion & 566 & 11 \\
+    .htaccess access & 572 & 223 \\
+    .idq access & 700 & 235 \\
+    iisadmin access & 502 & 133 \\
+    /system32/ in URI & 1,060 & 368 \\
+    Script tag in URI & 1,332 & 287 \\
     \bottomrule

### 2020-02-18 simulated-annealing/sendAttackVariation.sh
-k none option

@@ -1,1 +1,1 @@
-suricata -c inputs/suricata-$1.yaml -r Datasets/all-$1.pcap
+suricata -c inputs/suricata-$1.yaml -r Datasets/all-$1.pcap -k none

### 2020-02-20 paper/paper.tex
update table 3

--- 2020-02-08 paper/paper.tex
@@
   \centering
-  \begin{tabular}{lrrp{9cm}}
+  \begin{tabular}{lrrrp{9cm}}
     \toprule
@@
     \multicolumn{1}{c}{\#~Rules created} &
-    \multicolumn{1}{c}{Rank} \\
+    \multicolumn{1}{c}{Rank} &
+    \multicolumn{1}{c}{Rank fitness global} \\
     \midrule
-    Ping Scan & - & - \\
-    SYN Flood & - & - \\
-    JBoss JMX & 283 & 19 \\
-    WeBid & - & - \\    
-    ColdFusion & 566 & 37 \\
-    .htaccess access & 572 & 1 \\
-    .idq access & 700 & 136 \\
-    iisadmin access & 502 & 117 \\
-    /system32/ in URI & 1,060 & 191 \\
-    Script tag in URI & 1,332 & 304 \\
+    Ping Scan & - & - & - \\
+    SYN Flood & - & - & - \\
+    JBoss JMX & 382 & 208 & 23 \\
+    WeBid & - & - & - \\    
+    ColdFusion & 34 & 5 & 5\\
+    .htaccess access & 2107 & 365 & 2\\
+    .idq access & 1112 & 429 & 2 \\
+    iisadmin access & 862 & 372 & 1 \\
+    /system32/ in URI & - & - & - \\
+    Script tag in URI & - & - & - \\
     \bottomrule
@@
 
-\begin{table}[h!]
-  \small
-  \caption{\label{table:weight}Weight Combinations.}
-  \centering
-  \begin{tabular}{lrrp{15cm}}
-    \toprule
-    \multicolumn{1}{c}{Name} &
-    \multicolumn{1}{c}{Weights} &
-    \multicolumn{1}{c}{Rank} \\
-    \midrule
-    Ping Scan & - & - \\
-    SYN Flood & - & - \\
-    JBoss JMX & 1,0,1,1 & 107 \\
-    WeBid & - & - \\    
-    ColdFusion & 0,1,1,1 & 7 \\
-    .htaccess access & 0,1,1,1 & 11 \\
-    .idq access & 0,1,1,1 & 7 \\
-    iisadmin access & 0,1,1,1 & 2 \\
-    /system32/ in URI & 0,1,1,1 & 11 \\
-    Script tag in URI & 0,1,1,1 & 9 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+%%\begin{table}[h!]
+%%  \small
+%%  \caption{\label{table:weight}Weight Combinations.}
+%%  \centering
+%%  \begin{tabular}{lrrp{15cm}}
+%%    \toprule
+%%    \multicolumn{1}{c}{Name} &
+%%    \multicolumn{1}{c}{Weights} &
+%%    \multicolumn{1}{c}{Rank} \\
+%%    \midrule
+%%    Ping Scan & - & - \\
+%%    SYN Flood & - & - \\
+%%    JBoss JMX & 1,0,1,1 & 107 \\
+%%    WeBid & - & - \\    
+%%    ColdFusion & 0,1,1,1 & 7 \\
+%%    .htaccess access & 0,1,1,1 & 11 \\
+%%    .idq access & 0,1,1,1 & 7 \\
+%%    iisadmin access & 0,1,1,1 & 2 \\
+%%    /system32/ in URI & 0,1,1,1 & 11 \\
+%%    Script tag in URI & 0,1,1,1 & 9 \\
+%%    \bottomrule
+%%  \end{tabular}
+%%\end{table}
+%%
 

### 2020-02-20 paper/paper.tex
pass on section 2

@@
 Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
-the Open Information Security Foundation (OISF)~\cite{oisf}. A
+the Open Information Security Foundation (OISF)~\cite{oisf}.
+
+\vspace{-1ex}
+\begin{figure}[h!]
+  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
+  \vspace{-2ex}
+  \caption{Suricata rule for SYN Flood attack.}
+  \label{fig:synflood-example}
+\end{figure}
+\vspace{-2ex}
+
+A
 Suricata rule is divided in three parts---action, header, and rule
@@
 word in the rule description. An action denotes the task that needs to
-be executed if the rule pattern is satisfied. In this example, a
+be executed if the rule pattern is satisfied.
+In this example, a
 message will be sent to system administrators if the rule pattern is
@@
 
-\begin{table}[t!]
+\vspace{-1ex}
+\begin{table}[h!]
+  \footnotesize
   \caption{\label{table:rules}Example options of Suricata.}
+  \vspace{-2ex}
   \centering
@@
 \end{table}
+%\vspace{-1ex}
 
@@
 noting that Suricata supports a total of \totoptions\ distinct options
-covering \numproto\ distinct protocols. A total of 9 of these options
-are for documentation.\MyComment{For instance, the purpose of rule
-  option \CodeIn{msg} is to print on the output an informative message
-  indicating what kind of intrusion has been observed.}
-Figure~\ref{fig:distribution-rules-protocol} shows the distributions
-of these options per protocol. The sum of the number of protocol
-options is higher than \totoptions\ as some options are shared across
-different protocols. \tname{} does not make distinction among protocols.
-
-%\begin{wrapfigure}[14]{r}{0.5\textwidth}
+covering \numproto\ distinct protocols.
+
+%% A total of 9 of these options
+%% are for documentation.\MyComment{For instance, the purpose of rule
+%%   option \CodeIn{msg} is to print on the output an informative message
+%%   indicating what kind of intrusion has been observed.}
+%% Figure~\ref{fig:distribution-rules-protocol} shows the distributions
+%% of these options per protocol. The sum of the number of protocol
+%% options is higher than \totoptions\ as some options are shared across
+%% different protocols. \tname{} does not make distinction among protocols.
+
+%% %\begin{wrapfigure}[14]{r}{0.5\textwidth}
+%% \begin{figure}[t!]
+%%   \centering
+%% %  \vspace{-7ex}
+%%   \scalebox{0.85}{
+%%     \begin{tikzpicture}
+%%       \begin{axis}[
+%%           ybar stacked,
+%%           enlargelimits=0.15,
+%%           legend style={at={(0.5,-0.15)},
+%%             anchor=north,legend columns=-1},
+%%           %          ylabel={\#rules},
+%%           symbolic x coords={dns, ftp, http, icmp, ip, smb, smtp, ssh,
+%%           tcp, tls, udp},
+%%           xtick=data,
+%%           xticklabel style={rotate=45},          
+%%           nodes near coords,
+%%           every node near coord/.append style={font=\scriptsize},        
+%%           nodes near coords align={vertical},
+%%         ]
+%%         \addplot coordinates {(dns,10) (ftp,11) (http,20) (icmp,18)(ip,16) (smb,13) (smtp,11) (ssh, 10) (tcp, 32) (tls,18) (udp, 18)};
+%% %        \legend{relevant, irrelevant}
+%%       \end{axis}
+%%     \end{tikzpicture}
+%%   }
+%%   \vspace{-2ex}
+%%   \caption{\label{fig:distribution-rules-protocol}Number of relevant
+%%     options per protocol supported by Suricata.\Mar{not sure if this
+%%       is really helpful.}}
+%% %\end{wrapfigure}
+%% \end{figure}
+
+\subsection{Rules and Packets}
+\label{sec:rules-and-packets}
+
 \begin{figure}[t!]
 \centering
-%  \vspace{-7ex}
-  \scalebox{0.85}{
-    \begin{tikzpicture}
-      \begin{axis}[
-          ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
-          %          ylabel={\#rules},
-          symbolic x coords={dns, ftp, http, icmp, ip, smb, smtp, ssh,
-          tcp, tls, udp},
-          xtick=data,
-          xticklabel style={rotate=45},          
-          nodes near coords,
-          every node near coord/.append style={font=\scriptsize},        
-          nodes near coords align={vertical},
-        ]
-        \addplot coordinates {(dns,10) (ftp,11) (http,20) (icmp,18)(ip,16) (smb,13) (smtp,11) (ssh, 10) (tcp, 32) (tls,18) (udp, 18)};
-%        \legend{relevant, irrelevant}
-      \end{axis}
-    \end{tikzpicture}
-  }
+\includegraphics[scale=0.28]{figs/ICMP-packet-structure.png}
 \vspace{-2ex}
-  \caption{\label{fig:distribution-rules-protocol}Number of relevant
-    options per protocol supported by Suricata.}
-%\end{wrapfigure}
-\end{figure}
-
-\subsection{Rules and Packets}
-\label{sec:rules-and-packets}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.28]{figs/ICMP-packet-structure.png}
 \caption{Layout of an ICMP packet.}
@@
 \end{figure}
+%\vspace{-2ex}
 
@@
 ``Code'', declared in the ICMP header, are mapped to the options
-\CodeIn{itype} and \CodeIn{icode}. (Table~\ref{table:rules} lists some
-example rule options.) The values of the options ``ttl'', ``itype'',
-and ``icode'' are the same as the value of the corresponding fields in
-the ICMP packet. For the field ``Flags'' the value of the
-corresponding option is encoded in the bitvector. For example, if the
-``Flags'' field stores the value 1010, the corresponding rule option
-will be \CodeIn{flagbits:RM}, as each position of the vector encodes a
-different character in the \CodeIn{flagbits} rule option. The value of
-several options can be determined through transformation functions, as
-those described above. One expection to this is the option
-\CodeIn{threshold}, which is typically used in denial-of-service
-attacks. That option operates over sequence of messages as opposed to
-a single message. Consequently, it is not possible to determined its
-values from a single message. More precisely, to trigger an action on
-a rule containing that option a \nids\ tool needs to monitor several
-messages.
+\CodeIn{itype} and \CodeIn{icode}. The values of the options ``ttl'',
+``itype'', and ``icode'' are the same as the value of the
+corresponding fields in the ICMP packet. For the field ``Flags'' the
+value of the corresponding option is encoded in a bitvector. For
+example, if the ``Flags'' field stores the value 1010, the
+corresponding rule option will be \CodeIn{flagbits:RM}, as each
+position of the vector encodes a different character in the
+\CodeIn{flagbits} rule option. The value of several options can be
+determined through transformation functions, as those described
+above. One exception to this is the option \CodeIn{threshold}, which
+is typically used in denial-of-service attacks. That option operates
+over sequence of messages as opposed to a single
+message. Consequently, it is not possible to determined its values
+from a single message.
+
+%% More precisely, to trigger an action on a rule
+%% containing that option a \nids\ tool needs to monitor several
+%% messages.
 
@@
 legitimate requests cannot be satisfied due to resource exhaustion.
-
-\begin{figure}[h!]
-  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
-  \caption{Suricata rule for SYN Flood attack.}
-  \label{fig:synflood-example}
-\end{figure}
 

### 2020-02-21 paper/paper.tex
updated values on sec 3.1

@@
 extracts options expressed in that traffic. In this case, the
-malicious traffic is characterized as a sequence of 20
+malicious traffic is characterized as a sequence of 5000
 messages. \tname{} finds the following set of options at this stage:
@@
 in false negatives. For example, in this particular case, note that
-the option \CodeIn{windows: 512} is not included in the golden rule
+the option \CodeIn{window: 512} is not included in the golden rule
 from Figure~\ref{fig:synflood-example}.
@@
 extracted from a database of existing rules. One example heuristic
-function we used to rank the generated rules was
+function we used to rank the generated rules was \Luc{a quantidade de opcoes na regra}
 \Fix{...elaborate...}. The list below shows the rule options for the
 top-5 rules \tname{} produces for this attack.
-
-\Mar{listem 5 primeira regras neste caso}
 
@@
 the input messages should have covered cases with options ``1'' or
-``2'' set. Considering option \CodeIn{windows: 64}, \tname{} is able
+``2'' set. Considering option \CodeIn{window: 512}, \tname{} is able
 to generate rules without that option, as discussed above. Considering
@@
 
-\Luc{Apenas 3 regras tiveram f1 de 100\%. O total de regras nesse caso foi de 13.}
 \Fix{...listem 5 primeiras regras neste caso...}

### 2020-02-21 paper/paper.tex
pass

@@
 legitimate requests cannot be satisfied due to resource exhaustion.
-
 Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
@@
   both, track by\_dst, count 5000, seconds 5} indicates that a high
-volume of such packets should be requested in a short period of time.
-
-In the following, we illustrate how \tname{} proceeds for obtaining
-the rule for the SYN Flood attack. The challenge in synthesizing rules
-is to discover the set of options in a rule that makes the \nids\ to
-capture the malicious traffic, but allow the benign traffic to go
-through.  The other parts of the rule (\ie{}, the action and the
-header) are either inferred or left constant. For example, in this
-case, \tname{} sets the action header to \CodeIn{alert} and sets the
-header to \CodeIn{<proto> any any -> any any}, indicating that it
-infers the target protocol and uses the most general flow description
-possible, \ie{}, it instructs the \nids\ to analyze the traffic
-flowing from/to any address and port.  We understand that this is
-domain-specific information and assume that the system administrator
-would know how to customize it.
-
-\tname{} proceeds as follows to synthesize rules for this
-attack. First, it analyzes the positive (\ie, malicious) traffic and
-extracts options expressed in that traffic. In this case, the
-malicious traffic is characterized as a sequence of 5000
-messages. \tname{} finds the following set of options at this stage:
+volume of such packets should be requested in a short period of
+time. The other options are for documentation purposes.
+
+In the following, we discuss the rule produced by \tname{} to capture
+this attack. The action and the header of the rules that \tname{}
+creates are either inferred or left constant. For example, in this
+case, \tname{} sets the action to \CodeIn{alert} and sets the header
+to \CodeIn{<proto> any any -> any any}, indicating that it infers the
+target protocol and uses the most general flow description possible,
+\ie{}, it instructs the \nids\ to analyze the traffic flowing from/to
+any address and port.  We understand that the choice of these
+parameters are domain-specific and assume that the sys admin would
+know how (and want) to customize them.
+
+The main challenge in synthesizing rules is to discover the set of
+options in a rule that blocks the malicious traffic, but allows the
+benign traffic to pass. \tname{} proceeds as follows to synthesize
+rules for this attack.
+
+First, \tname{} analyzes the malicious traffic and extracts options
+expressed in that traffic. Recall from
+Section~\ref{sec:rules-and-packets} that rule options can be inferred
+from packet contents. In this case, however, that information alone is
+insufficient as the malicious traffic is characterized as a sequence
+of 5K\Mar{really 5K?}\Mar{...and where did they come from?}
+messages. \tname{} finds the following set of options at this stage by
+analyzing the flow and contents of the messages:
 
@@
 \noindent
-\tname{} only reports a rule if the \nids\ instantiated with that rule
-is able to capture the traffic passed on input. Consequently, by
-construction, a rule that uses this set of options captures the
+By construction, a rule that uses this set of options captures the
 malicious traffic. Unfortunately, this rule is potentially
@@
 
-Second, \tname{} uses negative (\ie{}, benign) traffic to mitigate the
-overfitting issue. In this step, \tname{} searches for alternative
-rules that preserves the invarianxt that a rule should capture the
-positive traffic but not capture the negative traffic. It uses the
-rule obtained in the previous step, containing three options, to
-bootstrap the search for rules. The search generates rules containing
-subsets of the options from the seed rule.
+Second, \tname{} uses benign traffic to mitigate the overfitting
+issue. In this step, \tname{} searches for alternative rules that
+preserves the invarianxt that a rule should capture the positive
+malicious traffic but not capture the negative benign traffic. It uses
+the rule obtained in the previous step, containing three options, to
+bootstrap the search for alternative rules. The search generates rules
+containing subsets of the options from the seed rule.
 
@@
 Finally, \tname{} ranks the inferred rules according to heuristics
-extracted from a database of existing rules. One example heuristic
-function we used to rank the generated rules was \Luc{a quantidade de opcoes na regra}
-\Fix{...elaborate...}. The list below shows the rule options for the
-top-5 rules \tname{} produces for this attack.
+learned from a database of existing rules. For example, one heuristic
+function \tname{} uses for ranking leverages the amount of options a
+rule of a given protocol typically has---the score of a rule on this
+function is proportional to the popularity of the number of options it
+declares. The list below shows the rule options for the top-5 rules
+\tname{} produces for this attack.\Mar{algum problema em fazer isto?
+  exemplos concretos ajudam muito!}
 
 To sum up, \tname{} uses the messages that manifest an attack in the
-first step (to create an overspecific seed rule), benign traffic in
-the second step (to create a set of plausible rules from the seed
-rule), and existing rules from popular sources in the third step (to
-rank the generated rules).
+first step to create an overspecific seed rule, benign traffic in the
+second step to create a set of plausible rules from the seed rule, and
+existing rules from popular sources in the third step to rank
+alternative plausible rules.
 
@@
 set while the second parameter indicates non-mandatory flags (\ie{},
-``don't cares''). The option \CodeIn{flags: S} indicates that the flag
-``S'' must be set and all flags not mentioned (``1'' and ``2'', in our
-case) must be unset for the \nids\ to capture the
-traffic. Consequently, that choice of option alone could result in
+``don't cares''). More specifically, the option \CodeIn{flags: S}
+indicates that the flag ``S'' must be set and all flags not mentioned
+(including ``1'' and ``2'') must be unset for the \nids\ to capture
+the traffic. Consequently, that choice of option alone could result in
 false negatives and false positives. An attack that satisfies all
-other conditions in a rule and have flags ``1'' or ``2'' set would be
-captured by the original rule but not the inferred rule---a case of
+other conditions in a rule and has flags ``1'' or ``2'' set would be
+captured by the original rule but not captured by the inferred rule---a case of
 false negative. Likewise, an attack that satisfies all conditions and
 have flags ``1'' and ``2'' unset would be captured by the inferred
-rule but not the original rule---a case of false positive. The set of
+rule but not captured by the original rule---a case of false positive. The set of
 positive examples is to blame in this case. To mitigate this issue,

### 2020-02-21 paper/paper.tex
pass on eval

@@
 
+\usepackage{subcaption}
 \usepackage{libertine}
@@
 \usepackage{listings}
-\usepackage[caption=false, font=footnotesize]{subfig}
+%\usepackage[caption=false, font=footnotesize]{subfig}
 \renewcommand{\ttdefault}{pcr}
@@
 
+More precisely, we calculate the contribution of size of a given rule
+$r$ with the following formula:
+
+\[\frac{\mathit{rules\_per\_size}[\mathit{len(r)}]}{\mathit{max(rules\_per\_size})}\]
+
+\noindent
+, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
+$\mathit{rules\_per\_size}$ encodes distribution of sizes as a
+dictionary mapping sizes to number of associated rules, and
+$\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
+for any dictionary entry. Consequently, the contribution of size is
+proportional to its popularity. For example, the contribution of this
+heuristic for a rule with three options would be 0.684 (=3,662/5,353)
+whereas the contribution of a rule with nine options would be 0.028
+(=154/5,353). Note that the measurement is normalized in the 0-1
+range.
+
+\subsubsection{Number of content options.} We empirically observed that content options are very
+common. \Luc{Se for preciso temos numeros para fundamentar isso
+  melhor, sao cerca de 70K contents nas 28K regras do
+  dataset}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
+  (\%) das 28K regras tem contents.} \Luc{sao 69417 contents nas 26094 regras q utilizam contents} As consequence, we defined a
+heuristic function that takes the number of content options into
+account. We calculate the contribution of number of content options
+with the following formula:
+
 \pgfplotsset{width=6cm,compat=1.8}
@@
 
-\begin{figure}[H]
+\begin{figure*}[t!]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \scalebox{1.2}{  
+    \begin{tikzpicture}
+      \begin{axis}[
+          bar width=6pt,
+          scaled ticks=false,
+          tick label style={/pgf/number format/fixed},
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          xtick=data,
+          nodes near coords,
+          every node near coord/.append style={font=\tiny},
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \vspace{-1ex}
+  \caption{\label{fig:distribution-contents}Number of content options.}
+\end{subfigure}% <---- don't forget this %
+%\vspace{1ex}
+%\\
+\begin{subfigure}{.5\textwidth}
   \centering
@@
   }
-  \vspace{-2ex}
-  \caption{\label{fig:distribution-rule-size}Distribution of rule sizes.}
-\end{figure}
-
-More precisely, we calculate the contribution of size of a given rule
-$r$ with the following formula:
-
-\[\frac{\mathit{rules\_per\_size}[\mathit{len(r)}]}{\mathit{max(rules\_per\_size})}\]
-
-\noindent
-, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
-$\mathit{rules\_per\_size}$ encodes distribution of sizes as a
-dictionary mapping sizes to number of associated rules, and
-$\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
-for any dictionary entry. Consequently, the contribution of size is
-proportional to its popularity. For example, the contribution of this
-heuristic for a rule with three options would be 0.684 (=3,662/5,353)
-whereas the contribution of a rule with nine options would be 0.028
-(=154/5,353). Note that the measurement is normalized in the 0-1
-range.
-
-\subsubsection{Number of content options.} We empirically observed that content options are very
-common. \Luc{Se for preciso temos numeros para fundamentar isso
-  melhor, sao cerca de 70K contents nas 28K regras do
-  dataset}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
-  (\%) das 28K regras tem contents.} \Luc{sao 69417 contents nas 26094 regras q utilizam contents} As consequence, we defined a
-heuristic function that takes the number of content options into
-account. We calculate the contribution of number of content options
-with the following formula:
-
-\begin{figure}[t!]
-  \centering
-  \scalebox{1.2}{  
-    \begin{tikzpicture}
-      \begin{axis}[
-          bar width=6pt,
-          scaled ticks=false,
-          tick label style={/pgf/number format/fixed},
-          ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
-          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
-          xtick=data,
-          nodes near coords,
-          every node near coord/.append style={font=\tiny},
-          nodes near coords align={vertical},
-        ]
-        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
-      \end{axis}
-    \end{tikzpicture}
-  }
-  \caption{\label{fig:distribution-contents}Distribution of number of
-    content options.}
-\end{figure}
-
-\[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
-
-\noindent
-Note that the only difference of this formula compared to the formula
-to compute contribution of size is the distribution used, which is
-given by the term $\mathit{rules\_per\_number\_of\_contents}$ denoting
-a dictionary of number of content options to number of rules that
-contain that number of content options.
-
-\subsubsection{Content rarity.} 
-This heuristic weighs higher rules that use popular strings in content
-options. For a given rule \tname{} takes the arithmetic mean of the
-popularity scores of each string used in contents in that rule. More
-precisely, we calculate the contribution of content rarity with the
-following formula:
-
-\[\frac{\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}}{N}\]
-
-\noindent
-, where the term $\mathit{content_i}$ denotes the $i$-th xcontent in a
-given rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
-denotes the frequency of the string used in the $i$-th content, the
-term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
-observed across all strings appearing as content parameters, and $N$
-is the number of content options in a given
-rule. Figure~\ref{fig:distribution-strings} shows the distribution of
-the ten most popular strings in our rule dataset. To illustrate this
-function, let us consider a rule with two content options, using the
-strings ``GET'' and ``ID'', respectively. The contribution of that
-rule for the computation of similarity score would be 0.601
-(=(1+.203)/2).
-
-\begin{figure}[t!]
+  \vspace{-1ex}
+  \caption{\label{fig:distribution-rule-size}Rule sizes.}
+\end{subfigure}%
+%\vspace{1ex}
+\\
+\begin{subfigure}{.5\textwidth}
   \centering
@@
   }
-  \caption{\label{fig:distribution-strings}Distribution of strings
-    used as parameters of content options.}
-\end{figure}
-
-\subsubsection{Frequency of content modifiers.} As previously
-mentioned, \percRulesWithContent\ (=26K/\numrulessuri) of Suricata
-rules involve content options. Furthermore, of the 26K rules that use
-content options, 13K use HTTP modifiers\Mar{@Lucas, pode dar o numero
-  desconsiderando HTTP? Esta parte eh generica.}. \Luc{Posso, mas nessa função nós consideramos apenas os modificadores http.} Content modifiers
-serve to indicate what part of a network packet the string mentioned
-in the content option should appear. For example, the modifier
-\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
-  http\_header} states the requirement that the string ``keep-alive''
-be present in the header of the message. (A keep-alive connection
-attribute indicates that a single TCP connection should remain open
-for multiple HTTP requests/responses.)
-Figure~\ref{fig:http-header-example} illustrates possible content
-options used to match against certain strings that appear in an HTTP
-message. A check (resp., cross) mark indicates a match (resp.,
-unmatch).
-
-\begin{figure}[t!]
-\centering
-\includegraphics[scale=0.5]{figs/http_header-example.png}
-\caption{\CodeIn{http\_header} example.}
-\label{fig:http-header-example}
-\end{figure}
-
-%% we observed---by analyzing
-%% the dataset of existing rules (see
-%% Section~\ref{sec:dataset-benign})---that these rules often indicate
-%% what part of the message the content string should be checked
-%% against. For example,
-%% For example, the
-%% creator of the rule may want to indicate that a specific string be
-%% present in the header of an HTTP request message. Content
-%% \emph{modifiers} serve this purpose.
-%% Suricata supports content modifiers for tcp and http.
-\Luc{Eh possivel utilizar os modificadores em qualquer protocolo.
-Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo.
-O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós.
-Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso.
-Se necessário posso explicar.} Given that
-%% content options often use modifiers\Mar{assumindo que eh
-%%   verdade. confirmar estat. solicitada acima...},
-
-\tname{} leverages the fact that these modifiers are commonly used to
-define rules as to better rank the rules it synthesizes. More
-precisely, it builds on the observation that strings acessible in
-positions associated with content modifiers (\eg{}, GET, Host,
-Connection) are more likely to produce useful rules. \tname\ uses the
-frequency of content modifiers to discriminate rules that are more or
-less likely to reflect actual rules. For example, historical data
-indicates that content options that use the \CodeIn{http\_connection}
-modifier are much less likely than content options that use the
-modifier
-\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
-shows the histogram of content modifiers for http. Although content
-modifiers are applicable to rules of any protocol supported by
-Suricata, currently, \tname\ only supports tcp and http. These are the
-protocols involving ~80\% of the rules from our dataset.
-
-%% and
-%% (2) the payload of messages of different protocols are not as
-%% segmented as the case of tcp and http. Consequently, rule writers do
-%% not refer to specific parts of the payload as often as the task o and
-%% more challenging for
-
-Similarly to the the previous heuristic functions, this function uses
-the frequency of content modifiers to calibrate the score of each
-rule. More precisely, we compute this function with the following
-formula:
-
-\[\frac{\sum_{i=1}^{N}g(i)}{N},~g(i)=\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]
-
-\noindent
-, where $N$ is the number of contents options, $g(i)$ is the
-contribution of content option $i$ to the score,
-$\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
-of a modifier $j$ to content option $i$,
-$\mathit{max(modifier\_frequency)}$ refers to the modifier with
-maximum popularity (for normalization), and $M_i$ is the number of
-content modifiers that could access the string associated with content
-$i$.
-
-%% \Mar{Qual o rationale para considerar as varias
-%%   alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}0
-%%   \Luc{Vc está supondo que as regras que geramos incluem os modificadores e que deveriamos pontuar somente de acordo com os que estão presentes?
-Não eh o caso.
-A ideia aqui basicamente eh: se um certo content do payload está numa posição q eh muito verificada nas regras, possivelmente esse content eh mais relevante do q outro q está numa posição q raramente eh verificada.
-Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no dataset.}
-%%   \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content.
-Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores que conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}
-
-%% \Luc{$M_i$ eh o numero de modificadores que podem referenciar a posição em q se encontra o $content_i$}\Mar{isto eh confuso. vc. esta dizendo que
-%%   o mesmo content pode ter mais de um modificador? ou a string poderia
-%%   ser acessada por mais de um modificador. se for este o caso,
-%%   sinceramente, achei isto bem confuso. afinal, a regra jah indica
-%%   qual o modificador foi escolhido para acessar a string.}.\Luc{Cada content em uma regra só pode possuir um modificador para indicar a posição onde ele deve ser procurado.
-Entretanto sim, uma "string poderia
-%%   ser acessada por mais de um modificador", foi o caso daquele exemplo com o http\_header.} \Luc{M eh
-%%   o numero de modificadores associados a cada content. Temos
-%%   content[1], content[2], ..., content[N]. Cada um desses contents tem
-%%   modificadores associados content[i][1], content[i][2], ...,
-%%   content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
-%%   content. a frequencia seria 0, correto?} \Luc{Sim}
-
-\begin{figure}[t!]
+  \vspace{-3ex}  
+  \caption{\label{fig:distribution-strings}Strings used as parameters of content options.}
+\end{subfigure}%
+%\vspace{1ex}
+%\\
+\begin{subfigure}{.5\textwidth}
   \centering
@@
   }
-  \vspace{-2ex}
+  \vspace{-1ex}
   \caption{\label{fig:distribution-content_modifiers}Content modifiers
     frequency.}
+\end{subfigure}%
+\vspace{-1ex}
+  \caption{Distribution of features.}
+\end{figure*}
+
+\[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
+
+\noindent
+Note that the only difference of this formula compared to the formula
+to compute contribution of size is the distribution used, which is
+given by the term $\mathit{rules\_per\_number\_of\_contents}$ denoting
+a dictionary of number of content options to number of rules that
+contain that number of content options.
+
+\subsubsection{Content rarity.} 
+This heuristic weighs higher rules that use popular strings in content
+options. For a given rule \tname{} takes the arithmetic mean of the
+popularity scores of each string used in contents in that rule. More
+precisely, we calculate the contribution of content rarity with the
+following formula:
+
+\[\frac{\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}}{N}\]
+
+\noindent
+, where the term $\mathit{content_i}$ denotes the $i$-th xcontent in a
+given rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
+denotes the frequency of the string used in the $i$-th content, the
+term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
+observed across all strings appearing as content parameters, and $N$
+is the number of content options in a given
+rule. Figure~\ref{fig:distribution-strings} shows the distribution of
+the ten most popular strings in our rule dataset. To illustrate this
+function, let us consider a rule with two content options, using the
+strings ``GET'' and ``ID'', respectively. The contribution of that
+rule for the computation of similarity score would be 0.601
+(=(1+.203)/2).
+
+\subsubsection{Frequency of content modifiers.} As previously
+mentioned, \percRulesWithContent\ (=26K/\numrulessuri) of Suricata
+rules involve content options. Furthermore, of the 26K rules that use
+content options, 13K use HTTP modifiers\Mar{@Lucas, pode dar o numero
+  desconsiderando HTTP? Esta parte eh generica.}. \Luc{Posso, mas nessa função nós consideramos apenas os modificadores http.} Content modifiers
+serve to indicate what part of a network packet the string mentioned
+in the content option should appear. For example, the modifier
+\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
+  http\_header} states the requirement that the string ``keep-alive''
+be present in the header of the message. (A keep-alive connection
+attribute indicates that a single TCP connection should remain open
+for multiple HTTP requests/responses.)
+Figure~\ref{fig:http-header-example} illustrates possible content
+options used to match against certain strings that appear in an HTTP
+message. A check (resp., cross) mark indicates a match (resp.,
+unmatch).
+
+\begin{figure}[t!]
+\centering
+\includegraphics[scale=0.5]{figs/http_header-example.png}
+\caption{\CodeIn{http\_header} example.}
+\label{fig:http-header-example}
 \end{figure}
+
+%% we observed---by analyzing
+%% the dataset of existing rules (see
+%% Section~\ref{sec:dataset-benign})---that these rules often indicate
+%% what part of the message the content string should be checked
+%% against. For example,
+%% For example, the
+%% creator of the rule may want to indicate that a specific string be
+%% present in the header of an HTTP request message. Content
+%% \emph{modifiers} serve this purpose.
+%% Suricata supports content modifiers for tcp and http.
+\Luc{Eh possivel utilizar os modificadores em qualquer protocolo.
+Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo.
+O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós.
+Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso.
+Se necessário posso explicar.} Given that
+%% content options often use modifiers\Mar{assumindo que eh
+%%   verdade. confirmar estat. solicitada acima...},
+
+\tname{} leverages the fact that these modifiers are commonly used to
+define rules as to better rank the rules it synthesizes. More
+precisely, it builds on the observation that strings acessible in
+positions associated with content modifiers (\eg{}, GET, Host,
+Connection) are more likely to produce useful rules. \tname\ uses the
+frequency of content modifiers to discriminate rules that are more or
+less likely to reflect actual rules. For example, historical data
+indicates that content options that use the \CodeIn{http\_connection}
+modifier are much less likely than content options that use the
+modifier
+\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
+shows the histogram of content modifiers for http. Although content
+modifiers are applicable to rules of any protocol supported by
+Suricata, currently, \tname\ only supports tcp and http. These are the
+protocols involving ~80\% of the rules from our dataset.
+
+%% and
+%% (2) the payload of messages of different protocols are not as
+%% segmented as the case of tcp and http. Consequently, rule writers do
+%% not refer to specific parts of the payload as often as the task o and
+%% more challenging for
+
+Similarly to the the previous heuristic functions, this function uses
+the frequency of content modifiers to calibrate the score of each
+rule. More precisely, we compute this function with the following
+formula:
+
+\[\frac{\sum_{i=1}^{N}g(i)}{N},~g(i)=\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]
+
+\noindent
+, where $N$ is the number of contents options, $g(i)$ is the
+contribution of content option $i$ to the score,
+$\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
+of a modifier $j$ to content option $i$,
+$\mathit{max(modifier\_frequency)}$ refers to the modifier with
+maximum popularity (for normalization), and $M_i$ is the number of
+content modifiers that could access the string associated with content
+$i$.
+
+%% \Mar{Qual o rationale para considerar as varias
+%%   alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}0
+%%   \Luc{Vc está supondo que as regras que geramos incluem os modificadores e que deveriamos pontuar somente de acordo com os que estão presentes?
+Não eh o caso.
+A ideia aqui basicamente eh: se um certo content do payload está numa posição q eh muito verificada nas regras, possivelmente esse content eh mais relevante do q outro q está numa posição q raramente eh verificada.
+Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no dataset.}
+%%   \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content.
+Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores que conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}
+
+%% \Luc{$M_i$ eh o numero de modificadores que podem referenciar a posição em q se encontra o $content_i$}\Mar{isto eh confuso. vc. esta dizendo que
+%%   o mesmo content pode ter mais de um modificador? ou a string poderia
+%%   ser acessada por mais de um modificador. se for este o caso,
+%%   sinceramente, achei isto bem confuso. afinal, a regra jah indica
+%%   qual o modificador foi escolhido para acessar a string.}.\Luc{Cada content em uma regra só pode possuir um modificador para indicar a posição onde ele deve ser procurado.
+Entretanto sim, uma "string poderia
+%%   ser acessada por mais de um modificador", foi o caso daquele exemplo com o http\_header.} \Luc{M eh
+%%   o numero de modificadores associados a cada content. Temos
+%%   content[1], content[2], ..., content[N]. Cada um desses contents tem
+%%   modificadores associados content[i][1], content[i][2], ...,
+%%   content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
+%%   content. a frequencia seria 0, correto?} \Luc{Sim}
 
@@
 
-\section{Evaluation}
-
-    %%Black Nurse \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;}\\
-    %%Ping Flood  \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;}\\  
-    %%Port Scan \MyComment{& single} & TCP \MyComment{& Discovers open ports in a host} & \scriptsize{flags: A; ack: 0;dsize: 0;} \\
-
-    %%UDP Flood \MyComment{& multiple} & UDP \MyComment{& Denial of Service} & \scriptsize{fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5;} \\
-
-\begin{table*}[h!]
+\begin{table*}[t!]
   \small
@@
 \end{table*}
+
+\section{Evaluation}
+
+This section reports on the evaluation of \tname{}.
+
+\newcommand{\textRQone}{How well \tname\ ranks the golden rule?}
+\vspace{0.2cm}
+\begin{itemize}[leftmargin=*]
+\item{\textbf{RQ1.}} \textRQone
+\end{itemize}
+
+\textbf{Rationale.}~The golden rule is a rule that is known to capture
+a given attack. \tname\ generates several rules that satisfy the
+invariant to only capture the malicious traffic, but the amount of
+data we use to assess that invariant, albeit large, is limited. In
+fact, assessing quality of each rule (\ie{}, fitness to the attack)
+requires domain knowledge. The position of the golden rule at the
+ranking produced by \tname\ is a proxy measure to the effort of a
+system administrator in finding the rule that fits the purpose.
+
+\newcommand{\textRQtwo}{What is the importance of heuristic functions in ranking?}
+\vspace{0.2cm}
+\begin{itemize}[leftmargin=*]
+\item{\textbf{RQ2.}} \textRQtwo\
+\end{itemize}
+
+    %%Black Nurse \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;}\\
+    %%Ping Flood  \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;}\\  
+    %%Port Scan \MyComment{& single} & TCP \MyComment{& Discovers open ports in a host} & \scriptsize{flags: A; ack: 0;dsize: 0;} \\
+
+    %%UDP Flood \MyComment{& multiple} & UDP \MyComment{& Denial of Service} & \scriptsize{fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5;} \\
 
@@
 
-This section reports on the evaluation of \tname{}.
-
 \subsection{Objects of Analysis}
 \label{sec:dataset-benign}
+\label{attack-reproduction}
 
@@
     Flood e se/porque o Nikto nao eh usado neste caso...Enfim, nao
-    precisa responder por aqui. Mova o texto deixando-o *consistente*.}
-
-\subsubsection{Benign traffic.}\tname{} uses public databases storing benign
-traffic for all protocols covered by Suricata (see
-Figure~\ref{fig:distribution-rules-protocol}). We used two datasets of
-benign traffic. The Bigflows.pcap dataset is part of the
-TcpReplay~\cite{tcpreplay} open source project. This dataset includes
-real network traffic of a busy private network access point to the
-Internet. According to the TcpReplay web site ``This capture is much
-larger and has a smaller average packet size than the previous capture
-(smallFlows.pcap). It also has many more flows and different
-applications.''. In addition to the Bigflows.pcap dataset, we also
-used datasets of normal traffic made available by Stratosphere
-Lab~\cite{stratosphere-normal}.
-
-\subsubsection{Ground Truth.}To evaluate precision of \tname\ we
-need rules that are considered correct to use as reference. The rules
-that detect the attacks from Table~\ref{table:attacks} were obtained
-from the Emerging Threats Open Ruleset~\cite{emerging-threats-open}.
-
-%% Note from Figure~\ref{fig:overview} that the positive traffic is an
-%% input to the technique. The negative non-malicious traffic, however,
-%% is not.
-
-\subsection{Methodology}
-
-\subsubsection{Attack reproduction.}\label{attack-reproduction} \tname{} requires the attack to
+    precisa responder por aqui. Mova o texto deixando-o
+    *consistente*.}
+
+  \tname{} requires the attack to
 be replayed in order to check whether or not the rule it generates
@@
 monitor~\cite{wireshark-net-monitor}.
+
+\subsubsection{Benign traffic.}\tname{} uses public databases storing benign
+traffic for all protocols covered by Suricata (see
+Figure~\ref{fig:distribution-rules-protocol}). We used two datasets of
+benign traffic. The Bigflows.pcap dataset is part of the
+TcpReplay~\cite{tcpreplay} open source project. This dataset includes
+real network traffic of a busy private network access point to the
+Internet. According to the TcpReplay web site ``This capture is much
+larger and has a smaller average packet size than the previous capture
+(smallFlows.pcap). It also has many more flows and different
+applications.''. In addition to the Bigflows.pcap dataset, we also
+used datasets of normal traffic made available by Stratosphere
+Lab~\cite{stratosphere-normal}.
+
+\subsubsection{Ground Truth.}To evaluate precision of \tname\ we
+need rules that are considered correct to use as reference. The rules
+that detect the attacks from Table~\ref{table:attacks} were obtained
+from the Emerging Threats Open Ruleset~\cite{emerging-threats-open}.
+
+%% Note from Figure~\ref{fig:overview} that the positive traffic is an
+%% input to the technique. The negative non-malicious traffic, however,
+%% is not.
+
+\subsection{Methodology}
 
@@
 
-\subsection{Results}
+\subsection{Answering RQ1: \textRQone}
+
+\subsection{Answering RQ2: \textRQtwo}
 

### 2020-02-22 paper/paper.tex
pass on table 3

@@
     iisadmin access & 862 & 372 & 1 \\
-    /system32/ in URI & - & - & - \\
+    /system32/ in URI & 3665 & 783 & 2 \\
     Script tag in URI & - & - & - \\

### 2020-02-22 paper/paper.tex
pass on example

@@
 
-Active reconnaissance is a method used by malicious individuals to
+Active Reconnaissance is a method used by malicious individuals to
 collect information about a computing system to determine potential
-vulnerabilities. Active reconnaissance is a preliminary step towards
-an actual attack. For example, Ping Scan is a very common type of
-attack that scans the network for IPs of accessible hosts. An attacker
-sends several ICMP Echo requests to a range of IP addresses in the
-network and checks which ones respond to the requests. This allow the
-attacker to know which machines are active in the network. Then (s)he
-can choose one of them to attack.
-
+vulnerabilities. This methods is used as a preliminary step towards an
+actual attack. For example, Ping Scan scans the network for IPs of
+accessible hosts. An attacker sends several ICMP Echo requests to a
+range of IP addresses in the network and checks which ones respond to
+these requests. This method enables an attacker to know which machines
+are active in the network.
+
+\vspace{-2ex}
 \begin{figure}[h!]
-  \vspace{1ex}
   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
+  \vspace{-2ex}    
   \caption{Suricata rule for Ping Scan attack.}
   \label{fig:pingscan-example}
-  \vspace{1ex}  
 \end{figure}
-
-\begin{figure*}[ht!]
-  \centering
-  \includegraphics[trim=0 350 100 0,clip,width=0.65\textwidth]{figs/nids-workflow}
-  \caption{\tname\ workflow.}
-  \label{fig:overview}
-  \vspace{-1ex}
-\end{figure*}
-
-Figure 4 shows a Suricata rule that captures a reproduction of a Ping
-Scan attack available from NMAP~\cite{netmap}, a network discovery and
-security auditing open-source tool. The relevant options are in
-bold---the option \CodeIn{dsize: 0} indicates that the packet has no
-bytes in the payload and the option \CodeIn{itype: 8} indicates that
-the packet to be captured are ICMP Echo requests.
-
-Again, to synthesize rules for a Ping Scan, \tname{} analyzes the
-positive malicious traffic and extracts the following rule options:
+\vspace{-4ex}
+
+Figure 4 shows a Suricata rule that captures a Ping Scan attempt, as
+documented in NMAP~\cite{netmap}, a network discovery and security
+auditing open-source tool. The relevant options are in bold---the
+option \CodeIn{dsize: 0} indicates that the packet has no bytes in the
+payload and the option \CodeIn{itype: 8} indicates that the packet to
+be captured are ICMP Echo requests. Again, to synthesize rules for a
+Ping Scan, \tname{} analyzes the positive malicious traffic and
+extracts the following rule options:
 
@@
 
-Again, note that rule \CodeIn{icode} was not present in the original
-rule.  \tname{} produces the following five rules at the top of the
-ranking after minimization and ranking.
+\noindent
+Note that, although options \CodeIn{dsize:0} and \CodeIn{itype:8} are
+present in the set of inferred options, the others are not.  This
+indicates, as expected, that the set of rule options is
+overspecified. \tname{} produces the following five rules at the top
+of the ranking after minimization and ranking.\Mar{eu ainda estou
+  aguardando isto. A tabela 3 tambem nao indica o resultado!}
 
@@
 \label{sec:technique}
+
+\begin{figure*}[ht!]
+  \centering
+  \includegraphics[trim=0 350 100 0,clip,width=0.65\textwidth]{figs/nids-workflow}
+  \caption{\tname\ workflow.}
+  \label{fig:overview}
+  \vspace{-1ex}
+\end{figure*}
 

### 2020-02-22 paper/paper.tex
pass on 3.1

@@
 insufficient as the malicious traffic is characterized as a sequence
-of 5K\Mar{really 5K?}\Mar{...and where did they come from?}
+of 5K\Mar{really 5K?}\Mar{...and where did they come from?} \Luc{Sim, floods podem ser compostos por milhares de pacotes, um dos datasets q encontrei possui mais de 500k pacotes de flood.
+Os datasets foram encontrados na internet.}
 messages. \tname{} finds the following set of options at this stage by
@@
 \tname{} produces for this attack.\Mar{algum problema em fazer isto?
-  exemplos concretos ajudam muito!}
+  exemplos concretos ajudam muito!} \Luc{No caso desse ataque, a regra inicial possui apenas duas opcoes (sem contar o threshold).
+Entao existem apenas 3 combinacoes possiveis, dessas 3 apenas 2 sao capazes de isolar o ataque do trafego normal.
+Portanto, o top-2 eh:
+
+  \CodeIn{...window:512; threshold: type both, track by\_dst, count 5000, seconds 3;...}
+
+  \CodeIn{...flags: S; window:512; threshold: type both, track by\_dst, count 5000, seconds 3;...}
+   }
 
@@
   aguardando isto. A tabela 3 tambem nao indica o resultado!}
+  \Luc{Irei adicionar}
 

### 2020-02-22 paper/paper.tex
pass on 3.1 3.2
section 3.1

--- 2020-02-22 paper/paper.tex
@@
 produces options of the second kind, which are \emph{relevant} to
-capturing the attack.
+capture the attack. For example, the option \CodeIn{content:<arg>} checks if
+the argument is present in the payload of the message.
 
@@
     8 & fragbits & checks if the fragmentation or reserved bits are set in the IP header\\
-    9 & threshold & controls the rule’s alert frequency\\
+    9 & threshold & minimum threshold for a rule before it generates alerts\\
     10 & content & checks if argument is present on the payload\\
@@
 benign traffic to pass. \tname{} proceeds as follows to synthesize
-rules for this attack.
-
-First, \tname{} analyzes the malicious traffic and extracts options
+rules for this attack. First, \tname{} analyzes the malicious traffic and extracts options
 expressed in that traffic. Recall from
@@
 insufficient as the malicious traffic is characterized as a sequence
-of 5K\Mar{really 5K?}\Mar{...and where did they come from?}
-messages. \tname{} finds the following set of options at this stage by
+of messages. \tname{} finds the following set of options at this stage by
 analyzing the flow and contents of the messages:
+
+%% \Mar{...and where did they come from?} \Luc{Sim, floods podem ser compostos por milhares de pacotes, um dos datasets q encontrei possui mais de 500k pacotes de flood.
+Os datasets foram encontrados na internet.}
 
@@
 \tname{} produces for this attack.\Mar{algum problema em fazer isto?
-  exemplos concretos ajudam muito!}
+  exemplos concretos ajudam muito!} \Luc{No caso desse ataque, a regra inicial possui apenas duas opcoes (sem contar o threshold).
+Entao existem apenas 3 combinacoes possiveis, dessas 3 apenas 2 sao capazes de isolar o ataque do trafego normal.
+Portanto, o top-2 eh:
+  \begin{figure}[h]
+    \vspace{-2ex}
+    \lstinputlisting[language=C,numbers=none,frame=none,keywords={window,flags}]{top-synflood.rules}
+    \vspace{-3ex}  
+  \end{figure}
+
+   }\Mar{Lucas, eu nao entendo porque vc. desconsiderou threshold da
+  discussao (afinal, a gente destaca thrshold como relevante no inicio
+  desta secao). tambem nao entendo porque window---que a gente disse
+  acima que *nao* era relevante---aparece nas duas regras
+  geradas. enfim, isto esta inconsistente com nosso discurso.}
 
 To sum up, \tname{} uses the messages that manifest an attack in the
-first step to create an overspecific seed rule, benign traffic in the
-second step to create a set of plausible rules from the seed rule, and
-existing rules from popular sources in the third step to rank
-alternative plausible rules.
+first step to create an overspecific seed rule, it uses the benign
+traffic in the second step to create a set of plausible rules from the
+seed rule, and it uses existing rules from popular sources in the
+third step to rank alternative plausible rules.
 

### 2020-02-22 paper/paper.tex
examples

@@
 
+\subsection{Active Reconnaissance}
+\label{sec:active-recon}
+
+%% \Mar{@Lucas, acima falamos de porta. Abaixo nao falamos de
+%%   porta, apenas de maquina. Vc. pode ser mais especifico? ->} \Luc{De
+%%   fato estamos falando de duas coisas diferentes, mas ambas são
+%%   exemplos de 'active reconnaissance'. Primeiramente citamos o port
+%%   scanning, q tem o objetivo de encontrar portas abertas em uma
+%%   maquina especifica. Depois falamos do ping scan, q tem o objetivo de
+%%   encontrar ip's ativos numa rede.}
+
+Active Reconnaissance is a method used by malicious individuals to
+collect information about a computing system to determine potential
+vulnerabilities. This methods is used as a preliminary step towards an
+actual attack. For example, Ping Scan scans the network for IPs of
+accessible hosts. An attacker sends several ICMP Echo requests to a
+range of IP addresses in the network and checks which ones respond to
+these requests. This method enables an attacker to know which machines
+are active in the network.
+
+\vspace{-1ex}
+\begin{figure}[h!]
+  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
+  \vspace{-2ex}  
+  \caption{Suricata rule for Ping Scan attack.}
+  \label{fig:pingscan-example}
+\end{figure}
+\vspace{-3ex}
+
+Figure 4 shows a Suricata rule that captures a Ping Scan attempt, as
+documented in NMAP~\cite{netmap}, a network discovery and security
+auditing open-source tool. The relevant options are in bold---the
+option \CodeIn{dsize: 0} indicates that the packet has no bytes in the
+payload and the option \CodeIn{itype: 8} indicates that the packet to
+be captured are ICMP Echo requests.
+
+In the following, we discuss the rule produced by \tname{} to capture
+this attack. The action and the header of the rules that \tname{}
+creates are either inferred or left constant. For example, in this
+case, \tname{} sets the action to \CodeIn{alert} and sets the header
+to \CodeIn{<proto> any any -> any any}, indicating that it infers the
+target protocol and uses the most general flow description possible,
+\ie{}, it instructs the \nids\ to analyze the traffic flowing from/to
+any address and port.  The choice of these parameters are
+domain-specific and subject to change by the system administrator.
+The main challenge in synthesizing rules is to discover the set of
+options in a rule that blocks the malicious traffic, but allows the
+benign traffic to pass.
+
+\tname{} proceeds as follows to synthesize rules. First, it analyzes the malicious
+traffic and extracts options expressed in that traffic. Recall from
+Section~\ref{sec:rules-and-packets} that rule options can be inferred
+from packet contents. \tname{} finds the following set of options at
+this stage:
+
+\begin{figure}[h]
+  \vspace{-2ex}
+  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{pingscan.suricata.synth}
+  \vspace{-2ex}  
+\end{figure}
+
+\noindent
+By construction, a rule that uses this set of options captures the
+malicious traffic. Unfortunately, the rule above is
+overspecified. Informally, the rule includes ``too specific''
+characteristics of the malicious traffic and could potentially result
+in false negatives. In this case, although options \CodeIn{dsize:0}
+and \CodeIn{itype:8}~(see Figure~\ref{fig:pingscan-example}) are
+present in the set of inferred options, the others are not.
+
+Second, \tname{} uses the benign traffic to mitigate the overfitting
+issue. In this step, \tname{} searches for alternative rules that
+preserve the invariant to capture the positive malicious traffic but
+not capture the negative benign traffic. It uses the rule obtained in
+the previous step as seed to bootstrap the search for alternative
+plausible rules, \ie{}, rules that satisfy the aforementioned
+invariant. The search is designed to generate rules containing subsets
+of the options from the seed rule. Considering this example, \tname{}
+produces a total of \Fix{XXXXX} plausible rules at the end of the
+search.
+
+Finally, \tname{} ranks plausible rules according to heuristics
+learned from a database of existing rules. The intuition is that new
+rules should share characteristics with pre-existing rules. For
+example, one heuristic function used by \tname{} leverages the amount
+of options a rule typically has. The score of a rule, given by this
+function, is proportional to the popularity of the number of options
+declared in rules of a given protocol. At the end of the process,
+\tname{} ranks the golden rule at the top of the ranking including
+\Fix{XXXXXX} other plausible rules.
+
+%% the following rules 
+%% at the top of the ranking produced by \tname{}:
+%% 1)~{\small\CodeIn{itype:8; dsize:0;}}, 2)~{\small\CodeIn{itype:8;
+%%     icmp\_id:23570;}}, 3)~{\small\CodeIn{icode:8; icmp\_id:23570;}},
+%% 4)~{\small\CodeIn{itype:8; icode:0; dsize:0;}},
+%% 5)~{\small\CodeIn{icmp\_id:23570}}. Note that the golden rule appears
+%% at the top position.
+
+%% \begin{figure}[h]
+%%   \vspace{-2ex}
+%%   \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{top-pingscan.rules}
+%%   \vspace{-3ex}  
+%% \end{figure}
+
+To sum up, \tname{} uses the messages that manifest an attack in the
+first step to create a potentially overspecified seed rule, it uses
+the benign traffic in the second step to create a set of plausible
+rules from the seed rule, and it uses public databases with existing
+rules in the third step to rank plausible rules.
+
+\subsection{Remote Code Execution (RCE)}
+\label{sec:rce}
+\label{sec:jboss}
+\label{sec:content-example}
+
+%% Adaptacao do texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution
+
+RCE refers to the family of attacks where hackers exploit a network
+vulnerability to run arbitrary code---that could damage the system or
+steal sensitive information---on a remote machine. JBoss~\cite{jboss}
+is popular open source Java application server. This attacks enables
+remote execution of code on the JBoss application server.
+\Mar{@Lucas, por favor, observe o padrao usado nos exemplos
+  anteriores. eh preciso explicar em alto nivel o ataque antes de
+  explicar a regra!}
+
+%% In an RCE attack, hackers
+%% intentionally exploit a vulnerability in the system to run malware
+%% By prompting the targeted device to perform code execution, a hacker
+%% can run arbritrary code to
+%% their own programming in its
+%% place. This programming can then enable them to gain full access,
+%% steal data, carry out a full distributed denial of service (DDoS)
+%% attack, destroy files and infrastructure, or engage in illegal
+%% activity. And as the term remote execution suggests, an RCE
+%% cyberattack can take place from any geophysical location."\\
+
+\begin{figure}[H]
+  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
+  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code Execution.}
+\end{figure}
+
+%A opção  define que o match só será feito em pacotes que fazem parte
+%de uma conexão já estabelecida entre um cliente e um servidor. Além
+%disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.
+
+Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
+JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
+indicates that a rule match will only occur on packets that (1) flow
+on an already-established connection and that (2) flow to the server
+as opposed to the client. For the content options the modifiers
+\CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
+the matching not be case sensitive and that the string should be
+located at the packet field ``Request URI''. \Mar{Lucas, ``nocase''
+  tambem eh considerado modificador? Isto nao eh confuso? Na secao
+  4.3.4 consideramos modificadores como sinonimo para qualificador de
+  campo, nao?}\Luc{Sim, ``nocase'' define que a busca pela string nao deve ser case sensitive. Em 4.3.4 estamos considerando apenas os modificadores que definem posições/campos de pacotes http.}
+
+\tname{} reproduces the malicious JBoss attack using the traffic from
+a Nikto script~\cite{nikto}. However, similar JBoss attack is also
+available from the Metasploit database~\cite{metasploit}. As in the
+other examples, \tname{} uses, as additional inputs, benign traffic
+from \Fix{...} and a set of existing Suricata rules from
+\Fix{...}. \tname{} initially produces one seed rule containing
+\Fix{xx} options, by analyzing the malicious traffic alone. Then, it
+uses that seed rule to derive \Fix{yy} rules, each rule containing
+subsets of the option in the initial rule. Finally, it ranks that
+list. The list below shows the top-5 rules that \tname{}
+produces.\Mar{<- check/revise}
+
+\Fix{list rules}
+
+Note that the golden rule is in position \Fix{...} on this
+case. \Mar{explain syntactical difference, if exist}
+
 \subsection{Denial-of-Service}
@@
 
-In the following, we discuss the rule produced by \tname{} to capture
-this attack. The action and the header of the rules that \tname{}
-creates are either inferred or left constant. For example, in this
-case, \tname{} sets the action to \CodeIn{alert} and sets the header
-to \CodeIn{<proto> any any -> any any}, indicating that it infers the
-target protocol and uses the most general flow description possible,
-\ie{}, it instructs the \nids\ to analyze the traffic flowing from/to
-any address and port.  We understand that the choice of these
-parameters are domain-specific and assume that the sys admin would
-know how (and want) to customize them.
-
-The main challenge in synthesizing rules is to discover the set of
-options in a rule that blocks the malicious traffic, but allows the
-benign traffic to pass. \tname{} proceeds as follows to synthesize
-rules for this attack. First, \tname{} analyzes the malicious traffic and extracts options
-expressed in that traffic. Recall from
-Section~\ref{sec:rules-and-packets} that rule options can be inferred
-from packet contents. In this case, however, that information alone is
+...........
+In this case, however, that information alone is
 insufficient as the malicious traffic is characterized as a sequence
-of messages. \tname{} finds the following set of options at this stage by
-analyzing the flow and contents of the messages:
+of messages. 
+...........
 
@@
 \noindent
-By construction, a rule that uses this set of options captures the
-malicious traffic. Unfortunately, this rule is potentially
-overspecified. Informally, the rule includes ``too specific''
-characteristics of the malicious traffic and could potentially result
-in false negatives. For example, in this particular case, note that
-the option \CodeIn{window: 512} is not included in the golden rule
-from Figure~\ref{fig:synflood-example}.
-
-Second, \tname{} uses benign traffic to mitigate the overfitting
-issue. In this step, \tname{} searches for alternative rules that
-preserves the invarianxt that a rule should capture the positive
-malicious traffic but not capture the negative benign traffic. It uses
-the rule obtained in the previous step, containing three options, to
-bootstrap the search for alternative rules. The search generates rules
-containing subsets of the options from the seed rule.
-
-%% \Luc{Ataque "JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit"
-%% . O payload do pacote é o seguinte:
-%% \begin{figure}[H]
-%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-rule.suricata}
-%%   \label{fig:rule-example}
-%% \end{figure}
-
-%% A partir disso produzimos uma regra com os seguintes contents:
-
-%% \begin{figure}[H]
-%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-contents-splitted.suricata}
-%%   \label{fig:rule-example}
-%% \end{figure}
-
-%% Depois da etapa da minimização, as top 5 regras são:
-
-%% \begin{figure}[H]
-%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-top5-rules.suricata}
-%%   \label{fig:rule-example}
-%% \end{figure}
-
-%% A regra de ouro é:
-
-%% \begin{figure}[H]
-%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-golden-rule.suricata}
-%%   \label{fig:rule-example}
-%% \end{figure}
-
-%% }
-
-Finally, \tname{} ranks the inferred rules according to heuristics
-learned from a database of existing rules. For example, one heuristic
-function \tname{} uses for ranking leverages the amount of options a
-rule of a given protocol typically has---the score of a rule on this
-function is proportional to the popularity of the number of options it
-declares. The list below shows the rule options for the top-5 rules
+
+.....
+
+ The list below shows the rule options for the top-5 rules
 \tname{} produces for this attack.\Mar{algum problema em fazer isto?
@@
   geradas. enfim, isto esta inconsistente com nosso discurso.}
-
-To sum up, \tname{} uses the messages that manifest an attack in the
-first step to create an overspecific seed rule, it uses the benign
-traffic in the second step to create a set of plausible rules from the
-seed rule, and it uses existing rules from popular sources in the
-third step to rank alternative plausible rules.
 
@@
 this option.
-
-\subsection{Active Reconnaissance}
-\label{sec:active-recon}
-
-%% \Mar{@Lucas, acima falamos de porta. Abaixo nao falamos de
-%%   porta, apenas de maquina. Vc. pode ser mais especifico? ->} \Luc{De
-%%   fato estamos falando de duas coisas diferentes, mas ambas são
-%%   exemplos de 'active reconnaissance'. Primeiramente citamos o port
-%%   scanning, q tem o objetivo de encontrar portas abertas em uma
-%%   maquina especifica. Depois falamos do ping scan, q tem o objetivo de
-%%   encontrar ip's ativos numa rede.}
-
-Active Reconnaissance is a method used by malicious individuals to
-collect information about a computing system to determine potential
-vulnerabilities. This methods is used as a preliminary step towards an
-actual attack. For example, Ping Scan scans the network for IPs of
-accessible hosts. An attacker sends several ICMP Echo requests to a
-range of IP addresses in the network and checks which ones respond to
-these requests. This method enables an attacker to know which machines
-are active in the network.
-
-\vspace{-2ex}
-\begin{figure}[h!]
-  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
-  \vspace{-2ex}    
-  \caption{Suricata rule for Ping Scan attack.}
-  \label{fig:pingscan-example}
-\end{figure}
-\vspace{-4ex}
-
-Figure 4 shows a Suricata rule that captures a Ping Scan attempt, as
-documented in NMAP~\cite{netmap}, a network discovery and security
-auditing open-source tool. The relevant options are in bold---the
-option \CodeIn{dsize: 0} indicates that the packet has no bytes in the
-payload and the option \CodeIn{itype: 8} indicates that the packet to
-be captured are ICMP Echo requests. Again, to synthesize rules for a
-Ping Scan, \tname{} analyzes the positive malicious traffic and
-extracts the following rule options:
-
-\begin{figure}[h]
-  \vspace{-2ex}
-  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{pingscan.suricata.synth}
-  \vspace{-3ex}  
-\end{figure}
-
-\noindent
-Note that, although options \CodeIn{dsize:0} and \CodeIn{itype:8} are
-present in the set of inferred options, the others are not.  This
-indicates, as expected, that the set of rule options is
-overspecified. \tname{} produces the following five rules at the top
-of the ranking after minimization and ranking.\Mar{eu ainda estou
-  aguardando isto. A tabela 3 tambem nao indica o resultado!}
-
-\begin{figure}[h]
-  \vspace{-2ex}
-  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{top-pingscan.rules}
-  \vspace{-3ex}  
-\end{figure}
-
-\Fix{...listem 5 primeiras regras neste caso...}
-
-\subsection{Remote Code Execution (RCE)}
-\label{sec:rce}
-\label{sec:jboss}
-\label{sec:content-example}
-
-%% Adaptacao do texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution
-
-RCE refers to the family of attacks where hackers exploit a network
-vulnerability to run arbitrary code---that could damage the system or
-steal sensitive information---on a remote machine. JBoss~\cite{jboss}
-is popular open source Java application server. This attacks enables
-remote execution of code on the JBoss application server.
-\Mar{@Lucas, por favor, observe o padrao usado nos exemplos
-  anteriores. eh preciso explicar em alto nivel o ataque antes de
-  explicar a regra!}
-
-%% In an RCE attack, hackers
-%% intentionally exploit a vulnerability in the system to run malware
-%% By prompting the targeted device to perform code execution, a hacker
-%% can run arbritrary code to
-%% their own programming in its
-%% place. This programming can then enable them to gain full access,
-%% steal data, carry out a full distributed denial of service (DDoS)
-%% attack, destroy files and infrastructure, or engage in illegal
-%% activity. And as the term remote execution suggests, an RCE
-%% cyberattack can take place from any geophysical location."\\
-
-\begin{figure}[H]
-  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
-  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code Execution.}
-\end{figure}
-
-%A opção  define que o match só será feito em pacotes que fazem parte
-%de uma conexão já estabelecida entre um cliente e um servidor. Além
-%disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.
-
-Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
-JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
-indicates that a rule match will only occur on packets that (1) flow
-on an already-established connection and that (2) flow to the server
-as opposed to the client. For the content options the modifiers
-\CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
-the matching not be case sensitive and that the string should be
-located at the packet field ``Request URI''. \Mar{Lucas, ``nocase''
-  tambem eh considerado modificador? Isto nao eh confuso? Na secao
-  4.3.4 consideramos modificadores como sinonimo para qualificador de
-  campo, nao?}\Luc{Sim, ``nocase'' define que a busca pela string nao deve ser case sensitive. Em 4.3.4 estamos considerando apenas os modificadores que definem posições/campos de pacotes http.}
-
-\tname{} reproduces the malicious JBoss attack using the traffic from
-a Nikto script~\cite{nikto}. However, similar JBoss attack is also
-available from the Metasploit database~\cite{metasploit}. As in the
-other examples, \tname{} uses, as additional inputs, benign traffic
-from \Fix{...} and a set of existing Suricata rules from
-\Fix{...}. \tname{} initially produces one seed rule containing
-\Fix{xx} options, by analyzing the malicious traffic alone. Then, it
-uses that seed rule to derive \Fix{yy} rules, each rule containing
-subsets of the option in the initial rule. Finally, it ranks that
-list. The list below shows the top-5 rules that \tname{}
-produces.\Mar{<- check/revise}
-
-\Fix{list rules}
-
-Note that the golden rule is in position \Fix{...} on this
-case. \Mar{explain syntactical difference, if exist}
 

### 2020-02-23 paper/paper.tex
pass

@@
 
-\section{Illustrative Examples}
+\section{Illustrative Example}
 \label{sec:suri-metas-coverage}
-
-This section briefly presents examples of attacks and illustrates the
-rules that \tname{} synthesizes to capture those attacks.
-
-\subsection{Active Reconnaissance}
 \label{sec:active-recon}
 
-%% \Mar{@Lucas, acima falamos de porta. Abaixo nao falamos de
-%%   porta, apenas de maquina. Vc. pode ser mais especifico? ->} \Luc{De
-%%   fato estamos falando de duas coisas diferentes, mas ambas são
-%%   exemplos de 'active reconnaissance'. Primeiramente citamos o port
-%%   scanning, q tem o objetivo de encontrar portas abertas em uma
-%%   maquina especifica. Depois falamos do ping scan, q tem o objetivo de
-%%   encontrar ip's ativos numa rede.}
-
-Active Reconnaissance is a method used by malicious individuals to
-collect information about a computing system to determine potential
-vulnerabilities. This methods is used as a preliminary step towards an
-actual attack. For example, Ping Scan scans the network for IPs of
-accessible hosts. An attacker sends several ICMP Echo requests to a
-range of IP addresses in the network and checks which ones respond to
-these requests. This method enables an attacker to know which machines
-are active in the network.
+This section briefly shows how \tname{} synthesizes rules for one
+example attack. Active Reconnaissance is a method used by malicious
+individuals to collect information about a computing system to
+determine potential vulnerabilities. This method is used as a
+preliminary step towards an actual attack. For example, Ping Scan
+scans the network for IPs of accessible hosts. An attacker sends
+several ICMP Echo requests to a range of IP addresses in the network
+and checks which ones respond to these requests. This method enables
+an attacker to know which machines are active in the network.
 
@@
 
-Figure 4 shows a Suricata rule that captures a Ping Scan attempt, as
-documented in NMAP~\cite{netmap}, a network discovery and security
-auditing open-source tool. The relevant options are in bold---the
+Figure~\ref{fig:pingscan-example} shows a Suricata rule that captures
+a Ping Scan attempt, as documented in NMAP~\cite{netmap}, a network
+discovery and security auditing open-source tool. The relevant options
+(see Section~\ref{sec:example-suricata-rules}) are in bold---the
 option \CodeIn{dsize: 0} indicates that the packet has no bytes in the
@@
 
-In the following, we discuss the rule produced by \tname{} to capture
+In the following, we briefly show how \tname{} synthesizes rules for
 this attack. The action and the header of the rules that \tname{}
@@
 \ie{}, it instructs the \nids\ to analyze the traffic flowing from/to
-any address and port.  The choice of these parameters are
-domain-specific and subject to change by the system administrator.
-The main challenge in synthesizing rules is to discover the set of
-options in a rule that blocks the malicious traffic, but allows the
-benign traffic to pass.
-
-\tname{} proceeds as follows to synthesize rules. First, it analyzes the malicious
-traffic and extracts options expressed in that traffic. Recall from
-Section~\ref{sec:rules-and-packets} that rule options can be inferred
-from packet contents. \tname{} finds the following set of options at
-this stage:
-
-\begin{figure}[h]
-  \vspace{-2ex}
-  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{pingscan.suricata.synth}
-  \vspace{-2ex}  
-\end{figure}
-
-\noindent
-By construction, a rule that uses this set of options captures the
-malicious traffic. Unfortunately, the rule above is
-overspecified. Informally, the rule includes ``too specific''
-characteristics of the malicious traffic and could potentially result
-in false negatives. In this case, although options \CodeIn{dsize:0}
-and \CodeIn{itype:8}~(see Figure~\ref{fig:pingscan-example}) are
-present in the set of inferred options, the others are not.
+any address and port.  The rationale for this decision is that the
+choice of these parameters are domain-specific and subject to change
+by the system administrator.  The main challenge in synthesizing rules
+is to discover the set of options in a rule that blocks the malicious
+traffic, but allows the benign traffic to pass.
+
+\tname{} proceeds as follows to synthesize rules. First, it analyzes
+the malicious traffic and extracts options expressed in that
+traffic. Recall from Section~\ref{sec:rules-and-packets} that some
+options can be inferred from packet contents. \tname{} finds the
+following options at this stage:
+{\scriptsize{\texttt{\textbf{dsize}:0; \textbf{itype:}8;
+      \textbf{icode:}0; \textbf{icmp\_id:}23570;
+      \textbf{icmp\_seq:}3439;}}}.  By construction, a rule that uses
+this set of options captures the malicious traffic. Unfortunately, the
+rule above is overspecified. Informally, the rule includes ``too
+specific'' characteristics of the malicious traffic and could
+potentially result in false negatives. In this case, although the
+options \CodeIn{dsize:0} and \CodeIn{itype:8} are present in the set
+of inferred options (as the golden rule from
+Figure~\ref{fig:pingscan-example}), the others are not.
+
+%% \vspace{-2.5ex}
+%% \begin{figure}[h]
+%%   \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{pingscan.suricata.synth}
+%% \end{figure}
+%% \vspace{-2.5ex}
+%% \noindent
 
@@
 produces a total of \Fix{XXXXX} plausible rules at the end of the
-search.
+search.\Mar{Technically, there are $32 (=2^5)$ subsets for a set with 5
+  elements. So, it is important to indicate why Table 3 shows sooo
+  many plausible rules and that this is is a very simple rule (for illustration).}
 
@@
 function, is proportional to the popularity of the number of options
-declared in rules of a given protocol. At the end of the process,
-\tname{} ranks the golden rule at the top of the ranking including
-\Fix{XXXXXX} other plausible rules.
-
-%% the following rules 
-%% at the top of the ranking produced by \tname{}:
-%% 1)~{\small\CodeIn{itype:8; dsize:0;}}, 2)~{\small\CodeIn{itype:8;
-%%     icmp\_id:23570;}}, 3)~{\small\CodeIn{icode:8; icmp\_id:23570;}},
-%% 4)~{\small\CodeIn{itype:8; icode:0; dsize:0;}},
-%% 5)~{\small\CodeIn{icmp\_id:23570}}. Note that the golden rule appears
-%% at the top position.
-
-%% \begin{figure}[h]
-%%   \vspace{-2ex}
-%%   \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{top-pingscan.rules}
-%%   \vspace{-3ex}  
-%% \end{figure}
+declared in rules of a given
+protocol. Figure~\ref{fig:distribution-contents} shows the
+distribution of the number of content options for the database of
+rules we used\Mar{reference}. Note that rules with 1 to 3 options are
+more common.  At the end of the process, \tname{} ranks the golden
+rule at the top of the ranking including \Fix{XXXXXX} other plausible
+rules.
+
+\Mar{add short par indicating that other problems can be much more
+  complex than this...}
 
 To sum up, \tname{} uses the messages that manifest an attack in the
-first step to create a potentially overspecified seed rule, it uses
-the benign traffic in the second step to create a set of plausible
-rules from the seed rule, and it uses public databases with existing
-rules in the third step to rank plausible rules.
-
-\subsection{Remote Code Execution (RCE)}
-\label{sec:rce}
-\label{sec:jboss}
-\label{sec:content-example}
-
-%% Adaptacao do texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution
-
-RCE refers to the family of attacks where hackers exploit a network
-vulnerability to run arbitrary code---that could damage the system or
-steal sensitive information---on a remote machine. JBoss~\cite{jboss}
-is popular open source Java application server. This attacks enables
-remote execution of code on the JBoss application server.
-\Mar{@Lucas, por favor, observe o padrao usado nos exemplos
-  anteriores. eh preciso explicar em alto nivel o ataque antes de
-  explicar a regra!}
-
-%% In an RCE attack, hackers
-%% intentionally exploit a vulnerability in the system to run malware
-%% By prompting the targeted device to perform code execution, a hacker
-%% can run arbritrary code to
-%% their own programming in its
-%% place. This programming can then enable them to gain full access,
-%% steal data, carry out a full distributed denial of service (DDoS)
-%% attack, destroy files and infrastructure, or engage in illegal
-%% activity. And as the term remote execution suggests, an RCE
-%% cyberattack can take place from any geophysical location."\\
-
-\begin{figure}[H]
-  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
-  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code Execution.}
-\end{figure}
-
-%A opção  define que o match só será feito em pacotes que fazem parte
-%de uma conexão já estabelecida entre um cliente e um servidor. Além
-%disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.
-
-Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
-JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
-indicates that a rule match will only occur on packets that (1) flow
-on an already-established connection and that (2) flow to the server
-as opposed to the client. For the content options the modifiers
-\CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
-the matching not be case sensitive and that the string should be
-located at the packet field ``Request URI''. \Mar{Lucas, ``nocase''
-  tambem eh considerado modificador? Isto nao eh confuso? Na secao
-  4.3.4 consideramos modificadores como sinonimo para qualificador de
-  campo, nao?}\Luc{Sim, ``nocase'' define que a busca pela string nao deve ser case sensitive. Em 4.3.4 estamos considerando apenas os modificadores que definem posições/campos de pacotes http.}
-
-\tname{} reproduces the malicious JBoss attack using the traffic from
-a Nikto script~\cite{nikto}. However, similar JBoss attack is also
-available from the Metasploit database~\cite{metasploit}. As in the
-other examples, \tname{} uses, as additional inputs, benign traffic
-from \Fix{...} and a set of existing Suricata rules from
-\Fix{...}. \tname{} initially produces one seed rule containing
-\Fix{xx} options, by analyzing the malicious traffic alone. Then, it
-uses that seed rule to derive \Fix{yy} rules, each rule containing
-subsets of the option in the initial rule. Finally, it ranks that
-list. The list below shows the top-5 rules that \tname{}
-produces.\Mar{<- check/revise}
-
-\Fix{list rules}
-
-Note that the golden rule is in position \Fix{...} on this
-case. \Mar{explain syntactical difference, if exist}
-
-\subsection{Denial-of-Service}
-\label{sec:dos}
-
-The SYN Flood attack is a denial-of-service attack that exploits a vulnerability in the TCP/IP handshake
-to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
-works as follows in normal circumstances. First, a client sends a SYN
-packet to the server, requesting a connection. Second, the server
-responds with a SYN-ACK packet to the client. Third, the client
-responds with an ACK message and the connection is established. Aware
-of the protocol, an attacker sends multiple SYN packets to different
-ports of a server, often using fake IP addresses. Then, after the
-server responds with a SYN-ACK packet, the client keeps sending other
-SYN packets to avoid the connection to time out. Without proper
-protection, the server accepts these malicious requests and eventually
-legitimate requests cannot be satisfied due to resource exhaustion.
-Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
-this attack. This rule was obtained from the Emerging Threats Open
-Ruleset~\cite{emerging-threats-open}. The relevant options in the
-rule appear in bold---the option \CodeIn{flags: S,12}, identifies a
-SYN packet in a TCP packet and the option \CodeIn{threshold: type
-  both, track by\_dst, count 5000, seconds 5} indicates that a high
-volume of such packets should be requested in a short period of
-time. The other options are for documentation purposes.
-
-...........
-In this case, however, that information alone is
-insufficient as the malicious traffic is characterized as a sequence
-of messages. 
-...........
-
-%% \Mar{...and where did they come from?} \Luc{Sim, floods podem ser compostos por milhares de pacotes, um dos datasets q encontrei possui mais de 500k pacotes de flood.
-Os datasets foram encontrados na internet.}
-
-\begin{figure}[h]
-  \vspace{-2ex}
-  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window}]{synflood.suricata.synth}
-%  \caption{Suricata rule for SYN Flood Attacks.}
-  %  \label{fig:synflood-example-synt}
-  \vspace{-2ex}  
-\end{figure}
-
-%% \Mar{@Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
-%%   Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
-%%   respondam isto logo.}  \Gui{O window:64 seria removido
-%%   sim.}\Mar{ok. vou explicar isto.}
-  %% \Luc{O segundo campo da opção flags define flags que serão
-  %%   ignoradas. Nesse caso a regra procura por pacotes onde apenas a
-  %%   flag S está 'setada', independentemente do valor das flags 1 e
-  %%   2. Portanto a regra gerada pode capturar coisas diferentes da
-  %%   regra original. A inferencia do "12" n pode ser feita apenas a
-  %%   partir do pacote de entrada, precisariamos tratar
-  %%   isso.}\Mar{vc.
-pode explicar isto **mais detalhadamente** \Luc{a opção flags pode ter dois campos separados por virgula (ex: flags:S,12).
-o primeiro campo indica quais flags devem estar ativas em um pacote para que o suricata gere um alerta.
-se uma ou mais flags forem irrelevantes para a detecção de um determinado ataque (ou seja, se n importa se a flag está ou não ativa), elas serão adicionadas ao segundo campo.} E (1)
-  %%   explicar que isto eh uma limitacao \Luc{n entendi exatamente oq explicar aqui.} e (2) indicar a possivel consequencia
-  %%   disto?
-\Luc{a consequencia sao possiveis falsos negativos.
-ao não incluir flags no segundo campo podemos descartar pacotes por conta da presença de flags q n são relevantes (ou seja, vamos levar em conta flags q n deveriam interferir na detecção).}}
-
-\noindent
-
-.....
-
- The list below shows the rule options for the top-5 rules
-\tname{} produces for this attack.\Mar{algum problema em fazer isto?
-  exemplos concretos ajudam muito!} \Luc{No caso desse ataque, a regra inicial possui apenas duas opcoes (sem contar o threshold).
-Entao existem apenas 3 combinacoes possiveis, dessas 3 apenas 2 sao capazes de isolar o ataque do trafego normal.
-Portanto, o top-2 eh:
-  \begin{figure}[h]
-    \vspace{-2ex}
-    \lstinputlisting[language=C,numbers=none,frame=none,keywords={window,flags}]{top-synflood.rules}
-    \vspace{-3ex}  
-  \end{figure}
-
-   }\Mar{Lucas, eu nao entendo porque vc. desconsiderou threshold da
-  discussao (afinal, a gente destaca thrshold como relevante no inicio
-  desta secao). tambem nao entendo porque window---que a gente disse
-  acima que *nao* era relevante---aparece nas duas regras
-  geradas. enfim, isto esta inconsistente com nosso discurso.}
-
-\vspace{1ex}
-\noindent
-\textbf{Differences.}~Note that there are differences between the
-options produced in the first step of \tname{} and the options from
-the golden rule, as per Figure~\ref{fig:synflood-example}. The option
-\CodeIn{flags} does not mention parameters \CodeIn{1 2}, there is an
-extra option \CodeIn{window}, and there are differences in the option
-\CodeIn{threshold}. Considering the option \CodeIn{flags}, the first
-parameter of that option indicates what flags in the packet must be
-set while the second parameter indicates non-mandatory flags (\ie{},
-``don't cares''). More specifically, the option \CodeIn{flags: S}
-indicates that the flag ``S'' must be set and all flags not mentioned
-(including ``1'' and ``2'') must be unset for the \nids\ to capture
-the traffic. Consequently, that choice of option alone could result in
-false negatives and false positives. An attack that satisfies all
-other conditions in a rule and has flags ``1'' or ``2'' set would be
-captured by the original rule but not captured by the inferred rule---a case of
-false negative. Likewise, an attack that satisfies all conditions and
-have flags ``1'' and ``2'' unset would be captured by the inferred
-rule but not captured by the original rule---a case of false positive. The set of
-positive examples is to blame in this case. To mitigate this issue,
-the input messages should have covered cases with options ``1'' or
-``2'' set. Considering option \CodeIn{window: 512}, \tname{} is able
-to generate rules without that option, as discussed above. Considering
-option \CodeIn{threshold}, the values used in the golden rule are
-domain-specific. \tname{} uses default values for these options. We
-conjecture that the system administrator would be able to configure
-this option.
-
-%% \begin{table}[h]
-%%   \caption{\label{table:rules}Rule evolution}  
-%%   \centering
-%%   \begin{tabular}{lllll}
-%%     \toprule
-%%     \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
-%%     \midrule     
-%%     0 & (ack:0; seq:0; window:0; flags:F;)\\
-%%     5 & (window:0; flags:F;)\\
-%%     68 & (window:64; flags:F;)\\
-%%     73 & (window:64; flags:S,12;)\\
-%%     94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
-%%     \bottomrule
-%%   \end{tabular}
-%% \end{table}
+first step to create a potentially overspecified seed rule. Then, it
+uses benign traffic in the second step to create a set of plausible
+rules from the seed rule. Finally, in the third step, \tname{} uses
+existing rules to learn heuristics to rank plausible rules.
 
@@
 
-External Validity. \Fix{elaborate ...how popular are Suricata/Metasploit?}
+External Validity. \Fix{elaborate ...how popular are
+  Suricata/Metasploit?}
+
+\section{Discussion}
+
+\subsection{Sample of other attacks}
+
+\subsubsection{Remote Code Execution (RCE)}
+\label{sec:rce}
+\label{sec:jboss}
+\label{sec:content-example}
+
+%% Adaptacao do texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution
+
+RCE refers to the family of attacks where hackers exploit a network
+vulnerability to run arbitrary code---that could damage the system or
+steal sensitive information---on a remote machine. JBoss~\cite{jboss}
+is popular open source Java application server. This attacks enables
+remote execution of code on the JBoss application server.
+\Mar{@Lucas, por favor, observe o padrao usado nos exemplos
+  anteriores. eh preciso explicar em alto nivel o ataque antes de
+  explicar a regra!}
+
+%% In an RCE attack, hackers
+%% intentionally exploit a vulnerability in the system to run malware
+%% By prompting the targeted device to perform code execution, a hacker
+%% can run arbritrary code to
+%% their own programming in its
+%% place. This programming can then enable them to gain full access,
+%% steal data, carry out a full distributed denial of service (DDoS)
+%% attack, destroy files and infrastructure, or engage in illegal
+%% activity. And as the term remote execution suggests, an RCE
+%% cyberattack can take place from any geophysical location."\\
+
+\begin{figure}[H]
+  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
+  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code Execution.}
+\end{figure}
+
+%A opção  define que o match só será feito em pacotes que fazem parte
+%de uma conexão já estabelecida entre um cliente e um servidor. Além
+%disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.
+
+Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
+JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
+indicates that a rule match will only occur on packets that (1) flow
+on an already-established connection and that (2) flow to the server
+as opposed to the client. For the content options the modifiers
+\CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
+the matching not be case sensitive and that the string should be
+located at the packet field ``Request URI''. \Mar{Lucas, ``nocase''
+  tambem eh considerado modificador? Isto nao eh confuso? Na secao
+  4.3.4 consideramos modificadores como sinonimo para qualificador de
+  campo, nao?}\Luc{Sim, ``nocase'' define que a busca pela string nao deve ser case sensitive. Em 4.3.4 estamos considerando apenas os modificadores que definem posições/campos de pacotes http.}
+
+\tname{} reproduces the malicious JBoss attack using the traffic from
+a Nikto script~\cite{nikto}. However, similar JBoss attack is also
+available from the Metasploit database~\cite{metasploit}. As in the
+other examples, \tname{} uses, as additional inputs, benign traffic
+from \Fix{...} and a set of existing Suricata rules from
+\Fix{...}. \tname{} initially produces one seed rule containing
+\Fix{xx} options, by analyzing the malicious traffic alone. Then, it
+uses that seed rule to derive \Fix{yy} rules, each rule containing
+subsets of the option in the initial rule. Finally, it ranks that
+list. The list below shows the top-5 rules that \tname{}
+produces.\Mar{<- check/revise}
+
+\Fix{list rules}
+
+Note that the golden rule is in position \Fix{...} on this
+case. \Mar{explain syntactical difference, if exist}
+
+\subsubsection{Denial-of-Service}
+\label{sec:dos}
+
+The SYN Flood attack is a denial-of-service attack that exploits a vulnerability in the TCP/IP handshake
+to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
+works as follows in normal circumstances. First, a client sends a SYN
+packet to the server, requesting a connection. Second, the server
+responds with a SYN-ACK packet to the client. Third, the client
+responds with an ACK message and the connection is established. Aware
+of the protocol, an attacker sends multiple SYN packets to different
+ports of a server, often using fake IP addresses. Then, after the
+server responds with a SYN-ACK packet, the client keeps sending other
+SYN packets to avoid the connection to time out. Without proper
+protection, the server accepts these malicious requests and eventually
+legitimate requests cannot be satisfied due to resource exhaustion.
+Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
+this attack. This rule was obtained from the Emerging Threats Open
+Ruleset~\cite{emerging-threats-open}. The relevant options in the
+rule appear in bold---the option \CodeIn{flags: S,12}, identifies a
+SYN packet in a TCP packet and the option \CodeIn{threshold: type
+  both, track by\_dst, count 5000, seconds 5} indicates that a high
+volume of such packets should be requested in a short period of
+time. The other options are for documentation purposes.
+
+...........
+In this case, however, that information alone is
+insufficient as the malicious traffic is characterized as a sequence
+of messages. 
+...........
+
+%% \Mar{...and where did they come from?} \Luc{Sim, floods podem ser compostos por milhares de pacotes, um dos datasets q encontrei possui mais de 500k pacotes de flood.
+Os datasets foram encontrados na internet.}
+
+\begin{figure}[h]
+  \vspace{-2ex}
+  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window}]{synflood.suricata.synth}
+%  \caption{Suricata rule for SYN Flood Attacks.}
+  %  \label{fig:synflood-example-synt}
+  \vspace{-2ex}  
+\end{figure}
+
+%% \Mar{@Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
+%%   Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
+%%   respondam isto logo.}  \Gui{O window:64 seria removido
+%%   sim.}\Mar{ok. vou explicar isto.}
+  %% \Luc{O segundo campo da opção flags define flags que serão
+  %%   ignoradas. Nesse caso a regra procura por pacotes onde apenas a
+  %%   flag S está 'setada', independentemente do valor das flags 1 e
+  %%   2. Portanto a regra gerada pode capturar coisas diferentes da
+  %%   regra original. A inferencia do "12" n pode ser feita apenas a
+  %%   partir do pacote de entrada, precisariamos tratar
+  %%   isso.}\Mar{vc.
+pode explicar isto **mais detalhadamente** \Luc{a opção flags pode ter dois campos separados por virgula (ex: flags:S,12).
+o primeiro campo indica quais flags devem estar ativas em um pacote para que o suricata gere um alerta.
+se uma ou mais flags forem irrelevantes para a detecção de um determinado ataque (ou seja, se n importa se a flag está ou não ativa), elas serão adicionadas ao segundo campo.} E (1)
+  %%   explicar que isto eh uma limitacao \Luc{n entendi exatamente oq explicar aqui.} e (2) indicar a possivel consequencia
+  %%   disto?
+\Luc{a consequencia sao possiveis falsos negativos.
+ao não incluir flags no segundo campo podemos descartar pacotes por conta da presença de flags q n são relevantes (ou seja, vamos levar em conta flags q n deveriam interferir na detecção).}}
+
+\noindent
+
+.....
+
+ The list below shows the rule options for the top-5 rules
+\tname{} produces for this attack.\Mar{algum problema em fazer isto?
+  exemplos concretos ajudam muito!} \Luc{No caso desse ataque, a regra inicial possui apenas duas opcoes (sem contar o threshold).
+Entao existem apenas 3 combinacoes possiveis, dessas 3 apenas 2 sao capazes de isolar o ataque do trafego normal.
+Portanto, o top-2 eh:
+  \begin{figure}[h]
+    \vspace{-2ex}
+    \lstinputlisting[language=C,numbers=none,frame=none,keywords={window,flags}]{top-synflood.rules}
+    \vspace{-3ex}  
+  \end{figure}
+
+   }\Mar{Lucas, eu nao entendo porque vc. desconsiderou threshold da
+  discussao (afinal, a gente destaca thrshold como relevante no inicio
+  desta secao). tambem nao entendo porque window---que a gente disse
+  acima que *nao* era relevante---aparece nas duas regras
+  geradas. enfim, isto esta inconsistente com nosso discurso.}
+
+\vspace{1ex}
+\noindent
+\textbf{Differences.}~Note that there are differences between the
+options produced in the first step of \tname{} and the options from
+the golden rule, as per Figure~\ref{fig:synflood-example}. The option
+\CodeIn{flags} does not mention parameters \CodeIn{1 2}, there is an
+extra option \CodeIn{window}, and there are differences in the option
+\CodeIn{threshold}. Considering the option \CodeIn{flags}, the first
+parameter of that option indicates what flags in the packet must be
+set while the second parameter indicates non-mandatory flags (\ie{},
+``don't cares''). More specifically, the option \CodeIn{flags: S}
+indicates that the flag ``S'' must be set and all flags not mentioned
+(including ``1'' and ``2'') must be unset for the \nids\ to capture
+the traffic. Consequently, that choice of option alone could result in
+false negatives and false positives. An attack that satisfies all
+other conditions in a rule and has flags ``1'' or ``2'' set would be
+captured by the original rule but not captured by the inferred rule---a case of
+false negative. Likewise, an attack that satisfies all conditions and
+have flags ``1'' and ``2'' unset would be captured by the inferred
+rule but not captured by the original rule---a case of false positive. The set of
+positive examples is to blame in this case. To mitigate this issue,
+the input messages should have covered cases with options ``1'' or
+``2'' set. Considering option \CodeIn{window: 512}, \tname{} is able
+to generate rules without that option, as discussed above. Considering
+option \CodeIn{threshold}, the values used in the golden rule are
+domain-specific. \tname{} uses default values for these options. We
+conjecture that the system administrator would be able to configure
+this option.
+
+%% \begin{table}[h]
+%%   \caption{\label{table:rules}Rule evolution}  
+%%   \centering
+%%   \begin{tabular}{lllll}
+%%     \toprule
+%%     \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
+%%     \midrule     
+%%     0 & (ack:0; seq:0; window:0; flags:F;)\\
+%%     5 & (window:0; flags:F;)\\
+%%     68 & (window:64; flags:F;)\\
+%%     73 & (window:64; flags:S,12;)\\
+%%     94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
+%%     \bottomrule
+%%   \end{tabular}
+%% \end{table}
 

### 2020-02-23 paper/paper.tex
pass on 3

@@
 distribution of the number of content options for the database of
-rules we used\Mar{reference}. Note that rules with 1 to 3 options are
-more common.  At the end of the process, \tname{} ranks the golden
-rule at the top of the ranking including \Fix{XXXXXX} other plausible
-rules.
+rules we used\Mar{reference} \Luc{essa distribuicao n eh utilizada 
+para pontuar as regras desse exemplo, n seria mais interessante citar outra?
+poderia ser a de tamanhos das regras.}. Note that rules with 1 to 3 options 
+are more common \Luc{1 to 3 \textbf{content} options}. At the end of the 
+process, \tname{} ranks the golden rule at the top of the ranking 
+including \Fix{XXXXXX} other plausible rules.
 

### 2020-02-23 paper/paper.tex
pass on table 3

@@
     \midrule
-    Ping Scan & - & - & - \\
-    SYN Flood & - & - & - \\
+    Ping Scan & 11 & 2 & - \\
+    SYN Flood & 2 & 1 & - \\
     JBoss JMX & 382 & 208 & 23 \\

### 2020-02-23 paper/paper.tex
done my pass on section 3. moving to section 4.

@@
 of the options from the seed rule. Considering this example, \tname{}
-produces a total of \Fix{XXXXX} plausible rules at the end of the
-search.\Mar{Technically, there are $32 (=2^5)$ subsets for a set with 5
-  elements. So, it is important to indicate why Table 3 shows sooo
-  many plausible rules and that this is is a very simple rule (for illustration).}
+produces a total of \pingscanplausible{} plausible rules at the end of
+the search.  In the worst case, \tname{} produces $2^n$ plausible
+rules, where $n$ is the number of options contained in the inferred
+rule ($n=5$, in this case). Rules are discared in the search process
+as they are found to be implausible.
+
+%% \Mar{Technically, there are $32 (=2^5)$ subsets for a set
+%%   with 5 elements. So, it is important to indicate why Table 3 shows
+%%   sooo many plausible rules and that this is is a very simple rule
+%%   (for illustration).}
 
@@
 rules should share characteristics with pre-existing rules. For
-example, one heuristic function used by \tname{} leverages the amount
-of options a rule typically has. The score of a rule, given by this
-function, is proportional to the popularity of the number of options
+example, one heuristic function used by \tname{} leverages the
+\emph{size of a rule} for that. The score of a rule given by such a
+function is proportional to the popularity of the number of options
 declared in rules of a given
-protocol. Figure~\ref{fig:distribution-contents} shows the
-distribution of the number of content options for the database of
-rules we used\Mar{reference} \Luc{essa distribuicao n eh utilizada 
-para pontuar as regras desse exemplo, n seria mais interessante citar outra?
-poderia ser a de tamanhos das regras.}. Note that rules with 1 to 3 options 
-are more common \Luc{1 to 3 \textbf{content} options}. At the end of the 
-process, \tname{} ranks the golden rule at the top of the ranking 
-including \Fix{XXXXXX} other plausible rules.
-
-\Mar{add short par indicating that other problems can be much more
-  complex than this...}
+protocol. Figure~\ref{fig:distribution-rule-size} shows the
+distribution of the number of content options associated with rules
+from a public data set~\cite{emerging-threats-open}. Note that rules
+with 2 to 6 options are more common. At the end of this step, \tname{}
+ranks the golden rule at the top of the ranking, including
+\pingscanplausible{} rules. It is worth noting that
+\tname\ consistently ranks the golden rule very high in most of the
+cases even when the seed rule, produced at the first step, contains a
+large number of options (see Table~\ref{table:results}).
 
@@
 the previous step. Third, \tname{} uses several heuristic functions
-derived from a dataset of rules to rank the rules generated in the
+derived from a data set of rules to rank the rules generated in the
 previous step.
@@
 
-\tname{} leverages public datasets~\cite{tcpreplay} of benign traffic
+\tname{} leverages public data sets~\cite{tcpreplay} of benign traffic
 to guide the search for more general rules, hence addressing the
@@
 negative traffic. In this case, however, \tname{} can use public
-datasets available online.
+data sets available online.
 
@@
 low. Figure~\ref{fig:distribution-rule-size} shows the histogram of
-rule sizes for our rule dataset. The x-axis shows the number of
+rule sizes for our rule data set. The x-axis shows the number of
 options included in a rule (up to 10) while the y-axis shows the
@@
   melhor, sao cerca de 70K contents nas 28K regras do
-  dataset}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
+  data set}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
   (\%) das 28K regras tem contents.} \Luc{sao 69417 contents nas 26094 regras q utilizam contents} As consequence, we defined a
@@
 rule. Figure~\ref{fig:distribution-strings} shows the distribution of
-the ten most popular strings in our rule dataset. To illustrate this
+the ten most popular strings in our rule data set. To illustrate this
 function, let us consider a rule with two content options, using the
@@
 %% we observed---by analyzing
-%% the dataset of existing rules (see
-%% Section~\ref{sec:dataset-benign})---that these rules often indicate
+%% the data set of existing rules (see
+%% Section~\ref{sec:data set-benign})---that these rules often indicate
 %% what part of the message the content string should be checked
@@
 Suricata, currently, \tname\ only supports tcp and http. These are the
-protocols involving ~80\% of the rules from our dataset.
+protocols involving ~80\% of the rules from our data set.
 
@@
 A ideia aqui basicamente eh: se um certo content do payload está numa posição q eh muito verificada nas regras, possivelmente esse content eh mais relevante do q outro q está numa posição q raramente eh verificada.
-Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no dataset.}
+Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no data set.}
 %%   \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content.
@@
 traffic for all protocols covered by Suricata (see
-Figure~\ref{fig:distribution-rules-protocol}). We used two datasets of
-benign traffic. The Bigflows.pcap dataset is part of the
-TcpReplay~\cite{tcpreplay} open source project. This dataset includes
+Figure~\ref{fig:distribution-rules-protocol}). We used two data sets of
+benign traffic. The Bigflows.pcap data set is part of the
+TcpReplay~\cite{tcpreplay} open source project. This data set includes
 real network traffic of a busy private network access point to the
@@
 (smallFlows.pcap). It also has many more flows and different
-applications.''. In addition to the Bigflows.pcap dataset, we also
-used datasets of normal traffic made available by Stratosphere
+applications.''. In addition to the Bigflows.pcap data set, we also
+used data sets of normal traffic made available by Stratosphere
 Lab~\cite{stratosphere-normal}.
@@
 
-%% \Mar{...and where did they come from?} \Luc{Sim, floods podem ser compostos por milhares de pacotes, um dos datasets q encontrei possui mais de 500k pacotes de flood.
-Os datasets foram encontrados na internet.}
+%% \Mar{...and where did they come from?} \Luc{Sim, floods podem ser
+%% compostos por milhares de pacotes, um dos data sets q encontrei
+%% possui mais de 500k pacotes de flood. Os data set  s foram encontrados na internet.}
 

### 2020-02-24 paper/paper.tex
updated table 3

@@
   \centering
-  \begin{tabular}{lrrrp{9cm}}
+  \begin{tabular}{lcrp{9cm}}
     \toprule
     \multicolumn{1}{c}{Name} &
-    \multicolumn{1}{c}{\#~Rules created} &
-    \multicolumn{1}{c}{Rank} &
-    \multicolumn{1}{c}{Rank fitness global} \\
+    \multicolumn{1}{c}{\#~Options Seed Rule} &
+    \multicolumn{1}{c}{Rank} \\
     \midrule
-    Ping Scan & 11 & 2 & - \\
-    SYN Flood & 2 & 1 & - \\
-    JBoss JMX & 382 & 208 & 23 \\
-    WeBid & - & - & - \\    
-    ColdFusion & 34 & 5 & 5\\
-    .htaccess access & 2107 & 365 & 2\\
-    .idq access & 1112 & 429 & 2 \\
-    iisadmin access & 862 & 372 & 1 \\
-    /system32/ in URI & 3665 & 783 & 2 \\
-    Script tag in URI & - & - & - \\
+    Ping Scan & 5 & - \\
+    SYN Flood & - & - \\
+    JBoss JMX & - & 23 \\
+    WeBid & - & -\\    
+    ColdFusion & - & 5\\
+    .htaccess access & - & 2\\
+    .idq access & - & 2 \\
+    iisadmin access & - & 1 \\
+    /system32/ in URI & - & 2 \\
+    Script tag in URI & - & - \\
     \bottomrule

### 2020-02-24 paper/paper.tex
table 3 values

@@
     Ping Scan & 5 & - \\
-    SYN Flood & - & - \\
-    JBoss JMX & - & 23 \\
-    WeBid & - & -\\    
-    ColdFusion & - & 5\\
-    .htaccess access & - & 2\\
-    .idq access & - & 2 \\
-    iisadmin access & - & 1 \\
-    /system32/ in URI & - & 2 \\
-    Script tag in URI & - & - \\
+    SYN Flood & 2 & - \\
+    JBoss JMX & 23 & 23 \\
+    WeBid & 23 & -\\    
+    ColdFusion & 19 & 5\\
+    .htaccess access & 17 & 2\\
+    .idq access & 19 & 2 \\
+    iisadmin access & 17 & 1 \\
+    /system32/ in URI & 20 & 2 \\
+    Script tag in URI & 22 & - \\
     \bottomrule

### 2020-02-24 paper/paper.tex
pass on section 4

@@
 
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[trim=0 395 50 0,clip,width=0.65\textwidth]{figs/nids-workflow}
+  \caption{\tname\ workflow.}
+  \label{fig:overview}
+  \vspace{-2ex}
+\end{figure*}
+
 \section{Technique}
 \label{sec:technique}
-
-\begin{figure*}[ht!]
-  \centering
-  \includegraphics[trim=0 350 100 0,clip,width=0.65\textwidth]{figs/nids-workflow}
-  \caption{\tname\ workflow.}
-  \label{fig:overview}
-  \vspace{-1ex}
-\end{figure*}
 
@@
 set of rules from public rulesets and produces on output a list of
-plausible rules to isolate the malicious from the benign
-traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
-pipeline of three components, which appear numbered in the figure.
-The inputs of the technique appear at the bottom of the figure whereas
-the output appears at the right.
+\emph{plausible rules}---\ie{}, rules that isolate the malicious
+traffic from the benign traffic---in order of their likelihood to
+reflect user's intent. Figure~\ref{fig:overview} shows the workflow of
+\tname{} as a pipeline of three components, which appear numbered in
+the figure.  Public data sets of benign
+traffic~\cite{tcpreplay,stratosphere-normal} and \nids\ rules
+exist~\cite{emerging-threats-open}. Consequently, a user of \tname{}
+only needs to provide the malicious traffic on input.
 
 \tname\ works as follows. First, it produces a potentially
-overspecified rule that captures the malicious traffic, provided on
-input in the form of pcap files~\cite{pcap}. In this step,
+overspecified rule that captures the malicious traffic, which is
+provided on input in the form of pcap files~\cite{pcap}. In this step,
 \tname\ detects rule options directly from the input messages (see
@@
 considered in this step, the rule produced is likely overspecified,
-\ie{}, it includes more options than the golden rule (see
+\ie{}, it may be more restrictive than the golden rule (see
 Section~\ref{sec:suri-metas-coverage}). Consequently, the use of that
-rule could lead to false negatives as not all manifestations of the
-same attack are identical. Second, \tname{} generates plausible rules
-that are similar to the seed rule produced in the previous step. It is
-worth noting that any subset of the original set of options captures
-the malicious traffic. \tname{} leverages that observation to search
-for rules containing a proper subset of rules defined in the original
-seed option. As the generation procedure discards options from the
-initial rule, the rules \tname{} creates satisfies the property of
-capturing malicious traffic. However, the goal of \tname\ is to
-produce rules that capture \emph{only} malicious traffic. To achieve
-that goal, \tname{} uses benign traffic to reduce the likelihood the
-\nids\ will report false positives on the synthesized rule. The
-minimization step uses a population-based gradient descent-like search
-to find all plausible rules derived from the seed rule, obtained in
-the previous step. Third, \tname{} uses several heuristic functions
-derived from a data set of rules to rank the rules generated in the
-previous step.
+rule could lead to false negatives. Second, \tname{} generates rules
+that weakens the constraints expressed in the inferred seed rule
+whilest still capturing the malicious traffic. Given that every option
+declared in a rule needs to be satisfied for a rule to capture an
+attack, any subset of the set of options---inferred in the first
+step---captures the malicious traffic by definition. \tname{}
+leverages that observation to search for rules containing a proper
+subset of rules defined in the original seed option. Although that
+approach captures the malicious traffic, it can result in false
+positives. Intuitively, there are no restrictions on to what extent a
+rule can be relaxed. \tname{} uses benign traffic for that
+purpose. The minimization step uses a population-based gradient
+descent-like search to find all plausible rules derived from the seed
+rule, obtained in the previous step. Finally, \tname{} uses several
+heuristic functions derived from a data set of existing rules to rank
+the plausible rules generated in the previous step.
+
+%%  As the generation
+%% procedure discards options from the initial rule, the rules \tname{}
+%% creates satisfy the property of capturing malicious traffic.
 
@@
 \subsection{Step 1: Reverse Engineering}
+
+\Mar{----------------------------------- PAREI AQUI}
 

### 2020-02-24 paper/paper.tex
another pass on technique overview

@@
 provided on input in the form of pcap files~\cite{pcap}. In this step,
-\tname\ detects rule options directly from the input messages (see
-Section~\ref{sec:rules-and-packets}). As only malicious traffic is
-considered in this step, the rule produced is likely overspecified,
-\ie{}, it may be more restrictive than the golden rule (see
-Section~\ref{sec:suri-metas-coverage}). Consequently, the use of that
-rule could lead to false negatives. Second, \tname{} generates rules
-that weakens the constraints expressed in the inferred seed rule
-whilest still capturing the malicious traffic. Given that every option
-declared in a rule needs to be satisfied for a rule to capture an
-attack, any subset of the set of options---inferred in the first
-step---captures the malicious traffic by definition. \tname{}
-leverages that observation to search for rules containing a proper
-subset of rules defined in the original seed option. Although that
-approach captures the malicious traffic, it can result in false
-positives. Intuitively, there are no restrictions on to what extent a
-rule can be relaxed. \tname{} uses benign traffic for that
-purpose. The minimization step uses a population-based gradient
-descent-like search to find all plausible rules derived from the seed
-rule, obtained in the previous step. Finally, \tname{} uses several
-heuristic functions derived from a data set of existing rules to rank
-the plausible rules generated in the previous step.
+\tname\ detects options directly from the input messages (see
+Section~\ref{sec:rules-and-packets}). More precisely, it produces a
+rule containing all possible options satisfied by the traffic. Let us
+refer to that rule as the \emph{seed rule} and its corresponding set
+of options as \emph{seed options}. Note that, as only malicious
+traffic is considered in this step, the seed rule is likely
+overspecified, \ie{}, it may be more restrictive than the golden rule
+(see Section~\ref{sec:suri-metas-coverage}). Consequently, the use of
+that rule to capture attacks associated with the input traffic could
+lead to false negatives.  Second, \tname\ generates rules that weaken
+the constraints associated with the seed options whilst still
+capturing the malicious traffic. Given that every option declared in a
+rule needs to be satisfied for a rule to capture an attack, any subset
+of the set of seed options must capture the malicious traffic by
+definition. \tname{} leverages that observation to search for rules
+containing a proper subset of options from the seed options. However,
+note that that approach alone could result in false positives as it
+imposes no limit on to what extent a rule can be relaxed. \tname{}
+uses benign traffic for that purpose. More precisely, it uses a
+gradient descent-like search to find all plausible rules derived from
+the seed rule, obtained in the previous step. Finally, \tname{} uses
+several heuristic functions derived from a data set of existing rules
+to rank the plausible rules generated in the previous step by their
+similarity to existing rules. The rationale is that many plausible
+rules can be generated in this step and they have different semantics,
+which can differ from user's intent. Similarity is used to approximate
+user's intent.
 

### 2020-02-24 paper/paper.tex
updating table

@@
 
-\Mar{----------------------------------- PAREI AQUI}
-
-The goal of the first step is to extract options from the positive
-traffic, which is characterized with pcap files~\cite{pcap}, a popular
-format to describe network packets.
-
-For that, \tname{} parses the message looking for options associated
-with the data. As described in Section~\ref{sec:rules-and-packets},
-there is a correspondence between packet fields and rule
-options. \tname{} leverages that mapping to infer options. The
-examples of sections~\ref{sec:dos} and~\ref{sec:active-recon} show
-cases where \tname{} infers options with this mapping. Unfortunately,
-inferring options at the field granularity level is insufficient to
-capture content options as those described on
-Section~\ref{sec:content-example}. These options refer to parts of the
-payload of the message and cannot be extracted directly from the
-fields. Handling this case is critically important as 93\% of the
-rules contain \CodeIn{content} options.
-
-Figure~\ref{fig:distribution-contents} shows the histogram of the
-number of these options per rule obtained from a public
-ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
-rules. We limited the number of content options per rule to 10 for
-space. Rules without \CodeIn{content} options is an exception
-(7\%) and half of the rules contain at least two
-\CodeIn{content} options.  To find the values associated with these
-options, \tname{} splits the payload of the malicious message in
-tokens using natural language delimiters, such as $\backslash$t,
-$\backslash$n, $\backslash$r, and spaces. We observed, by inspecting
-existing rules, that \CodeIn{content} options typically refer to
-sequence of characters in the payload of messages separated by these
-delimiters. This tokenization step can generate dozens of options.
-\tname{} produced a total of \Fix{xx} content options for the example
-Section~\ref{sec:content-example}.
-
-\Mar{Revisar.........................}
-\Fix{
-A avaliação que fazemos aqui pontua um dado content de acordo com
-quantas vezes o campo em que ele se encontra eh referenciado nas
-regras. Por exemplo, vamos supor que o tokenizer produziu a opção
-content: "/HtmlAdaptor". Essa string pertence ao campo "Request
-URI:". Esse campo pode ser acessado pelo modificador http\_uri. O
-tokenizer associa o content a esse modificador.
-}
+The first step of \tname\ is Reverse Engineering. The goal of this
+step is to extract options from the malicious traffic. A rule that
+contains these options, by construction, captures the malicious
+traffic. \tname\ uses two strategies to extract options: 1)~field and
+2)~payload. We elaborate these strategies in the following.
+
+For the field strategy, \tname{} parses the input messages looking for
+options associated with the data in the fields of a network
+message. Section~\ref{sec:rules-and-packets} shows the correspondence
+between packet fields and some of the rule options from
+\suri. Section~\ref{sec:active-recon} shows a concrete example where
+the values for options are directly extracted from packet fields. We
+constructed the mapping of network fields to options manually, from
+the \suri\ documentation.
+
+Unfortunately, the relevant data for the option may be at the message
+payload not the fields. Figure~\ref{fig:rule-jboss} shows an example
+of a rule that captures traffic containing the string
+\CodeIn{"/HtmlAdaptor"} in the payload.  The content strategy focuses
+on extracting (content) options from the payload. It is very important
+to note that content options are prevalent. For example,
+\percRulesWithContent{} of the \numrulessuri{} rules from the public
+data set~\cite{emerging-threats-open} we analyzed contain at least one
+option of this kind. Figure~\ref{fig:distribution-contents} shows the
+histogram of the number of these options per rule for the data set
+mentioned above. We limited the number of content options per rule to
+10 for space. Rules without \CodeIn{content} options are an exception
+and half of the rules contain at least two \CodeIn{content} options.
+To find the values associated with these options, \tname{} splits the
+payload of the malicious message in tokens using natural language
+delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
+spaces. We observed, by inspecting existing rules, that
+\CodeIn{content} options typically refer to sequence of characters in
+the payload of messages separated by these delimiters. This
+tokenization step can generate dozens of options.  For example,
+considering the example from Section~\ref{sec:content-example},
+\tname{} produces a total of \Fix{XXX} content options.
+
+\Mar{Nao esta muito claro como isto funciona. Rodamos sempre as duas
+  estrategias (para obter field options e content options)?
+  (Precisamos dizer se eh possivel ter regras com os dois tipos de
+  opcoes). Como isto funciona 1) quando uma ataque tem varias
+  mensagens (e.g., ping scan) e 2) quando temos varios ataques?}
+
+\textbf{Implementation.}\Fix{elaborate...}
+, characterized with pcap files~\cite{pcap}, a popular format
+to describe network packets.
+
+%% \Fix{
+%% A avaliação que fazemos aqui pontua um dado content de acordo com
+%% quantas vezes o campo em que ele se encontra eh referenciado nas
+%% regras. Por exemplo, vamos supor que o tokenizer produziu a opção
+%% content: "/HtmlAdaptor". Essa string pertence ao campo "Request
+%% URI:". Esse campo pode ser acessado pelo modificador http\_uri. O
+%% tokenizer associa o content a esse modificador.
+%% }
 
@@
 \label{sec:minimization}
+
+\Mar{------------------------------------------- Eu estou aqui}
 
@@
   \caption{\label{table:results}Results.}
+  \vspace{-2ex}
   \centering
-  \begin{tabular}{lcrp{9cm}}
+  \begin{tabular}{lrrr}
     \toprule
-    \multicolumn{1}{c}{Name} &
-    \multicolumn{1}{c}{\#~Options Seed Rule} &
+    \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
+    \#~Options &
+    \#~Plausible Rules &    
     \multicolumn{1}{c}{Rank} \\
+
+     &
+    \multicolumn{1}{c}{(step 1)} &
+    \multicolumn{1}{c}{(step 2)} &    
+    \multicolumn{1}{c}{(step 3)} \\
+
     \midrule
-    Ping Scan & 5 & - \\
-    SYN Flood & 2 & - \\
-    JBoss JMX & 23 & 23 \\
-    WeBid & 23 & -\\    
-    ColdFusion & 19 & 5\\
-    .htaccess access & 17 & 2\\
-    .idq access & 19 & 2 \\
-    iisadmin access & 17 & 1 \\
-    /system32/ in URI & 20 & 2 \\
-    Script tag in URI & 22 & - \\
+    Ping Scan & 5 & - & - \\
+    SYN Flood & 2 & - & - \\
+    JBoss JMX & 23 & - & 23 \\
+    WeBid & 23 & - & -\\    
+    ColdFusion & 19 & - & 5\\
+    .htaccess access & 17 & - & 2\\
+    .idq access & 19 & - & 2 \\
+    iisadmin access & 17 & - & 1 \\
+    /system32/ in URI & 20 & - & 2 \\
+    Script tag in URI & 22 & - & - \\
     \bottomrule

### 2020-02-25 paper/paper.tex
minor changes

@@
   \includegraphics[trim=0 395 50 0,clip,width=0.65\textwidth]{figs/nids-workflow}
-  \caption{\tname\ workflow.}
+  \caption{The \tname\ workflow.}
   \label{fig:overview}
@@
 
-The first step of \tname\ is Reverse Engineering. The goal of this
-step is to extract options from the malicious traffic. A rule that
-contains these options, by construction, captures the malicious
-traffic. \tname\ uses two strategies to extract options: 1)~field and
-2)~payload. We elaborate these strategies in the following.
+The first step in the \tname's workflow is Reverse Engineering. The
+goal of this step is to extract options from the malicious traffic. A
+rule that contains these options, by construction, captures the
+malicious traffic. \tname\ uses two strategies to extract options:
+1)~field (for any protocol) and 2)~payload (for http only). We
+elaborate these strategies in the following.
 
@@
 the values for options are directly extracted from packet fields. We
-constructed the mapping of network fields to options manually, from
-the \suri\ documentation.
+obtained the mapping of network fields to options manually, from the
+\suri\ documentation.
 
 Unfortunately, the relevant data for the option may be at the message
-payload not the fields. Figure~\ref{fig:rule-jboss} shows an example
-of a rule that captures traffic containing the string
+payload not the fields. We found empirically that this occurs
+frequently in http attacks. Figure~\ref{fig:rule-jboss} shows an
+example of a rule that captures traffic containing the string
 \CodeIn{"/HtmlAdaptor"} in the payload.  The content strategy focuses
-on extracting (content) options from the payload. It is very important
-to note that content options are prevalent. For example,
-\percRulesWithContent{} of the \numrulessuri{} rules from the public
-data set~\cite{emerging-threats-open} we analyzed contain at least one
-option of this kind. Figure~\ref{fig:distribution-contents} shows the
-histogram of the number of these options per rule for the data set
-mentioned above. We limited the number of content options per rule to
-10 for space. Rules without \CodeIn{content} options are an exception
-and half of the rules contain at least two \CodeIn{content} options.
-To find the values associated with these options, \tname{} splits the
+on extracting such kinds of options from the payload. It is very
+important to note that content options are prevalent in \suri\ rules
+and the majority of these rule prescribe defenses against http
+attacks. For example, \percRulesWithContent{} of the \numrulessuri{}
+rules from the public data set~\cite{emerging-threats-open} we
+analyzed contain at least one content
+option. Figure~\ref{fig:distribution-contents} shows the histogram of
+the number of these options per rule for the data set mentioned
+above. We limited the number of content options per rule to 10 for
+space. Furthermore, \Fix{xxx\%} of the rules are for
+http.\MyComment{Rules without \CodeIn{content} options are an
+  exception and half of the rules contain at least two
+  \CodeIn{content} options.}  Consequently, hangling content options
+for http is very important.
+
+We observed, by inspecting existing rules, that \CodeIn{content}
+options typically refer to sequence of characters in the payload. To
+find the values associated with these options, \tname{} splits the
 payload of the malicious message in tokens using natural language
 delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
-spaces. We observed, by inspecting existing rules, that
-\CodeIn{content} options typically refer to sequence of characters in
-the payload of messages separated by these delimiters. This
-tokenization step can generate dozens of options.  For example,
-considering the example from Section~\ref{sec:content-example},
-\tname{} produces a total of \Fix{XXX} content options.
-
-\Mar{Nao esta muito claro como isto funciona. Rodamos sempre as duas
-  estrategias (para obter field options e content options)?
-  (Precisamos dizer se eh possivel ter regras com os dois tipos de
-  opcoes). Como isto funciona 1) quando uma ataque tem varias
-  mensagens (e.g., ping scan) e 2) quando temos varios ataques?}
-
-\textbf{Implementation.}\Fix{elaborate...}
-, characterized with pcap files~\cite{pcap}, a popular format
-to describe network packets.
+spaces. Considering the example from
+Section~\ref{sec:content-example}, \tname{} produced a total of 23
+content options.
+
+It is worth noting that we found empirically that deciding which
+reverse engineering strategy to use based on the protocol was
+effective. We used the payload strategy for http and the field
+strategy for the other protocols. At the end of this step,
+\tname\ produces a seed rule that captures the malicious traffic by
+definition.
+
+%% \Mar{Nao esta muito claro como isto funciona. Rodamos sempre as duas
+%%   estrategias (para obter field options e content options)?
+%%   (Precisamos dizer se eh possivel ter regras com os dois tipos de
+%%   opcoes). Como isto funciona 1) quando uma ataque tem varias
+%%   mensagens (e.g., ping scan) e 2) quando temos varios ataques?}
+
+%% \textbf{Implementation.}~
+%% , characterized with pcap files~\cite{pcap}, a popular format
+%% to describe network packets.
 

### 2020-02-25 paper/paper.tex
comment to discuss flooding

@@
 effective. We used the payload strategy for http and the field
-strategy for the other protocols. At the end of this step,
+strategy for the other protocols.\Mar{discuss flood cases...} At the end of this step,
 \tname\ produces a seed rule that captures the malicious traffic by
@@
   \caption{Distribution of features.}
+  \Luc{@Marcelo, estamos utilizando mais uma distribuicao agora. Precisamos adicionar o grafico dela tb?}
 \end{figure*}
@@
     \midrule
-    Ping Scan & 5 & - & - \\
-    SYN Flood & 2 & - & - \\
-    JBoss JMX & 23 & - & 23 \\
+    Ping Scan & 5 & 2 & - \\
+    SYN Flood & 2 & 2 & - \\
+    JBoss JMX & 23 & 381 & 23 \\
     WeBid & 23 & - & -\\    
-    ColdFusion & 19 & - & 5\\
-    .htaccess access & 17 & - & 2\\
-    .idq access & 19 & - & 2 \\
-    iisadmin access & 17 & - & 1 \\
-    /system32/ in URI & 20 & - & 2 \\
+    ColdFusion & 19 & 39 & 5\\
+    .htaccess access & 17 & 2458 & 2\\
+    .idq access & 19 & 1111 & 2 \\
+    iisadmin access & 17 & 861 & 1 \\
+    /system32/ in URI & 20 & 3665 & 2 \\
     Script tag in URI & 22 & - & - \\

### 2020-02-25 simulated-annealing/sa.py
renaming output filenames

@@ -13,7 +13,7 @@ from functools import partial, reduce
 from itertools import combinations
 
 open("bad.rules", 'w').close()
-
+attacks_list = ["adaptor", "coldfusion", "htaccess", "idq", "issadmin", "system", "script", "pingscan", "synflood"]
 parser = argparse.ArgumentParser(description="Description.")
 parser.add_argument('attack', metavar='A')
 args = parser.parse_args()
@@ -66,7 +66,7 @@ else:
     pcapAttack = "Datasets/" + str(args.attack) + ".pcap"
 
 if args.attack == "pingscan":
-    pcapAttack = "Datasets/ping_scan.pcap"
+    pcapAttack = "Datasets/pingscan.pcap"
 
 pcapVariations = "Datasets/all-" + str(args.attack) + ".pcap"
 pkts = pyshark.FileCapture(pcapAttack)
@@ -730,10 +730,13 @@ class Rule:
         global max_fitness
         self.fitness = []
         self.fitness.append(ruleSizeFitness(self))
-        self.fitness.append(ruleContentsFitness(self))
-        self.fitness.append(rareContentsFitness(self))
-        self.fitness.append(ruleContentsModifiersFitness(self))
         self.fitness.append(ruleOptionsFitness(self))
+        
+        if self.protocol == "http":
+            self.fitness.append(ruleContentsFitness(self))
+            self.fitness.append(rareContentsFitness(self))
+            self.fitness.append(ruleContentsModifiersFitness(self))
+        
 
         for i in range(0, len(self.fitness)):
             if self.fitness[i] > max_fitness[i]:
@@ -754,7 +757,7 @@ def callGetFitness(rule, weights):
     return rule.getFitness(weights)
 
 def sortRules():
-    with open("all_rules_raw_0.out", "r") as all_rules:
+    with open("all_rules_raw_"+str(args.attack)+".out", "r") as all_rules:
         all_rules = all_rules.readlines()
 
     all_rules_list = []
@@ -786,27 +789,30 @@ def sortRules():
     
     for w0 in [0.01, 0.25, 0.5, 0.75, 1]:
         for w1 in [0.01, 0.25, 0.5, 0.75, 1]:
-            for w2 in [0.01, 0.25, 0.5, 0.75, 1]:
+            """for w2 in [0.01, 0.25, 0.5, 0.75, 1]:
                 for w3 in [0.01, 0.25, 0.5, 0.75, 1]:
[…]

### 2020-02-25 paper/paper.tex
updated table 3 and fig 7
done my pass on step 1

@@
   breaklines=true,           % sets automatic line breaking
-  breakatwhitespace=false,   % sets if automatic breaks should only happen at
+  breakatwhitespace=true,   % sets if automatic breaks should only happen at
   showspaces=false,
@@
   %extendedchars=true,         %
-  %breaklines=true,            % Zeilen werden Umgebrochen
   keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
@@
   \vspace{-2ex}
-  \caption{Suricata rule for SYN Flood attack.}
+  \caption{\suri\ rule for SYN Flood attack.}
   \label{fig:synflood-example}
@@
 Unfortunately, the relevant data for the option may be at the message
-payload not the fields. We found empirically that this occurs
-frequently in http attacks. Figure~\ref{fig:rule-jboss} shows an
-example of a rule that captures traffic containing the string
-\CodeIn{"/HtmlAdaptor"} in the payload.  The content strategy focuses
-on extracting such kinds of options from the payload. It is very
-important to note that content options are prevalent in \suri\ rules
-and the majority of these rule prescribe defenses against http
-attacks. For example, \percRulesWithContent{} of the \numrulessuri{}
-rules from the public data set~\cite{emerging-threats-open} we
-analyzed contain at least one content
+payload; not the fields. We found empirically that this occurs
+frequently in http attacks, which are prevalent. For example,
+\Fix{XXX\%} of the \numrulessuri{} rules from a public data
+set~\cite{emerging-threats-open} we analyzed prescribe defenses
+against http attacks. Figure~\ref{fig:rule-jboss} shows an example of
+a rule that captures traffic containing the string
+\CodeIn{"/HtmlAdaptor"} in the payload.  Content options are very
+common in http rules---\percRulesWithContent{} of the rules from the
+data set mentioned above contain at least one content
 option. Figure~\ref{fig:distribution-contents} shows the histogram of
 the number of these options per rule for the data set mentioned
-above. We limited the number of content options per rule to 10 for
-space. Furthermore, \Fix{xxx\%} of the rules are for
-http.\MyComment{Rules without \CodeIn{content} options are an
-  exception and half of the rules contain at least two
-  \CodeIn{content} options.}  Consequently, hangling content options
-for http is very important.
-
-We observed, by inspecting existing rules, that \CodeIn{content}
-options typically refer to sequence of characters in the payload. To
-find the values associated with these options, \tname{} splits the
-payload of the malicious message in tokens using natural language
-delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
+above. We limited the number of content options per rule to 10 in the
+histogram for space. \MyComment{Rules without \CodeIn{content} options
+  are an exception and half of the rules contain at least two
+  \CodeIn{content} options.}  Consequently, handling content options
+for http is important. The content strategy focuses on extracting such
+kinds of options from the payload.
+
+We found, inspecting existing rules, that \CodeIn{content} options
+typically refer to sequence of characters in the payload. To find the
+values associated with these options, \tname{} splits the payload of
+the malicious message in tokens using natural language delimiters,
+such as $\backslash$t, $\backslash$n, $\backslash$r, and
 spaces. Considering the example from
 Section~\ref{sec:content-example}, \tname{} produced a total of 23
-content options.
+tokens, considered content option candidates.
 
 It is worth noting that we found empirically that deciding which
-reverse engineering strategy to use based on the protocol was
-effective. We used the payload strategy for http and the field
-strategy for the other protocols.\Mar{discuss flood cases...} At the end of this step,
-\tname\ produces a seed rule that captures the malicious traffic by
-definition.
+reverse engineering strategy to use based on the protocol of the input
+(malicious) messages was effective. We used the payload strategy for
+http and the field strategy for the other protocols. For flooding
+attacks, \tname\ runs the same approach for each message and
+identifies common options across all messages. For that case,
+\tname\ reports, in addition, a \CodeIn{threshold} option, as
+discussed on Section~\ref{sec:rules-and-packets}.  At the end of the
+Reverse Engineering step, \tname\ produces a seed rule that captures
+the malicious traffic.
 
@@
     \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
-    \#~Options &
+    \#~Option Cand. &
     \#~Plausible Rules &    
@@
     ColdFusion & 19 & 39 & 5\\
-    .htaccess access & 17 & 2458 & 2\\
-    .idq access & 19 & 1111 & 2 \\
+    .htaccess access & 17 & 2,458 & 2\\
+    .idq access & 19 & 1,111 & 2 \\
     iisadmin access & 17 & 861 & 1 \\
-    /system32/ in URI & 20 & 3665 & 2 \\
+    /system32/ in URI & 20 & 3,665 & 2 \\
     Script tag in URI & 22 & - & - \\

### 2020-02-25 paper/paper.tex
pass

@@
 
-\vspace{2ex}
-\noindent\textbf{Limitations.} It is worth noting that although
+\textbf{Implementation details.} We implemented function
+\emph{isPlausible} by 1)~updating Suricata to only monitor the rule
+that is passed as parameter, 2)~replaying the negative traffic (see
+Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
+check if any benign message was captured by the rule. It is worth
+noting that the benign traffic is stored in ``pcap'' files~\cite{pcap}
+and Suricata can process those files very efficiently compared to
+analyzing network messages directly. Instead of sending traffic to a
+machine with a deployed server (\eg{}, web server) that
+\suri\ monitors, our setup calls Suricata directly through an
+API. Suricata is invoked multiple times in the process of minimizing
+the rule.
+\textbf{Limitations.} It is worth noting that although
 \tname{} is accurate for generating rules that discrimate positive and
@@
 of the training data it uses. For the positive traffic, \tname{} only
-requires one sample, but can benefit of several samples as the example
-from Section~\ref{sec:dos} illustrated. The same applies for the
-negative traffic. In this case, however, \tname{} can use public
-data sets available online.
-
-\vspace{2ex}
-\noindent\textbf{Implementation details.} We implemented function
-\emph{isPlausible} by 1)~updating Suricata to only use the rule that
-is passed as parameter to this function, 2)~replaying the negative
-traffic (see Section~\ref{sec:dataset-benign}), and 3)~reading the
-Suricata logs to check if any benign message was captured by the
-rule. It is worth noting that the benign traffic is stored in ``pcap''
-files~\cite{pcap} and Suricata can process those files very
-efficiently compared to analyzing network messages directly.
+requires one sample, but can benefit of several samples. The same
+applies for the negative traffic. In this case, however, \tname{} can
+use public data sets available online.
 
@@
 
-\textbf{Rationale.}~The golden rule is a rule that is known to capture
-a given attack. \tname\ generates several rules that satisfy the
-invariant to only capture the malicious traffic, but the amount of
-data we use to assess that invariant, albeit large, is limited. In
-fact, assessing quality of each rule (\ie{}, fitness to the attack)
-requires domain knowledge. The position of the golden rule at the
-ranking produced by \tname\ is a proxy measure to the effort of a
-system administrator in finding the rule that fits the purpose.
+\noindent
+\textbf{Rationale.}~The golden rule is a trusted human-written rule
+obtained from credible sources. \tname\ generates several rules that
+satisfy the invariant to only capture the malicious traffic, but the
+amount of data we use to evaluate that property, albeit large, is
+limited. Assessing the quality of each rule \tname\ produces (\ie{},
+fitness to the attack) requires domain knowledge and much effort. The
+position of the golden rule at the ranking produced by \tname\ is a
+proxy measure to the effort a system administrator would have in
+finding the rule that fits the purpose.
 
@@
 
-    %%Black Nurse \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:3; icode:3; detection\_filter:track by\_dst, count 250, seconds 1;}\\
-    %%Ping Flood  \MyComment{& multiple} & ICMP \MyComment{& Denial of Service} & \scriptsize{itype:8; icode:0; detection\_filter:track by\_src, count 30, seconds 1;}\\  
-    %%Port Scan \MyComment{& single} & TCP \MyComment{& Discovers open ports in a host} & \scriptsize{flags: A; ack: 0;dsize: 0;} \\
-
-    %%UDP Flood \MyComment{& multiple} & UDP \MyComment{& Denial of Service} & \scriptsize{fragbits:M; threshold: type both, track by\_dst, count 5000, seconds 5;} \\
+\noindent
+\textbf{Rationale.}~\tname\ proposes a variety of heuristic functions
+to rank the plausible rules produced during minimization. It is
+therefore important to understand how much each function contributes
+to the overall performance of the technique.
+
+\newcommand{\textRQthree}{What is the importance of the positive and
+  negative traffic in ranking?}
+\vspace{0.2cm}
+\begin{itemize}[leftmargin=*]
+\item{\textbf{RQ3.}} \textRQthree\
+\end{itemize}
+
+\noindent
+\textbf{Rationale.}~\tname\ uses both positive and negative examples
+to generate rules. This question investigates how sensible the
+technique is to the traffic that it uses.
+\noindent
+\vspace{1ex}
+
+The following sections describe the objects we used to evaluate the
+technique (Section~\ref{sec:dataset-benign}), answer the questions
+(sections~\ref{sec:answer-rqone},~\ref{sec:answer-rqothree}), and
+discuss threats to the validity of the
+experiments~(Section~\ref{sec:threats}).
 
@@
 
-\subsection{Methodology}
-
-Recall that Suricata is invoked multiple times in the process of
-minimizing the rule (see Section~\ref{sec:minimization}). Suricata
-supports a feature to process ``pcap'' files, enabling very efficient
-performance compared to processing actual network messages.
+%subsection{Methodology}
 
@@
 \subsection{Answering RQ1: \textRQone}
+\label{sec:answer-rqone}
 
 \subsection{Answering RQ2: \textRQtwo}
+\label{sec:answer-rqotwo}
+
+\subsection{Answering RQ3: \textRQthree}
+\label{sec:answer-rqothree}
 
@@
 \subsection{Threats to Validity}
+\label{sec:threats}
 

### 2020-02-25 paper/paper.tex
pass on eval (stop at 5.1)

@@
 check if any benign message was captured by the rule. It is worth
-noting that the benign traffic is stored in ``pcap'' files~\cite{pcap}
-and Suricata can process those files very efficiently compared to
-analyzing network messages directly. Instead of sending traffic to a
-machine with a deployed server (\eg{}, web server) that
-\suri\ monitors, our setup calls Suricata directly through an
-API. Suricata is invoked multiple times in the process of minimizing
-the rule.
-\textbf{Limitations.} It is worth noting that although
-\tname{} is accurate for generating rules that discrimate positive and
-negative traffic, the quality of these rules is limited by the quality
-of the training data it uses. For the positive traffic, \tname{} only
-requires one sample, but can benefit of several samples. The same
-applies for the negative traffic. In this case, however, \tname{} can
-use public data sets available online.
+noting that the benign traffic is stored in \emph{pcap}
+files~\cite{pcap} and Suricata can process those files very
+efficiently compared to analyzing network messages directly. Instead
+of sending traffic to a machine with a deployed server (\eg{}, web
+server) that \suri\ monitors, our setup calls Suricata directly
+through an API. Suricata is invoked multiple times in the process of
+minimizing the rule.  \textbf{Limitations.} It is worth noting that
+although \tname{} is accurate for generating rules that discrimate
+positive and negative traffic, the quality of these rules is limited
+by the quality of the training data it uses. For the positive traffic,
+\tname{} only requires one sample, but can benefit of several
+samples. The same applies for the negative traffic. In this case,
+however, \tname{} can use public data sets available online.
 
 \subsection{Step 3: Ranking}
+\label{sec:ranking}
 
@@
 \vspace{0.2cm}
-\begin{itemize}[leftmargin=*]
+\begin{itemize}[leftmargin=*,label={}]
 \item{\textbf{RQ1.}} \textRQone
@@
 \noindent
-\textbf{Rationale.}~The golden rule is a trusted human-written rule
-obtained from credible sources. \tname\ generates several rules that
-satisfy the invariant to only capture the malicious traffic, but the
-amount of data we use to evaluate that property, albeit large, is
-limited. Assessing the quality of each rule \tname\ produces (\ie{},
-fitness to the attack) requires domain knowledge and much effort. The
-position of the golden rule at the ranking produced by \tname\ is a
-proxy measure to the effort a system administrator would have in
-finding the rule that fits the purpose.
-
-\newcommand{\textRQtwo}{What is the importance of heuristic functions in ranking?}
+\textbf{Rationale.}~The golden rule is a trusted human-written
+\nids\ rule that we obtained from credible sources. \tname\ generates
+several rules that satisfy the invariant to only capture the malicious
+traffic, but the amount of data we used to evaluate that property,
+albeit large, is limited. Assessing the quality of each rule
+\tname\ produces (\ie{}, fitness to the attack) requires domain
+knowledge and much effort. The position of the golden rule at the
+ranking produced by \tname\ is a proxy measure to the effort of a
+system administrator in finding the rule that fits the purpose.
+
+\newcommand{\textRQtwo}{What is the importance of heuristic functions
+  on \tname's performance?}
 \vspace{0.2cm}
-\begin{itemize}[leftmargin=*]
+\begin{itemize}[leftmargin=*,label={}]
 \item{\textbf{RQ2.}} \textRQtwo\
@@
 \noindent
-\textbf{Rationale.}~\tname\ proposes a variety of heuristic functions
-to rank the plausible rules produced during minimization. It is
-therefore important to understand how much each function contributes
-to the overall performance of the technique.
-
-\newcommand{\textRQthree}{What is the importance of the positive and
-  negative traffic in ranking?}
+\textbf{Rationale.}~We proposed a variety of heuristic functions to
+rank the plausible rules that \tname\ created during minimization (as
+per step 2 in the workflow). Recall that these functions are based on
+the data from a public ruleset~\cite{emerging-threats-open} (see
+Section~\ref{sec:ranking}). It is therefore important to understand
+the extent to which each function contributes to the overall
+performance of the technique.
+
+\newcommand{\textRQthree}{What is the importance of the benign and
+  malicious traffic on \tname's performance?}
 \vspace{0.2cm}
-\begin{itemize}[leftmargin=*]
+\begin{itemize}[leftmargin=*,label={}]
 \item{\textbf{RQ3.}} \textRQthree\
@@
 to generate rules. This question investigates how sensible the
-technique is to the traffic that it uses.
+technique is to the selection of these inputs.
 \noindent
@@
 The following sections describe the objects we used to evaluate the
-technique (Section~\ref{sec:dataset-benign}), answer the questions
-(sections~\ref{sec:answer-rqone},~\ref{sec:answer-rqothree}), and
+technique (Section~\ref{sec:dataset-benign}), answer the posed
+research questions (sections~\ref{sec:answer-rqone},
+and~\ref{sec:answer-rqtwo}, and~\ref{sec:answer-rqthree}), and
 discuss threats to the validity of the
@@
 
-\subsubsection{Malicious traffic.}To reproduce the malicious traffic, we
-configured an HTTP server (Apache2) in one machine and ran
-Nikto2~\cite{nikto} in a different machine. We used this setup to
-replay each of the attacks from Table~\ref{table:attacks}.  On each
-execution, we captured the incoming traffic to the web server with
-Wireshark~\cite{wireshark-net-monitor} and used the corresponding pcap
-files as an input to \tname.\Mar{@Lucas, eu traduzi o que vc. falou
-  para ingles. Mas, aparentemente, isto esta **inconsistente** com o
-  que eh dito na Secao~\ref{attack-reproduction}. Por favor, esclareca!}
-  \Luc{Essa descrição é referente aos ataques HTTP/TCP da tabela. O
-    hping3 gera apenas o trafego do SYN Flood. Ambas as explicações
-    estão corretas. Creio q a descrição da seção 5.2.1 tenha q ser
-    movida para cá pois está explicando como capturamos o trafego
-    malicioso. Nós executamos o Nikto e o hping3 uma unica vez para
-    capturar com o wireshark o trafego q eles geram. O trafego
-    produzido pelo Nikto ja inclui todos os ataques que utilizamos,
-    nós separamos os pacotes de cada um deles manualmente utilizando o
-    wireshark.}\Mar{Fiquei na duvida se o hping soh serve para o SYN
-    Flood e se/porque o Nikto nao eh usado neste caso...Enfim, nao
-    precisa responder por aqui. Mova o texto deixando-o
-    *consistente*.}
-
-  \tname{} requires the attack to
-be replayed in order to check whether or not the rule it generates
-captures the attack.  See function $\mathit{isPlausible}$, used in the
-minimization algorithm from Section~\ref{sec:minimization}. For all
-attacks listed on Table~\ref{table:attacks}, we found scripts
-available on the web to replay the attacks. For example, the SYN Flood
-attack can be replayed with hping3~\cite{hping3}, using the following
-command: \CodeIn{hping3 -d 80 -w 64 -S -p 80 --flood --rand-source
-  <IP>}. With the script to reproduce the attack, we executed the
-following steps to obtain ``pcap'' files, which abstracts network
-messages~\cite{pcap}, to run \tname:~1)~instantiate the script (\eg{},
-define the IP, for the example above), 2)~execute the script,
-3)~monitor the traffic with the wireshark network
-monitor~\cite{wireshark-net-monitor}.
+\subsubsection{Malicious traffic.}To reproduce the malicious traffic
+for http and tcp attacks, we configured an http server (Apache2) in
+one machine and Nikto2~\cite{nikto} in a different machine. Nikto is
+an open source tool with a catalog of scripts to reproduce
+vulnerabilities. It is important to note that this setup is \emph{not}
+necessary if \emph{pcap} files---pcap is the input format used by
+\tname{}---for the attack already exist. As explained in
+Section~\ref{sec:minimization}, pcap files abstract network
+messages~\cite{pcap} and can be used to avoid network traffic during
+rule synthesis. We proceeded as follows to obtain pcap files for the
+the attacks we analyzed: we 1)~instantiated the script (\eg{}, defined
+the IP of the web server), 2)~executed the Nikto2 script, 3)~and
+monitored the traffic with the Wireshark network
+monitor~\cite{wireshark-net-monitor}. The Wireshark tool records the
+traffic and reports pcap files on output for replaying. The only
+difference in the setup of non-http non-tcp attack is in the first
+step of this procedure. For example, we used the hping3~\cite{hping3}
+command to reproduce the ICMP SYN Flood attack: \CodeIn{hping3 -d 80
+  -w 64 -S -p 80 --flood --rand-source <IP>}. This command alone
+reproduces a flooding attack on a given machine
+(Section~\ref{sec:dos}).\Mar{what about the variants? where do they
+  come from? this is confusing as the narrative above suggests the
+  attack is hardcoded in the tool.}
+
+%% On each execution, we captured
+%% the incoming traffic to the web server with
+%% Wireshark~\cite{wireshark-net-monitor} and used the corresponding pcap
+%% files as an input to
+
+%% \tname.\Mar{@Lucas, eu traduzi o que vc. falou
+%%   para ingles. Mas, aparentemente, isto esta **inconsistente** com o
+%%   que eh dito na Secao~\ref{attack-reproduction}. Por favor,
+%%   esclareca!}  \Luc{Essa descrição é referente aos ataques HTTP/TCP da
+%%   tabela. O hping3 gera apenas o trafego do SYN Flood. Ambas as
+%%   explicações estão corretas. Creio q a descrição da seção 5.2.1 tenha
+%%   q ser movida para cá pois está explicando como capturamos o trafego
+%%   malicioso. Nós executamos o Nikto e o hping3 uma unica vez para
+%%   capturar com o wireshark o trafego q eles geram. O trafego produzido
+%%   pelo Nikto ja inclui todos os ataques que utilizamos, nós separamos
+%%   os pacotes de cada um deles manualmente utilizando o
+%%   wireshark.}\Mar{Fiquei na duvida se o hping soh serve para o SYN
+%%   Flood e se/porque o Nikto nao eh usado neste caso...Enfim, nao
+%%   precisa responder por aqui. Mova o texto deixando-o *consistente*.}
+%%   \tname{} requires the attack to
+%% be replayed in order to check whether or not the rule it generates
+%% captures the attack.  See function $\mathit{isPlausible}$, used in the
+%% minimization algorithm from Section~\ref{sec:minimization}. For all
+%% attacks listed on Table~\ref{table:attacks}, we found scripts
+%available on the web to replay the attacks. For example,
+
+\Mar{--------------------------- estou aqui!}
 
@@
 \subsection{Answering RQ2: \textRQtwo}
-\label{sec:answer-rqotwo}
+\label{sec:answer-rqtwo}
 
 \subsection{Answering RQ3: \textRQthree}
-\label{sec:answer-rqothree}
+\label{sec:answer-rqthree}
 

### 2020-02-25 paper/paper.tex
table 2 missing categories

@@
     JBoss JMX\MyComment{ Console Beanshell Exploit} & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
-    WeBid \MyComment{Local File Inclusion} & - & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
-    ColdFusion \MyComment{Privilege Escalation} & - & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
+    WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
+    ColdFusion \MyComment{Privilege Escalation} & Privilege Escalation & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
     .htaccess access\MyComment{attempted-recon} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
@@
 command to reproduce the ICMP SYN Flood attack: \CodeIn{hping3 -d 80
-  -w 64 -S -p 80 --flood --rand-source <IP>}. This command alone
+  -w 512 -S -p 80 --flood --rand-source <IP>}. This command alone
 reproduces a flooding attack on a given machine

### 2020-02-25 paper/paper.tex
5.1.1

@@
     JBoss JMX\MyComment{ Console Beanshell Exploit} & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
-    WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
+    %% WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
     ColdFusion \MyComment{Privilege Escalation} & Privilege Escalation & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
@@
     JBoss JMX & 23 & 381 & 23 \\
-    WeBid & 23 & - & -\\    
+    %%   WeBid & 23 & - & -\\    
     ColdFusion & 19 & 39 & 5\\
@@
   come from? this is confusing as the narrative above suggests the
-  attack is hardcoded in the tool.}
+  attack is hardcoded in the tool.}\Luc{O trafego de um ataque 
+  normalmente possui varios pacotes, entre esses varios pacotes alguns 
+  deles possuem um padrao. As regras sao criadas descrevendo esse padrao,
+  com isso um unico ataque pode disparar varios alertas.
+
+  Porem, existem ataques q so produzem um alerta, ou seja, a regra criada 
+  so detecta um dos pacotes do trafego. Nesses casos, procuramos na internet 
+  datasets (arquivos pcap) que contivessem mais pacotes desses ataques.
+  Fizemos isso tambem para ataques q nos ja tinhamos variacoes, pois
+  (teoricamente) quanto mais variacoes, melhor. Existem alguns sites onde 
+  eh possivel encontrar pcaps com o trafego de alguns ataques, n consegui
+  adicionar nas referencias: 
+  https://www.netresec.com/?page=PcapFiles
+}
 

### 2020-02-25 paper/paper.tex
pass on 5.1.2

@@
 
-\newcommand{\textRQtwo}{What is the importance of heuristic functions
-  on \tname's performance?}
+\newcommand{\textRQtwo}{How important are the heuristic functions
+  to \tname's performance?}
 \vspace{0.2cm}
@@
 
-\newcommand{\textRQthree}{What is the importance of the benign and
-  malicious traffic on \tname's performance?}
+\newcommand{\textRQthree}{How important are the benign and
+  malicious traffic to \tname's performance?}
 \vspace{0.2cm}
@@
 
-%% On each execution, we captured
-%% the incoming traffic to the web server with
-%% Wireshark~\cite{wireshark-net-monitor} and used the corresponding pcap
-%% files as an input to
-
-%% \tname.\Mar{@Lucas, eu traduzi o que vc. falou
-%%   para ingles. Mas, aparentemente, isto esta **inconsistente** com o
-%%   que eh dito na Secao~\ref{attack-reproduction}. Por favor,
-%%   esclareca!}  \Luc{Essa descrição é referente aos ataques HTTP/TCP da
-%%   tabela. O hping3 gera apenas o trafego do SYN Flood. Ambas as
-%%   explicações estão corretas. Creio q a descrição da seção 5.2.1 tenha
-%%   q ser movida para cá pois está explicando como capturamos o trafego
-%%   malicioso. Nós executamos o Nikto e o hping3 uma unica vez para
-%%   capturar com o wireshark o trafego q eles geram. O trafego produzido
-%%   pelo Nikto ja inclui todos os ataques que utilizamos, nós separamos
-%%   os pacotes de cada um deles manualmente utilizando o
-%%   wireshark.}\Mar{Fiquei na duvida se o hping soh serve para o SYN
-%%   Flood e se/porque o Nikto nao eh usado neste caso...Enfim, nao
-%%   precisa responder por aqui. Mova o texto deixando-o *consistente*.}
-%%   \tname{} requires the attack to
-%% be replayed in order to check whether or not the rule it generates
-%% captures the attack.  See function $\mathit{isPlausible}$, used in the
-%% minimization algorithm from Section~\ref{sec:minimization}. For all
-%% attacks listed on Table~\ref{table:attacks}, we found scripts
-%available on the web to replay the attacks. For example,
-
-\Mar{--------------------------- estou aqui!}
-
-\subsubsection{Benign traffic.}\tname{} uses public databases storing benign
-traffic for all protocols covered by Suricata (see
-Figure~\ref{fig:distribution-rules-protocol}). We used two data sets of
-benign traffic. The Bigflows.pcap data set is part of the
-TcpReplay~\cite{tcpreplay} open source project. This data set includes
-real network traffic of a busy private network access point to the
-Internet. According to the TcpReplay web site ``This capture is much
-larger and has a smaller average packet size than the previous capture
-(smallFlows.pcap). It also has many more flows and different
-applications.''. In addition to the Bigflows.pcap data set, we also
-used data sets of normal traffic made available by Stratosphere
-Lab~\cite{stratosphere-normal}.
+\subsubsection{Benign traffic.}~Security
+tools (e.g., perforation testing tools) often use benign traffic to
+evaluate false positive rates. Consequently, different public options
+exist. We used two popular public sources. The Bigflows.pcap data set
+is part of the TcpReplay~\cite{tcpreplay} open source project. This
+data set includes real network traffic of a busy private network
+access point to the Internet. According to the TcpReplay web site
+``This capture is much larger and has a smaller average packet size
+than the previous capture (smallFlows.pcap). It also has many more
+flows and different applications.''. In addition to the Bigflows.pcap
+data set, we also used data sets\Mar{more than one?} made available by
+Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
+  of the traffic? as the description for Bigflows.pcap?}.
 

### 2020-02-26 paper/paper.tex
pass on 5.1.2

@@
   \small
-  \caption{\label{table:attacks}List of attacks.}
+  \caption{\label{table:attacks}List of attacks.\Mar{adicionar coluna ``Source''}}
   \centering
@@
     \multicolumn{1}{c}{Protocol} &
-    \multicolumn{1}{c}{Groud Truth} \\
+    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
     \midrule     
@@
 
-Table~\ref{table:attacks} shows the attacks we considered in our
+\subsubsection{Attacks.}Table~\ref{table:attacks} shows the attacks we considered in our
 evaluation. These attacks cover different protocols and features of
@@
 ``Protocol'' shows the protocol name, and column ``Ground Truth''
-shows the options of the rule that captures the attack.
-
-\subsubsection{Malicious traffic.}To reproduce the malicious traffic
-for http and tcp attacks, we configured an http server (Apache2) in
-one machine and Nikto2~\cite{nikto} in a different machine. Nikto is
-an open source tool with a catalog of scripts to reproduce
-vulnerabilities. It is important to note that this setup is \emph{not}
-necessary if \emph{pcap} files---pcap is the input format used by
-\tname{}---for the attack already exist. As explained in
-Section~\ref{sec:minimization}, pcap files abstract network
-messages~\cite{pcap} and can be used to avoid network traffic during
-rule synthesis. We proceeded as follows to obtain pcap files for the
-the attacks we analyzed: we 1)~instantiated the script (\eg{}, defined
-the IP of the web server), 2)~executed the Nikto2 script, 3)~and
-monitored the traffic with the Wireshark network
-monitor~\cite{wireshark-net-monitor}. The Wireshark tool records the
-traffic and reports pcap files on output for replaying. The only
-difference in the setup of non-http non-tcp attack is in the first
-step of this procedure. For example, we used the hping3~\cite{hping3}
-command to reproduce the ICMP SYN Flood attack: \CodeIn{hping3 -d 80
-  -w 512 -S -p 80 --flood --rand-source <IP>}. This command alone
-reproduces a flooding attack on a given machine
-(Section~\ref{sec:dos}).\Mar{what about the variants? where do they
-  come from? this is confusing as the narrative above suggests the
-  attack is hardcoded in the tool.}\Luc{O trafego de um ataque 
-  normalmente possui varios pacotes, entre esses varios pacotes alguns 
-  deles possuem um padrao. As regras sao criadas descrevendo esse padrao,
-  com isso um unico ataque pode disparar varios alertas.
-
-  Porem, existem ataques q so produzem um alerta, ou seja, a regra criada 
-  so detecta um dos pacotes do trafego. Nesses casos, procuramos na internet 
-  datasets (arquivos pcap) que contivessem mais pacotes desses ataques.
-  Fizemos isso tambem para ataques q nos ja tinhamos variacoes, pois
-  (teoricamente) quanto mais variacoes, melhor. Existem alguns sites onde 
-  eh possivel encontrar pcaps com o trafego de alguns ataques, n consegui
-  adicionar nas referencias: 
-  https://www.netresec.com/?page=PcapFiles
-}
+shows the options of the rule that captures the attack.\Mar{Explicar o
+  rationale para esta selecao de ataques!}
+
+%% It is important to note that this setup is \emph{not}
+%% necessary if \emph{pcap} files for the attack already exist. We
+%% proceeded as follows to obtain pcap files for the the attacks we
+%% analyzed: we 1)~instantiated the script (\eg{}, defined the IP of the
+%% web server), 2)~executed the script, 3)~and monitored the traffic with
+%% the Wireshark network monitor. The
+%% Wireshark tool records the traffic and reports pcap files on output
+%% for replaying it.
+
+\subsubsection{Malicious traffic.}~For some of the attacks listed on
+Table~\ref{table:attacks}, we obtained pcap files directly from
+trusted security websites (\eg{},
+NetReseC~\cite{pcap-attacks}). Recall that \tname\ takes pcap files on
+input. As explained in Section~\ref{sec:minimization}, these files
+abstract network messages~\cite{pcap} and can be used to avoid network
+traffic during rule synthesis. In addition to these sources, we used
+the Nikto2~\cite{nikto} open source security tool to reproduce
+attacks. To reproduce the http and tcp attacks listed on
+Table~\ref{table:attacks}, we configured an http server (Apache2) in
+one machine and ran the security tool in a different
+machine. Simultenously, we ran WireShark~\cite{wireshark-net-monitor}
+to record the network traffic and Suricata to analyze the traffic. We
+configured \suri\ with all rules from the ``Emerging Threats Open
+Ruleset''~\cite{emerging-threats-open} and ran Nikto2 with default
+parameters, which makes it replay all recorded attacks. After
+execution, we inspected the \suri\ log to identify which rules have
+been activated and which packets were involved in each
+activation. With the packet numbers, it is possible to recover the
+corresponding pcap files, using Wireshark, as to ran \tname{} on
+them. It is worth noting that some rules are triggered multiple times,
+indicating that there are multiple instances of the same attack
+encoded in Nikto2. RQ3 evaluates the importance of these variants to
+\tname{}'s performance. Finally, Nikto2 only supports tcp and http
+attacks.  We used the hping3~\cite{hping3} to reproduce the Ping Scan
+icmp attack (see Section~\ref{sec:dos}).  More precisely, we used the
+following hping3 command to reproduce that attack: \CodeIn{hping3 -d
+  80 -w 512 -S -p 80 --flood --rand-source <IP>}.
+
+%% \Mar{what about the variants? where do
+%%   they come from? this is confusing as the narrative above suggests
+%%   the attack is hardcoded in the tool.}\Luc{O trafego de um ataque
+%%   normalmente possui varios pacotes, entre esses varios pacotes alguns
+%%   deles possuem um padrao. As regras sao criadas descrevendo esse
+%%   padrao, com isso um unico ataque pode disparar varios alertas.
+
+%%   Porem, existem ataques q so produzem um alerta, ou seja, a regra criada 
+%%   so detecta um dos pacotes do trafego. Nesses casos, procuramos na internet 
+%%   datasets (arquivos pcap) que contivessem mais pacotes desses ataques.
+%%   Fizemos isso tambem para ataques q nos ja tinhamos variacoes, pois
+%%   (teoricamente) quanto mais variacoes, melhor. Existem alguns sites onde 
+%%   eh possivel encontrar pcaps com o trafego de alguns ataques, n consegui
+%%   adicionar nas referencias: 
+%%   https://www.netresec.com/?page=PcapFiles
+%% }
 
@@
 \label{sec:answer-rqthree}
-
-\Mar{Explique resumidamente o que cada ataque listado na tabela 2 faz...}
 

### 2020-02-26 paper/paper.tex
updated table 3

@@
   \centering
-  \begin{tabular}{lrrr}
+  \begin{tabular}{lrrrr}
     \toprule
     \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
+    \multicolumn{1}{c}{\multirow{2}{*}{\# Variants}} &
     \#~Option Cand. &
@@
      &
+    \multicolumn{1}{c}{} &
     \multicolumn{1}{c}{(step 1)} &
@@
     \midrule
-    Ping Scan & 5 & 2 & - \\
-    SYN Flood & 2 & 2 & - \\
-    JBoss JMX & 23 & 381 & 23 \\
-    %%   WeBid & 23 & - & -\\    
-    ColdFusion & 19 & 39 & 5\\
-    .htaccess access & 17 & 2,458 & 2\\
-    .idq access & 19 & 1,111 & 2 \\
-    iisadmin access & 17 & 861 & 1 \\
-    /system32/ in URI & 20 & 3,665 & 2 \\
-    Script tag in URI & 22 & - & - \\
+    Ping Scan & 5 & 5 & 2 & - \\
+    SYN Flood & 3 & 2 & 2 & - \\
+    JBoss JMX & 4 & 23 & 381 & 23 \\
+    %%   WeBid & - & 23 & - & -\\    
+    ColdFusion & 6 & 19 & 39 & 5\\
+    .htaccess access & 4 & 17 & 2,458 & 2\\
+    .idq access & 13 & 19 & 1,111 & 2 \\
+    iisadmin access & 3 & 17 & 861 & 1 \\
+    /system32/ in URI & 3 & 20 & 3,665 & 2 \\
+    Script tag in URI & 22 & 22 & - & - \\
     \bottomrule

### 2020-02-26 paper/paper.tex
updated table 2

@@
   \centering
-  \begin{tabular}{lllp{11cm}}
+  \begin{tabular}{llllp{11cm}}
     \toprule
     \multicolumn{1}{c}{Name} &
+    \multicolumn{1}{c}{Source} &
     \multicolumn{1}{c}{Category} &
@@
     \midrule     
-    Ping Scan & Active Recon. & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
-    SYN Flood & DoS & TCP & \textbf{flags}: S,12; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\
-
-    JBoss JMX\MyComment{ Console Beanshell Exploit} & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
+    Ping Scan & NMAP & Active Recon. & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
+    SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\
+
+    JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
     %% WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
-    ColdFusion \MyComment{Privilege Escalation} & Privilege Escalation & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
-    .htaccess access\MyComment{attempted-recon} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
-    .idq access\MyComment{EXPLOIT ISAPI} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
-    iisadmin access\MyComment{web-application-attack} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
-    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
-    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
+    ColdFusion \MyComment{Privilege Escalation} & Nikto2 & Privilege Escalation & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
+    .htaccess access\MyComment{attempted-recon} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
+    .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
+    iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
+    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
+    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
     \bottomrule

### 2020-02-26 paper/paper.tex
revising descriptioin of Malicious traffic

@@
 NetReseC~\cite{pcap-attacks}). Recall that \tname\ takes pcap files on
-input. As explained in Section~\ref{sec:minimization}, these files
+input. As explained on Section~\ref{sec:minimization}, these files
 abstract network messages~\cite{pcap} and can be used to avoid network
@@
 attacks. To reproduce the http and tcp attacks listed on
-Table~\ref{table:attacks}, we configured an http server (Apache2) in
-one machine and ran the security tool in a different
-machine. Simultenously, we ran WireShark~\cite{wireshark-net-monitor}
-to record the network traffic and Suricata to analyze the traffic. We
-configured \suri\ with all rules from the ``Emerging Threats Open
-Ruleset''~\cite{emerging-threats-open} and ran Nikto2 with default
-parameters, which makes it replay all recorded attacks. After
+Table~\ref{table:attacks}, we configured an http server (Apache2) and
+ran a security tool to attack that server. Simultenously, we ran
+WireShark~\cite{wireshark-net-monitor} to record the network traffic
+and Suricata to analyze the traffic. We configured \suri\ with all
+rules from the ``Emerging Threats Open
+Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
+default parameters, which makes it replay all recorded attacks. After
 execution, we inspected the \suri\ log to identify which rules have
@@
 indicating that there are multiple instances of the same attack
-encoded in Nikto2. RQ3 evaluates the importance of these variants to
+encoded in Nikto2\Mar{@Lucas, diga exatamente como vc. obtem as
+  variacoes. isto eh que eh importante para a proxima sentenca.}. rq3 evaluates the importance of these variants to
 \tname{}'s performance. Finally, Nikto2 only supports tcp and http

### 2020-02-26 paper/paper.tex
small changes on another pass on Section 5.1.2

@@
 traffic during rule synthesis. In addition to these sources, we used
-the Nikto2~\cite{nikto} open source security tool to reproduce
-attacks. To reproduce the http and tcp attacks listed on
-Table~\ref{table:attacks}, we configured an http server (Apache2) and
-ran a security tool to attack that server. Simultenously, we ran
+the Nikto2~\cite{nikto} open source security tool as follows to obtain
+pcap files for http and tcp attacks. We configured an http server
+(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
 WireShark~\cite{wireshark-net-monitor} to record the network traffic
@@
 execution, we inspected the \suri\ log to identify which rules have
-been activated and which packets were involved in each
+been activated and which packets were involved on each
 activation. With the packet numbers, it is possible to recover the
@@
 encoded in Nikto2\Mar{@Lucas, diga exatamente como vc. obtem as
-  variacoes. isto eh que eh importante para a proxima sentenca.}. rq3 evaluates the importance of these variants to
-\tname{}'s performance. Finally, Nikto2 only supports tcp and http
-attacks.  We used the hping3~\cite{hping3} to reproduce the Ping Scan
-icmp attack (see Section~\ref{sec:dos}).  More precisely, we used the
-following hping3 command to reproduce that attack: \CodeIn{hping3 -d
-  80 -w 512 -S -p 80 --flood --rand-source <IP>}.
+  variacoes. isto eh que eh importante para a proxima sentenca.}. RQ3
+evaluates the importance of these variants to \tname{}'s
+performance. Finally, we used the hping3~\cite{hping3} to reproduce
+the Ping Scan icmp attack (see Section~\ref{sec:dos}) as Nikto2 only
+supports tcp and http attacks. More precisely, we used the following
+hping3 command to reproduce that attack: \CodeIn{hping3 -d 80 -w 512
+  -S -p 80 --flood --rand-source <IP>}.
 

### 2020-02-26 paper/paper.tex
minor changes after revising section 5.1.1.

@@
   \small
-  \caption{\label{table:attacks}List of attacks.\Mar{adicionar coluna ``Source''}}
+  \caption{\label{table:attacks}List of attacks.}
   \centering
-  \begin{tabular}{llllp{11cm}}
+  \begin{tabular}{llllp{10cm}}
     \toprule
@@
     \multicolumn{1}{c}{Source} &
-    \multicolumn{1}{c}{Category} &
+    \multicolumn{1}{c}{Type} &
     \multicolumn{1}{c}{Protocol} &
@@
     \midrule     
-    Ping Scan & NMAP & Active Recon. & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
+    Ping Scan & NMAP & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
     SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\
@@
     %% WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
-    ColdFusion \MyComment{Privilege Escalation} & Nikto2 & Privilege Escalation & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
+    ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
     .htaccess access\MyComment{attempted-recon} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
@@
 
-\subsubsection{Attacks.}Table~\ref{table:attacks} shows the attacks we considered in our
-evaluation. These attacks cover different protocols and features of
-\tname{}. Column ``Name'' shows the name of the attack, column
-``Protocol'' shows the protocol name, and column ``Ground Truth''
-shows the options of the rule that captures the attack.\Mar{Explicar o
-  rationale para esta selecao de ataques!}
+\subsubsection{Attacks.}~Table~\ref{table:attacks} shows the selection
+of attacks considered in our evaluation.  Column ``Name'' shows the
+name of the attack, column ``Source'' shows the source where we
+obtained data to reproduce the attack (see
+Section~\ref{subsec:malicious-traffic}), column ``Type'' shows the
+kind of attack, column ``Protocol'' shows the name of the protocol
+explored in the attack, and column ``Ground Truth'' shows the rule
+options to capture the attack. The rationale for the selection were
+(1) to cover attacks with different characteristics, (2) to cover
+attacks of various protocols, and (3) to explore different features of
+\tname. As for (1), the selection includes attacks of the following
+types: Denial-of-Service (DoS), Active Reconnaissance (AREC),
+Privilege escalation (PESC), and Remote Code Execution (RCE). Note
+that the distribution of attacks is not uniform across types and
+protocols as they are not uniform in practice. For example, as
+per~\cite{emerging-threats-open}, RCE are the most prevalent type of
+attacks today.\Mar{please, confirm} Columns ``Protocol'' and ``Ground
+Truth'' indicate, respectively, that we analyzed attacks of different
+protocols and analyzed different features of \tname.
 
@@
 
-\subsubsection{Malicious traffic.}~For some of the attacks listed on
+\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~For some of the attacks listed on
 Table~\ref{table:attacks}, we obtained pcap files directly from
@@
 evaluates the importance of these variants to \tname{}'s
-performance. Finally, we used the hping3~\cite{hping3} to reproduce
-the Ping Scan icmp attack (see Section~\ref{sec:dos}) as Nikto2 only
-supports tcp and http attacks. More precisely, we used the following
-hping3 command to reproduce that attack: \CodeIn{hping3 -d 80 -w 512
-  -S -p 80 --flood --rand-source <IP>}.
+performance. We used NMAP~\cite{netmap} to reproduce the Ping Scan
+attack (see Section~\ref{sec:active-recon}) and hping3~\cite{hping3}
+to reproduce the SYN Flood attack (see
+Section~\ref{sec:dos}).
 
@@
 \label{sec:dos}
+
+\Mar{move from above --> We used the following hping3 command to
+  reproduce that attack: \CodeIn{hping3 -d 80 -w 512 -S -p 80 --flood
+    --rand-source <IP>}.}
 

### 2020-02-26 paper/paper.tex
change is results table layout

@@
 
+\setlength{\tabcolsep}{3pt}
 \begin{table}[h!]
@@
     \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
-    \multicolumn{1}{c}{\multirow{2}{*}{\# Variants}} &
+    \multicolumn{1}{c}{\multirow{2}{*}{\# Vars.}} &
     \#~Option Cand. &
@@
 
+This section describes the objects used in \tname{}'s evaluation.
+
+\vspace{1ex}
 \subsubsection{Attacks.}~Table~\ref{table:attacks} shows the selection
@@
 that the distribution of attacks is not uniform across types and
-protocols as they are not uniform in practice. For example, as
-per~\cite{emerging-threats-open}, RCE are the most prevalent type of
-attacks today.\Mar{please, confirm} Columns ``Protocol'' and ``Ground
-Truth'' indicate, respectively, that we analyzed attacks of different
-protocols and analyzed different features of \tname.
+protocols as these distributions are not uniform in practice. For
+example, as per~\cite{emerging-threats-open}, RCE are the most
+prevalent type of attacks today.\Mar{please, confirm} Columns
+``Protocol'' and ``Ground Truth'' indicate, respectively, that we
+analyzed attacks of different protocols and analyzed different
+features of \tname.
 

### 2020-02-26 paper/paper.tex
ground truth section

@@
 
-\subsubsection{Ground Truth.}To evaluate precision of \tname\ we
-need rules that are considered correct to use as reference. The rules
-that detect the attacks from Table~\ref{table:attacks} were obtained
-from the Emerging Threats Open Ruleset~\cite{emerging-threats-open}.
+\subsubsection{Ground Truth.}To evaluate \tname\ objectively we need
+to compare the rules it generates with human-written rules from a
+trusted source. We used the ``Emerging Threats Open
+Ruleset''~\cite{emerging-threats-open} data set for that. After
+finding the attack and corresponding rule using the procedure
+described in Section~\ref{subsec:malicious-traffic}, we reproduced the
+attack and confirmed that the rule captures the attack. Column
+``Ground Truth'' from Table~\ref{table:attacks} only shows the
+essential relevant options. Documentation options (\eg{}, ``msg'') are
+discared.
 

### 2020-02-26 paper/paper.tex
comments

@@
     ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
-    .htaccess access\MyComment{attempted-recon} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
+    .htaccess access\MyComment{attempted-recon} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
     .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
     iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
-    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
+    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
     Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
@@
 example, as per~\cite{emerging-threats-open}, RCE are the most
-prevalent type of attacks today.\Mar{please, confirm} Columns
+prevalent type of attacks today.\Mar{please, confirm} \Luc{Nao da
+para afirmar isso, as regras nao contem essa informacao. A classificacao
+das regras eh bem generica, por exemplo: "web-application-attack".} Columns
 ``Protocol'' and ``Ground Truth'' indicate, respectively, that we
@@
 encoded in Nikto2\Mar{@Lucas, diga exatamente como vc. obtem as 
-  variacoes. isto eh que eh importante para a proxima sentenca.}. RQ3
-evaluates the importance of these variants to \tname{}'s
+variacoes. isto eh que eh importante para a proxima sentenca.}. 
+\Luc{Eh feito da mesma forma q foi descrita acima. Quando um ataque 
+produz varios pacotes que geram alertas, nos criamos um pcap com um 
+dos pacotes e utilizamos ele como para gerar a regra maxima inicial. 
+Os outros pacotes sao salvos em um outro pcap e sao utilizados para 
+calcular o recall.} RQ3 evaluates the importance of these variants to \tname{}'s
 performance. We used NMAP~\cite{netmap} to reproduce the Ping Scan
@@
 Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
-  of the traffic? as the description for Bigflows.pcap?}.
+  of the traffic? as the description for Bigflows.pcap?}. \Luc{@Guilherme ...}
 

### 2020-02-26 paper/paper.tex
pass

@@
 Privilege escalation (PESC), and Remote Code Execution (RCE). Note
-that the distribution of attacks is not uniform across types and
-protocols as these distributions are not uniform in practice. For
-example, as per~\cite{emerging-threats-open}, RCE are the most
-prevalent type of attacks today.\Mar{please, confirm} \Luc{Nao da
-para afirmar isso, as regras nao contem essa informacao. A classificacao
-das regras eh bem generica, por exemplo: "web-application-attack".} Columns
-``Protocol'' and ``Ground Truth'' indicate, respectively, that we
-analyzed attacks of different protocols and analyzed different
-features of \tname.
+that the distribution of attacks is not uniform across protocols as
+these distributions are not uniform in practice---http attacks are
+prevalent.
 

### 2020-02-27 suricata/doc/userguide/conf.py
removed unused suricata folder

@@ -1,328 +1,0 @@
-# -*- coding: utf-8 -*-
-#
-# Suricata documentation build configuration file, created by
-# sphinx-quickstart on Fri Nov  6 10:12:25 2015.
-#
-# This file is execfile()d with the current directory set to its
-# containing dir.
-#
-# Note that not all possible configuration values are present in this
-# autogenerated file.
-#
-# All configuration values have a default; values that are commented out
-# serve to show the default.
-
-import sys
-import os
-import shlex
-import re
-
-on_rtd = os.environ.get('READTHEDOCS', None) == 'True'
-
-# If extensions (or modules to document with autodoc) are in another directory,
-# add these directories to sys.path here. If the directory is relative to the
-# documentation root, use os.path.abspath to make it absolute, like shown here.
-#sys.path.insert(0, os.path.abspath('.'))
-
-# -- General configuration ------------------------------------------------
-
-# If your documentation needs a minimal Sphinx version, state it here.
-#needs_sphinx = '1.0'
-
-# Add any Sphinx extension module names here, as strings. They can be
-# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
-# ones.
-extensions = []
-
-# Add any paths that contain templates here, relative to this directory.
-templates_path = ['_templates']
-
-# The suffix(es) of source filenames.
-# You can specify multiple suffix as a list of string:
-# source_suffix = ['.rst', '.md']
-source_suffix = '.rst'
-
-# The encoding of source files.
-#source_encoding = 'utf-8-sig'
-
-# The master toctree document.
-master_doc = 'index'
-
-# General information about the project.
-project = u'Suricata'
-copyright = u'2016, OISF'
-author = u'OISF'
-
-# The version info for the project you're documenting, acts as replacement for
-# |version| and |release|, also used in various other places throughout the
-# built documents.
-#
-# The short X.Y version.
-
-# Attempt to extract the version configure.ac.
-try:
-    version = os.environ.get('version', None)
-    if not version:
-        version = re.search(
-            "AC_INIT\(suricata,\s*(.*)?\)",
-            open("../../configure.ac").read()).groups()[0]
-    if not version:
-        version = "unknown"
-except:
-    version = "unknown"
-release = version
-
-# The language for content autogenerated by Sphinx. Refer to documentation
-# for a list of supported languages.
-#
-# This is also used if you do content translation via gettext catalogs.
[…]

### 2020-02-27 suricata/libhtp/htp/htp_base64.h

@@ -1,74 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- * 
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
[…]

### 2020-02-27 suricata/libhtp/htp/htp_content_handlers.c

@@ -1,299 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- * 
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
-#include "htp_config_auto.h"
-
-#include "htp_private.h"
-
-/**
[…]

### 2020-02-27 suricata/libhtp/htp/htp_cookies.c

@@ -1,119 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- * 
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
-#include "htp_config_auto.h"
-
-#include "htp_private.h"
-
-/**
[…]

### 2020-02-27 suricata/libhtp/htp/htp_multipart.h

@@ -1,345 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- * 
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
-#ifndef _HTP_MULTIPART_H
[…]

### 2020-02-27 suricata/libhtp/htp/htp_parsers.c

@@ -1,192 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- * 
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
-#include "htp_config_auto.h"
-
-#include "htp_private.h"
-
-/**
- * Determines protocol number from a textual representation (i.e., "HTTP/1.1"). This
[…]

### 2020-02-27 suricata/libhtp/htp/htp_private.h

@@ -1,253 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- *
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
[…]

### 2020-02-27 suricata/libhtp/htp/htp_request_generic.c

@@ -1,387 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- * 
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
-#include "htp_config_auto.h"
-
-#include "htp_private.h"
-
-/**
[…]

### 2020-02-27 suricata/libhtp/htp/htp_response.c

@@ -1,1207 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- * 
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
-#include "htp_config_auto.h"
-
[…]

### 2020-02-27 suricata/libhtp/htp/htp_table.h

@@ -1,184 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- *
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
-#ifndef HTP_TABLE_H
-#define	HTP_TABLE_H
-
-#ifdef	__cplusplus
-extern "C" {
-#endif
-
[…]

### 2020-02-27 suricata/libhtp/htp/htp_util.c

@@ -1,2609 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- * 
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- * 
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- * 
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
-#include "htp_config_auto.h"
-
-#include "htp_private.h"
-
-/**
[…]

### 2020-02-27 suricata/libhtp/htp_config_auto_gen.h

@@ -1,77 +1,0 @@
-/* htp_config_auto_gen.h.  Generated from htp_config_auto_gen.h.in by configure.  */
-/* htp_config_auto_gen.h.in.  Generated from configure.ac by autoheader.  */
-
-/* Define to 1 if you have the <dlfcn.h> header file. */
-#define HAVE_DLFCN_H 1
-
-/* Define if you have the iconv() function and it works. */
-#define HAVE_ICONV 1
-
-/* Define to 1 if you have the <inttypes.h> header file. */
-#define HAVE_INTTYPES_H 1
-
-/* Define to 1 if you have the `z' library (-lz). */
-#define HAVE_LIBZ 1
-
-/* Define to 1 if you have the <memory.h> header file. */
-#define HAVE_MEMORY_H 1
-
-/* Define to 1 if you have the <stdint.h> header file. */
-#define HAVE_STDINT_H 1
-
-/* Define to 1 if you have the <stdlib.h> header file. */
-#define HAVE_STDLIB_H 1
-
-/* Define to 1 if you have the <strings.h> header file. */
-#define HAVE_STRINGS_H 1
-
-/* Define to 1 if you have the <string.h> header file. */
-#define HAVE_STRING_H 1
-
-/* Define to 1 if you have the `strlcat' function. */
-/* #undef HAVE_STRLCAT */
-
-/* Define to 1 if you have the `strlcpy' function. */
-/* #undef HAVE_STRLCPY */
-
-/* Define to 1 if you have the <sys/stat.h> header file. */
-#define HAVE_SYS_STAT_H 1
-
-/* Define to 1 if you have the <sys/types.h> header file. */
-#define HAVE_SYS_TYPES_H 1
-
-/* Define to 1 if you have the <unistd.h> header file. */
-#define HAVE_UNISTD_H 1
-
-/* Define as const if the declaration of iconv() needs const. */
-#define ICONV_CONST 
-
-/* Define to the sub-directory where libtool stores uninstalled libraries. */
-#define LT_OBJDIR ".libs/"
-
-/* Name of package */
-#define PACKAGE "libhtp"
-
-/* Define to the address where bug reports for this package should be sent. */
-#define PACKAGE_BUGREPORT ""
-
-/* Define to the full name of this package. */
-#define PACKAGE_NAME "LibHTP"
-
-/* Define to the full name and version of this package. */
-#define PACKAGE_STRING "LibHTP 0.5.30"
-
[…]

### 2020-02-27 suricata/libhtp/test/test_bench.cpp

@@ -1,78 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- *
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
- ***************************************************************************/
-
-/**
- * @file
- *
- * @author Ivan Ristic <ivanr@webkreator.com>
- */
-
-#include <iostream>
-#include <gtest/gtest.h>
-#include <htp/htp_private.h>
[…]

### 2020-02-27 suricata/libhtp/test/test_gunzip.cpp

@@ -1,269 +1,0 @@
-/***************************************************************************
- * Copyright (c) 2009-2010 Open Information Security Foundation
- * Copyright (c) 2010-2013 Qualys, Inc.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are
- * met:
- *
- * - Redistributions of source code must retain the above copyright
- *   notice, this list of conditions and the following disclaimer.
-
- * - Redistributions in binary form must reproduce the above copyright
- *   notice, this list of conditions and the following disclaimer in the
- *   documentation and/or other materials provided with the distribution.
-
- * - Neither the name of the Qualys, Inc. nor the names of its
- *   contributors may be used to endorse or promote products derived from
- *   this software without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
- * DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
- * THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
- * (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
- * OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
[…]

### 2020-02-27 suricata/rules/count-proto-without-content.sh

@@ -1,25 +1,0 @@
-#!/bin/bash
-
-proto_list=($(suricata --list-app-layer-protos) tcp udp icmp ip modbus pkthdr ipv6)
-str_to_ignore_list=(Error opening file /usr/local/var/log/suricata/suricata.log =========Supported App Layer Protocols=========)
-ignore_str=0
-
-for proto in "${proto_list[@]}";
-do
-    ignore_str=0
-    
-    for str_to_ignore in "${str_to_ignore_list[@]}";
-    do 
-        if [ $proto == $str_to_ignore ]
-        then
-            ignore_str=1
-            break
-        fi
-    done
-
-    if [ $ignore_str == 0 ]
-    then
-        #echo $app_layer_proto
-        sudo echo $(find $1 *.rules -exec grep "$proto " {} \; | wc -l) $proto
-    fi
-done 

### 2020-02-27 suricata/rust/gen/c-headers/rust-dns-log-gen.h

@@ -1,30 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/*
- * DO NOT EDIT. This file is automatically generated.
- */
-
-#ifndef __RUST_DNS_LOG_GEN_H__
-#define __RUST_DNS_LOG_GEN_H__
-
-json_t * rs_dns_log_json_query(RSDNSTransaction * tx, uint16_t i, uint64_t flags);
-json_t * rs_dns_log_json_answer(RSDNSTransaction * tx, uint64_t flags);
-json_t * rs_dns_log_json_answer_v1(RSDNSTransaction * tx, uint16_t i, uint64_t flags);
-json_t * rs_dns_log_json_authority_v1(RSDNSTransaction * tx, uint16_t i, uint64_t flags);
-
-#endif /* ! __RUST_DNS_LOG_GEN_H__ */

### 2020-02-27 suricata/rust/gen/c-headers/rust-nfs-nfs-gen.h

@@ -1,57 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/*
- * DO NOT EDIT. This file is automatically generated.
- */
-
-#ifndef __RUST_NFS_NFS_GEN_H__
-#define __RUST_NFS_NFS_GEN_H__
-
-void * rs_nfs_state_new(void);
-void rs_nfs_state_free(void * state);
-int8_t rs_nfs_parse_request(Flow * _flow, NFSState * state, void * _pstate, uint8_t * input, uint32_t input_len, void * _data);
-int8_t rs_nfs_parse_request_tcp_gap(NFSState * state, uint32_t input_len);
-int8_t rs_nfs_parse_response(Flow * _flow, NFSState * state, void * _pstate, uint8_t * input, uint32_t input_len, void * _data);
-int8_t rs_nfs_parse_response_tcp_gap(NFSState * state, uint32_t input_len);
-int8_t rs_nfs_parse_request_udp(Flow * _flow, NFSState * state, void * _pstate, uint8_t * input, uint32_t input_len, void * _data);
-int8_t rs_nfs_parse_response_udp(Flow * _flow, NFSState * state, void * _pstate, uint8_t * input, uint32_t input_len, void * _data);
-uint64_t rs_nfs_state_get_tx_count(NFSState * state);
-NFSTransaction * rs_nfs_state_get_tx(NFSState * state, uint64_t tx_id);
-AppLayerGetTxIterTuple rs_nfs_state_get_tx_iterator(NFSState * state, uint64_t min_tx_id, uint64_t * istate);
-void rs_nfs_state_tx_free(NFSState * state, uint64_t tx_id);
[…]

### 2020-02-27 suricata/rust/gen/c-headers/rust-smb-detect-gen.h

@@ -1,31 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/*
- * DO NOT EDIT. This file is automatically generated.
- */
-
-#ifndef __RUST_SMB_DETECT_GEN_H__
-#define __RUST_SMB_DETECT_GEN_H__
-
-uint8_t rs_smb_tx_get_share(SMBTransaction * tx, uint8_t ** buffer, uint32_t * buffer_len);
-uint8_t rs_smb_tx_get_named_pipe(SMBTransaction * tx, uint8_t ** buffer, uint32_t * buffer_len);
-uint8_t rs_smb_tx_get_stub_data(SMBTransaction * tx, uint8_t direction, uint8_t ** buffer, uint32_t * buffer_len);
-uint8_t rs_smb_tx_get_dce_opnum(SMBTransaction * tx, uint16_t * opnum);
-uint8_t rs_smb_tx_get_dce_iface(SMBState * state, SMBTransaction * tx, uint8_t * uuid_ptr, uint16_t uuid_len, uint8_t ver_op, uint16_t ver_check);
-
-#endif /* ! __RUST_SMB_DETECT_GEN_H__ */

### 2020-02-27 suricata/rust/src/ftp/mod.rs

@@ -1,131 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-extern crate libc;
-extern crate nom;
-
-use nom::{digit};
-use std::str;
-use std;
-use std::str::FromStr;
-
-use log::*;
-
-// We transform an integer string into a i64, ignoring surrounding whitespaces
-// We look for a digit suite, and try to convert it.
-// If either str::from_utf8 or FromStr::from_str fail,
-// we fallback to the parens parser defined above
-named!(getu16<u16>,
-    map_res!(
-      map_res!(
-        ws!(digit),
-        str::from_utf8
-      ),
-      FromStr::from_str
-    )
-);
-
-// 227 Entering Passive Mode (212,27,32,66,221,243).
-named!(pub ftp_pasv_response<u16>,
-       do_parse!(
-            tag!("227") >>
-            take_until_and_consume!("(") >>
-            digit >> tag!(",") >> digit >> tag!(",") >>
-            digit >> tag!(",") >> digit >> tag!(",") >>
-            part1: verify!(getu16, |v| v <= std::u8::MAX as u16) >>
-            tag!(",") >>
-            part2: verify!(getu16, |v| v <= std::u8::MAX as u16) >>
-            alt! (tag!(").") | tag!(")")) >>
-            (
-                part1 * 256 + part2
-            )
-        )
-);
-
-
-#[no_mangle]
-pub extern "C" fn rs_ftp_pasv_response(input: *const u8, len: u32) -> u16 {
-    let buf = unsafe{std::slice::from_raw_parts(input, len as usize)};
-    match ftp_pasv_response(buf) {
-        nom::IResult::Done(_, dport) => {
-            return dport;
-        }
[…]

### 2020-02-27 suricata/rust/src/ikev2/ikev2.rs

@@ -1,723 +1,0 @@
-/* Copyright (C) 2017-2018 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-// written by Pierre Chifflier  <chifflier@wzdftpd.net>
-
-use ikev2::ipsec_parser::*;
-use ikev2::state::IKEV2ConnectionState;
-use core;
-use core::{AppProto,Flow,ALPROTO_UNKNOWN,ALPROTO_FAILED,STREAM_TOSERVER,STREAM_TOCLIENT};
-use applayer;
-use parser::*;
-use libc;
-use std;
-use std::ffi::{CStr,CString};
-
-use log::*;
-
-use nom::IResult;
-
-#[repr(u32)]
-pub enum IKEV2Event {
-    MalformedData = 0,
-    NoEncryption,
-    WeakCryptoEnc,
-    WeakCryptoPRF,
-    WeakCryptoDH,
-    WeakCryptoAuth,
-    WeakCryptoNoDH,
-    WeakCryptoNoAuth,
-    InvalidProposal,
-    UnknownProposal,
-}
-
-pub struct IKEV2State {
-    /// List of transactions for this session
-    transactions: Vec<IKEV2Transaction>,
-
-    /// tx counter for assigning incrementing id's to tx's
-    tx_id: u64,
-
-    /// The connection state
-    connection_state: IKEV2ConnectionState,
-
-    /// The transforms proposed by the initiator
-    pub client_transforms : Vec<Vec<IkeV2Transform>>,
-
-    /// The transforms selected by the responder
-    pub server_transforms : Vec<Vec<IkeV2Transform>>,
-
-    /// The encryption algorithm selected by the responder
-    pub alg_enc:  IkeTransformEncType,
-    /// The authentication algorithm selected by the responder
-    pub alg_auth: IkeTransformAuthType,
[…]

### 2020-02-27 suricata/rust/src/smb/smb.rs

@@ -1,2092 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/* TODO
- * - check all parsers for calls on non-SUCCESS status
- */
-
-/*  GAP processing:
- *  - if post-gap we've seen a succesful tx req/res: we consider "re-sync'd"
- */
-
-// written by Victor Julien
-extern crate libc;
-use std;
-use std::mem::transmute;
-use std::str;
-use std::ffi::CStr;
-
-use nom::IResult;
-use std::collections::HashMap;
-
-use core::*;
-use log::*;
-use applayer;
-use applayer::LoggerFlags;
-
-use smb::nbss_records::*;
-use smb::smb1_records::*;
-use smb::smb2_records::*;
-
-use smb::smb1::*;
-use smb::smb2::*;
-use smb::smb3::*;
-use smb::dcerpc::*;
-use smb::session::*;
-use smb::events::*;
-use smb::files::*;
-use smb::smb2_ioctl::*;
-
-pub static mut SURICATA_SMB_FILE_CONFIG: Option<&'static SuricataFileContext> = None;
-
-#[no_mangle]
-pub extern "C" fn rs_smb_init(context: &'static mut SuricataFileContext)
-{
-    unsafe {
-        SURICATA_SMB_FILE_CONFIG = Some(context);
-    }
-}
-
[…]

### 2020-02-27 suricata/rust/src/smb/smb2.rs

@@ -1,785 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-use core::*;
-use log::*;
-use nom::IResult;
-
-use smb::smb::*;
-use smb::smb2_records::*;
-use smb::smb2_session::*;
-use smb::smb2_ioctl::*;
-use smb::dcerpc::*;
-use smb::events::*;
-use smb::files::*;
-
-pub const SMB2_COMMAND_NEGOTIATE_PROTOCOL:      u16 = 0;
-pub const SMB2_COMMAND_SESSION_SETUP:           u16 = 1;
-pub const SMB2_COMMAND_SESSION_LOGOFF:          u16 = 2;
-pub const SMB2_COMMAND_TREE_CONNECT:            u16 = 3;
-pub const SMB2_COMMAND_TREE_DISCONNECT:         u16 = 4;
-pub const SMB2_COMMAND_CREATE:                  u16 = 5;
-pub const SMB2_COMMAND_CLOSE:                   u16 = 6;
-pub const SMB2_COMMAND_FLUSH:                   u16 = 7;
-pub const SMB2_COMMAND_READ:                    u16 = 8;
-pub const SMB2_COMMAND_WRITE:                   u16 = 9;
-pub const SMB2_COMMAND_LOCK:                    u16 = 10;
-pub const SMB2_COMMAND_IOCTL:                   u16 = 11;
-pub const SMB2_COMMAND_CANCEL:                  u16 = 12;
-pub const SMB2_COMMAND_KEEPALIVE:               u16 = 13;
-pub const SMB2_COMMAND_FIND:                    u16 = 14;
-pub const SMB2_COMMAND_CHANGE_NOTIFY:           u16 = 15;
[…]

### 2020-02-27 suricata/rust/src/smb/smb2_records.rs

@@ -1,538 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-use nom::{rest, le_u8, le_u16, le_u32, le_u64, AsBytes, IResult};
-use smb::smb::*;
-
-#[derive(Debug,PartialEq)]
-pub struct Smb2SecBlobRecord<'a> {
-    pub data: &'a[u8],
-}
-
-named!(pub parse_smb2_sec_blob<Smb2SecBlobRecord>,
-    do_parse!(
-         data: rest
-         >> ( Smb2SecBlobRecord {
-                data: data,
-            })
-));
-
-#[derive(Debug,PartialEq)]
-pub struct Smb2Record<'a> {
-    pub direction: u8,    // 0 req, 1 res
-    pub nt_status: u32,
-    pub command: u16,
-    pub message_id: u64,
-    pub tree_id: u32,
-    pub async_id: u64,
-    pub session_id: u64,
-    pub data: &'a[u8],
-}
-
-impl<'a> Smb2Record<'a> {
-    pub fn is_async(&self) -> bool {
-        self.async_id != 0
-    }
-}
-
-named!(pub parse_smb2_request_record<Smb2Record>,
-    do_parse!(
-            server_component: tag!(b"\xfeSMB")
-        >>  hlen: le_u16
-        >>  credit_charge: le_u16
-        >>  channel_seq: le_u16
-        >>  reserved: take!(2)
-        >>  command: le_u16
-        >>  credits_requested: le_u16
-        >>  flags: bits!(tuple!(
-                take_bits!(u8, 2),      // reserved / unused
-                take_bits!(u8, 1),      // replay op
-                take_bits!(u8, 1),      // dfs op
-                take_bits!(u32, 24),    // reserved / unused
[…]

### 2020-02-27 suricata/rust/src/smb/smb_records.rs

@@ -1,53 +1,0 @@
-/* Copyright (C) 2018 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-use nom::{IResult, ErrorKind};
-use log::*;
-
-/// parse a UTF16 string that is null terminated. Normally by 2 null
-/// bytes, but at the end of the data it can also be a single null.
-/// Skip every second byte.
-pub fn smb_get_unicode_string(blob: &[u8]) -> IResult<&[u8], Vec<u8>>
-{
-    SCLogDebug!("get_unicode_string: blob {} {:?}", blob.len(), blob);
-    let mut name : Vec<u8> = Vec::new();
-    let mut c = blob;
-    while c.len() >= 1 {
-        if c.len() == 1 && c[0] == 0 {
-            let rem = &c[1..];
-            SCLogDebug!("get_unicode_string: name {:?}", name);
-            return IResult::Done(rem, name)
-        } else if c.len() == 1 {
-            break;
-        } else if c[0] == 0 && c[1] == 0 {
-            let rem = &c[2..];
-            SCLogDebug!("get_unicode_string: name {:?}", name);
-            return IResult::Done(rem, name)
-        }
-        name.push(c[0]);
-        c = &c[2..];
-    }
-    IResult::Error(error_code!(ErrorKind::Custom(130)))
-}
-
-// parse an ASCII string that is null terminated
-named!(pub smb_get_ascii_string<Vec<u8>>,
-    do_parse!(
-            s: take_until_and_consume!("\x00")
-        >> ( s.to_vec() )
-));
-

### 2020-02-27 suricata/src/app-layer-dns-tcp-rust.c

@@ -1,312 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-
-#include "app-layer-protos.h"
-#include "app-layer-detect-proto.h"
-#include "app-layer-parser.h"
-#include "app-layer-dns-common.h"
-
-#include "util-unittest.h"
-
-#ifdef HAVE_RUST
-
-#include "app-layer-dns-tcp-rust.h"
-#include "rust-dns-dns-gen.h"
-
-#ifdef UNITTESTS
-static void RustDNSTCPParserRegisterTests(void);
-#endif
-
-static int RustDNSTCPParseRequest(Flow *f, void *state,
-        AppLayerParserState *pstate, uint8_t *input, uint32_t input_len,
-        void *local_data, const uint8_t flags)
-{
-    SCLogDebug("RustDNSTCPParseRequest");
-    return rs_dns_parse_request_tcp(f, state, pstate, input, input_len,
-            local_data);
-}
-
-static int RustDNSTCPParseResponse(Flow *f, void *state,
-        AppLayerParserState *pstate, uint8_t *input, uint32_t input_len,
[…]

### 2020-02-27 suricata/src/app-layer-dns-udp-rust.c

@@ -1,378 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-
-#include "app-layer-protos.h"
-#include "app-layer-detect-proto.h"
-#include "app-layer-parser.h"
-#include "app-layer-dns-common.h"
-
-#include "util-unittest.h"
-
-#ifdef HAVE_RUST
-
-#include "app-layer-dns-udp-rust.h"
-#include "rust-dns-dns-gen.h"
-
-#ifdef UNITTESTS
-static void RustDNSUDPParserRegisterTests(void);
-#endif
-
-static int RustDNSUDPParseRequest(Flow *f, void *state,
-        AppLayerParserState *pstate, uint8_t *input, uint32_t input_len,
-        void *local_data, const uint8_t flags)
-{
-    return rs_dns_parse_request(f, state, pstate, input, input_len,
-            local_data);
-}
-
-static int RustDNSUDPParseResponse(Flow *f, void *state,
-        AppLayerParserState *pstate, uint8_t *input, uint32_t input_len,
-        void *local_data, const uint8_t flags)
-{
[…]

### 2020-02-27 suricata/src/app-layer-dns-udp.c

@@ -1,985 +1,0 @@
-/* Copyright (C) 2013 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- * \author Victor Julien <victor@inliniac.net>
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-
-#include "conf.h"
-#include "util-misc.h"
-
-#include "debug.h"
-#include "decode.h"
-
-#include "flow-util.h"
-
-#include "threads.h"
-
-#include "util-print.h"
-#include "util-pool.h"
-#include "util-debug.h"
-
-#include "stream-tcp-private.h"
-#include "stream-tcp-reassemble.h"
-#include "stream-tcp.h"
-#include "stream.h"
-
-#include "app-layer-protos.h"
-#include "app-layer-parser.h"
-
-#include "util-spm.h"
-#include "util-unittest.h"
-
-#include "app-layer-dns-udp.h"
-
-#ifdef HAVE_RUST
-#include "app-layer-dns-udp-rust.h"
-#endif
-
-#ifndef HAVE_RUST
-/** \internal
- *  \brief Parse DNS request packet
- */
-static int DNSUDPRequestParse(Flow *f, void *dstate,
-                              AppLayerParserState *pstate,
-                              uint8_t *input, uint32_t input_len,
-                              void *local_data, const uint8_t flags)
-{
-    DNSState *dns_state = (DNSState *)dstate;
-
-    SCLogDebug("starting %u", input_len);
-
-    if (input == NULL && AppLayerParserStateIssetFlag(pstate, APP_LAYER_PARSER_EOF)) {
-        SCReturnInt(1);
-    }
-
-    /** \todo remove this when PP is fixed to enforce ipproto */
-    if (f != NULL && f->proto != IPPROTO_UDP)
[…]

### 2020-02-27 suricata/src/app-layer-events.c

@@ -1,141 +1,0 @@
-/* Copyright (C) 2014 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Victor Julien <victor@inliniac.net>
- * \author Anoop Saldanha <anoopsaldanha@gmail.com>
- */
-
-#include "suricata-common.h"
-#include "decode.h"
-#include "flow.h"
-#include "app-layer-events.h"
-#include "app-layer-parser.h"
-#include "util-enum.h"
-
-/* events raised during protocol detection are stored in the
- * packets storage, not in the flow. */
-SCEnumCharMap app_layer_event_pkt_table[ ] = {
-    { "APPLAYER_MISMATCH_PROTOCOL_BOTH_DIRECTIONS",
-      APPLAYER_MISMATCH_PROTOCOL_BOTH_DIRECTIONS },
-    { "APPLAYER_WRONG_DIRECTION_FIRST_DATA",
-      APPLAYER_WRONG_DIRECTION_FIRST_DATA },
-    { "APPLAYER_DETECT_PROTOCOL_ONLY_ONE_DIRECTION",
-      APPLAYER_DETECT_PROTOCOL_ONLY_ONE_DIRECTION },
-    { "APPLAYER_PROTO_DETECTION_SKIPPED",
-      APPLAYER_PROTO_DETECTION_SKIPPED },
-    { "APPLAYER_NO_TLS_AFTER_STARTTLS",
-      APPLAYER_NO_TLS_AFTER_STARTTLS },
-    { "APPLAYER_UNEXPECTED_PROTOCOL",
-      APPLAYER_UNEXPECTED_PROTOCOL },
-    { NULL,
-      -1 },
-};
-
-int AppLayerGetPktEventInfo(const char *event_name, int *event_id)
-{
[…]

### 2020-02-27 suricata/src/app-layer-htp-file.h

@@ -1,34 +1,0 @@
-/* Copyright (C) 2007-2011 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Victor Julien <victor@inliniac.net>
- *
- */
-
-#ifndef __APP_LAYER_HTP_FILE_H__
-#define __APP_LAYER_HTP_FILE_H__
-
-int HTPFileOpen(HtpState *, const uint8_t *, uint16_t, const uint8_t *, uint32_t, uint64_t, uint8_t);
-int HTPFileStoreChunk(HtpState *, const uint8_t *, uint32_t, uint8_t);
-int HTPFileClose(HtpState *, const uint8_t *, uint32_t, uint8_t, uint8_t);
-
-void HTPFileParserRegisterTests(void);
-
-#endif /* __APP_LAYER_HTP_FILE_H__ */

### 2020-02-27 suricata/src/app-layer-ikev2.c

@@ -1,52 +1,0 @@
-/* Copyright (C) 2015 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Pierre Chifflier <chifflier@wzdftpd.net>
- *
- * Parser for IKEv2 application layer running on UDP port 500.
- */
-
-#include "suricata-common.h"
-#include "stream.h"
-#include "conf.h"
-
-#include "util-unittest.h"
-
-#include "app-layer-detect-proto.h"
-#include "app-layer-parser.h"
-
-#include "app-layer-ikev2.h"
-
-#ifdef HAVE_RUST
-
-#include "rust-ikev2-ikev2-gen.h"
-
-void RegisterIKEV2Parsers(void)
-{
-    rs_register_ikev2_parser();
-}
-
-#else /* HAVE_RUST */
-
-void RegisterIKEV2Parsers(void)
-{
-}
-
-#endif /* HAVE_RUST */

### 2020-02-27 suricata/src/counters.h

@@ -1,150 +1,0 @@
-/* Copyright (C) 2007-2015 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Anoop Saldanha <anoopsaldanha@gmail.com>
- * \author Victor Julien <victor@inliniac.net>
- */
-
-#ifndef __COUNTERS_H__
-#define __COUNTERS_H__
-
-/* forward declaration of the ThreadVars structure */
-struct ThreadVars_;
-
-/**
- * \brief Container to hold the counter variable
- */
-typedef struct StatsCounter_ {
-    int type;
-
-    /* local id for this counter in this thread */
-    uint16_t id;
-
-    /* global id, used in output */
-    uint16_t gid;
-
-    /* counter value(s): copies from the 'private' counter */
-    uint64_t value;     /**< sum of updates/increments, or 'set' value */
-    uint64_t updates;   /**< number of updates (for avg) */
-
-    /* when using type STATS_TYPE_Q_FUNC this function is called once
-     * to get the counter value, regardless of how many threads there are. */
-    uint64_t (*Func)(void);
-
-    /* name of the counter */
-    const char *name;
-
-    /* the next perfcounter for this tv's tm instance */
-    struct StatsCounter_ *next;
-} StatsCounter;
-
-/**
- * \brief Stats Context for a ThreadVars instance
- */
-typedef struct StatsPublicThreadContext_ {
-    /* flag set by the wakeup thread, to inform the client threads to sync */
-    uint32_t perf_flag;
-
-    /* pointer to the head of a list of counters assigned under this context */
-    StatsCounter *head;
-
-    /* holds the total no of counters already assigned for this perf context */
-    uint16_t curr_id;
-
[…]

### 2020-02-27 suricata/src/decode-afl.c

@@ -1,169 +1,0 @@
-/* Copyright (C) 2007-2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-#include "conf.h"
-#include "decode.h"
-#include "util-debug.h"
-#include "util-mem.h"
-#include "app-layer-detect-proto.h"
-#include "app-layer.h"
-#include "tm-threads.h"
-#include "util-error.h"
-#include "util-print.h"
-#include "tmqh-packetpool.h"
-#include "util-profiling.h"
-#include "pkt-var.h"
-#include "util-mpm-ac.h"
-
-#include "output.h"
-#include "output-flow.h"
-
-#include "defrag.h"
-#include "flow.h"
-
-#ifdef AFLFUZZ_DECODER
-
-/* stateful processing of data as packets. Because AFL in case of a
- * crash will only safe the last input, we dump all the inputs to a
- * directory 'dump' with a unique timestamp for the serie and an
- * incrementing 'id' so that we can 'replay' it in
- * DecoderParseDataFromFileSerie().
- */
-int DecoderParseDataFromFile(char *filename, DecoderFunc Decoder) {
-    uint8_t buffer[65536];
-
-    struct timeval ts;
-    memset(&ts, 0, sizeof(ts));
-    gettimeofday(&ts, NULL);
-
-    uint32_t cnt = 0;
-
-    DefragInit();
-    FlowInitConfig(FLOW_QUIET);
-
-    ThreadVars tv;
-    memset(&tv, 0, sizeof(tv));
-    DecodeThreadVars *dtv = DecodeThreadVarsAlloc(&tv);
[…]

### 2020-02-27 suricata/src/decode-gre.c

@@ -1,413 +1,0 @@
-/* Copyright (C) 2007-2013 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \ingroup decode
- *
- * @{
- */
-
-
-/**
- * \file
- *
- * \author Breno Silva <breno.silva@gmail.com>
- *
- * Decodes GRE
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-#include "decode.h"
-#include "decode-events.h"
-#include "decode-gre.h"
-
-#include "util-unittest.h"
-#include "util-debug.h"
-
-/**
- * \brief Function to decode GRE packets
- */
-
-int DecodeGRE(ThreadVars *tv, DecodeThreadVars *dtv, Packet *p, uint8_t *pkt, uint32_t len, PacketQueue *pq)
-{
-    uint32_t header_len = GRE_HDR_LEN;
-    GRESreHdr *gsre = NULL;
-
-    StatsIncr(tv, dtv->counter_gre);
-
-    if(len < GRE_HDR_LEN)    {
-        ENGINE_SET_INVALID_EVENT(p, GRE_PKT_TOO_SMALL);
-        return TM_ECODE_FAILED;
-    }
-
-    p->greh = (GREHdr *)pkt;
-    if(p->greh == NULL)
-        return TM_ECODE_FAILED;
-
-    SCLogDebug("p %p pkt %p GRE protocol %04x Len: %d GRE version %x",
-        p, pkt, GRE_GET_PROTO(p->greh), len,GRE_GET_VERSION(p->greh));
-
-    switch (GRE_GET_VERSION(p->greh))
-    {
-        case GRE_VERSION_0:
-
-            /* GRE version 0 doenst support the fields below RFC 1701 */
-
-            /**
[…]

### 2020-02-27 suricata/src/detect-ack.c

@@ -1,355 +1,0 @@
-/* Copyright (C) 2007-2016 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Brian Rectanus <brectanu@gmail.com>
- * \author Victor Julien <victor@inliniac.net>
- *
- * Implements the "ack" keyword.
- */
-
-#include "suricata-common.h"
-#include "debug.h"
-#include "decode.h"
-#include "detect.h"
-
-#include "detect-parse.h"
-#include "detect-engine.h"
-#include "detect-engine-mpm.h"
-#include "detect-engine-prefilter.h"
-#include "detect-engine-prefilter-common.h"
-
-#include "detect-ack.h"
-
-#include "util-byte.h"
-#include "util-unittest.h"
-#include "util-unittest-helper.h"
-#include "util-debug.h"
-
-/* prototypes */
-static int DetectAckSetup(DetectEngineCtx *, Signature *, const char *);
-static int DetectAckMatch(ThreadVars *, DetectEngineThreadCtx *,
-                          Packet *, const Signature *, const SigMatchCtx *);
-static void DetectAckRegisterTests(void);
-static void DetectAckFree(void *);
-static int PrefilterSetupTcpAck(DetectEngineCtx *de_ctx, SigGroupHead *sgh);
-static _Bool PrefilterTcpAckIsPrefilterable(const Signature *s);
-
-void DetectAckRegister(void)
-{
-    sigmatch_table[DETECT_ACK].name = "ack";
[…]

### 2020-02-27 suricata/src/detect-base64-decode.h

@@ -1,33 +1,0 @@
-/* Copyright (C) 2015 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-#ifndef __DETECT_BASE64_DECODE_H__
-#define __DETECT_BASE64_DECODE_H__
-
-#include "app-layer-template.h"
-
-typedef struct DetectBase64Decode_ {
-    uint32_t bytes;
-    uint32_t offset;
-    uint8_t relative;
-} DetectBase64Decode;
-
-void DetectBase64DecodeRegister(void);
-int DetectBase64DecodeDoMatch(DetectEngineThreadCtx *, const Signature *,
-    const SigMatchData *, uint8_t *, uint32_t);
-
-#endif /* __DETECT_BASE64_DECODE_H__ */

### 2020-02-27 suricata/src/detect-classtype.c

@@ -1,321 +1,0 @@
-/* Copyright (C) 2007-2010 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Anoop Saldanha <anoopsaldanha@gmail.com>
- *
- * Implements classtype keyword.
- */
-
-#include "suricata-common.h"
-#include "decode.h"
-
-#include "detect.h"
-#include "detect-parse.h"
-#include "detect-engine.h"
-#include "detect-engine-mpm.h"
-#include "detect-classtype.h"
-#include "flow-var.h"
-#include "util-classification-config.h"
-#include "util-error.h"
-#include "util-debug.h"
-#include "util-unittest.h"
-
-#define PARSE_REGEX "^\\s*([a-zA-Z][a-zA-Z0-9-_]*)\\s*$"
-
-static pcre *regex = NULL;
-static pcre_extra *regex_study = NULL;
-
-static int DetectClasstypeSetup(DetectEngineCtx *, Signature *, const char *);
-static void DetectClasstypeRegisterTests(void);
-
-/**
- * \brief Registers the handler functions for the "Classtype" keyword.
- */
-void DetectClasstypeRegister(void)
-{
-    sigmatch_table[DETECT_CLASSTYPE].name = "classtype";
-    sigmatch_table[DETECT_CLASSTYPE].desc = "information about the classification of rules and alerts";
-    sigmatch_table[DETECT_CLASSTYPE].url = DOC_URL DOC_VERSION "/rules/meta.html#classtype";
-    sigmatch_table[DETECT_CLASSTYPE].Match = NULL;
-    sigmatch_table[DETECT_CLASSTYPE].Setup = DetectClasstypeSetup;
-    sigmatch_table[DETECT_CLASSTYPE].Free  = NULL;
[…]

### 2020-02-27 suricata/src/detect-csum.c

@@ -1,1076 +1,0 @@
-/* Copyright (C) 2007-2013 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Anoop Saldanha <anoopsaldanha@gmail.com>
- *
- * Implements checksum keyword.
- */
-
-#include "suricata-common.h"
-#include "debug.h"
-#include "decode.h"
-
-#include "detect.h"
-#include "detect-parse.h"
-
-#include "detect-csum.h"
-
-#include "util-unittest.h"
-#include "util-debug.h"
-
-#include "pkt-var.h"
-#include "host.h"
-#include "util-profiling.h"
-
-/* prototypes for the "ipv4-csum" rule keyword */
-static int DetectIPV4CsumMatch(ThreadVars *, DetectEngineThreadCtx *,
-        Packet *, const Signature *, const SigMatchCtx *);
-static int DetectIPV4CsumSetup(DetectEngineCtx *, Signature *, const char *);
-static void DetectIPV4CsumFree(void *);
-
-/* prototypes for the "tcpv4-csum" rule keyword */
-static int DetectTCPV4CsumMatch(ThreadVars *, DetectEngineThreadCtx *,
-        Packet *, const Signature *, const SigMatchCtx *);
-static int DetectTCPV4CsumSetup(DetectEngineCtx *, Signature *, const char *);
-static void DetectTCPV4CsumFree(void *);
-
-/* prototypes for the "tcpv6-csum" rule keyword */
-static int DetectTCPV6CsumMatch(ThreadVars *, DetectEngineThreadCtx *,
-        Packet *, const Signature *, const SigMatchCtx *);
[…]

### 2020-02-27 suricata/src/detect-detection-filter.c

@@ -1,655 +1,0 @@
-/* Copyright (C) 2007-2010 Open Information Security Foundation
-  *
-  * You can copy, redistribute or modify this Program under the terms of
-  * the GNU General Public License version 2 as published by the Free
-  * Software Foundation.
-  *
-  * This program is distributed in the hope that it will be useful,
-  * but WITHOUT ANY WARRANTY; without even the implied warranty of
-  * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-  * GNU General Public License for more details.
-  *
-  * You should have received a copy of the GNU General Public License
-  * version 2 along with this program; if not, write to the Free Software
-  * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
-  * 02110-1301, USA.
-  */
-
-/**
- * \file
- *
- * \author Gerardo Iglesias <iglesiasg@gmail.com>
- *
- * Implements the detection_filter keyword
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-#include "debug.h"
-#include "decode.h"
-#include "detect.h"
-
-#include "host.h"
-
-#include "detect-detection-filter.h"
-#include "detect-threshold.h"
-#include "detect-parse.h"
-
-#include "util-byte.h"
-#include "util-unittest.h"
-#include "util-unittest-helper.h"
-#include "util-debug.h"
-
-#define TRACK_DST      1
-#define TRACK_SRC      2
-
-/**
- *\brief Regex for parsing our detection_filter options
- */
-#define PARSE_REGEX "^\\s*(track|count|seconds)\\s+(by_src|by_dst|\\d+)\\s*,\\s*(track|count|seconds)\\s+(by_src|by_dst|\\d+)\\s*,\\s*(track|count|seconds)\\s+(by_src|by_dst|\\d+)\\s*$"
-
-static pcre *parse_regex;
-static pcre_extra *parse_regex_study;
-
-static int DetectDetectionFilterMatch(ThreadVars *, DetectEngineThreadCtx *,
-        Packet *, const Signature *, const SigMatchCtx *);
-static int DetectDetectionFilterSetup(DetectEngineCtx *, Signature *, const char *);
-static void DetectDetectionFilterRegisterTests(void);
-static void DetectDetectionFilterFree(void *);
-
-/**
- * \brief Registration function for detection_filter: keyword
[…]

### 2020-02-27 suricata/src/detect-dnp3.c

@@ -1,1042 +1,0 @@
-/* Copyright (C) 2015 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-#include "suricata-common.h"
-
-#include "stream.h"
-
-#include "detect.h"
-#include "detect-parse.h"
-#include "detect-dnp3.h"
-#include "detect-engine.h"
-#include "detect-engine-content-inspection.h"
-
-#include "app-layer-dnp3.h"
-
-static int g_dnp3_match_buffer_id = 0;
-static int g_dnp3_data_buffer_id = 0;
-
-/**
- * The detection struct.
- */
-typedef struct DetectDNP3_ {
-    union {
-        struct {
-            /* Function code for function code detection. */
-            uint8_t  function_code;
-        };
-        struct {
-            /* Internal indicator flags for IIN detection. */
-            uint16_t ind_flags;
-        };
-        struct {
-            /* Object info for object detection. */
-            uint8_t  obj_group;
-            uint8_t  obj_variation;
-        };
-    };
-} DetectDNP3;
-
-/**
- * Indicator names to value mappings (Snort compatible).
- */
-DNP3Mapping DNP3IndicatorsMap[] = {
-    {"device_restart",        0x8000},
-    {"device_trouble",        0x4000},
-    {"local_control",         0x2000},
-    {"need_time",             0x1000},
-    {"class_3_events",        0x0800},
-    {"class_2_events",        0x0400},
-    {"class_1_events",        0x0200},
-    {"all_stations",          0x0100},
-
[…]

### 2020-02-27 suricata/src/detect-engine-analyzer.c

@@ -1,1295 +1,0 @@
-/* Copyright (C) 2007-2018 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Eileen Donlon <emdonlo@gmail.com>
- * \author Victor Julien <victor@inliniac.net>
- *
- * Rule analyzers for the detection engine
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-#include "detect.h"
-#include "detect-parse.h"
-#include "detect-engine.h"
-#include "detect-engine-analyzer.h"
-#include "detect-engine-mpm.h"
-#include "conf.h"
-#include "detect-content.h"
-#include "detect-flow.h"
-#include "detect-flags.h"
-#include "util-print.h"
-
-static int rule_warnings_only = 0;
-static FILE *rule_engine_analysis_FD = NULL;
-static FILE *fp_engine_analysis_FD = NULL;
-static pcre *percent_re = NULL;
-static pcre_extra *percent_re_study = NULL;
-static char log_path[PATH_MAX];
-
-typedef struct FpPatternStats_ {
-    uint16_t min;
-    uint16_t max;
-    uint32_t cnt;
-    uint64_t tot;
-} FpPatternStats;
-
-static FpPatternStats fp_pattern_stats[DETECT_SM_LIST_MAX];
-
-static void FpPatternStatsAdd(int list, uint16_t patlen)
-{
-    if (list < 0 || list >= DETECT_SM_LIST_MAX)
-        return;
-
-    FpPatternStats *f = &fp_pattern_stats[list];
-
-    if (f->min == 0)
-        f->min = patlen;
-    else if (patlen < f->min)
-        f->min = patlen;
-
-    if (patlen > f->max)
[…]

### 2020-02-27 suricata/src/detect-engine-content-inspection.c

@@ -1,647 +1,0 @@
-/* Copyright (C) 2007-2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Anoop Saldanha <anoopsaldanha@gmail.com>
- *
- * Performs content inspection on any buffer supplied.
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-
-#include "decode.h"
-
-#include "detect.h"
-#include "detect-engine.h"
-#include "detect-parse.h"
-#include "detect-content.h"
-#include "detect-pcre.h"
-#include "detect-isdataat.h"
-#include "detect-bytetest.h"
-#include "detect-bytejump.h"
-#include "detect-byte-extract.h"
-#include "detect-replace.h"
-#include "detect-engine-content-inspection.h"
-#include "detect-uricontent.h"
-#include "detect-urilen.h"
-#include "detect-bsize.h"
-#include "detect-lua.h"
-#include "detect-base64-decode.h"
-#include "detect-base64-data.h"
-
-#include "app-layer-dcerpc.h"
-
-#include "util-spm.h"
-#include "util-debug.h"
-#include "util-print.h"
-
-#include "util-unittest.h"
-#include "util-unittest-helper.h"
-#include "util-profiling.h"
-
-#ifdef HAVE_LUA
-#include "util-lua.h"
-#endif
-
-/**
- * \brief Run the actual payload match functions
- *
- * The following keywords are inspected:
- * - content, including all the http and dce modified contents
- * - isdaatat
- * - pcre
- * - bytejump
- * - bytetest
- * - byte_extract
- * - urilen
- * -
- *
- * All keywords are evaluated against the buffer with buffer_len.
- *
- * For accounting the last match in relative matching the
- * det_ctx->buffer_offset int is used.
- *
- * \param de_ctx          Detection engine context
[…]

### 2020-02-27 suricata/src/detect-engine-prefilter.c

@@ -1,628 +1,0 @@
-/* Copyright (C) 2016 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Victor Julien <victor@inliniac.net>
- *
- * Prefilter engine
- *
- * Prefilter engines have as purpose to check for a critical common part of
- * a set of rules. If the condition is present in the traffic, the rules
- * will have to be inspected individually. Otherwise, the rules can be
- * skipped.
- *
- * The best example of this is the MPM. From each rule take a pattern and
- * add it to the MPM state machine. Inspect that in one step and only
- * individually inspect the rules that had a match in MPM.
- *
- * This prefilter API is designed to abstract this logic so that it becomes
- * easier to add other types of prefilters.
- *
- * The prefilter engines are structured as a simple list of engines. Each
- * engine checks for a condition using it's callback function and private
- * data. It then adds the rule match candidates to the PrefilterRuleStore
- * structure.
- *
- * After the engines have run the resulting list of match candidates is
- * sorted by the rule id's so that the individual inspection happens in
- * the correct order.
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-
-#include "detect-engine-prefilter.h"
-#include "detect-engine-mpm.h"
-
-#include "app-layer-parser.h"
[…]

### 2020-02-27 suricata/src/detect-filemagic.h

@@ -1,47 +1,0 @@
-/* Copyright (C) 2007-2011 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Victor Julien <victor@inliniac.net>
- */
-
-#ifndef __DETECT_FILEMAGIC_H__
-#define __DETECT_FILEMAGIC_H__
-
-#ifdef HAVE_MAGIC
-#include "util-spm-bm.h"
-
-typedef struct DetectFilemagicThreadData {
-    magic_t ctx;
-} DetectFilemagicThreadData;
-
-typedef struct DetectFilemagicData {
-    int thread_ctx_id;
-    uint8_t *name; /** name of the file to match */
-    BmCtx *bm_ctx; /** BM context */
-    uint16_t len; /** name length */
-    uint32_t flags;
-} DetectFilemagicData;
-
-/* prototypes */
-int FilemagicGlobalLookup(File *file);
-#endif
-void DetectFilemagicRegister (void);
-
-#endif /* __DETECT_FILEMAGIC_H__ */

### 2020-02-27 suricata/src/detect-ftpbounce.c

@@ -1,574 +1,0 @@
-/* Copyright (C) 2007-2010 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Pablo Rincon <pablo.rincon.crespo@gmail.com>
- *
- * ftpbounce keyword, part of the detection engine.
- */
-
-#include "suricata-common.h"
-#include "debug.h"
-#include "decode.h"
-#include "detect.h"
-#include "detect-parse.h"
-#include "detect-engine.h"
-#include "detect-engine-mpm.h"
-#include "detect-engine-state.h"
-#include "detect-content.h"
-
-#include "app-layer.h"
-#include "app-layer-parser.h"
-#include "app-layer-ftp.h"
-#include "util-unittest.h"
-#include "util-unittest-helper.h"
-#include "util-debug.h"
-#include "flow.h"
-#include "flow-var.h"
-#include "flow-util.h"
-#include "threads.h"
-#include "detect-ftpbounce.h"
-#include "stream-tcp.h"
-#include "util-byte.h"
-
-static int DetectFtpbounceALMatch(ThreadVars *, DetectEngineThreadCtx *,
-        Flow *, uint8_t, void *, void *,
-        const Signature *, const SigMatchCtx *);
-
-static int DetectFtpbounceSetup(DetectEngineCtx *, Signature *, const char *);
-static void DetectFtpbounceRegisterTests(void);
-static int g_ftp_request_list_id = 0;
-
-static int InspectFtpRequest(ThreadVars *tv,
-        DetectEngineCtx *de_ctx, DetectEngineThreadCtx *det_ctx,
[…]

### 2020-02-27 suricata/src/detect-http-headers.c

@@ -1,37 +1,0 @@
-/* Copyright (C) 2007-2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-#include "detect-http-accept.h"
-#include "detect-http-accept-enc.h"
-#include "detect-http-accept-lang.h"
-#include "detect-http-connection.h"
-#include "detect-http-content-len.h"
-#include "detect-http-content-type.h"
-#include "detect-http-referer.h"
-#include "detect-http-headers.h"
-
-void DetectHttpHeadersRegister(void)
-{
-    RegisterHttpHeadersAccept();
-    RegisterHttpHeadersAcceptEnc();
-    RegisterHttpHeadersAcceptLang();
-    RegisterHttpHeadersReferer();
-    RegisterHttpHeadersConnection();
-    RegisterHttpHeadersContentLen();
-    RegisterHttpHeadersContentType();
-}
-

### 2020-02-27 suricata/src/detect-http-method.c

@@ -1,914 +1,0 @@
-/* Copyright (C) 2007-2010 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \ingroup httplayer
- *
- * @{
- */
-
-
-/**
- * \file
- *
- * \author Brian Rectanus <brectanu@gmail.com>
- *
- * Implements the http_method keyword
- */
-
-#include "suricata-common.h"
-#include "threads.h"
-#include "debug.h"
-#include "decode.h"
-#include "detect.h"
-
-#include "detect-parse.h"
-#include "detect-engine.h"
-#include "detect-engine-mpm.h"
-#include "detect-engine-state.h"
-#include "detect-content.h"
-#include "detect-pcre.h"
-
-#include "flow.h"
-#include "flow-var.h"
-#include "flow-util.h"
-
-#include "util-debug.h"
-#include "util-unittest.h"
-#include "util-unittest-helper.h"
-#include "util-spm.h"
-
-#include "app-layer.h"
-#include "app-layer-parser.h"
-
-#include "app-layer-htp.h"
-#include "detect-http-method.h"
-#include "detect-engine-hmd.h"
-#include "stream-tcp.h"
-
-static int g_http_method_buffer_id = 0;
-static int DetectHttpMethodSetup(DetectEngineCtx *, Signature *, const char *);
-void DetectHttpMethodRegisterTests(void);
-void DetectHttpMethodFree(void *);
-static _Bool DetectHttpMethodValidateCallback(const Signature *s, const char **sigerror);
-
-/**
- * \brief Registration function for keyword: http_method
- */
-void DetectHttpMethodRegister(void)
-{
-    sigmatch_table[DETECT_AL_HTTP_METHOD].name = "http_method";
[…]

### 2020-02-27 suricata/src/detect-krb5-cname.c

@@ -1,218 +1,0 @@
-/* Copyright (C) 2018 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Pierre Chifflier <chifflier@wzdftpd.net>
- */
-
-#include "suricata-common.h"
-#include "util-unittest.h"
-
-#include "detect-parse.h"
-#include "detect-engine.h"
-#include "detect-engine-mpm.h"
-#include "detect-engine-state.h"
-#include "detect-engine-prefilter.h"
-#include "detect-engine-content-inspection.h"
-
-#include "detect-krb5-cname.h"
-
-#ifdef HAVE_RUST
-#include "rust.h"
-#include "app-layer-krb5.h"
-#include "rust-krb-detect-gen.h"
-
-static int g_krb5_cname_buffer_id = 0;
-
-struct Krb5PrincipalNameDataArgs {
-    int local_id;  /**< used as index into thread inspect array */
-    void *txv;
-};
-
-static int DetectKrb5CNameSetup(DetectEngineCtx *de_ctx, Signature *s, const char *arg)
-{
-    DetectBufferSetActiveList(s, g_krb5_cname_buffer_id);
-
-    if (DetectSignatureSetAppProto(s, ALPROTO_KRB5) != 0)
-        return -1;
-
-    return 0;
-}
-
-static InspectionBuffer *GetKrb5CNameData(DetectEngineThreadCtx *det_ctx,
-        const DetectEngineTransforms *transforms,
-        Flow *_f, const struct Krb5PrincipalNameDataArgs *cbdata,
-        int list_id, bool first)
-{
-    SCEnter();
-
[…]

### 2020-02-27 suricata/src/detect-ssh-proto.c

@@ -1,205 +1,0 @@
-/* Copyright (C) 2007-2016 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \ingroup sshlayer
- *
- * @{
- */
-
-
-/**
- * \file
- *
- * \author Victor Julien <victor@inliniac.net>
- *
- * Implements support ssh_proto sticky buffer
- */
-
-#include "suricata-common.h"
-#include "threads.h"
-#include "decode.h"
-
-#include "detect.h"
-#include "detect-parse.h"
-#include "detect-engine.h"
-#include "detect-engine-mpm.h"
-#include "detect-engine-state.h"
-#include "detect-engine-prefilter.h"
-#include "detect-engine-content-inspection.h"
-
-#include "app-layer.h"
-#include "app-layer-parser.h"
-#include "app-layer-ssh.h"
-#include "detect-ssh-proto.h"
-
-#define KEYWORD_NAME "ssh_proto"
-#define KEYWORD_DOC "ssh-keywords.html#ssh-proto"
-#define BUFFER_NAME "ssh_protocol"
-#define BUFFER_DESC "ssh protocol field"
-static int g_buffer_id = 0;
-
-/** \brief SSH Protocol Mpm prefilter callback
- *
- *  \param det_ctx detection engine thread ctx
- *  \param p packet to inspect
- *  \param f flow to inspect
- *  \param txv tx to inspect
- *  \param pectx inspection context
- */
-static void PrefilterTxSshRequestProtocol(DetectEngineThreadCtx *det_ctx,
[…]

### 2020-02-27 suricata/src/detect-tls-cert-serial.c

@@ -1,577 +1,0 @@
-/* Copyright (C) 2017 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Mats Klepsland <mats.klepsland@gmail.com>
- *
- * Implements support for tls_cert_serial keyword.
- */
-
-#include "suricata-common.h"
-#include "threads.h"
-#include "debug.h"
-#include "decode.h"
-#include "detect.h"
-
-#include "detect-parse.h"
-#include "detect-engine.h"
-#include "detect-engine-mpm.h"
-#include "detect-engine-prefilter.h"
-#include "detect-content.h"
-#include "detect-pcre.h"
-
-#include "flow.h"
-#include "flow-util.h"
-#include "flow-var.h"
-
-#include "util-debug.h"
-#include "util-unittest.h"
-#include "util-spm.h"
-#include "util-print.h"
-
-#include "stream-tcp.h"
-
-#include "app-layer.h"
-#include "app-layer-ssl.h"
-#include "detect-tls-cert-serial.h"
-
-#include "util-unittest.h"
-#include "util-unittest-helper.h"
-
-static int DetectTlsSerialSetup(DetectEngineCtx *, Signature *, const char *);
-static void DetectTlsSerialRegisterTests(void);
-static InspectionBuffer *GetData(DetectEngineThreadCtx *det_ctx,
-        const DetectEngineTransforms *transforms,
[…]

### 2020-02-27 suricata/src/detect-tls-version.c

@@ -1,514 +1,0 @@
-/* Copyright (C) 2007-2016 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Victor Julien <victor@inliniac.net>
- *
- * Implements the tls.version keyword
- */
-
-#include "suricata-common.h"
-#include "threads.h"
-#include "debug.h"
-#include "decode.h"
-
-#include "detect.h"
-#include "detect-parse.h"
-
-#include "detect-engine.h"
-#include "detect-engine-mpm.h"
-#include "detect-engine-state.h"
-
-#include "flow.h"
-#include "flow-var.h"
-#include "flow-util.h"
-
-#include "util-debug.h"
-#include "util-unittest.h"
-#include "util-unittest-helper.h"
-
-#include "app-layer.h"
-#include "app-layer-parser.h"
-
-#include "app-layer-ssl.h"
-#include "detect-tls-version.h"
-
-#include "stream-tcp.h"
-
-/**
- * \brief Regex for parsing "id" option, matching number or "number"
- */
-#define PARSE_REGEX  "^\\s*([A-z0-9\\.]+|\"[A-z0-9\\.]+\")\\s*$"
-
-static pcre *parse_regex;
-static pcre_extra *parse_regex_study;
-
-static int DetectTlsVersionMatch (ThreadVars *, DetectEngineThreadCtx *,
-        Flow *, uint8_t, void *, void *,
-        const Signature *, const SigMatchCtx *);
[…]

### 2020-02-27 suricata/src/detect.h

@@ -1,1459 +1,0 @@
-/* Copyright (C) 2007-2014 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Victor Julien <victor@inliniac.net>
- */
-
-#ifndef __DETECT_H__
-#define __DETECT_H__
-
-#include "suricata-common.h"
-
-#include "flow.h"
-
-#include "detect-engine-proto.h"
-#include "detect-reference.h"
-#include "detect-metadata.h"
-#include "detect-engine-register.h"
-#include "packet-queue.h"
-
-#include "util-prefilter.h"
-#include "util-mpm.h"
-#include "util-spm.h"
-#include "util-hash.h"
-#include "util-hashlist.h"
-#include "util-debug.h"
-#include "util-error.h"
-#include "util-radix-tree.h"
-#include "util-file.h"
-#include "reputation.h"
-
-#include "detect-mark.h"
-
-#include "stream.h"
-
-#include "util-var-name.h"
-
-#include "app-layer-events.h"
-
-#define DETECT_MAX_RULE_SIZE 8192
-
-#define DETECT_TRANSFORMS_MAX 16
-
-/* forward declarations for the structures from detect-engine-sigorder.h */
-struct SCSigOrderFunc_;
-struct SCSigSignatureWrapper_;
-
-/*
-
-  The detection engine groups similar signatures/rules together. Internally a
-  tree of different types of data is created on initialization. This is it's
-  global layout:
-
-   For TCP/UDP
-
-   - Flow direction
-   -- Protocol
-   -=- Src address
-   -==- Dst address
-   -===- Src port
-   -====- Dst port
[…]

### 2020-02-27 suricata/src/output-json-flow.h

@@ -1,32 +1,0 @@
-/* Copyright (C) 2007-2010 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * Victor Julien <victor@inliniac.net>
- */
-
-#ifndef __OUTPUT_JSON_FLOW_H__
-#define __OUTPUT_JSON_FLOW_H__
-
-void JsonFlowLogRegister(void);
-#ifdef HAVE_LIBJANSSON
-void JsonAddFlow(Flow *f, json_t *js, json_t *hjs);
-#endif /* HAVE_LIBJANSSON */
-
-#endif /* __OUTPUT_JSON_FLOW_H__ */

### 2020-02-27 suricata/src/output-json-http.c

@@ -1,755 +1,0 @@
-/* Copyright (C) 2007-2013 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Tom DeCanio <td@npulsetech.com>
- *
- * Implements HTTP JSON logging portion of the engine.
- */
-
-#include "suricata-common.h"
-#include "debug.h"
-#include "detect.h"
-#include "pkt-var.h"
-#include "conf.h"
-
-#include "threads.h"
-#include "threadvars.h"
-#include "tm-threads.h"
-
-#include "util-print.h"
-#include "util-unittest.h"
-
-#include "util-debug.h"
-
-#include "output.h"
-#include "app-layer-htp.h"
-#include "app-layer-htp-xff.h"
-#include "app-layer.h"
-#include "app-layer-parser.h"
-#include "util-privs.h"
-#include "util-buffer.h"
-#include "util-proto-name.h"
-#include "util-logopenfile.h"
-#include "util-time.h"
-#include "util-crypt.h"
-#include "output-json.h"
-#include "output-json-alert.h"
-#include "output-json-http.h"
-
-#ifdef HAVE_LIBJANSSON
-
-typedef struct LogHttpFileCtx_ {
-    LogFileCtx *file_ctx;
-    uint32_t flags; /** Store mode */
-    uint64_t fields;/** Store fields */
-    HttpXFFCfg *xff_cfg;
-    HttpXFFCfg *parent_xff_cfg;
-    OutputJsonCommonSettings cfg;
-} LogHttpFileCtx;
-
-typedef struct JsonHttpLogThread_ {
-    LogHttpFileCtx *httplog_ctx;
[…]

### 2020-02-27 suricata/src/output-json-template-rust.c

@@ -1,197 +1,0 @@
-/* Copyright (C) 2018 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/*
- * TODO: Update \author in this file and in output-json-template.h.
- * TODO: Remove SCLogNotice statements, or convert to debug.
- * TODO: Implement your app-layers logging.
- */
-
-/**
- * \file
- *
- * \author FirstName LastName <yourname@domain>
- *
- * Implement JSON/eve logging app-layer Template.
- */
-
-#include "suricata-common.h"
-#include "debug.h"
-#include "detect.h"
-#include "pkt-var.h"
-#include "conf.h"
-
-#include "threads.h"
-#include "threadvars.h"
-#include "tm-threads.h"
-
-#include "util-unittest.h"
-#include "util-buffer.h"
-#include "util-debug.h"
-#include "util-byte.h"
-
-#include "output.h"
-#include "output-json.h"
-
-#include "app-layer.h"
-#include "app-layer-parser.h"
-
-#include "app-layer-template-rust.h"
-#include "output-json-template-rust.h"
-
-#if defined(HAVE_LIBJANSSON) && defined(HAVE_RUST)
-
-#include "rust-applayertemplate-logger-gen.h"
-
-typedef struct LogTemplateFileCtx_ {
-    LogFileCtx *file_ctx;
-    uint32_t    flags;
-} LogTemplateFileCtx;
-
-typedef struct LogTemplateLogThread_ {
-    LogTemplateFileCtx *templatelog_ctx;
-    uint32_t            count;
-    MemBuffer          *buffer;
[…]

### 2020-02-27 suricata/src/output-json.h

@@ -1,97 +1,0 @@
-/* Copyright (C) 2007-2013 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Tom DeCanio <td@npulsetech.com>
- */
-
-#ifndef __OUTPUT_JSON_H__
-#define __OUTPUT_JSON_H__
-
-#include "suricata-common.h"
-#include "util-buffer.h"
-#include "util-logopenfile.h"
-#include "output.h"
-
-#include "app-layer-htp-xff.h"
-
-void OutputJsonRegister(void);
-
-#ifdef HAVE_LIBJANSSON
-
-enum OutputJsonLogDirection {
-    LOG_DIR_PACKET = 0,
-    LOG_DIR_FLOW,
-    LOG_DIR_FLOW_TOCLIENT,
-    LOG_DIR_FLOW_TOSERVER,
-};
-
-/* helper struct for OutputJSONMemBufferCallback */
-typedef struct OutputJSONMemBufferWrapper_ {
-    MemBuffer **buffer; /**< buffer to use & expand as needed */
-    size_t expand_by;   /**< expand by this size */
-} OutputJSONMemBufferWrapper;
-
-int OutputJSONMemBufferCallback(const char *str, size_t size, void *data);
-
-void CreateJSONFlowId(json_t *js, const Flow *f);
-void JsonTcpFlags(uint8_t flags, json_t *js);
-void JsonFiveTuple(const Packet *, enum OutputJsonLogDirection, json_t *);
-json_t *CreateJSONHeader(const Packet *p,
-        enum OutputJsonLogDirection dir, const char *event_type);
-json_t *CreateJSONHeaderWithTxId(const Packet *p,
[…]

### 2020-02-27 suricata/src/pesquisa.c

@@ -1,20 +1,0 @@
-#include "pesquisa.h"
-#include "string.h"
-
-void logFitness(void *key, int signatureID, double fitness){
-    FILE *fp = fopen("./rulesFitness.txt", "a");
-
-    char *keyword = (char *) key;
-    //fitness = abs(fitness);
-
-    fprintf(fp, " - %s: %lf", keyword, fitness);
-
-    if(strcmp(keyword, "threshold") == 0){
-        //fprintf(fp, "\n");
-    }
-
-    fflush(fp);
-
-    fclose(fp);
-    printf("SID: %d - %s keyword fitness: %lf\n", signatureID, keyword, fitness);
-}

### 2020-02-27 suricata/src/tm-threads-common.h

@@ -1,94 +1,0 @@
-/* Copyright (C) 2007-2013 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Victor Julien <victor@inliniac.net>
- * \author Anoop Saldanha <anoopsaldanha@gmail.com>
- */
-
-#ifndef __TM_THREADS_COMMON_H__
-#define __TM_THREADS_COMMON_H__
-
-/** \brief Thread Model Module id's.
- *
- *  \note anything added here should also be added to TmModuleTmmIdToString
- *        in tm-modules.c
- */
-typedef enum {
-    TMM_FLOWWORKER,
-    TMM_DECODENFQ,
-    TMM_VERDICTNFQ,
-    TMM_RECEIVENFQ,
-    TMM_RECEIVEPCAP,
-    TMM_RECEIVEPCAPFILE,
-    TMM_DECODEPCAP,
-    TMM_DECODEPCAPFILE,
-    TMM_RECEIVEPFRING,
-    TMM_DECODEPFRING,
-    TMM_RESPONDREJECT,
-    TMM_DECODEIPFW,
-    TMM_VERDICTIPFW,
-    TMM_RECEIVEIPFW,
-    TMM_RECEIVEERFFILE,
-    TMM_DECODEERFFILE,
-    TMM_RECEIVEERFDAG,
-    TMM_DECODEERFDAG,
-    TMM_RECEIVEAFP,
-    TMM_DECODEAFP,
-    TMM_RECEIVENETMAP,
-    TMM_DECODENETMAP,
-    TMM_ALERTPCAPINFO,
-    TMM_RECEIVEMPIPE,
[…]

### 2020-02-27 suricata/src/util-affinity.c

@@ -1,322 +1,0 @@
-/* Copyright (C) 2010-2016 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/** \file
- *
- *  \author Eric Leblond <eric@regit.org>
- *
- *  CPU affinity related code and helper.
- */
-
-#include "suricata-common.h"
-#define _THREAD_AFFINITY
-#include "util-affinity.h"
-#include "util-cpu.h"
-#include "conf.h"
-#include "threads.h"
-#include "queue.h"
-#include "runmodes.h"
-
-ThreadsAffinityType thread_affinity[MAX_CPU_SET] = {
-    {
-        .name = "receive-cpu-set",
-        .mode_flag = EXCLUSIVE_AFFINITY,
-        .prio = PRIO_MEDIUM,
-        .lcpu = 0,
-    },
-    {
-        .name = "worker-cpu-set",
-        .mode_flag = EXCLUSIVE_AFFINITY,
-        .prio = PRIO_MEDIUM,
-        .lcpu = 0,
-    },
-    {
-        .name = "verdict-cpu-set",
-        .mode_flag = BALANCED_AFFINITY,
-        .prio = PRIO_MEDIUM,
-        .lcpu = 0,
-    },
-    {
-        .name = "management-cpu-set",
-        .mode_flag = BALANCED_AFFINITY,
-        .prio = PRIO_MEDIUM,
-        .lcpu = 0,
-    },
-
-};
-
-int thread_affinity_init_done = 0;
-
-/**
- * \brief find affinity by its name
- * \retval a pointer to the affinity or NULL if not found
- */
-ThreadsAffinityType * GetAffinityTypeFromName(const char *name)
-{
-    int i;
-    for (i = 0; i < MAX_CPU_SET; i++) {
-        if (!strcmp(thread_affinity[i].name, name)) {
[…]

### 2020-02-27 suricata/src/util-mpm-hs.c

@@ -1,2176 +1,0 @@
-/* Copyright (C) 2007-2016 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Jim Xu <jim.xu@windriver.com>
- * \author Justin Viiret <justin.viiret@intel.com>
- *
- * MPM pattern matcher that calls the Hyperscan regex matcher.
- */
-
-#include "suricata-common.h"
-#include "suricata.h"
-
-#include "detect.h"
-#include "detect-parse.h"
-#include "detect-engine.h"
-
-#include "conf.h"
-#include "util-debug.h"
-#include "util-unittest.h"
-#include "util-unittest-helper.h"
-#include "util-memcmp.h"
-#include "util-mpm-hs.h"
-#include "util-memcpy.h"
-#include "util-hash.h"
-#include "util-hash-lookup3.h"
-#include "util-hyperscan.h"
-
-#ifdef BUILD_HYPERSCAN
-
-#include <hs.h>
-
-void SCHSInitCtx(MpmCtx *);
-void SCHSInitThreadCtx(MpmCtx *, MpmThreadCtx *);
-void SCHSDestroyCtx(MpmCtx *);
-void SCHSDestroyThreadCtx(MpmCtx *, MpmThreadCtx *);
-int SCHSAddPatternCI(MpmCtx *, uint8_t *, uint16_t, uint16_t, uint16_t,
-                     uint32_t, SigIntId, uint8_t);
-int SCHSAddPatternCS(MpmCtx *, uint8_t *, uint16_t, uint16_t, uint16_t,
-                     uint32_t, SigIntId, uint8_t);
-int SCHSPreparePatterns(MpmCtx *mpm_ctx);
-uint32_t SCHSSearch(const MpmCtx *mpm_ctx, MpmThreadCtx *mpm_thread_ctx,
[…]

### 2020-02-27 suricata/src/util-radix-tree.c

@@ -1,4093 +1,0 @@
-/* Copyright (C) 2007-2010 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Anoop Saldanha <anoopsaldanha@gmail.com>
- *
- * Implementation of radix trees
- */
-
-#include "suricata-common.h"
-#include "util-radix-tree.h"
-#include "util-debug.h"
-#include "util-error.h"
-#include "util-ip.h"
-#include "util-unittest.h"
-#include "util-memcmp.h"
-
-/**
- * \brief Allocates and returns a new instance of SCRadixUserData.
- *
- * \param netmask The netmask entry (cidr) that has to be made in the new
- *                SCRadixUserData instance
- * \param user    The user data that has to be set for the above
- *                netmask in the newly created SCRadixUserData instance.
- *
- * \retval user_data Pointer to a new instance of SCRadixUserData.
- */
-static SCRadixUserData *SCRadixAllocSCRadixUserData(uint8_t netmask, void *user)
-{
-    SCRadixUserData *user_data = SCMalloc(sizeof(SCRadixUserData));
-    if (unlikely(user_data == NULL)) {
-        SCLogError(SC_ERR_MEM_ALLOC, "Error allocating memory");
-        return NULL;
-    }
-
-    memset(user_data, 0, sizeof(SCRadixUserData));
-
-    user_data->netmask = netmask;
-    user_data->user = user;
-
-    return user_data;
-}
-
-/**
- * \brief Deallocates an instance of SCRadixUserData.
- *
- * \param user_data Pointer to the instance of SCRadixUserData that has to be
- *                  freed.
- */
-static void SCRadixDeAllocSCRadixUserData(SCRadixUserData *user_data)
-{
-    SCFree(user_data);
-
-    return;
-}
-
-/**
- * \brief Appends a user_data instance(SCRadixUserData) to a
[…]

### 2020-02-27 suricata/src/util-reference-config.c

@@ -1,808 +1,0 @@
-/* Copyright (C) 2007-2010 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/**
- * \file
- *
- * \author Anoop Saldanha <anoopsaldanha@gmail.com>
- */
-
-#include "suricata-common.h"
-#include "detect.h"
-#include "detect-engine.h"
-#include "util-hash.h"
-
-#include "util-reference-config.h"
-#include "conf.h"
-#include "util-unittest.h"
-#include "util-error.h"
-#include "util-debug.h"
-#include "util-fmemopen.h"
-
-/* Regex to parse each line from reference.config file.  The first substring
- * is for the system name and the second for the url */
-/*-----------------------------------------------------------system-------------------url----*/
-#define SC_RCONF_REGEX "^\\s*config\\s+reference\\s*:\\s*([a-zA-Z][a-zA-Z0-9-_]*)\\s+(.+)\\s*$"
-
-/* Default path for the reference.conf file */
-#define SC_RCONF_DEFAULT_FILE_PATH CONFIG_DIR "/reference.config"
-
-static pcre *regex = NULL;
-static pcre_extra *regex_study = NULL;
-
-/* the hash functions */
-uint32_t SCRConfReferenceHashFunc(HashTable *ht, void *data, uint16_t datalen);
-char SCRConfReferenceHashCompareFunc(void *data1, uint16_t datalen1,
-                                     void *data2, uint16_t datalen2);
-void SCRConfReferenceHashFree(void *ch);
-
-/* used to get the reference.config file path */
-static const char *SCRConfGetConfFilename(const DetectEngineCtx *de_ctx);
-
-void SCReferenceConfInit(void)
-{
-    const char *eb = NULL;
-    int eo;
-    int opts = 0;
-
[…]

### 2020-02-27 suricata/src/util-running-modes.c

@@ -1,56 +1,0 @@
-/* Copyright (C) 2013 Open Information Security Foundation
- *
- * You can copy, redistribute or modify this Program under the terms of
- * the GNU General Public License version 2 as published by the Free
- * Software Foundation.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * version 2 along with this program; if not, write to the Free Software
- * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA
- * 02110-1301, USA.
- */
-
-/** \file
- *
- *  \author Eric Leblond <eric@regit.org>
- */
-
-#include "suricata-common.h"
-#include "config.h"
-#include "app-layer-detect-proto.h"
-#include "app-layer.h"
-#include "app-layer-parser.h"
-#include "util-unittest.h"
-#include "util-debug.h"
-#include "conf-yaml-loader.h"
-#include "util-running-modes.h"
-
-int ListKeywords(const char *keyword_info)
-{
-    if (ConfYamlLoadFile(DEFAULT_CONF_FILE) != -1)
-        SCLogLoadConfig(0, 0);
-    MpmTableSetup();
-    SpmTableSetup();
-    AppLayerSetup();
-    SigTableSetup(); /* load the rule keywords */
-    SigTableList(keyword_info);
-    exit(EXIT_SUCCESS);
-}
-
-int ListAppLayerProtocols()
-{
-    if (ConfYamlLoadFile(DEFAULT_CONF_FILE) != -1)
-        SCLogLoadConfig(0, 0);
-    MpmTableSetup();
-    SpmTableSetup();
-    AppLayerSetup();
-    AppLayerListSupportedProtocols();
-
-    exit(EXIT_SUCCESS);
-}
-

### 2020-02-27 paper/paper.tex
pass on RQ1

@@
 \definecolor{Gray}{gray}{0.9}
+\usepackage[skins]{tcolorbox}
 
@@
 
-\Fix{summarize results}
+\Fix{summarize results} We believe these results provide initial
+evidence on the effectiveness of \tname{}.
 
@@
 
-\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~For some of the attacks listed on
-Table~\ref{table:attacks}, we obtained pcap files directly from
-trusted security websites (\eg{},
-NetReseC~\cite{pcap-attacks}). Recall that \tname\ takes pcap files on
-input. As explained on Section~\ref{sec:minimization}, these files
-abstract network messages~\cite{pcap} and can be used to avoid network
-traffic during rule synthesis. In addition to these sources, we used
-the Nikto2~\cite{nikto} open source security tool as follows to obtain
-pcap files for http and tcp attacks. We configured an http server
-(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
+\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~Recall that \tname\ takes
+pcap files on input. As explained on Section~\ref{sec:minimization},
+these files abstract network messages~\cite{pcap} and can be used to
+avoid network traffic during rule synthesis. For some of the attacks
+listed on Table~\ref{table:attacks}, we obtained pcap files directly
+from trusted security websites (\eg{},
+NetReseC~\cite{pcap-attacks}\Fix{confirm}).\Mar{reminder: we still
+  don't have any example of attack listed on Table~\ref{table:attacks}
+  whose pcap file was obtained from those websites.} In addition to
+these sources, we used the Nikto2~\cite{nikto} open source security
+tool as follows to obtain pcap files for (non-flooding) http and tcp
+attacks. We configured an http server (Apache2) and ran Nikto2 to
+attack that server. Simultenously, we ran
 WireShark~\cite{wireshark-net-monitor} to record the network traffic
@@
 them. It is worth noting that some rules are triggered multiple times,
-indicating that there are multiple instances of the same attack
-encoded in Nikto2\Mar{@Lucas, diga exatamente como vc. obtem as 
-variacoes. isto eh que eh importante para a proxima sentenca.}. 
-\Luc{Eh feito da mesma forma q foi descrita acima. Quando um ataque 
-produz varios pacotes que geram alertas, nos criamos um pcap com um 
-dos pacotes e utilizamos ele como para gerar a regra maxima inicial. 
-Os outros pacotes sao salvos em um outro pcap e sao utilizados para 
-calcular o recall.} RQ3 evaluates the importance of these variants to \tname{}'s
-performance. We used NMAP~\cite{netmap} to reproduce the Ping Scan
-attack (see Section~\ref{sec:active-recon}) and hping3~\cite{hping3}
-to reproduce the SYN Flood attack (see
+indicating that Nikto2 expresses multiple variants of the same
+attack. RQ3 evaluates the importance of these variants to \tname{}'s
+performance. We used similar methodology to obtain pcap files using
+other security testing tools. We used NMAP~\cite{netmap} to reproduce
+the Ping Scan attack (see Section~\ref{sec:active-recon}) and
+hping3~\cite{hping3} to reproduce the SYN Flood attack (see
 Section~\ref{sec:dos}).
+
+%% \Mar{@Lucas, diga exatamente como vc. obtem as variacoes. isto eh que
+%%   eh importante para a proxima sentenca.}.  \Luc{Eh feito da mesma
+%%   forma q foi descrita acima. Quando um ataque produz varios pacotes
+%%   que geram alertas, nos criamos um pcap com um dos pacotes e
+%%   utilizamos ele como para gerar a regra maxima inicial.  Os outros
+%%   pacotes sao salvos em um outro pcap e sao utilizados para calcular o
+%%   recall.}
 
@@
 
-\subsubsection{Benign traffic.}~Security
-tools (e.g., perforation testing tools) often use benign traffic to
-evaluate false positive rates. Consequently, different public options
-exist. We used two popular public sources. The Bigflows.pcap data set
-is part of the TcpReplay~\cite{tcpreplay} open source project. This
-data set includes real network traffic of a busy private network
-access point to the Internet. According to the TcpReplay web site
-``This capture is much larger and has a smaller average packet size
-than the previous capture (smallFlows.pcap). It also has many more
-flows and different applications.''. In addition to the Bigflows.pcap
-data set, we also used data sets\Mar{more than one?} made available by
+%% According to the TcpReplay web
+%% site ``This capture is much larger and has a smaller average packet
+%% size than the previous capture (smallFlows.pcap). It also has many
+%% more flows and different applications.''
+
+\subsubsection{Benign traffic.}~Public data sets of bening traffic
+exist; they are important to evaluate the rate of false positives of
+security tools. We used two popular public sources. The Bigflows.pcap
+data set is part of the TcpReplay~\cite{tcpreplay} open source
+project. It includes real network traffic of a busy private network
+access point to the Internet. In addition to the Bigflows.pcap data
+set, we also used data sets\Mar{more than one?}  made available by
 Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
-  of the traffic? as the description for Bigflows.pcap?}. \Luc{@Guilherme ...}
+  of the traffic? as the description for
+  Bigflows.pcap?}. \Luc{@Guilherme ...}
 
 \subsubsection{Ground Truth.}To evaluate \tname\ objectively we need
-to compare the rules it generates with human-written rules from a
-trusted source. We used the ``Emerging Threats Open
-Ruleset''~\cite{emerging-threats-open} data set for that. After
-finding the attack and corresponding rule using the procedure
-described in Section~\ref{subsec:malicious-traffic}, we reproduced the
-attack and confirmed that the rule captures the attack. Column
-``Ground Truth'' from Table~\ref{table:attacks} only shows the
-essential relevant options. Documentation options (\eg{}, ``msg'') are
-discared.
+to compare the rules it generates with \emph{trusted rules}. We used
+the ``Emerging Threats Open Ruleset''~\cite{emerging-threats-open} for
+that. After finding the attack and corresponding rule using the
+procedure described in Section~\ref{subsec:malicious-traffic}, we
+reproduced the attack and confirmed that the rule captures the
+attack. Column ``Ground Truth'' from Table~\ref{table:attacks} only
+shows the essential relevant options. Documentation options (\eg{},
+``msg'') are discared.
 
@@
 \label{sec:answer-rqone}
+
+This section reports on the performance evaluation of \tname. The
+evaluation metric used is the position of the golden rule at the
+ranked list of plausible rules produced by the tool implementing the
+technique (see Figure~\ref{fig:overview}). This metric is a proxy of
+human performance in analyzing the list of plausible rules that
+capture only the malicious traffic provided on input.
+
+Table~\ref{table:results} shows results. Column ``Name'' shows the
+name of the attack, column ``\#~Vars.'' shows the number of variants
+of the attack, column ``\#~Option Cand. (step 1)'' shows the number of
+option candidates reported by \tname\ at step 1, column ``\#~Plausible
+Rules (step 2)'' shows the number of plausible rules reported by
+\tname\ at step 2, column ``Rank (step 3)'' shows the position of the
+golden rule in the ranked list reported at step 3 of the technique.
+It is worth to indicate that column ``\#~Vars.'' shows the input of
+\tname\ that varies across different attacks. Although \tname\ also
+depends on the data sets of benign traffic and existing \suri\ rules,
+these inputs are constant across different executions of the tool. The
+rest of the columns show the progress of the tool at various stages
+and provide an indication of complexity of the synthesis problem.
+
+Observe that the number of option candidates is considerably higher
+for attacks captured with content options compared to attacks not
+involving content options. This happens because for content attacks
+the number of option candidates depend on the unbounded size of the
+http payload whereas for attacks of different protocols the number of
+option candidates depend on the layout of the messages. Also observe
+that the number of plausible rules is much smaller for non-content
+attacks compared to content attacks.
+
+\newcommand{\percTopFiveRanking}{\Fix{X\%}}
+
+Overall, results show that \tname\ performed very well in most of the
+cases. It reported the ground truth rule in \emph{all} cases and
+reported that rule within the top five rules of the ranking in
+\percTopFiveRanking\ of the cases. It is important to note that the
+number of options in the ground truth (Table~\ref{table:attacks},
+column ``Ground Truth'') is not an indication of difficulty to
+synthesize rules. Note that in \Fix{most?} of the cases where the
+ground truth rule contains only one content options the number of
+option candidates and plausible rules produced is high.
+
+\begin{center}
+\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
+\tname\ successfully synthesizes the golden rules in all cases and
+that rule is ranked among the first top five rules in
+\percTopFiveRanking\ of the cases.
+\end{tcolorbox}
+\end{center}
 

### 2020-02-28 simulated-annealing/sa.py
fixed weights

@@ -801,35 +801,35 @@ def sortRules():
 
     print("all rules len:", all_rules_len)
     
-    for w0 in [0.01, 0.25, 0.5, 0.75, 1]:
+    for w0 in [0, 0.25, 0.5, 0.75, 1]:
         #w.append(w0)
-        for w1 in [0.01, 0.25, 0.5, 0.75, 1]:
+        for w1 in [0, 0.25, 0.5, 0.75, 1]:
             #w.append(w1)
-            for w2 in [0.01, 0.25, 0.5, 0.75, 1]:
+            for w2 in [0, 0.25, 0.5, 0.75, 1]:
                 #w.append(w2)
-                for w3 in [0.01, 0.25, 0.5, 0.75, 1]:
+                for w3 in [0, 0.25, 0.5, 0.75, 1]:
                     #w.append(w3)
-                    for w4 in [0.01, 0.25, 0.5, 0.75, 1]:
+                    for w4 in [0, 0.25, 0.5, 0.75, 1]:
                         #w.append(w4)
                         w = [w0,w1,w2,w3,w4]
-                        
-                        print("weights:", str(w))
-                        all_rules_list = sorted(all_rules_list, key=partial(callGetFitness, weights=w))
-                        #exit()
-                        for x, rule in enumerate(all_rules_list):
-                            if rule.sid == 1099019:
-                                golden_rule_pos = all_rules_list.index(rule)
-                            else:
-                                rule.sid=x+1
+                        if w != [0,0,0,0]:
+                            print("weights:", str(w))
+                            all_rules_list = sorted(all_rules_list, key=partial(callGetFitness, weights=w))
+                            #exit()
+                            for x, rule in enumerate(all_rules_list):
+                                if rule.sid == 1099019:
+                                    golden_rule_pos = all_rules_list.index(rule)
+                                else:
+                                    rule.sid=x+1
 
-                        current_pos = all_rules_len-golden_rule_pos
-                        
-                        print(current_pos)
+                            current_pos = all_rules_len-golden_rule_pos
+                            
+                            print(current_pos)
 
-                        if current_pos <= best_pos:
-                            best_pos = current_pos
-                            best_rule_list = copy.deepcopy(all_rules_list)
-                            best_weights = copy.deepcopy(w)
+                            if current_pos <= best_pos:
+                                best_pos = current_pos
+                                best_rule_list = copy.deepcopy(all_rules_list)
+                                best_weights = copy.deepcopy(w)
                     #w.pop()
                 #w.pop()
            # w.pop()

### 2020-02-28 paper/paper.tex
pass on intro

@@
 
-This paper focuses on rule-based \nids{}, which is a very popular kind
-of NIDS used in industry (see Section~\ref{sec:background}). A
+This paper focuses on rule-based \nids{}, which are very popular in
+industry~\cite{proofpoint-etpro,snort-rule-subscriptions}. A
 \emph{rule-based intrusion detector}\footnote{\aka\ signature-based
   intrusion detector.} checks if the network traffic matches a fixed
-set of rules. Figure~\ref{fig:synflood-example} shows an example rule
+set of rules. Figure~\ref{fig:pingscan-example} shows an example rule
 of Suricata~\cite{suricata}, a popular open-source \nids{}. This
-particular rule prescribes a method to capture a denial-of-service
+particular rule prescribes a method to capture an Active
+Reconnaissance attack for discovering which ports of a publicly
+accessible machine are open for connection. Rulesets exist for
+Snort~\cite{snort} and Suricata~\cite{suricata}, providing protection
+against existing attacks. Unfortunately, attackers actively look for
+new mechanisms to exploit systems. For example, Figure~\Fix{plot}
+shows the increase in number of rules added to the ``Emerging Threats
+Open Ruleset''---a popular public data set of \suri\ rules---in the
+past 12 months.
+
+\Fix{add figure here}
+
+Intuitively, rulesets need to be constantly revised to protect a
+system against attackers. Unfortunately, creating these rules is
+time-consuming as the language for describing rules is non-trivial
+(see Section~\ref{sec:example-suricata-rules}) and it is error-prone
+as checking whether the rule captures only malicious traffic is
+non-trivial.
+
+%% IT-security companies capitalize on this phenomena and
+%% offer rulesets on the
+%% market.
+
+This paper proposes \tname{}\footnote{Abbreviation for
+  \textbf{Sy}nthesis of Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a
+technique that uses machine intelligence to automatically synthesize
+rules for rule-based \nids. Our goal is to facilitate the creation of
+rules and, consequently, harden the protection of \nids\ to network
+attacks. \tname{} synthesizes rules from malicious and benign traffic
+and from examples of correct rules.
+
+\Fix{
+denial-of-service
 attack to a server by matching specific conditions about the network
@@
 \nids\ are restricted to a fixed set of rules defined by the network
-system administrator. Updating these rules is tedious and error-prone,
-but it is important as attackers constantly create new strategies to
-circumvent existing rules. IT-security companies capitalize on this
-phenomena and offer rulesets on the
-market~\cite{proofpoint-etpro,snort-rule-subscriptions}.
-
-This paper proposes \tname{}\footnote{Abbreviation for
-  \textbf{Sy}nthesis of Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a
-technique that uses machine intelligence to automatically synthesize
-rules for rule-based \nids. Our goal is to facilitate the creation
-process of rules and, consequently, harden the protection of \nids\ to
-network attacks. \tname{} synthesizes rules from malicious and benign
-traffic and from examples of correct rules.
+system administrator.}\Mar{check...}
 
@@
 the Open Information Security Foundation (OISF)~\cite{oisf}.
-
-\vspace{-1ex}
-\begin{figure}[h!]
-  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
-  \vspace{-2ex}
-  \caption{\suri\ rule for SYN Flood attack.}
-  \label{fig:synflood-example}
-\end{figure}
-\vspace{-2ex}
-
-A
-Suricata rule is divided in three parts---action, header, and rule
+A Suricata rule is divided in three parts---action, header, and rule
 options~\cite{suri-rule-format}. The action part appears as the first
 word in the rule description. An action denotes the task that needs to
-be executed if the rule pattern is satisfied.
-In this example, a
+be executed if the rule pattern is satisfied.  In this example, a
 message will be sent to system administrators if the rule pattern is
@@
 description. It restricts the information flow covered by the
-rule. For this rule, the header is \CodeIn{tcp \$HOME\_NET any ->
-  \$EXTERNAL\_NET any}. It instructs Suricata to inspect \CodeIn{tcp}
-traffic flowing from any port in the home network to any other address
-outside the home network. The variables \CodeIn{\$HOME\_NET} and
-\CodeIn{\$EXTERNAL\_NET} are configurable. The rule options come after
-the header. It is a semi-colon-separated sequence of key-value
-pairs. Rule options serve to document the analyzed traffic (\eg{},
-``msg'', ``classtype'', and ``sid'') and to characterize the attack
-pattern (\eg, ``flags'', ``threshold'', and ``content''). \tname{}
-produces options of the second kind, which are \emph{relevant} to
-capture the attack. For example, the option \CodeIn{content:<arg>} checks if
-the argument is present in the payload of the message.
+rule. The header of this rule is \CodeIn{icmp any any -> any
+  any}. It instructs Suricata to inspect \CodeIn{icmp} traffic flowing
+from any machine and port to any machine and port.\MyComment{ The variables
+\CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are
+configurable.} The rule options come after the header. It is a
+semi-colon-separated sequence of key-value pairs. Rule options serve
+to document the analyzed traffic (\eg{}, ``msg'', ``classtype'', and
+``sid'') and to characterize the attack pattern (\eg, ``flags'',
+``threshold'', and ``content''). \tname{} produces options of the
+second kind, which are \emph{relevant} to capture the attack. For
+example, the option \CodeIn{content:<arg>} checks if the argument is
+present in the payload of the message.
 
@@
   \footnotesize
-  \caption{\label{table:rules}Example options of Suricata.}
+  \caption{\label{table:rules}Example of non-documentation options of Suricata.}
   \vspace{-2ex}
@@
 
-Table~\ref{table:rules} shows a sample of the options of
+Table~\ref{table:rules} shows a sample of non-documentation options of
 Suricata. Column ``\#'' shows the option number, column ``Name'' shows
-the name of the option, and column ``Description'' presents a short
-description of the option purpose. We found that the option
-\CodeIn{content} is particularly prevalent in rules from public
-databases\MyComment{ (see Figure~\ref{fig:distribution-contents})}. This
+the name of the option, and column ``Description'' summarizes
+purpose. The option \CodeIn{content} is very common in rules to
+capture http attacks, which are the most common kind of rules. This
 option takes a sequence of bytes as argument (\eg{},
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[scale=0.28]{figs/ICMP-packet-structure.png}
-\vspace{-2ex}
-\caption{Layout of an ICMP packet.}
-\label{fig:icmp-packet-layout}
-\end{figure}
-%\vspace{-2ex}
-
 Rule options are often associated with specific fields of network
@@
 
+\begin{figure}[t!]
+\centering
+\includegraphics[scale=0.28]{figs/ICMP-packet-structure.png}
+\vspace{-2ex}
+\caption{Layout of an ICMP packet.}
+\label{fig:icmp-packet-layout}
+\end{figure}
+%\vspace{-2ex}
+
 %% More precisely, to trigger an action on a rule
@@
 
-\vspace{-1ex}
+%\vspace{-1ex}
 \begin{figure}[h!]
@@
 legitimate requests cannot be satisfied due to resource exhaustion.
+
+\vspace{-1ex}
+\begin{figure}[h!]
+  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
+  \vspace{-2ex}
+  \caption{\suri\ rule for SYN Flood attack.}
+  \label{fig:synflood-example}
+\end{figure}
+\vspace{-2ex}
+
 Figure~\ref{fig:synflood-example} shows a Suricata rule that captures

### 2020-02-28 paper/paper.tex
pass

@@
 is to discover the set of options in a rule that blocks the malicious
-traffic, but allows the benign traffic to pass.
+traffic, but allows the benign traffic to pass. \Luc{O Suricata normalmente 
+nao bloqueia trafego, apenas gera alertas sobre ele. Para bloquear um 
+pacote ele precisa ser configurado no modo Intrusion Prevention System 
+(IPS) ao inves de IDS: https://blog.starti.com.br/ids-ips/}
 
@@
 frequently in http attacks, which are prevalent. For example,
-\Fix{XXX\%} of the \numrulessuri{} rules from a public data
+\Fix{\Luc{53\%}} of the \numrulessuri{} rules from a public data
 set~\cite{emerging-threats-open} we analyzed prescribe defenses
@@
   data set}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
-  (\%) das 28K regras tem contents.} \Luc{sao 69417 contents nas 26094 regras q utilizam contents} As consequence, we defined a
+  (\%) das 28K regras tem contents.} \Luc{92\%} As consequence, we defined a
 heuristic function that takes the number of content options into
@@
   \caption{Distribution of features.}
-  \Luc{@Marcelo, estamos utilizando mais uma distribuicao agora. Precisamos adicionar o grafico dela tb?}
 \end{figure*}
@@
     JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
-    %% WeBid \MyComment{Local File Inclusion} & RFI & TCP \MyComment{& Local File Inclusion} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
+    WeBid \MyComment{Local File Inclusion} & RFI & HTTP & \Fix{LFI} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
     ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
@@
     Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
+    Apache Struts 2 & \Fix{Source} & \Fix{RCE} & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
+    WordPress Login & \Fix{Source} & Policy Violation & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
     \bottomrule
@@
     SYN Flood & 3 & 2 & 2 & - \\
-    JBoss JMX & 4 & 23 & 381 & 23 \\
-    %%   WeBid & - & 23 & - & -\\    
-    ColdFusion & 6 & 19 & 39 & 5\\
-    .htaccess access & 4 & 17 & 2,458 & 2\\
-    .idq access & 13 & 19 & 1,111 & 2 \\
-    iisadmin access & 3 & 17 & 861 & 1 \\
-    /system32/ in URI & 3 & 20 & 3,665 & 2 \\
-    Script tag in URI & 22 & 22 & - & - \\
+    JBoss JMX & 4 & 23 & 228 & \Fix{23} \\
+    WeBid & 1 & 23 & - & -\\    
+    ColdFusion & 6 & 19 & 34 & \Fix{5}\\
+    .htaccess access & 4 & 17 & 984 & \Fix{2}\\
+    .idq access & 13 & 19 & 511 & \Fix{2} \\
+    iisadmin access & 3 & 17 & \Fix{861} & \Fix{1} \\
+    /system32/ in URI & 3 & 20 & 1,913 & \Fix{2} \\
+    Script tag in URI & 22 & 22 & 786 & - \\
+    Apache Struts 2 & - & - & - & - \\
+    Wordpress Login & - & - & - & - \\
     \bottomrule
@@
 types: Denial-of-Service (DoS), Active Reconnaissance (AREC),
-Privilege escalation (PESC), and Remote Code Execution (RCE). Note
-that the distribution of attacks is not uniform across protocols as
-these distributions are not uniform in practice---http attacks are
-prevalent.
+Privilege escalation (PESC), Remote Code Execution (RCE), \Luc{Local 
+File Inclusion (LFI) and Policy Violation (PV)}. Note that the 
+distribution of attacks is not uniform across protocols as these 
+distributions are not uniform in practice---http attacks are prevalent.
 
@@
   don't have any example of attack listed on Table~\ref{table:attacks}
-  whose pcap file was obtained from those websites.} In addition to
+  whose pcap file was obtained from those websites.} \Luc{Ataques Apache 
+  e WordPress. @Guilherme adicionar fontes} In addition to
 these sources, we used the Nikto2~\cite{nikto} open source security

### 2020-02-28 simulated-annealing/sa.py
update

@@ -859,7 +859,7 @@ def sortMultipleAttacks():
             if i == 0:
                 print("No file was loaded, something is wrong.")
                 exit()
-            break
+            #break
  
 
     all_rules_list = []

### 2020-02-28 paper/paper.tex
pass on intro

@@
 Reconnaissance attack for discovering which ports of a publicly
-accessible machine are open for connection. Rulesets exist for
-Snort~\cite{snort} and Suricata~\cite{suricata}, providing protection
-against existing attacks. Unfortunately, attackers actively look for
-new mechanisms to exploit systems. For example, Figure~\Fix{plot}
-shows the increase in number of rules added to the ``Emerging Threats
-Open Ruleset''---a popular public data set of \suri\ rules---in the
-past 12 months.
+accessible machine are open for connection (see
+Section~\ref{sec:active-recon}). Rulesets exist for Snort~\cite{snort}
+and Suricata~\cite{suricata}, providing protection against existing
+attacks. Unfortunately, attackers actively look for new mechanisms to
+exploit systems. For example, Figure~\Fix{plot} shows the increase in
+number of rules added to the ``Emerging Threats Open Ruleset''---a
+popular public data set of \suri\ rules---in the past 12 months.
 
@@
 system against attackers. Unfortunately, creating these rules is
-time-consuming as the language for describing rules is non-trivial
-(see Section~\ref{sec:example-suricata-rules}) and it is error-prone
-as checking whether the rule captures only malicious traffic is
-non-trivial.
+time-consuming and error-prone. The language for describing rules is
+non-trivial (see Section~\ref{sec:example-suricata-rules}) and testing
+correctness of a rule requires data and careful inspection.
 
@@
 rules for rule-based \nids. Our goal is to facilitate the creation of
-rules and, consequently, harden the protection of \nids\ to network
-attacks. \tname{} synthesizes rules from malicious and benign traffic
-and from examples of correct rules.
+rules and, consequently, help system administrators to harden the
+protection of \nids\ to network attacks. \tname{} synthesizes rules
+from malicious and benign traffic and from examples of correct rules.
 

### 2020-02-28 paper/paper.tex
pass

@@
     Ping Scan & NMAP & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
-    SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\
-
+    SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12;
+    \textbf{threshold}: type both, track by\_dst, count 5000, seconds
+    5;\\
+    \midrule
     JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
-    WeBid \MyComment{Local File Inclusion} & RFI & HTTP & \Fix{LFI} & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
+    WeBid \MyComment{Local File Inclusion} & RFI & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
     ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
@@
     Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
-    Apache Struts 2 & \Fix{Source} & \Fix{RCE} & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
-    WordPress Login & \Fix{Source} & Policy Violation & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
+    Apache Struts 2 & \Fix{Source} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
+    WordPress Login & \Fix{Source} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
     \bottomrule
@@
 \subsubsection{Attacks.}~Table~\ref{table:attacks} shows the selection
-of attacks considered in our evaluation.  Column ``Name'' shows the
-name of the attack, column ``Source'' shows the source where we
-obtained data to reproduce the attack (see
-Section~\ref{subsec:malicious-traffic}), column ``Type'' shows the
-kind of attack, column ``Protocol'' shows the name of the protocol
-explored in the attack, and column ``Ground Truth'' shows the rule
-options to capture the attack. The rationale for the selection were
-(1) to cover attacks with different characteristics, (2) to cover
-attacks of various protocols, and (3) to explore different features of
-\tname. As for (1), the selection includes attacks of the following
-types: Denial-of-Service (DoS), Active Reconnaissance (AREC),
-Privilege escalation (PESC), Remote Code Execution (RCE), \Luc{Local 
-File Inclusion (LFI) and Policy Violation (PV)}. Note that the 
-distribution of attacks is not uniform across protocols as these 
-distributions are not uniform in practice---http attacks are prevalent.
+of attacks we analyzed. Column ``Name'' shows the name of the attack,
+column ``Source'' shows the source where we obtained data to reproduce
+the attack (see Section~\ref{subsec:malicious-traffic}), column
+``Type'' shows the kind of attack, column ``Protocol'' shows the name
+of the protocol explored in the attack, and column ``Ground Truth''
+shows the rule options to capture the attack. The rationale for the
+selection were (1) to cover attacks with different characteristics,
+(2) to cover attacks of various protocols, and (3) to explore
+different features of \tname. As for (1), the selection includes
+attacks of the following types: Denial-of-Service
+(DoS)~\cite{denial-of-service}, Active Reconnaissance
+(AREC)~\cite{active-reconnaissance}, Privilege escalation
+(PVE)~\cite{privilege-escalation}, Remote Code Execution
+(RCE)~\cite{remote-code-execution}, Local File Inclusion
+(LFI)~\cite{local-file-inclusion}, and Policy Violation
+(PV)~\Fix{local-file-inclusion}. Note that the distribution of attacks
+is not uniform across protocols as these distributions are not uniform
+in practice---http attacks are prevalent.
 

### 2020-02-28 paper/paper.tex
pass on table

@@
     5;\\
+    UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
+    UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
+    5; \\
+    Black Nurse & NetReseC & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3; \\
     \midrule
@@
     Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
-    Apache Struts 2 & \Fix{Source} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
-    WordPress Login & \Fix{Source} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
+    Apache Struts 2 & WRCCDC & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
+    WordPress Login & WRCCDC & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
     \bottomrule
@@
     \midrule
-    Ping Scan & 5 & 5 & 2 & - \\
-    SYN Flood & 3 & 2 & 2 & - \\
-    JBoss JMX & 4 & 23 & 228 & \Fix{23} \\
+    Ping Scan & 6 & 5 & 2 & - \\
+    SYN Flood & 4 & 2 & 2 & - \\
+    JBoss JMX & 5 & 23 & 228 & \Fix{23} \\
     WeBid & 1 & 23 & - & -\\    
-    ColdFusion & 6 & 19 & 34 & \Fix{5}\\
-    .htaccess access & 4 & 17 & 984 & \Fix{2}\\
-    .idq access & 13 & 19 & 511 & \Fix{2} \\
-    iisadmin access & 3 & 17 & \Fix{861} & \Fix{1} \\
-    /system32/ in URI & 3 & 20 & 1,913 & \Fix{2} \\
-    Script tag in URI & 22 & 22 & 786 & - \\
+    ColdFusion & 7 & 19 & 34 & \Fix{5}\\
+    .htaccess access & 5 & 17 & 984 & \Fix{2}\\
+    .idq access & 14 & 19 & 511 & \Fix{2} \\
+    iisadmin access & 4 & 17 & \Fix{861} & \Fix{1} \\
+    /system32/ in URI & 4 & 20 & 1,913 & \Fix{2} \\
+    Script tag in URI & 23 & 22 & 786 & - \\
     Apache Struts 2 & - & - & - & - \\
@@
   whose pcap file was obtained from those websites.} \Luc{Ataques Apache 
-  e WordPress. @Guilherme adicionar fontes} In addition to
+  e WordPress.
+@Guilherme adicionar fontes} \Gui{O NetReseC oferece uma lista de sites que oferecem datasets, como o proprio Stratosphere, que estamos utilizando para trafego normal.
+Para esses dois ataques novos com content, peguei do WRCCDC - Western Regional Collegiate Cyber Defense Competition: https://archive.wrccdc.org/pcaps/}In addition to
 these sources, we used the Nikto2~\cite{nikto} open source security
@@
   of the traffic? as the description for
-  Bigflows.pcap?}. \Luc{@Guilherme ...}
+  Bigflows.pcap?}.
+\Luc{@Guilherme ...} \Gui{The Stratosphere IPS Project has a sister project called the Malware Capture Facility Project that is responsible for making the long-term captures.
+This project is continually obtaining malware and normal data to feed the Stratosphere IPS (copy paste do site).
+No site tem uma lista de datasets, e embaixo dele descricao da origem desses datasets, como de uma rede universitaria, um usuario normal, etc.
+https://www.stratosphereips.org/datasets-normal}
 

### 2020-02-28 simulated-annealing/sa.py
wordpress attack

@@ -43,6 +43,8 @@ contents_dict["idq"] = {'GET':[], '/scripts':[], '/samples':[], '/search':[], '/
 
 contents_dict["system"] = {'GET':[], '/c':[], '/winnt':[], '/system32/':["http_uri", "nocase"], 'cmd.exe?':[], '/c+dir+':[], '/OG':[], 'HTTP/1.1':[], '192.168.1.108':[], 'User-Agent:':[], 'Mozilla':[], '/5.00':[], '(Nikto':[], '/2.1.5)':[],  '(Evasions:':[], 'None)':[], '(Test:':[], '000121)':[], 'Connection:':[], 'Keep-Alive':[]}
 
+contents_dict["wordpress"] = {'POST':[], 'deflate': [], 'Hungry4Apples%21': [], '&pwd=': [], '58.0)': [], '10.47.26.186': [], 'Gecko': [], 'Win64': [], 'log=': [], 'application': [], '1': [], '/x-www-form-urlencoded': [], 'Accept-Language:': [], '&wp-submit=': [], 'Firefox': [], '0.9,*': [], 'WP+Cookie+check': [], 'admin': [], '%2Fwordpress%2Fwp-admin%2F': [], 'testcookie=': [], '/html,application': [], 'Connection:': [], 'gzip,': [], '/58.0': [], '/wordpress': [], '/20100101': [], '/xml': [], '0.5': [], 'wordpress_test_cookie=': [], 'x64': [], '10.0': [], 'ess': [], '(Windows': [], 'text': [], '/5.0': [], 'Log+In': [], 'Content-Type:': [], 'Cookie:': [], '/xhtml+xml,application': [], 'Referer:': [], '/10.47.26.186': [], 'HTTP/1.1': [], 'Host:': [], 'en-US,en': [], 'Mozilla': [], 'NT': [], '99': [], 'Accept-Encoding:': [], 'rv:': [], 'Accept:': [], 'redirect_to=': [], 'Content-Length:': [], '/wp-login.php': [], 'User-Agent:': [], 'http:': [], '0.8': [], 'q=': [], 'keep-alive': [], 'Upgrade-Insecure-Requests:': []}
+
 if args.attack in contents_dict:
     contents = contents_dict[args.attack]
 else:
@@ -65,9 +67,6 @@ if str(args.attack) in ["adaptor", "coldfusion", "htaccess", "cron", "jsp", "scr
 else:
[…]

### 2020-02-28 simulated-annealing/sendGoodTraffic.sh

@@ -1,2 +1,2 @@
 suricata -c inputs/suricata-$1.yaml -r Datasets/positive-http.pcap
-suricata -c inputs/suricata-$1.yaml -r Datasets/nikto-without-$1.pcap
+suricata -c inputs/suricata-$1.yaml -r Datasets/nikto.pcap

### 2020-02-28 paper/paper.tex
fix on table

@@
     5; \\
-    Black Nurse & NetReseC & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3; \\
+    Black Nurse & NetReseC & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3;\textbf{threshold}: type both, track by\_dst, count 250, seconds
+    1; \\
     \midrule
@@
     SYN Flood & 4 & 2 & 2 & - \\
+    UDP Teardrop & - & - & - & - \\
+    UDP Flood & - & - & - & - \\
+    Black Nurse & - & - & - & - \\
     JBoss JMX & 5 & 23 & 228 & \Fix{23} \\
@@
     Apache Struts 2 & - & - & - & - \\
-    Wordpress Login & - & - & - & - \\
+    Wordpress Login & 8 & 59 & - & - \\
     \bottomrule
@@
   e WordPress.
-@Guilherme adicionar fontes} \Gui{O NetReseC oferece uma lista de sites que oferecem datasets, como o proprio Stratosphere, que estamos utilizando para trafego normal.
+@Guilherme adicionar fontes} \Gui{O NetReseC possui uma lista de sites (alem dos pcaps que o proprio site ja oferece, como foi obtido o Black Nurse) que oferecem datasets, como o proprio Stratosphere, que estamos utilizando para trafego normal.
 Para esses dois ataques novos com content, peguei do WRCCDC - Western Regional Collegiate Cyber Defense Competition: https://archive.wrccdc.org/pcaps/}In addition to

### 2020-02-28 paper/paper.tex
resolving conflicts

--- 2020-02-28 paper/paper.tex
@@
   \small
-  \caption{\label{table:attacks}List of attacks.}
+  \caption{\label{table:attacks}List of attacks. Non-http attacks
+    appear at the top andcontent-based http attacks appear at the bottom.}
   \centering
@@
     \midrule
-    Ping Scan & NMAP & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
-    SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12;
+    Black Nurse & \cite{pcap-attacks} & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3; \textbf{threshold}: type both, track by\_dst, count 250, seconds 1;\\    
+    Ping Scan & \cite{netmap} & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
+    SYN Flood & \cite{hping3} & DoS & TCP & \textbf{flags}: S,12;
     \textbf{threshold}: type both, track by\_dst, count 5000, seconds
     5;\\
+    UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
+    UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
+    5; \\
     \midrule
-    JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
-    WeBid \MyComment{Local File Inclusion} & RFI & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
-    ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
-    .htaccess access\MyComment{attempted-recon} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
-    .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
-    iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
-    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
-    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
-    Apache Struts 2 & \Fix{Source} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
-    WordPress Login & \Fix{Source} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
+    Apache Struts 2 & \cite{wrccdc} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
+    ColdFusion \MyComment{PESC} & \cite{nikto} & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
+    iisadmin access\MyComment{web-application-attack} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
+    JBoss JMX\MyComment{ Console Beanshell Exploit} & \cite{nikto} & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
+    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
+    WeBid \MyComment{Local File Inclusion} & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
+    WordPress Login & \cite{wrccdc} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
+    .htaccess access\MyComment{attempted-recon} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
+    .idq access\MyComment{EXPLOIT ISAPI} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
+    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
     \bottomrule
@@
     \midrule
-    Ping Scan & 5 & 5 & 2 & - \\
-    SYN Flood & 3 & 2 & 2 & - \\
-    JBoss JMX & 4 & 23 & 228 & \Fix{23} \\
+    Black Nurse & - & - & - & - \\    
+    Ping Scan & 6 & 5 & 2 & - \\
+    SYN Flood & 4 & 2 & 2 & - \\
+    UDP Teardrop & - & - & - & - \\
+    UDP Flood & - & - & - & - \\
+    \midrule
+    Apache Struts 2 & - & - & - & - \\
+    ColdFusion & 7 & 19 & 34 & \Fix{5}\\
+    iisadmin access & 4 & 17 & \Fix{861} & \Fix{1} \\        
+    JBoss JMX & 5 & 23 & 228 & \Fix{23} \\
+    Script tag in URI & 23 & 22 & 786 & - \\
     WeBid & 1 & 23 & - & -\\    
-    ColdFusion & 6 & 19 & 34 & \Fix{5}\\
-    .htaccess access & 4 & 17 & 984 & \Fix{2}\\
-    .idq access & 13 & 19 & 511 & \Fix{2} \\
-    iisadmin access & 3 & 17 & \Fix{861} & \Fix{1} \\
-    /system32/ in URI & 3 & 20 & 1,913 & \Fix{2} \\
-    Script tag in URI & 22 & 22 & 786 & - \\
-    Apache Struts 2 & - & - & - & - \\
-    Wordpress Login & - & - & - & - \\
+    Wordpress Login & 8 & 59 & - & - \\
+    .htaccess access & 5 & 17 & 984 & \Fix{2}\\
+    .idq access & 14 & 19 & 511 & \Fix{2} \\
+    /system32/ in URI & 4 & 20 & 1,913 & \Fix{2} \\
     \bottomrule
@@
 
-%% It is important to note that this setup is \emph{not}
-%% necessary if \emph{pcap} files for the attack already exist. We
-%% proceeded as follows to obtain pcap files for the the attacks we
-%% analyzed: we 1)~instantiated the script (\eg{}, defined the IP of the
-%% web server), 2)~executed the script, 3)~and monitored the traffic with
-%% the Wireshark network monitor. The
-%% Wireshark tool records the traffic and reports pcap files on output
-%% for replaying it.
-
+\vspace{1ex}
 \subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~Recall that \tname\ takes
@@
 from trusted security websites (\eg{},
-NetReseC~\cite{pcap-attacks}\Fix{confirm}).\Mar{reminder: we still
-  don't have any example of attack listed on Table~\ref{table:attacks}
-  whose pcap file was obtained from those websites.} \Luc{Ataques Apache 
-  e WordPress. @Guilherme adicionar fontes} In addition to
-these sources, we used the Nikto2~\cite{nikto} open source security
-tool as follows to obtain pcap files for (non-flooding) http and tcp
-attacks. We configured an http server (Apache2) and ran Nikto2 to
-attack that server. Simultenously, we ran
+NetReseC~\cite{pcap-attacks}). In addition to these sources, we used
+the Nikto2~\cite{nikto} open source security tool as follows to obtain
+pcap files for some of the http attacks. We configured an http server
+(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
 WireShark~\cite{wireshark-net-monitor} to record the network traffic
@@
 Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
-default parameters, which makes it replay all recorded attacks. After
-execution, we inspected the \suri\ log to identify which rules have
-been activated and which packets were involved on each
-activation. With the packet numbers, it is possible to recover the
-corresponding pcap files, using Wireshark, as to ran \tname{} on
-them. It is worth noting that some rules are triggered multiple times,
-indicating that Nikto2 expresses multiple variants of the same
-attack. RQ3 evaluates the importance of these variants to \tname{}'s
-performance. We used similar methodology to obtain pcap files using
-other security testing tools. We used NMAP~\cite{netmap} to reproduce
-the Ping Scan attack (see Section~\ref{sec:active-recon}) and
-hping3~\cite{hping3} to reproduce the SYN Flood attack (see
-Section~\ref{sec:dos}).
-
-%% \Mar{@Lucas, diga exatamente como vc. obtem as variacoes. isto eh que
-%%   eh importante para a proxima sentenca.}.  \Luc{Eh feito da mesma
-%%   forma q foi descrita acima. Quando um ataque produz varios pacotes
-%%   que geram alertas, nos criamos um pcap com um dos pacotes e
-%%   utilizamos ele como para gerar a regra maxima inicial.  Os outros
-%%   pacotes sao salvos em um outro pcap e sao utilizados para calcular o
-%%   recall.}
-
-%% \Mar{what about the variants? where do
-%%   they come from? this is confusing as the narrative above suggests
-%%   the attack is hardcoded in the tool.}\Luc{O trafego de um ataque
-%%   normalmente possui varios pacotes, entre esses varios pacotes alguns
-%%   deles possuem um padrao. As regras sao criadas descrevendo esse
-%%   padrao, com isso um unico ataque pode disparar varios alertas.
-
-%%   Porem, existem ataques q so produzem um alerta, ou seja, a regra criada 
-%%   so detecta um dos pacotes do trafego. Nesses casos, procuramos na internet 
-%%   datasets (arquivos pcap) que contivessem mais pacotes desses ataques.
-%%   Fizemos isso tambem para ataques q nos ja tinhamos variacoes, pois
-%%   (teoricamente) quanto mais variacoes, melhor. Existem alguns sites onde 
-%%   eh possivel encontrar pcaps com o trafego de alguns ataques, n consegui
-%%   adicionar nas referencias: 
-%%   https://www.netresec.com/?page=PcapFiles
-%% }
-
-%% According to the TcpReplay web
-%% site ``This capture is much larger and has a smaller average packet
-%% size than the previous capture (smallFlows.pcap). It also has many
-%% more flows and different applications.''
-
+default parameters, which makes it replay all recorded attacks from
+its benchmark. After execution, we inspected the \suri\ log to
+identify which rules have been activated and which packets were
+involved on each activation. With the packet numbers, it is possible
+to recover the corresponding pcap files, using Wireshark, as to ran
+\tname{} on them separately. It is worth noting that some rules are
+triggered multiple times, indicating that Nikto2 expresses multiple
+variants of the same attack. RQ3 evaluates the importance of these
+variants to \tname{}'s performance. We used similar methodology to
+obtain pcap files using other security testing tools. For example, we
+used NMAP~\cite{netmap} to reproduce the Ping Scan attack (see
+Section~\ref{sec:active-recon}) and hping3~\cite{hping3} to reproduce
+the SYN Flood attack (see Section~\ref{sec:dos}).
+
+\vspace{1ex}
 \subsubsection{Benign traffic.}~Public data sets of bening traffic
@@
 project. It includes real network traffic of a busy private network
-access point to the Internet. In addition to the Bigflows.pcap data
+access point to the Internet. We also used the
+
+\Fix{---------------- estou aqui}
+
+In addition to the Bigflows.pcap data
 set, we also used data sets\Mar{more than one?}  made available by

### 2020-02-28 paper/paper.tex
pass

@@
 \subsection{Step 1: Reverse Engineering}
+\label{sec:reverse-engineering}
 
@@
 
+\newcommand{\numNonContentAttacks}{5}
+\newcommand{\numContentAttacks}{10}
+
 \subsection{Objects of Analysis}
@@
 \subsubsection{Attacks.}~Table~\ref{table:attacks} shows the selection
-of attacks we analyzed. Column ``Name'' shows the name of the attack,
-column ``Source'' shows the source where we obtained data to reproduce
-the attack (see Section~\ref{subsec:malicious-traffic}), column
-``Type'' shows the kind of attack, column ``Protocol'' shows the name
-of the protocol explored in the attack, and column ``Ground Truth''
-shows the rule options to capture the attack. The rationale for the
-selection were (1) to cover attacks with different characteristics,
-(2) to cover attacks of various protocols, and (3) to explore
-different features of \tname. As for (1), the selection includes
-attacks of the following types: Denial-of-Service
+of attacks we analyzed.  Non-http attacks are listed at the top of the
+table whereas http attacks are listed at the bottom. Recall from
+Section~\ref{sec:reverse-engineering} that (1) http attacks are
+prevalent, (2) most http attacks use only content options, and (3)
+most non-http attacks only use field options. We selected from various
+sources \numNonContentAttacks\ non-http attacks and
+\numContentAttacks\ http attacks. The higher number of http attacks
+reflects their popularity. Column ``Name'' shows the name of the
+attack, column ``Source'' shows the source where we obtained data to
+reproduce the attack (see Section~\ref{subsec:malicious-traffic}),
+column ``Type'' shows the kind of attack, column ``Protocol'' shows
+the name of the protocol explored in the attack, and column ``Ground
+Truth'' shows the rule options to capture the attack. The rationale
+for the selection were (1) to cover various kinds of attacks and (2)
+to cover attacks of various protocols. Considering (1), the selection
+includes attacks of the following types: Denial-of-Service
 (DoS)~\cite{denial-of-service}, Active Reconnaissance
 (AREC)~\cite{active-reconnaissance}, Privilege escalation
-(PVE)~\cite{privilege-escalation}, Remote Code Execution
+(PESC)~\cite{privilege-escalation}, Remote Code Execution
 (RCE)~\cite{remote-code-execution}, Local File Inclusion
 (LFI)~\cite{local-file-inclusion}, and Policy Violation
-(PV)~\Fix{local-file-inclusion}. Note that the distribution of attacks
-is not uniform across protocols as these distributions are not uniform
-in practice---http attacks are prevalent.
+(PVE)~\Fix{local-file-inclusion}.
 

### 2020-02-28 paper/paper.tex
done with section 5.1 (for now)

@@
 \subsubsection{Benign traffic.}~Public data sets of bening traffic
-exist; they are important to evaluate the rate of false positives of
+exist and can be used to evaluate the rate of false positives of
 security tools. We used two popular public sources. The Bigflows.pcap
@@
 project. It includes real network traffic of a busy private network
-access point to the Internet. We also used the
-
-\Fix{---------------- estou aqui}
-
-In addition to the Bigflows.pcap data
-set, we also used data sets\Mar{more than one?}  made available by
-Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
-  of the traffic? as the description for
-  Bigflows.pcap?}.
+access point to the Internet. We also used data sets from the
+Stratosphere IPS Project~\cite{stratosphere-normal} including, for
+example, the regular traffic of a university network.
+
+%% In addition to the Bigflows.pcap data
+%% set, we also used data sets\Mar{more than one?}  made available by
+%% Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
+%%   of the traffic? as the description for
+%%   Bigflows.pcap?}.
 \Luc{@Guilherme ...} \Gui{The Stratosphere IPS Project has a sister project called the Malware Capture Facility Project that is responsible for making the long-term captures.
@@
 the ``Emerging Threats Open Ruleset''~\cite{emerging-threats-open} for
-that. After finding the attack and corresponding rule using the
-procedure described in Section~\ref{subsec:malicious-traffic}, we
-reproduced the attack and confirmed that the rule captures the
-attack. Column ``Ground Truth'' from Table~\ref{table:attacks} only
-shows the essential relevant options. Documentation options (\eg{},
-``msg'') are discared.
+that. After finding a matching pair of attack and corresponding rule
+using the procedure described in
+Section~\ref{subsec:malicious-traffic}, we reproduced the attack to
+confirm the rule captures the attack in isolation (and no noise was
+introduced in the process). Column ``Ground Truth'' from
+Table~\ref{table:attacks} only shows the essential
+options. Documentation options (\eg{}, ``msg'') are discared.
 

### 2020-02-28 simulated-annealing/sa.py
removed sortRules

@@ -733,8 +733,8 @@ class Rule:
         str_options = str_options + ' ' + "sid:" + str(self.sid) + ';'
         
         str_protocol = str(self.protocol)
-        if str_protocol == "http":
-            str_protocol = "tcp"
+        #if str_protocol == "http":
+        #    str_protocol = "tcp"
 
         return (str(self.action) + ' ' + str(str_protocol) + ' ' + str(self.header) + ' ' + '(' + str(self.message) + str_options + ')')
 
@@ -799,7 +799,7 @@ def sortRules():
 
 
     exit()
-"""
+    """
     current_pos = 0
     best_pos = math.inf
     best_rule_list = []
@@ -836,7 +836,9 @@ def sortRules():
 
                             if current_pos <= best_pos:
                                 best_pos = current_pos
+                                print("deepcopy start")
                                 best_rule_list = copy.deepcopy(all_rules_list)
+                                print("deepcopy end")
                                 best_weights = copy.deepcopy(w)
                     #w.pop()
                 #w.pop()
@@ -1190,6 +1192,7 @@ def optimizeRule(rule):
         golden_rule.options = {'dsize':0, 'itype':8}
     
     print("golden_rule:", golden_rule)
+    print("all_rules_list len:", len(all_rule_list))
     #print("fit1: ", ruleSizeFitness(golden_rule), "fit2: ", ruleContentsFitness(golden_rule), "fit3: ", rareContentsFitness(golden_rule), "fit4: ", ruleContentsModifiersFitness(golden_rule))
     all_rule_list.insert(0, golden_rule)
     #all_rule_list.append(final_rule)
@@ -1203,12 +1206,14 @@ def optimizeRule(rule):
             writer.write(str(all_rule_list[i])+'\n')
             raw_writer.write(str(all_rule_list[i].getAllAttributesRaw())+'\n')
 
+    regrafit=[]
+
+    """
     all_rule_list, best_weights, best_pos = sortRules()
 
     print("Best Weights:", str(best_weights))
     print("Best Pos: " + str(best_pos) + ',' + str(len(all_rule_list)))
 
-    regrafit=[]
 
     print("pegando precision")
     precision=checkPrecision(all_rule_list)
@@ -1216,7 +1221,7 @@ def optimizeRule(rule):
     recall=checkFalseNegative(all_rule_list)
     i=0
     golden_index = 0
-
+    """
     """for rule in all_rule_list:
[…]

### 2020-02-28 paper/paper.tex
pass on section 3 (example)

@@
 message will be sent to system administrators if the rule pattern is
-satisfied. The header comes after the action in the rule
+satisfied\footnote{Technically, an IPS (Intrusion Prevention System)
+  differs from an IDS in the action taken when an intrusion is
+  observed. IPS systems block malicous traffic. However, IDS and IPS
+  do not differ in the approaches that can be used to detect the
+  intrusion~\cite{ids-ips}.\Mar{check}}. The header comes after the action in the rule
 description. It restricts the information flow covered by the
@@
 This section briefly shows how \tname{} synthesizes rules for one
-example attack. Active Reconnaissance is a method used by malicious
-individuals to collect information about a computing system to
-determine potential vulnerabilities. This method is used as a
-preliminary step towards an actual attack. For example, Ping Scan
-scans the network for IPs of accessible hosts. An attacker sends
-several ICMP Echo requests to a range of IP addresses in the network
-and checks which ones respond to these requests. This method enables
-an attacker to know which machines are active in the network.
+example attack. Active Reconnaissance is typically used as a
+preliminary step towards an actual attack. It is a method to collect
+information about a computing system to determine potential
+vulnerabilities. For example, Ping Scan scans the network for IPs of
+accessible hosts. An attacker sends several ICMP Echo requests to a
+range of IP addresses in the network and checks which ones respond to
+these requests. This method enables an attacker to discover which
+machines are active in the network.
 
@@
 Figure~\ref{fig:pingscan-example} shows a Suricata rule that captures
-a Ping Scan attempt, as documented in NMAP~\cite{netmap}, a network
-discovery and security auditing open-source tool. The relevant options
-(see Section~\ref{sec:example-suricata-rules}) are in bold---the
-option \CodeIn{dsize: 0} indicates that the packet has no bytes in the
-payload and the option \CodeIn{itype: 8} indicates that the packet to
-be captured are ICMP Echo requests.
+a Ping Scan, as documented in NMAP~\cite{netmap}, a network discovery
+and security auditing open-source tool. The essential options to
+capture the attack (see Section~\ref{sec:example-suricata-rules}) are
+in bold---the option \CodeIn{dsize: 0} indicates that the packet has
+no bytes in the payload and the option \CodeIn{itype: 8} indicates
+that the packet to be captured are ICMP Echo requests.
 
 In the following, we briefly show how \tname{} synthesizes rules for
-this attack. The action and the header of the rules that \tname{}
-creates are either inferred or left constant. For example, in this
-case, \tname{} sets the action to \CodeIn{alert} and sets the header
-to \CodeIn{<proto> any any -> any any}, indicating that it infers the
-target protocol and uses the most general flow description possible,
-\ie{}, it instructs the \nids\ to analyze the traffic flowing from/to
-any address and port.  The rationale for this decision is that the
-choice of these parameters are domain-specific and subject to change
-by the system administrator.  The main challenge in synthesizing rules
-is to discover the set of options in a rule that blocks the malicious
-traffic, but allows the benign traffic to pass. \Luc{O Suricata normalmente 
-nao bloqueia trafego, apenas gera alertas sobre ele. Para bloquear um 
-pacote ele precisa ser configurado no modo Intrusion Prevention System 
-(IPS) ao inves de IDS: https://blog.starti.com.br/ids-ips/}
-
-\tname{} proceeds as follows to synthesize rules. First, it analyzes
-the malicious traffic and extracts options expressed in that
-traffic. Recall from Section~\ref{sec:rules-and-packets} that some
-options can be inferred from packet contents. \tname{} finds the
-following options at this stage:
-{\scriptsize{\texttt{\textbf{dsize}:0; \textbf{itype:}8;
+this attack. The action is left constant (\CodeIn{alert}) and the
+header is inferred from the protocol. For example, in this case,
+\tname{} sets the header to \CodeIn{icmp any any -> any any},
+indicating that the rule applies to \CodeIn{icmp} messages and applies
+to the most general flow description, \ie{}, it instructs the
+\nids\ to analyze the traffic flowing from/to any address and port.
+The rationale for this decision is that the choice of these parameters
+are domain-specific and subject to change by the system administrator.
+
+The main challenge in synthesizing rules is to discover the set of
+rule options that capture the malicious traffic, but allow the benign
+traffic to pass. \tname{} proceeds as follows to synthesize rule
+options. First, it analyzes the malicious traffic to extract options
+expressed in that traffic. Recall from
+Section~\ref{sec:rules-and-packets} that some options can be inferred
+from packet contents. \tname{} finds the following five options at
+this stage: {\scriptsize{\texttt{\textbf{dsize}:0; \textbf{itype:}8;
       \textbf{icode:}0; \textbf{icmp\_id:}23570;
@@
 Second, \tname{} uses the benign traffic to mitigate the overfitting
-issue. In this step, \tname{} searches for alternative rules that
+problem. In this step, \tname{} searches for alternative rules that
 preserve the invariant to capture the positive malicious traffic but
@@
 invariant. The search is designed to generate rules containing subsets
-of the options from the seed rule. Considering this example, \tname{}
-produces a total of \pingscanplausible{} plausible rules at the end of
-the search.  In the worst case, \tname{} produces $2^n$ plausible
-rules, where $n$ is the number of options contained in the inferred
-rule ($n=5$, in this case). Rules are discared in the search process
-as they are found to be implausible.
+of the options declared in the seed rule. Considering this example,
+\tname{} produces a total of \pingscanplausible{}\Fix{check!}
+plausible rules at the end of the search.  Note that there are $2^n$
+possible plausible rules, where $n$ is the number of options contained
+in the seed rule ($n=5$, in this case). \tname{} is able to prune the
+search space and discard many implausible rules. It is worth noting
+that the number of plausible rules that \tname\ produces at this step
+can be very high for other example attacks (see column ``\#~Plausible
+Rules'' on Table~\ref{table:results}).
 
@@
 Finally, \tname{} ranks plausible rules according to heuristics
-learned from a database of existing rules. The intuition is that new
-rules should share characteristics with pre-existing rules. For
-example, one heuristic function used by \tname{} leverages the
-\emph{size of a rule} for that. The score of a rule given by such a
-function is proportional to the popularity of the number of options
-declared in rules of a given
-protocol. Figure~\ref{fig:distribution-rule-size} shows the
-distribution of the number of content options associated with rules
-from a public data set~\cite{emerging-threats-open}. Note that rules
-with 2 to 6 options are more common. At the end of this step, \tname{}
-ranks the golden rule at the top of the ranking, including
-\pingscanplausible{} rules. It is worth noting that
-\tname\ consistently ranks the golden rule very high in most of the
-cases even when the seed rule, produced at the first step, contains a
-large number of options (see Table~\ref{table:results}).
+obtained by analyzing a database of existing rules. The general
+intuition is that new rules should share characteristics with
+pre-existing rules. For example, one heuristic function used by
+\tname{} leverages the \emph{size of a rule} for ranking.  The score
+of a rule $r$ that this function produces is proportional to the
+popularity of the number of options in
+$r$. Figure~\ref{fig:distribution-rule-size} shows the distribution of
+the number of content options associated with rules from a public data
+set~\cite{emerging-threats-open}. Note that rules with 2 to 6 options
+are more common. At the end of this step, \tname{} ranks the golden
+rule at the top of the ranking, including \pingscanplausible{}
+rules. It is worth noting that \tname\ consistently ranks the golden
+rule very high in most of the cases even when the technique produces a
+large number of plausible rules (see Table~\ref{table:results}).
 

### 2020-02-28 paper/paper.tex
pass table 3

@@
     Black Nurse & - & - & - & - \\    
-    Ping Scan & 6 & 5 & 2 & - \\
-    SYN Flood & 4 & 2 & 2 & - \\
+    Ping Scan & 6 & 5 & 2 & 1 \\
+    SYN Flood & 4 & 2 & 2 & 2 \\
     UDP Teardrop & - & - & - & - \\
@@
     Apache Struts 2 & - & - & - & - \\
-    ColdFusion & 7 & 19 & 34 & \Fix{5}\\
+    ColdFusion & 7 & 19 & 34 & 1\\
     iisadmin access & 4 & 17 & \Fix{861} & \Fix{1} \\        
-    JBoss JMX & 5 & 23 & 228 & \Fix{23} \\
-    Script tag in URI & 23 & 22 & 786 & - \\
+    JBoss JMX & 5 & 23 & 228 & 5 \\
+    Script tag in URI & 23 & 22 & 786 & 1 \\
     WeBid & 1 & 23 & - & -\\    
     Wordpress Login & 8 & 59 & - & - \\
-    .htaccess access & 5 & 17 & 984 & \Fix{2}\\
-    .idq access & 14 & 19 & 511 & \Fix{2} \\
-    /system32/ in URI & 4 & 20 & 1,913 & \Fix{2} \\
+    .htaccess access & 5 & 17 & 984 & 2\\
+    .idq access & 14 & 19 & 511 & 1 \\
+    /system32/ in URI & 4 & 20 & 1,913 & 1 \\
     \bottomrule

### 2020-02-28 paper/paper.tex
minor change

@@
 \emph{plausible rules}---\ie{}, rules that isolate the malicious
-traffic from the benign traffic---in order of their likelihood to
-reflect user's intent. Figure~\ref{fig:overview} shows the workflow of
+traffic from the benign traffic---ranked in order of similarity with
+existing rules. Figure~\ref{fig:overview} shows the workflow of
 \tname{} as a pipeline of three components, which appear numbered in

### 2020-02-28 paper/paper.tex
pass on 3.0

@@
 
-%% \Mar{Technically, there are $32 (=2^5)$ subsets for a set
-%%   with 5 elements. So, it is important to indicate why Table 3 shows
-%%   sooo many plausible rules and that this is is a very simple rule
-%%   (for illustration).}
-
 Finally, \tname{} ranks plausible rules according to heuristics
@@
 
-\begin{figure*}[ht]
-  \centering
-  \includegraphics[trim=0 395 50 0,clip,width=0.65\textwidth]{figs/nids-workflow}
-  \caption{The \tname\ workflow.}
-  \label{fig:overview}
-  \vspace{-2ex}
-\end{figure*}
-
 \section{Technique}
@@
 
-\tname\ works as follows. First, it produces a potentially
-overspecified rule that captures the malicious traffic, which is
-provided on input in the form of pcap files~\cite{pcap}. In this step,
-\tname\ detects options directly from the input messages (see
-Section~\ref{sec:rules-and-packets}). More precisely, it produces a
-rule containing all possible options satisfied by the traffic. Let us
-refer to that rule as the \emph{seed rule} and its corresponding set
-of options as \emph{seed options}. Note that, as only malicious
-traffic is considered in this step, the seed rule is likely
-overspecified, \ie{}, it may be more restrictive than the golden rule
-(see Section~\ref{sec:suri-metas-coverage}). Consequently, the use of
-that rule to capture attacks associated with the input traffic could
-lead to false negatives.  Second, \tname\ generates rules that weaken
-the constraints associated with the seed options whilst still
-capturing the malicious traffic. Given that every option declared in a
-rule needs to be satisfied for a rule to capture an attack, any subset
-of the set of seed options must capture the malicious traffic by
-definition. \tname{} leverages that observation to search for rules
-containing a proper subset of options from the seed options. However,
-note that that approach alone could result in false positives as it
-imposes no limit on to what extent a rule can be relaxed. \tname{}
-uses benign traffic for that purpose. More precisely, it uses a
-gradient descent-like search to find all plausible rules derived from
-the seed rule, obtained in the previous step. Finally, \tname{} uses
-several heuristic functions derived from a data set of existing rules
-to rank the plausible rules generated in the previous step by their
-similarity to existing rules. The rationale is that many plausible
-rules can be generated in this step and they have different semantics,
-which can differ from user's intent. Similarity is used to approximate
-user's intent.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[trim=0 395 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
+  \caption{The \tname\ workflow.}
+  \label{fig:overview}
+  \vspace{-2ex}
+\end{figure}
+
+\tname\ works as follows. First, it produces a rule that captures the
+malicious traffic. In this step, \tname\ detects options directly from
+the input messages (see Section~\ref{sec:rules-and-packets}). More
+precisely, it produces a rule containing all possible options
+satisfied by the traffic. Let us refer to that rule as the \emph{seed
+  rule} and its corresponding set of options as \emph{seed
+  options}. Note that, as only the malicious traffic is considered in
+this step, the seed rule is likely overspecified, \ie{}, it may be
+more restrictive than the golden rule. Consequently, the use of that
+rule could result in false negatives. Second, \tname\ generates
+alternative rules that 1) weaken the constraints expressed in the seed
+rule while 2) still capturing the malicious traffic. Given that every
+option declared in a rule needs to be satisfied for a rule to capture
+an attack, any subset of the set of seed options, by definition,
+captures the attack characterized by the malicious traffic. \tname{}
+leverages that observation to search for rules containing a proper
+subset of options from the seed options. However, note that that
+approach alone could result in false positives as it imposes no limit
+on to what extent a rule can be relaxed. \tname{} uses benign traffic
+for that purpose. More precisely, it uses a gradient descent-like
+search to find plausible rules derived from the seed rule, obtained in
+the previous step. Finally, \tname{} uses several heuristic functions
+derived from a data set of existing rules to rank the generated
+plausible rules by their similarity with existing rules. Similarity is
+used to approximate user's intent.
 

### 2020-02-28 paper/paper.tex
adding a comment

@@
   \caption{\label{table:attacks}List of attacks. Non-http attacks
-    appear at the top andcontent-based http attacks appear at the bottom.}
+    appear at the top; http attacks appear at the bottom.}
   \centering
@@
     Black Nurse & - & - & - & - \\    
-    Ping Scan & 6 & 5 & 2 & 1 \\
+    Ping Scan & 6 & 5 & 2\Mar{18?} & 1 \\
     SYN Flood & 4 & 2 & 2 & 2 \\

### 2020-02-29 paper/paper.tex
fix pingscan plusible rules

@@
     Black Nurse & - & - & - & - \\    
-    Ping Scan & 6 & 5 & 2\Mar{18?} & 1 \\
+    Ping Scan & 6 & 5 & 9 & 1 \\
     SYN Flood & 4 & 2 & 2 & 2 \\

### 2020-02-29 simulated-annealing/sa.py
updated contents dict

@@ -13,7 +13,7 @@ from functools import partial, reduce
 from itertools import combinations
 
 #open("bad.rules", 'w').close()
-attacks_list = ["adaptor", "coldfusion", "htaccess", "idq", "issadmin", "system", "script", "pingscan", "synflood"]
+attacks_list = ["adaptor", "coldfusion", "htaccess", "idq", "issadmin", "system", "script", "synflood", "pingscan"]
 parser = argparse.ArgumentParser(description="Description.")
 parser.add_argument('attack', metavar='A')
 args = parser.parse_args()
@@ -43,7 +43,7 @@ contents_dict["idq"] = {'GET':[], '/scripts':[], '/samples':[], '/search':[], '/
 
 contents_dict["system"] = {'GET':[], '/c':[], '/winnt':[], '/system32/':["http_uri", "nocase"], 'cmd.exe?':[], '/c+dir+':[], '/OG':[], 'HTTP/1.1':[], '192.168.1.108':[], 'User-Agent:':[], 'Mozilla':[], '/5.00':[], '(Nikto':[], '/2.1.5)':[],  '(Evasions:':[], 'None)':[], '(Test:':[], '000121)':[], 'Connection:':[], 'Keep-Alive':[]}
 
[…]

### 2020-02-29 paper/paper.tex
iisadmin results on table 2

@@
     ColdFusion & 7 & 19 & 34 & 1\\
-    iisadmin access & 4 & 17 & \Fix{861} & \Fix{1} \\        
+    iisadmin access & 3 & 17 & 408 & 1 \\        
     JBoss JMX & 5 & 23 & 228 & 5 \\

### 2020-02-29 paper/paper.tex
pass on section 4.1

@@
 The first step in the \tname's workflow is Reverse Engineering. The
-goal of this step is to extract options from the malicious traffic. A
-rule that contains these options, by construction, captures the
-malicious traffic. \tname\ uses two strategies to extract options:
-1)~field (for any protocol) and 2)~payload (for http only). We
-elaborate these strategies in the following.
+goal of this step is to extract options from the malicious
+traffic. Consequently, a rule that contains these options captures the
+malicious traffic. \tname\ uses different strategies to extract
+options based on the protocol detected in the input traffic: 1)~the
+field strategy applies to any protocol and 2)~the payload strategy
+applies only to http. We elaborate these strategies in the following.
 
 For the field strategy, \tname{} parses the input messages looking for
-options associated with the data in the fields of a network
-message. Section~\ref{sec:rules-and-packets} shows the correspondence
-between packet fields and some of the rule options from
-\suri. Section~\ref{sec:active-recon} shows a concrete example where
-the values for options are directly extracted from packet fields. We
-obtained the mapping of network fields to options manually, from the
+options associated with the network message field
+data. Section~\ref{sec:rules-and-packets} shows the correspondence
+between packet fields and some of the rule options from \suri and
+Section~\ref{sec:active-recon} shows a concrete example where the
+values for options are directly extracted from the fields of
+messages. We obtained the option-field mapping manunually, from the
 \suri\ documentation.
@@
 frequently in http attacks, which are prevalent. For example,
-\Fix{\Luc{53\%}} of the \numrulessuri{} rules from a public data
+\percHttp\ of the \numrulessuri{} rules from a public data
 set~\cite{emerging-threats-open} we analyzed prescribe defenses
-against http attacks. Figure~\ref{fig:rule-jboss} shows an example of
-a rule that captures traffic containing the string
-\CodeIn{"/HtmlAdaptor"} in the payload.  Content options are very
-common in http rules---\percRulesWithContent{} of the rules from the
-data set mentioned above contain at least one content
-option. Figure~\ref{fig:distribution-contents} shows the histogram of
-the number of these options per rule for the data set mentioned
+against http attacks and the majority of these rules use content
+options. Figure~\ref{fig:rule-jboss} shows an example of a rule that
+captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the
+payload.  Content options are very common in http
+rules---\percRulesWithContent{} of the rules from the data set
+mentioned above contain at least one content
+option. Figure~\ref{fig:distribution-contents} shows the distribution
+of number of content options per rule for the data set mentioned
 above. We limited the number of content options per rule to 10 in the
@@
   are an exception and half of the rules contain at least two
-  \CodeIn{content} options.}  Consequently, handling content options
-for http is important. The content strategy focuses on extracting such
-kinds of options from the payload.
-
-We found, inspecting existing rules, that \CodeIn{content} options
-typically refer to sequence of characters in the payload. To find the
-values associated with these options, \tname{} splits the payload of
-the malicious message in tokens using natural language delimiters,
-such as $\backslash$t, $\backslash$n, $\backslash$r, and
-spaces. Considering the example from
-Section~\ref{sec:content-example}, \tname{} produced a total of 23
-tokens, considered content option candidates.
+  \CodeIn{content} options.}  To sum up, handling content options for
+http is very important.
+
+The content strategy focuses on extracting content options from the
+message payload. Inspecting existing rules, we found that these
+options refer, in the payload, to sequences of characters separated by
+typical natural language delimiters such as $\backslash$t,
+$\backslash$n, $\backslash$r, and spaces. To find the values
+associated with these options, \tname{} splits the payload of the
+malicious message in tokens using these delimiters.  Considering the
+example from Section~\ref{sec:content-example}, \tname{} produced a
+total of 23 tokens, considered content option candidates.
 

### 2020-02-29 paper/paper.tex
quick pass on Section 4.2

@@
 
-\Mar{------------------------------------------- Eu estou aqui}
-
-The rule reported in the first step satisfies the criterion to only
-capture positive traffic. Unfortunately, that rule is overconstrained.
-The rule can include options that are irrelevant for the input
-message. Consequently, the \nids\ may be unable to capture a different
-manifestation of the same attack.
+The seed rule produced in the first step satisfies the criterion to
+only capture positive traffic. However, that rule is overconstrained.
+It could include options that are irrelevant to capturing the attack
+and could lead to false negatives---the \nids\ may be unable to
+capture manifestations of the same attack whose messages do not
+exhibit the data satisfying irrelevant options.
 
@@
 capturing the positive malicious traffic. We say that a rule is
-minimal if discarding any option results in capturing negative (\ie{},
-benign) traffic.  The minimization procedure we propose in this step
-builds on the observation that all options from the rule produced in
-the previous step are satisifed by the positive traffic. Consequently,
-any subset of these options also captures the positive traffic. The
-procedure iteratively discards options from the candidate rule until
-it detects that the rule is minimal, \ie{}, it finds that no other
-option can be removed. The procedure uses negative traffic to guide
-the search.
+minimal if discarding any option results in the capture of negative
+(\ie{}, benign) traffic.  The minimization procedure we propose in
+this step builds on the observation that all options from the rule
+produced in the previous step are satisifed by the positive
+traffic. Consequently, any subset of these options also captures the
+positive traffic. The minimization procedure that we describe in the
+following iteratively discards options from visited rules---derived
+from the seed rule---using the negative traffic to guide the search
+for minimal rules.
 
@@
 \end{algorithm}
+
+\Mar{--------------------------------------- estou aqui}
 

### 2020-02-29 paper/paper.tex
pass on section 4.2

@@
 \algnewcommand\And{\textbf{and} }%
+\setlength{\textfloatsep}{3pt}% Remove \textfloatsep
 \begin{algorithm}[t!]
@@
            \For {$i \gets 1$ to $N$}\Comment{mutate each rule $N$ times.}\label{start-for}
-               \State $r \gets copy(ro)$      
-               \State $j\gets \mathit{rand}(0,\mathit{len}(r))$
-               \State $r[j]\gets 0$\Comment{discard option $j$ in $r$}
-               \If{$\mathit{isNew(r)} \And \mathit{isPlausible}(r)$} 
+               \State $r \gets copy(ro)$\label{copy}
+               \State $j\gets \mathit{rand}(0,\mathit{len}(r))$\label{choose-index}
+               \State $r[j]\gets 0$\Comment{discard option $j$ in $r$}\label{delete-option}
+               \If{$\mathit{isNew(r)} \And\mathit{isPlausible}(r)$} 
                    \State $\mathcal R' \gets \mathcal R' \cup \{r\}$\label{line-progress}
@@
 
-\Mar{--------------------------------------- estou aqui}
-
-Algorithm~\ref{algo1} shows the pseudocode of the \tname{}
-minimization algorithm. The algorithm takes as input the
-overconstrained rule produced by the reverse engineering step and
-produces a set $\mathcal R$ of minimized rules. Line~\ref{line-init}
-initializes the set $\mathcal{R}$ of candidate solutions with the seed
+Algorithm~\ref{algo1} shows the pseudocode for \tname{}'s rule
+minimization. The algorithm takes as input an overconstrained rule
+produced by the reverse engineering step and produces a set $\mathcal
+O$ of minimized rules. Line~\ref{line-init} initializes the current
+population of individuals (\ie{}, rules) $\mathcal{R}$ with the seed
 rule $sr$. We encode a rule as a bitvector where each index in the
 vector represents one option; the value assigned at a given position
-indicates whether or not the option is present in the rule. Each
+indicates whether or not the option is present in the rule. One
 iteration of the outer loop
-(lines~\ref{start-do-while}-\ref{end-do-while}) represents one
-generation of the population of individuals produced by the
-search. The algorithm terminates when no rule in the current
+(lines~\ref{start-do-while}-\ref{end-do-while}) processes the current
+population of individuals $\mathcal R$ and creates the next generation
+$\mathcal R'$. Line~\ref{init-rprime} initializes the population of
+individuals for the next generation. The inner loop
+(lines~\ref{start-foreach}--\ref{end-foreach}) iterates through each
+candidate rule in $\mathcal R$. The algorithm mutates each of these
+rules for $N$ times, discarding one option of the rule each
+time. Line~\ref{copy} assigns a copy of the bitvector encoding a rule
+in $\mathcal R$ to variable $r$. Function $\mathit{len}$ returns the
+length of a vector and function $\mathit{rand}$ randomly chooses one
+integer value in the range defined by its parameters. So, the
+expression $\mathit{rand(0, len(r))}$, appearing at
+line~\ref{choose-index}, returns an index of the option in $r$ to
+delete at line~\ref{delete-option}. The call to $\mathit{isNew}$
+checks if $r$ has not been previously generated through another path
+whereas the call to $\mathit{isPlausible}$ checks if the rule only
+captures positive traffic. If $r$ satisfies both criteria, it is added
+to $\mathcal R'$, which denotes the next generation of individuals in
+the search.
+
+The algorithm terminates when no rule in the current
 generation of individuals could be further reduced without violating
@@
 traffic. This means that execution of one iteration has not reached
-the statement declared at line~\ref{line-progress}, which adds an
-element to the set $\mathcal{R'}$. Consequently, that set will be
-empty at the end of one iteration of the outer
-loop. Line~\ref{init-rprime} initializes the new generation of
-candidate rules. The inner loop
-(lines~\ref{start-foreach}--\ref{end-foreach}) iterates through each
-candidate rule in $\mathcal R$. Note that these rules have been found
-in the previous generation. The algorithm mutates each of these rules
-for $N$ times, discarding one option of the rule each time. The call
-to $\mathit{isNew}$ checks if $r$ has not been previously generated
-through another path whereas the call to $\mathit{isPlausible}$ checks
-if the rule only captures positive traffic. Function $\mathit{copy}$
-creates a copy of the bitvector of a rule, function $\mathit{len}$
-returns the length of the vector, and function $\mathit{rand}$
-randomly chooses one index of the vector.
+the statement declared at line~\ref{line-progress}, which adds one
+element to the set $\mathcal{R'}$. Consequently, as $\mathcal{R'}$ is
+initialized empty that set will remain empty at the end of an
+iteration of the outer loop.
+
+\Mar{--------------------------------------- estou aqui}
 

### 2020-02-29 paper/paper.tex
pass on section 4.2

@@
 
+\Fix{consider changing the name of this step}
 The seed rule produced in the first step satisfies the criterion to
@@
 
-\Mar{--------------------------------------- estou aqui}
-
-Note that no rules are produced in the last iteration of the
-algorithm. Minimal rules are produced in the second-to-last
-iteration. In the other iterations, intermediate rules are
+Note that the algorithm produces no rules in the last iteration. Rules
+with a single option are produced in the second-to-last iteration. In
+other iterations, ules with more than one option are
 produced. Conceptually, the search induces a search tree with the seed
-rule $sr$ as the root, minimal rules as child nodes, and intermediate
-rules as internal nodes. The edges of the tree denote the ancestor
-relation between two rules. Note also that the algorithm reports every
-rule in the tree instead of only minimal rules. The reason is that the
-correct rule may be non-minimal. In summary, this algorithm reports
-many rules that are plausible by using benign traffic to minimize an
-overconstrained rule that is focused on the malicious traffic. In the
-following step, \tname{} ranks the rules this algorithm reports by
-their similarity with previous rules.
-
-\begin{theorem}
-  The minimization algorithm computes rules that capture only
-  the positive traffic.
-\end{theorem}
+rule $sr$ as the root, single-option rules as child nodes, and rules
+with more options as internal nodes. The edges of the tree denote the
+ancestor relation between two rules.
+
+It is also worth noting that the algorithm reports every rule in the
+tree instead of only those that are surely minimal, \ie{} rules with
+one option. The reason is that the the traffic used to check
+plausibility is limited by definition---reporting only minimal rules
+with respect to that traffic could discard the correct rule. In
+summary, this algorithm reports many rules that are plausible by using
+the benign traffic to reduce the overconstrained rule produced at the
+\reveng\ step. In the following step, \tname{} ranks the rules this
+algorithm reports by their similarity to existing trusted
+rules. Intuitively, rules that do not look alike to rules written by
+humans and previously used rank low.
+
+\begin{proposition}
+  The minimization algorithm reports rules that capture only the
+  positive traffic.
+\end{proposition}
 \begin{proof}
-  Every rule created during the search captures the positive traffic
-  as every rule contains a proper subset of the option from the
-  original rule and all those options are satisfied by the positive
-  traffic. By definition of function $\mathit{isPlausible}$, these rules
-  also capture \emph{only} the positive traffic. If some negative
-  traffic was captured, a rule would not be added to $\mathcal
-  R'$.
+  Every rule created during the search captures the positive malicious
+  traffic as every rule contains a proper subset of the options from
+  the original rule and all those options are satisfied by the
+  malicious traffic. By definition of function $\mathit{isPlausible}$,
+  these rules also capture \emph{only} the positive traffic. If
+  negative traffic is captured, a rule is not added to $\mathcal R'$;
+  consequently, it cannot be an element of $\mathcal O$ and reported
+  on the output.
 \end{proof}
 
-\begin{proposition}
-  The minimization algorithm may not produce minimal rules.
-\end{proposition}
-
-\begin{proof}
-The output of the minimization algorithm may contain rules that are
-not minimal given the non-determinism of the algorithm. In principle,
-it is possible that options that could be discarded are not selected
-in the loop from lines~\ref{start-for}--\ref{end-for}. However, we
-found empirically that a small value of $N$ suffices to discard those
-options from a rule.
-\end{proof}
+%% \begin{proposition}
+%%   The minimization algorithm may not produce minimal rules.
+%% \end{proposition}
+
+%% \begin{proof}
+%% The output of the minimization algorithm may contain rules that are
+%% not minimal given the non-determinism of the algorithm. In principle,
+%% it is possible that options that could be discarded are not selected
+%% in the loop from lines~\ref{start-for}--\ref{end-for}. However, we
+%% found empirically that a small value of $N$ suffices to discard those
+%% options from a rule.
+%% \end{proof}
 
@@
 \emph{isPlausible} by 1)~updating Suricata to only monitor the rule
-that is passed as parameter, 2)~replaying the negative traffic (see
+that is passed as parameter\Mar{isto realmente eh necessario? se sim,
+  explique. me pareceu que o que vc. precisa ate aqui jah estaria no
+  log do \suri}, 2)~replaying the negative benign traffic (see
 Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
@@
 of sending traffic to a machine with a deployed server (\eg{}, web
-server) that \suri\ monitors, our setup calls Suricata directly
-through an API. Suricata is invoked multiple times in the process of
+server) that \suri\ would monitor, our setup calls \suri\ directly
+through an API. \suri\ is invoked multiple times in the process of
 minimizing the rule.  \textbf{Limitations.} It is worth noting that
@@
 \tname{} only requires one sample, but can benefit of several
-samples. The same applies for the negative traffic. In this case,
-however, \tname{} can use public data sets available online.
+samples\Mar{como? nao estah claro. por favor expliquem.}. The same
+applies to the negative traffic. However, in this case, \tname{} uses
+different public data sets available online (see
+Section~\ref{sec:objectanalysis-benign}).
 
@@
 \label{sec:ranking}
+
+\Mar{----------------------------- estou aqui}
 
@@
 \vspace{1ex}
-\subsubsection{Benign traffic.}~Public data sets of bening traffic
+\subsubsection{\label{sec:objectanalysis-benign}Benign traffic.}~Public data sets of bening traffic
 exist and can be used to evaluate the rate of false positives of

### 2020-02-29 paper/paper.tex
name of subsections

@@
 
-\subsubsection{Content rarity.} 
+\subsubsection{Popularity of content strings.} 
 This heuristic weighs higher rules that use popular strings in content
@@
 
-\subsubsection{Frequency of content modifiers.} As previously
+\subsubsection{Popularity of content modifiers.} As previously
 mentioned, \percRulesWithContent\ (=26K/\numrulessuri) of Suricata

### 2020-02-29 paper/paper.tex
heuristic 5 histogram and comment

@@
 \end{subfigure}%
+%\vspace{1ex}
+\\
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \scalebox{1.2}{  
+    \begin{tikzpicture}
+      \begin{axis}[
+          bar width=6pt,
+          scaled ticks=false,
+          tick label style={/pgf/number format/fixed},
+          ybar stacked,
+          enlargelimits=0.15,
+          legend style={at={(0.5,-0.15)},
+            anchor=north,legend columns=-1},
+          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode, flags, tls.fingerprint},
+          xtick=data,
+          xticklabel style={rotate=45},          
+          nodes near coords,
+          every node near coord/.append style={font=\tiny},
+          nodes near coords align={vertical},
+        ]
+        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155) (icode, 108) (flags, 97) (tls.fingerprint, 61)};
+      \end{axis}
+    \end{tikzpicture}
+  }
+  \vspace{-3ex}  
+  \caption{\label{fig:distribution-strings}Option Frequency.}
+\end{subfigure}%
 \vspace{-1ex}
@@
 would be 0.35 (=(0.2+0.5)/2).
+\Gui{A quinta heuristica que utilizamos considera a popularidade das opcoes.
+Primeiro pegamos a quantidade de vezes que uma opcao eh utilizada (sem considerar repeticoes numa unica regra).
+Dividimos o valor encontrado para cada opcao pelo maior valor encontrado, encontrando sempre valores de no maximo 1.
+Entao na regra nos somamos os valores de cada opcao, dividido pelo numero de opcoes, chegando a um numero de no maximo 1, que representa a popularidade das opcoes na regra.
+\[\frac{\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}}{N}\]}
 

### 2020-02-29 simulated-annealing/sa.py
fix keyword freq, added teardrop

@@ -52,9 +52,9 @@ else:
 
 keyword_list=("app-layer-protocol", "uricontent", "ack", "seq", "window", "ipopts", "flags", "fragbits", "fragoffset", "ttl", "tos", "itype", "icode", "icmp_id", "icmp_seq", "dsize", "flow", "threshold", "tag", "content", "pcre", "replace", "rawbytes", "byte_test", "byte_jump", "sameip", "geoip", "ip_proto", "ftpbounce", "id", "rpc", "flowvar", "flowint", "pktvar", "flowbits", "hostbits", "ipv4-csum", "tcpv4-csum", "tcpv6-csum", "udpv4-csum", "udpv6-csum", "icmpv4-csum", "icmpv6-csum", "stream_size", "detection_filter", "decode-event", "nfq_set_mark", "bsize", "tls.version", "tls.subject", "tls.issuerdn", "tls_cert_notbefore", "tls_cert_notafter", "tls_cert_expired", "tls_cert_valid", "tls.fingerprint", "tls_store", "http_protocol", "http_start", "urilen", "http_header_names", "http_accept", "http_accept_lang", "http_accept_enc", "http_connection", "http_content_len", "http_content_type", "http_referer", "http_request_line", "http_response_line", "nfs_procedure", "nfs_version", "ssh_proto", "ssh.protoversion", "ssh_software", "ssh.softwareversion", "ssl_version", "ssl_state", "byte_extract", "file_data", "pkt_data", "app-layer-event", "dce_iface", "dce_opnum", "dce_stub_data", "smb_named_pipe", "smb_share", "asn1", "engine-event", "stream-event", "filename", "fileext", "filestore", "filemagic", "filemd5", "filesha1", "filesha256", "filesize", "l3_proto", "lua", "iprep", "dns_query", "tls_sni", "tls_cert_issuer", "tls_cert_subject", "tls_cert_serial", "tls_cert_fingerprint", "ja3_hash", "ja3_string", "modbus", "cip_service", "enip_command", "dnp3_data", "dnp3_func", "dnp3_ind", "dnp3_obj", "xbits", "base64_decode", "base64_data", "krb5_err_code", "krb5_msg_type", "krb5_cname", "krb5_sname", "template2", "ftpdata_command", "bypass", "prefilter", "compress_whitespace", "strip_whitespace", "to_sha256", "depth", "distance", "within", "offset", "nocase", "fast_pattern", "startswith", "endswith", "distance", "noalert", "http_cookie", "ht
[…]

### 2020-02-29 paper/paper.tex
included udp teardrop on table 3

@@
     SYN Flood & 4 & 2 & 2 & 2 \\
-    UDP Teardrop & - & - & - & - \\
+    UDP Teardrop & 1 & 4 & 8 & - \\
     UDP Flood & - & - & - & - \\

### 2020-03-01 paper/paper.tex
added webid plausible rules

@@
     Script tag in URI & 23 & 22 & 786 & 1 \\
-    WeBid & 1 & 23 & - & -\\    
+    WeBid & 1 & 23 & 11100 & -\\    
     Wordpress Login & 8 & 59 & - & - \\

### 2020-03-01 paper/paper.tex
add rank teardrop and webid

--- 2020-02-29 paper/paper.tex
@@
     SYN Flood & 4 & 2 & 2 & 2 \\
-    UDP Teardrop & 1 & 4 & 8 & - \\
+    UDP Teardrop & 1 & 4 & 8 & 3 \\
     UDP Flood & - & - & - & - \\
@@
     Script tag in URI & 23 & 22 & 786 & 1 \\
-    WeBid & 1 & 23 & - & -\\    
+    WeBid & 1 & 23 & 11100 & 6\\    
     Wordpress Login & 8 & 59 & - & - \\

### 2020-03-01 paper/paper.tex
pass on table 2 and 3

@@
     /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
+    Tomcat server snoop access & \cite{nikto} & AREC & HTTP & \textbf{content}:"/jsp/snp/"; http\_uri; \textbf{content}:".snp"; http\_uri;\\  
     \bottomrule
@@
     /system32/ in URI & 4 & 20 & 1,913 & 1 \\
+    Tomcat server snoop access & 1 & - & - & - \\
     \bottomrule

### 2020-03-01 paper/paper.tex
pass table 3

--- 2020-03-01 paper/paper.tex
@@
     Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
-    WeBid \MyComment{Local File Inclusion} & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
+    WeBid LFI \MyComment{Local File Inclusion} & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
     WordPress Login & \cite{wrccdc} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
@@
     \midrule
-    Black Nurse & - & - & - & - \\    
+    Black Nurse & 1 & 4 & 9 & - \\    
     Ping Scan & 6 & 5 & 9 & 1 \\
@@
     \midrule
-    Apache Struts 2 & - & - & - & - \\
+    Apache Struts 2 & 2 & 65 & - & - \\
     ColdFusion & 7 & 19 & 34 & 1\\
@@
     Script tag in URI & 23 & 22 & 786 & 1 \\
-    WeBid & 1 & 23 & 11100 & 6\\    
+    WeBid LFI & 1 & 23 & 11100 & 6\\    
     Wordpress Login & 8 & 59 & - & - \\
@@
     /system32/ in URI & 4 & 20 & 1,913 & 1 \\
+    Tomcat Snoop Access & 1 & 24 & 29830 & - \\
     \bottomrule

### 2020-03-01 simulated-annealing/sa.py
new attacks

@@ -13,7 +13,7 @@ from functools import partial, reduce
 from itertools import combinations
 
 #open("bad.rules", 'w').close()
-attacks_list = ["adaptor", "coldfusion", "htaccess", "idq", "issadmin", "system", "script", "synflood", "pingscan"]
+attacks_list = ["adaptor", "coldfusion", "htaccess", "idq", "issadmin", "system", "script", "synflood", "pingscan", "cron", "teardrop"]
 parser = argparse.ArgumentParser(description="Description.")
 parser.add_argument('attack', metavar='A')
 args = parser.parse_args()
@@ -25,11 +25,11 @@ fitnessFile_path = "./suricata-logs/" + str(args.attack) + ".log"
 time_begin = time.time()
 
 contents_dict = {}
-contents_dict["cron"] = {'GET':[], '/cron.php?':["http_uri", "nocase"], 'include_path=':["http_uri", "nocase"], 'http:':[], '/cirt.net':[], '/rfiinc':[], '../':[], '.txt??':[], 'HTTP':[], '/1.1':[], 'Connection:':[], 'Keep-Alive':[], 'User-Agent':[], 'Mozilla':[], '5.00':[], '(Nikto':[], '/2.1.5)':[], '(Evasions:':[], 'None)':[], '(Test':[], '004603)':[], 'Host:':[], '192.168.1.108': []}
+contents_dict["cron"] = {'GET':[], '/cron.php?':["http_uri", "nocase"], 'include_path=':["http_uri", "nocase"], 'http:':[], '/cirt.net':[], '/rfiinc':[], '.txt??':[], 'HTTP':[], '/1.1':[], 'Connection:':[], 'Keep-Alive':[], 'User-Agent':[], 'Mozilla':[], '5.00':[], '(Nikto':[], '/2.1.5)':[], '(Evasions:':[], 'None)':[], '(Test':[], '004603)':[], 'Host:':[], '192.168.1.108': []}
 
 contents_dict["htaccess"] = {'GET':[], '/Ud3uMSnb':[], '.htaccess':["http_uri", "nocase"], 'HTTP':[], '/1.1':[], 'User-Agent:':[], 'Mozilla':[], '/5.00':[], '(Nikto':[], '/2.1.5)':[], '(Evasions:':[], 'None)':[], '(Test:':[], 'map_codes)':[], 'Connection:':[], 'Keep-Alive':[], 'Host:':[]}
 
[…]

### 2020-03-01 simulated-annealing/sa.py
memory optimization - checkPrecision and checkFalseNegative

@@ -636,11 +636,12 @@ def checkFalseNegative(rules):
         output.append(0)
 
     with open(fitnessFile_path, "r") as fitnessFile:
-        fitnessFile = fitnessFile.read()
-    
-    for i in range(len(rules)):
-        if fitnessFile.count('[1:'+str(rules[i].sid)+':') >= 1:
-            output[i] = fitnessFile.count('[1:'+str(rules[i].sid)+':')
+        line = fitnessFile.readline()
+
+        while line:
+            if line.count('[1:'+str(rules[i].sid)+':') >= 1:
+                output[i] = line.count('[1:'+str(rules[i].sid)+':')
+            line = fitnessFile.readline()
             #if str(rules[i].sid) == "1099019":
         #else:
             #print("fitness file count", ':'+str(rules[i].sid)+':', fitnessFile.count(str(rules[i].sid)))
@@ -669,11 +670,12 @@ def checkPrecision(rules):
         output.append(0)
 
     with open(fitnessFile_path, "r") as fitnessFile:
-        fitnessFile = fitnessFile.read()
-    
-    for i in range(len(rules)):
-        if fitnessFile.count('[1:'+str(rules[i].sid)+':') >= 1:
-            output[i] = fitnessFile.count('[1:'+str(rules[i].sid)+':')
+        line = fitnessFile.readline()
+
+        while line:   
+            if line.count('[1:'+str(rules[i].sid)+':') >= 1:
+                output[i] = fitnessFile.count('[1:'+str(rules[i].sid)+':')
+            line = fitnessFile.readline()
         #else:
             #print("fitness file count", ':'+str(rules[i].sid)+':', fitnessFile.count(str(rules[i].sid)))
             #print("False negative rule:", rules[i])

### 2020-03-01 simulated-annealing/sa.py
fix - remove exit()

@@ -1412,7 +1412,7 @@ else:
         final_rule.options = teardrop_options
         
     print("initial rule:\n", final_rule)
-    exit()
+    #exit()
     #print("fit1: ", ruleSizeFitness(final_rule), "fit2: ", ruleContentsFitness(final_rule), "fit3: ", rareContentsFitness(final_rule), "fit4: ", #ruleContentsModifiersFitness(final_rule))
     #quit()
     final_rule = optimizeRule(final_rule)

### 2020-03-01 simulated-annealing/sa.py
memory optimization - evalContents

@@ -697,15 +697,18 @@ def evalContents(rules):
         output.append(0)
 
     with open(fitnessFile_path, "r") as fitnessFile:
-        fitnessFile = fitnessFile.read()
-    
-    if len(fitnessFile) == 0 :
-        return output
+        line = fitnessFile.readline()
 
-    for i in range(len(rules)):
-        s="Testing rule {} ".format(i)
-        if s in fitnessFile:
-            output[i]=1
+        if not line :
+            return output
+
+        while line:
+            for i in range(len(rules)):
+                s="Testing rule {} ".format(i)
+                if s in line:
+                    output[i]=1
+
+            line = fitnessFile.readline()
 
     #print(output)
     return output

### 2020-03-01 paper/paper.tex
pass on technique

@@
 
-\Mar{----------------------------- estou aqui}
-
-\tname{} uses a composition of heurisitc functions to sort the rules
-it produces. Each function uses a different criterion to estimate how
-similar a rule, created in the previous step, is to rules from a
-public database~\cite{emerging-threats-open} including a total of
-\numrulessuri\ rules. We describe in the following each of these
-heuristic functions. The codomain of each function ranges in the 0-1
-interval. The score of a rule is the average across the values
-produced by each function.
+\tname{} uses a composition of heurisitic functions to sort the rules
+produced in the previous step. Each of these functions uses a
+different criterion to estimate how similar a created rule is to those
+from public rulesets~\cite{emerging-threats-open}. The
+intuition is that rules that look more similar to existing trusted
+rules rank higher and could catch the attention of a system
+administrator. The score of a rule is the weighted average across the
+values produced by each heuristic function. The codomain of each
+function ranges on the 0-1 interval. Hence, the score of a rule also
+ranges on the 0-1 interval. We describe in the following each of these
+heuristic functions.
 
@@
 \subsubsection{Size of a rule}
+\label{sec:heuristic-size-of-rule}
 One element that \tname{} uses to measure similarity between a
@@
 rule, \ie{}, the number of options it contains. Informally, rules with
-popular sizes score high whereas inpopular rules score
+popular sizes score high whereas unpopular rules score
 low. Figure~\ref{fig:distribution-rule-size} shows the histogram of
@@
 options. \tname{} values rules in that size range higher compared to
-other sizes.
-
-%\tname{} uses this information to rank rules.
-
-More precisely, we calculate the contribution of size of a given rule
-$r$ with the following formula:
-
-\[\frac{\mathit{rules\_per\_size}[\mathit{len(r)}]}{\mathit{max(rules\_per\_size})}\]
-
-\noindent
-, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
-$\mathit{rules\_per\_size}$ encodes distribution of sizes as a
+other sizes. We calculate the contribution of size of a given rule
+$r$ with the following formula $\mathit{rules\_per\_size}[\mathit{len(r)}]/\mathit{max(rules\_per\_size)}$, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
+$\mathit{rules\_per\_size}$ encodes the distribution of sizes as a
 dictionary mapping sizes to number of associated rules, and
 $\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
-for any dictionary entry. Consequently, the contribution of size is
-proportional to its popularity. For example, the contribution of this
-heuristic for a rule with three options would be 0.684 (=3,662/5,353)
-whereas the contribution of a rule with nine options would be 0.028
-(=154/5,353). Note that the measurement is normalized in the 0-1
-range.
-
-\subsubsection{Number of content options.} We empirically observed that content options are very
-common. \Luc{Se for preciso temos numeros para fundamentar isso
-  melhor, sao cerca de 70K contents nas 28K regras do
-  data set}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
-  (\%) das 28K regras tem contents.} \Luc{92\%} As consequence, we defined a
-heuristic function that takes the number of content options into
-account. We calculate the contribution of number of content options
-with the following formula:
+for any dictionary entry as to normalize the measurement in the 0-1
+range. Consequently, the contribution of size is proportional to its
+popularity. The contribution of this heuristic would be 0.684
+(=3,662/5,353) to a rule with three options and 0.028 (=154/5,353) to
+a rule with nine options.
+
+%% \vspace{-2ex}
+%% \[\]
+%% \vspace{-2ex}
+%\noindent
+
+%% \Luc{Se for preciso temos numeros para fundamentar isso
+%%   melhor, sao cerca de 70K contents nas 28K regras do
+%%   data set}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
+%%   (\%) das 28K regras tem contents.} \Luc{92\%}
 
@@
   \centering
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \scalebox{1.2}{
+      \begin{tikzpicture}
+        \begin{axis}[
+            bar width=6pt,
+            scaled ticks=false,
+            tick label style={/pgf/number format/fixed},
+            ybar stacked,
+            enlargelimits=0.15,
+            legend style={at={(0.5,-0.15)},
+              anchor=north,legend columns=-1},
+            symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+            xtick=data,
+            nodes near coords,
+            every node near coord/.append style={font=\tiny},
+            nodes near coords align={vertical},
+          ]
+          \addplot coordinates {(0,11) (1,1301) (2,4910) (3,3662) (4,3250) (5,5353) (6,2911) (7, 1337) (8, 1246) (9,763) (10, 495)};
+        \end{axis}
+      \end{tikzpicture}
+    }
+    \vspace{-1ex}
+    \caption{\label{fig:distribution-rule-size}Rule sizes.}
+  \end{subfigure}%  
   \begin{subfigure}{.5\textwidth}
@@
   \end{subfigure}% <---- don't forget this %
-%\vspace{1ex}
-%\\
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \scalebox{1.2}{
-    \begin{tikzpicture}
-      \begin{axis}[
-          bar width=6pt,
-          scaled ticks=false,
-          tick label style={/pgf/number format/fixed},
-          ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
-          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
-          xtick=data,
-          nodes near coords,
-          every node near coord/.append style={font=\tiny},
-          nodes near coords align={vertical},
-        ]
-        \addplot coordinates {(0,11) (1,1301) (2,4910) (3,3662) (4,3250) (5,5353) (6,2911) (7, 1337) (8, 1246) (9,763) (10, 495)};
-      \end{axis}
-    \end{tikzpicture}
-  }
-  \vspace{-1ex}
-  \caption{\label{fig:distribution-rule-size}Rule sizes.}
-\end{subfigure}%
-%\vspace{1ex}
   \\
@@
 
-\[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
-
-\noindent
-Note that the only difference of this formula compared to the formula
-to compute contribution of size is the distribution used, which is
-given by the term $\mathit{rules\_per\_number\_of\_contents}$ denoting
-a dictionary of number of content options to number of rules that
-contain that number of content options.
-
+\subsubsection{Number of content options.} We observed
+that \percContentOptions\ of the rules from the ruleset we are
+analyzing contain content options, but the distribution of those
+contents per rule is not uniform. Figure~\ref{} shows the
+corresponding distribution. Given the importance of this kind of
+option to the definition of rules, we defined a heuristic function
+focused on the typical number of content options in a rule. We
+calculate the contribution of this function as described on previous
+Section~\ref{sec:heuristic-size-of-rule}.
+
+%% to the score of a rule
+%% with the following formula:
+
+%% \vspace{-2ex}
+%% \[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
+%% \vspace{-2ex}
+
+%% \noindent
+%% The only difference of this formula compared to the one defined on
+%% Section~\ref{sec:heuristic-size-of-rule} is the distribution used. The
+%% distribution is given by the term
+%% $\mathit{rules\_per\_number\_of\_contents}$ denoting a dictionary of
+%% number of content options to number of rules that contain that number
+%% of content options.
+
+\Mar{------------------------------------------- estou aqui}
 \subsubsection{Popularity of content strings.} 

### 2020-03-01 paper/paper.tex
organizing plots + 4.3.3

@@
 \begin{proposition}
+  \vspace{-1ex}  
   The minimization algorithm reports rules that capture only the
@@
 \begin{proof}
+  \vspace{-1.5ex}
   Every rule created during the search captures the positive malicious
@@
   on the output.
+  \vspace{-1ex}    
 \end{proof}
@@
 
-\begin{figure*}[t!]
+\begin{figure}[t!]
   \centering
-  \begin{subfigure}{.5\textwidth}
+  \begin{subfigure}{.25\textwidth}
     \centering
-    \scalebox{1.2}{
+    \scalebox{0.8}{
       \begin{tikzpicture}
         \begin{axis}[
-            bar width=6pt,
-            scaled ticks=false,
-            tick label style={/pgf/number format/fixed},
+          bar width=4pt,
+          scaled ticks=true,
+          ymajorticks=false,
           ybar stacked,
-            enlargelimits=0.15,
-            legend style={at={(0.5,-0.15)},
-              anchor=north,legend columns=-1},
-            symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          enlargelimits=0.08,
+%          enlargelimits=false,
+%          axis equal,
+          clip=false,
+          legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
           xtick=data,
+%          xticklabel style={rotate=45},          
           nodes near coords,
@@
           nodes near coords align={vertical},
+          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
           ]
@@
   \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
+  \begin{subfigure}{.25\textwidth}
     \centering
-    \scalebox{1.2}{  
+    \scalebox{0.8}{  
       \begin{tikzpicture}
         \begin{axis}[
-            bar width=6pt,
-            scaled ticks=false,
-            tick label style={/pgf/number format/fixed},
+bar width=4pt,
+          scaled ticks=true,
+          ymajorticks=false,
           ybar stacked,
-            enlargelimits=0.15,
-            legend style={at={(0.5,-0.15)},
-              anchor=north,legend columns=-1},
-            symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          enlargelimits=0.08,
+%          enlargelimits=false,
+%          axis equal,
+          clip=false,
+          legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
           xtick=data,
+%          xticklabel style={rotate=45},          
           nodes near coords,
@@
   \\
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.25\textwidth}
   \centering
-  \scalebox{1.2}{  
+  \scalebox{0.8}{  
     \begin{tikzpicture}
       \begin{axis}[
-          bar width=6pt,
-          scaled ticks=false,
-          tick label style={/pgf/number format/fixed},
+          bar width=4pt,
+          scaled ticks=true,
+          ymajorticks=false,
           ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
-          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},
+          enlargelimits=0.08,
+%          enlargelimits=false,
+%          axis equal,
+          clip=false,
+          legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
           xtick=data,
-          xticklabel style={rotate=45},          
+          xticklabel style={rotate=35},          
           nodes near coords,
@@
           nodes near coords align={vertical},
+          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},          
         ]
@@
   }
-  \vspace{-3ex}  
-  \caption{\label{fig:distribution-strings}Strings used as parameters of content options.}
+  \vspace{-2ex}  
+  \caption{\label{fig:distribution-strings}Strings used in content options.}
 \end{subfigure}%
-%\vspace{1ex}
-%\\
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.25\textwidth}
   \centering
-  \scalebox{1.2}{
+  \scalebox{0.8}{
     \begin{tikzpicture}
       \begin{axis}[
+          bar width=4pt,
+          scaled ticks=true,
+          ymajorticks=false,
           ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
-          %          ylabel={\#rules},
-          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection, http\_request\_line},
+          enlargelimits=0.08,
+%          enlargelimits=false,
+%          axis equal,
+          clip=false,
+          legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
           xtick=data,
-          xticklabel style={rotate=45},          
+          xticklabel style={rotate=35},          
           nodes near coords,
@@
           nodes near coords align={vertical},
+          symbolic x coords={method, uri, user\_agent, header, host, connection, request\_line},          
         ]
-        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29) (http\_request\_line,11)};
+        \addplot coordinates {(method,3821) (uri,10111) (user\_agent,1022) (header,4089) (host,224) (connection,29) (request\_line,11)};
       \end{axis}
@@
   }
-  \vspace{-1ex}
-  \caption{\label{fig:distribution-content_modifiers}Content modifiers
-    frequency.}
+  \vspace{-2ex}
+  \caption{\label{fig:distribution-content_modifiers}Http content modifiers.}
 \end{subfigure}%
-%\vspace{1ex}
 \\
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.2\textwidth}
   \centering
-  \scalebox{1.2}{  
+  \scalebox{0.8}{  
     \begin{tikzpicture}
       \begin{axis}[
-          bar width=6pt,
-          scaled ticks=false,
-          tick label style={/pgf/number format/fixed},
+          bar width=4pt,
+          scaled ticks=true,
+          ymajorticks=false,
           ybar stacked,
-          enlargelimits=0.15,
-          legend style={at={(0.5,-0.15)},
-            anchor=north,legend columns=-1},
+          enlargelimits=0.08,
+%          enlargelimits=false,
+%          axis equal,
+          clip=false,
+          legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
+          xtick=data,
+          xticklabel style={rotate=35},          
+          nodes near coords,
+          every node near coord/.append style={font=\scriptsize},
+          nodes near coords align={vertical},
           symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode, flags, tls.fingerprint},          
-          xtick=data,
-          xticklabel style={rotate=45},          
-          nodes near coords,
-          every node near coord/.append style={font=\tiny},
-          nodes near coords align={vertical},
         ]
@@
   }
-  \vspace{-3ex}  
-  \caption{\label{fig:distribution-strings}Option Frequency.}
+  \vspace{-5ex}      
+  \caption{\label{fig:distribution-strings}Options.}
+%  \vspace{-1ex}    
 \end{subfigure}%
@@
 \caption{Distribution of features.}
-\end{figure*}
+\end{figure}
 
@@
 analyzing contain content options, but the distribution of those
-contents per rule is not uniform. Figure~\ref{} shows the
+contents per rule is not uniform. Figure~\ref{fig:distribution-contents} shows the
 corresponding distribution. Given the importance of this kind of
@@
 
-\Mar{------------------------------------------- estou aqui}
 \subsubsection{Popularity of content strings.} 
-This heuristic weighs higher rules that use popular strings in content
-options. For a given rule \tname{} takes the arithmetic mean of the
-popularity scores of each string used in contents in that rule. More
-precisely, we calculate the contribution of content rarity with the
+The heuristic described in this section leverages content rarity to
+discriminate rules---it weighs higher (respectively, lower) rules that
+use popular (respectively, unpopular) strings in content options. The
+score of a rule $r$ on this function is given by the arithmetic mean
+of the popularity scores of each string used in contents in $r$. More
+precisely, the contribution of content rarity is given by the
 following formula:
 
+\vspace{-2.5ex}
 \[\frac{\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}}{N}\]
+\vspace{-2.5ex}
 
 \noindent
-, where the term $\mathit{content_i}$ denotes the $i$-th xcontent in a
-given rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
+, where $\mathit{content_i}$ denotes the $i$-th content in a given
+rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
 denotes the frequency of the string used in the $i$-th content, the
@@
 rule. Figure~\ref{fig:distribution-strings} shows the distribution of
-the ten most popular strings in our rule data set. To illustrate this
+the ten most popular strings in our ruleset. To illustrate this
 function, let us consider a rule with two content options, using the
@@
 rule for the computation of similarity score would be 0.601
-(=(1+.203)/2).
-
+(=(1+.203)/2). The relatively high weight reflects the fact that
+``GET'' and ``ID'' are popular contents.
+
+\Mar{----------------------------------------------------------- estou
+aqui}
 \subsubsection{Popularity of content modifiers.} As previously

### 2020-03-01 simulated-annealing/sa.py
fix - checkPrecision

@@ -674,7 +674,7 @@ def checkPrecision(rules):
 
         while line:   
             if line.count('[1:'+str(rules[i].sid)+':') >= 1:
-                output[i] = fitnessFile.count('[1:'+str(rules[i].sid)+':')
+                output[i] = line.count('[1:'+str(rules[i].sid)+':')
             line = fitnessFile.readline()
         #else:
             #print("fitness file count", ':'+str(rules[i].sid)+':', fitnessFile.count(str(rules[i].sid)))

### 2020-03-01 paper/paper.tex
pass on technique

@@
           nodes near coords align={vertical},
-          symbolic x coords={method, uri, user\_agent, header, host, connection, request\_line},          
+          symbolic x coords={uri, header, method, user\_agent, host, connection, request\_line},          
         ]
-        \addplot coordinates {(method,3821) (uri,10111) (user\_agent,1022) (header,4089) (host,224) (connection,29) (request\_line,11)};
+        \addplot coordinates {(uri,10111) (method,3821) (user\_agent,1022) (header,4089) (host,224) (connection,29) (request\_line,11)};
       \end{axis}
@@
 
-\Mar{----------------------------------------------------------- estou
-aqui}
-\subsubsection{Popularity of content modifiers.} As previously
-mentioned, \percRulesWithContent\ (=26K/\numrulessuri) of Suricata
-rules involve content options. Furthermore, of the 26K rules that use
-content options, 13K use HTTP modifiers\Mar{@Lucas, pode dar o numero
-  desconsiderando HTTP? Esta parte eh generica.}. \Luc{Posso, mas nessa função nós consideramos apenas os modificadores http.} Content modifiers
-serve to indicate what part of a network packet the string mentioned
-in the content option should appear. For example, the modifier
-\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
-  http\_header} states the requirement that the string ``keep-alive''
-be present in the header of the message. (A keep-alive connection
-attribute indicates that a single TCP connection should remain open
-for multiple HTTP requests/responses.)
-Figure~\ref{fig:http-header-example} illustrates possible content
-options used to match against certain strings that appear in an HTTP
-message. A check (resp., cross) mark indicates a match (resp.,
-unmatch).
+\subsubsection{Popularity of http content modifiers.} As previously
+mentioned, \percRulesWithContent\ of \suri\ rules involve content
+options. Furthermore, nearly half of the rules that use content
+options, use content modifiers on http message payloads. Content
+modifiers serve to indicate what part of a network packet the string
+mentioned in the content option should appear. For example, the
+modifier \CodeIn{http\_header} in the option \CodeIn{content:
+  ``keep-alive''; http\_header} indicates that the string
+``keep-alive'' should be present in the header of the
+message\footnote{A keep-alive connection attribute indicates that a
+  single TCP connection should remain open for multiple HTTP
+  requests/responses.}.  Figure~\ref{fig:http-header-example}
+illustrates possible content options used to match against certain
+strings that appear in an http message. A check mark (resp., cross)
+indicates a match (resp., no match).
 
@@
 \includegraphics[scale=0.5]{figs/http_header-example.png}
-\caption{\CodeIn{http\_header} example.}
+\caption{\CodeIn{http\_header} example.\Mar{Lucas, simplifique esta
+    figura (reduza o tamanho da imagem). Mostre apenas os casos onde ha um match, por exemplo,
+    com host=www.google.com e connection=keep-alive.}}
 \label{fig:http-header-example}
@@
 
-\tname{} leverages the fact that these modifiers are commonly used to
-define rules as to better rank the rules it synthesizes. More
-precisely, it builds on the observation that strings acessible in
-positions associated with content modifiers (\eg{}, GET, Host,
-Connection) are more likely to produce useful rules. \tname\ uses the
-frequency of content modifiers to discriminate rules that are more or
-less likely to reflect actual rules. For example, historical data
-indicates that content options that use the \CodeIn{http\_connection}
-modifier are much less likely than content options that use the
-modifier
+%\tname\ uses the
+%frequency of content modifiers to discriminate rules that are more or
+%less likely to reflect actual rules.
+
+\tname{} builds on our observation that these modifiers are commonly
+used to define existing rules. More precisely, \tname\ builds on the
+observation that the strings accessible in positions associated with
+content modifiers (\eg{}, GET, Host, Connection) are more common. For
+example, historical data indicates that content options that use the
+modifier \CodeIn{http\_connection} are much less likely than content
+options that use the modifier
 \CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
-shows the histogram of content modifiers for http. Although content
-modifiers are applicable to rules of any protocol supported by
-Suricata, currently, \tname\ only supports tcp and http. These are the
-protocols involving ~80\% of the rules from our data set.
+shows the histogram of content modifiers for http. (We removed the
+prefix ``http'' on the modifier for space.) Although content modifiers
+are applicable to rules of any protocol supported by Suricata,
+currently, \tname\ only supports this features for tcp and http. But
+these rules comprise $\sim$80\% of the rules of our data set.
 
@@
 
-Similarly to the the previous heuristic functions, this function uses
-the frequency of content modifiers to calibrate the score of each
-rule. More precisely, we compute this function with the following
-formula:
+We compute the heuristic function for content modifiers with the
+following formula:
 
@@
 \noindent
-, where $N$ is the number of contents options, $g(i)$ is the
-contribution of content option $i$ to the score,
+, where $N$ is the number of contents options present in a rule,
+$g(i)$ is the contribution of content option $i$ to the score,
 $\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
@@
 
-Consider a rule containing the option \CodeIn{content:
-  ``/HtmlAdaptor''}. The contribution of this option to the score of
-the rule that uses it would be $0.5$
-(=$\frac{10,111+11}{10,111*2}$). The value of $M_i$ for this case is 2
-as the field of the HTTP message \CodeIn{Request URI}---holding the
-string ``/HtmlAdaptor''---can be accessed through the modifiers
-\CodeIn{http\_uri} and \CodeIn{http\_request\_line}. The relatively
-high heuristic value given to this option suggests it is a plausible
-option to be used in rules---the string used in the content option
-(``/HtmlAdaptor'') is stored in a specific field of the HTTP message
-(\CodeIn{Request URI}) and can be accessed by two content modifiers,
-namely (\CodeIn{http\_uri} and \CodeIn{http\_request\_line}). For the
-option \CodeIn{content: ``keep-alive''}, either the modifier
-\CodeIn{http\_header} or the modifier \CodeIn{http\_connection} could
-be used to locate the string ``keep-alive'' stored in the field
-\CodeIn{Connection} (see Figurgure~\ref{fig:http-header-example}). Consequently, for this
-case, the contribution of modifiers to this content option would be
-$0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
-used in a rule, the aggregated contribution of this heuristic function
-would be 0.35 (=(0.2+0.5)/2).
-\Gui{A quinta heuristica que utilizamos considera a popularidade das opcoes.
-Primeiro pegamos a quantidade de vezes que uma opcao eh utilizada (sem considerar repeticoes numa unica regra).
-Dividimos o valor encontrado para cada opcao pelo maior valor encontrado, encontrando sempre valores de no maximo 1.
-Entao na regra nos somamos os valores de cada opcao, dividido pelo numero de opcoes, chegando a um numero de no maximo 1, que representa a popularidade das opcoes na regra.
+To illustrate this function, consider a rule containing the option
+\CodeIn{content: ``/HtmlAdaptor''}. The contribution of this option to
+the score of the rule is $0.5$ (=$\frac{10,111+11}{10,111*2}$). The
+value of $M_i$ for this case is 2 as the string ``/HtmlAdaptor'' could
+be accessed through the modifiers \CodeIn{http\_uri} and
+\CodeIn{http\_request\_line} as that string is stored on the
+\CodeIn{Request URI} field of the http message\Mar{how do we know
+  that?}. The relatively high heuristic value reflects the fact that
+the string ``/HtmlAdaptor'', stored at \CodeIn{Request URI}, could be
+accessed by two content modifiers, namely (\CodeIn{http\_uri} and
+\CodeIn{http\_request\_line}).  For the option \CodeIn{content:
+  ``keep-alive''}, either the modifier \CodeIn{http\_header} or the
+modifier \CodeIn{http\_connection} could be used to locate the string
+``keep-alive'' stored in the field \CodeIn{Connection} (see
+Figure~\ref{fig:http-header-example}). Consequently, for this case,
+the contribution of modifiers to this content option would be $0.2$
+(=$\frac{4089+29}{10111*2}$). Assuming that both options are used in a
+rule, the aggregated contribution of this heuristic function would be
+0.35 (=(0.2+0.5)/2).
+
+\subsubsection{Popularity of options.}
+
+\Mar{--------------------------------------------------------- estou aqui}
+
+\Gui{A quinta heuristica que utilizamos
+  considera a popularidade das opcoes. Primeiro pegamos a quantidade
+  de vezes que uma opcao eh utilizada (sem considerar repeticoes numa
+  unica regra). Dividimos o valor encontrado para cada opcao pelo
+  maior valor encontrado, encontrando sempre valores de no maximo
+  1. Entao na regra nos somamos os valores de cada opcao, dividido
+  pelo numero de opcoes, chegando a um numero de no maximo 1, que
+  representa a popularidade das opcoes na regra.
 \[\frac{\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}}{N}\]}
@@
     UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
-    UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
+    %%UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
     5; \\
     \midrule
-    Apache Struts 2 & \cite{wrccdc} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
+    %%Apache Struts 2 & \cite{wrccdc} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
     ColdFusion \MyComment{PESC} & \cite{nikto} & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
@@
     WeBid LFI \MyComment{Local File Inclusion} & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
-    WordPress Login & \cite{wrccdc} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
+    %%WordPress Login & \cite{wrccdc} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
     .htaccess access\MyComment{attempted-recon} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
@@
     \midrule
-    Black Nurse & 1 & 4 & 9 & - \\    
-    Ping Scan & 6 & 5 & 9 & 1 \\
+    Black Nurse & 1 & 4 & 9 & 2 \\    
+    Ping Scan & 6 & 5 & 9 & 2 \\
     SYN Flood & 4 & 2 & 2 & 2 \\
-    UDP Teardrop & 1 & 4 & 8 & 3 \\
-    UDP Flood & - & - & - & - \\
+    UDP Teardrop & 1 & 4 & 8 & 4 \\
+    %%UDP Flood & - & - & - & - \\
     \midrule
-    Apache Struts 2 & 2 & 65 & - & - \\
-    ColdFusion & 7 & 19 & 34 & 1\\
+    %%Apache Struts 2 & 2 & 65 & - & - \\
+    ColdFusion & 7 & 19 & 34 & 3\\
     iisadmin access & 3 & 17 & 408 & 1 \\        
-    JBoss JMX & 5 & 23 & 228 & 5 \\
+    JBoss JMX & 5 & 23 & 228 & 45 \\
     Script tag in URI & 23 & 22 & 786 & 1 \\
-    WeBid LFI & 1 & 23 & 11100 & 6\\    
-    Wordpress Login & 8 & 59 & - & - \\
+    WeBid LFI & 1 & 23 & 11100 & 10\\    
+    %%Wordpress Login & 8 & 59 & - & - \\
     .htaccess access & 5 & 17 & 984 & 2\\
-    .idq access & 14 & 19 & 511 & 1 \\
-    /system32/ in URI & 4 & 20 & 1,913 & 1 \\
-    Tomcat Snoop Access & 1 & 24 & 29830 & - \\
+    .idq access & 14 & 19 & 511 & 2 \\
+    /system32/ in URI & 4 & 20 & 1,913 & 4 \\
+    Basebuilder LFI & 1 & 22 & 18133 & 34 \\
+    Tomcat Snoop Access & 1 & 24 & 29830 & 32 \\
     \bottomrule

### 2020-03-03 paper/paper.tex
pass on algo
pass

--- 2020-03-01 paper/paper.tex
@@
   \includegraphics[trim=0 395 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
-  \caption{The \tname\ workflow.}
+  \caption{The \tname\ workflow.\Mar{revisar}}
   \label{fig:overview}
@@
 
-\subsection{Step 2: Rule Minimization}
+\subsection{Step 2: Rule Creation}
 \label{sec:minimization}
 
-\Fix{consider changing the name of this step}
 The seed rule produced in the first step satisfies the criterion to
@@
 
-\tname{} leverages public data sets~\cite{tcpreplay} of benign traffic
-to guide the search for more general rules, hence addressing the
-overffiting issue. To that end, \tname{} \emph{minimizes} the rules
-produced in the previous step, preserving the invariant of \emph{only}
-capturing the positive malicious traffic. We say that a rule is
-minimal if discarding any option results in the capture of negative
-(\ie{}, benign) traffic.  The minimization procedure we propose in
-this step builds on the observation that all options from the rule
-produced in the previous step are satisifed by the positive
-traffic. Consequently, any subset of these options also captures the
-positive traffic. The minimization procedure that we describe in the
-following iteratively discards options from visited rules---derived
-from the seed rule---using the negative traffic to guide the search
-for minimal rules.
+\tname{} leverages public data sets of benign traffic~\cite{tcpreplay}
+and additional examples of the attack to guide the search for more
+general rules to address the overffiting issue. To that end, \tname{}
+\emph{minimizes} the rules produced in the previous step, preserving
+the invariant of \emph{only} capturing the positive malicious
+traffic. We say that a rule is minimal if discarding any option
+results in the capture of negative (\ie{}, benign) traffic.  The
+procedure we propose in this step builds on the observation that all
+options from the rule produced in the previous step are satisifed by
+the traffic produced by the attack. Consequently, any subset of these
+options also captures the attack. The rule creation procedure
+iteratively discards options from visited rules---derived from the
+seed rule---using the negative traffic to guide the search for minimal
+rules.
 
@@
 \begin{algorithm}[t!]
-  \noindent\hspace{-25ex}
-  \textbf{INPUT:} (overconstrained) seed rule $sr$ \\
-  \noindent\hspace{-27.5ex} 
+  %\noindent\hspace{-2ex}
+  \flushleft{}\textbf{INPUT:} (overconstrained) seed rule $sr$, bening traffic
+  $\mathit{bTr}$, malli-\\
+  cious traffic $\mathit{mTr}$\\
+  %\noindent\hspace{-27.5ex} 
   \textbf{OUTPUT:} Set of minimized rules $\mathcal O$\\
 %\vspace{-3ex}
-\captionof{algorithm}{Minimization algorithm}\label{algo1}
+\captionof{algorithm}{Rule creation algorithm}\label{algo1}
 \begin{algorithmic}[1]
-  \State $\mathcal R \gets \{\mathit{sr}\}$\Comment{initial population
-    of candidate solutions contains only the input
-    rule.}\label{line-init}
+  \State $\mathcal R \gets \{\mathit{sr}\}$\Comment{initial population of candidate solutions.}\label{line-init}
   \State $\mathcal O \gets \emptyset$
@@
                \State $r[j]\gets 0$\Comment{discard option $j$ in $r$}\label{delete-option}
-               \If{$\mathit{isNew(r)} \And\mathit{isPlausible}(r)$} 
+               \If{$\mathit{isNew(r)} \And\mathit{isPlausible}(r, \mathit{bTr})$} 
                    \State $\mathcal R' \gets \mathcal R' \cup \{r\}$\label{line-progress}
@@
        \EndFor\label{end-foreach}
-       \State $\mathcal R \gets \mathcal R'$\Comment{update new generation}
-       \State $\mathcal O \gets \mathcal O \cup \mathcal
-       R$\Comment{remember all rules}
-  \doWhile{$\mathcal R\not= \{\}$}\Comment{repeat until reaching a fixpoint}\label{end-do-while}
+       \State $\mathcal R \gets \mathcal R'$\Comment{update new generation}\label{define-next-gen}
+       \State $\mathcal O \gets \mathcal O \cup \mathcal  R$\Comment{remember all rules}\label{output-accum}
+  \doWhile{$\mathcal R\not= \{\}$}\Comment{repeat until reaching a
+    fixpoint}\label{end-do-while}\\
+  \Return $\mathit{filter(\mathcal O, mTr)}$\label{filtering}
 \end{algorithmic}
@@
 
-Algorithm~\ref{algo1} shows the pseudocode for \tname{}'s rule
-minimization. The algorithm takes as input an overconstrained rule
-produced by the reverse engineering step and produces a set $\mathcal
-O$ of minimized rules. Line~\ref{line-init} initializes the current
-population of individuals (\ie{}, rules) $\mathcal{R}$ with the seed
-rule $sr$. We encode a rule as a bitvector where each index in the
-vector represents one option; the value assigned at a given position
-indicates whether or not the option is present in the rule. One
-iteration of the outer loop
+Algorithm~\ref{algo1} shows the pseudocode for deriving rules from the
+seed rule produced at the reverse engineering step. The algorithm
+takes as input an overconstrained rule $sr$, produced by the reverse
+engineering step, and produces on output a set of alternative rules to
+$sr$ that captures every instance of the attack ($\mathit{mTr}$), but
+does not capture any message from the benign traffic ($\mathit{bTr}$).
+
+Line~\ref{line-init} initializes the current population of individuals
+(\ie{}, rules) $\mathcal{R}$ with the seed rule $sr$. We encode a rule
+as a bitvector where each index in the vector represents one option;
+the value assigned at a given position indicates whether or not the
+option is present in the rule. One iteration of the outer loop
 (lines~\ref{start-do-while}-\ref{end-do-while}) processes the current
@@
 rules for $N$ times, discarding one option of the rule each
-time. Line~\ref{copy} assigns a copy of the bitvector encoding a rule
-in $\mathcal R$ to variable $r$. Function $\mathit{len}$ returns the
+time. Line~\ref{copy} assigns a copy of the bitvector, encoding
+$\mathcal R$, to variable $r$. Function $\mathit{len}$ returns the
 length of a vector and function $\mathit{rand}$ randomly chooses one
@@
 whereas the call to $\mathit{isPlausible}$ checks if the rule only
-captures positive traffic. If $r$ satisfies both criteria, it is added
-to $\mathcal R'$, which denotes the next generation of individuals in
-the search.
-
-The algorithm terminates when no rule in the current
-generation of individuals could be further reduced without violating
-the property that a candidate solution should only capture positive
-traffic. This means that execution of one iteration has not reached
-the statement declared at line~\ref{line-progress}, which adds one
-element to the set $\mathcal{R'}$. Consequently, as $\mathcal{R'}$ is
-initialized empty that set will remain empty at the end of an
-iteration of the outer loop.
+captures positive traffic, \ie{} no message present in $\mathit{bTr}$
+is captured by $r$. If $r$ satisfies both criteria, it is added to
+$\mathcal R'$, which denotes the next generation of individuals in the
+search. After processing every candidate rule in $\mathcal R$, the
+next generation of individuals is defined (line~\ref{define-next-gen})
+and the set of discovered rules is updated (at
+line~\ref{output-accum}).
+
+The algorithm terminates when no rule in the current generation of
+individuals can be further reduced without violating the property that
+a candidate solution should only capture positive traffic. This means
+that execution of one iteration has not reached the statement declared
+at line~\ref{line-progress}, which adds one element to the set
+$\mathcal{R'}$. As $\mathcal{R'}$ is initialized empty that set will
+remain empty at the end of an iteration of the outer loop.
 
@@
 with a single option are produced in the second-to-last iteration. In
-other iterations, ules with more than one option are
+other iterations, rules with more than one option are
 produced. Conceptually, the search induces a search tree with the seed
@@
 the benign traffic to reduce the overconstrained rule produced at the
-\reveng\ step. In the following step, \tname{} ranks the rules this
-algorithm reports by their similarity to existing trusted
-rules. Intuitively, rules that do not look alike to rules written by
-humans and previously used rank low.
+\reveng\ step. In the Raking step (Section~\ref{sec:ranking}),
+\tname{} ranks the rules this algorithm reports by their similarity to
+existing trusted rules. Intuitively, rules that do not look like those
+trusted rules rank low.
 
@@
   \vspace{-1ex}  
-  The minimization algorithm reports rules that capture only the
+  The rule generation algorithm reports rules that capture only the
   positive traffic.
@@
 \begin{proof}
-  \vspace{-1.5ex}
-  Every rule created during the search captures the positive malicious
-  traffic as every rule contains a proper subset of the options from
-  the original rule and all those options are satisfied by the
-  malicious traffic. By definition of function $\mathit{isPlausible}$,
-  these rules also capture \emph{only} the positive traffic. If
-  negative traffic is captured, a rule is not added to $\mathcal R'$;
-  consequently, it cannot be an element of $\mathcal O$ and reported
-  on the output.
+  \vspace{-1.5ex} Every rule created during the search captures the
+  attack, provided as example on step one, as every rule contains a
+  proper subset of the options from the original rule and all those
+  options are satisfied by the traffic produced by the attack. By
+  definition of function $\mathit{isPlausible}$, these rules also
+  capture \emph{only} the positive traffic. If negative traffic is
+  captured, a rule is not added to $\mathcal R'$. Consequently, it
+  cannot be an element of $\mathcal O$ and reported on the
+  output. Finally, only the rules in $\mathcal O$ that capture all the
+  instances of the attack (\ie{}, variants) are reported on output.
   \vspace{-1ex}    
@@
 minimizing the rule.  \textbf{Limitations.} It is worth noting that
-although \tname{} is accurate for generating rules that discrimate
-positive and negative traffic, the quality of these rules is limited
-by the quality of the training data it uses. For the positive traffic,
-\tname{} only requires one sample, but can benefit of several
-samples\Mar{como? nao estah claro. por favor expliquem.}. The same
-applies to the negative traffic. However, in this case, \tname{} uses
-different public data sets available online (see
+the performance of \tname{} depends on the quality of inputs. For the
+positive malicious traffic, \tname{} only requires one sample (as
+input to step one), but can benefit of several samples (see
+line~\ref{filtering} of Algorithm~\ref{algo1}). The same applies to
+the negative traffic. However, in this case, \tname{} can use a large
+variety of public data sets available online (see
 Section~\ref{sec:objectanalysis-benign}).

### 2020-03-03 paper/paper.tex
added distribution rules per month

@@
 
+\begin{figure}
+    \centering
+    \scalebox{1.5}{  
+      \begin{tikzpicture}
+        \begin{axis}[
+bar width=4pt,
+            scaled ticks=true,
+            ymajorticks=false,
+            ybar stacked,
+            enlargelimits=0.08,
+    %          enlargelimits=false,
+    %          axis equal,
+            clip=false,
+            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
+            xtick=data,
+            xticklabel style={rotate=45, font=\tiny},          
+            nodes near coords,
+            every node near coord/.append style={font=\tiny},
+            nodes near coords align={vertical},
+            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
+            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
+        \end{axis}
+      \end{tikzpicture}
+    }
+    \vspace{-1ex}
+    \caption{\label{fig:distribution-rules-per-month}Rules per month - 2019.}
+\end{figure}% <---- don't forget this %
+
 \subsubsection{Number of content options.} We observed

### 2020-03-03 paper/paper.tex
table RQ2

@@
 
+\setlength{\tabcolsep}{3pt}
+\begin{table}[h!]
+  \small
+  \caption{\label{table:results}Results.}
+  \vspace{-2ex}
+  \centering
+  \begin{tabular}{lrrrrrr}
+    \toprule
+    Name &
+    Opt weights &
+    4.3.2 &
+    4.3.3 &    
+    4.3.4 &
+    4.3.5 \\ 
+    Black Nurse & 2 & 2 & 2 & 2 & 7 \\    
+    Ping Scan & 2 & 2 & 2 & 2 & 8 \\
+    SYN Flood & 2 & 2 & 2 & 2 & 2 \\
+    UDP Teardrop & 4 & 4 & 4 & 4 & 4 \\
+    %%UDP Flood & - & - & - & - & - \\
+
+    %%Apache Struts 2 & - & - & - & - & - & -\\
+    ColdFusion & 3 & 2 & 12 & 5 & 3 \\
+    iisadmin access & 1 & 1 & 1 & 8 & 1 \\        
+    JBoss JMX & 45 & 5 & 45 & 111 & 45 \\
+    Script tag in URI & 1 & 4 & 1 & 8 & 1 \\
+    WeBid LFI & 10 & 77 & 28 & 32 & 10 \\    
+    %%Wordpress Login & 8 & - & - & - & - \\
+    .htaccess access & 2 & 2 & 2 & 17 & 2 \\
+    .idq access & 2 & 2 & 2 & 12 & 2 \\
+    /system32/ in URI & 4 & 20 & 1 & 58 & 4 \\
+    Basebuilder LFI & 34 & 117 & 57 & 183 & 34 \\
+    Tomcat Snoop Access & 32 & 828 & 2 & 2564 & 32 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
 \subsection{Answering RQ3: \textRQthree}

### 2020-03-03 paper/paper.tex
adding plot

@@
 
-\Fix{add figure here}
+\begin{wrapfigure}[9in]{r}{0.22\textwidth}
+    \vspace{-2ex}  
+    \centering
+    \scalebox{0.9}{  
+      \begin{tikzpicture}
+        \begin{axis}[
+            width = 0.3\textwidth,
+            height = 0.2\textwidth,            
+            bar width=4pt,
+            scaled ticks=false,
+            ymajorticks=false,
+            ybar stacked,
+            enlargelimits=0.05,
+            clip=true,
+            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
+            xtick=data,
+            xticklabel style={rotate=45, font=\tiny},
+            tick pos=left,
+            nodes near coords,
+            every node near coord/.append style={font=\tiny},
+            nodes near coords align={vertical},
+            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
+            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
+        \end{axis}
+      \end{tikzpicture}
+    }
+    \vspace{-3ex}
+    \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
+\end{wrapfigure}% <---- don't forget this %
 
@@
 
-\begin{figure}
-    \centering
-    \scalebox{1.5}{  
-      \begin{tikzpicture}
-        \begin{axis}[
-bar width=4pt,
-            scaled ticks=true,
-            ymajorticks=false,
-            ybar stacked,
-            enlargelimits=0.08,
-    %          enlargelimits=false,
-    %          axis equal,
-            clip=false,
-            legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
-            xtick=data,
-            xticklabel style={rotate=45, font=\tiny},          
-            nodes near coords,
-            every node near coord/.append style={font=\tiny},
-            nodes near coords align={vertical},
-            symbolic x coords={Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec},]
-            \addplot coordinates {(Jan, 114) (Feb, 155) (Mar, 117) (Apr, 159) (May, 110) (Jun, 231) (Jul, 122) (Aug, 167) (Sep, 107) (Oct, 225) (Nov, 148) (Dec, 127)};
-        \end{axis}
-      \end{tikzpicture}
-    }
-    \vspace{-1ex}
-    \caption{\label{fig:distribution-rules-per-month}Rules per month - 2019.}
-\end{figure}% <---- don't forget this %
-
 \subsubsection{Number of content options.} We observed

### 2020-03-03 paper/paper.tex
pass on weights table

@@
 
-\setlength{\tabcolsep}{3pt}
+\setlength{\tabcolsep}{2pt}
 \begin{table}[h!]
@@
 
-\setlength{\tabcolsep}{3pt}
+\setlength{\tabcolsep}{5pt}
 \begin{table}[h!]
   \small
-  \caption{\label{table:results}Results.}
+  \caption{\label{table:results}Weights}
   \vspace{-2ex}
@@
     Name &
-    Opt weights &
-    4.3.2 &
-    4.3.3 &    
-    4.3.4 &
-    4.3.5 \\ 
-    Black Nurse & 2 & 2 & 2 & 2 & 7 \\    
-    Ping Scan & 2 & 2 & 2 & 2 & 8 \\
+    Opt. &
+    -f2 &
+    -f3 &    
+    -f4 &
+    -f5 \\ 
+    Black Nurse & 2 & 2 & 2 & 2 & 7$\uparrow$ \\    
+    Ping Scan & 2 & 2 & 2 & 2 &  8$\uparrow$\\
     SYN Flood & 2 & 2 & 2 & 2 & 2 \\
@@
     %%Apache Struts 2 & - & - & - & - & - & -\\
-    ColdFusion & 3 & 2 & 12 & 5 & 3 \\
-    iisadmin access & 1 & 1 & 1 & 8 & 1 \\        
-    JBoss JMX & 45 & 5 & 45 & 111 & 45 \\
-    Script tag in URI & 1 & 4 & 1 & 8 & 1 \\
-    WeBid LFI & 10 & 77 & 28 & 32 & 10 \\    
+    ColdFusion & 3 & 2\cellcolor{lightgray}$\downarrow$ & 12$\uparrow$ & 5$\uparrow$ & 3 \\
+    iisadmin access & 1 & 1 & 1 & 8$\uparrow$ & 1 \\        
+    JBoss JMX & 45 & 5\cellcolor{lightgray}$\downarrow$ & 45 & 111$\uparrow$ & 45 \\
+    Script tag in URI & 1 & 4$\uparrow$ & 1 & 8$\uparrow$ & 1 \\
+    WeBid LFI & 10 & 77$\uparrow$ & 28$\uparrow$ & 32$\uparrow$ & 10 \\    
     %%Wordpress Login & 8 & - & - & - & - \\
-    .htaccess access & 2 & 2 & 2 & 17 & 2 \\
-    .idq access & 2 & 2 & 2 & 12 & 2 \\
-    /system32/ in URI & 4 & 20 & 1 & 58 & 4 \\
-    Basebuilder LFI & 34 & 117 & 57 & 183 & 34 \\
-    Tomcat Snoop Access & 32 & 828 & 2 & 2564 & 32 \\
+    .htaccess access & 2 & 2 & 2 & 17$\uparrow$ & 2 \\
+    .idq access & 2 & 2 & 2 & 12$\uparrow$ & 2 \\
+    /system32/ in URI & 4 & 20$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 58$\uparrow$ & 4 \\
+    Basebuilder LFI & 34 & 117$\uparrow$ & 57$\uparrow$ & 183$\uparrow$ & 34 \\
+    Tomcat Snoop Access & 32 & 828$\uparrow$ & 2\cellcolor{lightgray}$\downarrow$ & 2564$\uparrow$ & 32 \\
     \bottomrule

### 2020-03-03 paper/paper.tex
pass table 2

@@
     SYN Flood & \cite{hping3} & DoS & TCP & \textbf{flags}: S,12;
-    \textbf{threshold}: type both, track by\_dst, count 5000, seconds
-    5;\\
+    \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\
     UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
-    %%UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
-    5; \\
+    %%UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5; \\
     \midrule
@@
     /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
+    Basebuilder LFI & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET"; http\_method; \textbf{content}:"/main.inc.php?"; nocase; http\_uri; \textbf{content}:"mj\_config[src\_path]="; nocase; http\_uri; \\
     Tomcat server snoop access & \cite{nikto} & AREC & HTTP & \textbf{content}:"/jsp/snp/"; http\_uri; \textbf{content}:".snp"; http\_uri;\\  

### 2020-03-03 paper/paper.tex
pass on intro

@@
 particular rule prescribes a method to capture an Active
-Reconnaissance attack for discovering which ports of a publicly
-accessible machine are open for connection (see
+Reconnaissance attack for discovering IPS of hosts in a network (see
 Section~\ref{sec:active-recon}). Rulesets exist for Snort~\cite{snort}
@@
 attacks. Unfortunately, attackers actively look for new mechanisms to
-exploit systems. For example, Figure~\Fix{plot} shows the increase in
-number of rules added to the ``Emerging Threats Open Ruleset''---a
-popular public data set of \suri\ rules---in the past 12 months.
+exploit systems. For example,
+Figure~\ref{fig:distribution-rules-per-month} shows the number of
+rules added per month in the year of 2019 to the ``Emerging Threats
+Open Ruleset'', a popular public data set of \suri\ rules. We used the
+``creation-date'' field associated with each entry in the ruleset. The
+data provides some evidence that new rules are created often.
 
@@
 rules and, consequently, help system administrators to harden the
-protection of \nids\ to network attacks. \tname{} synthesizes rules
-from malicious and benign traffic and from examples of correct rules.
+protection of \nids\ to network attacks. \tname{} synthesizes rules to
+protect against a given attack from 1) malicious traffic (\ie{},
+examples of the attack to protect against), 2) benign traffic, and 3)
+trusted rules.
 

### 2020-03-04 paper/paper.tex
pass on intro
pass on eval

@@
 \usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
+\usepgfplotslibrary{statistics}
 \usepackage{listings}
@@
 
-Intuitively, rulesets need to be constantly revised to protect a
-system against attackers. Unfortunately, creating these rules is
-time-consuming and error-prone. The language for describing rules is
-non-trivial (see Section~\ref{sec:example-suricata-rules}) and testing
-correctness of a rule requires data and careful inspection.
+Intuitively, rulesets need to be continuously revised because new
+vulnerabilities are discovered and attackers can take advantage of
+them. Unfortunately, creating these rules is time-consuming and
+error-prone. The language for describing rules is non-trivial (see
+Section~\ref{sec:example-suricata-rules}) and testing correctness of a
+rule requires data and careful inspection.
 
@@
 
-\subsection{Step 2: Rule Creation}
+\subsection{Step 2: Rules Creation}
 \label{sec:minimization}
@@
 
-\newcommand{\textRQthree}{How important are the benign and
-  malicious traffic to \tname's performance?}
+\newcommand{\textRQthree}{How sensitive \tname\ is to the amount of
+  positive examples?}
 \vspace{0.2cm}
@@
 
+This section evaluates the importance of each proposed heuristic
+function. More precisely, we evaluated the use of \tname{} without
+each of those functions with the goal of checking whether or not each
+of them contributes to the overall performance of the technique. We
+conducted that experiment by assigning the weight of the corresponding
+function to zero---effectively, disabling the function---and comparing
+the result with \tname\ using default weights, obtained with the
+procedure described on Section~\Fix{xxx}.
+
+Table~\ref{table:weights} shows results. Column ``Name'' shows the
+name of the attack and column ``All'' shows the result of \tname\ with
+all heuristic functions enabled (same as column ``Rank (step 3)'' from
+Table~\Fix{xxx}). The remaining columns show results supressing one
+heuristic function. An up arrow indicates that, considering that
+particular data point, the performance of \tname{} on that
+configuration is worse when compared to \tname's default
+configuration. A down arrow, appearing highlighted in the table,
+indicates the contrary.
+
 \setlength{\tabcolsep}{5pt}
@@
   \small
-  \caption{\label{table:results}Weights}
+  \caption{\label{table:weights}Contribution of heuristic functions.}
   \vspace{-2ex}
@@
     Name &
-    Opt. &
+    All &
     -f2 &
@@
 
+Note that the number of up arrows is relatively high and the magnitude
+of the degradation can be very high in some cases. Basebuilder and
+Tomcat are two cases where results get much worse when supressing one
+of the functions. Also note that the number of down arrows is
+relatively small. Recall that the optimization of weights looks for
+global optimum and cannot avoid those cases. Finally, note that every
+function contributed. Some functions contributed in more cases than
+others, namely \Fix{xxx}, and one of the functions, namely \Fix{xxx},
+only contributed to the performance of \tname{} on attacks involving
+non-http protocols.
+
+\begin{center}
+\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
+..........................
+\end{tcolorbox}
+\end{center}
+
 \subsection{Answering RQ3: \textRQthree}
 \label{sec:answer-rqthree}
+
+This section evaluates sensitivity of \tname{} to the number of
+attacks provided on input. Recall that \tname\ uses one example attack
+to generate the seed rule (at the Reverse Engineering step) and the
+remaining attacks to evaluate soundness of the rule (at the Rules
+Creations step). It is expected that the number and variety of
+attacks---of the same kind---plays an important role in the
+performance of the technique. To answer this research question, we
+selected the cases with the highest number of examples of attacks (see
+coolumn ``\#~Vars'' on Table~\ref{table:results}), varied the number of
+examples, and measured the position of the golden rule at the
+rank. For each number of variants $N$, we sampled five sets of size
+$N$ to provide as input to the technique.
+x
+
+\Mar{coloquem os numeros aqui para que eu gere o boxplot amanha,
+  depois da aula.
+
+  *** posicao no ranking (numeros artificais. por favor, substituam) ***
+
+  ataque 1:
+
+    1 variacao:  10, 20, 4, 2, 5
+
+    5 variacoes: 5, 4, 4, 2, 3
+
+    10 variacoes: 1, 2, 1, 2, 1
+
+    (ate numero de variacoes que aparece na Tabela 3)
+
+  ataque 2:
+     ...
+
+}
+
+%% \pgfplotsset{
+%%   box plot width/.initial=0.5em
+%% }
+%% \begin{tikzpicture}
+%%   \begin{axis}
+%%     [
+%%     ytick={1,2,3},
+%%     yticklabels={Index 0, Index 1, Index 2},
+%%     ]
+%%     \addplot+[
+%%     boxplot prepared={
+%%       median=1,
+%%       upper quartile=1.2,
+%%       lower quartile=0.4,
+%%       upper whisker=1.5,
+%%       lower whisker=0.2
+%%     },
+%%     ] coordinates {};
+%%     \addplot+[
+%%     boxplot prepared={
+%%       median=2,
+%%       upper quartile=2.3,
+%%       lower quartile=1.5,
+%%       upper whisker=2.7,
+%%       lower whisker=1
+%%     },
+%%     ] coordinates {};
+%%     \addplot+[
+%%     boxplot prepared={
+%%       median=0.7,
+%%       upper quartile=1.4,
+%%       lower quartile=0.5,
+%%       upper whisker=1.9,
+%%       lower whisker=0.1
+%%     },
+%%     ] coordinates {};
+%%   \end{axis}
+%% \end{tikzpicture}
 

### 2020-03-04 paper/R/scripttag/subject.r
adding R (boxplots)

@@ -1,0 +1,19 @@
+#!/usr/bin/env Rscript
+
+pdf("scripttag.pdf", width = 4.1, height = 2.9)
+
+df1 <- read.table("data1.txt", header = FALSE)
+df2 <- read.table("data5.txt", header = FALSE)
+df3 <- read.table("data10.txt", header = FALSE)
+df4 <- read.table("data23.txt", header = FALSE)
+
+# adjusting margins
+par(mar = c(5, 4, 4, 0), 
+    mgp = c(1, 1, 0),
+    mai = c(0.4, 0.6, 0.1, 0),
+    mfrow = c(1,2))
+
+bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,names=c("1","5","10","23"), boxwex=0.2, cex.axis=1, las=1)
+
+dev.off()
+

### 2020-03-04 paper/paper.tex
adding R (boxplots)
pass
pass on intro adn figure 1

@@
 \begin{abstract}
+  \Fix{revise}
 Network Intrusion Detection Systems (\nids{}) are a popular mechanism
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{\nids, synthesis, search}
+% \keywords{\nids, synthesis, search}
+
+%% remove on camera-ready
+\settopmatter{printacmref=false}
 
@@
 
-\begin{wrapfigure}[9in]{r}{0.22\textwidth}
-    \vspace{-2ex}  
+\begin{wrapfigure}[8in]{r}{0.24\textwidth}
+    \vspace{-5ex}  
     \centering
@@
         \begin{axis}[
+%            label style={font=\scriptsize},
+%            tick label style={font=\scriptsize}
             width = 0.3\textwidth,
@@
             ybar stacked,
-            enlargelimits=0.05,
+            enlargelimits=0.09,
             clip=true,
@@
             xtick=data,
-            xticklabel style={rotate=45, font=\tiny},
+            xticklabel style={rotate=45, font=\footnotesize},
             tick pos=left,
             nodes near coords,
-            every node near coord/.append style={font=\tiny},
+            every node near coord/.append style={font=\footnotesize},
             nodes near coords align={vertical},
@@
     }
-    \vspace{-3ex}
+    \vspace{-4ex}
     \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
@@
 $N$ to provide as input to the technique.
-x
+
+Table~\ref{fig:impact-number-attacks} shows
+results. \Fix{...elaborate...}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[scale=0.4]{R/scripttag/scripttag.pdf}
+\caption{Impact of number of positive examples.}
+\label{fig:impact-number-attacks}
+\end{figure}
 

### 2020-03-04 paper/paper.tex
pass on 2.1

@@
 
+This section presents background used in the next sections.
+
 \subsection{Rule-based and Anomaly-based \nids}
@@
 \sloppy \nids{} are typically categorized in two
-groups~\cite{kumar2007survey}: rule-based (see
-Section~\ref{sec:intro}) and anomaly-based \nids. An
-\emph{anomaly-based intrusion detector} learn usage behavior from
-benign traffic and, based on that, alert uncommon
+groups~\cite{kumar2007survey}: rule-based, as briefly described on
+Section~\ref{sec:intro}, and anomaly-based \nids. An
+\emph{anomaly-based intrusion detector} learns usage behavior from
+benign traffic and, based on that, alerts uncommon
 behavior~\cite{7579764,kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}. More
@@
 traffic manifests different characteristics compared to those of the
-benign traffic.
-
-Rule-based \nids\ can miss attacks as rulesets can become outdated
-whereas anomaly-based \nids\ can report false alarms as learning
-algorithms that generalize observations from bening traffic are
-approximate. To sum up, rule-based \nids{} and anomaly-based \nids{}
-are complementary---a rule-based \nids{} focuses on known attacks
-whereas anomaly-based \nids{} focuses on unknown potential attacks.
+benign traffic. Rule-based \nids{} and anomaly-based \nids{} are
+complementary. A rule-based \nids{} focuses on known attacks whereas
+anomaly-based \nids{} focuses on unknown potential attacks. Rule-based
+\nids\ can miss attacks as rulesets can become obsolete whereas
+anomaly-based \nids\ can report false alarms as learning algorithms
+that generalize observations from bening traffic are approximate.
 This paper focuses on rule-based \nids\ for its tremendous popularity
-in industry. Although we focused on the Suricata \nids, the principles
-we used are general to any other signature-based \nids~(\eg{},
-Snort~\cite{snort}).
+in industry\Fix{cite or tone down}. Although we used Suricata, the
+principles we used are general to any other signature-based
+\nids~(\eg{}, Snort~\cite{snort}).
 

### 2020-03-04 paper/paper.tex
pass on 2.2

@@
 
-Figure~\ref{fig:synflood-example} shows an example rule of
+Figure~\ref{fig:pingscan-example} shows an example rule of
 Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
-the Open Information Security Foundation (OISF)~\cite{oisf}.
-A Suricata rule is divided in three parts---action, header, and rule
-options~\cite{suri-rule-format}. The action part appears as the first
-word in the rule description. An action denotes the task that needs to
-be executed if the rule pattern is satisfied.  In this example, a
-message will be sent to system administrators if the rule pattern is
+the Open Information Security Foundation (OISF)~\cite{oisf}.  A
+Suricata rule is divided in three parts---action, header, and rule
+options~\cite{suri-rule-format}. The \emph{action} appears as the
+first word in the rule description (an \CodeIn{alert}, in this
+case). An action denotes the task that needs to be executed if the
+rule pattern is satisfied.  In this example, an alert message will be
+sent to system administrators if the rule pattern is
 satisfied\footnote{Technically, an IPS (Intrusion Prevention System)
   differs from an IDS in the action taken when an intrusion is
-  observed. IPS systems block malicous traffic. However, IDS and IPS
-  do not differ in the approaches that can be used to detect the
-  intrusion~\cite{ids-ips}.\Mar{check}}. The header comes after the action in the rule
-description. It restricts the information flow covered by the
-rule. The header of this rule is \CodeIn{icmp any any -> any
-  any}. It instructs Suricata to inspect \CodeIn{icmp} traffic flowing
-from any machine and port to any machine and port.\MyComment{ The variables
-\CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are
-configurable.} The rule options come after the header. It is a
-semi-colon-separated sequence of key-value pairs. Rule options serve
-to document the analyzed traffic (\eg{}, ``msg'', ``classtype'', and
-``sid'') and to characterize the attack pattern (\eg, ``flags'',
-``threshold'', and ``content''). \tname{} produces options of the
-second kind, which are \emph{relevant} to capture the attack. For
-example, the option \CodeIn{content:<arg>} checks if the argument is
-present in the payload of the message.
+  observed. IPS systems block malicous traffic whereas IDS system only
+  alert them. However, IDS and IPS do not differ in the approaches
+  used to detect the intrusion~\cite{ids-ips}.\Mar{check}}. The
+\emph{header} comes after the action in the rule description. It
+restricts the information flow covered by the rule. The header of this
+rule is \CodeIn{icmp any any -> any any}. It instructs Suricata to
+inspect \CodeIn{icmp} traffic flowing from any machine and port to any
+machine and port.\MyComment{ The variables \CodeIn{\$HOME\_NET} and
+  \CodeIn{\$EXTERNAL\_NET} are configurable.} Finally, \emph{rule
+  options} come after the header. It is a semi-colon-separated
+sequence of key-value pairs. Rule options serve to document the
+analyzed traffic (\eg{}, ``msg'', ``classtype'', and ``sid'') and to
+characterize the attack pattern (\eg, ``flags'', ``threshold'', and
+``content''). \tname{} produces options of the second kind, which are
+\emph{relevant} to capture the attack. For example, the option
+\CodeIn{content:<arg>} checks if the argument is present in the
+payload of the message.
 

### 2020-03-04 paper/R/scripttag/subject.r
data

@@ -5,7 +5,7 @@ pdf("scripttag.pdf", width = 4.1, height = 2.9)
 df1 <- read.table("data1.txt", header = FALSE)
 df2 <- read.table("data5.txt", header = FALSE)
 df3 <- read.table("data10.txt", header = FALSE)
-df4 <- read.table("data23.txt", header = FALSE)
+df4 <- read.table("data14.txt", header = FALSE)
 
 # adjusting margins
 par(mar = c(5, 4, 4, 0), 
@@ -13,7 +13,7 @@ par(mar = c(5, 4, 4, 0),
     mai = c(0.4, 0.6, 0.1, 0),
     mfrow = c(1,2))
 
-bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,names=c("1","5","10","23"), boxwex=0.2, cex.axis=1, las=1)
+bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,names=c("1","5","10","14"), boxwex=0.2, cex.axis=1, las=1)
 
 dev.off()
 

### 2020-03-04 paper/R/idqaccess/subject.r
pass on boxplots

@@ -1,11 +1,11 @@
 #!/usr/bin/env Rscript
 
-pdf("scripttag.pdf", width = 4.1, height = 2.9)
+pdf("idqaccess.pdf", width = 4.1, height = 2.9)
 
 df1 <- read.table("data1.txt", header = FALSE)
 df2 <- read.table("data5.txt", header = FALSE)
 df3 <- read.table("data10.txt", header = FALSE)
-df4 <- read.table("data23.txt", header = FALSE)
+df4 <- read.table("data14.txt", header = FALSE)
 
 # adjusting margins
 par(mar = c(5, 4, 4, 0), 
@@ -13,7 +13,7 @@ par(mar = c(5, 4, 4, 0),
     mai = c(0.4, 0.6, 0.1, 0),
     mfrow = c(1,2))
 
-bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,names=c("1","5","10","23"), boxwex=0.2, cex.axis=1, las=1)
+bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,names=c("1","5","10","14"), boxwex=0.2, cex.axis=1, las=1)
 
 dev.off()
 

### 2020-03-04 paper/paper.tex
pass on boxplots
done with section 2

@@
 rules added per month in the year of 2019 to the ``Emerging Threats
-Open Ruleset'', a popular public data set of \suri\ rules. We used the
-``creation-date'' field associated with each entry in the ruleset. The
-data provides some evidence that new rules are created often.
+Open Ruleset''~\cite{emerging-threats-open}, a popular public data set
+of \suri\ rules. We used the ``creation-date'' field associated with
+each entry in the ruleset. The data provides some evidence that new
+rules are created often.
 
@@
 \CodeIn{content:<arg>} checks if the argument is present in the
-payload of the message.
+payload of a message.
 
@@
 the name of the option, and column ``Description'' summarizes
-purpose. The option \CodeIn{content} is very common in rules to
-capture http attacks, which are the most common kind of rules. This
-option takes a sequence of bytes as argument (\eg{},
-\CodeIn{content:''SELECT ''}) and checks if that sequence is present
-in the payload of the message. Detailed description of the Suricata
-rules can be found elsewhere~\cite{suri-rule-format}. It is worth
-noting that Suricata supports a total of \totoptions\ distinct options
-covering \numproto\ distinct protocols.
+purpose. The option \CodeIn{content} is common in rules to capture
+http attacks and http rules are the most common in the Emerging
+Threats Open Ruleset~\cite{emerging-threats-open}. The content option
+takes a sequence of bytes as argument and checks if that sequence is
+present in the payload of the message.  For example, the option
+\CodeIn{content:''SELECT''} checks if the string ``SELECT'' is present
+in the payload.  Detailed description of the Suricata rules can be
+found elsewhere~\cite{suri-rule-format}. It is worth noting that
+Suricata supports a total of \totoptions\ distinct options covering
+\numproto\ distinct network protocols.
 
@@
 
-Rule options are often associated with specific fields of network
-packets. Let's consider the ICMP protocol as an
-example. Figure~\ref{fig:icmp-packet-layout} shows the layout of an
-ICMP packet. The packet fields ``Flags'' and ``Time to Live'',
-declared in the IP header, are mapped to the options \CodeIn{flagbits}
-and \CodeIn{ttl}, respectively. Likewise, the fields ``Type'' and
-``Code'', declared in the ICMP header, are mapped to the options
-\CodeIn{itype} and \CodeIn{icode}. The values of the options ``ttl'',
-``itype'', and ``icode'' are the same as the value of the
-corresponding fields in the ICMP packet. For the field ``Flags'' the
-value of the corresponding option is encoded in a bitvector. For
-example, if the ``Flags'' field stores the value 1010, the
-corresponding rule option will be \CodeIn{flagbits:RM}, as each
+We mentioned above that http rules often used content options, which
+associate with the message payload. For non-http rules, options are
+often associated with packet fields. Let us consider a rule for an
+ICMP attack as example. Figure~\ref{fig:icmp-packet-layout} shows the
+layout of an ICMP packet. The packet fields ``Flags'' and ``Time to
+Live'', declared in the IP header, are mapped to the \suri\ options
+\CodeIn{flagbits} and \CodeIn{ttl}, respectively. Likewise, the fields
+``Type'' and ``Code'', declared in the ICMP header, are mapped to the
+options \CodeIn{itype} and \CodeIn{icode}, respectively. The values of
+the options ``ttl'', ``itype'', and ``icode'' are the same as the
+value of the corresponding fields in the ICMP packet. For the field
+``Flags'' the value of the corresponding option is encoded in a
+bitvector. For example, if the ``Flags'' field stores the value 1010,
+the corresponding rule option will be \CodeIn{flagbits:RM}, as each
 position of the vector encodes a different character in the
-\CodeIn{flagbits} rule option. The value of several options can be
-determined through transformation functions, as those described
-above. One exception to this is the option \CodeIn{threshold}, which
-is typically used in denial-of-service attacks. That option operates
-over sequence of messages as opposed to a single
-message. Consequently, it is not possible to determined its values
-from a single message.
+\CodeIn{flagbits} rule option. In summary, the value of several
+options can be determined through transformation functions, as those
+informally described above. One exception to this is the option
+\CodeIn{threshold}, which is typically used in denial-of-service
+attacks. That option operates over sequence of messages as opposed to
+a single message. Consequently, it is not possible to determine its
+values from a single message.
 
@@
           nodes near coords,
-          every node near coord/.append style={font=\tiny},
+          every node near coord/.append style={font=\scriptsize},
+          xticklabel style={font=\footnotesize},          
           nodes near coords align={vertical},
@@
           every node near coord/.append style={font=\scriptsize},
+          xticklabel style={font=\footnotesize},          
           nodes near coords align={vertical},
@@
   \centering
-\includegraphics[scale=0.4]{R/scripttag/scripttag.pdf}
+  \begin{subfigure}{.2\textwidth}
+    \centering
+    \includegraphics[scale=0.4,trim=20 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
+    \caption{idqaccess}
+  \end{subfigure}%
+  \begin{subfigure}{.2\textwidth}
+    \centering
+    \includegraphics[scale=0.4,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
+    \caption{scripttag}    
+  \end{subfigure}%
 \caption{Impact of number of positive examples.}

### 2020-03-04 paper/paper.tex
done section 3 (for now)

@@
 \label{sec:suri-metas-coverage}
-\label{sec:active-recon}
+\label{sec-:active-recon}
 
@@
 information about a computing system to determine potential
-vulnerabilities. For example, Ping Scan scans the network for IPs of
-accessible hosts. An attacker sends several ICMP Echo requests to a
-range of IP addresses in the network and checks which ones respond to
-these requests. This method enables an attacker to discover which
-machines are active in the network.
+vulnerabilities. For example, the Ping Scan attack~\cite{ping-scan}
+scans the network for IPs of accessible hosts. An attacker sends
+several ICMP Echo requests to a range of IP addresses in the network
+and checks which ones respond to these requests. This method enables
+an attacker to discover which machines are active in the network.
 
@@
 and security auditing open-source tool. The essential options to
-capture the attack (see Section~\ref{sec:example-suricata-rules}) are
-in bold---the option \CodeIn{dsize: 0} indicates that the packet has
-no bytes in the payload and the option \CodeIn{itype: 8} indicates
-that the packet to be captured are ICMP Echo requests.
+capture the attack are in bold---the option \CodeIn{dsize: 0}
+indicates that the packet has no bytes in the payload and the option
+\CodeIn{itype: 8} indicates that the packet to be captured are ICMP
+Echo requests.
 
 In the following, we briefly show how \tname{} synthesizes rules for
-this attack. The action is left constant (\CodeIn{alert}) and the
-header is inferred from the protocol. For example, in this case,
-\tname{} sets the header to \CodeIn{icmp any any -> any any},
-indicating that the rule applies to \CodeIn{icmp} messages and applies
-to the most general flow description, \ie{}, it instructs the
-\nids\ to analyze the traffic flowing from/to any address and port.
-The rationale for this decision is that the choice of these parameters
-are domain-specific and subject to change by the system administrator.
+this attack. In short, \tname{} takes one or more examples of an
+attack---in the form of network messages---and produces rules to
+capture that attack. Considering the rule for the Ping Scan attack,
+whose example was reproduced with NMAP, \tname{} leaves the action of
+the corresponding rule constant (as \CodeIn{alert}) and infers the
+header by checking the protocol of the input messages. For example, in
+this case, \tname{} sets the header to \CodeIn{icmp any any -> any
+  any}, indicating that the rule applies to \CodeIn{icmp} messages and
+the most general flow description, \ie{}, it instructs the \nids\ to
+analyze the icmp traffic flowing from/to any address and port.  The
+rationale for this decision is that the choice of these parameters are
+domain-specific and subject to change by the system administrator.
 
 The main challenge in synthesizing rules is to discover the set of
-rule options that capture the malicious traffic, but allow the benign
-traffic to pass. \tname{} proceeds as follows to synthesize rule
-options. First, it analyzes the malicious traffic to extract options
-expressed in that traffic. Recall from
+rule options (see Table~\ref{table:rules}) that captures the malicious
+traffic, but allows the benign traffic to pass. \tname{} proceeds as
+follows to synthesize rule options. First, it analyzes the malicious
+traffic to extract options expressed in that traffic. Recall from
 Section~\ref{sec:rules-and-packets} that some options can be inferred
@@
 options \CodeIn{dsize:0} and \CodeIn{itype:8} are present in the set
-of inferred options (as the golden rule from
-Figure~\ref{fig:pingscan-example}), the others are not.
+of inferred options---as the golden rule from
+Figure~\ref{fig:pingscan-example} shows---the others are not.
 
@@
 problem. In this step, \tname{} searches for alternative rules that
-preserve the invariant to capture the positive malicious traffic but
-not capture the negative benign traffic. It uses the rule obtained in
-the previous step as seed to bootstrap the search for alternative
+preserve the invariant to capture the positive (/malicious) traffic
+but not capture the negative benign traffic. It uses the rule obtained
+in the previous step as seed to bootstrap the search for alternative
 plausible rules, \ie{}, rules that satisfy the aforementioned
@@
 of the options declared in the seed rule. Considering this example,
-\tname{} produces a total of \pingscanplausible{}\Fix{check!}
-plausible rules at the end of the search.  Note that there are $2^n$
-possible plausible rules, where $n$ is the number of options contained
-in the seed rule ($n=5$, in this case). \tname{} is able to prune the
+\tname{} produces a total of \pingscanplausible{} plausible rules at
+the end of the search. This is a smaller number when compared to $2^n$
+option combinations, where $n$ is the number of options contained in
+the seed rule ($n=5$, in this case). \tname{} is able to prune the
 search space and discard many implausible rules. It is worth noting
@@
 obtained by analyzing a database of existing rules. The general
-intuition is that new rules should share characteristics with
-pre-existing rules. For example, one heuristic function used by
-\tname{} leverages the \emph{size of a rule} for ranking.  The score
-of a rule $r$ that this function produces is proportional to the
-popularity of the number of options in
-$r$. Figure~\ref{fig:distribution-rule-size} shows the distribution of
-the number of content options associated with rules from a public data
-set~\cite{emerging-threats-open}. Note that rules with 2 to 6 options
-are more common. At the end of this step, \tname{} ranks the golden
-rule at the top of the ranking, including \pingscanplausible{}
-rules. It is worth noting that \tname\ consistently ranks the golden
-rule very high in most of the cases even when the technique produces a
-large number of plausible rules (see Table~\ref{table:results}).
+intuition is that new rules should share characteristics with existing
+human-written rules. For example, one heuristic function used by
+\tname{} leverages the popularity of options for ranking.  The score
+of a rule $r$, according to that heuristic alone, is proportional to
+the popularity of the options declared in
+$r$. Figure~\ref{fig:distribution-options} shows the distribution of
+options associated with rules from a public
+ruleset~\cite{emerging-threats-open} containing
+\numrulessuri\ rules. Note that rules with options \CodeIn{content}
+and \CodeIn{flow} are the most common. The options \CodeIn{dsize} and
+\CodeIn{itype}, the ones associated with the golden rule, appear as
+the 6th and 7th most popular, respectively. But, among the options
+inferred by \tname{} at the first step, \CodeIn{dsize} and
+\CodeIn{itype} are the most popular. At the end of this step, \tname{}
+ranks the golden rule second in the list including
+\pingscanplausible{} rules. It is worth noting that
+\tname\ consistently ranks the golden rule very high in most of the
+cases even when the technique produces a large number of plausible
+rules (see Table~\ref{table:results}).
 
@@
   \vspace{-5ex}      
-  \caption{\label{fig:distribution-strings}Options.}
+  \caption{\label{fig:distribution-options}Options.}
 %  \vspace{-1ex}    
@@
 \vspace{-1ex}
-\caption{Distribution of features.}
+\caption{Distributions.}
 \end{figure}

### 2020-03-04 paper/paper.tex
pass

@@
 \label{sec:suri-metas-coverage}
-\label{sec-:active-recon}
+\label{sec:active-recon}
 
@@
 
-\tname{} is a technique to synthesize rules for signature-based
+\tname{} is a technique to synthesize rules for rule-based
 \nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
-goal is to create rules that capture \emph{only} the
+goal is to automatically create rules that capture \emph{only} the
 malicious traffic.
@@
 \vspace{1ex}
-\noindent\textbf{Overview.}~To create rules, \tname\ leverages (1)
-benign and malicious network traffic and (2) rules from public
-rulesets. \tname\ takes on input malicious and benign traffic and a
-set of rules from public rulesets and produces on output a list of
+\noindent\textbf{Overview.}~\tname\ takes on input one or more
+examples of attacks and produces on output a ranked list of
 \emph{plausible rules}---\ie{}, rules that isolate the malicious
-traffic from the benign traffic---ranked in order of similarity with
-existing rules. Figure~\ref{fig:overview} shows the workflow of
-\tname{} as a pipeline of three components, which appear numbered in
-the figure.  Public data sets of benign
-traffic~\cite{tcpreplay,stratosphere-normal} and \nids\ rules
+traffic produced by the attacks from the benign
+traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
+pipeline of three components, appearing numbered in the figure.
+Public data sets of benign
+traffic~\cite{tcpreplay,stratosphere-normal} and rules
 exist~\cite{emerging-threats-open}. Consequently, a user of \tname{}
-only needs to provide the malicious traffic on input.
-
-\begin{figure}[h!]
+only needs to provide examples of attacks on input.
+
+%% To create rules, \tname\ leverages (1)
+%% benign and malicious network traffic and (2) rules from public
+%% rulesets.
+
+\begin{figure}[t!]
   \centering
   \includegraphics[trim=0 395 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
+  \vspace{-4ex}  
   \caption{The \tname\ workflow.\Mar{revisar}}
   \label{fig:overview}
-  \vspace{-2ex}
+  \vspace{-3ex}
 \end{figure}
 
-\tname\ works as follows. First, it produces a rule that captures the
-malicious traffic. In this step, \tname\ detects options directly from
-the input messages (see Section~\ref{sec:rules-and-packets}). More
-precisely, it produces a rule containing all possible options
-satisfied by the traffic. Let us refer to that rule as the \emph{seed
-  rule} and its corresponding set of options as \emph{seed
-  options}. Note that, as only the malicious traffic is considered in
-this step, the seed rule is likely overspecified, \ie{}, it may be
-more restrictive than the golden rule. Consequently, the use of that
-rule could result in false negatives. Second, \tname\ generates
-alternative rules that 1) weaken the constraints expressed in the seed
-rule while 2) still capturing the malicious traffic. Given that every
-option declared in a rule needs to be satisfied for a rule to capture
-an attack, any subset of the set of seed options, by definition,
-captures the attack characterized by the malicious traffic. \tname{}
-leverages that observation to search for rules containing a proper
-subset of options from the seed options. However, note that that
-approach alone could result in false positives as it imposes no limit
-on to what extent a rule can be relaxed. \tname{} uses benign traffic
-for that purpose. More precisely, it uses a gradient descent-like
-search to find plausible rules derived from the seed rule, obtained in
-the previous step. Finally, \tname{} uses several heuristic functions
-derived from a data set of existing rules to rank the generated
-plausible rules by their similarity with existing rules. Similarity is
-used to approximate user's intent.
+\tname\ works as follows. First, it creates a rule that captures the
+malicious traffic produced by the examples provided on input. In this
+step, \tname\ detects options directly from the input messages (see
+Section~\ref{sec:rules-and-packets}). More precisely, it creates a
+rule containing options that are guaranteed to be satisfied by the
+malicious traffic. Let us refer to that rule as the \emph{seed rule}
+and its corresponding set of options as \emph{seed options}. Note
+that, as only the malicious traffic is considered in this step, the
+seed rule is likely overspecified, \ie{}, it may be more restrictive
+than the golden rule. Consequently, the use of that rule could result
+in false negatives. Second, \tname\ generates alternative rules that
+1) weaken the constraints expressed in the seed rule while 2) still
+capturing the malicious traffic. Given that every option declared in a
+rule needs to be satisfied for a rule to capture an attack, any subset
+of the set of seed options, by definition, captures the attack
+characterized by the malicious traffic. \tname{} leverages that
+observation to search for plausible rules containing a proper subset
+of options from the seed options. However, note that that approach
+alone could result in false positives as it imposes no limit on to
+what extent a the constraints induced from the rule can be
+weakened. \tname{} uses benign traffic for that purpose. More
+precisely, it searches for plausible rules derived from the seed rule
+obtained in the previous step. Finally, \tname{} uses heuristic
+functions obtained from a data set of existing rules to rank the
+generated plausible rules by their similarity to existing
+rules. Similarity is used to approximate user's intent. The following
+sections detail each component of \tname.
 
@@
 
-The first step in the \tname's workflow is Reverse Engineering. The
-goal of this step is to extract options from the malicious
-traffic. Consequently, a rule that contains these options captures the
-malicious traffic. \tname\ uses different strategies to extract
-options based on the protocol detected in the input traffic: 1)~the
+The goal of the Reverse Engineering step is to extract options from
+the malicious traffic produced by the input example attacks. A rule
+that contains these options captures the malicious traffic by
+construction. \tname\ uses two strategies to extract options: 1)~the
 field strategy applies to any protocol and 2)~the payload strategy
-applies only to http. We elaborate these strategies in the following.
+applies only to http.  The decision of which strategy to use is based
+on the protocol associated with input traffic--the technique uses the
+payload strategy if http is detected; otherwise, it uses the field
+strategy. We elaborate these strategies in the following.
 
 For the field strategy, \tname{} parses the input messages looking for
-options associated with the network message field
-data. Section~\ref{sec:rules-and-packets} shows the correspondence
-between packet fields and some of the rule options from \suri and
-Section~\ref{sec:active-recon} shows a concrete example where the
-values for options are directly extracted from the fields of
-messages. We obtained the option-field mapping manunually, from the
-\suri\ documentation.
+options associated with the data in the fields of
+messages. Section~\ref{sec:rules-and-packets} shows the correspondence
+between packet fields and some of the rule options from \suri\ and
+Section~\ref{sec:active-recon} shows a concrete example illustrating a
+case of this strategy. We obtained the option-field mapping
+manunually, from the \suri\ documentation.
 
 Unfortunately, the relevant data for the option may be at the message
-payload; not the fields. We found empirically that this occurs
-frequently in http attacks, which are prevalent. For example,
-\percHttp\ of the \numrulessuri{} rules from a public data
-set~\cite{emerging-threats-open} we analyzed prescribe defenses
-against http attacks and the majority of these rules use content
-options. Figure~\ref{fig:rule-jboss} shows an example of a rule that
-captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the
-payload.  Content options are very common in http
-rules---\percRulesWithContent{} of the rules from the data set
-mentioned above contain at least one content
-option. Figure~\ref{fig:distribution-contents} shows the distribution
-of number of content options per rule for the data set mentioned
+payload; not the fields. By analyzing our data set of rules, we found
+that this occurs frequently in http attacks, which are prevalent. For
+example, \percHttp\ of the \numrulessuri{} rules from the ruleset we
+analyzed~\cite{emerging-threats-open} prescribe defenses against http
+attacks. In addition, \percRulesWithContent{} of these rules contain
+at least one content option. Figure~\ref{fig:rule-jboss} shows an
+example of a rule that captures traffic containing the string
+\CodeIn{"/HtmlAdaptor"} in the payload.
+Figure~\ref{fig:distribution-contents} shows the distribution of
+number of content options per rule for the data set mentioned
 above. We limited the number of content options per rule to 10 in the
@@
 The content strategy focuses on extracting content options from the
-message payload. Inspecting existing rules, we found that these
-options refer, in the payload, to sequences of characters separated by
-typical natural language delimiters such as $\backslash$t,
-$\backslash$n, $\backslash$r, and spaces. To find the values
-associated with these options, \tname{} splits the payload of the
-malicious message in tokens using these delimiters.  Considering the
-example from Section~\ref{sec:content-example}, \tname{} produced a
-total of 23 tokens, considered content option candidates.
-
-It is worth noting that we found empirically that deciding which
-reverse engineering strategy to use based on the protocol of the input
+message payload. By inspecting existing rules, we found that these
+options refer to sequences of characters separated by typical natural
+language delimiters such as $\backslash$t, $\backslash$n,
+$\backslash$r, and spaces. To find the values associated with these
+options, \tname{} splits the payload in tokens using these delimiters.
+Considering the example from Section~\ref{sec:content-example},
+\tname{} produced a total of 23 tokens\Mar{@Lucas, isto estah certo?
+  Se sim, como isto cai para 5 (pela Tabela 3)}, considered content
+option candidates.
+
+It is worth noting that we found that deciding which reverse
+engineering strategy to use based on the protocol of the input
 (malicious) messages was effective. We used the payload strategy for

### 2020-03-05 paper/paper.tex
on section 4 (up to 4.3)

@@
 \nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
-goal is to automatically create rules that capture \emph{only} the
-malicious traffic.
+goal is to automatically create rules that capture an attack of
+interest.
 
@@
 \noindent\textbf{Overview.}~\tname\ takes on input one or more
-examples of attacks and produces on output a ranked list of
+examples of the same attack and produces on output a ranked list of
 \emph{plausible rules}---\ie{}, rules that isolate the malicious
@@
 The seed rule produced in the first step satisfies the criterion to
-only capture positive traffic. However, that rule is overconstrained.
-It could include options that are irrelevant to capturing the attack
-and could lead to false negatives---the \nids\ may be unable to
-capture manifestations of the same attack whose messages do not
-exhibit the data satisfying irrelevant options.
+capture the traffic produced by the example attacks. However, that
+rule is overconstrained.  It may include options that are irrelevant
+to capture the attack and lead to false negatives---the \nids\ may be
+unable to capture manifestations of the same attack whose messages do
+not exhibit the data satisfying irrelevant options. For example, the
+seed rule produced after the Reverse Engineering step on the Ping Scan
+example from Section~\ref{sec:active-recon} contains options
+\CodeIn{icode:0}, \CodeIn{icmp\_id:23570}, and
+\CodeIn{icmp\_seq:3439}, which are irrelevant to capture that attack.
 
@@
 
-\tname{} leverages public data sets of benign traffic~\cite{tcpreplay}
-and additional examples of the attack to guide the search for more
-general rules to address the overffiting issue. To that end, \tname{}
-\emph{minimizes} the rules produced in the previous step, preserving
-the invariant of \emph{only} capturing the positive malicious
-traffic. We say that a rule is minimal if discarding any option
-results in the capture of negative (\ie{}, benign) traffic.  The
-procedure we propose in this step builds on the observation that all
-options from the rule produced in the previous step are satisifed by
-the traffic produced by the attack. Consequently, any subset of these
-options also captures the attack. The rule creation procedure
+To address the overffiting issue, \tname{} leverages public data sets
+of benign traffic~\cite{tcpreplay} and additional examples of the
+attack to guide the search for more general rules. To that end,
+\tname{} \emph{minimizes} the rules produced in the previous step,
+preserving the invariant of \emph{only} capturing the positive
+malicious traffic. We say that a rule is minimal if discarding any
+option results in the capture of negative (\ie{}, benign) traffic.
+The procedure we propose in this step builds on the observation that
+all options from the rule produced in the previous step are satisifed
+by the traffic produced by the attack. Consequently, any subset of
+these options also captures the attack. The rule creation procedure
 iteratively discards options from visited rules---derived from the
@@
   %\noindent\hspace{-2ex}
-  \flushleft{}\textbf{INPUT:} (overconstrained) seed rule $sr$, bening traffic
-  $\mathit{bTr}$, malli-\\
-  cious traffic $\mathit{mTr}$\\
+  \flushleft{}\textbf{INPUT:} (overconstrained) seed rule $sr$, pcap
+  files for bening traffic\\
+  ($\mathit{bTr}$), pcap files for mallicious traffic ($\mathit{mTr}$)\\
   %\noindent\hspace{-27.5ex} 
@@
 Algorithm~\ref{algo1} shows the pseudocode for deriving rules from the
-seed rule produced at the reverse engineering step. The algorithm
-takes as input an overconstrained rule $sr$, produced by the reverse
-engineering step, and produces on output a set of alternative rules to
-$sr$ that captures every instance of the attack ($\mathit{mTr}$), but
-does not capture any message from the benign traffic ($\mathit{bTr}$).
+seed rule produced at the Reverse Engineering step. The algorithm
+takes as input an overconstrained rule $sr$ and pcap files~\cite{pcap}
+to reconstruct the benign ($\mathit{bTr}$) and malicious traffic
+($\mathit{mTr}$). The algorithm produces on output a set of
+alternative rules to $sr$ that captures every instance of the attack
+(as per $\mathit{mTr}$), but does not capture any message from the
+benign traffic (as per $\mathit{bTr}$).
 
@@
 time. Line~\ref{copy} assigns a copy of the bitvector, encoding
-$\mathcal R$, to variable $r$. Function $\mathit{len}$ returns the
+$\mathit{ro}$, to variable $r$. Function $\mathit{len}$ returns the
 length of a vector and function $\mathit{rand}$ randomly chooses one
@@
 delete at line~\ref{delete-option}. The call to $\mathit{isNew}$
-checks if $r$ has not been previously generated through another path
+checks if $r$ has not been previously generated in another iteration
 whereas the call to $\mathit{isPlausible}$ checks if the rule only
-captures positive traffic, \ie{} no message present in $\mathit{bTr}$
-is captured by $r$. If $r$ satisfies both criteria, it is added to
-$\mathcal R'$, which denotes the next generation of individuals in the
-search. After processing every candidate rule in $\mathcal R$, the
-next generation of individuals is defined (line~\ref{define-next-gen})
-and the set of discovered rules is updated (at
-line~\ref{output-accum}).
-
-The algorithm terminates when no rule in the current generation of
-individuals can be further reduced without violating the property that
-a candidate solution should only capture positive traffic. This means
-that execution of one iteration has not reached the statement declared
-at line~\ref{line-progress}, which adds one element to the set
-$\mathcal{R'}$. As $\mathcal{R'}$ is initialized empty that set will
-remain empty at the end of an iteration of the outer loop.
+captures positive traffic, \ie{}, it checks that no message present in
+$\mathit{bTr}$ is captured by $r$. If $r$ satisfies both criteria, it
+is added to $\mathcal R'$, which denotes the next generation of
+individuals in the search. After processing every candidate rule in
+$\mathcal R$, the next generation of individuals is defined
+(line~\ref{define-next-gen}) and the set of discovered rules is
+updated (at line~\ref{output-accum}).
+
+The do-while loop terminates when no rule in the current generation of
+individuals can be further reduced without violating the invariant
+that a candidate solution should only capture positive traffic. This
+means that execution of one iteration of the do-while loop has not
+reached the statement declared at line~\ref{line-progress}, which adds
+one element to the set $\mathcal{R'}$. As $\mathcal{R'}$ is
+initialized empty, that set will remain empty at the end of the
+execution of the loop from
+lines~\ref{start-foreach}-\ref{end-foreach}. The algorithm returns the
+rules from $\mathcal O$ that capture \emph{all} example attacks
+provided on input (line~\ref{filtering}).
 
@@
 produced. Conceptually, the search induces a search tree with the seed
-rule $sr$ as the root, single-option rules as child nodes, and rules
-with more options as internal nodes. The edges of the tree denote the
-ancestor relation between two rules.
-
-It is also worth noting that the algorithm reports every rule in the
-tree instead of only those that are surely minimal, \ie{} rules with
-one option. The reason is that the the traffic used to check
+rule $sr$ as root, single-option rules as child nodes, and rules with
+more options as internal nodes. The edges of the tree denote the
+ancestor relation between a pair of rules. It is also worth noting
+that the algorithm reports every rule in the tree instead of only
+those with one option. The reason is that the traffic used to check
 plausibility is limited by definition---reporting only minimal rules
-with respect to that traffic could discard the correct rule. In
-summary, this algorithm reports many rules that are plausible by using
-the benign traffic to reduce the overconstrained rule produced at the
-\reveng\ step. In the Raking step (Section~\ref{sec:ranking}),
+with respect to that traffic could discard the correct rule.
+
+In summary, this algorithm reports many rules that are plausible by
+using the benign traffic to reduce the overconstrained rule produced
+at the \reveng\ step. In the Raking step (Section~\ref{sec:ranking}),
 \tname{} ranks the rules this algorithm reports by their similarity to
@@
   \vspace{-1.5ex} Every rule created during the search captures the
-  attack, provided as example on step one, as every rule contains a
-  proper subset of the options from the original rule and all those
-  options are satisfied by the traffic produced by the attack. By
-  definition of function $\mathit{isPlausible}$, these rules also
-  capture \emph{only} the positive traffic. If negative traffic is
-  captured, a rule is not added to $\mathcal R'$. Consequently, it
-  cannot be an element of $\mathcal O$ and reported on the
-  output. Finally, only the rules in $\mathcal O$ that capture all the
-  instances of the attack (\ie{}, variants) are reported on output.
+  attack provided as example on step one as every created rule
+  contains a proper subset of the options from the original rule and
+  all those options are, by definition, satisfied by the traffic
+  produced by the original attack. In addition, from the definition of
+  function $\mathit{isPlausible}$, reported rules capture \emph{only}
+  the positive traffic. If negative traffic is captured, a rule is not
+  added to $\mathcal R'$. Consequently, it cannot be an element of
+  $\mathcal O$ and reported on the output. Finally, from the
+  definition of function $\mathit{filter}$, only the rules in
+  $\mathcal O$ that capture all the instances of the attack (\ie{},
+  variants) are reported on output.
   \vspace{-1ex}    

### 2020-03-05 paper/R/scripttag/subject.r
adjusting scripts

@@ -3,9 +3,12 @@
 pdf("scripttag.pdf", width = 4.1, height = 2.9)
 
 df1 <- read.table("data1.txt", header = FALSE)
-df2 <- read.table("data5.txt", header = FALSE)
-df3 <- read.table("data10.txt", header = FALSE)
-df4 <- read.table("data23.txt", header = FALSE)
+df2 <- read.table("data2.txt", header = FALSE)
+df3 <- read.table("data3.txt", header = FALSE)
+df4 <- read.table("data4.txt", header = FALSE)
+df5 <- read.table("data5.txt", header = FALSE)
+df6 <- read.table("data10.txt", header = FALSE)
+df7 <- read.table("data23.txt", header = FALSE)
 
 # adjusting margins
 par(mar = c(5, 4, 4, 0), 
@@ -13,7 +16,7 @@ par(mar = c(5, 4, 4, 0),
     mai = c(0.4, 0.6, 0.1, 0),
     mfrow = c(1,2))
 
-bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,names=c("1","5","10","23"), boxwex=0.2, cex.axis=1, las=1)
+bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,df6$V1,df7$V1,names=c("1","2","3","4","5","10","23"), boxwex=0.2, cex.axis=1, las=1)
 
 dev.off()
 

### 2020-03-05 paper/paper.tex
pass on Section 4.3

@@
 Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
-check if any benign message was captured by the rule. It is worth
+check if any benign message was captured by the rule. \Mar{discuss
+  filter function.} It is worth
 noting that the benign traffic is stored in \emph{pcap}
@@
 different criterion to estimate how similar a created rule is to those
-from public rulesets~\cite{emerging-threats-open}. The
-intuition is that rules that look more similar to existing trusted
-rules rank higher and could catch the attention of a system
-administrator. The score of a rule is the weighted average across the
-values produced by each heuristic function. The codomain of each
-function ranges on the 0-1 interval. Hence, the score of a rule also
-ranges on the 0-1 interval. We describe in the following each of these
-heuristic functions.
+from public rulesets~\cite{emerging-threats-open}. The intuition is
+that rules that look more similar to existing trusted rules are more
+likely to be useful. The aggregated score of a rule is the weighted
+average across the values produced by each heuristic function. The
+codomain of each function ranges on the 0-1 interval. Hence, the score
+of a rule also ranges on the 0-1 interval. We describe in the
+following each of these heuristic functions.
 
@@
 
-\subsubsection{Size of a rule}
-\label{sec:heuristic-size-of-rule}
-One element that \tname{} uses to measure similarity between a
-synthesized and a human-created rule is the size of the synthetic
-rule, \ie{}, the number of options it contains. Informally, rules with
-popular sizes score high whereas unpopular rules score
-low. Figure~\ref{fig:distribution-rule-size} shows the histogram of
-rule sizes for our rule data set. The x-axis shows the number of
-options included in a rule (up to 10) while the y-axis shows the
-number of rules with the corresponding number of options. Note that
-there are 11 rules with no options. These rules only use the header to
-signal a traffic as malicious. Also note that most rules have 2 to 6
-options. \tname{} values rules in that size range higher compared to
-other sizes. We calculate the contribution of size of a given rule
-$r$ with the following formula $\mathit{rules\_per\_size}[\mathit{len(r)}]/\mathit{max(rules\_per\_size)}$, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
-$\mathit{rules\_per\_size}$ encodes the distribution of sizes as a
-dictionary mapping sizes to number of associated rules, and
-$\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
-for any dictionary entry as to normalize the measurement in the 0-1
-range. Consequently, the contribution of size is proportional to its
-popularity. The contribution of this heuristic would be 0.684
-(=3,662/5,353) to a rule with three options and 0.028 (=154/5,353) to
-a rule with nine options.
+%% Note that there are 11 rules with no
+%% options. These rules only use the header to signal a traffic as
+%% malicious. Also note that most rules have 2 to 6 options.
+
+\subsubsection{Popularity of options.}
+One heuristic function that \tname{} uses to measure similarity
+between a synthesized rule and a human-created rule is the popularity
+of the options used in those rules. According to this heuristic, rules
+containing popular options score higher compared to those containing
+unpopular rules. Figure~\ref{fig:distribution-options} shows the
+histogram of rule options for the ruleset we
+used~\cite{emerging-threats-open}. The x-axis shows the options used
+in rules (the 10 most common) and the y-axis shows the number of rules
+using the corresponding option. Note that the options
+\CodeIn{content}, \CodeIn{flow}, and \CodeIn{prce} are very
+popular. \tname\ uses the following formula to compute the
+contribution of option popularity to the score of a rule $r$:
+
+\vspace{-2ex}
+\[\frac{\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}}{N}\]
+\vspace{-2ex}
+
+\noindent
+The term $\mathit{option_i}$ denotes the $i$th option in the bitvector
+encoding $r$. The term $\mathit{option\_frequency}$ encodes the
+distribution of option popularity as a dictionary mapping
+numbers---indicating popularity---to corresponding
+options. Consequently, the term $\mathit{option\_frequency[option_i]}$
+denotes the number of rules, from our dataset, containing
+$\mathit{option\_i}$. The term $\mathit{max(option\_frequency)}$
+refers to the maximum number of rules for any dictionary entry (to
+normalize the measurement in the 0-1 range), and $N$ is the number of
+options in $r$. To illustrate, the contribution of this function to
+the score of a rule would be $0.26$ ($=(10181+3049)/(2*25699)$) for a
+rule with options \CodeIn{prce} and \CodeIn{flowbits} and $0.01$
+($=(155+108)/(2*25699)$) for a rule with options \CodeIn{itype} and
+\CodeIn{icode}.
+
+%% \Gui{A quinta heuristica que utilizamos
+%%   considera a popularidade das opcoes. Primeiro pegamos a quantidade
+%%   de vezes que uma opcao eh utilizada (sem considerar repeticoes numa
+%%   unica regra). Dividimos o valor encontrado para cada opcao pelo
+%%   maior valor encontrado, encontrando sempre valores de no maximo
+%%   1. Entao na regra nos somamos os valores de cada opcao, dividido
+%%   pelo numero de opcoes, chegando a um numero de no maximo 1, que
+%%   representa a popularidade das opcoes na regra.}
+
+%% \subsubsection{Size of a rule}
+%% \label{sec:heuristic-size-of-rule}
+%% \tname{} values rules in that size range higher compared to
+%% other sizes.
+
+%% We calculate the contribution of size of a given rule
+%% $r$ with the  $\mathit{rules\_per\_size}[\mathit{len(r)}]/\mathit{max(rules\_per\_size)}$, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
+%% $\mathit{rules\_per\_size}$ encodes the distribution of sizes as a
+%% dictionary mapping sizes to number of associated rules, and
+%% $\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
+%% for any dictionary entry as to normalize the measurement in the 0-1
+%% range. 
+%% Consequently, the contribution of size is proportional to its
+%% popularity. The contribution of this heuristic would be 0.684
+%% (=3,662/5,353) to a rule with three options and 0.028 (=154/5,353) to
+%% a rule with nine options.
 
@@
 \pgfplotsset{every tick label/.append style={font=\tiny}}
+
+%%   \begin{subfigure}{.25\textwidth}
+%%     \centering
+%%     \scalebox{0.8}{
+%%       \begin{tikzpicture}
+%%         \begin{axis}[
+%%           bar width=4pt,
+%%           scaled ticks=true,
+%%           ymajorticks=false,
+%%           ybar stacked,
+%%           enlargelimits=0.08,
+%% %          enlargelimits=false,
+%% %          axis equal,
+%%           clip=false,
+%%           legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
+%%           xtick=data,
+%% %          xticklabel style={rotate=45},          
+%%           nodes near coords,
+%%           every node near coord/.append style={font=\scriptsize},
+%%           xticklabel style={font=\footnotesize},          
+%%           nodes near coords align={vertical},
+%%           symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+%%           ]
+%%           \addplot coordinates {(0,11) (1,1301) (2,4910) (3,3662) (4,3250) (5,5353) (6,2911) (7, 1337) (8, 1246) (9,763) (10, 495)};
+%%         \end{axis}
+%%       \end{tikzpicture}
+%%     }
+%%     \vspace{-1ex}
+%%     \caption{\label{fig:distribution-rule-size}Rule sizes.}
+%%   \end{subfigure}%
 
@@
           xtick=data,
-%          xticklabel style={rotate=45},          
+          xticklabel style={rotate=30},
           nodes near coords,
@@
           nodes near coords align={vertical},
-          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode, flags, tls.fingerprint},          
         ]
-          \addplot coordinates {(0,11) (1,1301) (2,4910) (3,3662) (4,3250) (5,5353) (6,2911) (7, 1337) (8, 1246) (9,763) (10, 495)};
+        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155) (icode, 108) (flags, 97) (tls.fingerprint, 61)};
       \end{axis}
@@
   }
-    \vspace{-1ex}
-    \caption{\label{fig:distribution-rule-size}Rule sizes.}
+  \vspace{-5ex}
+  \hspace{4ex}
+  \caption{\label{fig:distribution-options}Option popularity.}
+%  \vspace{-1ex}    
 \end{subfigure}%  
@@
 \end{subfigure}%
-\\
-\begin{subfigure}{.2\textwidth}
-  \centering
-  \scalebox{0.8}{  
-    \begin{tikzpicture}
-      \begin{axis}[
-          bar width=4pt,
-          scaled ticks=true,
-          ymajorticks=false,
-          ybar stacked,
-          enlargelimits=0.08,
-%          enlargelimits=false,
-%          axis equal,
-          clip=false,
-          legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
-          xtick=data,
-          xticklabel style={rotate=35},          
-          nodes near coords,
-          every node near coord/.append style={font=\scriptsize},
-          xticklabel style={font=\footnotesize},          
-          nodes near coords align={vertical},
-          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode, flags, tls.fingerprint},          
-        ]
-        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155) (icode, 108) (flags, 97) (tls.fingerprint, 61)};
-      \end{axis}
-    \end{tikzpicture}
-  }
-  \vspace{-5ex}      
-  \caption{\label{fig:distribution-options}Options.}
-%  \vspace{-1ex}    
-\end{subfigure}%
 \vspace{-1ex}
@@
 
+\Mar{------------------------------------------ estou aqui}
 \subsubsection{Number of content options.} We observed
@@
 
-\subsubsection{Popularity of options.}
-
-\Mar{--------------------------------------------------------- estou aqui}
-
-\Gui{A quinta heuristica que utilizamos
-  considera a popularidade das opcoes. Primeiro pegamos a quantidade
-  de vezes que uma opcao eh utilizada (sem considerar repeticoes numa
-  unica regra). Dividimos o valor encontrado para cada opcao pelo
-  maior valor encontrado, encontrando sempre valores de no maximo
-  1. Entao na regra nos somamos os valores de cada opcao, dividido
-  pelo numero de opcoes, chegando a um numero de no maximo 1, que
-  representa a popularidade das opcoes na regra.
-\[\frac{\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}}{N}\]}
+\subsubsection{Optimization}
 
@@
 
+\setlength{\tabcolsep}{5pt}
 \begin{table*}[t!]
@@
   \centering
-  \begin{tabular}{llllp{10cm}}
+  \begin{tabular}{llllp{8cm}}
     \toprule

### 2020-03-05 paper/paper.tex
pass on 4.3.2 and 4.3.3

@@
 
-\subsubsection{Popularity of options.}
+\subsubsection{\label{sec:popularity-options}Popularity of options.}
 One heuristic function that \tname{} uses to measure similarity
@@
 \vspace{-2ex}
-\[\frac{\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}}{N}\]
+\[\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}/N\]
 \vspace{-2ex}
@@
 
-\Mar{------------------------------------------ estou aqui}
 \subsubsection{Number of content options.} We observed
-that \percContentOptions\ of the rules from the ruleset we are
-analyzing contain content options, but the distribution of those
-contents per rule is not uniform. Figure~\ref{fig:distribution-contents} shows the
-corresponding distribution. Given the importance of this kind of
-option to the definition of rules, we defined a heuristic function
-focused on the typical number of content options in a rule. We
-calculate the contribution of this function as described on previous
-Section~\ref{sec:heuristic-size-of-rule}.
-
-%% to the score of a rule
-%% with the following formula:
-
-%% \vspace{-2ex}
-%% \[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]
-%% \vspace{-2ex}
+that \percContentOptions\ of the rules from the ruleset we analyzed
+contain content options, but the distribution of those contents per
+rule is not uniform. Figure~\ref{fig:distribution-contents} shows the
+distribution of number of contents in a rule. Results indicate nearly
+73\% of the rules from our ruleset contain one to three content
+options. Hence, we found that that information could help
+\tname\ discriminate useful rules. \tname{} computes the contribution
+of the number of content options to the score of a rule $r$ with the
+fraction
+$\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}$,
+where $\mathit{len(r)}$ denotes the number of content options in $r$
+and $\mathit{rules\_per\_number\_of\_contents}$ denotes the
+distribution of number of contents (encoded with a dictionary as in
+the function described on
+Section~\ref{sec:popularity-options}). Consequently the term
+$\mathit{rules\_per\_number\_of\_contents[len(r)]}$ corresponds to the
+number of rules in the ruleset with the same number of contents as
+$r$. The denominator of the fraction normalizes the contribution in
+the 0-1 range. To illustrate, the contribution of this heuristic
+function to the score a rule would be $.64$ $(=5841/9164)$ to a rule
+with 3 content options and $.18$ $(=1622/9164)$ to a rule with 5
+content options.
 
@@
 \subsubsection{Popularity of content strings.} 
-The heuristic described in this section leverages content rarity to
-discriminate rules---it weighs higher (respectively, lower) rules that
-use popular (respectively, unpopular) strings in content options. The
-score of a rule $r$ on this function is given by the arithmetic mean
-of the popularity scores of each string used in contents in $r$. More
+This heuristic leverages content rarity to discriminate rules---it
+weighs higher (respectively, lower) rules that use popular
+(respectively, unpopular) strings in content options. The score of a
+rule $r$ on this function is given by the arithmetic mean of the
+popularity scores of each string used in contents in $r$. More
 precisely, the contribution of content rarity is given by the
@@
 
-\vspace{-2.5ex}
-\[\frac{\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}}{N}\]
-\vspace{-2.5ex}
+\vspace{-1ex}
+\[\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}/N\]
+\vspace{-1ex}
 
 \noindent
-, where $\mathit{content_i}$ denotes the $i$-th content in a given
+, where $\mathit{content_i}$ denotes the $i$th content in a given
 rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
-denotes the frequency of the string used in the $i$-th content, the
+denotes the frequency of the string used in the $i$th content, the
 term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
@@
 rule. Figure~\ref{fig:distribution-strings} shows the distribution of
-the ten most popular strings in our ruleset. To illustrate this
-function, let us consider a rule with two content options, using the
-strings ``GET'' and ``ID'', respectively. The contribution of that
-rule for the computation of similarity score would be 0.601
-(=(1+.203)/2). The relatively high weight reflects the fact that
-``GET'' and ``ID'' are popular contents.
+the ten most popular strings in our ruleset. To illustrate, let us
+consider a rule with whose content options use the strings ``GET'' and
+``ID'', respectively. The contribution of that rule for the
+computation of similarity score would be 0.601 (=(1+.203)/2). The
+relatively high weight reflects the fact that ``GET'' and ``ID'' are
+popular content strings.
+
+\Mar{------------------------------------------------ estou aqui}
 

### 2020-03-05 paper/paper.tex
minor change

@@
 \subsubsection{Optimization}
+\label{sec:weight-optimization}
 
@@
 %%    JBoss JMX & 1,0,1,1 & 107 \\
-%%    WeBid & - & - \\    
+%%    WetedhhtedBid & - & - \\    
 %%    ColdFusion & 0,1,1,1 & 7 \\
@@
 This section evaluates the importance of each proposed heuristic
-function. More precisely, we evaluated the use of \tname{} without
-each of those functions with the goal of checking whether or not each
-of them contributes to the overall performance of the technique. We
-conducted that experiment by assigning the weight of the corresponding
-function to zero---effectively, disabling the function---and comparing
-the result with \tname\ using default weights, obtained with the
-procedure described on Section~\Fix{xxx}.
+function. To that end, we evaluated the use of \tname{} without each
+of these functions with the goal of checking how much each of them
+contributes to the overall performance of the technique. Recall that
+the score a function $r$ is given by the weigthed average of the
+values given by each heuristic function, \ie{} the score of $r$ is
+obtained with the aggregate function $f(r)=\sum_{i}^{N}
+w_i*f_i(r)/N;$, where $w_i$ is the weight of heuristic $i$, $f_i(r)$
+is the score of heuristic $i$ on $r$, and $N$ is the number of
+heuristic---four, in our case. To assess the value of each heuristic
+function, we assigned its weight to zero---effectively, disabling the
+function---and compared the result with \tname\ using default weights,
+obtained with the procedure described on
+Section~\ref{sec:weight-optimization}.
 
@@
 name of the attack and column ``All'' shows the result of \tname\ with
-all heuristic functions enabled (same as column ``Rank (step 3)'' from
-Table~\Fix{xxx}). The remaining columns show results supressing one
-heuristic function. An up arrow indicates that, considering that
-particular data point, the performance of \tname{} on that
-configuration is worse when compared to \tname's default
+all heuristic functions enabled; as in column ``Rank (step 3)'' from
+Table~\ref{table:results}. The remaining columns show results obtained
+by running the same experiment as in Section~\ref{sec:answer-rqone},
+but supressing one heuristic function. An up arrow indicates that,
+considering that particular data point, the performance of \tname{} on
+that configuration is worse when compared to \tname's default
 configuration. A down arrow, appearing highlighted in the table,
 indicates the contrary.
+
+\Mar{--------------------------------- estou aqui}
 

### 2020-03-05 paper/paper.tex
pass

@@
 
+{\small
 \vspace{-2ex}
@@
 \vspace{-2ex}
+}
 
@@
 
+{\small
 \vspace{-1ex}
@@
 \vspace{-1ex}
-
+}
 \noindent
@@
 
-\Mar{------------------------------------------------ estou aqui}
-
 \subsubsection{Popularity of http content modifiers.} As previously
@@
 options. Furthermore, nearly half of the rules that use content
-options, use content modifiers on http message payloads. Content
-modifiers serve to indicate what part of a network packet the string
-mentioned in the content option should appear. For example, the
-modifier \CodeIn{http\_header} in the option \CodeIn{content:
-  ``keep-alive''; http\_header} indicates that the string
-``keep-alive'' should be present in the header of the
-message\footnote{A keep-alive connection attribute indicates that a
-  single TCP connection should remain open for multiple HTTP
-  requests/responses.}.  Figure~\ref{fig:http-header-example}
+options use content modifiers to qualify the search. Http modifiers
+typically serve to specify parts of the message payloads that the
+input string should be matched against. For example, the modifier
+\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
+  http\_header} indicates that the string ``keep-alive'' should be
+present in the header of the http message\footnote{A keep-alive
+  connection attribute indicates that a single tcp connection should
+  remain open for multiple http requests/responses.}. Note that, in
+this case, the content modifier refers to a field of an http message,
+which appears in the payload of a tcp message, which is the target of
+the option \CodeIn{content}. Figure~\ref{fig:http-header-example}
 illustrates possible content options used to match against certain
-strings that appear in an http message. A check mark (resp., cross)
-indicates a match (resp., no match).
+strings that appear in an http message. A check mark (respectively,
+cross) indicates a match (respectively, no match).
 
@@
 
-%% we observed---by analyzing
-%% the data set of existing rules (see
-%% Section~\ref{sec:data set-benign})---that these rules often indicate
-%% what part of the message the content string should be checked
-%% against. For example,
-%% For example, the
-%% creator of the rule may want to indicate that a specific string be
-%% present in the header of an HTTP request message. Content
-%% \emph{modifiers} serve this purpose.
-%% Suricata supports content modifiers for tcp and http.
-\Luc{Eh possivel utilizar os modificadores em qualquer protocolo.
-Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo.
-O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós.
-Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso.
-Se necessário posso explicar.} Given that
-%% content options often use modifiers\Mar{assumindo que eh
-%%   verdade. confirmar estat. solicitada acima...},
-
-%\tname\ uses the
-%frequency of content modifiers to discriminate rules that are more or
-%less likely to reflect actual rules.
-
-\tname{} builds on our observation that these modifiers are commonly
-used to define existing rules. More precisely, \tname\ builds on the
-observation that the strings accessible in positions associated with
-content modifiers (\eg{}, GET, Host, Connection) are more common. For
-example, historical data indicates that content options that use the
-modifier \CodeIn{http\_connection} are much less likely than content
-options that use the modifier
-\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
-shows the histogram of content modifiers for http. (We removed the
-prefix ``http'' on the modifier for space.) Although content modifiers
+\tname\ builds on the observation that the strings commonly accessed
+at positions associated with content modifiers (\eg{}, host,
+connection) are more common. For example, historical data indicates
+that content options that use the modifier \CodeIn{connection} are
+much less likely than content options that use the modifier
+\CodeIn{uri}. Figure~\ref{fig:distribution-content_modifiers} shows
+the histogram of content modifiers for http. (We removed the prefix
+\CodeIn{http} on the modifier for space.) Although content modifiers
 are applicable to rules of any protocol supported by Suricata,
-currently, \tname\ only supports this features for tcp and http. But
-these rules comprise $\sim$80\% of the rules of our data set.
-
-%% and
-%% (2) the payload of messages of different protocols are not as
-%% segmented as the case of tcp and http. Consequently, rule writers do
-%% not refer to specific parts of the payload as often as the task o and
-%% more challenging for
-
-We compute the heuristic function for content modifiers with the
-following formula:
+currently, \tname\ only supports this features for tcp and http; these
+rules comprise $\sim$80\% of the rules of our data set. We compute the
+heuristic function for content modifiers with the following formula:
 
@@
 \noindent
-, where $N$ is the number of contents options present in a rule,
-$g(i)$ is the contribution of content option $i$ to the score,
+, where $g(i)$ is the contribution of content option $i$ to the score,
+$N$ is the number of contents options present in a rule,
 $\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
@@
 
-%% \Mar{Qual o rationale para considerar as varias
-%%   alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}0
-%%   \Luc{Vc está supondo que as regras que geramos incluem os modificadores e que deveriamos pontuar somente de acordo com os que estão presentes?
-Não eh o caso.
-A ideia aqui basicamente eh: se um certo content do payload está numa posição q eh muito verificada nas regras, possivelmente esse content eh mais relevante do q outro q está numa posição q raramente eh verificada.
-Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no data set.}
-%%   \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content.
-Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores que conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}
-
-%% \Luc{$M_i$ eh o numero de modificadores que podem referenciar a posição em q se encontra o $content_i$}\Mar{isto eh confuso. vc. esta dizendo que
-%%   o mesmo content pode ter mais de um modificador? ou a string poderia
-%%   ser acessada por mais de um modificador. se for este o caso,
-%%   sinceramente, achei isto bem confuso. afinal, a regra jah indica
-%%   qual o modificador foi escolhido para acessar a string.}.\Luc{Cada content em uma regra só pode possuir um modificador para indicar a posição onde ele deve ser procurado.
-Entretanto sim, uma "string poderia
-%%   ser acessada por mais de um modificador", foi o caso daquele exemplo com o http\_header.} \Luc{M eh
-%%   o numero de modificadores associados a cada content. Temos
-%%   content[1], content[2], ..., content[N]. Cada um desses contents tem
-%%   modificadores associados content[i][1], content[i][2], ...,
-%%   content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
-%%   content. a frequencia seria 0, correto?} \Luc{Sim}
-
-%\Mar{what about ``none'' (no modifier)?}}
-%  \Luc{Uma regra pode sim ter contents sem modificadores, mas isso n
-    %% vem ao caso aqui. Nessa heuristica toda string do payload vai ter
-    %% associada a ela os modificadores q podem acessar a sua
-    %% posição. Portanto, não vai existir o caso de não haver
-    %% modificador. Atenção: isso não quer dizer q as regras q geramos
-%% incluem os modificadores, elas n incluem.
-%\Luc{Inclui esse modificador pq ele tambem pode acessar o campo Request URI.
-Isso eh um problema?
-O http\_uri eh o modificador mais comum, mas o fato de haver o modificador http\_request\_line que tambem accessa a mesma posição faz com que o fitness maximo seja 0.5, pois a ocorrencia de http\_request\_line eh irrelevante no calculo.}
-
-To illustrate this function, consider a rule containing the option
-\CodeIn{content: ``/HtmlAdaptor''}. The contribution of this option to
+To illustrate, consider a rule containing the option
+\CodeIn{content:``/HtmlAdaptor''} (see Section~\ref{sec:jboss}). The
+contribution of this option---\ie{}, the value of $\mathit{g(i)}$---to
 the score of the rule is $0.5$ (=$\frac{10,111+11}{10,111*2}$). The
-value of $M_i$ for this case is 2 as the string ``/HtmlAdaptor'' could
-be accessed through the modifiers \CodeIn{http\_uri} and
-\CodeIn{http\_request\_line} as that string is stored on the
-\CodeIn{Request URI} field of the http message\Mar{how do we know
-  that?}. The relatively high heuristic value reflects the fact that
-the string ``/HtmlAdaptor'', stored at \CodeIn{Request URI}, could be
-accessed by two content modifiers, namely (\CodeIn{http\_uri} and
-\CodeIn{http\_request\_line}).  For the option \CodeIn{content:
-  ``keep-alive''}, either the modifier \CodeIn{http\_header} or the
-modifier \CodeIn{http\_connection} could be used to locate the string
-``keep-alive'' stored in the field \CodeIn{Connection} (see
-Figure~\ref{fig:http-header-example}). Consequently, for this case,
-the contribution of modifiers to this content option would be $0.2$
-(=$\frac{4089+29}{10111*2}$). Assuming that both options are used in a
-rule, the aggregated contribution of this heuristic function would be
-0.35 (=(0.2+0.5)/2).
+value of $M_i$, in this case, is 2 as the string ``/HtmlAdaptor''
+could be accessed through the modifiers \CodeIn{uri} and
+\CodeIn{request\_line}. The reason is that \tname\ identifies that
+that string appears in the field \CodeIn{Request URI} (of an http
+message from the malicous traffic) and that field is accessible
+through these two modifiers. The relatively high heuristic value
+reflects the fact that the string could be accessed by two popular
+content modifiers.  As another example, consider the option
+\CodeIn{content: ``keep-alive''}. In this case, either the modifier
+\CodeIn{header} or the modifier \CodeIn{connection} could be used to
+locate the string ``keep-alive'', stored in the field
+\CodeIn{Connection} from an http message (see
+Figure~\ref{fig:http-header-example}). Consequently the contribution
+of this heuristic function to the corresponding rule would be $0.2$
+(=$\frac{4089+29}{10111*2}$), in this case. Assuming that a rule uses
+the two content options mentioned aboved, the aggregated contribution
+of this heuristic function to the rule would be 0.35 (=(0.2+0.5)/2).
 
@@
 O total de regras criadas é igual ao da Tabela 3.}
-
-%% \section{Modificadores}É possivel deixar que uma regra verifique todo o payload do pacote em busca de um match com um content ou especificar partes do payload onde esse content deve ser procurado.
-Essa definição é feita utilizando os chamados modificadores de content, eles adicionam mais detalhes/informações sobre a detecção de um determinado content.
-Os modificadores que estamos considerando são validos apenas para regras dos protocolos tcp e http.
-
-%% Exemplo do payload "cru" de um pacote http:\\
-%% Aqui, todas as informações do pacote estão juntas.
-
-%% \begin{figure}[h!]
-%%   \lstinputlisting[language=C,numbers=none,keywords={content}]{http-example.payload}
-%% \end{figure}
-
-%% O mesmo payload interpretado pelo Wireshark:\\
-%% Aqui, o wireshark interpreta as informações dos pacotes e as divide para melhor visualização e entendimento.
-
-%% \begin{figure}[H]
-%%   \lstinputlisting[language=C,numbers=none,keywords={content}]{http-example-wireshark.payload}
-%% \end{figure}
-
-%% O que o suricata faz eh identificar os campos e seus valores e com isso permite que as regras façam referencia direta a eles.
-Por exemplo, na figura do payload "cru" não conseguimos identificar facilmente oq eh o "Request URI Path:", já na segunda figura isso está explícito.
-
-%% Eu disse que o "GET" estava associado tambem ao http\_header, mas não está, me enganei.
-Mas sim, pode haver interseção entre os campos que dois ou mais modificadores referenciam.
-O seguinte trecho do payload é acessado pelo modificador http\_header:
-
-%% \begin{figure}[H]
-%%   \lstinputlisting[language=C,numbers=none,keywords={content}]{http-header.payload}
-%% \end{figure}
-
-%% Entretando, tambem existem os modificadores http\_host,
-%% http\_accept\_enc, http\_connection e http\_accept que permitem
-%% acessar diretamente cada uma das linhas acima.
-
-%% \Mar{Lucas, use uma regra concreta com a regra da Secao 3.3 (assumindo
-%% que vai mante-la) e mostre como eh feito o calculo da regra desta
-%% funcao heuristica para regra inteira.}\\
-%% \Luc{-------------}
-
-%% A avaliação que fazemos aqui pontua um dado content de acordo com quantas vezes o campo em que ele se encontra eh referenciado nas regras.
-Por exemplo, vamos supor que o tokenizer produziu a opção content: "/HtmlAdaptor".
-Essa string pertence ao campo "Request URI:".
-Esse campo pode ser acessado pelo modificador http\_uri.
-O tokenizer associa o content a esse modificador.
-%% Posteriormente, esse content vai receber uma pontuação
-%% correspondente a popularidade desse modificador:\\
-
-%% A multiplicação por 1 no denominador eh referente a quantidade de
-%% modificadores que estão associadas a esse content.\Luc{O correto eh ser dividido pela quantidade de modificadores.
-Atualizei a formula para tentar representar isso.
-Considere que no exemplo acima o content pudesse ser acessado por mais de um modificador.
-A pontuação dele então seria $\frac{10,111+x}{10,111*1}$ > 1.
-Para manter a pontuação entre 0-1, o certo nesse caso seria fazer $\frac{10,111+x}{10,111*2}$}\Mar{nao foi isto
-%%   que vc. disse acima, Lucas. foi dito que N (=2) era o numero de
-%%   contents e o exemplo acima tem apenas 1 content.}
-
-%% Content options are
-%% associated with specific packets fields\Mar{or modifiers?}.\Mar{why
-%%   GET is associated with http\_method and http\_header?}\Mar{are all
-%%   content associated with one or many fields or positions in a packet?
-%%   can you give examples other than GET to explain this? if yes,
-%%   shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
-%%   is important to justify (to explain rationale)---how is it different
-%%   from counting rarity?}
-
-%% Esse componente pontua um content de acordo com o campo ao qual ele pertence no pacote.
-Contents que pertencem a campos mais comuns (campos que são mais referenciados nas regras) recebem maior pontuação.
-%% Modificadores diferentes podem acessar campos iguais em um pacote. Por
-%% exemplo os modificadores http\_method e http\_header, o http\_method
-%% acessa apenas o metodo do pacote http, mas o http\_header acessa todo
-%% o cabeçalho (incluindo o proprio metodo). Com isso, a string "GET"
-%% receberia a pontuação correspondente a esses dois modificadores.
 

### 2020-03-05 paper/paper.tex
balancing columns
pending changes on section 4.3.4

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Learning to Synthesize Rules for Network Intrusion Detectors}
+\title{Learning to Synthesize Rules for Rule-based Intrusion Detectors}
 
@@
 \includegraphics[scale=0.5]{figs/http_header-example.png}
-\caption{\CodeIn{http\_header} example.\Mar{Lucas, simplifique esta
-    figura (reduza o tamanho da imagem). Mostre apenas os casos onde ha um match, por exemplo,
-    com host=www.google.com e connection=keep-alive.}}
+\caption{\CodeIn{http\_header} example.}
 \label{fig:http-header-example}
@@
 
+\balance
 \bibliographystyle{ACM-Reference-Format}

### 2020-03-05 paper/paper.tex
weight optimization

@@
 {\small
-\vspace{-1ex}
+\vspace{-1.5ex}
 \[\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}/N\]
-\vspace{-1ex}
+\vspace{-0.5ex}
 }
@@
 \includegraphics[scale=0.5]{figs/http_header-example.png}
-\caption{\CodeIn{http\_header} example.}
+\vspace{-2ex}
+\caption{Content modifiers.}
 \label{fig:http-header-example}
@@
 
-\subsubsection{Optimization}
+\subsubsection{Weight Optimization}
 \label{sec:weight-optimization}
-
-\Gui{Fizemos um teste para encontrar melhores pesos para cada ataque.
+We optimized the weights of these heuristic functions as follows. We
+ran \tname{} using different combinations of weights on a sample of
+attacks and chose the combination that offered the best overall
+performance, which we computed with the sum of the rankings of the
+golden rule on each attack. The weight combination that led to the
+smaller sum of positions (of the golden rule) in the rakings was
+considered the best choice. We considered \emph{all} combinations of
+the following weights 0, 0.25, 0.5, 0.75, and 1. Given that we have
+four different functions and five possible weight values, we needed to
+run \tname{} on the sample of attacks for 625 (=$5^4$) times. In the
+end of the process, the weight combination $\langle{}x,y,z,w\rangle$
+was the best we found. \Mar{$\leftarrow$indique aqui quais foram os
+  pesos que a gente usar para cada funcao.} Although empirical results
+showed that this strategy results in good performance, we remain to
+evaluate other optimization strategies, such as evolutionary
+alternatives as AVM~\cite{10.1109/32.57624}, to find better weights.
+
+%\Gui{Fizemos um teste para encontrar melhores pesos para cada ataque.
 Calculamos o fitness de cada regra atribuindo diferentes pesos para cada função heurística.

### 2020-03-05 paper/paper.tex
pass on technique

@@
 \noindent\textbf{Overview.}~\tname\ takes on input one or more
-examples of the same attack and produces on output a ranked list of
+examples of an attack and produces on output a ranked list of
 \emph{plausible rules}---\ie{}, rules that isolate the malicious
@@
   \centering
-  \includegraphics[trim=0 395 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
+  \includegraphics[trim=0 340 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
   \vspace{-4ex}  
-  \caption{The \tname\ workflow.\Mar{revisar}}
+  \caption{The \tname\ workflow.}
   \label{fig:overview}
@@
 seed rule is likely overspecified, \ie{}, it may be more restrictive
-than the golden rule. Consequently, the use of that rule could result
-in false negatives. Second, \tname\ generates alternative rules that
-1) weaken the constraints expressed in the seed rule while 2) still
+than necessary. Consequently, the use of that rule could result in
+false negatives. Second, \tname\ generates alternative rules that 1)
+weaken the constraints expressed in the seed rule while 2) still
 capturing the malicious traffic. Given that every option declared in a
@@
 of options from the seed options. However, note that that approach
-alone could result in false positives as it imposes no limit on to
-what extent a the constraints induced from the rule can be
-weakened. \tname{} uses benign traffic for that purpose. More
-precisely, it searches for plausible rules derived from the seed rule
-obtained in the previous step. Finally, \tname{} uses heuristic
-functions obtained from a data set of existing rules to rank the
-generated plausible rules by their similarity to existing
-rules. Similarity is used to approximate user's intent. The following
-sections detail each component of \tname.
+alone could result in false positives as it imposes no limit on the
+extent the constraints induced from the rule can be weakened. \tname{}
+uses benign traffic for that purpose. More precisely, it searches for
+plausible rules derived from the seed rule obtained in the previous
+step. Finally, \tname{} uses heuristic functions, obtained from a data
+set of existing rules, to rank the generated plausible rules by their
+similarity to existing rules. Similarity is used to approximate user's
+intent. The following sections detail each component of \tname.
 

### 2020-03-05 paper/paper.tex
minor changes

@@
   \centering
-  \begin{tabular}{clp{5.5cm}}
+  \begin{tabular}{clp{6cm}}
     \toprule
@@
     7 & flags & matches on TCP flags\\
-    8 & fragbits & checks if the fragmentation or reserved bits are set in the IP header\\
+    8 & fragbits & checks if the frag. or reserved bits are set in the IP header\\
     9 & threshold & minimum threshold for a rule before it generates alerts\\
@@
 \centering
-\includegraphics[scale=0.28]{figs/ICMP-packet-structure.png}
+\includegraphics[scale=0.27]{figs/ICMP-packet-structure.png}
 \vspace{-2ex}
 \caption{Layout of an ICMP packet.}
+\vspace{-2ex}
 \label{fig:icmp-packet-layout}

### 2020-03-05 paper/R/idqaccess/subject.r
idqaccess plot

@@ -3,9 +3,12 @@
 pdf("idqaccess.pdf", width = 4.1, height = 2.9)
 
 df1 <- read.table("data1.txt", header = FALSE)
-df2 <- read.table("data5.txt", header = FALSE)
-df3 <- read.table("data10.txt", header = FALSE)
-df4 <- read.table("data14.txt", header = FALSE)
+df2 <- read.table("data2.txt", header = FALSE)
+df3 <- read.table("data3.txt", header = FALSE)
+df4 <- read.table("data4.txt", header = FALSE)
+df5 <- read.table("data5.txt", header = FALSE)
+df6 <- read.table("data10.txt", header = FALSE)
+df7 <- read.table("data14.txt", header = FALSE)
 
 # adjusting margins
 par(mar = c(5, 4, 4, 0), 
@@ -13,7 +16,7 @@ par(mar = c(5, 4, 4, 0),
     mai = c(0.4, 0.6, 0.1, 0),
     mfrow = c(1,2))
 
-bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,names=c("1","5","10","14"), boxwex=0.2, cex.axis=1, las=1)
+bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,df6$V1,df7$V1,names=c("1","2","3","4","5","10","14"), boxwex=0.2, cex.axis=1, las=1)
 
 dev.off()
 

### 2020-03-06 paper/paper.tex
pass on RQ3

@@
 
-\subsubsection{Popularity of http content modifiers.} As previously
+\subsubsection{\label{sec:popularity-content-modifiers}Popularity of http content modifiers.} As previously
 mentioned, \percRulesWithContent\ of \suri\ rules involve content
@@
 
-This section evaluates the importance of each proposed heuristic
-function. To that end, we evaluated the use of \tname{} without each
-of these functions with the goal of checking how much each of them
+RQ2 evaluates the importance of each proposed heuristic function. To
+answer that question, we evaluated the use of \tname{} without each of
+these functions with the goal of checking how much each of them
 contributes to the overall performance of the technique. Recall that
-the score a function $r$ is given by the weigthed average of the
-values given by each heuristic function, \ie{} the score of $r$ is
-obtained with the aggregate function $f(r)=\sum_{i}^{N}
-w_i*f_i(r)/N;$, where $w_i$ is the weight of heuristic $i$, $f_i(r)$
-is the score of heuristic $i$ on $r$, and $N$ is the number of
-heuristic---four, in our case. To assess the value of each heuristic
-function, we assigned its weight to zero---effectively, disabling the
-function---and compared the result with \tname\ using default weights,
-obtained with the procedure described on
+the score a rule $r$ is given by the weigthed average of the values
+given by each heuristic function, \ie{} the score of $r$ is obtained
+with the function $f(r)=\sum_{i}^{N} w_i*f_i(r)/N$, where $w_i$ is the
+weight of heuristic $i$, $f_i(r)$ is the score of heuristic $i$ on
+$r$, and $N$ is the number of heuristics considered---four, in our
+case. To assess the value of each heuristic function, we assigned its
+weight to zero---effectively, disabling the function---and compared
+the result with \tname\ using default weights, obtained with the
+optimization procedure described on
 Section~\ref{sec:weight-optimization}.
@@
 Table~\ref{table:weights} shows results. Column ``Name'' shows the
-name of the attack and column ``All'' shows the result of \tname\ with
-all heuristic functions enabled; as in column ``Rank (step 3)'' from
-Table~\ref{table:results}. The remaining columns show results obtained
-by running the same experiment as in Section~\ref{sec:answer-rqone},
-but supressing one heuristic function. An up arrow indicates that,
-considering that particular data point, the performance of \tname{} on
-that configuration is worse when compared to \tname's default
-configuration. A down arrow, appearing highlighted in the table,
-indicates the contrary.
-
-\Mar{--------------------------------- estou aqui}
+name of the attack and column ``base'' shows the result of
+\tname\ with all heuristic functions enabled, as in column ``Rank
+(step 3)'' from Table~\ref{table:results}. The remaining columns show
+results obtained by running the same experiment, as on
+Section~\ref{sec:answer-rqone}, but supressing one heuristic
+function. The number corresponds to the order the heuristic function
+is introduced on Section~\ref{sec:ranking}. The symbol of an up arrow
+($\uparrow$) indicates that the performance of \tname{} on that
+configuration is worse when compared to \tname's default
+configuration. A down arrow ($\downarrow$), appearing highlighted in
+the table, indicates the contrary.
 
@@
     Name &
-    All &
+    base &
+    -f1 &
     -f2 &    
     -f3 &
-    -f4 &
-    -f5 \\ 
-    Black Nurse & 2 & 2 & 2 & 2 & 7$\uparrow$ \\    
-    Ping Scan & 2 & 2 & 2 & 2 &  8$\uparrow$\\
+    -f4 \\
+    \midrule
+    Black Nurse & 2 & 7$\uparrow$ & 2 & 2 & 2 \\    
+    Ping Scan & 2 &  8$\uparrow$& 2 & 2 & 2 \\
     SYN Flood & 2 & 2 & 2 & 2 & 2 \\
@@
     %%UDP Flood & - & - & - & - & - \\
-
+    \midrule
     %%Apache Struts 2 & - & - & - & - & - & -\\
-    ColdFusion & 3 & 2\cellcolor{lightgray}$\downarrow$ & 12$\uparrow$ & 5$\uparrow$ & 3 \\
-    iisadmin access & 1 & 1 & 1 & 8$\uparrow$ & 1 \\        
-    JBoss JMX & 45 & 5\cellcolor{lightgray}$\downarrow$ & 45 & 111$\uparrow$ & 45 \\
-    Script tag in URI & 1 & 4$\uparrow$ & 1 & 8$\uparrow$ & 1 \\
-    WeBid LFI & 10 & 77$\uparrow$ & 28$\uparrow$ & 32$\uparrow$ & 10 \\    
+    ColdFusion & 3 & 3 & 2\cellcolor{lightgray}$\downarrow$ & 12$\uparrow$ & 5$\uparrow$ \\
+    iisadmin access & 1  & 1 & 1 & 1 & 8$\uparrow$\\        
+    JBoss JMX & 45 & 45 & 5\cellcolor{lightgray}$\downarrow$ & 45 & 111$\uparrow$ \\
+    Script tag in URI & 1 & 1 & 4$\uparrow$ & 1 & 8$\uparrow$ \\
+    WeBid LFI & 10 & 10 & 77$\uparrow$ & 28$\uparrow$ & 32$\uparrow$ \\    
     %%Wordpress Login & 8 & - & - & - & - \\
-    .htaccess access & 2 & 2 & 2 & 17$\uparrow$ & 2 \\
-    .idq access & 2 & 2 & 2 & 12$\uparrow$ & 2 \\
-    /system32/ in URI & 4 & 20$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 58$\uparrow$ & 4 \\
-    Basebuilder LFI & 34 & 117$\uparrow$ & 57$\uparrow$ & 183$\uparrow$ & 34 \\
-    Tomcat Snoop Access & 32 & 828$\uparrow$ & 2\cellcolor{lightgray}$\downarrow$ & 2564$\uparrow$ & 32 \\
+    .htaccess access & 2  & 2 & 2 & 2 & 17$\uparrow$\\
+    .idq access & 2  & 2 & 2 & 2 & 12$\uparrow$ \\
+    /system32/ in URI & 4  & 4 & 20$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 58$\uparrow$\\
+    Basebuilder LFI & 34  & 34 & 117$\uparrow$ & 57$\uparrow$ & 183$\uparrow$\\
+    Tomcat Snoop Access & 32  & 32 & 828$\uparrow$ & 2\cellcolor{lightgray}$\downarrow$ & 2564$\uparrow$\\
     \bottomrule
@@
 global optimum and cannot avoid those cases. Finally, note that every
-function contributed. Some functions contributed in more cases than
-others, namely \Fix{xxx}, and one of the functions, namely \Fix{xxx},
-only contributed to the performance of \tname{} on attacks involving
-non-http protocols.
+function contributed. Function ``Popularity of content modifiers''
+(Section~\ref{sec:popularity-content-modifiers}) was the function that
+contributed in more cases whereas function ``Popularity of options''
+(see Section~{sec:popularity-options}) only contributed to the
+performance of attacks involving non-http protocols.
 
@@
 \begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
-..........................
+Every heuristic function contributed to the overall performance of
+\tname. The function that builds on the popularity of content
+modifiers was the one that affected positively the higher number of
+cases we analyzed.
 \end{tcolorbox}
@@
 attacks---of the same kind---plays an important role in the
-performance of the technique. To answer this research question, we
-selected the cases with the highest number of examples of attacks (see
-coolumn ``\#~Vars'' on Table~\ref{table:results}), varied the number of
-examples, and measured the position of the golden rule at the
-rank. For each number of variants $N$, we sampled five sets of size
-$N$ to provide as input to the technique.
-
-Table~\ref{fig:impact-number-attacks} shows
-results. \Fix{...elaborate...}
-
-\begin{figure}[t!]
+performance of the technique.
+
+To answer this research question, we selected two cases with the
+highest number of examples (see column ``\#~Vars'' on
+Table~\ref{table:results})---idqaccess and scripttag---and the case
+where \tname\ performs worst---jboss. We measured the position of the
+golden rule at the rank for an increasing number of of examples
+provided to \tname. For each number of variants, we sampled five sets
+of that size to provide as input to the
+technique. Figure~\ref{fig:impact-number-attacks} shows the results
+for each attack we selected. The y-axis of each plot denotes the rank
+of the golden rule and x-axis denotes the number of attack examples
+used to synthesize rules. Each point in the x-axis is associated with
+a boxplot for the distribution of results for the corresponding number
+of variants. Note that 1) the number of examples is important, 2) the
+choice of examples is important, and 3) a small number of examples
+suffices to reach the best results.
+
+\begin{figure}[h!]
   \centering

### 2020-03-06 paper/paper.tex
fix synflood results

@@
     Ping Scan & 6 & 5 & 9 & 2 \\
-    SYN Flood & 4 & 2 & 2 & 2 \\
+    SYN Flood & 4 & 2 & 2 & 1 \\
     UDP Teardrop & 1 & 4 & 8 & 4 \\
@@
     Ping Scan & 2 &  8$\uparrow$& 2 & 2 & 2 \\
-    SYN Flood & 2 & 2 & 2 & 2 & 2 \\
+    SYN Flood & 1 & 1 & 1 & 1 & 1 \\
     UDP Teardrop & 4 & 4 & 4 & 4 & 4 \\
@@
     \vspace{-2ex}
-    \lstinputlisting[language=C,numbers=none,frame=none,keywords={window,flags}]{top-synflood.rules}
+    \lstinputlisting[language=C,numbers=none,frame=none,keywords={window,flags, threshold}]{top-synflood.rules}
     \vspace{-3ex}  
@@
   geradas. enfim, isto esta inconsistente com nosso discurso.}
+  \Luc{Coloquei o threshold em destaque tambem. O window nas duas regras foi um erro, corrigi.}
 

### 2020-03-06 paper/paper.tex
snapshot

@@
 
-\newcommand{\numNonContentAttacks}{5}
+\newcommand{\numNonContentAttacks}{4}
 \newcommand{\numContentAttacks}{10}
@@
 of attacks we analyzed.  Non-http attacks are listed at the top of the
-table whereas http attacks are listed at the bottom. Recall from
+table and http attacks are listed at the bottom. Recall from
 Section~\ref{sec:reverse-engineering} that (1) http attacks are
 prevalent, (2) most http attacks use only content options, and (3)
-most non-http attacks only use field options. We selected from various
-sources \numNonContentAttacks\ non-http attacks and
-\numContentAttacks\ http attacks. The higher number of http attacks
-reflects their popularity. Column ``Name'' shows the name of the
-attack, column ``Source'' shows the source where we obtained data to
-reproduce the attack (see Section~\ref{subsec:malicious-traffic}),
-column ``Type'' shows the kind of attack, column ``Protocol'' shows
-the name of the protocol explored in the attack, and column ``Ground
-Truth'' shows the rule options to capture the attack. The rationale
-for the selection were (1) to cover various kinds of attacks and (2)
-to cover attacks of various protocols. Considering (1), the selection
-includes attacks of the following types: Denial-of-Service
-(DoS)~\cite{denial-of-service}, Active Reconnaissance
-(AREC)~\cite{active-reconnaissance}, Privilege escalation
-(PESC)~\cite{privilege-escalation}, Remote Code Execution
-(RCE)~\cite{remote-code-execution}, Local File Inclusion
-(LFI)~\cite{local-file-inclusion}, and Policy Violation
-(PVE)~\Fix{local-file-inclusion}.
+most non-http attacks only use field options. We selected
+\numNonContentAttacks\ non-http attacks and \numContentAttacks\ http
+attacks. The higher number of http attacks reflects their
+popularity. Column ``Name'' shows the name of the attack, column
+``Source'' shows the source where we obtained data to reproduce the
+attack (see Section~\ref{subsec:malicious-traffic}), column ``Type''
+shows the kind of attack, column ``Protocol'' shows the name of the
+protocol explored in the attack, and column ``Ground Truth'' shows the
+rule options to capture the attack. The rationale for the selection
+were (1) to cover various kinds of attacks and (2) to cover attacks of
+various protocols. Considering (1), the selection includes attacks of
+the following types: Denial-of-Service (DoS)~\cite{denial-of-service},
+Active Reconnaissance (AREC)~\cite{active-reconnaissance}, Privilege
+escalation (PESC)~\cite{privilege-escalation}, Remote Code Execution
+(RCE)~\cite{remote-code-execution}, and Local File Inclusion
+(LFI)~\cite{local-file-inclusion}.
+
+%, and Policy Violation (PVE).
 
@@
 used to synthesize rules. Each point in the x-axis is associated with
-a boxplot for the distribution of results for the corresponding number
-of variants. Note that 1) the number of examples is important, 2) the
-choice of examples is important, and 3) a small number of examples
-suffices to reach the best results.
+the boxplot of the distribution of results for the corresponding
+number of variants.
 
@@
 
-\Mar{coloquem os numeros aqui para que eu gere o boxplot amanha,
-  depois da aula.
-
-  *** posicao no ranking (numeros artificais. por favor, substituam) ***
-
-  ataque 1:
-
-    1 variacao:  10, 20, 4, 2, 5
-
-    5 variacoes: 5, 4, 4, 2, 3
-
-    10 variacoes: 1, 2, 1, 2, 1
-
-    (ate numero de variacoes que aparece na Tabela 3)
-
-  ataque 2:
-     ...
-
-}
+Results indicate that 1) the number of examples is important, 2) the
+choice of examples is important, and 3) a small number of examples
+suffices to reach the best results.
+
+\begin{center}
+\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
+As expected, \tname\ is sensitive to the number and choice of example
+input attacks. However, a small number of examples suffices to achieve
+best results.
+\end{tcolorbox}
+\end{center}
 
@@
 
-External Validity. \Fix{elaborate ...how popular are
-  Suricata/Metasploit?}
+Our experiments is affected by the usual threats. Considering external
+validity. Our results cannot be generalized beyond the subjects we
+selected. To mitigate that threat we selected a variety of attacks
+with different characteristics. Considering internal validity, the
+implementation of \tname\ could contain errors. We mitigated that
+threat by carefully inspecting code and results. Considering construct
+validity, the choice of metric used to evaluate the approach may not
+be applicable in practice. We considering the use of ranking of the
+golden rule a reasonable choice as the approach places the golden rule
+close to the top in most cases.
 

### 2020-03-06 paper/paper.tex
pass

@@
 \nids\ are restricted to a fixed set of rules defined by the network
-system administrator.}\Mar{check...}
+system administrator.}
 
@@
   alert them. However, IDS and IPS do not differ in the approaches
-  used to detect the intrusion~\cite{ids-ips}.\Mar{check}}. The
+  used to detect the intrusion~\cite{ids-ips}.}. The
 \emph{header} comes after the action in the rule description. It
@@
 \tname{} produced a total of 23 tokens\Mar{@Lucas, isto estah certo?
-  Se sim, como isto cai para 5 (pela Tabela 3)}, considered content
+  Se sim, como isto cai para 5 (pela Tabela 3)} \Luc{6.1.1 eh o ataque 
+JBoss. Na tabela constam 23 tokens.}, considered content
 option candidates.
@@
 was the best we found. \Mar{$\leftarrow$indique aqui quais foram os
-  pesos que a gente usar para cada funcao.} Although empirical results
+  pesos que a gente usar para cada funcao.} \Luc{f1: 1, f2: 0.25, f3: 0.25, f4: 1} Although empirical results
 showed that this strategy results in good performance, we remain to

### 2020-03-06 paper/paper.tex
pass on RQs

@@
 often associated with packet fields. Let us consider a rule for an
-ICMP attack as example. Figure~\ref{fig:icmp-packet-layout} shows the
-layout of an ICMP packet. The packet fields ``Flags'' and ``Time to
-Live'', declared in the IP header, are mapped to the \suri\ options
+icmp attack as example. Figure~\ref{fig:icmp-packet-layout} shows the
+layout of an icmp packet. The packet fields ``Flags'' and ``Time to
+Live'', declared in the ip header, are mapped to the \suri\ options
 \CodeIn{flagbits} and \CodeIn{ttl}, respectively. Likewise, the fields
-``Type'' and ``Code'', declared in the ICMP header, are mapped to the
+``Type'' and ``Code'', declared in the icmp header, are mapped to the
 options \CodeIn{itype} and \CodeIn{icode}, respectively. The values of
 the options ``ttl'', ``itype'', and ``icode'' are the same as the
-value of the corresponding fields in the ICMP packet. For the field
+value of the corresponding fields in the icmp packet. For the field
 ``Flags'' the value of the corresponding option is encoded in a
@@
 \vspace{-2ex}
-\caption{Layout of an ICMP packet.}
+\caption{Layout of an icmp packet.}
 \vspace{-2ex}
@@
 
-\setlength{\tabcolsep}{5pt}
+%\setlength{\tabcolsep}{5pt}
+\renewcommand{\arraystretch}{0.9}
 \begin{table*}[t!]
@@
     appear at the top; http attacks appear at the bottom.}
+  \vspace{-2ex}
   \centering
-  \begin{tabular}{llllp{8cm}}
+  \begin{tabular}{lcccp{10cm}}
     \toprule
@@
     \midrule
-    Black Nurse & \cite{pcap-attacks} & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3; \textbf{threshold}: type both, track by\_dst, count 250, seconds 1;\\    
-    Ping Scan & \cite{netmap} & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
-    SYN Flood & \cite{hping3} & DoS & TCP & \textbf{flags}: S,12;
+    Black Nurse & \cite{pcap-attacks} & DoS & icmp & \textbf{icode}:3; \textbf{itype}:3; \textbf{threshold}: type both, track by\_dst, count 250, seconds 1;\\    
+    Ping Scan & \cite{netmap} & AREC & icmp & \textbf{dsize}:0; \textbf{itype}:8; \\
+    SYN Flood & \cite{hping3} & DoS & tcp & \textbf{flags}: S,12;
     \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\
-    UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
+    UDP Teardrop & \cite{udp-teardrop-source} & DoS & udp & \textbf{fragbits}:M; \textbf{id}:124; \\
     %%UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5; \\
     \midrule
-    %%Apache Struts 2 & \cite{wrccdc} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
-    ColdFusion \MyComment{PESC} & \cite{nikto} & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
-    iisadmin access\MyComment{web-application-attack} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
-    JBoss JMX\MyComment{ Console Beanshell Exploit} & \cite{nikto} & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
-    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
-    WeBid LFI \MyComment{Local File Inclusion} & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
-    %%WordPress Login & \cite{wrccdc} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
-    .htaccess access\MyComment{attempted-recon} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
-    .idq access\MyComment{EXPLOIT ISAPI} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
-    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
-    Basebuilder LFI & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET"; http\_method; \textbf{content}:"/main.inc.php?"; nocase; http\_uri; \textbf{content}:"mj\_config[src\_path]="; nocase; http\_uri; \\
-    Tomcat server snoop access & \cite{nikto} & AREC & HTTP & \textbf{content}:"/jsp/snp/"; http\_uri; \textbf{content}:".snp"; http\_uri;\\  
+    %%Apache Struts 2 & \cite{wrccdc} & RCE & http & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
+    ColdFusion \MyComment{PESC} & \cite{nikto} & PESC & http \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
+    iisadmin access\MyComment{web-application-attack} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
+    JBoss JMX\MyComment{ Console Beanshell Exploit} & \cite{nikto} & RCE & http \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
+    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
+    WeBid LFI \MyComment{Local File Inclusion} & \cite{nikto} & LFI & http & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
+    %%WordPress Login & \cite{wrccdc} & PVE & http & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
+    .htaccess access\MyComment{attempted-recon} & \cite{nikto} & AREC & http \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
+    .idq access\MyComment{EXPLOIT ISAPI} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
+    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & http \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
+    Basebuilder LFI & \cite{nikto} & LFI & http & \textbf{content}:"GET"; http\_method; \textbf{content}:"/main.inc.php?"; nocase; http\_uri; \textbf{content}:"mj\_config[src\_path]="; nocase; http\_uri; \\
+    Tomcat server snoop access & \cite{nikto} & AREC & http & \textbf{content}:"/jsp/snp/"; http\_uri; \textbf{content}:".snp"; http\_uri;\\  
     \bottomrule
@@
 
+\renewcommand{\arraystretch}{0.9}
 \setlength{\tabcolsep}{2pt}
@@
     Script tag in URI & 23 & 22 & 786 & 1 \\
-    WeBid LFI & 1 & 23 & 11100 & 10\\    
+    WeBid LFI & 1 & 23 & 11,100 & 10\\    
     %%Wordpress Login & 8 & 59 & - & - \\
@@
     /system32/ in URI & 4 & 20 & 1,913 & 4 \\
-    Basebuilder LFI & 1 & 22 & 18133 & 34 \\
-    Tomcat Snoop Access & 1 & 24 & 29830 & 32 \\
+    Basebuilder LFI & 1 & 22 & 18,133 & 34 \\
+    Tomcat Snoop Access & 1 & 24 & 29,830 & 32 \\
     \bottomrule
@@
 
-%% Note from Figure~\ref{fig:overview} that the positive traffic is an
-%% input to the technique. The negative non-malicious traffic, however,
-%% is not.
-
-%subsection{Methodology}
-
-%% \Gui{
-%% O procedimento que já tínhamos representa a etapa de treinamento.
-Toda aquela parte que foi explicada na apresentação para os outros professores e descrito no paper.
-Descreverei novamente pra confirmar, e também foi adicionado uma pequena etapa.
-
-%% TREINAMENTO
-
-%% ->Syrius recebe como input um pcap contendo o ataque (no caso de ataques com content, usamos ataques do Nikto), um pcap benígno (usamos o bigflows.pcap) e um pcap contendo as variações de ataque (isso pode ser otimizado unificando com o pcap de ataque, mas por agora está assim)
-%% ->É gerado a regra máxima através do tokenizer e engenharia reversa.
-%% ->Syrius remove UM content de uma regra da lista de regras atuais (inicialmente apenas a regra máxima) e checa se a regra criada já existe na lista.
-Se não, a regra é adicionada à lista de regras-não-testadas.
-Syrius faz isso duas vezes por regra, para todas as regras da lista.
-%% ->A lista de regras-não-testadas é inicializada no suricata e então o rodamos com o pcap benígno de treinamento (75porcento do total).
-Verificando no arquivo de log do suricata, caso a regra alerte, ela é removida da lista.
-As regras restantes vão para a lista de regras atuais.
-%% ->Quando não houver alterações, retornamos a lista com todas as regras que deram certo.
-%% ->Opcionalmente, ainda rodamos no final uma variação de ataque (no momento apenas um pacote devido ao número limitado que temos do ataque).
-Ao contrário do método com pacotes benígnos, se a regra NÃO alertar, ela é removida.
-%% ->Calculamos o fitness de cada regra e ordenamos de acordo com o fitness resultante. Esta lista ordenada é o fim do treinamento.
-
-%% TESTE
-
-%% ->Inicializamos a lista de regras ordenadas no suricata e rodamos o pcap benígno de teste (25porcento restantes) para cálculo do precision.
-O cálculo ficou basicamente, por regra: [100 - (numero-de-alertas/total-de-pacotes)]
-
-%% ->Inicializamos a lista de regras ordenadas no suricata e rodamos o pcap com variações de ataque (caso a etapa opcional ocorrer, não usamos aquele pacote especifico) para cálculo do recall.
-O cálculo ficou basicamente, por regra: [(100/total-de-pacotes) * n-alertas].
-
-%% ->Por final calculamos o F1 Score por regra: [2*((precision*recall)/(precision+recall))]
-
-%% ->O teste retorna um csv contendo a lista de regras ordenadas com seus respectivos precision, recall e f1 score.
-%% }
-
-%% \Gui{We are using the bigflows.pcap provided by tcpreplay on http://tcpreplay.appneta.com/wiki/captures.html "This is a capture of real network traffic on a busy private network’s access point to the Internet"
-%%  }
-
-%% \Luc{entrada: utilizamos a ferramenta hping3 para executar o ataque em um alvo arbitrario dentro da rede (n ha necessidade de haver um alvo real para executar o hping3) e capturamos o trafego gerado com o wireshark.
-o comando utilizado foi:\\*
-%% \texttt{\# hping3 -d 80 -w 64 -S -p 80 --flood --rand-source 192.138.1.115} \\*
-%% que envia pacotes de 80 bytes (-d 80), com a flag SYN habilitada (-S), tamanho de janela de 64 bytes (-w 64), direcionados a porta 80 (-p 80).
-A flag --flood indica q o hping3 vai enviar os pacotes o mais rápido possível.
-a flag --rand-source indica que cada pacote vai ter um ip de origem aleatorio.
-192.138.1.115 é o alvo.\\*
-%% O trafego armazenado contem 40 mil pacotes SYN, enviados no intervalo de 1 segundo,  pegamos uma amostra de 20 desses pacotes para representar o flood e servir como entrada.\\*
-%% O wireshark armazena os pacotes no formato pcap, utilizamos a ferramenta tcpreplay para reproduzir o trafego com o comando:\\*
-%% \texttt{\# tcpreplay --pps 200 --intf1=wlp2s0 Datasets/syn-flood-sample.pcap}\\*
-%% Nesse caso, estamos enviando os pacotes armazenados em \texttt{Datasets/syn-flood-sample.pcap} a uma taxa de 200 pacotes por segundo (\texttt{-pps 200}) utilizando a interface wlp2s0 (\texttt{--intf1 wlp2s0})\\*
-%% Há diferenças entre os pacotes na opção ack.\\* 
-%% Por que 20?
-Apenas para gerar a regra mais rapidamente, poderia ser qualquer valor maior ou menor.
-No final, o parâmetro \texttt{count} da opção \texttt{threshold} vai ser alterado pelo adm da rede.\\*
-%% Intervalo de reprodução: 0.005 segundo
-
-%% }
-
-%% \Fix{any other detail we should consider?}
-
 \subsection{Answering RQ1: \textRQone}
@@
 
-This section reports on the performance evaluation of \tname. The
-evaluation metric used is the position of the golden rule at the
-ranked list of plausible rules produced by the tool implementing the
-technique (see Figure~\ref{fig:overview}). This metric is a proxy of
-human performance in analyzing the list of plausible rules that
-capture only the malicious traffic provided on input.
+RQ1 addresses the performance of \tname. The metric we use is the
+position of the golden rule at the ranked list of plausible rules
+reported by \tname\ (see Figure~\ref{fig:overview}). This metric is a
+proxy of human performance in analyzing the list of plausible rules.
 
@@
 Rules (step 2)'' shows the number of plausible rules reported by
-\tname\ at step 2, column ``Rank (step 3)'' shows the position of the
-golden rule in the ranked list reported at step 3 of the technique.
-It is worth to indicate that column ``\#~Vars.'' shows the input of
+\tname\ at step 2, and column ``Rank (step 3)'' shows the position of the
+golden rule in the ranked list reported at step 3.
+It is important to indicate that column ``\#~Vars.'' shows the input of
 \tname\ that varies across different attacks. Although \tname\ also
@@
 
-\newcommand{\percTopFiveRanking}{\Fix{X\%}}
+\newcommand{\percTopFiveRanking}{66.7\% (=10/15)}
 
@@
 reported that rule within the top five rules of the ranking in
-\percTopFiveRanking\ of the cases. It is important to note that the
-number of options in the ground truth (Table~\ref{table:attacks},
-column ``Ground Truth'') is not an indication of difficulty to
-synthesize rules. Note that in \Fix{most?} of the cases where the
-ground truth rule contains only one content options the number of
-option candidates and plausible rules produced is high.
+\percTopFiveRanking\ of the cases.
+
+%% It is important to note that the
+%% number of options in the ground truth (Table~\ref{table:attacks},
+%% column ``Ground Truth'') is not an indication of difficulty to
+%% synthesize rules. In facto, note that in  of the cases where the
+%% ground truth rule contains only one content options the number of
+%% option candidates and plausible rules produced is high.
 
@@
   \end{subfigure}%
+\vspace{-2ex}  
 \caption{Impact of number of positive examples.}
 \label{fig:impact-number-attacks}
+\vspace{-2ex}
 \end{figure}
@@
 
-The SYN Flood attack is a denial-of-service attack that exploits a vulnerability in the TCP/IP handshake
-to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
+The SYN Flood attack is a denial-of-service attack that exploits a
+vulnerability in the tcp/ip handshake
+to establish a tcp connection~\cite{cloudfare-synflood}. The handshake
 works as follows in normal circumstances. First, a client sends a SYN
@@
 rule appear in bold---the option \CodeIn{flags: S,12}, identifies a
-SYN packet in a TCP packet and the option \CodeIn{threshold: type
+SYN packet in a tcp packet and the option \CodeIn{threshold: type
   both, track by\_dst, count 5000, seconds 5} indicates that a high

### 2020-03-06 paper/paper.tex
pass 4.2

@@
   explique. me pareceu que o que vc. precisa ate aqui jah estaria no
-  log do \suri}, 2)~replaying the negative benign traffic (see
+  log do \suri} \Luc{Nao precisa desse ponto.
+Para essa etapa foi preciso modificar o codigo do suricata para q ele fosse capaz de gerar mais de 15 alertas para um unico pacote.
+Durante a execução do Syrius pode ser gerada uma grande quantidade de regras q causam falsos positivos, eventualmente podem haver 15 delas q detectam erroneamente um mesmo pacote.
+Sem essa alteração, algumas regras q geram falso positivo n seriam eliminadas.}, 2)~replaying the negative benign traffic (see
 Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to

### 2020-03-06 paper/paper.tex
pass on pending fixes

--- 2020-03-06 paper/paper.tex
@@
 options, \tname{} splits the payload in tokens using these delimiters.
-Considering the example from Section~\ref{sec:content-example},
-\tname{} produced a total of 23 tokens\Mar{@Lucas, isto estah certo?
-  Se sim, como isto cai para 5 (pela Tabela 3)} \Luc{6.1.1 eh o ataque 
-JBoss. Na tabela constam 23 tokens.}, considered content
-option candidates.
+
+%Considering the example from Section~\ref{sec:active-recon},
+%\tname{} produced a total of 5 tokens.
 
@@
 engineering strategy to use based on the protocol of the input
-(malicious) messages was effective. We used the payload strategy for
-http and the field strategy for the other protocols. For flooding
-attacks, \tname\ runs the same approach for each message and
+(malicious) messages showed to be effective. We used the payload
+strategy for http and the field strategy for the other protocols. For
+flooding attacks, \tname\ runs the same approach for each message and
 identifies common options across all messages. For that case,
@@
 \textbf{Implementation details.} We implemented function
-\emph{isPlausible} by 1)~updating Suricata to only monitor the rule
-that is passed as parameter\Mar{isto realmente eh necessario? se sim,
-  explique. me pareceu que o que vc. precisa ate aqui jah estaria no
-  log do \suri}, 2)~replaying the negative benign traffic (see
+\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
+rule, 2)~replaying the negative benign traffic (see
 Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
-check if any benign message was captured by the rule. \Mar{discuss
-  filter function.} It is worth
-noting that the benign traffic is stored in \emph{pcap}
-files~\cite{pcap} and Suricata can process those files very
-efficiently compared to analyzing network messages directly. Instead
-of sending traffic to a machine with a deployed server (\eg{}, web
-server) that \suri\ would monitor, our setup calls \suri\ directly
-through an API. \suri\ is invoked multiple times in the process of
-minimizing the rule.  \textbf{Limitations.} It is worth noting that
-the performance of \tname{} depends on the quality of inputs. For the
-positive malicious traffic, \tname{} only requires one sample (as
-input to step one), but can benefit of several samples (see
-line~\ref{filtering} of Algorithm~\ref{algo1}). The same applies to
-the negative traffic. However, in this case, \tname{} can use a large
-variety of public data sets available online (see
-Section~\ref{sec:objectanalysis-benign}).
+check if any benign message was captured by the rule. The function
+\emph{filter} is implemented similarly. It returns a projection of the
+set $\mathcal O$ where every rule captures the attacks characterized
+by $\mathit{mTr}$. It is worth noting that Suricata can process
+\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
+network messages directly. Instead of sending traffic to a machine
+with a deployed server (\eg{}, web server) that \suri\ would monitor,
+our setup calls \suri\ directly through an API. \suri\ is invoked
+multiple times for creating rules.
+
+%% \textbf{Limitations.} It is worth
+%% noting that the performance of \tname{} depends on the quality of
+%% inputs. For the positive malicious traffic, \tname{} only requires one
+%% sample (as input to step one), but can benefit of several samples (see
+%% line~\ref{filtering} of Algorithm~\ref{algo1}). The same applies to
+%% the negative traffic. However, in this case, \tname{} can use a large
+%% variety of public data sets available online (see
+%% Section~\ref{sec:objectanalysis-benign}).
 
@@
 run \tname{} on the sample of attacks for 625 (=$5^4$) times. In the
-end of the process, the weight combination $\langle{}x,y,z,w\rangle$
-was the best we found. \Mar{$\leftarrow$indique aqui quais foram os
-  pesos que a gente usar para cada funcao.} \Luc{f1: 1, f2: 0.25, f3: 0.25, f4: 1} Although empirical results
-showed that this strategy results in good performance, we remain to
-evaluate other optimization strategies, such as evolutionary
-alternatives as AVM~\cite{10.1109/32.57624}, to find better weights.
+end of the process, the weight combination
+$\langle{}1,0.25,0.25,1\rangle$ was the best choice we found. Although
+results showed that this strategy results in good performance, we
+remain to evaluate other optimization strategies to find better
+weights.
+
+%% , such as evolutionary alternatives as
+%% AVM~\cite{10.1109/32.57624}.
 

### 2020-03-06 paper/paper.tex
pass on discussion

@@
 
-\section{Network Intrusion Detection Systems (\nids)}
+\section{\nids}
 \label{sec:background}
@@
 
-\sloppy \nids{} are typically categorized in two
+\sloppy Network Intrusion Detection Systems (\nids{}) are typically categorized in two
 groups~\cite{kumar2007survey}: rule-based, as briefly described on
@@
 
+%\vspace{-1ex}
 \subsection{Suricata Rules}
@@
 sent to system administrators if the rule pattern is
-satisfied\footnote{Technically, an IPS (Intrusion Prevention System)
-  differs from an IDS in the action taken when an intrusion is
-  observed. IPS systems block malicous traffic whereas IDS system only
-  alert them. However, IDS and IPS do not differ in the approaches
-  used to detect the intrusion~\cite{ids-ips}.}. The
-\emph{header} comes after the action in the rule description. It
-restricts the information flow covered by the rule. The header of this
-rule is \CodeIn{icmp any any -> any any}. It instructs Suricata to
-inspect \CodeIn{icmp} traffic flowing from any machine and port to any
-machine and port.\MyComment{ The variables \CodeIn{\$HOME\_NET} and
-  \CodeIn{\$EXTERNAL\_NET} are configurable.} Finally, \emph{rule
-  options} come after the header. It is a semi-colon-separated
-sequence of key-value pairs. Rule options serve to document the
-analyzed traffic (\eg{}, ``msg'', ``classtype'', and ``sid'') and to
-characterize the attack pattern (\eg, ``flags'', ``threshold'', and
-``content''). \tname{} produces options of the second kind, which are
-\emph{relevant} to capture the attack. For example, the option
-\CodeIn{content:<arg>} checks if the argument is present in the
-payload of a message.
+satisfied\footnote{IPS (Intrusion Prevention System) and IDS differ in
+  the action taken when an intrusion is observed. IPS systems block
+  malicous traffic whereas IDS systems only alert them. However, IDS
+  and IPS are similar in the approaches used to detect
+  intrusions~\cite{ids-ips}.}. The \emph{header} comes after the
+action in the rule description. It restricts the information flow
+covered by the rule. The header of this rule is \CodeIn{icmp any any
+  -> any any}. It instructs Suricata to inspect \CodeIn{icmp} traffic
+flowing from any machine and port to any machine and port.\MyComment{
+  The variables \CodeIn{\$HOME\_NET} and \CodeIn{\$EXTERNAL\_NET} are
+  configurable.} Finally, \emph{rule options} come after the
+header. It is a semi-colon-separated sequence of key-value pairs. Rule
+options serve to document the analyzed traffic (\eg{}, ``msg'',
+``classtype'', and ``sid'') and to characterize the attack pattern
+(\eg, ``flags'', ``threshold'', and ``content''). \tname{} produces
+options of the second kind, which are \emph{relevant} to capture the
+attack. For example, the option \CodeIn{content:<arg>} checks if the
+argument is present in the payload of a message.
 
 \vspace{-1ex}
-\begin{table}[h!]
+\begin{table}[t!]
   \footnotesize
@@
   \end{tabular}
+  \vspace{-3ex}  
 \end{table}
-%\vspace{-1ex}
 
@@
 \newcommand{\textRQone}{How well \tname\ ranks the golden rule?}
-\vspace{0.2cm}
+%\vspace{0.2cm}
 \begin{itemize}[leftmargin=*,label={}]
@@
   to \tname's performance?}
-\vspace{0.2cm}
+%\vspace{0.2cm}
 \begin{itemize}[leftmargin=*,label={}]
@@
   positive examples?}
-\vspace{0.2cm}
+%\vspace{0.2cm}
 \begin{itemize}[leftmargin=*,label={}]
@@
 \tname\ successfully synthesizes the golden rules in all cases and
-that rule is ranked among the first top five rules in
+that rule is ranked among the top five rules in
 \percTopFiveRanking\ of the cases.
@@
 
-\begin{figure}[h!]
+\begin{figure}[t!]
   \centering
@@
   \end{subfigure}%
-\vspace{-2ex}  
+\vspace{-2.5ex}  
 \caption{Impact of number of positive examples.}
 \label{fig:impact-number-attacks}
-\vspace{-2ex}
+\vspace{1ex}
 \end{figure}
@@
 
-\section{Discussion}
-
-\subsection{Sample of other attacks}
-
-\subsubsection{Remote Code Execution (RCE)}
+\section{Discussion of other attacks}
 \label{sec:rce}
@@
 \label{sec:content-example}
-
-%% Adaptacao do texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution
-
-RCE refers to the family of attacks where hackers exploit a network
-vulnerability to run arbitrary code---that could damage the system or
-steal sensitive information---on a remote machine. JBoss~\cite{jboss}
-is popular open source Java application server. This attacks enables
-remote execution of code on the JBoss application server.
-\Mar{@Lucas, por favor, observe o padrao usado nos exemplos
-  anteriores. eh preciso explicar em alto nivel o ataque antes de
-  explicar a regra!}
+\label{sec:dos}
+
+This section elaborates how \tname\ performs on two other attacks.
+
+\textbf{Remote Code Execution (RCE).}  RCE refers to the family of
+attacks where hackers exploit a network vulnerability to run arbitrary
+code on a remote machine. The code to be executed could damage the
+system or steal sensitive information. JBoss~\cite{jboss} is a popular
+open source Java application server. This attack enables remote
+execution of code on the JBoss application server.
+
+%\Mar{@Lucas, por
+%  favor, observe o padrao usado nos exemplos anteriores. eh preciso
+%  explicar em alto nivel o ataque antes de explicar a regra!}
 
@@
   \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
-  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code Execution.}
+  \vspace{-2ex}  
+  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code
+    Execution.}
+  \vspace{-3ex}
 \end{figure}
@@
 \CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
-the matching not be case sensitive and that the string should be
-located at the packet field ``Request URI''. \Mar{Lucas, ``nocase''
-  tambem eh considerado modificador? Isto nao eh confuso? Na secao
-  4.3.4 consideramos modificadores como sinonimo para qualificador de
-  campo, nao?}\Luc{Sim, ``nocase'' define que a busca pela string nao deve ser case sensitive. Em 4.3.4 estamos considerando apenas os modificadores que definem posições/campos de pacotes http.}
-
-\tname{} reproduces the malicious JBoss attack using the traffic from
-a Nikto script~\cite{nikto}. However, similar JBoss attack is also
-available from the Metasploit database~\cite{metasploit}. As in the
-other examples, \tname{} uses, as additional inputs, benign traffic
-from \Fix{...} and a set of existing Suricata rules from
-\Fix{...}. \tname{} initially produces one seed rule containing
-\Fix{xx} options, by analyzing the malicious traffic alone. Then, it
-uses that seed rule to derive \Fix{yy} rules, each rule containing
-subsets of the option in the initial rule. Finally, it ranks that
-list. The list below shows the top-5 rules that \tname{}
-produces.\Mar{<- check/revise}
-
-\Fix{list rules}
-
-Note that the golden rule is in position \Fix{...} on this
-case. \Mar{explain syntactical difference, if exist}
-
-\subsubsection{Denial-of-Service}
-\label{sec:dos}
-
-\Mar{move from above --> We used the following hping3 command to
-  reproduce that attack: \CodeIn{hping3 -d 80 -w 512 -S -p 80 --flood
-    --rand-source <IP>}.}
-
+the matching is not case sensitive and that the string should be
+located at the packet field ``Request URI''. We reproduced the
+malicious JBoss attack using the traffic from a Nikto
+script~\cite{nikto}. Similar attacks are also available on the
+Metasploit database~\cite{metasploit}. This is a case where
+\tname\ performed poorly. Of the 228 plausible rules it generated for
+this attack, the ground truth was located at position 45 in the ranked
+list. The reason for that is that \Fix{...xxx...}
+
+%% \Mar{move from above --> We used the following hping3 command to
+%%   reproduce that attack: \CodeIn{hping3 -d 80 -w 512 -S -p 80 --flood
+%%     --rand-source <IP>}.}
+
+\textbf{Denial-of-Service.}~
 The SYN Flood attack is a denial-of-service attack that exploits a
@@
 volume of such packets should be requested in a short period of
-time. The other options are for documentation purposes.
-
-...........
-In this case, however, that information alone is
-insufficient as the malicious traffic is characterized as a sequence
-of messages. 
-...........
-
-%% \Mar{...and where did they come from?} \Luc{Sim, floods podem ser
-%% compostos por milhares de pacotes, um dos data sets q encontrei
-%% possui mais de 500k pacotes de flood. Os data set  s foram encontrados na internet.}
-
-\begin{figure}[h]
-  \vspace{-2ex}
-  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold,window}]{synflood.suricata.synth}
-%  \caption{Suricata rule for SYN Flood Attacks.}
-  %  \label{fig:synflood-example-synt}
-  \vspace{-2ex}  
-\end{figure}
-
-%% \Mar{@Lucas/Guilheme, há algo o que remover? talvez ``window:64''?
-%%   Qual a correspondencia de ``flags:S,12'' com ``flags:S''? Por favor,
-%%   respondam isto logo.}  \Gui{O window:64 seria removido
-%%   sim.}\Mar{ok. vou explicar isto.}
-  %% \Luc{O segundo campo da opção flags define flags que serão
-  %%   ignoradas. Nesse caso a regra procura por pacotes onde apenas a
-  %%   flag S está 'setada', independentemente do valor das flags 1 e
-  %%   2. Portanto a regra gerada pode capturar coisas diferentes da
-  %%   regra original. A inferencia do "12" n pode ser feita apenas a
-  %%   partir do pacote de entrada, precisariamos tratar
-  %%   isso.}\Mar{vc.
-pode explicar isto **mais detalhadamente** \Luc{a opção flags pode ter dois campos separados por virgula (ex: flags:S,12).
-o primeiro campo indica quais flags devem estar ativas em um pacote para que o suricata gere um alerta.
-se uma ou mais flags forem irrelevantes para a detecção de um determinado ataque (ou seja, se n importa se a flag está ou não ativa), elas serão adicionadas ao segundo campo.} E (1)
-  %%   explicar que isto eh uma limitacao \Luc{n entendi exatamente oq explicar aqui.} e (2) indicar a possivel consequencia
-  %%   disto?
-\Luc{a consequencia sao possiveis falsos negativos.
-ao não incluir flags no segundo campo podemos descartar pacotes por conta da presença de flags q n são relevantes (ou seja, vamos levar em conta flags q n deveriam interferir na detecção).}}
-
-\noindent
-
-.....
-
- The list below shows the rule options for the top-5 rules
-\tname{} produces for this attack.\Mar{algum problema em fazer isto?
-  exemplos concretos ajudam muito!} \Luc{No caso desse ataque, a regra inicial possui apenas duas opcoes (sem contar o threshold).
-Entao existem apenas 3 combinacoes possiveis, dessas 3 apenas 2 sao capazes de isolar o ataque do trafego normal.
-Portanto, o top-2 eh:
-  \begin{figure}[h]
-    \vspace{-2ex}
-    \lstinputlisting[language=C,numbers=none,frame=none,keywords={window,flags, threshold}]{top-synflood.rules}
-    \vspace{-3ex}  
-  \end{figure}
-
-   }\Mar{Lucas, eu nao entendo porque vc. desconsiderou threshold da
-  discussao (afinal, a gente destaca thrshold como relevante no inicio
-  desta secao). tambem nao entendo porque window---que a gente disse
-  acima que *nao* era relevante---aparece nas duas regras
-  geradas. enfim, isto esta inconsistente com nosso discurso.}
-  \Luc{Coloquei o threshold em destaque tambem. O window nas duas regras foi um erro, corrigi.}
-
-\vspace{1ex}
-\noindent
-\textbf{Differences.}~Note that there are differences between the
-options produced in the first step of \tname{} and the options from
-the golden rule, as per Figure~\ref{fig:synflood-example}. The option
-\CodeIn{flags} does not mention parameters \CodeIn{1 2}, there is an
-extra option \CodeIn{window}, and there are differences in the option
-\CodeIn{threshold}. Considering the option \CodeIn{flags}, the first
-parameter of that option indicates what flags in the packet must be
-set while the second parameter indicates non-mandatory flags (\ie{},
-``don't cares''). More specifically, the option \CodeIn{flags: S}
-indicates that the flag ``S'' must be set and all flags not mentioned
-(including ``1'' and ``2'') must be unset for the \nids\ to capture
-the traffic. Consequently, that choice of option alone could result in
-false negatives and false positives. An attack that satisfies all
-other conditions in a rule and has flags ``1'' or ``2'' set would be
-captured by the original rule but not captured by the inferred rule---a case of
-false negative. Likewise, an attack that satisfies all conditions and
-have flags ``1'' and ``2'' unset would be captured by the inferred
-rule but not captured by the original rule---a case of false positive. The set of
-positive examples is to blame in this case. To mitigate this issue,
-the input messages should have covered cases with options ``1'' or
-``2'' set. Considering option \CodeIn{window: 512}, \tname{} is able
-to generate rules without that option, as discussed above. Considering
-option \CodeIn{threshold}, the values used in the golden rule are
-domain-specific. \tname{} uses default values for these options. We
-conjecture that the system administrator would be able to configure
-this option.
-
-%% \begin{table}[h]
-%%   \caption{\label{table:rules}Rule evolution}  
-%%   \centering
-%%   \begin{tabular}{lllll}
-%%     \toprule
-%%     \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
-%%     \midrule     
-%%     0 & (ack:0; seq:0; window:0; flags:F;)\\
-%%     5 & (window:0; flags:F;)\\
-%%     68 & (window:64; flags:F;)\\
-%%     73 & (window:64; flags:S,12;)\\
-%%     94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
-%%     \bottomrule
-%%   \end{tabular}
-%% \end{table}
+time. The other options are for documentation purposes. This is a case
+where \tname\ works very well, reporting the golden rule at the top
+of the ranking.
 

### 2020-03-06 paper/paper.tex
pass related works

@@
 
-\Fix{...}
+\Luc{\textbf{Autonomous Rule Creation for Intrusion Detection} - Usam o Snort e geram regras apenas para pacotes ICMP.
+O pacote maligno tb eh sabido previamente e, assim como nos, eles identificam todas as opcoes q podem ser usadas para identificar esse pacote.
+A tecnica proposta utiliza algoritmos geneticos onde a populacao inicial eh formada por regras q combinam aleatoriamente todas essas opcoes.
+Eles foram capazes de gerar regras q n geram falsos negativos e apenas 3 positivos entre 33k pacotes.}
 

### 2020-03-06 paper/paper.tex
Guilherme pass related works

@@
 
+\Gui{\textbf{Efficient Snort Rule Generation using Evolutionary computing for Network Intrusion Detection} O metodo proposto por Raghavan eh baseado em Clustering utilizando algoritmos geneticos e Hill climbing para otimizar os clusters formados.
+Outros papers, citados no related works, tambem propuseram uma ideia semelhante, no [1][4][6] do paper de Raghban.
+A arquitetura proposta pega traffic files em intervalos de tempo constante e clusteriza os pacotes utilizando algoritmo hibrido de k means e partitioning around medoids.
+Com clusters formados, ele otimiza usando algoritmos geneticos e hill climbing.
+Eh entao criada uma regra que isola perfeitamente o cluster considerado como ataque.
+Quanto as opcoes, ele nao especifica quais, e nao deixa claro que considera content.
+O que ele fala eh que para lidar com dados continuos, como numero de bytes enviados numa conexao tcp, ele utiliza ranges com > e < ( maior que menor que).
+E para opcoes discretas sao consideradas todas e colocados como um OR.
+Ou seja, a IA serve para separar o ataque do trafego normal, e a regra eh gerada "do nada" e aparentemente eh bem especifica.}
+
 \section{Conclusions}

### 2020-03-06 paper/paper.tex
pass on intro

@@
 We evaluated \tname{} on a diverse
-set of attacks. Results indicate that \Fix{...}
+set of attacks. Results indicate that 
 \end{abstract}
@@
 not block. Various open-source (\eg{}, Snort~\cite{snort} and
-Suricata~\cite{suricata}) and commercial \nids\ implementations (\eg{},
-SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar}) exist
-today. These systems are very popular in industry to secure local
-computer networks given the amount of potential malicious traffic that
-exist on the Internet.
+Suricata~\cite{suricata}) and commercial (\eg{},
+SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar})
+implementations exist.
 
 This paper focuses on rule-based \nids{}, which are very popular in
-industry~\cite{proofpoint-etpro,snort-rule-subscriptions}. A
+industry~\cite{proofpoint-etpro,snort-rule-subscriptions} given the
+amount of potential malicious traffic that exist on the Internet. A
 \emph{rule-based intrusion detector}\footnote{\aka\ signature-based
@@
 set of rules. Figure~\ref{fig:pingscan-example} shows an example rule
-of Suricata~\cite{suricata}, a popular open-source \nids{}. This
-particular rule prescribes a method to capture an Active
-Reconnaissance attack for discovering IPS of hosts in a network (see
-Section~\ref{sec:active-recon}). Rulesets exist for Snort~\cite{snort}
-and Suricata~\cite{suricata}, providing protection against existing
-attacks. Unfortunately, attackers actively look for new mechanisms to
-exploit systems. For example,
-Figure~\ref{fig:distribution-rules-per-month} shows the number of
-rules added per month in the year of 2019 to the ``Emerging Threats
-Open Ruleset''~\cite{emerging-threats-open}, a popular public data set
-of \suri\ rules. We used the ``creation-date'' field associated with
-each entry in the ruleset. The data provides some evidence that new
+of Suricata~\cite{suricata}, a popular open-source \nids{}. A rule
+describes a pattern in the traffic. The observation of that pattern in
+the traffic typically indicates an attempt of a mallicious individual
+to attack a local network. This particular rule prescribes a method to
+capture an Active Reconnaissance attack for discovering accessible
+hosts in a network (see Section~\ref{sec:active-recon}). Rulesets
+exist for Snort~\cite{snort} and Suricata~\cite{suricata}, providing
+protection against existing attacks. Unfortunately, attackers actively
+look for new mechanisms to exploit systems through the
+network. Figure~\ref{fig:distribution-rules-per-month} shows the
+number of rules added per month in the year of 2019 to the ``Emerging
+Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
+data set of \suri\ rules. This data provides some evidence that new
 rules are created often.
 
-\begin{wrapfigure}[8in]{r}{0.24\textwidth}
-    \vspace{-5ex}  
+\begin{wrapfigure}[8]{r}{0.24\textwidth}
+    \vspace{-3ex}  
     \centering
@@
         \begin{axis}[
-%            label style={font=\scriptsize},
-%            tick label style={font=\scriptsize}
             width = 0.3\textwidth,
@@
     \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
-\end{wrapfigure}% <---- don't forget this %
-
-Intuitively, rulesets need to be continuously revised because new
-vulnerabilities are discovered and attackers can take advantage of
-them. Unfortunately, creating these rules is time-consuming and
-error-prone. The language for describing rules is non-trivial (see
-Section~\ref{sec:example-suricata-rules}) and testing correctness of a
-rule requires data and careful inspection.
+\end{wrapfigure}
+
+Intuitively, rulesets need to be continuously revised because the
+community finds and advertizes vulnerabilities and attackers can take
+advantage of them. Unfortunately, creating these rules is
+time-consuming and error-prone. The language for describing rules is
+non-trivial (see Section~\ref{sec:example-suricata-rules}) and testing
+correctness of a rule requires data and careful inspection.
 
@@
 rules and, consequently, help system administrators to harden the
-protection of \nids\ to network attacks. \tname{} synthesizes rules to
-protect against a given attack from 1) malicious traffic (\ie{},
-examples of the attack to protect against), 2) benign traffic, and 3)
-trusted rules.
-
-\Fix{
-denial-of-service
-attack to a server by matching specific conditions about the network
-traffic~\cite{understanding-dos}. Relevant properties about the
-traffic of interest appear in bold in this rule (see
-Section~\ref{sec:suri-metas-coverage}). Deployments of rule-based
-\nids\ are restricted to a fixed set of rules defined by the network
-system administrator.}
+protection of \nids\ against network attacks.  \tname{} leverages
+different sources of information to synthesize rules for a given
+attack. It uses 1) positive examples, \ie{}, scenarios that reproduce
+the attack, 2) negative examples in the form of a data set of benign
+messages, and 3) a data set of human-written trusted rules. The
+positive and negative examples constrain the search space of rules
+that can be used to capture the attack. Intuitively, positive examples
+indicate what should be captured whereas the negative examples
+indicate what should not be captured. It is worth noting that vast
+amounts of negative data exist~\cite{tcpreplay,stratosphere-normal},
+but positive data is scarce. Consequently, classical machine learning
+techniques that require considerable amounts of training data is not
+directly applicable in this context. Furthermore, \tname\ uses public
+data sets of rules to differentiate rules by the similarity to
+human-written rules.
+
+\tname{} takes on input one or more examples of an attack and produces
+on output a ranked list of rules for that attack. It uses public data
+sets of benign messages~\cite{tcpreplay,stratosphere-normal} and
+rulesets~\cite{emerging-threats-open} to create those rules. The
+technique is organized as three pipelined components. First,
+\tname\ uses the messages that manifest an attack to create a
+potentially overspecified seed rule. Then, it uses benign traffic in
+to derive from the seed rule a set of plausible rules that isolate the
+benign traffic from the malicious traffic. Finally, it uses existing
+rules to learn heuristics to rank the plausible rules.
+
+We evaluated \tname\ against 14 attacks of various kinds (see
+Table~\ref{table:attacks}). Results are encouraging--- \tname\ ranked
+the correct rule among the top five rules in \percTopFiveRanking\ of
+the cases and ranked the correct rule among the top ten rules in 11 of
+the 14 cases.
+
+%% \Fix{
+%% denial-of-service
+%% attack to a server by matching specific conditions about the network
+%% traffic~\cite{understanding-dos}. Relevant properties about the
+%% traffic of interest appear in bold in this rule (see
+%% Section~\ref{sec:suri-metas-coverage}). Deployments of rule-based
+%% \nids\ are restricted to a fixed set of rules defined by the network
+%% system administrator.}
 
@@
 
-\Fix{summarize how the technique works}
-
-\Fix{summarize results} We believe these results provide initial
-evidence on the effectiveness of \tname{}.
-
 This paper makes the following contributions.
+
+\newcommand{\Contrib}[1]{$\star$#1}
+\begin{itemize}[topsep=.2ex,itemsep=.2ex,leftmargin=0.8em]
+
+\item[\Contrib{}]\textbf{Approach.} A novel approach to synthesize
+  rules for \nids.
+
+\item[\Contrib{}]\textbf{Evaluation.} An evaluation of our approach on
+  attacks of various kinds. Results indicate that \tname\ is effective
+  in creating plausible rules and performs very well in ranking the
+  correct rule among those generated.
+
+\item[\Contrib{}]\textbf{Implementation.} A public implementation of
+  \tname. (Available on request to preserve double-blindness.)
+
+\item[\Contrib{}]\textbf{Data.} A data set of artifacts involved in
+  the evaluation of \tname\ (\eg{}, pcap files of example attacks, list
+  of plausible rules generated, etc.). The data set is available at
+  the following URL~\url{https://zenodo.org/record/3698366}.
+
+\end{itemize}
 
@@
 
-\newcommand{\percTopFiveRanking}{66.7\% (=10/15)}
-
 Overall, results show that \tname\ performed very well in most of the

### 2020-03-06 paper/paper.tex
Guilherme pass related works

@@
 
-\Gui{\textbf{Efficient Snort Rule Generation using Evolutionary computing for Network Intrusion Detection} O metodo proposto por Raghavan eh baseado em Clustering utilizando algoritmos geneticos e Hill climbing para otimizar os clusters formados.
+\Gui{\textbf{Efficient Snort Rule Generation using Evolutionary computing for Network Intrusion Detection -} O metodo proposto por Raghavan eh baseado em Clustering utilizando algoritmos geneticos e Hill climbing para otimizar os clusters formados.
 Outros papers, citados no related works, tambem propuseram uma ideia semelhante, no [1][4][6] do paper de Raghban.
@@
 
+\Gui{\textbf{Automatic NIDS Rule Generating System for Detecting HTTP-like Malware Communication -} Apresenta um sintetizador de regras para trafego HTTP.
+O metodo eh bem parecido com o nosso no que se trata em utilizar trafego maligno separado para a criacao de regra, e utilizacao de trafego benigno para detectar falso positivo.
+O ponto eh que eles nao descrevem o metodo que utilizam para separar os contents.
+Alem disso, caso um falso positivo seja capturado, a regra deve ser manualmente ajustadas.
+O paper possui algo interessante que o nosso nao possui. Ele eh capaz de detectar clusters nas regras criadas e pre existentes, sendo capaz de otimiza-las.}
+
 \section{Conclusions}

### 2020-03-06 paper/paper.tex
Guilherme pass related works

@@
 
+\Gui{\textbf{Automated Repair By Example for Firewalls -} Ruzica repara e modifica regras de firewall a partir de uma entrada contendo exemplos ou behaviour.
+Embora firewalls e NIDS nao sejam exatamente a mesma coisa, a estrutura de regras contem muitas semelhancas, e a utilizacao de exemplos para modificar a regra pode ser comparado com nosso metodo de criacao de regras a partir de um pacote maligno (para ser sincero, nao vi muitas semelhancas com nosso paper, nao sei descrever mais)}
+
+\Gui{\textbf{Search-based test and improvement of machine-learning-based anomaly detection systems -} o autor desenvolve uma ferramenta search-based que ira testar e explorar automaticamente vulnerabilidades em Anomaly-Based NIDS, e entao apresentando medidas preventivas contra o ataque recem criado atravez de attack schemes, que detectam apenas o ataque e com o minimo de falsos positivos possivel.
+Semelhante a nos, eles tambem utilizam trafego benigno para simular uma rede normal.
+(Os metodos preventivos sao para Anomaly-Based NIDS, nao sei como se encaixa no nosso paper tambem)}
+
 \section{Conclusions}

### 2020-03-06 paper/paper.tex
first pass on related work

@@
 
-\Luc{\textbf{Autonomous Rule Creation for Intrusion Detection} - Usam o Snort e geram regras apenas para pacotes ICMP.
-O pacote maligno tb eh sabido previamente e, assim como nos, eles identificam todas as opcoes q podem ser usadas para identificar esse pacote.
-A tecnica proposta utiliza algoritmos geneticos onde a populacao inicial eh formada por regras q combinam aleatoriamente todas essas opcoes.
-Eles foram capazes de gerar regras q n geram falsos negativos e apenas 3 positivos entre 33k pacotes.}
-
-\Gui{\textbf{Efficient Snort Rule Generation using Evolutionary computing for Network Intrusion Detection -} O metodo proposto por Raghavan eh baseado em Clustering utilizando algoritmos geneticos e Hill climbing para otimizar os clusters formados.
+This section elaborates on the most similar work.
+
+\noindent
+\textbf{Alternative methods for rule synthesis.} We discuss in the
+following techniques similar in goal to ours. Unfortunately, we could
+not access the tools implementing the techniques for side
+comparison. Nevertheless, we found that there are important
+differences between our work and related work. Vollmer
+\etal{}~\cite{vollmer-etal-cics2011} propose the use of search to
+derive rules for Snort, which are very similar to \suri's.  Their
+approach focuses on icmp attacks. It models every field of the icmp
+packet in the chromosome of candidate solutions (\ie{}, a rule) of a
+genetic search. The initial population is a set of candidate solutions
+that randomly combine options, modelling the packet fields. They show
+that they were able to generate rules without false negatives and only
+3 false positives across 33K benign packets. The key differences
+between our work and Vollmer~\etal{}'s are 1) the support of different
+protocols and 2) the use of ranking to discriminate alternative
+plausible rules. As for 1, our approach supports a variety of
+protocols and options that refer to message payloads which
+significantly increase the search space. Note that icmp rules only
+refer to message fields (see Section~\ref{sec:rules-and-packets}). As
+for 2, our solution additionally ranks plausible options, which
+contain \emph{no} false positives or negatives with respect to the
+benign and mallicious traffic considered. Muthuregunathan
+\etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
+create rules. They characterize candidate solutions as a set of field
+values and use a clustering algorithm---hybrid k-means partitioned
+around menoids---to isolate mallicious from benign
+traffic. Subsequently, they optimize the formed clusters using genetic
+algorithms and hill-climbing search. Ou approach differs from theirs
+in the method used, results, and support for content options.
+
+%% \Luc{\textbf{Autonomous Rule Creation for Intrusion Detection} -
+%% Usam o Snort e geram regras apenas para pacotes ICMP. O pacote
+%% maligno tb eh sabido previamente e, assim como nos, eles
+%% identificam todas as opcoes q podem ser usadas para identificar
+%% esse pacote. A tecnica proposta utiliza algoritmos geneticos onde a
+%% populacao inicial eh formada por regras q combinam aleatoriamente
+%% todas essas opcoes. Eles foram capazes de gerar regras q n geram
+%% falsos negativos e apenas 3 positivos entre 33k pacotes.}
+
+%% \Gui{\textbf{Efficient Snort Rule Generation using Evolutionary computing for Network Intrusion Detection -} O metodo proposto por Raghavan eh baseado em Clustering utilizando algoritmos geneticos e Hill climbing para otimizar os clusters formados.
 Outros papers, citados no related works, tambem propuseram uma ideia semelhante, no [1][4][6] do paper de Raghban.
-A arquitetura proposta pega traffic files em intervalos de tempo constante e clusteriza os pacotes utilizando algoritmo hibrido de k means e partitioning around medoids.
+%% A arquitetura proposta pega traffic files em intervalos de tempo constante e clusteriza os pacotes utilizando algoritmo hibrido de k means e partitioning around medoids.
 Com clusters formados, ele otimiza usando algoritmos geneticos e hill climbing.
-Eh entao criada uma regra que isola perfeitamente o cluster considerado como ataque.
+%% Eh entao criada uma regra que isola perfeitamente o cluster considerado como ataque.
 Quanto as opcoes, ele nao especifica quais, e nao deixa claro que considera content.
@@
 E para opcoes discretas sao consideradas todas e colocados como um OR.
-Ou seja, a IA serve para separar o ataque do trafego normal, e a regra eh gerada "do nada" e aparentemente eh bem especifica.}
+%% Ou seja, a IA serve para separar o ataque do trafego normal, e a regra eh gerada "do nada" e aparentemente eh bem especifica.}
 

### 2020-03-06 paper/paper.tex
discussing another paper

@@
 algorithms and hill-climbing search. Ou approach differs from theirs
-in the method used, results, and support for content options.
+in the method used, results, and support for content
+options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed ARGS, a system
+to synthesize http rules that take mallicious traffic as input. ARGS
+looks for a common pattern across the different inputs to define
+rules. It does not take existing rules as input or benign
+traffic. Consequently, false positive can be produced and need to be
+handled manually by users. Our approach, in contrast, does not
+generate false positives w.r.t. the input traffic.
 
@@
 
-\Gui{\textbf{Automatic NIDS Rule Generating System for Detecting HTTP-like Malware Communication -} Apresenta um sintetizador de regras para trafego HTTP.
+%% \Gui{\textbf{Automatic NIDS Rule Generating System for Detecting HTTP-like Malware Communication -} Apresenta um sintetizador de regras para trafego HTTP.
 O metodo eh bem parecido com o nosso no que se trata em utilizar trafego maligno separado para a criacao de regra, e utilizacao de trafego benigno para detectar falso positivo.
@@
 Alem disso, caso um falso positivo seja capturado, a regra deve ser manualmente ajustadas.
-O paper possui algo interessante que o nosso nao possui. Ele eh capaz de detectar clusters nas regras criadas e pre existentes, sendo capaz de otimiza-las.}
+%% O paper possui algo interessante que o nosso nao possui. Ele eh capaz de detectar clusters nas regras criadas e pre existentes, sendo capaz de otimiza-las.}
 

### 2020-03-06 paper/paper.tex
pass on related work

@@
 Alem disso, caso um falso positivo seja capturado, a regra deve ser manualmente ajustadas.
-%% O paper possui algo interessante que o nosso nao possui. Ele eh capaz de detectar clusters nas regras criadas e pre existentes, sendo capaz de otimiza-las.}
-
-\Gui{\textbf{Automated Repair By Example for Firewalls -} Ruzica repara e modifica regras de firewall a partir de uma entrada contendo exemplos ou behaviour.
+%% O paper possui algo interessante que o nosso nao possui. Ele eh
+%% capaz de detectar clusters nas regras criadas e pre existentes,
+%% sendo capaz de otimiza-las.}
+
+\noindent
+\textbf{Rule repair.}~Hallahan~\etal{}~\cite{8102263} propose to
+repair firewall rules based on false positive detected by the
+rules. Although firewall and \nids\ rules have differences and the
+problem addressed is different---repair not synthesis---the work is
+similar in spirit and inspiring. We remain to evaluate how \tname\ can
+be used for repair.
+
+\noindent
+\textbf{Anomaly-based intrusion detection.}
+Cordy~\etal{}\cite{cordy-etal-issta19} propose the use of
+\Fix{...what...} to \Fix{...calibrate parameters of?....}
+anomaly-based \nids. Although our approach and theirs use machine
+intelligence to detect attacks, as discussed on
+Section~\ref{sec:background} anomaly-based and rule-based \nids\ are
+complemetary.
+
+%% \Gui{\textbf{Automated Repair By Example for Firewalls -} Ruzica repara e modifica regras de firewall a partir de uma entrada contendo exemplos ou behaviour.
 Embora firewalls e NIDS nao sejam exatamente a mesma coisa, a estrutura de regras contem muitas semelhancas, e a utilizacao de exemplos para modificar a regra pode ser comparado com nosso metodo de criacao de regras a partir de um pacote maligno (para ser sincero, nao vi muitas semelhancas com nosso paper, nao sei descrever mais)}
 
-\Gui{\textbf{Search-based test and improvement of machine-learning-based anomaly detection systems -} o autor desenvolve uma ferramenta search-based que ira testar e explorar automaticamente vulnerabilidades em Anomaly-Based NIDS, e entao apresentando medidas preventivas contra o ataque recem criado atravez de attack schemes, que detectam apenas o ataque e com o minimo de falsos positivos possivel.
+%% \Gui{\textbf{Search-based test and improvement of machine-learning-based anomaly detection systems -} o autor desenvolve uma ferramenta search-based que ira testar e explorar automaticamente vulnerabilidades em Anomaly-Based NIDS, e entao apresentando medidas preventivas contra o ataque recem criado atravez de attack schemes, que detectam apenas o ataque e com o minimo de falsos positivos possivel.
 Semelhante a nos, eles tambem utilizam trafego benigno para simular uma rede normal.

### 2020-03-06 paper/paper.tex
drafting conclusion

@@
 
-\Fix{...}
+Network Intrusion Detectors are an important defense mechanism against
+networks attacks. This paper presents \tname{}, a novel approach to
+generate rules for Rule-based \nids. \tname\ leverages positive and
+negative traffic and examples of trusted rules to synthesize
+rules. \tname{} is organized as a pipeline of three components that
+1)~create an overspecified seed rule, 2)~derive plausible rules from
+the seed, and 3)~rank plausible rules. Experimental results provide
+early evidence showing that \tname\ is very promising. It was able to
+rank the correct rule among the first five rules of the reported
+ranking in \percTopFiveRanking\ of the cases we analyzed.
 

### 2020-03-06 paper/paper.tex
pass section 7

@@
 significantly increase the search space. Note that icmp rules only
-refer to message fields (see Section~\ref{sec:rules-and-packets}). As
-for 2, our solution additionally ranks plausible options, which
+refer to message fields (see Section~\ref{sec:rules-and-packets})
+\Luc{eh possivel incluir contents em regras icmp tb. o q a secao 2.3 
+diz q \textbf{geralmente} as regras nao-http possuem apenas field options}. 
+As for 2, our solution additionally ranks plausible options, which
 contain \emph{no} false positives or negatives with respect to the

### 2020-03-06 paper/paper.tex
fix jboss

@@
 Table~\ref{table:results})---idqaccess and scripttag---and the case
-where \tname\ performs worst---jboss. We measured the position of the
+where \tname\ performs worst---jboss \Fix{Remover jboss}. We measured the position of the
 golden rule at the rank for an increasing number of of examples

### 2020-03-06 paper/paper.tex
pass on abstract

@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
 \documentclass[sigconf,review, anonymous]{acmart}
@@
 \renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
 \begin{abstract}
-  \Fix{revise}
-Network Intrusion Detection Systems (\nids{}) are a popular mechanism
-used by system administrators to defend against network attacks. These
-systems monitor the network traffic and flag suspicious network
-behavior. Signature-based \nids\ do that by checking the\MyComment{ network}
-traffic against a pre-defined set of rules, which become obsolete
-as attackers learn new strategies to circumvent existing defenses.
-
-This paper proposes \tname{}, a technique that automatically
-synthesizes \nids\ rules from positive and negative examples, \ie{},
-malicious and benign traffic. \tname{} formulates synthesis as an
-optimization problem whose candidate solutions are rules. Optimal
-solutions maximize the capture of positive traffic and minimize the
-capture of negative traffic. \tname{} bootstraps the search with a
-candidate solution inferred from the payload of the malicious message.
-It uses that initial overfitting solution to search for alternative
-rules that only capture the malicious traffic. Then, it ranks those
-rules according to heuristics---drawn from a data set of existing
-rules---denoting the likelihood of those rules of being useful.
-
-We evaluated \tname{} on a diverse
-set of attacks. Results indicate that 
+
+Network Intrusion Detection Systems (\nids{}) are popular tools to
+defend against network attacks. These systems monitor the network
+traffic and flag suspicious network behavior. Rule-based \nids\ do
+that by checking the network traffic against a set of rules, which
+become obsolete as attackers learn new strategies to circumvent
+existing defenses. This paper proposes \tname{}, a novel approach to
+synthesize rules for Rule-based \nids. \tname\ leverages positive and
+negative traffic and examples of trusted rules to create
+rules. \tname{} is organized as a pipeline of three components that
+1)~create an overspecified seed rule, 2)~derive plausible rules from
+the seed, and 3)~rank plausible rules. Experimental results provide
+early evidence showing that \tname\ is very promising. For example, it
+was able to rank the correct rule among the first five rules of the
+reported ranking in \percTopFiveRanking\ of the cases we analyzed.
+
 \end{abstract}
@@
 %%
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+%% \begin{CCSXML}
+%% <ccs2012>
+%%  <concept>
+%%   <concept_id>10010520.10010553.10010562</concept_id>
+%%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+%%   <concept_significance>500</concept_significance>
+%%  </concept>
+%%  <concept>
+%%   <concept_id>10010520.10010575.10010755</concept_id>
+%%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
+%%   <concept_significance>300</concept_significance>
+%%  </concept>
+%%  <concept>
+%%   <concept_id>10010520.10010553.10010554</concept_id>
+%%   <concept_desc>Computer systems organization~Robotics</concept_desc>
+%%   <concept_significance>100</concept_significance>
+%%  </concept>
+%%  <concept>
+%%   <concept_id>10003033.10003083.10003095</concept_id>
+%%   <concept_desc>Networks~Network reliability</concept_desc>
+%%   <concept_significance>100</concept_significance>
+%%  </concept>
+%% </ccs2012>
+%% \end{CCSXML}
+
+%% \ccsdesc[500]{Computer systems organization~Embedded systems}
+%% \ccsdesc[300]{Computer systems organization~Redundancy}
+%% \ccsdesc{Computer systems organization~Robotics}
+%% \ccsdesc[100]{Networks~Network reliability}
 
@@
 amount of potential malicious traffic that exist on the Internet. A
-\emph{rule-based intrusion detector}\footnote{\aka\ signature-based
-  intrusion detector.} checks if the network traffic matches a fixed
+\emph{rule-based intrusion detector}\MyComment{\footnote{\aka\ signature-based
+  intrusion detector.}} checks if the network traffic matches a fixed
 set of rules. Figure~\ref{fig:pingscan-example} shows an example rule
@@
 
-\begin{wrapfigure}[8]{r}{0.24\textwidth}
-    \vspace{-3ex}  
+Intuitively, rulesets need to be continuously revised because the
+community finds and advertizes vulnerabilities and attackers can take
+advantage of them. Unfortunately, creating these rules is
+\begin{wrapfigure}[8]{r}{0.21\textwidth}
+    \vspace{-2ex}  
     \centering
@@
 \end{wrapfigure}
-
-Intuitively, rulesets need to be continuously revised because the
-community finds and advertizes vulnerabilities and attackers can take
-advantage of them. Unfortunately, creating these rules is
 time-consuming and error-prone. The language for describing rules is
@@
 complemetary.
+\Gui{Anomaly-based intrusion detection. Cordy and colleagues[28] 
+propose the use of a search-based genetic algorithm to not only 
+generate attack shcemes capable of penetrating the tested anomaly-based 
+NIDS, but also repairing the NIDS to identify the given attack scheme. 
+Although our approach and theirs use machine intelligence to detect 
+attacks, as discussed on Section 2 anomaly-based and rule-based NIDS 
+are complementary.}
 

### 2020-03-06 paper/paper.tex
spell checking

--- 2020-03-06 paper/paper.tex
@@
 rules. \tname{} is organized as a pipeline of three components that
-1)~create an overspecified seed rule, 2)~derive plausible rules from
+1)~create an over-specified seed rule, 2)~derive plausible rules from
 the seed, and 3)~rank plausible rules. Experimental results provide
@@
 describes a pattern in the traffic. The observation of that pattern in
-the traffic typically indicates an attempt of a mallicious individual
+the traffic typically indicates an attempt of a malicious individual
 to attack a local network. This particular rule prescribes a method to
@@
 Intuitively, rulesets need to be continuously revised because the
-community finds and advertizes vulnerabilities and attackers can take
+community finds and advertises vulnerabilities and attackers can take
 advantage of them. Unfortunately, creating these rules is
@@
 \tname\ uses the messages that manifest an attack to create a
-potentially overspecified seed rule. Then, it uses benign traffic in
+potentially over-specified seed rule. Then, it uses benign traffic in
 to derive from the seed rule a set of plausible rules that isolate the
@@
 anomaly-based \nids\ can report false alarms as learning algorithms
-that generalize observations from bening traffic are approximate.
+that generalize observations from benign traffic are approximate.
 This paper focuses on rule-based \nids\ for its tremendous popularity
@@
   the action taken when an intrusion is observed. IPS systems block
-  malicous traffic whereas IDS systems only alert them. However, IDS
+  malicious traffic whereas IDS systems only alert them. However, IDS
   and IPS are similar in the approaches used to detect
@@
 this set of options captures the malicious traffic. Unfortunately, the
-rule above is overspecified. Informally, the rule includes ``too
+rule above is over-specified. Informally, the rule includes ``too
 specific'' characteristics of the malicious traffic and could
@@
 
-Second, \tname{} uses the benign traffic to mitigate the overfitting
+Second, \tname{} uses the benign traffic to mitigate the over-fitting
 problem. In this step, \tname{} searches for alternative rules that
@@
 To sum up, \tname{} uses the messages that manifest an attack in the
-first step to create a potentially overspecified seed rule. Then, it
+first step to create a potentially over-specified seed rule. Then, it
 uses benign traffic in the second step to create a set of plausible
@@
 that, as only the malicious traffic is considered in this step, the
-seed rule is likely overspecified, \ie{}, it may be more restrictive
+seed rule is likely over-specified, \ie{}, it may be more restrictive
 than necessary. Consequently, the use of that rule could result in
@@
 case of this strategy. We obtained the option-field mapping
-manunually, from the \suri\ documentation.
+manually, from the \suri\ documentation.
 
@@
 capture the traffic produced by the example attacks. However, that
-rule is overconstrained.  It may include options that are irrelevant
+rule is over-constrained.  It may include options that are irrelevant
 to capture the attack and lead to false negatives---the \nids\ may be
@@
 
-To address the overffiting issue, \tname{} leverages public data sets
+To address the overfiting issue, \tname{} leverages public data sets
 of benign traffic~\cite{tcpreplay} and additional examples of the
@@
   %\noindent\hspace{-2ex}
-  \flushleft{}\textbf{INPUT:} (overconstrained) seed rule $sr$, pcap
-  files for bening traffic\\
-  ($\mathit{bTr}$), pcap files for mallicious traffic ($\mathit{mTr}$)\\
+  \flushleft{}\textbf{INPUT:} (over-constrained) seed rule $sr$, pcap
+  files for benign traffic\\
+  ($\mathit{bTr}$), pcap files for malicious traffic ($\mathit{mTr}$)\\
   %\noindent\hspace{-27.5ex} 
@@
 seed rule produced at the Reverse Engineering step. The algorithm
-takes as input an overconstrained rule $sr$ and pcap files~\cite{pcap}
+takes as input an over-constrained rule $sr$ and pcap files~\cite{pcap}
 to reconstruct the benign ($\mathit{bTr}$) and malicious traffic
@@
 In summary, this algorithm reports many rules that are plausible by
-using the benign traffic to reduce the overconstrained rule produced
+using the benign traffic to reduce the over-constrained rule produced
 at the \reveng\ step. In the Raking step (Section~\ref{sec:ranking}),
@@
 
-\tname{} uses a composition of heurisitic functions to sort the rules
+\tname{} uses a composition of heuristic functions to sort the rules
 produced in the previous step. Each of these functions uses a
@@
 average across the values produced by each heuristic function. The
-codomain of each function ranges on the 0-1 interval. Hence, the score
+co-domain of each function ranges on the 0-1 interval. Hence, the score
 of a rule also ranges on the 0-1 interval. We describe in the
@@
 that string appears in the field \CodeIn{Request URI} (of an http
-message from the malicous traffic) and that field is accessible
+message from the malicious traffic) and that field is accessible
 through these two modifiers. The relatively high heuristic value
@@
 (=$\frac{4089+29}{10111*2}$), in this case. Assuming that a rule uses
-the two content options mentioned aboved, the aggregated contribution
+the two content options mentioned above, the aggregated contribution
 of this heuristic function to the rule would be 0.35 (=(0.2+0.5)/2).
@@
 golden rule on each attack. The weight combination that led to the
-smaller sum of positions (of the golden rule) in the rakings was
+smaller sum of positions (of the golden rule) in the rankings was
 considered the best choice. We considered \emph{all} combinations of
@@
 pcap files for some of the http attacks. We configured an http server
-(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
+(Apache2) and ran Nikto2 to attack that server. Simultaneously, we ran
 WireShark~\cite{wireshark-net-monitor} to record the network traffic
@@
 involved on each activation. With the packet numbers, it is possible
-to recover the corresponding pcap files, using Wireshark, as to ran
+to recover the corresponding pcap files, using WireShark, as to ran
 \tname{} on them separately. It is worth noting that some rules are
@@
 \vspace{1ex}
-\subsubsection{\label{sec:objectanalysis-benign}Benign traffic.}~Public data sets of bening traffic
+\subsubsection{\label{sec:objectanalysis-benign}Benign traffic.}~Public data sets of benign traffic
 exist and can be used to evaluate the rate of false positives of
@@
 Table~\ref{table:attacks} only shows the essential
-options. Documentation options (\eg{}, ``msg'') are discared.
+options. Documentation options (\eg{}, ``msg'') are discarded.
 
@@
 contributes to the overall performance of the technique. Recall that
-the score a rule $r$ is given by the weigthed average of the values
+the score a rule $r$ is given by the weighted average of the values
 given by each heuristic function, \ie{} the score of $r$ is obtained
@@
 results obtained by running the same experiment, as on
-Section~\ref{sec:answer-rqone}, but supressing one heuristic
+Section~\ref{sec:answer-rqone}, but suppressing one heuristic
 function. The number corresponds to the order the heuristic function
@@
 of the degradation can be very high in some cases. Basebuilder and
-Tomcat are two cases where results get much worse when supressing one
+Tomcat are two cases where results get much worse when suppressing one
 of the functions. Also note that the number of down arrows is
@@
 genetic search. The initial population is a set of candidate solutions
-that randomly combine options, modelling the packet fields. They show
+that randomly combine options, model-ling the packet fields. They show
 that they were able to generate rules without false negatives and only
@@
 values and use a clustering algorithm---hybrid k-means partitioned
-around menoids---to isolate mallicious from benign
+around menoids---to isolate malicious from benign
 traffic. Subsequently, they optimize the formed clusters using genetic
@@
 options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed ARGS, a system
-to synthesize http rules that take mallicious traffic as input. ARGS
+to synthesize http rules that take malicious traffic as input. ARGS
 looks for a common pattern across the different inputs to define
@@
 Section~\ref{sec:background} anomaly-based and rule-based \nids\ are
-complemetary.
+complementary.
 \Gui{Anomaly-based intrusion detection. Cordy and colleagues[28] 
 propose the use of a search-based genetic algorithm to not only 
-generate attack shcemes capable of penetrating the tested anomaly-based 
+generate attack schemes capable of penetrating the tested anomaly-based 
 NIDS, but also repairing the NIDS to identify the given attack scheme. 
@@
 %% End of file `sample-sigconf.tex'.
+
+%%  LocalWords:  Suricata SolarWinds QRadar Rulesets Ruleset rulesets
+%%  LocalWords:  pipelined pcap OISF IPS icmp msg classtype sid arg
+%%  LocalWords:  clp http ip bitvector flagbits IPs NMAP dsize itype
+%%  LocalWords:  icode ruleset th HtmlAdaptor pseudocode prce dataset
+%%  LocalWords:  flowbits len tcp uri RQ lrrrr Cand UDP ColdFusion
+%%  LocalWords:  iisadmin JBoss JMX WeBid LFI htaccess idq DoS AREC
+%%  LocalWords:  Basebuilder PESC RCE NetReseC Nikto WireShark hping
+%%  LocalWords:  Bigflows TcpReplay lrrrrrr boxplot Metasploit ACK
+%%  LocalWords:  Vollmer Hallahan Cordy NIDS

### 2020-03-06 paper/paper.tex
handling comments

@@
   of plausible rules generated, etc.). The data set is available at
-  the following URL~\url{https://zenodo.org/record/3698366}.
+  the following URL~\ourdataset.
 
@@
 that generalize observations from benign traffic are approximate.
-This paper focuses on rule-based \nids\ for its tremendous popularity
-in industry\Fix{cite or tone down}. Although we used Suricata, the
+This paper focuses on rule-based \nids\ for its popularity
+in industry. Although we used Suricata, the
 principles we used are general to any other signature-based
@@
   \end{tabular}
-  \vspace{-3ex}  
+%  \vspace{-3ex}  
 \end{table}
@@
 often associated with packet fields. Let us consider a rule for an
-icmp attack as example. Figure~\ref{fig:icmp-packet-layout} shows the
+icmp attack as example.
+
+\begin{figure}[h!]
+  \centering
+  \vspace{-2ex}  
+  \includegraphics[scale=0.27]{figs/ICMP-packet-structure.png}
+  \vspace{-2ex}
+  \caption{Layout of an icmp packet.}
+  \vspace{-2ex}
+  \label{fig:icmp-packet-layout}
+\end{figure}
+
+Figure~\ref{fig:icmp-packet-layout} shows the
 layout of an icmp packet. The packet fields ``Flags'' and ``Time to
@@
 attacks. That option operates over sequence of messages as opposed to
-a single message. Consequently, it is not possible to determine its
-values from a single message.
-
-\begin{figure}[t!]
-\centering
-\includegraphics[scale=0.27]{figs/ICMP-packet-structure.png}
-\vspace{-2ex}
-\caption{Layout of an icmp packet.}
-\vspace{-2ex}
-\label{fig:icmp-packet-layout}
-\end{figure}
-%\vspace{-2ex}
+single messages.
+
+%Consequently, it is not possible to determine its
+%values from a single message.
 
@@
   \caption{Suricata rule for Ping Scan attack.}
+  \vspace{-4ex}  
   \label{fig:pingscan-example}
 \end{figure}
-\vspace{-3ex}
 
@@
 
+%%---and the case where \tname\ performs worst---jboss \Fix{Remover jboss}
+
 To answer this research question, we selected two cases with the
 highest number of examples (see column ``\#~Vars'' on
-Table~\ref{table:results})---idqaccess and scripttag---and the case
-where \tname\ performs worst---jboss \Fix{Remover jboss}. We measured the position of the
-golden rule at the rank for an increasing number of of examples
-provided to \tname. For each number of variants, we sampled five sets
-of that size to provide as input to the
+Table~\ref{table:results})---idqaccess and scripttag. We measured the
+position of the golden rule at the rank for an increasing number of
+examples provided to \tname. For each number of variants, we sampled
+five sets of that size to provide as input to the
 technique. Figure~\ref{fig:impact-number-attacks} shows the results
@@
 the boxplot of the distribution of results for the corresponding
-number of variants.
+number of variants. Results indicate that 1) the number of examples is
+important, 2) the choice of examples is important, and 3) a small
+number of examples suffices to reach the best results.
 
@@
 \end{figure}
-
-Results indicate that 1) the number of examples is important, 2) the
-choice of examples is important, and 3) a small number of examples
-suffices to reach the best results.
 
@@
 
-\begin{figure}[H]
-  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
-  \vspace{-2ex}  
-  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code
-    Execution.}
-  \vspace{-3ex}
-\end{figure}
-
 %A opção  define que o match só será feito em pacotes que fazem parte
@@
 this attack, the ground truth was located at position 45 in the ranked
-list. The reason for that is that \Fix{...xxx...}
-
-%% \Mar{move from above --> We used the following hping3 command to
-%%   reproduce that attack: \CodeIn{hping3 -d 80 -w 512 -S -p 80 --flood
-%%     --rand-source <IP>}.}
+list.
+
+\begin{figure}[t!]
+%  \vspace{-1ex}    
+  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
+  \vspace{-2ex}  
+  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code
+    Execution.}
+%  \vspace{-1ex}
+\end{figure}
 
@@
 genetic search. The initial population is a set of candidate solutions
-that randomly combine options, model-ling the packet fields. They show
+that randomly combine options, modelling the packet fields. They show
 that they were able to generate rules without false negatives and only
@@
 protocols and options that refer to message payloads which
-significantly increase the search space. Note that icmp rules only
-refer to message fields (see Section~\ref{sec:rules-and-packets})
-\Luc{eh possivel incluir contents em regras icmp tb. o q a secao 2.3 
-diz q \textbf{geralmente} as regras nao-http possuem apenas field options}. 
+significantly increase the search space. The icmp rules they generate only
+refer to message fields (see Section~\ref{sec:rules-and-packets}).
+\MyComment{\Luc{eh possivel incluir contents em regras icmp tb. o q a secao 2.3 
+diz q \textbf{geralmente} as regras nao-http possuem apenas field options}. }
 As for 2, our solution additionally ranks plausible options, which
@@
 \textbf{Anomaly-based intrusion detection.}
-Cordy~\etal{}\cite{cordy-etal-issta19} propose the use of
-\Fix{...what...} to \Fix{...calibrate parameters of?....}
-anomaly-based \nids. Although our approach and theirs use machine
-intelligence to detect attacks, as discussed on
-Section~\ref{sec:background} anomaly-based and rule-based \nids\ are
-complementary.
-\Gui{Anomaly-based intrusion detection. Cordy and colleagues[28] 
-propose the use of a search-based genetic algorithm to not only 
-generate attack schemes capable of penetrating the tested anomaly-based 
-NIDS, but also repairing the NIDS to identify the given attack scheme. 
-Although our approach and theirs use machine intelligence to detect 
-attacks, as discussed on Section 2 anomaly-based and rule-based NIDS 
-are complementary.}
+Cordy~\etal{}\cite{cordy-etal-issta19} propose the use of a genetic
+algorithm to not only generate attack schemes capable of penetrating
+the tested anomaly-based NIDS, but also repairing the NIDS to identify
+the given attack scheme. Although our approach and theirs use machine
+intelligence to detect attacks, anomaly-based and rule-based
+\nids\ are complementary, as discussed on
+Section~\ref{sec:background}.
 
@@
 rank the correct rule among the first five rules of the reported
-ranking in \percTopFiveRanking\ of the cases we analyzed.
+ranking in \percTopFiveRanking\ of the cases we analyzed. The data set
+containing our experimental artifacts is publicly accessible at \ourdataset.
 

### 2020-03-06 paper/paper.tex
fixing attack names

@@
     JBoss JMX\MyComment{ Console Beanshell Exploit} & \cite{nikto} & RCE & http \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
-    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
+    scripttag\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
     WeBid LFI \MyComment{Local File Inclusion} & \cite{nikto} & LFI & http & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
@@
     Basebuilder LFI & \cite{nikto} & LFI & http & \textbf{content}:"GET"; http\_method; \textbf{content}:"/main.inc.php?"; nocase; http\_uri; \textbf{content}:"mj\_config[src\_path]="; nocase; http\_uri; \\
-    Tomcat server snoop access & \cite{nikto} & AREC & http & \textbf{content}:"/jsp/snp/"; http\_uri; \textbf{content}:".snp"; http\_uri;\\  
+    Tomcat Snoop Access & \cite{nikto} & AREC & http & \textbf{content}:"/jsp/snp/"; http\_uri; \textbf{content}:".snp"; http\_uri;\\  
     \bottomrule
@@
     JBoss JMX & 5 & 23 & 228 & 45 \\
-    Script tag in URI & 23 & 22 & 786 & 1 \\
+    scripttag & 23 & 22 & 786 & 1 \\
     WeBid LFI & 1 & 23 & 11,100 & 10\\    
@@
 %%    /system32/ in URI & 0,1,1,1 & 11 \\
-%%    Script tag in URI & 0,1,1,1 & 9 \\
+%%    scripttag & 0,1,1,1 & 9 \\
 %%    \bottomrule
@@
     JBoss JMX & 45 & 45 & 5\cellcolor{lightgray}$\downarrow$ & 45 & 111$\uparrow$ \\
-    Script tag in URI & 1 & 1 & 4$\uparrow$ & 1 & 8$\uparrow$ \\
+    scripttag & 1 & 1 & 4$\uparrow$ & 1 & 8$\uparrow$ \\
     WeBid LFI & 10 & 10 & 77$\uparrow$ & 28$\uparrow$ & 32$\uparrow$ \\    

### 2020-05-11 paper/paper.tex
pass before Rui can jump in

@@
-\documentclass[sigconf,review, anonymous]{acmart}
-\acmConference[ESEC/FSE 2020]{The 28th ACM Joint European Software Engineering Conference and Symposium on the Foundations of Software Engineering}{8 - 13 November, 2020}{Sacramento, California, United States}
-
-\pagenumbering{arabic} % Remove for camera-ready
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first
+% footnote. If that is unneeded, please comment it out.
 
@@
 
-%% 
 \input{macros}
 
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-%% \setcopyright{acmcopyright}
-%% \copyrightyear{2018}
-%% \acmYear{2018}
-%% \acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-%% \acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
-%%   Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
-%% \acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%   June 03--05, 2018, Woodstock, NY}
-%% \acmPrice{15.00}
-%% \acmISBN{978-1-4503-XXXX-X/18/06}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-%%
-%% end of the preamble, start of the body of the document source.
+%\usepackage{theorem}
+\usepackage{amsmath,amssymb,amsfonts}
+%\usepackage{algorithmic}
+%% \usepackage{graphicx}
+%% \usepackage{textcomp}
+%% \usepackage{xcolor}
+
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 \begin{document}
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
 \title{Learning to Synthesize Rules for Rule-based Intrusion Detectors}
 
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-\author{Ben Trovato}
-\authornote{Both authors contributed equally to this research.}
-\email{trovato@corporation.com}
-\orcid{1234-5678-9012}
-\author{G.K.M. Tobin}
-\authornotemark[1]
-\email{webmaster@marysville-ohio.com}
-\affiliation{%
-  \institution{Institute for Clarity in Documentation}
-  \streetaddress{P.O. Box 1212}
-  \city{Dublin}
-  \state{Ohio}
-  \postcode{43017-6221}
-}
-
-\author{Lars Th{\o}rv{\"a}ld}
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{larst@affiliation.org}
-
-\author{Valerie B\'eranger}
-\affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
-}
-
-\author{Aparna Patel}
-\affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh}
- \state{Arunachal Pradesh}
- \country{India}}
-
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu}
-  \state{Beijing Shi}
-  \country{China}}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas}
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P. Kumquat}
-\affiliation{\institution{The Kumquat Consortium}}
-\email{jpkumquat@consortium.net}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
+%% \title{Conference Paper Title*\\
+%% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
+%% should not be used}
+%% \thanks{Identify applicable funding agency here. If none, delete this.}
+%% }
+
+%% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% \and
+%% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
+%% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+%% \textit{name of organization (of Aff.)}\\
+%% City, Country \\
+%% email address or ORCID}
+%% }
+
+\pagestyle{plain}
+\maketitle
+
 \begin{abstract}
-
 Network Intrusion Detection Systems (\nids{}) are popular tools to
@@
 the seed, and 3)~rank plausible rules. Experimental results provide
-early evidence showing that \tname\ is very promising. For example, it
-was able to rank the correct rule among the first five rules of the
-reported ranking in \percTopFiveRanking\ of the cases we analyzed.
-
+early evidence showing that \tname\ is promising. For example, it was
+able to synthesize the correct rule in all cases we analyzed and
+ranked the correct rule among the first five rules of the reported
+ranking in \percTopFiveRanking\ of the cases.
 \end{abstract}
 
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-%% \begin{CCSXML}
-%% <ccs2012>
-%%  <concept>
-%%   <concept_id>10010520.10010553.10010562</concept_id>
-%%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%%   <concept_significance>500</concept_significance>
-%%  </concept>
-%%  <concept>
-%%   <concept_id>10010520.10010575.10010755</concept_id>
-%%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%%   <concept_significance>300</concept_significance>
-%%  </concept>
-%%  <concept>
-%%   <concept_id>10010520.10010553.10010554</concept_id>
-%%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%%   <concept_significance>100</concept_significance>
-%%  </concept>
-%%  <concept>
-%%   <concept_id>10003033.10003083.10003095</concept_id>
-%%   <concept_desc>Networks~Network reliability</concept_desc>
-%%   <concept_significance>100</concept_significance>
-%%  </concept>
-%% </ccs2012>
-%% \end{CCSXML}
-
-%% \ccsdesc[500]{Computer systems organization~Embedded systems}
-%% \ccsdesc[300]{Computer systems organization~Redundancy}
-%% \ccsdesc{Computer systems organization~Robotics}
-%% \ccsdesc[100]{Networks~Network reliability}
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-% \keywords{\nids, synthesis, search}
-
-%% remove on camera-ready
-\settopmatter{printacmref=false}
-
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-\maketitle
+%% \begin{IEEEkeywords}
+%% component, formatting, style, styling, insert
+%% \end{IEEEkeywords}
 
@@
 SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar})
-implementations exist.
+\nids\ implementations exist.
+
+% given the amount of potential malicious traffic that exist on the Internet
 
 This paper focuses on rule-based \nids{}, which are very popular in
-industry~\cite{proofpoint-etpro,snort-rule-subscriptions} given the
-amount of potential malicious traffic that exist on the Internet. A
+industry~\cite{proofpoint-etpro,snort-rule-subscriptions}. A
 \emph{rule-based intrusion detector}\MyComment{\footnote{\aka\ signature-based
@@
 describes a pattern in the traffic. The observation of that pattern in
-the traffic typically indicates an attempt of a malicious individual
-to attack a local network. This particular rule prescribes a method to
+the traffic indicates an attempt of a malicious individual
+to attack a local network. The example rule prescribes a method to
 capture an Active Reconnaissance attack for discovering accessible
-hosts in a network (see Section~\ref{sec:active-recon}). Rulesets
-exist for Snort~\cite{snort} and Suricata~\cite{suricata}, providing
-protection against existing attacks. Unfortunately, attackers actively
-look for new mechanisms to exploit systems through the
-network. Figure~\ref{fig:distribution-rules-per-month} shows the
-number of rules added per month in the year of 2019 to the ``Emerging
-Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
-data set of \suri\ rules. This data provides some evidence that new
-rules are created often.
-
-Intuitively, rulesets need to be continuously revised because the
-community finds and advertises vulnerabilities and attackers can take
-advantage of them. Unfortunately, creating these rules is
+hosts in a network (see Section~\ref{sec:active-recon}).
+
+Open-source \nids{}, like Snort~\cite{snort} and
+Suricata~\cite{suricata}, provide rulesets for many existing
+attacks. Unfortunately, attackers actively look for new mechanisms to
 \begin{wrapfigure}[8]{r}{0.21\textwidth}
@@
     }
-    \vspace{-4ex}
+    \vspace{-5ex}
     \caption{\label{fig:distribution-rules-per-month}2019 new rules.}
 \end{wrapfigure}
+exploit systems through the
+network. Figure~\ref{fig:distribution-rules-per-month} shows the
+number of rules added per month in the year of 2019 to the ``Emerging
+Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
+data set of \suri\ rules. A closer observation on the
+changelogs of this ruleset~\cite{emerging-threats-changelogs} shows that new rules are
+often created.  Rulesets need to be continuously revised because the
+community finds and advertises vulnerabilities and attackers can take
+advantage of them. Unfortunately, creating these rules is
 time-consuming and error-prone. The language for describing rules is
@@
 
-This paper proposes \tname{}\footnote{Abbreviation for
-  \textbf{Sy}nthesis of Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a
-technique that uses machine intelligence to automatically synthesize
-rules for rule-based \nids. Our goal is to facilitate the creation of
-rules and, consequently, help system administrators to harden the
-protection of \nids\ against network attacks.  \tname{} leverages
-different sources of information to synthesize rules for a given
-attack. It uses 1) positive examples, \ie{}, scenarios that reproduce
-the attack, 2) negative examples in the form of a data set of benign
-messages, and 3) a data set of human-written trusted rules. The
-positive and negative examples constrain the search space of rules
-that can be used to capture the attack. Intuitively, positive examples
-indicate what should be captured whereas the negative examples
-indicate what should not be captured. It is worth noting that vast
-amounts of negative data exist~\cite{tcpreplay,stratosphere-normal},
-but positive data is scarce. Consequently, classical machine learning
-techniques that require considerable amounts of training data is not
-directly applicable in this context. Furthermore, \tname\ uses public
-data sets of rules to differentiate rules by the similarity to
-human-written rules.
+This paper proposes \tname{} (for \textbf{Sy}nthesis of
+Su\textbf{ri}cata R\textbf{u}le\textbf{s}), a technique to
+automatically synthesize rules for rule-based \nids. Our goal is to
+facilitate the creation of rules and, consequently, help system
+administrators to harden the protection against network attacks.
+\tname{} leverages different sources of information to synthesize
+rules for a given attack. It uses 1) positive examples, \ie{},
+scenarios that reproduce the attack, 2) negative examples in the form
+of a data set of benign messages, and 3) a data set of human-written
+trusted rules. The positive and negative examples constrain the search
+space of rules that can be used to capture the attack. Intuitively,
+positive examples indicate what should be captured whereas negative
+examples indicate what should not be captured. It is worth noting that
+vast amounts of negative data
+exist~\cite{tcpreplay,stratosphere-normal}, but positive data is
+scarce as it is not common to publish attacks openly on the
+web. Classical machine learning techniques \emph{are not directly
+  applicable in this context} for two main reasons: 1)~they require a
+large amount of training data to build models (recall that positive
+data is scarce) and 2)~they typically have limited support to string
+and structured data, which are prevalent in this domain.
+
+%Furthermore, \tname\ uses
+%public data sets of rules to differentiate rules by the similarity to
+%human-written rules.
 
@@
 rulesets~\cite{emerging-threats-open} to create those rules. The
-technique is organized as three pipelined components. First,
+technique is organized as a pipeline of three components. First,
 \tname\ uses the messages that manifest an attack to create a
-potentially over-specified seed rule. Then, it uses benign traffic in
-to derive from the seed rule a set of plausible rules that isolate the
+potentially over-specified seed rule. Then, it uses that seed rule and
+benign traffic to derive a set of plausible rules that isolate the
 benign traffic from the malicious traffic. Finally, it uses existing
@@
 We evaluated \tname\ against 14 attacks of various kinds (see
-Table~\ref{table:attacks}). Results are encouraging--- \tname\ ranked
+Table~\ref{table:attacks}). Results are encouraging. For example, \tname\ ranked
 the correct rule among the top five rules in \percTopFiveRanking\ of
@@
 
-\sloppy Network Intrusion Detection Systems (\nids{}) are typically categorized in two
-groups~\cite{kumar2007survey}: rule-based, as briefly described on
-Section~\ref{sec:intro}, and anomaly-based \nids. An
-\emph{anomaly-based intrusion detector} learns usage behavior from
-benign traffic and, based on that, alerts uncommon
-behavior~\cite{7579764,kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}. More
-precisely, an anomaly-based \nids{} flags an attack when the network
-traffic manifests different characteristics compared to those of the
-benign traffic. Rule-based \nids{} and anomaly-based \nids{} are
-complementary. A rule-based \nids{} focuses on known attacks whereas
-anomaly-based \nids{} focuses on unknown potential attacks. Rule-based
-\nids\ can miss attacks as rulesets can become obsolete whereas
-anomaly-based \nids\ can report false alarms as learning algorithms
-that generalize observations from benign traffic are approximate.
-This paper focuses on rule-based \nids\ for its popularity
-in industry. Although we used Suricata, the
-principles we used are general to any other signature-based
-\nids~(\eg{}, Snort~\cite{snort}).
+\sloppy Network Intrusion Detection Systems (\nids{}) are typically
+categorized in two groups~\cite{kumar2007survey}: rule-based, as
+briefly described on Section~\ref{sec:intro}, and anomaly-based
+\nids. An \emph{anomaly-based intrusion detector} learns usage
+behavior from benign traffic and, based on that, alerts uncommon
+traffic~\cite{7579764,kumar2007survey,Mitchell:2014:SID:2597757.2542049,cordy-etal-issta19}. Rule-based
+\nids{} and anomaly-based \nids{} are complementary. A rule-based
+\nids{} focuses on known attacks whereas an anomaly-based \nids{}
+focuses on unknown potential attacks. Rule-based \nids\ can miss
+attacks as rulesets can become obsolete whereas anomaly-based
+\nids\ can report false alarms as learning algorithms that generalize
+observations from benign traffic are approximate.  This paper focuses
+on rule-based \nids\ for its popularity in industry. Although we used
+Suricata, the principles we used are general to any other
+signature-based \nids~(\eg{}, Snort~\cite{snort}).
+
+%%  It
+%% has been successfully used to identify denial-of-service
+%% attacks. 
+%% More
+%% precisely, an anomaly-based \nids{} flags an attack when the traffic
+%% on the network shows a different pattern compared with the historical
+%% traffic.
 
@@
 
-\vspace{-1ex}
+%\vspace{-1ex}
 \begin{table}[t!]
@@
   \end{tabular}
-%  \vspace{-3ex}  
+  \vspace{-3ex}  
 \end{table}
@@
   \includegraphics[scale=0.27]{figs/ICMP-packet-structure.png}
-  \vspace{-2ex}
+  \vspace{-1ex}
   \caption{Layout of an icmp packet.}
-  \vspace{-2ex}
+  \vspace{-1ex}
   \label{fig:icmp-packet-layout}
@@
   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
+  \vspace{-1ex}  
+  \caption{Suricata rule for Ping Scan attack.}
   \vspace{-2ex}  
-  \caption{Suricata rule for Ping Scan attack.}
-  \vspace{-4ex}  
   \label{fig:pingscan-example}
@@
 Figure~\ref{fig:pingscan-example} shows a Suricata rule that captures
-a Ping Scan, as documented in NMAP~\cite{netmap}, a network discovery
-and security auditing open-source tool. The essential options to
-capture the attack are in bold---the option \CodeIn{dsize: 0}
-indicates that the packet has no bytes in the payload and the option
-\CodeIn{itype: 8} indicates that the packet to be captured are ICMP
-Echo requests.
+Ping Scan, as documented on NMAP~\cite{netmap}, a network security
+auditing tool. The essential options to capture the attack are in
+bold---the option \CodeIn{dsize: 0} indicates that the packet has no
+bytes in the payload and the option \CodeIn{itype: 8} indicates that
+the packet to be captured are ICMP Echo requests.
 
@@
   \includegraphics[trim=0 340 50 0,clip,width=0.48\textwidth]{figs/nids-workflow}
-  \vspace{-4ex}  
+%  \vspace{-2ex}  
   \caption{The \tname\ workflow.}
   \label{fig:overview}
-  \vspace{-3ex}
+  \vspace{-4ex}
 \end{figure}
@@
 capture the traffic produced by the example attacks. However, that
-rule is over-constrained.  It may include options that are irrelevant
-to capture the attack and lead to false negatives---the \nids\ may be
+rule is over-constrained.  It may include options irrelevant to
+capture the attack and lead to false negatives---the \nids\ may be
 unable to capture manifestations of the same attack whose messages do
@@
 seed rule produced after the Reverse Engineering step on the Ping Scan
-example from Section~\ref{sec:active-recon} contains options
+example (Section~\ref{sec:active-recon}) contains options
 \CodeIn{icode:0}, \CodeIn{icmp\_id:23570}, and
@@
 
-In summary, this algorithm reports many rules that are plausible by
+In summary, the algorithm reports many rules that are plausible by
 using the benign traffic to reduce the over-constrained rule produced
@@
 \tname{} ranks the rules this algorithm reports by their similarity to
-existing trusted rules. Intuitively, rules that do not look like those
-trusted rules rank low.
-
-\begin{proposition}
-  \vspace{-1ex}  
-  The rule generation algorithm reports rules that capture only the
-  positive traffic.
-\end{proposition}
-\begin{proof}
-  \vspace{-1.5ex} Every rule created during the search captures the
-  attack provided as example on step one as every created rule
-  contains a proper subset of the options from the original rule and
-  all those options are, by definition, satisfied by the traffic
+trusted rules. Intuitively, rules dissimilar to those trusted rules
+rank low.
+
+\textbf{Correctness argument.}~The algorithm reports rules that
+capture only the positive traffic. Every rule created during the
+search captures the attack associated with the input as every created
+rule contains a proper subset of the options from the original rule
+and all those options are, by definition, satisfied by the traffic
 produced by the original attack. In addition, from the definition of
-  function $\mathit{isPlausible}$, reported rules capture \emph{only}
-  the positive traffic. If negative traffic is captured, a rule is not
-  added to $\mathcal R'$. Consequently, it cannot be an element of
-  $\mathcal O$ and reported on the output. Finally, from the
-  definition of function $\mathit{filter}$, only the rules in
-  $\mathcal O$ that capture all the instances of the attack (\ie{},
-  variants) are reported on output.
-  \vspace{-1ex}    
-\end{proof}
+$\mathit{isPlausible}$, reported rules capture \emph{only} the
+positive traffic. If negative traffic is captured, a rule is not added
+to $\mathcal R'$. Consequently, it cannot be an element of $\mathcal
+O$ and reported on the output. Finally, from the definition of
+function $\mathit{filter}$, only the rules in $\mathcal O$ that
+capture all the instances of the attack (\ie{}, variants) are reported
+on output.
 
@@
 \begin{table*}[t!]
-  \small
+  \footnotesize
   \caption{\label{table:attacks}List of attacks. Non-http attacks
@@
 
-\renewcommand{\arraystretch}{0.9}
-\setlength{\tabcolsep}{2pt}
-\begin{table}[h!]
-  \small
-  \caption{\label{table:results}Results.}
-  \vspace{-2ex}
-  \centering
-  \begin{tabular}{lrrrr}
-    \toprule
-    \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
-    \multicolumn{1}{c}{\multirow{2}{*}{\# Vars.}} &
-    \#~Option Cand. &
-    \#~Plausible Rules &    
-    \multicolumn{1}{c}{Rank} \\
-
-     &
-    \multicolumn{1}{c}{} &
-    \multicolumn{1}{c}{(step 1)} &
-    \multicolumn{1}{c}{(step 2)} &    
-    \multicolumn{1}{c}{(step 3)} \\
-
-    \midrule
-    Black Nurse & 1 & 4 & 9 & 2 \\    
-    Ping Scan & 6 & 5 & 9 & 2 \\
-    SYN Flood & 4 & 2 & 2 & 1 \\
-    UDP Teardrop & 1 & 4 & 8 & 4 \\
-    %%UDP Flood & - & - & - & - \\
-    \midrule
-    %%Apache Struts 2 & 2 & 65 & - & - \\
-    ColdFusion & 7 & 19 & 34 & 3\\
-    iisadmin access & 3 & 17 & 408 & 1 \\        
-    JBoss JMX & 5 & 23 & 228 & 45 \\
-    scripttag & 23 & 22 & 786 & 1 \\
-    WeBid LFI & 1 & 23 & 11,100 & 10\\    
-    %%Wordpress Login & 8 & 59 & - & - \\
-    .htaccess access & 5 & 17 & 984 & 2\\
-    .idq access & 14 & 19 & 511 & 2 \\
-    /system32/ in URI & 4 & 20 & 1,913 & 4 \\
-    Basebuilder LFI & 1 & 22 & 18,133 & 34 \\
-    Tomcat Snoop Access & 1 & 24 & 29,830 & 32 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
 %%\begin{table}[h!]
@@
 \label{attack-reproduction}
-
-This section describes the objects used in \tname{}'s evaluation.
-
-\vspace{1ex}
-\subsubsection{Attacks.}~Table~\ref{table:attacks} shows the selection
+\label{subsec:malicious-traffic}
+\label{sec:objectanalysis-benign}
+
+%This section describes the objects used in the evaluation.
+
+%\vspace{0.5ex}
+\noindent\textbf{Attacks.}~Table~\ref{table:attacks} shows the selection
 of attacks we analyzed.  Non-http attacks are listed at the top of the
@@
 (LFI)~\cite{local-file-inclusion}.
-
-%, and Policy Violation (PVE).
-
-\vspace{1ex}
-\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~Recall that \tname\ takes
+\noindent\textbf{Malicious traffic.}~Recall that \tname\ takes
 pcap files on input. As explained on Section~\ref{sec:minimization},
@@
 the SYN Flood attack (see Section~\ref{sec:dos}).
-
-\vspace{1ex}
-\subsubsection{\label{sec:objectanalysis-benign}Benign traffic.}~Public data sets of benign traffic
+\noindent\textbf{Benign traffic.}~Public data sets of benign traffic
 exist and can be used to evaluate the rate of false positives of
@@
 example, the regular traffic of a university network.
+\noindent\textbf{Ground Truth.}~To objectively evaluate \tname{}'s performance we
+compared the rules it generates with \emph{trusted rules}. We used
+the ``Emerging Threats Open Ruleset''~\cite{emerging-threats-open} for
+that. After finding a matching pair of attack and corresponding rule
+using the procedure described in
+Section~\ref{subsec:malicious-traffic}, we reproduced the attack to
+confirm the rule captures the attack in isolation (and no noise was
+introduced in the process). Column ``Ground Truth'' from
+Table~\ref{table:attacks} only shows the essential
+options. Documentation options (\eg{}, ``msg'') are discarded.
 
@@
 
-\subsubsection{Ground Truth.}To evaluate \tname\ objectively we need
-to compare the rules it generates with \emph{trusted rules}. We used
-the ``Emerging Threats Open Ruleset''~\cite{emerging-threats-open} for
-that. After finding a matching pair of attack and corresponding rule
-using the procedure described in
-Section~\ref{subsec:malicious-traffic}, we reproduced the attack to
-confirm the rule captures the attack in isolation (and no noise was
-introduced in the process). Column ``Ground Truth'' from
-Table~\ref{table:attacks} only shows the essential
-options. Documentation options (\eg{}, ``msg'') are discarded.
-
 \subsection{Answering RQ1: \textRQone}
@@
 proxy of human performance in analyzing the list of plausible rules.
+
+\renewcommand{\arraystretch}{0.7}
+\setlength{\tabcolsep}{2pt}
+\begin{table}[t!]
+  \footnotesize
+  \caption{\label{table:results}Results.}
+  \vspace{-2ex}
+  \centering
+  \begin{tabular}{lrrrr}
+    \toprule
+    \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
+    \multicolumn{1}{c}{\multirow{2}{*}{\# Vars.}} &
+    \#~Option Cand. &
+    \#~Plausible Rules &    
+    \multicolumn{1}{c}{Rank} \\
+
+     &
+    \multicolumn{1}{c}{} &
+    \multicolumn{1}{c}{(step 1)} &
+    \multicolumn{1}{c}{(step 2)} &    
+    \multicolumn{1}{c}{(step 3)} \\
+
+    \midrule
+    Black Nurse & 1 & 4 & 9 & 2 \\    
+    Ping Scan & 6 & 5 & 9 & 2 \\
+    SYN Flood & 4 & 2 & 2 & 1 \\
+    UDP Teardrop & 1 & 4 & 8 & 4 \\
+    %%UDP Flood & - & - & - & - \\
+    \midrule
+    %%Apache Struts 2 & 2 & 65 & - & - \\
+    ColdFusion & 7 & 19 & 34 & 3\\
+    iisadmin access & 3 & 17 & 408 & 1 \\        
+    JBoss JMX & 5 & 23 & 228 & 45 \\
+    scripttag & 23 & 22 & 786 & 1 \\
+    WeBid LFI & 1 & 23 & 11,100 & 10\\    
+    %%Wordpress Login & 8 & 59 & - & - \\
+    .htaccess access & 5 & 17 & 984 & 2\\
+    .idq access & 14 & 19 & 511 & 2 \\
+    /system32/ in URI & 4 & 20 & 1,913 & 4 \\
+    Basebuilder LFI & 1 & 22 & 18,133 & 34 \\
+    Tomcat Snoop Access & 1 & 24 & 29,830 & 32 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
 
-\setlength{\tabcolsep}{5pt}
+\setlength{\tabcolsep}{8pt}
 \begin{table}[h!]
-  \small
+  \footnotesize
   \caption{\label{table:weights}Contribution of heuristic functions.}
@@
 \begin{center}
-\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
-Every heuristic function contributed to the overall performance of
-\tname. The function that builds on the popularity of content
-modifiers was the one that affected positively the higher number of
-cases we analyzed.
+\begin{tcolorbox}[enhanced,width=3.4in,center upper,drop shadow southwest,sharp corners]
+Every function showed impact on the performance of \tname. The function that builds on popularity of content
+modifiers showed highest impact.
+%The function that builds on  was the one that affected positively the higher number of
+%cases we analyzed.
 \end{tcolorbox}
@@
 attacks---of the same kind---plays an important role in the
-performance of the technique.
-
-%%---and the case where \tname\ performs worst---jboss \Fix{Remover jboss}
-
-To answer this research question, we selected two cases with the
+performance of the technique. To answer this research question, we selected two cases with the
 highest number of examples (see column ``\#~Vars'' on
-Table~\ref{table:results})---idqaccess and scripttag. We measured the
+Table~\ref{table:results})---idqaccess and scripttag and measured the
 position of the golden rule at the rank for an increasing number of
@@
     \centering
-    \includegraphics[scale=0.4,trim=20 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
+    \includegraphics[scale=0.35,trim=20 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
     \caption{idqaccess}
@@
     \centering
-    \includegraphics[scale=0.4,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
+    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
     \caption{scripttag}    
   \end{subfigure}%
-\vspace{-2.5ex}  
+%\vspace{-2.5ex}  
 \caption{Impact of number of positive examples.}
 \label{fig:impact-number-attacks}
-\vspace{1ex}
+%\vspace{1ex}
 \end{figure}
@@
 \begin{center}
-\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
+\begin{tcolorbox}[enhanced,width=3.4in,center upper,drop shadow southwest,sharp corners]
 As expected, \tname\ is sensitive to the number and choice of example
@@
 
-This section elaborates how \tname\ performs on two other attacks.
-
-\textbf{Remote Code Execution (RCE).}  RCE refers to the family of
-attacks where hackers exploit a network vulnerability to run arbitrary
-code on a remote machine. The code to be executed could damage the
-system or steal sensitive information. JBoss~\cite{jboss} is a popular
-open source Java application server. This attack enables remote
-execution of code on the JBoss application server.
-
 %\Mar{@Lucas, por
@@
 
+\begin{figure}[t!]
+%  \vspace{-1ex}    
+  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
+%  \vspace{-2ex}  
+  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code
+    Execution.}
+%  \vspace{-1ex}
+\end{figure}
+
+\Mar{We may want to comment part of this to make room for the 
+  experiments that Lucas and Guilherme are currently running.
+  Side-comparison with prior work based on paper (not actual tool) results.}
+This section elaborates how \tname\ performs on two other attacks.
+\textbf{Remote Code Execution (RCE).}~RCE refers to the family of
+attacks where hackers exploit a network vulnerability to run arbitrary
+code on a remote machine. The code to be executed could damage the
+system or steal sensitive information. JBoss~\cite{jboss} is a popular
+open source Java application server. This attack enables remote
+execution of code on the JBoss application server.
 Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
@@
 this attack, the ground truth was located at position 45 in the ranked
-list.
-
-\begin{figure}[t!]
-%  \vspace{-1ex}    
-  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
-  \vspace{-2ex}  
-  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code
-    Execution.}
-%  \vspace{-1ex}
-\end{figure}
-
-\textbf{Denial-of-Service.}~
+list. \textbf{Denial-of-Service.}~
 The SYN Flood attack is a denial-of-service attack that exploits a
@@
 
-\vspace{-1ex}
+%\vspace{-1ex}
 \begin{figure}[h!]
   \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
-  \vspace{-2ex}
+%  \vspace{-2ex}
   \caption{\suri\ rule for SYN Flood attack.}
@@
 \end{figure}
-\vspace{-2ex}
+%\vspace{-2ex}
 
@@
 ranking in \percTopFiveRanking\ of the cases we analyzed. The data set
-containing our experimental artifacts is publicly accessible at \ourdataset.
-
-%% \section*{Acknowledgments}
-%% This work is partially supported by RNP grant number 002951.
+containing our experimental artifacts is publicly accessible at
+\ourdataset.
 
 \balance
-\bibliographystyle{ACM-Reference-Format}
+%\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{IEEEtran}
 \bibliography{references}
@@
 \end{document}
-\endinput
-%%
-%% End of file `sample-sigconf.tex'.
-
-%%  LocalWords:  Suricata SolarWinds QRadar Rulesets Ruleset rulesets
-%%  LocalWords:  pipelined pcap OISF IPS icmp msg classtype sid arg
-%%  LocalWords:  clp http ip bitvector flagbits IPs NMAP dsize itype
-%%  LocalWords:  icode ruleset th HtmlAdaptor pseudocode prce dataset
-%%  LocalWords:  flowbits len tcp uri RQ lrrrr Cand UDP ColdFusion
-%%  LocalWords:  iisadmin JBoss JMX WeBid LFI htaccess idq DoS AREC
-%%  LocalWords:  Basebuilder PESC RCE NetReseC Nikto WireShark hping
-%%  LocalWords:  Bigflows TcpReplay lrrrrrr boxplot Metasploit ACK
-%%  LocalWords:  Vollmer Hallahan Cordy NIDS

### 2020-05-12 paper/paper.tex
pass on section 2

@@
 
+%% \footnote{IPS (Intrusion Prevention System) and IDS differ in
+%%   the action taken when an intrusion is observed. IPS systems block
+%%   malicious traffic whereas IDS systems only alert them. However, IDS
+%%   and IPS are similar in the approaches used to detect
+%%   intrusions~\cite{ids-ips}.}
+
 Figure~\ref{fig:pingscan-example} shows an example rule of
@@
 options~\cite{suri-rule-format}. The \emph{action} appears as the
-first word in the rule description (an \CodeIn{alert}, in this
-case). An action denotes the task that needs to be executed if the
+first word in the rule description---an \CodeIn{alert}, in this
+case. An action denotes the task that needs to be executed if the
 rule pattern is satisfied.  In this example, an alert message will be
 sent to system administrators if the rule pattern is
-satisfied\footnote{IPS (Intrusion Prevention System) and IDS differ in
-  the action taken when an intrusion is observed. IPS systems block
-  malicious traffic whereas IDS systems only alert them. However, IDS
-  and IPS are similar in the approaches used to detect
-  intrusions~\cite{ids-ips}.}. The \emph{header} comes after the
+satisfied. The \emph{header} comes after the
 action in the rule description. It restricts the information flow
-covered by the rule. The header of this rule is \CodeIn{icmp any any
+monitored by the rule. The header of this rule is \CodeIn{icmp any any
   -> any any}. It instructs Suricata to inspect \CodeIn{icmp} traffic
@@
 the name of the option, and column ``Description'' summarizes
-purpose. The option \CodeIn{content} is common in rules to capture
-http attacks and http rules are the most common in the Emerging
-Threats Open Ruleset~\cite{emerging-threats-open}. The content option
-takes a sequence of bytes as argument and checks if that sequence is
-present in the payload of the message.  For example, the option
-\CodeIn{content:''SELECT''} checks if the string ``SELECT'' is present
-in the payload.  Detailed description of the Suricata rules can be
-found elsewhere~\cite{suri-rule-format}. It is worth noting that
-Suricata supports a total of \totoptions\ distinct options covering
-\numproto\ distinct network protocols.
+purpose. The option \CodeIn{content} deserves attention. It is common
+in rules to capture http attacks and http rules are the most common in
+the Emerging Threats Open Ruleset~\cite{emerging-threats-open}. The
+content option takes a sequence of bytes as argument and checks if
+that sequence is in the payload of the message.  For example, the
+option \CodeIn{content:''SELECT''} checks if the string ``SELECT'' is
+present in the payload.  Detailed description of the Suricata rules
+can be found elsewhere~\cite{suri-rule-format}. Suricata supports a
+total of \totoptions\ distinct options covering \numproto\ distinct
+network protocols.
 
@@
 
-We mentioned above that http rules often used content options, which
-associate with the message payload. For non-http rules, options are
-often associated with packet fields. Let us consider a rule for an
-icmp attack as example.
+%% We mentioned above that http rules often used content options, which
+%% associate with the message payload.
+
+We observed that the options used in non-http rules typically
+associate with packet fields. Let us consider a rule for an icmp
+attack as example.
 
@@
 Live'', declared in the ip header, are mapped to the \suri\ options
-\CodeIn{flagbits} and \CodeIn{ttl}, respectively. Likewise, the fields
+\CodeIn{fragbits} and \CodeIn{ttl}, respectively. Likewise, the fields
 ``Type'' and ``Code'', declared in the icmp header, are mapped to the
@@
 ``Flags'' the value of the corresponding option is encoded in a
-bitvector. For example, if the ``Flags'' field stores the value 1010,
-the corresponding rule option will be \CodeIn{flagbits:RM}, as each
+bitvector. For example, if the ``Flags'' field stores the value 101,
+the corresponding rule option will be \CodeIn{fragbits:RM}, as each
 position of the vector encodes a different character in the
-\CodeIn{flagbits} rule option. In summary, the value of several
+\CodeIn{fragbits} rule option. In summary, the value of several
 options can be determined through transformation functions, as those

### 2020-05-15 paper/paper.tex
pass

--- 2020-05-11 paper/paper.tex
@@
 
-\title{Learning to Synthesize Rules for Rule-based Intrusion Detectors}
+%
+% Add comments in the text
+%
+\newboolean{showcomments}
+\setboolean{showcomments}{true}
+%\setboolean{showcomments}{false}
+
+\ifthenelse{\boolean{showcomments}}
+  {\newcommand{\nb}[3]{
+  {\color{#2}\small\fbox{\bfseries\sffamily\scriptsize#1}}
+  {\color{#2}\sffamily\small$\triangleright~$\textit{\small #3}$~\triangleleft$}
+  }
+  }
+  {\newcommand{\nb}[3]{}
+  }
+
+\newcommand\rui[1]{\nb{Rui}{blue}{#1}}
+
+\title{Rule Synthesis for Rule-based Intrusion Detectors}
 
@@
 existing defenses. This paper proposes \tname{}, a novel approach to
-synthesize rules for Rule-based \nids. \tname\ leverages positive and
-negative traffic and examples of trusted rules to create
-rules. \tname{} is organized as a pipeline of three components that
+synthesize rules for \textit{rule-based} \nids. \tname\ leverages positive and
+negative traffic 
+\rui{positive and negative traffic are first introduced here, withuot being explaiend.} 
+and examples of trusted rules to create
+rules not previously captured. \tname{} is organized as a pipeline of three components that
 1)~create an over-specified seed rule, 2)~derive plausible rules from
 the seed, and 3)~rank plausible rules. Experimental results provide
-early evidence showing that \tname\ is promising. For example, it was
+evidence that \tname\ is promising. \rui{the example is interesting, but this gives the 
+feeling that the work is too preliminary} For example, it was
 able to synthesize the correct rule in all cases we analyzed and
@@
 often created.  Rulesets need to be continuously revised because the
-community finds and advertises vulnerabilities and attackers can take
-advantage of them. Unfortunately, creating these rules is
-time-consuming and error-prone. The language for describing rules is
+community finds and advertises vulnerabilities, which eventually can be exploited by attackers. 
+Unfortunately, creating these rules is
+time-consuming and error-prone \rui{is there a citation we could add here?}. The language for describing rules is
 non-trivial (see Section~\ref{sec:example-suricata-rules}) and testing
@@
 
-\tname{} takes on input one or more examples of an attack and produces
+\tname{} takes on input one or more examples \rui{We ought to explain how does one get hold on examples. How diffcult are they to obtain?} of an attack and produces
 on output a ranked list of rules for that attack. It uses public data
@@
 
-We evaluated \tname\ against 14 attacks of various kinds (see
-Table~\ref{table:attacks}). Results are encouraging. For example, \tname\ ranked
-the correct rule among the top five rules in \percTopFiveRanking\ of
-the cases and ranked the correct rule among the top ten rules in 11 of
-the 14 cases.
+We evaluated \tname\ against 14 attacks of various kinds, such as
+
+\rui{add few examples}. The observed results are encouraging. For 
+example, \tname\ ranked the correct rule among the top five rules in 
+\percTopFiveRanking\ of the cases and ranked the correct rule among 
+the top ten rules in 11 of the 14 cases. \rui{to discuss, I find 
+this a bit weak.}
 
@@
 
-Table~\ref{table:rules} shows a sample of non-documentation options of
+Table~\ref{table:rules} shows a sample of non-documentation options \rui{what is the meaning of non-documentation in this context?} of
 Suricata. Column ``\#'' shows the option number, column ``Name'' shows
@@
 
-Figure~\ref{fig:pingscan-example} shows a Suricata rule that captures
+Figure~\ref{fig:pingscan-example} \rui{any reason for the bolds?} shows a Suricata rule that captures
 Ping Scan, as documented on NMAP~\cite{netmap}, a network security
@@
 \noindent
-, where $\mathit{content_i}$ denotes the $i$th content in a given
+where $\mathit{content_i}$ denotes the $i$th content in a given
 rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
@@
 \noindent
-, where $g(i)$ is the contribution of content option $i$ to the score,
+where $g(i)$ is the contribution of content option $i$ to the score,
 $N$ is the number of contents options present in a rule,
@@
 \textbf{Rationale.}~The golden rule is a trusted human-written
-\nids\ rule that we obtained from credible sources. \tname\ generates
+\nids\ rule that we obtained from credible sources \rui{which are?}. \tname\ generates
 several rules that satisfy the invariant to only capture the malicious
@@
 \newcommand{\textRQthree}{How sensitive \tname\ is to the amount of
-  positive examples?}
+  positive examples?} \rui{should we also investigate the impact of the negative ones?}
 %\vspace{0.2cm}
@@
 
+\rui{thinking out loud here: one of the motivations is that it is time-consuming
+and troublesome to generate new rules. Time-consuming is not being addressed in 
+the RQs and I think that, albeit it is difficult to conduct such a study, we 
+shuold discuss it. I would think that mentinoing the human-on-the-loop (not in-the-loop)
+should also be dscussed (this is more about the troublesomeness)}
+
 \balance

### 2020-05-15 paper/FSE_reviews_rebuttal.txt

We sincerely hope that reviewers can carefully read this rebuttal as some comments revealed important misconceptions.

R1: “manually capture the options from the example attack to form the seed rule”

The first step is an automatic extraction of options from the attack message, generating the seed rule. The option-field mapping was written manually, but as this is protocol information, there’s no need for rewriting it.

R1: “- the topic is not very relevant”

Please consider that (a) Intrusion Detection is a multi-million dollar business (type “IDS solutions” on Google), (b) detecting intrusion with rule-based IDS is the most commonly-adopted solution for intrusion detection in industry, and (c) generating these rules is a very common activity in this business. We explained this on intro, providing statistics.

R1: on bias toward existing rules. (...In my opinion, as a rule learner, it is better to learn new rules.)

This comment was unclear to us. The goal of Syrius is to create new rules. Rules that were not seen before to help the Suricata/Snort community. This goal is reflected in our experimental setup.

R1 and R2: approach is not general enough

As mentioned above, synthesizing rules is a challenging problem. Our goal is to handle most cases--and we provided statistical data showing that we do handle. The more rules that can be generated the better. It is unclear to us how general is “general enough” (as per comment), especially considering the fundamental limitation of existing learning algorithms.

R2: “...[Syrius] does not guarantee correctness, which is needed in the context of IDS.”

Even manual generation of rules can produce rules with false positives and negatives.  Classical ML-based approaches for Anomaly-based Intrusion Detection suffer from this problem the same way we do. Having said that, Syrius does guarantee correctness (as per false positives and negatives counts) relative to the very large dataset we use.

R2. “...what happens if the set of common options is empty? “ How the threshold would help in this case?”

If the set of common options is empty, the input is invalid for Syrius, and the threshold option is not suitable, since this option aims to count the occurrences of packets that have common characteristics.

R2) How do you define N in line 6 of algorithm 1? Why not considering all possible rules that canoe generated from the seed rule?

Using smaller values for N is a choice purely computational. Making N=number_of_options would find all possible rules, but also would take more time. We chose 2 as value for N as it was enough for finding the “Golden Rule”.

R2 and R3: “...No empirical comparison with other existing methods.”
R3: “...the authors could apply their own system to the same dataset(s) and compare to the published results.”

We admit this is a problem. We could not access alternative tools. We think that to fairly evaluate our tool against others we would need to access and run alternative tools. This would enable us to confirm that these tools actually produce the results they report.

R3. on Significance “Much of the recent research in this area applies real machine learning to NIDS and generates classifiers for malicious vs. benign.”

We know. We explained the rationale for using a non-traditional “learning” mechanism. Existing ML approaches deal well with numerical and categorical data, but fail to handle structured and string data, which are central to rule synthesis. Packets contents (the most common option in rules) can potentially have an infinite number of permutations [https://arxiv.org/pdf/1903.12101]. Anomaly-based NIDS monitor features like ratio of accesses by a given protocol that can be modeled with numerical features. Rule-base NIDS are complementary to anomaly-based NIDS.

R3. “...Please explain this, I don't understand how 1010 maps to RM”

typos:
lines 207, 214, 215: flagbits -> fragbits
line 213: 1010 -> 101

The fragbits option maps each bit of the Flags field to a character M, D or R, where M represents the most and R the least significant. So 101 means that the bits M and R are set and that the bit D is not, therefore the fragbits value is MR (or RM).

-------------------------------------
Dear Marcelo d'Amorim,

Thank you for your submission to ESEC/FSE 2020. The HotCrp site is now open
for rebuttal (https://esecfse2020.hotcrp.com) The rebuttal site will be
open until April 30th, 23:59 AoE.

During this time, you will have access to the current state of your reviews
and have the opportunity to submit a response of up to 750 words. This is a
soft limit, you may submit longer responses, however, reviewers are not
required to read beyond the 750 word limit.

Please keep in mind the following during this process:

* The response must focus on any factual errors in the reviews and any
questions posed by the reviewers. It must not provide new research results
or reformulate the presentation. Try to be as concise and to the point as
possible.

* The rebuttal period is an opportunity to react to the reviews, but not a
requirement to do so. Thus, if you feel the reviews are accurate and the
reviewers have not asked any questions, then you do not have to respond.

* The reviews are submitted by the PC members, without any coordination
between them. Thus, there may be inconsistencies. Furthermore, these are
not the final versions of the reviews. The reviews can be updated later to
take into account the discussions among the program committee, and we may
find it necessary to solicit additional reviews after the rebuttal period.

* Your response will be seen by all PC members who have access to the
discussion of your paper, so please try to be polite and constructive.

The reviews and scores for your paper are attached to this email. To submit
your response you should log on the HotCrp website and select your
submission on the menu. If you decide to withdraw your paper, please edit
the submission in HotCrp to mark the paper as withdrawn.

       Title: Learning to Synthesize Rules for Rule-based Intrusion
              Detectors
       Site: https://esecfse2020.hotcrp.com/paper/909

Best Regards,
Myra Cohen and Tom Zimmermann

--
Review #909A
===========================================================================

Overall merit
-------------
2. Weak reject

Paper summary
-------------
This paper proposes to learn the rules for rule based intrusion detectors. It proposes a three step process for this purpose: 1) manually capture the options from the example attack to form the seed rule; 2) create the variants of the seed rule by doing random mutations starting from the seed rule; and 3) ranking the obtained variants. The benign examples are used filter the negative examples in step 2 for making the mutation minimal. Some experiments have been conducted for the evaluation.

Strengths and weaknesses
------------------------
Strengths:
- the process is complete
- have used the well-known attack database

Weaknesses:
- the topic is not very relevant
- it does not seem to be a general method, the learning process is case by case and seems to depend heavily on the manual operation step.

Comments for the authors
------------------------
Soundness:
This paper introduces a three-step process for learning the rules for the rule-based intrusion detectors. But there is no methodological guidance for this process. For example, the first step is manually extracting the options from the attack message. There is not any guidance for the option extraction. In particular, the seed rule extracted by this process will directly affect the rules generated in the second step. Another questions about the soundness are the heuristics for ranking the generated rules. First, why the similarity between the generated rule and the existing rule is a suitable criteria? Is this more biased towards existing rules than newly discovered ones? Is it reasonable to define similarity as the four indicators in the section 4.3?

Significance:
As mentioned before, the ranking strategy proposed in this paper seems to be more biased towards learning existing rules. Is it right? In my opinion, as a rule learner, it is better to learn new rules. Learning new rules  might be more valuable and contributes more to the community. Of course, the newly-learning rule needs to be confirmed later.

Novelty:
There are some other strategy that the idea is similar to this work, e.g. genetic algorithm based rule learning. It seems having larger search space for rule learning than the proposal in this paper.

Recoverability, Replicability and Reproducibility:
As mentioned before, the learning result depends on the manual extraction of the seed rule. And also, it does not tell the formulation of the input message and fails to tell how to extract the necessary information from the input.

Presentation:
The presentation is too detailed. The operation process is case-by-case. It lacks of refinement, generalization,  and formalization.

* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *

Review #909B
===========================================================================

Overall merit
-------------
2. Weak reject

Paper summary
-------------
The paper proposes a method to learn rules for a rule-based network intrusion detection system (NIDS). The method is a pipeline of three main steps: reverse engineer of an over-specified rule from a given example of attack, then synthesis new rules that are possibly more general but at he same time are guaranteed to reflect benign behaviour, i.e. not to have false positives. The final set of constructed rules, referred to as plausible rules, are then ranked using heuristics that compare the level of similarity between each of these rule with ground truth rules defined by humans. Experimental results show that the proposed approach is capable of ranking 70% of the time the correct rule among top five plausible rules.

Strengths and weaknesses
------------------------
Strength:
+ well written paper with a clear simplified example at the beginning
+ experimental results show advantage of the approach

Weaknesses:
- The approach is not general enough
- Use of heuristics does not guarantee correctness, which is needed in the context of IDS.
- Heuristics are dataset dependent
- No empirical comparison with other existing methods.

Comments for the authors
------------------------
The problem of network intrusion detection is very relevant in the SE community and automated learning of intrusion detection rules is an open problem. The paper is well written and self-contained, with a very clear illustrative example that successfully introduce the main steps of the approach. I have, however, various concerns which stop me from recommending acceptance.

The approach is more a form of rule synthesis than learning. I would change the title of the paper to “Rule Synthesis for Rule-Based Intrusion Detection” and remove the word “learning” as the learning process is in tshie case very minimal.

Firstly, the authors state that the input to the approach is a set of examples of attacks. The rule creation algorithm however starts from a over-specialised rule that is reverse engineered from a single attack example. So it is not clear how the other maligned attack examples (mTrs) are used during the generalisation process. It is, on the other hand, clear how the benign traffics (bTrs) are used during the second phase of the approach.  Regarding the novelty of the algorithm, the idea of creating rules starting from an over-specialised rule is very similar to what well established symbolic rule learning methods have proposed in the literature. The “rule creation” step would correspond to a process of generalisation which, in the context of symbolic rule learning, guarantee to be sound and able to learn the most general rule, instead of an approximating it through heuristics as proposed in this paper. I don’t really see the advantage of using this approach instead of well known sound symbolic rule learning methods. Examples of symbolic rule learning approaches are those algorithms and systems that follow the “inverse entailment” learning method [1].

As for the significance, the heuristics seem to be dependent on the type of dataset and positive and negative examples considered by the algorithm. This makes the approach even more dataset specific. Changing type of datasets might require changing the heuristics as there is no evidence to believe that the approach will perform well. The authors claim that the approach is applicable to any IDS type, including Snort. It would’ve been good to show that indeed this is the case. The lack of empirical evaluation on this important aspect of the work make the significance of this work very limited.

As for the soundness:

- the outcomes of the algorithm is a rank of rules. This is not necessarily ideal as in practice an IDS has to use specific rules that correctly recognise attacks. Related to my first point, the metric used for evaluating the outcomes is not in practice very useful. The results in Table 3 seem to show that in the case of very limited number of variation of an attack and high number of seed options many plausible rules are generated and the correct rule results to be quite low in the ranking.

- although I do realise that starting the creation of rules from an over-specialised rule is computationally better than starting from just the attack example, the number of plausible rules does grow wrt the number of options included in the seed rule, affecting the level of ranking of the correct rules.

- it is also not clear if the “second step” of the algorithm works though a "tree of plausible rules" or a “lattice” of plausible rules. They can be many “incomparable” plausible rules in terms of number of options used as conditions.

- It is also not clear how the value of “N” is picked in the algorithm and if the algorithm does indeed expire all the possible plausible rules.

-  I’m not sure if the proof of Proposition 4.1 captures the statement of the Proposition. It seems to be stating the opposite of what has been defined to be the output of the algorithm.

Reference:
[1] “Inverse Entailment and Progol”, S. Muggleton, New Generation Computing 13 (1995), 245-286.

Typo:
1) Include in Table 1 the description for ttl.
2) Line 207: use “flag” instead of “flagbits’ to match with the name given in Table  or change the name in Table 1.
3) Line 505: Raking —> Ranking

Questions for the authors’ response
-------------------------------------
1) On page 4, line 396 the authors state “Syrius runs the same approach for each message and identifies common options across all messages”. But what happens if the set of common options is empty? How the threshold would help in this case?

2) How do you define N in line 6 of algorithm 1? Why not considering all possible rules that canoe generated from the seed rule?

* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *

Review #909C
===========================================================================

Overall merit
-------------
1. Reject

Paper summary
-------------
The paper proposes an approach to generating rules for signature-based network intrusion detection systems (NIDS).  Since there are relatively few examples of attacks relative to benign messages, the authors claim that machine learning is not applicable. So, despite the paper's title, the approach does not involve 'learning' in the machine learning sense.  Instead, the tool (called Syrius) goes through three stages: use the available attack messages to create an overly specific seed rule; use benign messages to derive a set of rules from that seed that still capture the attack but do not capture the benign messages; rank the set of rules according to their 'similarity' to human-written rules using heuristics such as frequently using those options that occur most commonly in human-written rules (options generally correspond to fields in network packets or content patterns in http payloads).  The rules in this case are targeted to Suricata, an open-source NIDS, but the approach is said to be applicable to other NIDS where a large number of sample human-written rules are available.

Strengths and weaknesses
------------------------
strengths:

The three-stage pipeline is an interesting approach.

The research questions are good (although they did not go far enough, as noted below there is no experimental comparison to alternative techniques).

It is good that the paper includes an example where Syrius performed poorly (section 6).  I wish more papers included such examples!

weaknesses:

There's been a lot of research on automatically generating rules for signature-based network intrusion detection, but the paper cites only three relevant papers, references 30, 33 and 35, the most recent of which is from 2015.  Some examples of more recent related work that generates rules: https://arxiv.org/abs/1903.12101, https://ieeexplore.ieee.org/abstract/document/7585840, https://ieeexplore.ieee.org/abstract/document/7544894.

The experimental evaluation (section 5) does not compare Syrius directly to any other work. The paper says that the implementations of related work are not available (1088-1090). However, much of the research in this area reports evaluation results on public datasets, so even when the tool implementation is not available, the authors could apply their own system to the same dataset(s) and compare to the published results.

The title and body of the paper often use the terms 'learning' and 'learn' in a misleading way, since the technical details in section 4 do not involve machine learning.

Comments for the authors
------------------------
soundness:

The heuristics used for ranking candidate rules (section 4.3) are not well-justified, they seem to be chosen based on what is easy to calculate.  This is mitigated to some extent by RQ2.

"Considering construct validity, the choice of metric used to evaluate the approach may not be applicable in practice. We considering the use of ranking of the golden rule a reasonable choice as the approach places the golden rule close to the top in most cases." (lines 1023-1028) This is a self-serving argument, not an appropriate mitigation.

significance:

Much of the recent research in this area applies real machine learning to NIDS and generates classifiers for malicious vs. benign.  (There is also a lot of work on embodying NIDS in FPGA, since software NIDS is deemed too slow.)

novelty:

As noted in "weaknesses" above, there are other techniques that may be as good or better for generating rules for signature-based NIDS, but the paper does not evaluate in comparison.

recoverability, replicability, reproducibility:

A public implementation is available by request.  The authors should post a public link in the final version of the paper.

The dataset is already publicly available.

presentation:

Section 3 should include the actual generated seed rule and the 9 generated plausible rules, in the generated rank order, for the PingScan example. The authors could also add some examples of discarded rules.

The first part of section 4 repeats some of the material in section 3.

The term "golden rule", used throughout the paper, is not defined until section 5.

other:

The paper has no keywords.

It is odd that 25 of the paper's 36 references are to (non-research) websites. Many of these might be better provided as footnotes rather than cites.

typos:

line 505: Raking -> Ranking

line 748: sensible -> sensitive

line 1018: is -> are

Questions for the authors’ response
-------------------------------------
"For example, if the “Flags” field stores the value 1010, the corresponding rule option will be flagbits:RM, as each position of the vector encodes a different character in the flagbits rule option." (lines 211-215) Please explain this, I don't understand how 1010 maps to RM.

### 2020-05-18 paper/paper.tex
pass

@@
 existing defenses. This paper proposes \tname{}, a novel approach to
-synthesize rules for \textit{rule-based} \nids. \tname\ leverages positive and
-negative traffic 
-\rui{positive and negative traffic are first introduced here, withuot being explaiend.} 
+synthesize rules for \textit{rule-based} \nids. \tname\ leverages
+malicious (positive) and benign (negative) traffic 
+%\rui{positive and negative traffic are first introduced here, withuot being explaiend.} 
 and examples of trusted rules to create
-rules not previously captured. \tname{} is organized as a pipeline of three components that
+rules for new attacks. \tname{} is organized as a pipeline of three components that
 1)~create an over-specified seed rule, 2)~derive plausible rules from
 the seed, and 3)~rank plausible rules. Experimental results provide
-evidence that \tname\ is promising. \rui{the example is interesting, but this gives the 
-feeling that the work is too preliminary} For example, it was
+evidence that \tname\ is promising.
+ \rui{the example is interesting, but this gives the 
+feeling that the work is too preliminary} 
+For example, it was
 able to synthesize the correct rule in all cases we analyzed and
@@
 the traffic indicates an attempt of a malicious individual
-to attack a local network. The example rule prescribes a method to
+to attack a local network. This particular rule prescribes a method to
 capture an Active Reconnaissance attack for discovering accessible
@@
 changelogs of this ruleset~\cite{emerging-threats-changelogs} shows that new rules are
-often created.  Rulesets need to be continuously revised because the
-community finds and advertises vulnerabilities, which eventually can be exploited by attackers. 
+often created.  Rulesets need to be continuously revised as the
+community finds and advertises vulnerabilities, which eventually can be exploited by hackers. 
 Unfortunately, creating these rules is
@@
 
-This paper proposes \tname{} (for \textbf{Sy}nthesis of
+\sloppy This paper proposes \tname{} (for \textbf{Sy}nthesis of
 Su\textbf{ri}cata R\textbf{u}le\textbf{s}), a technique to
@@
 exist~\cite{tcpreplay,stratosphere-normal}, but positive data is
-scarce as it is not common to publish attacks openly on the
-web. Classical machine learning techniques \emph{are not directly
+scarce as it is not common (and considered unethical~\Mar{look for
+  citation for ethics code on this}) to publish attacks on the
+web\Mar{is there a term for non-deep web?}. Classical machine learning techniques \emph{are not directly
   applicable in this context} for two main reasons: 1)~they require a
@@
 
-We evaluated \tname\ against 14 attacks of various kinds, such as
-
-\rui{add few examples}. The observed results are encouraging. For 
+We evaluated \tname\ against \Mar{@Lucas, please look for constant
+  numbers as this *in text* and use macros instead. We do not want to
+  leave inconsistent data in the paper!}14 attacks of various kinds, such as \Fix{...}\rui{add few examples}. The observed results are encouraging. For 
 example, \tname\ ranked the correct rule among the top five rules in 

### 2020-05-18 paper/paper.tex
pass on section 1 and 2

@@
   \footnotesize
-  \caption{\label{table:rules}Example of non-documentation options of Suricata.}
+  \caption{\label{table:rules}Example of capture-relevant options of Suricata.}
   \vspace{-2ex}
@@
 
-Table~\ref{table:rules} shows a sample of non-documentation options \rui{what is the meaning of non-documentation in this context?} of
+Table~\ref{table:rules} shows a sample of capture-relevant options of
 Suricata. Column ``\#'' shows the option number, column ``Name'' shows
 the name of the option, and column ``Description'' summarizes
-purpose. The option \CodeIn{content} deserves attention. It is common
-in rules to capture http attacks and http rules are the most common in
-the Emerging Threats Open Ruleset~\cite{emerging-threats-open}. The
-content option takes a sequence of bytes as argument and checks if
-that sequence is in the payload of the message.  For example, the
-option \CodeIn{content:''SELECT''} checks if the string ``SELECT'' is
-present in the payload.  Detailed description of the Suricata rules
-can be found elsewhere~\cite{suri-rule-format}. Suricata supports a
-total of \totoptions\ distinct options covering \numproto\ distinct
-network protocols.
+purpose. The option \CodeIn{content} deserves special attention. It is
+common in rules to capture http attacks and http rules are the most
+common in the Emerging Threats Open
+Ruleset~\cite{emerging-threats-open}. The content option takes a
+sequence of bytes as argument and checks if that sequence appears in
+the payload of the message.  For example, the option
+\CodeIn{content:''SELECT''} checks if the string ``SELECT'' is present
+in the payload.  Detailed description of the Suricata rules can be
+found elsewhere~\cite{suri-rule-format}. Suricata supports a total of
+\totoptions\ distinct options covering \numproto\ distinct network
+protocols.
 
@@
 
-\subsection{Rules and Packets}
+\subsection{Non-http Rules and Packets}
 \label{sec:rules-and-packets}
@@
 
-We observed that the options used in non-http rules typically
-associate with packet fields. Let us consider a rule for an icmp
-attack as example.
+Options of non-http rules typically associate with packet fields. Let
+us consider a rule for an icmp attack as example.
 
@@
 position of the vector encodes a different character in the
-\CodeIn{fragbits} rule option. In summary, the value of several
-options can be determined through transformation functions, as those
-informally described above. One exception to this is the option
+\CodeIn{fragbits} option.
+
+In summary, the value of several options \emph{of non-http rules} can
+be determined through transformation functions, as those informally
+described above. One exception to this is the option
 \CodeIn{threshold}, which is typically used in denial-of-service

### 2020-05-18 paper/paper.tex
pass

--- 2020-05-18 paper/paper.tex
@@
 
-Table~\ref{table:rules} shows a sample of non-documentation options \rui{what is the meaning of non-documentation in this context?} of
-Suricata. Column ``\#'' shows the option number, column ``Name'' shows
-the name of the option, and column ``Description'' summarizes
+Table~\ref{table:rules} shows a sample of non-documentation options \rui{what 
+is the meaning of non-documentation in this context?} \Luc{documentation options
+are options that gives information about the rule itself and are not used for 
+matching on packet data. So, non-documentation options are the ones that really
+matters for detection.} of Suricata. Column ``\#'' shows the option number, column 
+``Name'' shows the name of the option, and column ``Description'' summarizes
 purpose. The option \CodeIn{content} deserves attention. It is common

### 2020-05-18 paper/paper.tex
moved Rui's newcommand from paper to macros, changed color
pass

@@
   }
-
-\newcommand\rui[1]{\nb{Rui}{blue}{#1}}
 
@@
 
-Figure~\ref{fig:pingscan-example} \rui{any reason for the bolds?} shows a Suricata rule that captures
+Figure~\ref{fig:pingscan-example} \rui{any reason for the bolds?} \Luc{these are
+the options that are important for detection, the other ones are documentation
+options} shows a Suricata rule that captures
 Ping Scan, as documented on NMAP~\cite{netmap}, a network security

### 2020-05-18 paper/paper.tex
pass on section 3 (halfway through)

--- 2020-05-18 paper/paper.tex
@@
 
-Figure~\ref{fig:pingscan-example} \rui{any reason for the bolds?} shows a Suricata rule that captures
+Figure~\ref{fig:pingscan-example} shows a Suricata rule that captures
 Ping Scan, as documented on NMAP~\cite{netmap}, a network security
@@
 
-In the following, we briefly show how \tname{} synthesizes rules for
-this attack. In short, \tname{} takes one or more examples of an
-attack---in the form of network messages---and produces rules to
-capture that attack. Considering the rule for the Ping Scan attack,
-whose example was reproduced with NMAP, \tname{} leaves the action of
-the corresponding rule constant (as \CodeIn{alert}) and infers the
-header by checking the protocol of the input messages. For example, in
+%---in the form of network messages---
+% one or more examples of an attack
+%, indicating that the rule applies to \CodeIn{icmp}messages
+
+In the following, we show how \tname{} synthesizes rules for this
+attack. \tname{} takes on input benign and malicious traffic and
+produces on output rules to capture only the malicious
+traffic. Considering the Ping Scan attack, \tname{} replays the
+attack---with NMAP, in this case---and monitors the traffic.  It
+infers the header by checking the protocol of the input messages. In
 this case, \tname{} sets the header to \CodeIn{icmp any any -> any
-  any}, indicating that the rule applies to \CodeIn{icmp} messages and
-the most general flow description, \ie{}, it instructs the \nids\ to
-analyze the icmp traffic flowing from/to any address and port.  The
-rationale for this decision is that the choice of these parameters are
-domain-specific and subject to change by the system administrator.
+  any}. It uses the most general flow description instructing the
+\nids\ to analyze the icmp traffic flowing from/to any address and
+port.  The rationale is that the choice of these parameters are
+domain-specific and subject to change by the sys admins. Likewise,
+\tname{} leaves the action of the corresponding rule constant, as
+\CodeIn{alert}.
+
+\Mar{--------------------- I am here...}
 

### 2020-05-18 paper/paper.tex
pass
added new attacks on table ii

--- 2020-05-18 paper/paper.tex
@@
     UDP Teardrop & \cite{udp-teardrop-source} & DoS & udp & \textbf{fragbits}:M; \textbf{id}:124; \\
+      2100384 & - & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:8; \\
+      2100399 & - & ICMP Info & icmp & \textbf{icode}:1; \textbf{itype}:3; \\
+      2100401 & - & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:3; \\
+      2100402 & - & ICMP Info & icmp & \textbf{icode}:3; \textbf{itype}:3; \\
+      2100408 & - & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:0; \\
+      2100485 & - & ICMP Info & icmp & \textbf{icode}:13; \textbf{itype}:3; \\
+      2100486 & - & ICMP Info & icmp & \textbf{icode}:10; \textbf{itype}:3; \\
     %%UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5; \\
@@
 \textbf{Rationale.}~The golden rule is a trusted human-written
-\nids\ rule that we obtained from credible sources \rui{which are?}. \tname\ generates
+\nids\ rule that we obtained from credible sources \rui{which are?}\Luc{\cite{emerging-threats-open}}. \tname\ generates
 several rules that satisfy the invariant to only capture the malicious

### 2020-05-18 paper/paper.tex
added new attacks on table iii

@@
     UDP Teardrop & 1 & 4 & 8 & 4 \\
+    2100384 & 13 & 8 & 18 & 2 \\
+    2100399 & 132 & 6 & 3 & 1 \\
+    2100401 & 12 & 6 & 25 & 1 \\
+    2100402 & 3 & 6 & 12 & 1 \\
+    2100408 & 15 & 8 & 9 & 2 \\
+    2100485 & 4 & 6 & 3 & 1 \\
+    2100486 & 195 & 6 & 2 & 1 \\
     %%UDP Flood & - & - & - & - \\

### 2020-05-18 paper/paper.tex

@@
 
-Table~\ref{table:rules} shows a sample of non-documentation options \rui{what 
-is the meaning of non-documentation in this context?} \Luc{documentation options
-are options that gives information about the rule itself and are not used for 
-matching on packet data. So, non-documentation options are the ones that really
-matters for detection.} of Suricata. Column ``\#'' shows the option number, column 
-``Name'' shows the name of the option, and column ``Description'' summarizes
-purpose. The option \CodeIn{content} deserves attention. It is common
-in rules to capture http attacks and http rules are the most common in
-the Emerging Threats Open Ruleset~\cite{emerging-threats-open}. The
-content option takes a sequence of bytes as argument and checks if
-that sequence is in the payload of the message.  For example, the
-option \CodeIn{content:''SELECT''} checks if the string ``SELECT'' is
-present in the payload.  Detailed description of the Suricata rules
-can be found elsewhere~\cite{suri-rule-format}. Suricata supports a
-total of \totoptions\ distinct options covering \numproto\ distinct
-network protocols.
+Table~\ref{table:rules} shows a sample of capture-relevant options of
+Suricata. Column ``\#'' shows the option number, column ``Name'' shows
+the name of the option, and column ``Description'' summarizes
+purpose. The option \CodeIn{content} deserves special attention. It is
+common in rules to capture http attacks and http rules are the most
+common in the Emerging Threats Open
+Ruleset~\cite{emerging-threats-open}. The content option takes a
+sequence of bytes as argument and checks if that sequence appears in
+the payload of the message.  For example, the option
+\CodeIn{content:''SELECT''} checks if the string ``SELECT'' is present
+in the payload.  Detailed description of the Suricata rules can be
+found elsewhere~\cite{suri-rule-format}. Suricata supports a total of
+\totoptions\ distinct options covering \numproto\ distinct network
+protocols.
 

### 2020-05-18 paper/paper.tex
pass

@@
 
-Figure~\ref{fig:pingscan-example} \rui{any reason for the bolds?} \Luc{these are
-the options that are important for detection, the other ones are documentation
-options} shows a Suricata rule that captures
+Figure~\ref{fig:pingscan-example} shows a Suricata rule that captures
 Ping Scan, as documented on NMAP~\cite{netmap}, a network security
@@
 In the following, we show how \tname{} synthesizes rules for this
-attack. \tname{} takes on input benign and malicious traffic and
-produces on output rules to capture only the malicious
+attack. The main challenge in synthesizing rules is to discover the
+set of rule options (see Table~\ref{table:rules}) that captures the
+malicious traffic, but allows the benign traffic to pass. \tname{}
+proceeds as follows to synthesize rule options.
+
+%% \tname{} takes on input benign and malicious traffic and
+%% produces on output rules to capture only the malicious
+%% traffic.
+
+First, it analyzes the malicious traffic to extract rule's action,
+header, and \emph{initial} options expressed in the
 traffic. Considering the Ping Scan attack, \tname{} replays the
@@
 \tname{} leaves the action of the corresponding rule constant, as
-\CodeIn{alert}.
-
-\Mar{--------------------- I am here...}
-
-The main challenge in synthesizing rules is to discover the set of
-rule options (see Table~\ref{table:rules}) that captures the malicious
-traffic, but allows the benign traffic to pass. \tname{} proceeds as
-follows to synthesize rule options. First, it analyzes the malicious
-traffic to extract options expressed in that traffic. Recall from
+\CodeIn{alert}. Furthermore, recall from
 Section~\ref{sec:rules-and-packets} that some options can be inferred
@@
 that the number of plausible rules that \tname\ produces at this step
-can be very high for other example attacks (see column ``\#~Plausible
-Rules'' on Table~\ref{table:results}).
+can be very high for other example attacks. For illustration, see
+column ``\#~Plausible Rules'' on Table~\ref{table:results}.
 
@@
 human-written rules. For example, one heuristic function used by
-\tname{} leverages the popularity of options for ranking.  The score
-of a rule $r$, according to that heuristic alone, is proportional to
-the popularity of the options declared in
+\tname{} leverages the popularity of options for ranking. According to
+that heuristic alone, the score of a rule $r$ is proportional to the
+popularity of the options declared in
 $r$. Figure~\ref{fig:distribution-options} shows the distribution of
@@
 \CodeIn{itype}, the ones associated with the golden rule, appear as
-the 6th and 7th most popular, respectively. But, among the options
-inferred by \tname{} at the first step, \CodeIn{dsize} and
+the 6th and 7th most popular, respectively. Restricted to the set of
+five options obtained at the first step, \CodeIn{dsize} and
 \CodeIn{itype} are the most popular. At the end of this step, \tname{}
@@
 \label{sec:technique}
+
+\Mar{--------------------- I am here...}
 

### 2020-05-18 paper/paper.tex
pass

@@
 Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
-data set of \suri\ rules. A closer observation on the
-changelogs of this ruleset~\cite{emerging-threats-changelogs} shows that new rules are
-often created.  Rulesets need to be continuously revised as the
-community finds and advertises vulnerabilities, which eventually can be exploited by hackers. 
-Unfortunately, creating these rules is
-time-consuming and error-prone \rui{is there a citation we could add here?}. The language for describing rules is
-non-trivial (see Section~\ref{sec:example-suricata-rules}) and testing
-correctness of a rule requires data and careful inspection.
+data set of \suri\ rules. A closer observation on the changelogs of
+this ruleset~\cite{emerging-threats-changelogs} shows that new rules
+are often created.  Rulesets need to be continuously revised as the
+community finds and advertises vulnerabilities, which eventually can
+be exploited by hackers.  Unfortunately, creating these rules is
+time-consuming and error-prone~\cite{Ring_2019}. The language for
+describing rules is non-trivial (see
+Section~\ref{sec:example-suricata-rules}) and testing correctness of a
+rule requires data and careful inspection.
 
@@
 examples indicate what should not be captured. It is worth noting that
-vast amounts of negative data
-exist~\cite{tcpreplay,stratosphere-normal}, but positive data is
-scarce as it is not common (and considered unethical~\Mar{look for
-  citation for ethics code on this}) to publish attacks on the
-web\Mar{is there a term for non-deep web?}. Classical machine learning techniques \emph{are not directly
-  applicable in this context} for two main reasons: 1)~they require a
-large amount of training data to build models (recall that positive
-data is scarce) and 2)~they typically have limited support to string
-and structured data, which are prevalent in this domain.
+classical machine learning techniques \emph{are not directly
+  applicable in this context} as they typically have limited support
+to string and structured data, which are prevalent in this domain.
+
+%% vast amounts of negative data
+%% exist~\cite{tcpreplay,stratosphere-normal}, but positive data is
+%% scarce~\cite{} as it is not common (and considered unethical) to publish
+%% attacks on the web\Mar{is there a term for non-deep web?}.
 
@@
 the top ten rules in 11 of the 14 cases. \rui{to discuss, I find 
-this a bit weak.}
+this a bit weak.}\Mar{agree. let us try to add more examples and data.}
 
@@
 
+\Mar{mention something on ethical hacking.. ethical-hacking}
+
 This section elaborates on the most similar work.

### 2020-05-19 paper/paper.tex
pass on intro
new macros

@@
 \usepackage{booktabs}
-%% \usepackage{cite}
+\usepackage{cite}
 \usepackage{url}
@@
 be exploited by hackers.  Unfortunately, creating these rules is
-time-consuming and error-prone~\cite{Ring_2019}. The language for
+time-consuming and error-prone~\cite{vollmer-etal-cics2011,alparslan-blog-suri}. The language for
 describing rules is non-trivial (see
@@
 \percTopFiveRanking\ of the cases and ranked the correct rule among 
-the top ten rules in 11 of the 14 cases. \rui{to discuss, I find 
+the top ten rules in \topTenAttacks of the \totalAttacks cases. \rui{to discuss, I find 
 this a bit weak.}\Mar{agree. let us try to add more examples and data.}

### 2020-05-19 paper/paper.tex
pass on IV (preamble)

@@
 
-\Mar{--------------------- I am here...}
-
 \tname{} is a technique to synthesize rules for rule-based
@@
 examples of an attack and produces on output a ranked list of
-\emph{plausible rules}---\ie{}, rules that isolate the malicious
-traffic produced by the attacks from the benign
-traffic. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
+\emph{plausible rules}, \ie{}, rules that captures only the malicious
+traffic produced by the attacks. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
 pipeline of three components, appearing numbered in the figure.
@@
 
+% (see Section~\ref{sec:rules-and-packets})
+
+%% However, note that that
+%% approach alone could result in false positives as it imposes no limit
+%% on the extent that the rules can be weakened.  More precisely, it
+%% searches for plausible rules derived from the seed rule obtained in
+%% the previous step.
+
 \tname\ works as follows. First, it creates a rule that captures the
-malicious traffic produced by the examples provided on input. In this
-step, \tname\ detects options directly from the input messages (see
-Section~\ref{sec:rules-and-packets}). More precisely, it creates a
-rule containing options that are guaranteed to be satisfied by the
-malicious traffic. Let us refer to that rule as the \emph{seed rule}
-and its corresponding set of options as \emph{seed options}. Note
-that, as only the malicious traffic is considered in this step, the
-seed rule is likely over-specified, \ie{}, it may be more restrictive
-than necessary. Consequently, the use of that rule could result in
-false negatives. Second, \tname\ generates alternative rules that 1)
-weaken the constraints expressed in the seed rule while 2) still
-capturing the malicious traffic. Given that every option declared in a
-rule needs to be satisfied for a rule to capture an attack, any subset
-of the set of seed options, by definition, captures the attack
-characterized by the malicious traffic. \tname{} leverages that
-observation to search for plausible rules containing a proper subset
-of options from the seed options. However, note that that approach
-alone could result in false positives as it imposes no limit on the
-extent the constraints induced from the rule can be weakened. \tname{}
-uses benign traffic for that purpose. More precisely, it searches for
-plausible rules derived from the seed rule obtained in the previous
-step. Finally, \tname{} uses heuristic functions, obtained from a data
-set of existing rules, to rank the generated plausible rules by their
-similarity to existing rules. Similarity is used to approximate user's
-intent. The following sections detail each component of \tname.
+malicious traffic associated with example attacks provided on
+input. In this step, \tname\ detects options directly from the input
+messages. More precisely, it creates a rule containing options that
+are guaranteed to be satisfied by the malicious traffic. We refer to
+that rule as the \emph{seed rule} and its corresponding set of options
+as \emph{seed options}. It is important to note that, because only the
+malicious traffic is considered in this step, the seed rule is
+potentially over-specified, \ie{}, it may be more restrictive than
+necessary. Consequently, the use of that rule could result in false
+negatives. Second, \tname\ uses benign traffic to generate alternative
+rules that 1) weaken the constraints expressed in the seed rule while
+2) still capturing the malicious traffic. Given that every option
+declared in a rule needs to be satisfied for a rule to capture an
+attack, any subset of the set of seed options, by definition, captures
+the attack characterized by the malicious traffic. \tname{} leverages
+that observation to search for plausible rules containing a proper
+subset of options from the seed options that captures \emph{only} the
+malicious traffic. Finally, \tname{} uses heuristic functions,
+obtained from a data set of existing rules, to rank the generated
+plausible rules by their similarity to existing rules. Similarity is
+used to approximate user's intent. The following sections detail each
+component of \tname.
+
+\Mar{--------------------- I am here...}
 

### 2020-05-19 paper/paper.tex
added source on table ii

@@
   numbers as this *in text* and use macros instead. We do not want to
-  leave inconsistent data in the paper!}14 attacks of various kinds, such as \Fix{...}\rui{add few examples}. The observed results are encouraging. For 
+  leave inconsistent data in the paper!}\totalAttacks attacks of various kinds, such as \Fix{...}\rui{add few examples}. The observed results are encouraging. For 
 example, \tname\ ranked the correct rule among the top five rules in 
@@
     UDP Teardrop & \cite{udp-teardrop-source} & DoS & udp & \textbf{fragbits}:M; \textbf{id}:124; \\
-      2100384 & - & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:8; \\
-      2100399 & - & ICMP Info & icmp & \textbf{icode}:1; \textbf{itype}:3; \\
-      2100401 & - & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:3; \\
-      2100402 & - & ICMP Info & icmp & \textbf{icode}:3; \textbf{itype}:3; \\
-      2100408 & - & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:0; \\
-      2100485 & - & ICMP Info & icmp & \textbf{icode}:13; \textbf{itype}:3; \\
-      2100486 & - & ICMP Info & icmp & \textbf{icode}:10; \textbf{itype}:3; \\
+      2100384 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:8; \\
+      2100399 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:1; \textbf{itype}:3; \\
+      2100401 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:3; \\
+      2100402 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:3; \textbf{itype}:3; \\
+      2100408 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:0; \\
+      2100485 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:13; \textbf{itype}:3; \\
+      2100486 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:10; \textbf{itype}:3; \\
     %%UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5; \\

### 2020-05-19 paper/paper.tex
pass

@@
 
-We evaluated \tname\ against \Mar{@Lucas, please look for constant
-  numbers as this *in text* and use macros instead. We do not want to
-  leave inconsistent data in the paper!}\totalAttacks attacks of various kinds, such as \Fix{...}\rui{add few examples}. The observed results are encouraging. For 
+%\Mar{@Lucas, please look for constant
+  %numbers as this *in text* and use macros instead. We do not want to
+  %leave inconsistent data in the paper!}
+We evaluated \tname\ against \totalAttacks attacks of various kinds, such as 
+\Luc{This is already on section V, should it also be here?} Denial-of-Service (DoS)~\cite{denial-of-service},
+Active Reconnaissance (AREC)~\cite{active-reconnaissance} and Privilege
+escalation (PESC)~\cite{privilege-escalation}. %\Fix{...}\rui{add few examples}. 
+The observed results are encouraging. For 
 example, \tname\ ranked the correct rule among the top five rules in 

### 2020-05-19 paper/paper.tex
up to IV.B

@@
 
-\Mar{--------------------- I am here...}
-
 %%  As the generation
@@
 
-The goal of the Reverse Engineering step is to extract options from
-the malicious traffic produced by the input example attacks. A rule
-that contains these options captures the malicious traffic by
-construction. \tname\ uses two strategies to extract options: 1)~the
-field strategy applies to any protocol and 2)~the payload strategy
-applies only to http.  The decision of which strategy to use is based
-on the protocol associated with input traffic--the technique uses the
-payload strategy if http is detected; otherwise, it uses the field
-strategy. We elaborate these strategies in the following.
+The goal of the Reverse Engineering step is to infer a \emph{seed
+  rule} that captures the attack. To that end, \tname{} monitors the
+traffic produced by the input example attack(s) and extracts options
+from related messages. A rule that contains these options captures the
+malicious traffic by construction. \tname\ uses two strategies to
+extract options: 1)~field and 2)~payload. The field strategy applies
+to any protocol whereas the payload strategy applies only to http.
+The decision of which strategy to use is based on the protocol
+associated with input traffic--\tname{} uses the payload strategy if
+http is detected; otherwise, it uses the field strategy. We elaborate
+these strategies in the following.
 
@@
 between packet fields and some of the rule options from \suri\ and
-Section~\ref{sec:active-recon} shows a concrete example illustrating a
-case of this strategy. We obtained the option-field mapping
-manually, from the \suri\ documentation.
-
-Unfortunately, the relevant data for the option may be at the message
-payload; not the fields. By analyzing our data set of rules, we found
-that this occurs frequently in http attacks, which are prevalent. For
-example, \percHttp\ of the \numrulessuri{} rules from the ruleset we
-analyzed~\cite{emerging-threats-open} prescribe defenses against http
-attacks. In addition, \percRulesWithContent{} of these rules contain
-at least one content option. Figure~\ref{fig:rule-jboss} shows an
-example of a rule that captures traffic containing the string
-\CodeIn{"/HtmlAdaptor"} in the payload.
-Figure~\ref{fig:distribution-contents} shows the distribution of
-number of content options per rule for the data set mentioned
-above. We limited the number of content options per rule to 10 in the
-histogram for space. \MyComment{Rules without \CodeIn{content} options
-  are an exception and half of the rules contain at least two
-  \CodeIn{content} options.}  To sum up, handling content options for
-http is very important.
+Section~\ref{sec:active-recon} shows a concrete example where \tname{}
+uses this strategy. ALthough we built the option-field mapping
+manually, from the \suri\ documentation, this is a one-time activity.
+Unfortunately, for http attacks, the relevant data for the options is
+likely at the message payload; not the fields. By analyzing our data
+set of rules, we found that this occurs frequently in http attacks,
+which are prevalent. For example, \percHttp\ of the \numrulessuri{}
+rules from the ruleset we analyzed~\cite{emerging-threats-open}
+prescribe defenses against http attacks. In addition,
+\percRulesWithContent{} of these rules contain at least one content
+option. Figure~\ref{fig:rule-jboss} shows an example of a rule that
+captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the
+payload.  Figure~\ref{fig:distribution-contents} shows the
+distribution of number of content options per rule for the data set
+mentioned above. We limited the number of content options per rule to
+10 in the histogram for space. \MyComment{Rules without
+  \CodeIn{content} options are an exception and half of the rules
+  contain at least two \CodeIn{content} options.}  To sum up, handling
+content options for http is very important.
 
@@
 message payload. By inspecting existing rules, we found that these
-options refer to sequences of characters separated by typical natural
-language delimiters such as $\backslash$t, $\backslash$n,
-$\backslash$r, and spaces. To find the values associated with these
-options, \tname{} splits the payload in tokens using these delimiters.
+options refer to sequences of characters separated by natural language
+delimiters such as $\backslash$t, $\backslash$n, $\backslash$r, and
+spaces. To find the values associated with these options, \tname{}
+splits the payload in tokens using these delimiters.
 
@@
 %% }
+
+\Mar{--------------------------------------------- I am here}
 

### 2020-05-19 paper/paper.tex
pass
pass on Step 2 (RUles Creation)

@@
 
-\Mar{--------------------------------------------- I am here}
-
 \subsection{Step 2: Rules Creation}
@@
 by the traffic produced by the attack. Consequently, any subset of
-these options also captures the attack. The rule creation procedure
-iteratively discards options from visited rules---derived from the
-seed rule---using the negative traffic to guide the search for minimal
-rules.
+these options also captures the attack.
 
@@
 
-Algorithm~\ref{algo1} shows the pseudocode for deriving rules from the
-seed rule produced at the Reverse Engineering step. The algorithm
-takes as input an over-constrained rule $sr$ and pcap files~\cite{pcap}
-to reconstruct the benign ($\mathit{bTr}$) and malicious traffic
+The rule creation procedure \tname{} uses iteratively discards options
+from rules, derived from the seed rule, using the negative traffic to
+guide the search for minimal rules.  Algorithm~\ref{algo1} shows the
+pseudocode for deriving these rules. The algorithm takes as input an
+over-constrained rule $sr$, denoting the seed rule produced at the
+Reverse Engineering step, and pcap files~\cite{pcap} to reconstruct
+the benign traffic ($\mathit{bTr}$) and the malicious traffic
 ($\mathit{mTr}$). The algorithm produces on output a set of
@@
 candidate rule in $\mathcal R$. The algorithm mutates each of these
-rules for $N$ times, discarding one option of the rule each
-time. Line~\ref{copy} assigns a copy of the bitvector, encoding
-$\mathit{ro}$, to variable $r$. Function $\mathit{len}$ returns the
-length of a vector and function $\mathit{rand}$ randomly chooses one
-integer value in the range defined by its parameters. So, the
-expression $\mathit{rand(0, len(r))}$, appearing at
-line~\ref{choose-index}, returns an index of the option in $r$ to
-delete at line~\ref{delete-option}. The call to $\mathit{isNew}$
-checks if $r$ has not been previously generated in another iteration
-whereas the call to $\mathit{isPlausible}$ checks if the rule only
-captures positive traffic, \ie{}, it checks that no message present in
-$\mathit{bTr}$ is captured by $r$. If $r$ satisfies both criteria, it
-is added to $\mathcal R'$, which denotes the next generation of
-individuals in the search. After processing every candidate rule in
-$\mathcal R$, the next generation of individuals is defined
-(line~\ref{define-next-gen}) and the set of discovered rules is
-updated (at line~\ref{output-accum}).
+rules for $N$ times, discarding one rule option each
+time. Line~\ref{copy} assigns a copy of the bitvector encoding
+$\mathit{ro}$ to variable $r$. Function $\mathit{len}$ returns the
+length of a vector and function $\mathit{rand(lo, hi)}$ randomly
+chooses one integer value in the range $[lo, hi]$ defined by its
+parameters. So, the expression $\mathit{rand(0, len(r))}$, appearing
+at line~\ref{choose-index}, returns an index of the option in $r$ and
+the statement at line~\ref{delete-option} deletes that option. The
+call to $\mathit{isNew}$ checks if $r$ has not been previously
+generated in another iteration whereas the call to
+$\mathit{isPlausible}$ checks if the rule only captures positive
+traffic, \ie{}, it checks that no message present in $\mathit{bTr}$ is
+captured by $r$. If $r$ satisfies both criteria, it is added to
+$\mathcal R'$, which denotes the next generation of individuals in the
+search. After processing every candidate rule in $\mathcal R$, the
+next generation of individuals is defined (line~\ref{define-next-gen})
+and the set of discovered rules is updated (at
+line~\ref{output-accum}).
 
@@
 with a single option are produced in the second-to-last iteration. In
-other iterations, rules with more than one option are
-produced. Conceptually, the search induces a search tree with the seed
-rule $sr$ as root, single-option rules as child nodes, and rules with
-more options as internal nodes. The edges of the tree denote the
-ancestor relation between a pair of rules. It is also worth noting
-that the algorithm reports every rule in the tree instead of only
-those with one option. The reason is that the traffic used to check
-plausibility is limited by definition---reporting only minimal rules
-with respect to that traffic could discard the correct rule.
+other iterations, rules with more than one option are produced. The
+search induces a search tree with the seed rule $sr$ as root,
+single-option rules as child nodes, and rules with more options as
+internal nodes. The edges of the tree denote the ancestor relation
+between a pair of rules. It is also worth noting that the algorithm
+reports every rule in the tree instead of only those with one
+option. The reason is that the traffic used to check plausibility is
+limited by definition---reporting only minimal rules with respect to
+that traffic could discard the correct rule.
 
@@
 at the \reveng\ step. In the Raking step (Section~\ref{sec:ranking}),
-\tname{} ranks the rules this algorithm reports by their similarity to
-trusted rules. Intuitively, rules dissimilar to those trusted rules
-rank low.
+\tname{} ranks the rules.
+
+%Intuitively, rules dissimilar to (a very
+%large set of) trusted rules rank low.
+
+% this algorithm reports by their similarity to trusted rules
 
@@
 \emph{isPlausible} by 1)~spawning Suricata to monitor a specified
-rule, 2)~replaying the negative benign traffic (see
+rule, 2)~replaying the benign traffic (see
 Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
@@
 \label{sec:ranking}
+
+\Mar{-------------------------- I am here}
 
@@
 escalation (PESC)~\cite{privilege-escalation}, Remote Code Execution
-(RCE)~\cite{remote-code-execution}, and Local File Inclusion
-(LFI)~\cite{local-file-inclusion}.
+(RCE)~\cite{remote-code-execution}, Local File Inclusion
+(LFI)~\cite{local-file-inclusion} and ICMP Traffic Alerts \cite{icmp-types}.
 \noindent\textbf{Malicious traffic.}~Recall that \tname\ takes

### 2020-05-19 paper/paper.tex
opening space for Lucas/Guilherme

@@
 
-\subsection{Threats to Validity}
-\label{sec:threats}
-
-Our experiments is affected by the usual threats. Considering external
-validity. Our results cannot be generalized beyond the subjects we
-selected. To mitigate that threat we selected a variety of attacks
-with different characteristics. Considering internal validity, the
-implementation of \tname\ could contain errors. We mitigated that
-threat by carefully inspecting code and results. Considering construct
-validity, the choice of metric used to evaluate the approach may not
-be applicable in practice. We considering the use of ranking of the
-golden rule a reasonable choice as the approach places the golden rule
-close to the top in most cases.
-
-\section{Discussion of other attacks}
+\subsection{Discussion}
 \label{sec:rce}
@@
 
-%\Mar{@Lucas, por
-%  favor, observe o padrao usado nos exemplos anteriores. eh preciso
-%  explicar em alto nivel o ataque antes de explicar a regra!}
-
-%% In an RCE attack, hackers
-%% intentionally exploit a vulnerability in the system to run malware
-%% By prompting the targeted device to perform code execution, a hacker
-%% can run arbritrary code to
-%% their own programming in its
-%% place. This programming can then enable them to gain full access,
-%% steal data, carry out a full distributed denial of service (DDoS)
-%% attack, destroy files and infrastructure, or engage in illegal
-%% activity. And as the term remote execution suggests, an RCE
-%% cyberattack can take place from any geophysical location."\\
-
-%A opção  define que o match só será feito em pacotes que fazem parte
-%de uma conexão já estabelecida entre um cliente e um servidor. Além
-%disso, define também que os pacotes devem estar direcionados ao servidor, não ao cliente.
-
-\begin{figure}[t!]
-%  \vspace{-1ex}    
-  \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
-%  \vspace{-2ex}  
-  \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code
-    Execution.}
-%  \vspace{-1ex}
-\end{figure}
-
-\Mar{We may want to comment part of this to make room for the 
-  experiments that Lucas and Guilherme are currently running.
-  Side-comparison with prior work based on paper (not actual tool) results.}
-This section elaborates how \tname\ performs on two other attacks.
-\textbf{Remote Code Execution (RCE).}~RCE refers to the family of
-attacks where hackers exploit a network vulnerability to run arbitrary
-code on a remote machine. The code to be executed could damage the
-system or steal sensitive information. JBoss~\cite{jboss} is a popular
-open source Java application server. This attack enables remote
-execution of code on the JBoss application server.
-Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
-JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
-indicates that a rule match will only occur on packets that (1) flow
-on an already-established connection and that (2) flow to the server
-as opposed to the client. For the content options the modifiers
-\CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
-the matching is not case sensitive and that the string should be
-located at the packet field ``Request URI''. We reproduced the
-malicious JBoss attack using the traffic from a Nikto
-script~\cite{nikto}. Similar attacks are also available on the
-Metasploit database~\cite{metasploit}. This is a case where
-\tname\ performed poorly. Of the 228 plausible rules it generated for
-this attack, the ground truth was located at position 45 in the ranked
-list. \textbf{Denial-of-Service.}~
-The SYN Flood attack is a denial-of-service attack that exploits a
-vulnerability in the tcp/ip handshake
-to establish a tcp connection~\cite{cloudfare-synflood}. The handshake
-works as follows in normal circumstances. First, a client sends a SYN
-packet to the server, requesting a connection. Second, the server
-responds with a SYN-ACK packet to the client. Third, the client
-responds with an ACK message and the connection is established. Aware
-of the protocol, an attacker sends multiple SYN packets to different
-ports of a server, often using fake IP addresses. Then, after the
-server responds with a SYN-ACK packet, the client keeps sending other
-SYN packets to avoid the connection to time out. Without proper
-protection, the server accepts these malicious requests and eventually
-legitimate requests cannot be satisfied due to resource exhaustion.
-
-%\vspace{-1ex}
-\begin{figure}[h!]
-  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
-%  \vspace{-2ex}
-  \caption{\suri\ rule for SYN Flood attack.}
-  \label{fig:synflood-example}
-\end{figure}
-%\vspace{-2ex}
-
-Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
-this attack. This rule was obtained from the Emerging Threats Open
-Ruleset~\cite{emerging-threats-open}. The relevant options in the
-rule appear in bold---the option \CodeIn{flags: S,12}, identifies a
-SYN packet in a tcp packet and the option \CodeIn{threshold: type
-  both, track by\_dst, count 5000, seconds 5} indicates that a high
-volume of such packets should be requested in a short period of
-time. The other options are for documentation purposes. This is a case
-where \tname\ works very well, reporting the golden rule at the top
-of the ranking.
+\subsubsection{Comparison with other tools}\Mar{Lucas and Guilherme,
+  please summarize results here!}
+
+\subsubsection{Running time}\Mar{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Add this in a journal submission in section to discuss
+%% limitations...
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% \begin{figure}[t!]
+%%   \lstinputlisting[language=C,numbers=none,keywords={content,flow}]{adaptor-golden-rule.suricata}
+%%   \caption{\label{fig:rule-jboss}Suricata rule for JBoss Remote Code
+%%     Execution.}
+%% \end{figure}
+
+%% \Mar{We may want to comment part of this to make room for the 
+%%   experiments that Lucas and Guilherme are currently running.
+%%   Side-comparison with prior work based on paper (not actual tool) results.}
+%% This section elaborates how \tname\ performs on two other attacks.
+%% \textbf{Remote Code Execution (RCE).}~RCE refers to the family of
+%% attacks where hackers exploit a network vulnerability to run arbitrary
+%% code on a remote machine. The code to be executed could damage the
+%% system or steal sensitive information. JBoss~\cite{jboss} is a popular
+%% open source Java application server. This attack enables remote
+%% execution of code on the JBoss application server.
+%% Figure~\ref{fig:rule-jboss} shows a Suricata rule that captures a
+%% JBoss RCE attack.  The option \CodeIn{flow: established, to\_server;}
+%% indicates that a rule match will only occur on packets that (1) flow
+%% on an already-established connection and that (2) flow to the server
+%% as opposed to the client. For the content options the modifiers
+%% \CodeIn{nocase} and \CodeIn{http\_uri} indicate, respectively, that
+%% the matching is not case sensitive and that the string should be
+%% located at the packet field ``Request URI''. We reproduced the
+%% malicious JBoss attack using the traffic from a Nikto
+%% script~\cite{nikto}. Similar attacks are also available on the
+%% Metasploit database~\cite{metasploit}. This is a case where
+%% \tname\ performed poorly. Of the 228 plausible rules it generated for
+%% this attack, the ground truth was located at position 45 in the ranked
+%% list.
+
+%% \textbf{Denial-of-Service.}~
+%% The SYN Flood attack is a denial-of-service attack that exploits a
+%% vulnerability in the tcp/ip handshake
+%% to establish a tcp connection~\cite{cloudfare-synflood}. The handshake
+%% works as follows in normal circumstances. First, a client sends a SYN
+%% packet to the server, requesting a connection. Second, the server
+%% responds with a SYN-ACK packet to the client. Third, the client
+%% responds with an ACK message and the connection is established. Aware
+%% of the protocol, an attacker sends multiple SYN packets to different
+%% ports of a server, often using fake IP addresses. Then, after the
+%% server responds with a SYN-ACK packet, the client keeps sending other
+%% SYN packets to avoid the connection to time out. Without proper
+%% protection, the server accepts these malicious requests and eventually
+%% legitimate requests cannot be satisfied due to resource exhaustion.
+
+%% %\vspace{-1ex}
+%% \begin{figure}[h!]
+%%   \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
+%% %  \vspace{-2ex}
+%%   \caption{\suri\ rule for SYN Flood attack.}
+%%   \label{fig:synflood-example}
+%% \end{figure}
+%% %\vspace{-2ex}
+
+%% Figure~\ref{fig:synflood-example} shows a Suricata rule that captures
+%% this attack. This rule was obtained from the Emerging Threats Open
+%% Ruleset~\cite{emerging-threats-open}. The relevant options in the
+%% rule appear in bold---the option \CodeIn{flags: S,12}, identifies a
+%% SYN packet in a tcp packet and the option \CodeIn{threshold: type
+%%   both, track by\_dst, count 5000, seconds 5} indicates that a high
+%% volume of such packets should be requested in a short period of
+%% time. The other options are for documentation purposes. This is a case
+%% where \tname\ works very well, reporting the golden rule at the top
+%% of the ranking.
+
+\subsection{Threats to Validity}
+\label{sec:threats}
+
+Our experiments is affected by the usual threats. Considering external
+validity. Our results cannot be generalized beyond the subjects we
+selected. To mitigate that threat we selected a variety of attacks
+with different characteristics. Unfortunately, we could not access
+tools for related techniques. For that reason, we made a
+side-comparison based on the common examples and results reported in
+the original paper. Considering internal validity, the implementation
+of \tname\ could contain errors. We mitigated that threat by carefully
+inspecting code and results. Considering construct validity, the
+choice of ranking to evaluate the approach may not be useful in
+practice. We considered the use of ranking of the golden rule a
+reasonable choice as the approach places the golden rule close to the
+top in most cases.
 

### 2020-05-20 paper/paper.tex
comparison with Vollmer

@@
 
-\subsubsection{Comparison with other tools}\Mar{Lucas and Guilherme,
-  please summarize results here!}
+\subsubsection{Comparison with other tools}
+ \Luc{Vollmer \cite{vollmer-etal-cics2011} used and evolutionary approach to 
+autonomously generate rules to detect ICMP packets using the Snort IDS. They built their own dataset 
+using packet creation tools. Ten different packets were 
+created to match ten existing rules, and then these packets were used as input 
+to the rule generation. As a result, they were able to produce rules with 
+100\% detection rate and that produced 3 false positives out of 33k packets.
+On our experiments with Syrius, we used 381 \Fix{use macros}packets that matches nine existing 
+rules and it was able to generate rules that detected all the packets with 
+no false positives.  
+ }
+
+ \Luc{Guruprasad \cite{guruprasad}}
 

### 2020-05-20 paper/paper.tex
fix
pass V.E

--- 2020-05-19 paper/paper.tex
@@
 
-\subsubsection{Comparison with other tools}\Mar{Lucas and Guilherme,
-  please summarize results here!}
-
+\subsubsection{Comparison with other tools}
+ \Luc{Vollmer \cite{vollmer-etal-cics2011} and Guruprasad \cite{guruprasad} used 
+ evolutionaries approaches to 
+autonomously generate rules using the Snort IDS. They built 
+their own ICMP datasets by using packet creation tools or by capturing network traffic. 
+Guruprasad also used DARPA 1999 and ISCX 2012 \cite{iscx} datasets for non-ICMP attacks.
+
+Vollmer created ten different packets to match ten existing rules, and then used 
+these packets as input to the rule generation process. As a result, the produced 
+rules achieved a 100\% detection rate and produced 3 false positives out of 33k packets.
+
+Guruprasad manually created a set of rules by analyzing the network patterns and 
+the datasets, this set of rules was used as the initial population of the algorithm. 
+The produced rules achieved a detection rate up to 96\%, the false positive rate 
+was not evaluated.
+
+On our experiments with Syrius, we used a total of \totalPackets packets that 
+matches \totalAttacks different existing rules and it was able to generate rules that detected 
+all the packets with no false positives.  
+ }  
 \subsubsection{Running time}\Mar{}

### 2020-05-20 paper/paper.tex
fix missing reference
reminder

@@
 The observed results are encouraging. For 
-example, \tname\ ranked the correct rule among the top five rules in 
+example, \tname\ ranked the correct rule among the top five rules in \Fix{fix values}
 \percTopFiveRanking\ of the cases and ranked the correct rule among 
@@
 their own ICMP datasets by using packet creation tools or by capturing network traffic. 
-Guruprasad also used DARPA 1999 and ISCX 2012 \cite{iscx} datasets for non-ICMP attacks.
+Guruprasad also used DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets 
+for non-ICMP attacks.
 

### 2020-05-21 paper/paper.tex
comparison

@@
 \subsubsection{Comparison with other tools}
+ \Gui{There are a few reasons that makes the comparison between Syrius and other tools \Fix{impossible}
+
+\textbf{Tools availability.} We tried contact with authors of 5 different papers \Fix{[10][34][37][iran][bayesian]}, but got no response from them.
+Since their tools are not available online or at least not publicly, we could not run them.
+
+\textbf{Different output and specific rules.} We considered using the results they provides with results that Syrius provide, but concluded it wouldn't be possible as the results are essentially different.
+Commonly the  other results focus on showing the ratio of false negatives and false positives as a way to show the efficiency of the learning algorithm \Fix{citation needed}, but never comparing the generated rules to a Ground Truth, and most of times not showing individually the attacks used.
+Syrius' results, on the other hand, shows the generated rules that produce no false positives and negatives to the given packets, and always compares the results with a Ground Truth to verify its efficiency.
+
+\textbf{'content' option support.} Another limitation is the support of 'content' option.
+Most papers do not support rule generation with the 'content' option, using only classic DDoS attacks in the rule generation, like ICMP and SYN flood, which only use numerical and categorical options.
+Some papers do support 'content', but don't explain well how they handle those rules with exclusive 'content'.\Fix{citation needed}}
  \Luc{Vollmer \cite{vollmer-etal-cics2011} and Guruprasad \cite{guruprasad} used 

### 2020-05-21 paper/paper.tex
related works

@@
 
-\subsubsection{Comparison with other tools}
+\subsubsection{\Fix{Lack of }Comparison with other tools}
  \Gui{There are a few reasons that makes the comparison between Syrius and other tools \Fix{impossible}
@@
 Some papers do support 'content', but don't explain well how they handle those rules with exclusive 'content'.\Fix{citation needed}}
- \Luc{Vollmer \cite{vollmer-etal-cics2011} and Guruprasad \cite{guruprasad} used 
- evolutionaries approaches to 
-autonomously generate rules using the Snort IDS. They built 
-their own ICMP datasets by using packet creation tools or by capturing network traffic. 
-Guruprasad also used DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets 
-for non-ICMP attacks.
-
-Vollmer created ten different packets to match ten existing rules, and then used 
-these packets as input to the rule generation process. As a result, the produced 
-rules achieved a 100\% detection rate and produced 3 false positives out of 33k packets.
-
-Guruprasad manually created a set of rules by analyzing the network patterns and 
-the datasets, this set of rules was used as the initial population of the algorithm. 
-The produced rules achieved a detection rate up to 96\%, the false positive rate 
-was not evaluated.
-
-On our experiments with Syrius, we used a total of \totalPackets packets that 
-matches \totalAttacks different existing rules and it was able to generate rules that detected 
-all the packets with no false positives.  
- }  
+%%  \Luc{Vollmer \cite{vollmer-etal-cics2011} and Guruprasad \cite{guruprasad} used 
+%%  evolutionaries approaches to 
+%% autonomously generate rules using the Snort IDS. They built 
+%% their own ICMP datasets by using packet creation tools or by capturing network traffic. 
+%% Guruprasad also used DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets 
+%% for non-ICMP attacks.
+%% 
+%% Vollmer created ten different packets to match ten existing rules, and then used 
+%% these packets as input to the rule generation process. As a result, the produced 
+%% rules achieved a 100\% detection rate and produced 3 false positives out of 33k packets.
+%% 
+%% Guruprasad manually created a set of rules by analyzing the network patterns and 
+%% the datasets, this set of rules was used as the initial population of the algorithm. 
+%% The produced rules achieved a detection rate up to 96\%, the false positive rate 
+%% was not evaluated. 
+%% 
+%% On our experiments with Syrius, we used a total of \totalPackets packets that 
+%% matches \totalAttacks different existing rules and it was able to generate rules that detected 
+%% all the packets with no false positives.  
+%%  }  
 \subsubsection{Running time}\Mar{}
@@
 derive rules for Snort, which are very similar to \suri's.  Their
-approach focuses on icmp attacks. It models every field of the icmp
+approach focuses on icmp attacks \Gui{, creating their own packets manually}. It models every field of the icmp
 packet in the chromosome of candidate solutions (\ie{}, a rule) of a
@@
 contain \emph{no} false positives or negatives with respect to the
-benign and mallicious traffic considered. Muthuregunathan
+benign and mallicious traffic considered.\Gui{Guruprasad \cite{guruprasad} also proposes an evolutionary framework, creating their own ICMP dataset as Vollmer, but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets for non-ICMP attacks.
+The produced rules reported a detection rate up to 96\%.
+Although Guruprasad supports more protocols than Vollmer, our approach still supports more options and protocols, making both key differences considered when comparing with Vollmer also relevant with Guruprasad.} Muthuregunathan
 \etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to

### 2020-05-22 paper/paper.tex
adding some references to V.E

@@
 
-\subsubsection{\Fix{Lack of }Comparison with other tools}
- \Gui{There are a few reasons that makes the comparison between Syrius and other tools \Fix{impossible}
-
-\textbf{Tools availability.} We tried contact with authors of 5 different papers \Fix{[10][34][37][iran][bayesian]}, but got no response from them.
+\subsubsection{Lack of comparison with other tools}
+ \Gui{There are a few reasons that makes the comparison between Syrius and other tools impossible
+
+\textbf{Tools availability.} We tried contact with authors of \Fix{3} different papers \cite{vollmer-etal-cics2011}\cite{muthuregunathan-etal2009}\cite{fallahi}, but got no response from them.
 Since their tools are not available online or at least not publicly, we could not run them.

### 2020-05-22 paper/paper.tex
pass V.E

@@
 
-\textbf{Tools availability.} We tried contact with authors of \Fix{3} different papers \cite{vollmer-etal-cics2011}\cite{muthuregunathan-etal2009}\cite{fallahi}, but got no response from them.
+\textbf{Tools availability.} We tried contact with authors of 4 different papers \cite{vollmer-etal-cics2011}\cite{muthuregunathan-etal2009}\cite{fallahi}\cite{Kao2015AutomaticNR}, but got no response from them.
 Since their tools are not available online or at least not publicly, we could not run them.
@@
 \textbf{Different output and specific rules.} We considered using the results they provides with results that Syrius provide, but concluded it wouldn't be possible as the results are essentially different.
-Commonly the  other results focus on showing the ratio of false negatives and false positives as a way to show the efficiency of the learning algorithm \Fix{citation needed}, but never comparing the generated rules to a Ground Truth, and most of times not showing individually the attacks used.
+Commonly the  other results focus on showing the ratio of false negatives and false positives as a way to show the efficiency of the learning algorithm, but never comparing the generated rules to a Ground Truth, and most of times not showing individually the attacks used.
 Syrius' results, on the other hand, shows the generated rules that produce no false positives and negatives to the given packets, and always compares the results with a Ground Truth to verify its efficiency.
@@
 Most papers do not support rule generation with the 'content' option, using only classic DDoS attacks in the rule generation, like ICMP and SYN flood, which only use numerical and categorical options.
-Some papers do support 'content', but don't explain well how they handle those rules with exclusive 'content'.\Fix{citation needed}}
+Some papers do support 'content', but don't explain well how they handle those rules with exclusive 'content'.}
 %%  \Luc{Vollmer \cite{vollmer-etal-cics2011} and Guruprasad \cite{guruprasad} used 

### 2020-05-22 paper/paper.tex
comment on introduction

@@
 
-\tname{} takes on input one or more examples \rui{We ought to explain how does one get hold on examples. How diffcult are they to obtain?} of an attack and produces
+\tname{} takes on input one or more examples \rui{We ought to explain how does one get hold on examples.
+How diffcult are they to obtain?} \Gui{Given a captured network traffic, an expert must identify which packets can be used to generate the rule.
+These packets are separated and given as input to Syrius.}of an attack and produces
 on output a ranked list of rules for that attack. It uses public data
@@
 \etal{}~\cite{vollmer-etal-cics2011} propose the use of search to
-derive rules for Snort, which are very similar to \suri's.  Their
-approach focuses on icmp attacks \Gui{, creating their own packets manually}. It models every field of the icmp
+derive rules for Snort, which are very similar to \suri's.  \Fix{Their
+approach focuses on icmp attacks.} \Gui{The dataset used by them consists of manually produced ICMP packets using packet creation tools.}. It models every field of the icmp
 packet in the chromosome of candidate solutions (\ie{}, a rule) of a

### 2020-05-22 paper/paper.tex
pass table iii and table iv

@@
     %%Apache Struts 2 & 2 & 65 & - & - \\
-    ColdFusion & 7 & 19 & 34 & 3\\
-    iisadmin access & 3 & 17 & 408 & 1 \\        
-    JBoss JMX & 5 & 23 & 228 & 45 \\
-    scripttag & 23 & 22 & 786 & 1 \\
-    WeBid LFI & 1 & 23 & 11,100 & 10\\    
+    ColdFusion & 7 & 19 & 41 & 4\\
+    iisadmin access & 3 & 17 & 606 & 1 \\        
+    JBoss JMX & 5 & 23 & 925 & 88 \\
+    scripttag & 23 & 22 & 2,222 & 1 \\
+    WeBid LFI & 1 & 23 & 35,445 & 18\\    
     %%Wordpress Login & 8 & 59 & - & - \\
-    .htaccess access & 5 & 17 & 984 & 2\\
-    .idq access & 14 & 19 & 511 & 2 \\
-    /system32/ in URI & 4 & 20 & 1,913 & 4 \\
-    Basebuilder LFI & 1 & 22 & 18,133 & 34 \\
-    Tomcat Snoop Access & 1 & 24 & 29,830 & 32 \\
+    .htaccess access & 5 & 17 & 853 & 1\\
+    .idq access & 14 & 19 & 807 & 1 \\
+    /system32/ in URI & 4 & 20 & 1882 & 2 \\
+    Basebuilder LFI & 1 & 22 & 60223 & 49 \\
+    Tomcat Snoop Access & 1 & 24 & 58,123 & 39 \\
     \bottomrule
@@
     SYN Flood & 1 & 1 & 1 & 1 & 1 \\
-    UDP Teardrop & 4 & 4 & 4 & 4 & 4 \\
+    2100384 & 2 & \Fix{} & 2 & 2 & 2 \\
+    2100399 & 1 & \Fix{} & 1 & 1 & 1 \\
+    2100401 & 1 & \Fix{} & 1 & 1 & 1 \\
+    2100402 & 1 & \Fix{} & 1 & 1 & 1 \\
+    2100408 & 2 & \Fix{} & 2 & 2 & 2 \\
+    2100485 & 1 & \Fix{} & 1 & 1 & 1 \\
+    2100486 & 1 & \Fix{} & 1 & 1 & 1 \\
     %%UDP Flood & - & - & - & - & - \\
@@
     %%Apache Struts 2 & - & - & - & - & - & -\\
-    ColdFusion & 3 & 3 & 2\cellcolor{lightgray}$\downarrow$ & 12$\uparrow$ & 5$\uparrow$ \\
+    ColdFusion & 4 & 4 & 2\cellcolor{lightgray}$\downarrow$ & 12$\uparrow$ & 5$\uparrow$ \\
     iisadmin access & 1  & 1 & 1 & 1 & 8$\uparrow$\\        
-    JBoss JMX & 45 & 45 & 5\cellcolor{lightgray}$\downarrow$ & 45 & 111$\uparrow$ \\
+    JBoss JMX & 88 & 88 & 5\cellcolor{lightgray}$\downarrow$ & 45 & 111$\uparrow$ \\
     scripttag & 1 & 1 & 4$\uparrow$ & 1 & 8$\uparrow$ \\
-    WeBid LFI & 10 & 10 & 77$\uparrow$ & 28$\uparrow$ & 32$\uparrow$ \\    
+    WeBid LFI & 18 & 18 & 77$\uparrow$ & 28$\uparrow$ & 32$\uparrow$ \\    
     %%Wordpress Login & 8 & - & - & - & - \\
-    .htaccess access & 2  & 2 & 2 & 2 & 17$\uparrow$\\
-    .idq access & 2  & 2 & 2 & 2 & 12$\uparrow$ \\
-    /system32/ in URI & 4  & 4 & 20$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 58$\uparrow$\\
-    Basebuilder LFI & 34  & 34 & 117$\uparrow$ & 57$\uparrow$ & 183$\uparrow$\\
-    Tomcat Snoop Access & 32  & 32 & 828$\uparrow$ & 2\cellcolor{lightgray}$\downarrow$ & 2564$\uparrow$\\
+    .htaccess access & 1  & 1 & 1 & 1 & 17$\uparrow$\\
+    .idq access & 1  & 1 & 1 & 1 & 12$\uparrow$ \\
+    /system32/ in URI & 2  & 2 & 20$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 58$\uparrow$\\
+    Basebuilder LFI & 49  & 49 & 117$\uparrow$ & 57$\uparrow$ & 183$\uparrow$\\
+    Tomcat Snoop Access & 39  & 39 & 828$\uparrow$ & 2\cellcolor{lightgray}$\downarrow$ & 2564$\uparrow$\\
     \bottomrule

### 2020-05-24 paper/paper.tex
pass table iii and table iv

@@
     %%Apache Struts 2 & 2 & 65 & - & - \\
-    ColdFusion & 7 & 19 & 41 & 4\\
-    iisadmin access & 3 & 17 & 606 & 1 \\        
-    JBoss JMX & 5 & 23 & 925 & 88 \\
-    scripttag & 23 & 22 & 2,222 & 1 \\
-    WeBid LFI & 1 & 23 & 35,445 & 18\\    
+    ColdFusion & 7 & 19 & 40 & 4\\
+    iisadmin access & 3 & 17 & 605 & 1 \\        
+    JBoss JMX & 5 & 23 & 925 & 87 \\
+    scripttag & 23 & 22 & 2,221 & 1 \\
+    WeBid LFI & 1 & 23 & 35,444 & 17\\    
     %%Wordpress Login & 8 & 59 & - & - \\
-    .htaccess access & 5 & 17 & 853 & 1\\
-    .idq access & 14 & 19 & 807 & 1 \\
-    /system32/ in URI & 4 & 20 & 1882 & 2 \\
-    Basebuilder LFI & 1 & 22 & 60223 & 49 \\
-    Tomcat Snoop Access & 1 & 24 & 58,123 & 39 \\
+    .htaccess access & 5 & 17 & 852 & 1\\
+    .idq access & 14 & 19 & 806 & 2 \\
+    /system32/ in URI & 4 & 20 & 1881 & 2 \\
+    Basebuilder LFI & 1 & 22 & 60222 & 48 \\
+    Tomcat Snoop Access & 1 & 24 & 58,122 & 37 \\
     \bottomrule
@@
     %%Apache Struts 2 & - & - & - & - & - & -\\
-    ColdFusion & 4 & 4 & 2\cellcolor{lightgray}$\downarrow$ & 12$\uparrow$ & 5$\uparrow$ \\
-    iisadmin access & 1  & 1 & 1 & 1 & 8$\uparrow$\\        
-    JBoss JMX & 88 & 88 & 5\cellcolor{lightgray}$\downarrow$ & 45 & 111$\uparrow$ \\
-    scripttag & 1 & 1 & 4$\uparrow$ & 1 & 8$\uparrow$ \\
-    WeBid LFI & 18 & 18 & 77$\uparrow$ & 28$\uparrow$ & 32$\uparrow$ \\    
+    ColdFusion & 4 & 4 & 2\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ & 5$\uparrow$ \\
+    iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
+    JBoss JMX & 87 & 87 & 15\cellcolor{lightgray}$\downarrow$ & 84 & 269$\uparrow$ \\
+    scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
+    WeBid LFI & 17 & 17 & 144$\uparrow$ & 27$\uparrow$ & 41$\uparrow$ \\    
     %%Wordpress Login & 8 & - & - & - & - \\
-    .htaccess access & 1  & 1 & 1 & 1 & 17$\uparrow$\\
-    .idq access & 1  & 1 & 1 & 1 & 12$\uparrow$ \\
-    /system32/ in URI & 2  & 2 & 20$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 58$\uparrow$\\
-    Basebuilder LFI & 49  & 49 & 117$\uparrow$ & 57$\uparrow$ & 183$\uparrow$\\
-    Tomcat Snoop Access & 39  & 39 & 828$\uparrow$ & 2\cellcolor{lightgray}$\downarrow$ & 2564$\uparrow$\\
+    .htaccess access & 1  & 1 & 1 & 1 & 11$\uparrow$\\
+    .idq access & 2  & 2 & 1 & 1 & 15$\uparrow$ \\
+    /system32/ in URI & 2  & 2 & 6$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$\\
+    Basebuilder LFI & 48  & 48 & 846$\uparrow$ & 65$\uparrow$ & 664$\uparrow$\\
+    Tomcat Snoop Access & 37  & 37 & 1797$\uparrow$ & 6\cellcolor{lightgray}$\downarrow$ & 3580$\uparrow$\\
     \bottomrule

### 2020-05-24 paper/paper.tex
pass table iv

@@
     SYN Flood & 1 & 1 & 1 & 1 & 1 \\
-    2100384 & 2 & \Fix{} & 2 & 2 & 2 \\
-    2100399 & 1 & \Fix{} & 1 & 1 & 1 \\
-    2100401 & 1 & \Fix{} & 1 & 1 & 1 \\
-    2100402 & 1 & \Fix{} & 1 & 1 & 1 \\
-    2100408 & 2 & \Fix{} & 2 & 2 & 2 \\
-    2100485 & 1 & \Fix{} & 1 & 1 & 1 \\
-    2100486 & 1 & \Fix{} & 1 & 1 & 1 \\
+    2100384 & 2 & 18$\uparrow$& 2 & 2 & 2 \\
+    2100399 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
+    2100401 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
+    2100402 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
+    2100408 & 2 & 9$\uparrow$ & 2 & 2 & 2 \\
+    2100485 & 1 & 2$\uparrow$ & 1 & 1 & 1 \\
+    2100486 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
     %%UDP Flood & - & - & - & - & - \\

### 2020-05-24 paper/paper.tex
pass on section iv

@@
 
-\Mar{-------------------------- I am here}
-
-\tname{} uses a composition of heuristic functions to sort the rules
+\tname{} uses a composition of heuristic functions to rank the rules
 produced in the previous step. Each of these functions uses a
-different criterion to estimate how similar a created rule is to those
-from public rulesets~\cite{emerging-threats-open}. The intuition is
-that rules that look more similar to existing trusted rules are more
-likely to be useful. The aggregated score of a rule is the weighted
-average across the values produced by each heuristic function. The
-co-domain of each function ranges on the 0-1 interval. Hence, the score
-of a rule also ranges on the 0-1 interval. We describe in the
-following each of these heuristic functions.
+different criterion to estimate how similar a new rule is to the rules
+from public rulesets (\eg{},~\cite{emerging-threats-open}). The
+intuition is that rules that look more similar to existing trusted
+rules are more likely to reflect user's intention and be more
+useful. The aggregated score of a rule is the weighted average across
+the values produced by each heuristic function. The co-domain of each
+function ranges on the 0-1 interval. Hence, the score of a rule also
+ranges on the 0-1 interval. We describe in the following each of these
+heuristic functions.
 
@@
 \subsubsection{\label{sec:popularity-options}Popularity of options.}
-One heuristic function that \tname{} uses to measure similarity
-between a synthesized rule and a human-created rule is the popularity
-of the options used in those rules. According to this heuristic, rules
-containing popular options score higher compared to those containing
-unpopular rules. Figure~\ref{fig:distribution-options} shows the
-histogram of rule options for the ruleset we
-used~\cite{emerging-threats-open}. The x-axis shows the options used
-in rules (the 10 most common) and the y-axis shows the number of rules
-using the corresponding option. Note that the options
-\CodeIn{content}, \CodeIn{flow}, and \CodeIn{prce} are very
-popular. \tname\ uses the following formula to compute the
+
+\tname{} use the popularity of the options used in trusted rules as a
+heuristic function---rules containing popular options score higher
+compared to those containing unpopular
+rules. Figure~\ref{fig:distribution-options} shows the histogram of
+rule options for the ruleset we used~\cite{emerging-threats-open}. The
+x-axis shows the options used in rules (the 10 most common) and the
+y-axis shows the number of rules using the corresponding option. Note
+that the options \CodeIn{content}, \CodeIn{flow}, and \CodeIn{prce}
+are very popular. \tname\ uses the following formula to compute the
 contribution of option popularity to the score of a rule $r$:
-
-{\small
-\vspace{-2ex}
-\[\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}/N\]
-\vspace{-2ex}
-}
+\indent
+$\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}/N$.
+
+%% {\small
+%% \vspace{-2ex}
+%% \[
+
+%% \]
+%% \vspace{-2ex}
+%% }
 
@@
 encoding $r$. The term $\mathit{option\_frequency}$ encodes the
-distribution of option popularity as a dictionary mapping
-numbers---indicating popularity---to corresponding
-options. Consequently, the term $\mathit{option\_frequency[option_i]}$
-denotes the number of rules, from our dataset, containing
-$\mathit{option\_i}$. The term $\mathit{max(option\_frequency)}$
-refers to the maximum number of rules for any dictionary entry (to
-normalize the measurement in the 0-1 range), and $N$ is the number of
-options in $r$. To illustrate, the contribution of this function to
-the score of a rule would be $0.26$ ($=(10181+3049)/(2*25699)$) for a
-rule with options \CodeIn{prce} and \CodeIn{flowbits} and $0.01$
-($=(155+108)/(2*25699)$) for a rule with options \CodeIn{itype} and
-\CodeIn{icode}.
+distribution of option popularity as a dictionary mapping numbers,
+indicating popularity, to corresponding options. The expression
+$\mathit{option\_frequency[option_i]}$ denotes the number of rules,
+from our dataset, containing $\mathit{option\_i}$. The expression
+$\mathit{max(option\_frequency)}$ refers to the maximum number of
+rules for any dictionary entry (to normalize the measurement in the
+0-1 range), and $N$ is the number of options in $r$. To illustrate,
+the contribution of this function to the score of a rule would be
+$0.26=$($10181$+$3049)/(2$*$25699)$ for a rule with options
+\CodeIn{prce} and \CodeIn{flowbits} and
+$0.01$=$(155$+$108)/(2$*$25699)$ for a rule with options
+\CodeIn{itype} and \CodeIn{icode}.
 
@@
 rule is not uniform. Figure~\ref{fig:distribution-contents} shows the
-distribution of number of contents in a rule. Results indicate nearly
-73\% of the rules from our ruleset contain one to three content
+distribution of number of contents in a rule. Results indicate that
+nearly 73\% of the rules from our ruleset contain one to three content
 options. Hence, we found that that information could help
@@
 where $\mathit{len(r)}$ denotes the number of content options in $r$
-and $\mathit{rules\_per\_number\_of\_contents}$ denotes the
-distribution of number of contents (encoded with a dictionary as in
-the function described on
-Section~\ref{sec:popularity-options}). Consequently the term
-$\mathit{rules\_per\_number\_of\_contents[len(r)]}$ corresponds to the
-number of rules in the ruleset with the same number of contents as
-$r$. The denominator of the fraction normalizes the contribution in
-the 0-1 range. To illustrate, the contribution of this heuristic
-function to the score a rule would be $.64$ $(=5841/9164)$ to a rule
-with 3 content options and $.18$ $(=1622/9164)$ to a rule with 5
-content options.
+and $\mathit{rules\_per\_number\_of\_contents}$ is the distribution of
+number of contents. The expression in the numerator
+%$\mathit{rules\_per\_number\_of\_contents[len(r)]}$
+calculates the number of rules in the ruleset with the same number of
+contents as $r$ whereas the denominator normalizes the value in the
+0-1 range. To illustrate, the contribution of this function would be
+$.64$~($=5841/9164$) and $.18$~($=1622/9164$) to the score of rules
+with 3 and 5 options, respectively.
 
@@
 
+%% {\small
+%% \vspace{-1.5ex}
+%% \[
+%% \]
+%% \vspace{-0.5ex}
+%% }
+
 \subsubsection{Popularity of content strings.} 
-This heuristic leverages content rarity to discriminate rules---it
+This heuristic leverages content rarity to discriminate rules. It
 weighs higher (respectively, lower) rules that use popular
 (respectively, unpopular) strings in content options. The score of a
-rule $r$ on this function is given by the arithmetic mean of the
-popularity scores of each string used in contents in $r$. More
-precisely, the contribution of content rarity is given by the
-following formula:
-
-{\small
-\vspace{-1.5ex}
-\[\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}/N\]
-\vspace{-0.5ex}
-}
+rule $r$ on this function is the arithmetic mean of the popularity
+scores of each string used in contents in $r$. More precisely, content
+rarity is obtained with the following formula:
+\indent
+$\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}/N$
+
 \noindent
-where $\mathit{content_i}$ denotes the $i$th content in a given
+, where $\mathit{content_i}$ denotes the $i$th content in a given
 rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
@@
 rule. Figure~\ref{fig:distribution-strings} shows the distribution of
-the ten most popular strings in our ruleset. To illustrate, let us
-consider a rule with whose content options use the strings ``GET'' and
-``ID'', respectively. The contribution of that rule for the
-computation of similarity score would be 0.601 (=(1+.203)/2). The
-relatively high weight reflects the fact that ``GET'' and ``ID'' are
-popular content strings.
+the ten most popular strings in our ruleset. To illustrate, consider a
+rule whose content options use the strings ``GET'' and ``ID''. The
+contribution of that rule for the computation of similarity score
+would be $0.601$ $(=(1$+$.203)$/$2)$. The relatively high weight
+reflects the fact that ``GET'' and ``ID'' are popular content strings.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[scale=0.5]{figs/http_header-example.png}
+\vspace{-2ex}
+\caption{Content modifiers.}
+\label{fig:http-header-example}
+\end{figure}
 
@@
   http\_header} indicates that the string ``keep-alive'' should be
-present in the header of the http message\footnote{A keep-alive
-  connection attribute indicates that a single tcp connection should
-  remain open for multiple http requests/responses.}. Note that, in
-this case, the content modifier refers to a field of an http message,
-which appears in the payload of a tcp message, which is the target of
-the option \CodeIn{content}. Figure~\ref{fig:http-header-example}
-illustrates possible content options used to match against certain
-strings that appear in an http message. A check mark (respectively,
-cross) indicates a match (respectively, no match).
-
-\begin{figure}[t!]
-\centering
-\includegraphics[scale=0.5]{figs/http_header-example.png}
-\vspace{-2ex}
-\caption{Content modifiers.}
-\label{fig:http-header-example}
-\end{figure}
-
+present in the header of the http message. (A keep-alive connection
+attribute indicates that a single tcp connection should remain open
+for multiple http requests/responses.) In this case, the content
+modifier refers to an http field that appears in the payload of a tcp
+message, which is the target of the option
+\CodeIn{content}. Figure~\ref{fig:http-header-example} illustrates
+possible content options used to match against certain strings that
+appear in an http message. A check mark (respectively, cross)
+indicates a match (respectively, no match).
 \tname\ builds on the observation that the strings commonly accessed
@@
 
-\[\frac{\sum_{i=1}^{N}g(i)}{N},~g(i)=\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]
+\indent
+$\sum_{i=1}^{N}g(i)/N,~g(i)$=$\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}/M_i$
 
 \noindent
-where $g(i)$ is the contribution of content option $i$ to the score,
+, where $g(i)$ is the contribution of content option $i$ to the score,
 $N$ is the number of contents options present in a rule,
@@
 $i$.
-
 To illustrate, consider a rule containing the option
 \CodeIn{content:``/HtmlAdaptor''} (see Section~\ref{sec:jboss}). The
-contribution of this option---\ie{}, the value of $\mathit{g(i)}$---to
+contribution of this option (\ie{}, the value of $\mathit{g(i)}$) to
 the score of the rule is $0.5$ (=$\frac{10,111+11}{10,111*2}$). The
@@
 reflects the fact that the string could be accessed by two popular
-content modifiers.  As another example, consider the option
-\CodeIn{content: ``keep-alive''}. In this case, either the modifier
-\CodeIn{header} or the modifier \CodeIn{connection} could be used to
-locate the string ``keep-alive'', stored in the field
-\CodeIn{Connection} from an http message (see
-Figure~\ref{fig:http-header-example}). Consequently the contribution
-of this heuristic function to the corresponding rule would be $0.2$
-(=$\frac{4089+29}{10111*2}$), in this case. Assuming that a rule uses
-the two content options mentioned above, the aggregated contribution
-of this heuristic function to the rule would be 0.35 (=(0.2+0.5)/2).
+content modifiers.
+
+%%%%%%%%%%%%%%
+%% Removed for space
+%%%%%%%%%%%%%%
+
+%% As another example, consider the option
+%% \CodeIn{content: ``keep-alive''}. In this case, either the modifier
+%% \CodeIn{header} or the modifier \CodeIn{connection} could be used to
+%% locate the string ``keep-alive'', stored in the field
+%% \CodeIn{Connection} from an http message (see
+%% Figure~\ref{fig:http-header-example}). Consequently the contribution
+%% of this heuristic function to the corresponding rule would be $0.2$
+%% (=$\frac{4089+29}{10111*2}$), in this case. Assuming that a rule uses
+%% the two content options mentioned above, the aggregated contribution
+%% of this heuristic function to the rule would be 0.35 (=(0.2+0.5)/2).
 
@@
 golden rule on each attack. The weight combination that led to the
-smaller sum of positions (of the golden rule) in the rankings was
-considered the best choice. We considered \emph{all} combinations of
-the following weights 0, 0.25, 0.5, 0.75, and 1. Given that we have
-four different functions and five possible weight values, we needed to
-run \tname{} on the sample of attacks for 625 (=$5^4$) times. In the
-end of the process, the weight combination
+smaller sum of rankings was considered the best choice. We considered
+\emph{all} combinations of the following weights 0, 0.25, 0.5, 0.75,
+and 1. Given that there are four different functions and five possible
+weight values, we needed to run \tname{} on the sample of attacks for
+625 (=$5^4$) times. In the end of the process, the weight combination
 $\langle{}1,0.25,0.25,1\rangle$ was the best choice we found. Although
@@
 handled manually by users. Our approach, in contrast, does not
-generate false positives w.r.t. the input traffic.
+generate false positives w.r.t.
+the input traffic.
+\Gui{Fallahi and colleagues proposes a flow-based solution for automated rule generation.
+Instead of analyzing the packets individually, their algorithm investigates and creates the rules for flows in the network.
+The algorithm was tested using the ISCX 2012 dataset and showed good results in terms of precision, but it is limited to the type of attack.
+As it won't analyse individual details of each packet, it wouldn't be able to identify and generate rules for attacks that require detailed options, such as the HTTP PESC, RCE and LFI.
+https://ieeexplore.ieee.org/abstract/document/7585840}
 
@@
 be used for repair.
+\Gui{Ganesan and colleagues introduces an idea of generating new sets of rules based on modifications of existing rules.
+They used the community edition of snort rules from 2012 for the modifications and tested with the the pcap file from the MACCDC 2012 dataset.
+It was reported an increase in the number of alerts, but it wasn't analysed the false alarmrates for abduced rules.
+Their technique is parallel to the rule creation strategies, as the objective of the modified rules is to capture false positives, not new attacks.
+https://arxiv.org/abs/1903.12101 \Fix{I'm not sure if we can consider it 'rule repair', but I don't know where to put it.}}
 

### 2020-05-25 paper/paper.tex
fix on related works
pass related work

--- 2020-05-24 paper/paper.tex
@@
 contain \emph{no} false positives or negatives with respect to the
-benign and mallicious traffic considered.\Gui{Guruprasad \cite{guruprasad} also proposes an evolutionary framework, creating their own ICMP dataset as Vollmer, but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets for non-ICMP attacks.
+benign and mallicious traffic considered.
+
+\Gui{Guruprasad \etal{}\cite{guruprasad} also proposes an evolutionary framework, creating their own ICMP dataset as Vollmer, but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets for non-ICMP attacks.
 The produced rules reported a detection rate up to 96\%.
-Although Guruprasad supports more protocols than Vollmer, our approach still supports more options and protocols, making both key differences considered when comparing with Vollmer also relevant with Guruprasad.} Muthuregunathan
-\etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
+Although Guruprasad supports more protocols than Vollmer, our approach still supports more options and protocols, making both key differences considered when comparing with Vollmer also relevant with Guruprasad.}
+
+Muthuregunathan \etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
 create rules. They characterize candidate solutions as a set of field
@@
 handled manually by users. Our approach, in contrast, does not
-generate false positives w.r.t.
-the input traffic.
-\Gui{Fallahi and colleagues proposes a flow-based solution for automated rule generation.
+generate false positives w.r.t. the input traffic.
+
+\Gui{Fallahi \etal{}\cite{fallahi} proposes a flow-based solution for automated rule generation.
 Instead of analyzing the packets individually, their algorithm investigates and creates the rules for flows in the network.
 The algorithm was tested using the ISCX 2012 dataset and showed good results in terms of precision, but it is limited to the type of attack.
-As it won't analyse individual details of each packet, it wouldn't be able to identify and generate rules for attacks that require detailed options, such as the HTTP PESC, RCE and LFI.
-https://ieeexplore.ieee.org/abstract/document/7585840}
+As it won't analyse individual details of each packet, it wouldn't be able to identify and generate rules for attacks that require detailed options, such as the HTTP PESC, RCE and LFI.}
 
@@
 be used for repair.
-\Gui{Ganesan and colleagues introduces an idea of generating new sets of rules based on modifications of existing rules.
+
+\Gui{Ganesan \etal{} introduces an idea of generating new sets of rules based on modifications of existing rules.
 They used the community edition of snort rules from 2012 for the modifications and tested with the the pcap file from the MACCDC 2012 dataset.
 It was reported an increase in the number of alerts, but it wasn't analysed the false alarmrates for abduced rules.
-Their technique is parallel to the rule creation strategies, as the objective of the modified rules is to capture false positives, not new attacks.
-https://arxiv.org/abs/1903.12101 \Fix{I'm not sure if we can consider it 'rule repair', but I don't know where to put it.}}
+Their technique is parallel to the rule creation strategies, as the objective of the modified rules is to capture variations and evolutions of attacks.}
 

### 2020-05-25 paper/paper.tex
pass related work
running time

@@
 %%  }  
-\subsubsection{Running time}\Mar{}
+\subsubsection{Running time}\Mar{}\Gui{We ran all our tests using an Intel(R) Xeon(R) CPU E3-1240 v6 and 8GB RAM.
+For attacks with only categorical and numerical options, the average running time was 6 seconds per attack.
+For attacks with the content option, the average running time was 1445 seconds.}
 
@@
 
-\Gui{Ganesan \etal{} introduces an idea of generating new sets of rules based on modifications of existing rules.
+\Gui{Ganesan \etal{}\cite{ganesan} introduces an idea of generating new sets of rules based on modifications of existing rules.
 They used the community edition of snort rules from 2012 for the modifications and tested with the the pcap file from the MACCDC 2012 dataset.

### 2020-05-26 paper/paper.tex
pass in S.IV

@@
 examples of an attack and produces on output a ranked list of
-\emph{plausible rules}, \ie{}, rules that captures only the malicious
-traffic produced by the attacks. Figure~\ref{fig:overview} shows the workflow of \tname{} as a
-pipeline of three components, appearing numbered in the figure.
-Public data sets of benign
+\emph{plausible rules}, \ie{}, rules capturing only the malicious
+traffic produced by the attacks. Figure~\ref{fig:overview} shows 
+the workflow of \tname{} as a pipeline of three components, appearing 
+numbered in the figure. As public data sets of benign
 traffic~\cite{tcpreplay,stratosphere-normal} and rules
-exist~\cite{emerging-threats-open}. Consequently, a user of \tname{}
-only needs to provide examples of attacks on input.
+exist~\cite{emerging-threats-open}, a user of \tname{}
+only needs to provide examples of attacks.
 
@@
 http is detected; otherwise, it uses the field strategy. We elaborate
-these strategies in the following.
+on these strategies in the following.
 
@@
 Section~\ref{sec:active-recon} shows a concrete example where \tname{}
-uses this strategy. ALthough we built the option-field mapping
+uses this strategy. Although we built the option-field mapping
 manually, from the \suri\ documentation, this is a one-time activity.
@@
 engineering strategy to use based on the protocol of the input
-(malicious) messages showed to be effective. We used the payload
-strategy for http and the field strategy for the other protocols. For
+(malicious) messages showed to be effective. For
 flooding attacks, \tname\ runs the same approach for each message and
@@
 candidate rule in $\mathcal R$. The algorithm mutates each of these
-rules for $N$ times, discarding one rule option each
+rules $N$ times \rui{is N an input? How does one determine it in practice?}, discarding one rule option each
 time. Line~\ref{copy} assigns a copy of the bitvector encoding
@@
 %% Section~\ref{sec:objectanalysis-benign}).
+
+\rui{paragraph on complexity? I wonder whether that is a good ideia because
+this is already a long section.}
 
@@
 \textbf{Rationale.}~The golden rule is a trusted human-written
-\nids\ rule that we obtained from credible sources \rui{which are?}\Luc{\cite{emerging-threats-open}}. \tname\ generates
-several rules that satisfy the invariant to only capture the malicious
+\nids\ rule that we obtained from credible sources~\cite{emerging-threats-open}. 
+\tname\ generates several rules that satisfy the invariant to only capture the malicious
 traffic, but the amount of data we used to evaluate that property,

### 2020-05-26 paper/paper.tex
pass in S.IV

@@
 
-\subsubsection{\label{sec:popularity-options}Popularity of options.}
+\subsubsection{\label{sec:popularity-options}Popularity of options}
 
@@
 
-\subsubsection{Number of content options.} We observed
+\subsubsection{Number of content options} We observed
 that \percContentOptions\ of the rules from the ruleset we analyzed
@@
 
-\subsubsection{Popularity of content strings.} 
+\subsubsection{Popularity of content strings} 
 This heuristic leverages content rarity to discriminate rules. It
@@
 
-\subsubsection{\label{sec:popularity-content-modifiers}Popularity of http content modifiers.} As previously
+\subsubsection{\label{sec:popularity-content-modifiers}Popularity of http content modifiers} As previously
 mentioned, \percRulesWithContent\ of \suri\ rules involve content

### 2020-05-26 paper/paper.tex
pass

@@
     .idq access & 14 & 19 & 806 & 2 \\
-    /system32/ in URI & 4 & 20 & 1881 & 2 \\
-    Basebuilder LFI & 1 & 22 & 60222 & 48 \\
+    /system32/ in URI & 4 & 20 & 1,881 & 2 \\
+    Basebuilder LFI & 1 & 22 & 60,222 & 48 \\
     Tomcat Snoop Access & 1 & 24 & 58,122 & 37 \\

### 2020-05-31 paper/paper.tex
table 5

@@
 
+\renewcommand{\arraystretch}{0.9}
+\begin{table*}[t!]
+  \footnotesize
+  \caption{\label{table:attacks}-}
+  \vspace{-2ex}
+  \centering
+    \begin{tabular}{lp{2cm}p{2.1cm}ccp{2cm}c}
+    \toprule
+    \multicolumn{1}{c}{Paper} &
+    \multicolumn{1}{c}{Input} &
+    \multicolumn{1}{c}{Dataset} &
+    \multicolumn{1}{c}{Protocol} &
+    \multicolumn{1}{c}{Output} &
+    \multicolumn{1}{c}{Evaluation} &
+    \multicolumn{1}{c}{Availability} \\
+    \midrule
+    \cite{vollmer-etal-cics2011} & Malign Traffic & Manually crafted & ICMP & Rule List & TP, FP & No\\
+    \midrule
+    \cite{fallahi} & Mixed Traffic & \cite{iscx} & HTTP, SSH, TCP & Single Rule & TP, FP, Precision, Recall, F1 & No\\    
+    \midrule
+    \cite{Kao2015AutomaticNR} & Malign Traffic, Benign Traffic, Rule Dataset & Manually crafted, \cite{emerging-threats-open} & HTTP & Single Rule & Generating time, FP & No \\
+    \midrule
+    \cite{guruprasad}& Malign Traffic & \cite{iscx} \cite{darpa1999} & ICMP, TCP & Rule List & FP & No\\
+    \midrule
+    Syrius & Malign Traffic, Benign Traffic, Rule Dataset & Manually crafted, \cite{emerging-threats-open}, \cite{iscx} & HTTP, TCP, ICMP & Rule List & Rule Ranking & Yes \\
+    \bottomrule
+  \end{tabular}
+\end{table*}
+
 Note that the number of up arrows is relatively high and the magnitude

### 2020-05-31 paper/paper.tex
fix plots

@@
             height = 0.2\textwidth,            
-            bar width=4pt,
+            bar width=3pt,
             scaled ticks=false,
@@
             ybar stacked,
+            ymax=250,
             enlargelimits=0.09,
@@
   \centering
-  \scalebox{0.8}{  
+  \scalebox{0.9}{  
     \begin{tikzpicture}
@@
           ybar stacked,
-          enlargelimits=0.08,
+          ymax=28000,
+          enlargelimits=0.12,
 %          enlargelimits=false,
@@
           xtick=data,
-          xticklabel style={rotate=30},
+          xticklabel style={rotate=35},
           nodes near coords,
@@
     \centering
-    \scalebox{0.8}{  
+    \scalebox{0.9}{  
       \begin{tikzpicture}
@@
           ybar stacked,
-          enlargelimits=0.08,
+          ymax=10000,
+          enlargelimits=0.1,
 %          enlargelimits=false,
@@
           ybar stacked,
-          enlargelimits=0.08,
+          ymax=2200,
+          enlargelimits=0.1,
 %          enlargelimits=false,
@@
   \centering
-  \scalebox{0.8}{
+  \scalebox{0.9}{
     \begin{tikzpicture}
@@
           ybar stacked,
-          enlargelimits=0.08,
+          ymax=11000,
+          enlargelimits=0.12,
 %          enlargelimits=false,

### 2020-05-31 paper/paper.tex
cleaner plots

@@
           ybar stacked,
-          ymax=28000,
+          ymax=27000,
           enlargelimits=0.12,
@@
           nodes near coords align={vertical},
-          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode, flags, tls.fingerprint},          
+          %symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode, flags, tls.fingerprint},          
+          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype},          
         ]
-        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155) (icode, 108) (flags, 97) (tls.fingerprint, 61)};
+        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155)};
       \end{axis}
@@
           nodes near coords align={vertical},
-          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          %symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
+          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7},
           ]
-          \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
+          \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535)};
         \end{axis}
@@
           nodes near coords align={vertical},
-          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},          
+          %symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},          
+          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE},          
         ]
-        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387) (INSERT, 386) (200, 352) (/index.php?, 273)};
+        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387)};
       \end{axis}

### 2020-05-31 paper/paper.tex
impact of negative examples

@@
     \centering
-    \includegraphics[scale=0.35,trim=20 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
+    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
     \caption{idqaccess}
@@
   \end{subfigure}%
+\vspace{2ex}  
+  \begin{subfigure}{.2\textwidth}
+    \centering
+    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
+    \caption{coldfusion}
+  \end{subfigure}%
+    \hspace{-8ex}
+  \begin{subfigure}{.2\textwidth}
+    \centering
+    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/inc/inc.pdf}
+    \caption{inc}    
+  \end{subfigure}%
+    \hspace{-8ex}
+  \begin{subfigure}{.2\textwidth}
+    \centering
+    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
+    \caption{jsp}    
+  \end{subfigure}%
+
 %\vspace{-2.5ex}  
-\caption{Impact of number of positive examples.}
+\caption{Impact of number of examples.}
 \label{fig:impact-number-attacks}

### 2020-05-31 paper/R/coldfusion/subject.r

@@ -1,0 +1,21 @@
+#!/usr/bin/env Rscript
+
+pdf("coldfusion.pdf", width = 5, height = 2.9)
+
+
+df1 <- read.table("data0.txt", header = FALSE)
+df2 <- read.table("data1.txt", header = FALSE)
+df3 <- read.table("data100.txt", header = FALSE)
+df4 <- read.table("data10000.txt", header = FALSE)
+df5 <- read.table("data100000.txt", header = FALSE)
+
+# adjusting margins
+par(mar = c(5, 4, 4, 0), 
+    mgp = c(1, 1, 0),
+    mai = c(0.4, 0.6, 0.1, 0),
+    mfrow = c(1,2))
+
+    bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,names=c("0","","100","","100k"), boxwex=0.2, cex.axis=1, las=1)
+
+dev.off()
+

### 2020-05-31 paper/R/idqaccess/subject.r

@@ -1,6 +1,6 @@
 #!/usr/bin/env Rscript
 
-pdf("idqaccess.pdf", width = 4.1, height = 2.9)
+pdf("idqaccess.pdf", width = 5, height = 2.9)
 
 df1 <- read.table("data1.txt", header = FALSE)
 df2 <- read.table("data2.txt", header = FALSE)

### 2020-05-31 paper/R/inc/subject.r

@@ -1,0 +1,21 @@
+#!/usr/bin/env Rscript
+
+pdf("inc.pdf", width = 5, height = 2.9)
+
+
+df1 <- read.table("data0.txt", header = FALSE)
+df2 <- read.table("data1.txt", header = FALSE)
+df3 <- read.table("data100.txt", header = FALSE)
+df4 <- read.table("data10000.txt", header = FALSE)
+df5 <- read.table("data100000.txt", header = FALSE)
+
+# adjusting margins
+par(mar = c(5, 4, 4, 0), 
+    mgp = c(1, 1, 0),
+    mai = c(0.4, 0.6, 0.1, 0),
+    mfrow = c(1,2))
+
+bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,names=c("0","","100","","100k"), boxwex=0.2, cex.axis=1, las=1)
+
+dev.off()
+

### 2020-05-31 paper/R/jsp/subject.r

@@ -1,0 +1,21 @@
+#!/usr/bin/env Rscript
+
+pdf("jsp.pdf", width = 5, height = 2.9)
+
+
+df1 <- read.table("data0.txt", header = FALSE)
+df2 <- read.table("data1.txt", header = FALSE)
+df3 <- read.table("data100.txt", header = FALSE)
+df4 <- read.table("data10000.txt", header = FALSE)
+df5 <- read.table("data100000.txt", header = FALSE)
+
+# adjusting margins
+par(mar = c(5, 4, 4, 0), 
+    mgp = c(1, 1, 0),
+    mai = c(0.4, 0.6, 0.1, 0),
+    mfrow = c(1,2))
+
+bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,names=c("0","","100","","100k"), boxwex=0.2, cex.axis=1, las=1)
+
+dev.off()
+

### 2020-05-31 paper/R/scripttag/subject.r

@@ -1,6 +1,6 @@
 #!/usr/bin/env Rscript
 
-pdf("scripttag.pdf", width = 4.1, height = 2.9)
+pdf("scripttag.pdf", width = 5, height = 2.9)
 
 df1 <- read.table("data1.txt", header = FALSE)
 df2 <- read.table("data2.txt", header = FALSE)

### 2020-06-01 paper/paper.tex
impact of negative examples
pass

@@
     7 & flags & matches on TCP flags\\
-    8 & fragbits & checks if the frag. or reserved bits are set in the IP header\\
-    9 & threshold & minimum threshold for a rule before it generates alerts\\
+    8 & fragbits & checks if the frag. bits are set in the IP header\\ % or reserved bits
+    9 & threshold & min \# messages for a rule before it generates alerts\\
     10 & content & checks if argument is present on the payload\\
@@
 
-\setlength{\tabcolsep}{8pt}
 \begin{table}[h!]
   \footnotesize
+  \setlength{\tabcolsep}{8pt}  
   \caption{\label{table:weights}Contribution of heuristic functions.}
@@
 
-\renewcommand{\arraystretch}{0.9}
 \begin{table*}[t!]
   \footnotesize
-  \caption{\label{table:attacks}-}
+  \setlength{\tabcolsep}{5pt}
+  \renewcommand{\arraystretch}{-0.01}
+  \caption{\label{table:comparison}Alternative techniques.}
   \vspace{-2ex}
   \centering
-    \begin{tabular}{lp{2cm}p{2.1cm}ccp{2cm}c}
+    \begin{tabular}{lcccccc}
     \toprule
@@
     \multicolumn{1}{c}{Input} &
+    \multicolumn{1}{c}{Output} &    
     \multicolumn{1}{c}{Dataset} &
-    \multicolumn{1}{c}{Protocol} &
-    \multicolumn{1}{c}{Output} &
+    \multicolumn{1}{c}{Protocols} &
     \multicolumn{1}{c}{Evaluation} &
-    \multicolumn{1}{c}{Availability} \\
+    \multicolumn{1}{c}{Avail.?} \\
     \midrule
-    \cite{vollmer-etal-cics2011} & Malign Traffic & Manually crafted & ICMP & Rule List & TP, FP & No\\
-    \midrule
-    \cite{fallahi} & Mixed Traffic & \cite{iscx} & HTTP, SSH, TCP & Single Rule & TP, FP, Precision, Recall, F1 & No\\    
-    \midrule
-    \cite{Kao2015AutomaticNR} & Malign Traffic, Benign Traffic, Rule Dataset & Manually crafted, \cite{emerging-threats-open} & HTTP & Single Rule & Generating time, FP & No \\
-    \midrule
-    \cite{guruprasad}& Malign Traffic & \cite{iscx} \cite{darpa1999} & ICMP, TCP & Rule List & FP & No\\
-    \midrule
-    Syrius & Malign Traffic, Benign Traffic, Rule Dataset & Manually crafted, \cite{emerging-threats-open}, \cite{iscx} & HTTP, TCP, ICMP & Rule List & Rule Ranking & Yes \\
+    \cite{vollmer-etal-cics2011} & Malign Traffic & List  & Manually crafted & ICMP & TP, FP & No\\
+%    \midrule
+    \cite{fallahi} & Mixed Traffic & One Rule  & \cite{iscx} & HTTP, SSH, TCP & TP, FP, Precision, Recall, F1 & No\\    
+%    \midrule
+    \cite{Kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
+    One Rule & Manually crafted, \cite{emerging-threats-open} & HTTP & Generating time, FP & No \\
+%    \midrule
+    \cite{guruprasad}& Malign Traffic & List  & \cite{iscx} \cite{darpa1999} &
+    ICMP, TCP & FP & No\\
+%    \midrule
+    Syrius & Malign/Benign Traffic, Rule Dataset & Ranking & Manually crafted, \cite{emerging-threats-open}, \cite{iscx} & HTTP, TCP, ICMP & Rule List & Yes \\
     \bottomrule
@@
 
-This section evaluates sensitivity of \tname{} to the number of
-attacks provided on input. Recall that \tname\ uses one example attack
-to generate the seed rule (at the Reverse Engineering step) and the
-remaining attacks to evaluate soundness of the rule (at the Rules
-Creations step). It is expected that the number and variety of
-attacks---of the same kind---plays an important role in the
-performance of the technique. To answer this research question, we selected two cases with the
-highest number of examples (see column ``\#~Vars'' on
+This section evaluates sensitivity of \tname{} to the amount of
+examples provided on input. Recall that \tname\ uses one example
+attack to generate the seed rule (at the Reverse Engineering step) and
+the remaining attacks to evaluate soundness of the rule (at the Rules
+Creations step). It is expected that the number and variety of attacks
+play a role in the performance of the technique.
+
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{.2\textwidth}
+    \centering
+    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
+    \caption{idqaccess}
+  \end{subfigure}%
+  \begin{subfigure}{.2\textwidth}
+    \centering
+    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}       \caption{scripttag}    
+  \end{subfigure}%
+%\vspace{1ex}  
+\caption{Impact of number of positive examples.}
+\label{fig:impact-number-attacks}
+%\vspace{1ex}
+\end{figure}
+
+We answer this question in two parts. Firts, we analyzed the impact of
+the amount of positive examples in results. For that, we selected two
+cases with the highest number of examples (see column ``\#~Vars'' on
 Table~\ref{table:results})---idqaccess and scripttag and measured the
@@
 
-\begin{figure}[t!]
+\begin{figure}[h!]
  \centering
-  \begin{subfigure}{.2\textwidth}
-    \centering
-    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
-    \caption{idqaccess}
-  \end{subfigure}%
-  \begin{subfigure}{.2\textwidth}
-    \centering
-    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
-    \caption{scripttag}    
-  \end{subfigure}%
-\vspace{2ex}  
  \begin{subfigure}{.2\textwidth}
@@
   \end{subfigure}%
-
 %\vspace{-2.5ex}  
-\caption{Impact of number of examples.}
-\label{fig:impact-number-attacks}
-%\vspace{1ex}
+\caption{Impact of number of negative examples.}
+\label{fig:impact-negative}
+\vspace{1ex}
 \end{figure}
 
+We conducted a similar experiment with negative examples on three
+different attacks. Figure~\ref{fig:impact-negative} shows
+results. Note that, compared with the plots of
+Figure~\ref{fig:impact-number-attacks}, the rank of the golden rule
+improves at a much slower pace. Fortunately, negative traffic is much
+easier to access.
+
 \begin{center}
-\begin{tcolorbox}[enhanced,width=3.4in,center upper,drop shadow southwest,sharp corners]
-As expected, \tname\ is sensitive to the number and choice of example
-input attacks. However, a small number of examples suffices to achieve
-best results.
+\begin{tcolorbox}[enhanced,width=3.5in,center upper,drop shadow southwest,sharp corners]
+\tname\ is sensitive to the number and choice of examples. However, a
+small number of positive examples suffices to achieve good results.
 \end{tcolorbox}
@@
 \subsubsection{Lack of comparison with other tools}
- \Gui{There are a few reasons that makes the comparison between Syrius and other tools impossible
-
-\textbf{Tools availability.} We tried contact with authors of 4 different papers \cite{vollmer-etal-cics2011}\cite{muthuregunathan-etal2009}\cite{fallahi}\cite{Kao2015AutomaticNR}, but got no response from them.
+Table~\ref{table:comparison} summarizes the characteristics of
+alternative techniques and provides details about the evaluation used
+in their original paper. Unfortunately, side comparison was not
+possible for three key reasons. First, we did not have access to the
+tools. We tried, unsuccssefully, making contact with authors to obtain
+the tools. Second, the tools output different objects. For that
+reason, a comparison based on published results did not make sense. It
+is important to recall that we did not report precision and recall as
+it is 100\% always. \tname captures, by construction, all positive
+traffic and does not capture any negative traffic. We used a stronger
+criteria instead---the position of the correct rule in the ranking. It
+is possible that other rules, higher in the ranking, are useful, but
+only domain experts can identify that. Third, we are the only
+technique to support content options, which are prevalent and
+challenging to handle. A side comparison would be unfair if we
+consider rules with those options.
+
+%%  \Gui{There are a few reasons that makes the comparison between Syrius and other tools impossible
+
+%% \textbf{Tools availability.} We tried contact with authors of 4 different papers \cite{vollmer-etal-cics2011}\cite{muthuregunathan-etal2009}\cite{fallahi}\cite{Kao2015AutomaticNR}, but got no response from them.
 Since their tools are not available online or at least not publicly, we could not run them.
 
-\textbf{Different output and specific rules.} We considered using the results they provides with results that Syrius provide, but concluded it wouldn't be possible as the results are essentially different.
+%% \textbf{Different output and specific rules.} We considered using the results they provides with results that Syrius provide, but concluded it wouldn't be possible as the results are essentially different.
 Commonly the  other results focus on showing the ratio of false negatives and false positives as a way to show the efficiency of the learning algorithm, but never comparing the generated rules to a Ground Truth, and most of times not showing individually the attacks used.
@@
 
-\textbf{'content' option support.} Another limitation is the support of 'content' option.
+%% \textbf{'content' option support.} Another limitation is the support of 'content' option.
 Most papers do not support rule generation with the 'content' option, using only classic DDoS attacks in the rule generation, like ICMP and SYN flood, which only use numerical and categorical options.

### 2020-06-01 paper/paper.tex
pass

@@
 
+%\subsection{Threats to Validity}
+
 \subsection{Discussion}
@@
 \label{sec:dos}
-
-\subsubsection{Lack of comparison with other tools}
+\label{sec:threats}
+
+%\subsubsection{Lack of comparison with other tools}
+\noindent\textbf{Lack of comparison with other tools}
 Table~\ref{table:comparison} summarizes the characteristics of
 alternative techniques and provides details about the evaluation used
-in their original paper. Unfortunately, side comparison was not
+in corresponding papers. Unfortunately, side comparison was not
 possible for three key reasons. First, we did not have access to the
@@
 consider rules with those options.
+\noindent\textbf{Running times.}~We ran all our tests using an
+Intel(R) Xeon(R) CPU E3-1240 v6 and 8GB RAM. For attacks with only
+categorical and numerical options, the average running time was 6
+seconds per attack. For attacks with the content option, the average
+running time was 1,445 seconds (=24.1 minutes).  \textbf{Threats to
+  Validity.}~Considering external validity. Our results cannot be
+generalized beyond the subjects we selected. To mitigate that threat
+we selected a variety of attacks with different
+characteristics. Unfortunately, we could not access tools for related
+techniques, but there is strong evidence that \tname\ performs very
+well and is the only technique to support content options, which are
+very important in this domain. Considering internal validity, in
+principle, the implementation of \tname\ could contain errors. We
+mitigated that threat by carefully inspecting code and
+results. Considering construct validity, the choice of ranking to
+evaluate the approach may not be useful in practice. We considered the
+use of ranking of the golden rule a reasonable choice as the approach
+places the golden rule close to the top in most cases.
+
+% For that reason, we made a side-comparison based on the common
+% examples and results reported in the original paper.
 
@@
 %%  }
-\subsubsection{Running time}\Mar{}\Gui{We ran all our tests using an Intel(R) Xeon(R) CPU E3-1240 v6 and 8GB RAM.
-For attacks with only categorical and numerical options, the average running time was 6 seconds per attack.
-For attacks with the content option, the average running time was 1445 seconds.}
 
@@
 %% of the ranking.
-
-\subsection{Threats to Validity}
-\label{sec:threats}
-
-Our experiments is affected by the usual threats. Considering external
-validity. Our results cannot be generalized beyond the subjects we
-selected. To mitigate that threat we selected a variety of attacks
-with different characteristics. Unfortunately, we could not access
-tools for related techniques. For that reason, we made a
-side-comparison based on the common examples and results reported in
-the original paper. Considering internal validity, the implementation
-of \tname\ could contain errors. We mitigated that threat by carefully
-inspecting code and results. Considering construct validity, the
-choice of ranking to evaluate the approach may not be useful in
-practice. We considered the use of ranking of the golden rule a
-reasonable choice as the approach places the golden rule close to the
-top in most cases.
 

### 2020-06-01 paper/paper.tex
minor changes

@@
 
-\begin{figure}[t!]
+\begin{figure*}[t!]
   \centering
@@
   \centering
-  \scalebox{0.9}{  
+  \scalebox{0.8}{  
     \begin{tikzpicture}
@@
     \centering
-    \scalebox{0.9}{  
+    \scalebox{0.85}{  
       \begin{tikzpicture}
@@
   \end{subfigure}% <---- don't forget this %
-  \\
+%  \\
 \begin{subfigure}{.25\textwidth}
   \centering
-  \scalebox{0.9}{  
+  \scalebox{0.8}{  
     \begin{tikzpicture}
@@
   \centering
-  \scalebox{0.9}{
+  \scalebox{0.8}{
     \begin{tikzpicture}
@@
 \caption{Distributions.}
-\end{figure}
+\end{figure*}
 
@@
 %\setlength{\tabcolsep}{5pt}
-\renewcommand{\arraystretch}{0.9}
+%renewcommand{\arraystretch}{0.9}
 \begin{table*}[t!]
@@
   \centering
-  \begin{tabular}{lcccp{10cm}}
+  \setlength{\tabcolsep}{4.5pt}
+  \renewcommand{\arraystretch}{0.35}
+  \begin{tabular}{lcccp{11cm}}
     \toprule
@@
   \end{tabular}
+  \vspace{-2ex}
 \end{table*}
@@
 
-\renewcommand{\arraystretch}{0.7}
-\setlength{\tabcolsep}{2pt}
+%\renewcommand{\arraystretch}{0.7}
+%\setlength{\tabcolsep}{2pt}
 \begin{table}[t!]
   \footnotesize
+  \setlength{\tabcolsep}{2pt}
+  \renewcommand{\arraystretch}{0.5}
   \caption{\label{table:results}Results.}
@@
 
-\begin{center}
-\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
-\tname\ successfully synthesizes the golden rules in all cases and
-that rule is ranked among the top five rules in
-\percTopFiveRanking\ of the cases.
-\end{tcolorbox}
-\end{center}
+%% \begin{center}
+%% \begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
+%% \tname\ successfully synthesizes the golden rules in all cases and
+%% that rule is ranked among the top five rules in
+%% \percTopFiveRanking\ of the cases.
+%% \end{tcolorbox}
+%% \end{center}
+
+\emph{ \textbf{Summary of results:}~\tname\ successfully synthesizes
+  the golden rules in all cases and that rule is ranked among the top
+  five rules in \percTopFiveRanking\ of the cases.  }
 
@@
   \setlength{\tabcolsep}{8pt}
+  %\setlength{\tabcolsep}{5pt}
+  \renewcommand{\arraystretch}{-0.01}
   \caption{\label{table:weights}Contribution of heuristic functions.}
@@
 
-\begin{center}
-\begin{tcolorbox}[enhanced,width=3.4in,center upper,drop shadow southwest,sharp corners]
-Every function showed impact on the performance of \tname. The function that builds on popularity of content
-modifiers showed highest impact.
-%The function that builds on  was the one that affected positively the higher number of
-%cases we analyzed.
-\end{tcolorbox}
-\end{center}
+%% \begin{center}
+%% \begin{tcolorbox}[enhanced,width=3.4in,center upper,drop shadow southwest,sharp corners]
+%% %The function that builds on  was the one that affected positively the higher number of
+%% %cases we analyzed.
+%% \end{tcolorbox}
+%% \end{center}
+
+\emph{\textbf{Summary of results:}~Every function showed impact on the
+  performance of \tname. The function that builds on popularity of
+  content modifiers showed highest impact.}
 
@@
 Creations step). It is expected that the number and variety of attacks
-play a role in the performance of the technique.
+play a role in the performance of the technique.  We answered this
+question in two parts. First, we analyzed the impact of the amount of
+positive examples in results. For that, we selected two cases with the
+highest number of examples---idqaccess and scripttag---and measured
+the position of the golden rule at the rank for an increasing number
+of examples provided to \tname. For each number of variants, we
+sampled five sets of that size to provide as input to the technique.
+
+% (see column ``\#~Vars'' on Table~\ref{table:results})
 
 \begin{figure}[h]
+  \vspace{-2ex}
   \centering
@@
   \end{subfigure}%
-%\vspace{1ex}  
-\caption{Impact of number of positive examples.}
-\label{fig:impact-number-attacks}
-%\vspace{1ex}
+  \caption{\label{fig:impact-number-attacks}Impact of number of positive examples.}
+  \vspace{-2ex}
 \end{figure}
 
-We answer this question in two parts. Firts, we analyzed the impact of
-the amount of positive examples in results. For that, we selected two
-cases with the highest number of examples (see column ``\#~Vars'' on
-Table~\ref{table:results})---idqaccess and scripttag and measured the
-position of the golden rule at the rank for an increasing number of
-examples provided to \tname. For each number of variants, we sampled
-five sets of that size to provide as input to the
-technique. Figure~\ref{fig:impact-number-attacks} shows the results
-for each attack we selected. The y-axis of each plot denotes the rank
-of the golden rule and x-axis denotes the number of attack examples
-used to synthesize rules. Each point in the x-axis is associated with
-the boxplot of the distribution of results for the corresponding
-number of variants. Results indicate that 1) the number of examples is
-important, 2) the choice of examples is important, and 3) a small
-number of examples suffices to reach the best results.
+Figure~\ref{fig:impact-number-attacks} shows the results for each
+attack we selected. The y-axis of each plot denotes the rank of the
+golden rule and x-axis denotes the number of attack examples used to
+synthesize rules. Each point in the x-axis is associated with the
+boxplot of the distribution for the corresponding number of
+variants. Results indicate that the number of examples is important,
+the choice of examples is important, and a small number of examples
+suffices to reach the best results.
 
@@
 
-\begin{center}
-\begin{tcolorbox}[enhanced,width=3.5in,center upper,drop shadow southwest,sharp corners]
-\tname\ is sensitive to the number and choice of examples. However, a
-small number of positive examples suffices to achieve good results.
-\end{tcolorbox}
-\end{center}
+%% \begin{center}
+%% \begin{tcolorbox}[enhanced,width=3.5in,center upper,drop shadow southwest,sharp corners]
+
+%% \end{tcolorbox}
+%% \end{center}
+
+\emph{ \textbf{Summary of results:}~\tname\ is sensitive to the number
+  and choice of examples. However, a small number of positive examples
+  suffices to achieve good results.}
 

### 2020-06-01 paper/paper.tex
saving space
table 3 e 4

@@
 \caption{Distributions.}
+\vspace{-4ex}
 \end{figure*}
@@
 
-\begin{figure}[t!]
+\begin{wrapfigure}[9]{r}{0.21\textwidth}
+  \vspace{-2ex}
 \centering
-\includegraphics[scale=0.5]{figs/http_header-example.png}
-\vspace{-2ex}
+\includegraphics[scale=0.6]{figs/http_header-example.png}
+\vspace{-1ex}
 \caption{Content modifiers.}
 \label{fig:http-header-example}
-\end{figure}
+\end{wrapfigure}
 
@@
     %%Apache Struts 2 & 2 & 65 & - & - \\
-    ColdFusion & 7 & 19 & 40 & 4\\
+    ColdFusion & 7 & 19 & 40 & 2\\
     iisadmin access & 3 & 17 & 605 & 1 \\        
-    JBoss JMX & 5 & 23 & 925 & 87 \\
+    JBoss JMX & 5 & 23 & 925 & 36 \\
     scripttag & 23 & 22 & 2,221 & 1 \\
-    WeBid LFI & 1 & 23 & 35,444 & 17\\    
+    WeBid LFI & 1 & 23 & 35,444 & 29\\    
     %%Wordpress Login & 8 & 59 & - & - \\
@@
     .idq access & 14 & 19 & 806 & 2 \\
-    /system32/ in URI & 4 & 20 & 1,881 & 2 \\
-    Basebuilder LFI & 1 & 22 & 60,222 & 48 \\
-    Tomcat Snoop Access & 1 & 24 & 58,122 & 37 \\
+    /system32/ in URI & 4 & 20 & 1,881 & 1 \\
+    Basebuilder LFI & 1 & 22 & 60,222 & 55 \\
+    Tomcat Snoop Access & 1 & 24 & 58,122 & 6 \\
     \bottomrule
@@
 
-\begin{table}[h!]
+\begin{table}[t!]
   \footnotesize
@@
     %%Apache Struts 2 & - & - & - & - & - & -\\
-    ColdFusion & 4 & 4 & 2\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ & 5$\uparrow$ \\
+    ColdFusion & 2 & 2 & 2\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ & 5$\uparrow$ \\
     iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
-    JBoss JMX & 87 & 87 & 15\cellcolor{lightgray}$\downarrow$ & 84 & 269$\uparrow$ \\
+    JBoss JMX & 36 & 36 & 15\cellcolor{lightgray}$\downarrow$ & 84 & 269$\uparrow$ \\
     scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
-    WeBid LFI & 17 & 17 & 144$\uparrow$ & 27$\uparrow$ & 41$\uparrow$ \\    
+    WeBid LFI & 29 & 29 & 144$\uparrow$ & 27$\uparrow$ & 41$\uparrow$ \\    
     %%Wordpress Login & 8 & - & - & - & - \\
@@
     .idq access & 2  & 2 & 1 & 1 & 15$\uparrow$ \\
-    /system32/ in URI & 2  & 2 & 6$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$\\
-    Basebuilder LFI & 48  & 48 & 846$\uparrow$ & 65$\uparrow$ & 664$\uparrow$\\
-    Tomcat Snoop Access & 37  & 37 & 1797$\uparrow$ & 6\cellcolor{lightgray}$\downarrow$ & 3580$\uparrow$\\
+    /system32/ in URI & 1  & 1 & 6$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$\\
+    Basebuilder LFI & 55  & 55 & 846$\uparrow$ & 65$\uparrow$ & 664$\uparrow$\\
+    Tomcat Snoop Access & 6 & 6 & 1797$\uparrow$ & 6\cellcolor{lightgray}$\downarrow$ & 3580$\uparrow$\\
     \bottomrule
@@
 \begin{figure}[h!]
+  \vspace{-1.5ex}
   \centering
@@
   \end{subfigure}%
-%\vspace{-2.5ex}  
   \caption{Impact of number of negative examples.}
   \label{fig:impact-negative}
-\vspace{1ex}
+  \vspace{-2ex}
 \end{figure}

### 2020-06-01 paper/paper.tex
table 4

@@
     %%Apache Struts 2 & - & - & - & - & - & -\\
-    ColdFusion & 2 & 2 & 2\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ & 5$\uparrow$ \\
+    ColdFusion & 2 & 2 & 2 & 7$\uparrow$ & 5$\uparrow$ \\
     iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
-    JBoss JMX & 36 & 36 & 15\cellcolor{lightgray}$\downarrow$ & 84 & 269$\uparrow$ \\
+    JBoss JMX & 36 & 36 & 14\cellcolor{lightgray}$\downarrow$ & 34\cellcolor{lightgray}$\downarrow$ & 269$\uparrow$ \\
     scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
-    WeBid LFI & 29 & 29 & 144$\uparrow$ & 27$\uparrow$ & 41$\uparrow$ \\    
+    WeBid LFI & 29 & 29 & 72$\uparrow$ & 39$\uparrow$ & 41$\uparrow$ \\    
     %%Wordpress Login & 8 & - & - & - & - \\
     .htaccess access & 1  & 1 & 1 & 1 & 11$\uparrow$\\
-    .idq access & 2  & 2 & 1 & 1 & 15$\uparrow$ \\
-    /system32/ in URI & 1  & 1 & 6$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$\\
-    Basebuilder LFI & 55  & 55 & 846$\uparrow$ & 65$\uparrow$ & 664$\uparrow$\\
-    Tomcat Snoop Access & 6 & 6 & 1797$\uparrow$ & 6\cellcolor{lightgray}$\downarrow$ & 3580$\uparrow$\\
+    .idq access & 2  & 2 & 2 & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ \\
+    /system32/ in URI & 1  & 1 & 4$\uparrow$ & 1 & 15$\uparrow$\\
+    Basebuilder LFI & 55  & 55 & 320$\uparrow$ & 112$\uparrow$ & 58$\uparrow$\\
+      Tomcat Snoop Access & 6 & 6 & 3$\cellcolor{lightgray}\downarrow$ & 6 & 2665$\uparrow$\\
     \bottomrule

### 2020-06-01 paper/paper.tex
pass on related work

@@
 
-\Mar{mention something on ethical hacking.. ethical-hacking}
+%\Mar{mention something on ethical hacking.. ethical-hacking}
 
@@
 
+%We discuss in the following techniques similar in goal to ours. Unfortunately
+
 \noindent
-\textbf{Alternative methods for rule synthesis.} We discuss in the
-following techniques similar in goal to ours. Unfortunately, we could
-not access the tools implementing the techniques for side
+\textbf{Alternative methods for rule synthesis.} As detailed before,
+we could not access the tools implementing the techniques for
 comparison. Nevertheless, we found that there are important
-differences between our work and related work. Vollmer
-\etal{}~\cite{vollmer-etal-cics2011} propose the use of search to
-derive rules for Snort, which are very similar to \suri's.  \Fix{Their
-approach focuses on icmp attacks.} \Gui{The dataset used by them consists of manually produced ICMP packets using packet creation tools.}. It models every field of the icmp
-packet in the chromosome of candidate solutions (\ie{}, a rule) of a
+differences between our work and prior work. Vollmer
+\etal{}~\cite{vollmer-etal-cics2011} proposed the use of search to
+derive rules for Snort, which are very similar to \suri's.  Their
+solution focused on icmp attacks for which plausible attacks can be
+created directly from packets (see
+Section~\ref{sec:rules-and-packets}). They modeled every field of the
+icmp packet in chromosomes of candidate solutions (\ie{}, a rule) of a
 genetic search. The initial population is a set of candidate solutions
-that randomly combine options, modelling the packet fields. They show
-that they were able to generate rules without false negatives and only
-3 false positives across 33K benign packets. The key differences
-between our work and Vollmer~\etal{}'s are 1) the support of different
-protocols and 2) the use of ranking to discriminate alternative
-plausible rules. As for 1, our approach supports a variety of
-protocols and options that refer to message payloads which
-significantly increase the search space. The icmp rules they generate only
-refer to message fields (see Section~\ref{sec:rules-and-packets}).
-\MyComment{\Luc{eh possivel incluir contents em regras icmp tb. o q a secao 2.3 
-diz q \textbf{geralmente} as regras nao-http possuem apenas field options}. }
-As for 2, our solution additionally ranks plausible options, which
-contain \emph{no} false positives or negatives with respect to the
-benign and mallicious traffic considered.
-
-\Gui{Guruprasad \etal{}\cite{guruprasad} also proposes an evolutionary framework, creating their own ICMP dataset as Vollmer, but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets for non-ICMP attacks.
-The produced rules reported a detection rate up to 96\%.
-Although Guruprasad supports more protocols than Vollmer, our approach still supports more options and protocols, making both key differences considered when comparing with Vollmer also relevant with Guruprasad.}
-
-Muthuregunathan \etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
+that randomly combine options, modelling the packet fields. They
+showed that they were able to generate rules without false negatives
+and only 3 false positives across 33K benign packets. The key
+differences between our work and Vollmer~\etal{}'s are 1) the support
+of different protocols and 2) the use of ranking to discriminate
+alternative plausible rules. As for 1, our approach supports a variety
+of protocols and options (\eg{}, options that refer to message
+payloads). The icmp rules they generate only refer to message fields.
+\MyComment{\Luc{eh possivel incluir contents em regras icmp tb. o q a
+    secao 2.3 diz q \textbf{geralmente} as regras nao-http possuem
+    apenas field options}. } As for 2, our solution additionally ranks
+plausible options, which contain \emph{no} false positives or
+negatives with respect to the benign and mallicious traffic
+considered. Guruprasad \etal{}\cite{guruprasad} also proposes an
+evolutionary framework, creating their own ICMP dataset as Vollmer,
+but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx}
+datasets for non-ICMP attacks. Although Guruprasad supports more
+protocols than Vollmer, our approach still supports a larger set of
+protocols and the ability to ranks rules. Muthuregunathan
+\etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
 create rules. They characterize candidate solutions as a set of field
@@
 
+%% \Gui{Guruprasad \etal{}\cite{guruprasad} also proposes an evolutionary framework, creating their own ICMP dataset as Vollmer, but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets for non-ICMP attacks.
+The produced rules reported a detection rate up to 96\%.
+Although Guruprasad supports more protocols than Vollmer, our approach still supports more options and protocols, making both key differences considered when comparing with Vollmer also relevant with Guruprasad.}
+
 \Gui{Fallahi \etal{}\cite{fallahi} proposes a flow-based solution for automated rule generation.

### 2020-06-01 paper/paper.tex
saving space

@@
 attacks. Unfortunately, attackers actively look for new mechanisms to
+
 \begin{wrapfigure}[8]{r}{0.21\textwidth}
@@
 \emph{plausible rules}, \ie{}, rules capturing only the malicious
-traffic produced by the attacks. Figure~\ref{fig:overview} shows 
-the workflow of \tname{} as a pipeline of three components, appearing 
-numbered in the figure. As public data sets of benign
+traffic produced by the attacks. Figure~\ref{fig:overview} shows the
+workflow of \tname{} as a pipeline of three components, numbered in
+the figure. As public data sets of benign
 traffic~\cite{tcpreplay,stratosphere-normal} and rules
-exist~\cite{emerging-threats-open}, a user of \tname{}
-only needs to provide examples of attacks.
+exist~\cite{emerging-threats-open}, \tname{}'s users only needs to
+provide examples of attacks.
 
@@
 that observation to search for plausible rules containing a proper
-subset of options from the seed options that captures \emph{only} the
+subset of options (from the seed options) that captures \emph{only} the
 malicious traffic. Finally, \tname{} uses heuristic functions,
@@
 
-%\setlength{\tabcolsep}{5pt}
-%renewcommand{\arraystretch}{0.9}
 \begin{table*}[t!]
@@
 \newcommand{\textRQone}{How well \tname\ ranks the golden rule?}
-%\vspace{0.2cm}
 \begin{itemize}[leftmargin=*,label={}]
@@
 %\vspace{0.2cm}
+
 \begin{itemize}[leftmargin=*,label={}]
@@
 \newcommand{\textRQthree}{How sensitive \tname\ is to the amount of
-  positive examples?} \rui{should we also investigate the impact of the negative ones?}
-%\vspace{0.2cm}
+  positive and negative examples?}
+
 \begin{itemize}[leftmargin=*,label={}]
@@
     \end{tabular}
+    \vspace{-3ex}
 \end{table*}
@@
 
-\begin{figure}[h]
+\begin{wrapfigure}[11]{r}{0.28\textwidth}  
   \vspace{-2ex}
   \centering
-  \begin{subfigure}{.2\textwidth}
+  \begin{subfigure}{.15\textwidth}
     \centering
     \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
+    \vspace{-3ex}
     \caption{idqaccess}
   \end{subfigure}%
-  \begin{subfigure}{.2\textwidth}
+  \begin{subfigure}{.15\textwidth}
     \centering
-    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}       \caption{scripttag}    
+    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
+    \vspace{-3ex}        
+    \caption{scripttag}
   \end{subfigure}%
@@
   \vspace{-2ex}
-\end{figure}
-
-Figure~\ref{fig:impact-number-attacks} shows the results for each
+\end{wrapfigure}
+
+Figure~\ref{fig:impact-number-attacks} shows results for each
 attack we selected. The y-axis of each plot denotes the rank of the
@@
     \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
+    \vspace{-1ex}        
     \caption{coldfusion}
@@
     \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/inc/inc.pdf}
+    \vspace{-1ex}    
     \caption{inc}    
@@
     \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
+    \vspace{-1ex}    
     \caption{jsp}    
@@
 handled manually by users. Our approach, in contrast, does not
-generate false positives w.r.t. the input traffic.
+generate false positives w.r.t. the input traffic. Fallahi
+\etal{}\cite{fallahi} proposed flow-based synthesis approach to rule
+generation. Instead of analyzing packets individually, their algorithm
+investigates and creates the rules for flows in the network. The
+algorithm was tested using the ISCX 2012 dataset and showed good
+results in terms of precision, but it is limited to that type of
+attack. As it won't analyse individual details of each packet, it
+wouldn't be able to identify and generate rules for attacks that
+require detailed options, such as the HTTP PESC, RCE and LFI.
+\textbf{Rule repair.}~Hallahan~\etal{}~\cite{8102263} propose to
+repair firewall rules based on false positive detected by the
+rules. Although firewall and \nids\ rules have differences and the
+problem addressed is different---repair not synthesis---the work is
+similar in spirit and inspiring. We remain to evaluate how \tname\ can
+be used for repair.  Ganesan \etal{}\cite{ganesan} applied similar
+ideas in the context of IDS, \ie{}, they proposed to adjust existing
+rules with the goal of capturing variations of existing attacks. They
+used the community edition of snort rules from 2012 for the
+modifications and tested them with pcap file from the MACCDC 2012
+dataset. Their technique is complementary to \tname{}.
+\textbf{Anomaly-based intrusion detection.}
+Cordy~\etal{}\cite{cordy-etal-issta19} used genetic search to not only
+to generate attacks capable of penetrating the tested anomaly-based
+NIDS, but also to repair the NIDS to identify these
+attacks. \tname\ and their approach are complementary as anomaly-based
+and rule-based \nids\ are complementary (see
+Section~\ref{sec:background}).
+
+%% It was reported an increase in the number of alerts, but it
+%% wasn't analysed the false alarmrates for abduced rules.
+%% parallel to the rule creation strategies, as the
+%% objective of the modified rules is to capture variations and
+%% evolutions of attacks.
+
+\noindent
 
@@
 Although Guruprasad supports more protocols than Vollmer, our approach still supports more options and protocols, making both key differences considered when comparing with Vollmer also relevant with Guruprasad.}
-
-\Gui{Fallahi \etal{}\cite{fallahi} proposes a flow-based solution for automated rule generation.
-Instead of analyzing the packets individually, their algorithm investigates and creates the rules for flows in the network.
-The algorithm was tested using the ISCX 2012 dataset and showed good results in terms of precision, but it is limited to the type of attack.
-As it won't analyse individual details of each packet, it wouldn't be able to identify and generate rules for attacks that require detailed options, such as the HTTP PESC, RCE and LFI.}
 
@@
 
-\noindent
-\textbf{Rule repair.}~Hallahan~\etal{}~\cite{8102263} propose to
-repair firewall rules based on false positive detected by the
-rules. Although firewall and \nids\ rules have differences and the
-problem addressed is different---repair not synthesis---the work is
-similar in spirit and inspiring. We remain to evaluate how \tname\ can
-be used for repair.
-
-\Gui{Ganesan \etal{}\cite{ganesan} introduces an idea of generating new sets of rules based on modifications of existing rules.
-They used the community edition of snort rules from 2012 for the modifications and tested with the the pcap file from the MACCDC 2012 dataset.
-It was reported an increase in the number of alerts, but it wasn't analysed the false alarmrates for abduced rules.
-Their technique is parallel to the rule creation strategies, as the objective of the modified rules is to capture variations and evolutions of attacks.}
-
-\noindent
-\textbf{Anomaly-based intrusion detection.}
-Cordy~\etal{}\cite{cordy-etal-issta19} propose the use of a genetic
-algorithm to not only generate attack schemes capable of penetrating
-the tested anomaly-based NIDS, but also repairing the NIDS to identify
-the given attack scheme. Although our approach and theirs use machine
-intelligence to detect attacks, anomaly-based and rule-based
-\nids\ are complementary, as discussed on
-Section~\ref{sec:background}.
-
 %% \Gui{\textbf{Automated Repair By Example for Firewalls -} Ruzica repara e modifica regras de firewall a partir de uma entrada contendo exemplos ou behaviour.
@@
 Network Intrusion Detectors are an important defense mechanism against
-networks attacks. This paper presents \tname{}, a novel approach to
-generate rules for Rule-based \nids. \tname\ leverages positive and
-negative traffic and examples of trusted rules to synthesize
-rules. \tname{} is organized as a pipeline of three components that
-1)~create an over-specified seed rule, 2)~derive plausible rules from
-the seed, and 3)~rank plausible rules. Experimental results provide
-early evidence showing that \tname\ is very promising. It was able to
-rank the correct rule among the first five rules of the reported
-ranking in \percTopFiveRanking\ of the cases we analyzed. The data set
-containing our experimental artifacts is publicly accessible at
-\ourdataset.
-
-\rui{thinking out loud here: one of the motivations is that it is time-consuming
-and troublesome to generate new rules. Time-consuming is not being addressed in 
-the RQs and I think that, albeit it is difficult to conduct such a study, we 
-shuold discuss it. I would think that mentinoing the human-on-the-loop (not in-the-loop)
-should also be dscussed (this is more about the troublesomeness)}
+networks attacks. Generating rules to capture new attacks can be
+time-consuming and error-prone. This paper presents \tname{}, a novel
+approach to generate rules for Rule-based \nids. \tname\ leverages
+positive and negative traffic and examples of trusted rules to
+synthesize rules. \tname{} is organized as a pipeline of three
+components that 1)~create an over-specified seed rule, 2)~derive
+plausible rules from the seed, and 3)~rank plausible
+rules. Experimental results provide early evidence showing that
+\tname\ is promising. It was able to rank the correct rule among the
+first five rules of the reported ranking in \percTopFiveRanking\ of
+the cases we analyzed, for example. The data set containing our
+experimental artifacts is publicly accessible at \ourdataset.
+
+%\rui{thinking out loud here: one of the motivations is that it is
+%time-consuming and troublesome to generate new rules. Time-consuming
+%is not being addressed in the RQs and I think that, albeit it is
+%difficult to conduct such a study, we shuold discuss it. I would
+%think that mentinoing the human-on-the-loop (not in-the-loop) should also be dscussed (this is more about the troublesomeness)}
 

### 2020-06-01 paper/paper.tex
fix missing figure

@@
 
+\begin{figure}[h!]
+  \lstinputlisting[language=C,numbers=none,keywords={content}]{adaptor-golden-rule.suricata}
+  \vspace{-1ex}  
+  \caption{Suricata rule for JBoss JMX attack.}
+  \vspace{-2ex}  
+  \label{fig:adaptor-golden-rule.suricata}
+\end{figure}
+
 For the field strategy, \tname{} parses the input messages looking for
@@
 \percRulesWithContent{} of these rules contain at least one content
-option. Figure~\ref{fig:rule-jboss} shows an example of a rule that
+option. Figure~\ref{fig:adaptor-golden-rule.suricata} shows an example of a rule that
 captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the

### 2020-06-01 paper/paper.tex
macros update

@@
 able to synthesize the correct rule in all cases we analyzed and
-ranked the correct rule among the first five rules of the reported
+ranked the correct rule among the first ten rules of the reported
 ranking in \percTopFiveRanking\ of the cases.
@@
 The observed results are encouraging. For 
-example, \tname\ ranked the correct rule among the top five rules in \Fix{fix values}
-\percTopFiveRanking\ of the cases and ranked the correct rule among 
-the top ten rules in \topTenAttacks of the \totalAttacks cases. \rui{to discuss, I find 
+example, \tname\ ranked the correct rule among the top ten rules in \Fix{fix values}
+\percTopFiveRanking\ of the cases \Fix{and ranked the correct rule among 
+the top ten rules in \topTenAttacks of the \totalAttacks cases.} \rui{to discuss, I find 
 this a bit weak.}\Mar{agree. let us try to add more examples and data.}
@@
 cases. It reported the ground truth rule in \emph{all} cases and
-reported that rule within the top five rules of the ranking in
+reported that rule within the top ten rules of the ranking in
 \percTopFiveRanking\ of the cases.
@@
   the golden rules in all cases and that rule is ranked among the top
-  five rules in \percTopFiveRanking\ of the cases.  }
+  ten rules in \percTopFiveRanking\ of the cases.  }
 
@@
 \tname\ is promising. It was able to rank the correct rule among the
-first five rules of the reported ranking in \percTopFiveRanking\ of
+first ten rules of the reported ranking in \percTopFiveRanking\ of
 the cases we analyzed, for example. The data set containing our

### 2020-06-01 paper/paper.tex
comment on V.E

@@
 challenging to handle. A side comparison would be unfair if we
-consider rules with those options.
+consider rules with those options.\Gui{I think that it's not good to say we are the only techinique to support content options.
+The other papers do not explain well how they support it, if they support for http (which is the most relevant one) or even if they support it at all.
+\cite{Kao2015AutomaticNR} manages contents in a similar way to ours, but it requires a manual step and their provided email doesn't even exist anymore.
+\cite{vollmer-etal-cics2011} and \cite{guruprasad} do not support http and don't explain well how they deal with ICMP contents.
+\cite{fallahi} technique deals with flows and not individual packets, so it doesn't even mention supporting contents.}
 \noindent\textbf{Running times.}~We ran all our tests using an

### 2020-06-02 paper/paper.tex
pass on IV.B

@@
 The observed results are encouraging. For 
-example, \tname\ ranked the correct rule among the top ten rules in \Fix{fix values}
-\percTopFiveRanking\ of the cases \Fix{and ranked the correct rule among 
-the top ten rules in \topTenAttacks of the \totalAttacks cases.} \rui{to discuss, I find 
+example, \tname\ ranked the correct rule among the top ten rules in 
+\percTopFiveRanking\ of the cases. \rui{to discuss, I find 
 this a bit weak.}\Mar{agree. let us try to add more examples and data.}
@@
 candidate rule in $\mathcal R$. The algorithm mutates each of these
-rules $N$ times \rui{is N an input? How does one determine it in practice?}, discarding one rule option each
+rules $N$ times \rui{is N an input?
+How does one determine it in practice?} \Gui{Yes.
+Higher N values increases the chances of creating every possible rule for the attack (decreasing the negative effects of the non-determinism), at the same time that increases the computational cost.
+So the value can be increased in case the desired rule is not being created, or it can be reduced if the code is creating too many rules.
+In our code, we chose manually N=2, as it was already enough for showing the desired results.}, discarding one rule option each
 time. Line~\ref{copy} assigns a copy of the bitvector encoding

### 2020-06-02 paper/paper.tex
table v

@@
     \multicolumn{1}{c}{Dataset} &
-    \multicolumn{1}{c}{Protocols} &
+    \multicolumn{1}{c}{Evaluated Protocols} &
     \multicolumn{1}{c}{Evaluation} &
@@
     \midrule
-    \cite{vollmer-etal-cics2011} & Malign Traffic & List  & Manually crafted & ICMP & TP, FP & No\\
+    Vollmer\cite{vollmer-etal-cics2011} & Malign Traffic & List  & Manually crafted & ICMP & TP, FP & No\\
 %    \midrule
-    \cite{fallahi} & Mixed Traffic & One Rule  & \cite{iscx} & HTTP, SSH, TCP & TP, FP, Precision, Recall, F1 & No\\    
+    Fallahi\cite{fallahi} & Mixed Traffic & One Rule  & \cite{iscx} & HTTP, SSH, TCP & TP, FP, Precision, Recall, F1 & No\\    
 %    \midrule
-    \cite{Kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
+    Kao\cite{kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
     One Rule & Manually crafted, \cite{emerging-threats-open} & HTTP & Generating time, FP & No \\
 %    \midrule
-    \cite{guruprasad}& Malign Traffic & List  & \cite{iscx} \cite{darpa1999} &
+    Guruprasad\cite{guruprasad}& Malign Traffic & List  & \cite{iscx} \cite{darpa1999} &
     ICMP, TCP & FP & No\\

### 2020-06-02 paper/paper.tex
minor changes

@@
 \noindent
-\textbf{Rationale.}~The golden rule is a trusted human-written
-\nids\ rule that we obtained from credible sources~\cite{emerging-threats-open}. 
-\tname\ generates several rules that satisfy the invariant to only capture the malicious
-traffic, but the amount of data we used to evaluate that property,
-albeit large, is limited. Assessing the quality of each rule
+\textbf{Rationale.}~Assessing the quality of each rule
 \tname\ produces (\ie{}, fitness to the attack) requires domain
-knowledge and much effort. The position of the golden rule at the
-ranking produced by \tname\ is a proxy measure to the effort of a
-system administrator in finding the rule that fits the purpose.
+knowledge and effort. The ranking position of the golden rule is a
+proxy to measure the effort of a system administrator in finding the
+rule that fits the purpose.  The golden rule is a trusted
+human-written \nids\ rule that we obtained from credible
+sources~\cite{emerging-threats-open}.
+
+%\tname\ generates several rules that satisfy the invariant to only
+%capture the malicious traffic, but the amount of data we used to evaluate that property, albeit large, is limited.
 
@@
 \textbf{Rationale.}~We proposed a variety of heuristic functions to
-rank the plausible rules that \tname\ created during minimization (as
-per step 2 in the workflow). Recall that these functions are based on
-the data from a public ruleset~\cite{emerging-threats-open} (see
+rank the plausible rules that \tname\ created during
+minimization. Recall that these functions are based on the data from a
+public ruleset~\cite{emerging-threats-open} (see
 Section~\ref{sec:ranking}). It is therefore important to understand
@@
 
-The following sections describe the objects we used to evaluate the
-technique (Section~\ref{sec:dataset-benign}), answer the posed
-research questions (sections~\ref{sec:answer-rqone},
-and~\ref{sec:answer-rqtwo}, and~\ref{sec:answer-rqthree}), and
-discuss threats to the validity of the
-experiments~(Section~\ref{sec:threats}).
+The following sections describe the objects we used in the evaluation
+(Section~\ref{sec:dataset-benign}), answer the research questions
+(sections~\ref{sec:answer-rqone}, C, D), and discuss
+results~(Section~\ref{sec:threats}).
+%and~\ref{sec:answer-rqtwo}, and~\ref{sec:answer-rqthree}),
 
@@
 the position of the golden rule at the rank for an increasing number
-of examples provided to \tname. For each number of variants, we
-sampled five sets of that size to provide as input to the technique.
-
-% (see column ``\#~Vars'' on Table~\ref{table:results})
-
-\begin{wrapfigure}[11]{r}{0.28\textwidth}  
+\begin{wrapfigure}[10]{r}{0.28\textwidth}  
   \vspace{-2ex}
@@
     \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
-    \vspace{-3ex}        
+    \vspace{-2ex}        
     \caption{scripttag}
   \end{subfigure}%
-  \caption{\label{fig:impact-number-attacks}Impact of number of positive examples.}
-  \vspace{-2ex}
+    \vspace{-1ex}          
+  \caption{\label{fig:impact-number-attacks}Positive examples.}
+  \vspace{-1ex}
 \end{wrapfigure}
-
+of examples provided to \tname. For each number of variants, we
+sampled five sets of that size to provide as input to the technique.
 Figure~\ref{fig:impact-number-attacks} shows results for each
@@
 
-\begin{figure}[h!]
+\begin{wrapfigure}[11]{r}{0.28\textwidth}  
   \vspace{-1.5ex}
   \centering
-  \begin{subfigure}{.2\textwidth}
+  %% \begin{subfigure}{.2\textwidth}
+  %%   \centering
+  %%   \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
+  %%   \vspace{-1ex}        
+  %%   \caption{coldfusion}
+  %% \end{subfigure}%
+  %% \hspace{-8ex}
+  \begin{subfigure}{.14\textwidth}
     \centering
-    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
-    \vspace{-1ex}        
-    \caption{coldfusion}
-  \end{subfigure}%
-  \hspace{-8ex}
-  \begin{subfigure}{.2\textwidth}
-    \centering
-    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/inc/inc.pdf}
+    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/inc/inc.pdf}
     \vspace{-1ex}    
@@
   \end{subfigure}%
-  \hspace{-8ex}
-  \begin{subfigure}{.2\textwidth}
+  \begin{subfigure}{.14\textwidth}
     \centering
-    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
+    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
     \vspace{-1ex}    
@@
   \end{subfigure}%
-  \caption{Impact of number of negative examples.}
+  \caption{Negative examples.}
   \label{fig:impact-negative}
   \vspace{-2ex}
-\end{figure}
+\end{wrapfigure}
 

### 2020-06-02 paper/paper.tex
pass on related work

@@
 
+%% Third, we are the only
+%% technique to support content options, which are prevalent and
+%% challenging to handle. A side comparison would be unfair if we
+%% consider rules with those options.\Gui{I think that it's not good to
+%%   say we are the only techinique to support content options. The other
+%%   papers do not explain well how they support it, if they support for
+%%   http (which is the most relevant one) or even if they support it at
+%%   all. \cite{Kao2015AutomaticNR} manages contents in a similar way to
+%%   ours, but it requires a manual step and their provided email doesn't
+%%   even exist anymore. \cite{vollmer-etal-cics2011} and
+%%   \cite{guruprasad} do not support http and don't explain well how
+%%   they deal with ICMP contents. \cite{fallahi} technique deals with
+%%   flows and not individual packets, so it doesn't even mention
+%%   supporting contents.}
+
 %\subsubsection{Lack of comparison with other tools}
@@
 in corresponding papers. Unfortunately, side comparison was not
-possible for three key reasons. First, we did not have access to the
+possible for two central reasons. First, we did not have access to the
 tools. We tried, unsuccssefully, making contact with authors to obtain
 the tools. Second, the tools output different objects. For that
-reason, a comparison based on published results did not make sense. It
+reason, a comparison based on published results was not realistic. It
 is important to recall that we did not report precision and recall as
@@
 is possible that other rules, higher in the ranking, are useful, but
-only domain experts can identify that. Third, we are the only
-technique to support content options, which are prevalent and
-challenging to handle. A side comparison would be unfair if we
-consider rules with those options.\Gui{I think that it's not good to say we are the only techinique to support content options.
-The other papers do not explain well how they support it, if they support for http (which is the most relevant one) or even if they support it at all.
-\cite{Kao2015AutomaticNR} manages contents in a similar way to ours, but it requires a manual step and their provided email doesn't even exist anymore.
-\cite{vollmer-etal-cics2011} and \cite{guruprasad} do not support http and don't explain well how they deal with ICMP contents.
-\cite{fallahi} technique deals with flows and not individual packets, so it doesn't even mention supporting contents.}
+only domain experts can identify that.
 \noindent\textbf{Running times.}~We ran all our tests using an

### 2020-06-02 paper/paper.tex
done V.0 and V.A

@@
 data set is part of the TcpReplay~\cite{tcpreplay} open source
-project. It includes real network traffic of a busy private network
-access point to the Internet. We also used data sets from the
-Stratosphere IPS Project~\cite{stratosphere-normal} including, for
-example, the regular traffic of a university network.
-\noindent\textbf{Ground Truth.}~To objectively evaluate \tname{}'s performance we
-compared the rules it generates with \emph{trusted rules}. We used
-the ``Emerging Threats Open Ruleset''~\cite{emerging-threats-open} for
-that. After finding a matching pair of attack and corresponding rule
-using the procedure described in
-Section~\ref{subsec:malicious-traffic}, we reproduced the attack to
-confirm the rule captures the attack in isolation (and no noise was
-introduced in the process). Column ``Ground Truth'' from
-Table~\ref{table:attacks} only shows the essential
-options. Documentation options (\eg{}, ``msg'') are discarded.
+project. It includes real network traffic of a private network. We
+also used data sets from the Stratosphere IPS
+Project~\cite{stratosphere-normal} including, for example, the regular
+traffic of a university network.
+\noindent\textbf{Ground Truth.}~To objectively evaluate \tname{}'s
+performance we compared the rules it generates with \emph{trusted
+  rules}. We used the ``Emerging Threats Open
+Ruleset''~\cite{emerging-threats-open} for that. After finding a
+matching pair of attack and corresponding rule using the procedure
+described in Section~\ref{subsec:malicious-traffic}, we reproduced the
+attack to confirm the rule captures the attack in isolation. Column
+``Ground Truth'' from Table~\ref{table:attacks} only shows the
+essential options. Documentation options (\eg{}, ``msg'') are
+discarded.
 

### 2020-06-02 paper/paper.tex
macro update

--- 2020-06-02 paper/paper.tex
@@
 \noindent
-\textbf{Rationale.}~Assessing the quality of each rule
+\textbf{Rationale.}~The golden rule is a trusted human-written
+\nids\ rule that we obtained from credible sources~\cite{emerging-threats-open}. 
+\tname\ generates several rules that satisfy the invariant to only capture the malicious
+traffic, but the amount of data we used to evaluate that property,
+albeit large, is limited. Assessing the quality of each rule
 \tname\ produces (\ie{}, fitness to the attack) requires domain
-knowledge and effort. The ranking position of the golden rule is a
-proxy to measure the effort of a system administrator in finding the
-rule that fits the purpose.  The golden rule is a trusted
-human-written \nids\ rule that we obtained from credible
-sources~\cite{emerging-threats-open}.
-
-%\tname\ generates several rules that satisfy the invariant to only
-%capture the malicious traffic, but the amount of data we used to evaluate that property, albeit large, is limited.
+knowledge and much effort. The position of the golden rule at the
+ranking produced by \tname\ is a proxy measure to the effort of a
+system administrator in finding the rule that fits the purpose.
 
@@
 \textbf{Rationale.}~We proposed a variety of heuristic functions to
-rank the plausible rules that \tname\ created during
-minimization. Recall that these functions are based on the data from a
-public ruleset~\cite{emerging-threats-open} (see
+rank the plausible rules that \tname\ created during minimization (as
+per step 2 in the workflow). Recall that these functions are based on
+the data from a public ruleset~\cite{emerging-threats-open} (see
 Section~\ref{sec:ranking}). It is therefore important to understand
@@
 
-The following sections describe the objects we used in the evaluation
-(Section~\ref{sec:dataset-benign}), answer the research questions
-(sections~\ref{sec:answer-rqone}, C, D), and discuss
-results~(Section~\ref{sec:threats}).
-%and~\ref{sec:answer-rqtwo}, and~\ref{sec:answer-rqthree}),
+The following sections describe the objects we used to evaluate the
+technique (Section~\ref{sec:dataset-benign}), answer the posed
+research questions (sections~\ref{sec:answer-rqone},
+and~\ref{sec:answer-rqtwo}, and~\ref{sec:answer-rqthree}), and
+discuss threats to the validity of the
+experiments~(Section~\ref{sec:threats}).
 
@@
 
-\newcommand{\numNonContentAttacks}{4}
+\newcommand{\numNonContentAttacks}{11}
 \newcommand{\numContentAttacks}{10}
@@
 the position of the golden rule at the rank for an increasing number
-\begin{wrapfigure}[10]{r}{0.28\textwidth}  
+of examples provided to \tname. For each number of variants, we
+sampled five sets of that size to provide as input to the technique.
+
+% (see column ``\#~Vars'' on Table~\ref{table:results})
+
+\begin{wrapfigure}[11]{r}{0.28\textwidth}  
   \vspace{-2ex}
@@
     \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
-    \vspace{-2ex}        
+    \vspace{-3ex}        
     \caption{scripttag}
   \end{subfigure}%
-    \vspace{-1ex}          
-  \caption{\label{fig:impact-number-attacks}Positive examples.}
-  \vspace{-1ex}
+  \caption{\label{fig:impact-number-attacks}Impact of number of positive examples.}
+  \vspace{-2ex}
 \end{wrapfigure}
-of examples provided to \tname. For each number of variants, we
-sampled five sets of that size to provide as input to the technique.
+
 Figure~\ref{fig:impact-number-attacks} shows results for each
@@
 
-\begin{wrapfigure}[11]{r}{0.28\textwidth}  
+\begin{figure}[h!]
   \vspace{-1.5ex}
   \centering
-  %% \begin{subfigure}{.2\textwidth}
-  %%   \centering
-  %%   \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
-  %%   \vspace{-1ex}        
-  %%   \caption{coldfusion}
-  %% \end{subfigure}%
-  %% \hspace{-8ex}
-  \begin{subfigure}{.14\textwidth}
+  \begin{subfigure}{.2\textwidth}
     \centering
-    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/inc/inc.pdf}
+    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
+    \vspace{-1ex}        
+    \caption{coldfusion}
+  \end{subfigure}%
+  \hspace{-8ex}
+  \begin{subfigure}{.2\textwidth}
+    \centering
+    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/inc/inc.pdf}
     \vspace{-1ex}    
@@
   \end{subfigure}%
-  \begin{subfigure}{.14\textwidth}
+  \hspace{-8ex}
+  \begin{subfigure}{.2\textwidth}
     \centering
-    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
+    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
     \vspace{-1ex}    
@@
   \end{subfigure}%
-  \caption{Negative examples.}
+  \caption{Impact of number of negative examples.}
   \label{fig:impact-negative}
   \vspace{-2ex}
-\end{wrapfigure}
+\end{figure}
 

### 2020-06-02 paper/paper.tex
merging
RQ1

--- 2020-06-02 paper/paper.tex
@@
 
-\newcommand{\numNonContentAttacks}{4}
-\newcommand{\numContentAttacks}{10}
-
 \subsection{Objects of Analysis}
@@
 
-RQ1 addresses the performance of \tname. The metric we use is the
-position of the golden rule at the ranked list of plausible rules
-reported by \tname\ (see Figure~\ref{fig:overview}). This metric is a
-proxy of human performance in analyzing the list of plausible rules.
+RQ1 assesses \tname's performance. The metric we use is the position
+of the golden rule at the ranked list of plausible rules reported by
+\tname\ (see Figure~\ref{fig:overview}). This metric is a proxy of
+human performance in analyzing the list of plausible rules.
 
@@
 name of the attack, column ``\#~Vars.'' shows the number of variants
-of the attack, column ``\#~Option Cand. (step 1)'' shows the number of
+of the attack, column ``\#~Option Cand. (step~1)'' shows the number of
 option candidates reported by \tname\ at step 1, column ``\#~Plausible
-Rules (step 2)'' shows the number of plausible rules reported by
-\tname\ at step 2, and column ``Rank (step 3)'' shows the position of the
-golden rule in the ranked list reported at step 3.
-It is important to indicate that column ``\#~Vars.'' shows the input of
-\tname\ that varies across different attacks. Although \tname\ also
-depends on the data sets of benign traffic and existing \suri\ rules,
-these inputs are constant across different executions of the tool. The
-rest of the columns show the progress of the tool at various stages
-and provide an indication of complexity of the synthesis problem.
+Rules (step~2)'' shows the number of rules reported by \tname\ at step
+2, and column ``Rank (step~3)'' shows the position of the golden rule
+in the reported ranking.  Column ``\#~Vars.''  shows the number of
+variants of a given attack. The other data that \tname\ depends upon
+are not mentioned as they are constant. The rest of the columns show
+the progress of the tool at various stages and provide an indication
+of complexity of the synthesis problem.
 
@@
 cases. It reported the ground truth rule in \emph{all} cases and
-reported that rule within the top ten rules of the ranking in
-\percTopFiveRanking\ of the cases.
+reported that rule within the top-1, top-3, top-5, and top-10 rules of
+the ranking, respectively, in \percTopOneRanking\,
+\percTopThreeRanking, \percTopFiveRanking, and \percTopTenRanking\ of
+the cases.
 
@@
   the golden rules in all cases and that rule is ranked among the top
-  ten rules in \percTopFiveRanking\ of the cases.  }
+  three rules of the ranking in \percTopThreeRanking\ of the cases.  }
 

### 2020-06-02 paper/paper.tex
fixing bib ref

@@
 %    \midrule
-    Kao\cite{kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
+    Kao\cite{Kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
     One Rule & Manually crafted, \cite{emerging-threats-open} & HTTP & Generating time, FP & No \\

### 2020-06-02 paper/paper.tex
saving space fig 9
pass

@@
 
+%% The other data that \tname\ depends upon are
+%% not mentioned as they are constant. The rest of the columns show the
+%% progress of the tool at various stages and provide an indication of
+%% complexity of the synthesis problem.
+
 Table~\ref{table:results} shows results. Column ``Name'' shows the
@@
 2, and column ``Rank (step~3)'' shows the position of the golden rule
-in the reported ranking.  Column ``\#~Vars.''  shows the number of
-variants of a given attack. The other data that \tname\ depends upon
-are not mentioned as they are constant. The rest of the columns show
-the progress of the tool at various stages and provide an indication
-of complexity of the synthesis problem.
-
-Observe that the number of option candidates is considerably higher
-for attacks captured with content options compared to attacks not
-involving content options. This happens because for content attacks
-the number of option candidates depend on the unbounded size of the
-http payload whereas for attacks of different protocols the number of
-option candidates depend on the layout of the messages. Also observe
-that the number of plausible rules is much smaller for non-content
-attacks compared to content attacks.
+in the reported ranking. Observe that the number of option candidates
+is considerably higher for attacks captured with content options
+compared to attacks not involving content options. This happens
+because for content attacks the number of option candidates depend on
+the unbounded size of the http payload whereas for attacks of
+different protocols the number of option candidates depend on the
+layout of the messages. Also observe that the number of plausible
+rules is much smaller for non-content attacks compared to content
+attacks.
 
@@
 the cases.
+
+%Results suggest that \tname\ either performs extremely well (in most of the cases) or poorly.
 
@@
     \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
-    \vspace{-2ex}        
+    \vspace{-3ex}        
     \caption{scripttag}
@@
 \begin{wrapfigure}[11]{r}{0.28\textwidth}  
-  \vspace{-1.5ex}
+  \vspace{-2ex}
   \centering
@@
     \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/inc/inc.pdf}
-    \vspace{-1ex}    
+    \vspace{-3ex}    
     \caption{inc}    
@@
     \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
-    \vspace{-1ex}    
+    \vspace{-3ex}    
     \caption{jsp}    
   \end{subfigure}%
+  \vspace{-1ex}
   \caption{Negative examples.}
+  \vspace{-1ex}
   \label{fig:impact-negative}
-  \vspace{-2ex}
 \end{wrapfigure}
@@
 in the method used, results, and support for content
-options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed ARGS, a system
+options. Kao~\etal~\cite{kao2015AutomaticNR} proposed ARGS, a system
 to synthesize http rules that take malicious traffic as input. ARGS

### 2020-06-02 paper/paper.tex
minor changesRQ1-RQ3

@@
 
-\subsection{Step 2: Rules Creation}
+\subsection{Step 2: Rules Generation}
 \label{sec:minimization}
@@
 
-RQ2 evaluates the importance of each proposed heuristic function. To
-answer that question, we evaluated the use of \tname{} without each of
-these functions with the goal of checking how much each of them
-contributes to the overall performance of the technique. Recall that
-the score a rule $r$ is given by the weighted average of the values
-given by each heuristic function, \ie{} the score of $r$ is obtained
-with the function $f(r)=\sum_{i}^{N} w_i*f_i(r)/N$, where $w_i$ is the
-weight of heuristic $i$, $f_i(r)$ is the score of heuristic $i$ on
-$r$, and $N$ is the number of heuristics considered---four, in our
-case. To assess the value of each heuristic function, we assigned its
-weight to zero---effectively, disabling the function---and compared
-the result with \tname\ using default weights, obtained with the
-optimization procedure described on
+RQ2 assesses the importance of each heuristic function. To that end,
+we compared \tname{}, with all functions enabled, against
+\tname\ without each one of those functions. Recall that the score a
+rule $r$ is given by the weighted average of the values given by each
+heuristic function, \ie{}, the score of $r$ is obtained with the
+function $f(r)=\sum_{i}^{N} w_i*f_i(r)/N$, where $w_i$ is the weight
+of heuristic $i$, $f_i(r)$ is the score of heuristic $i$ on $r$, and
+$N$ is the number of heuristics considered. To assess the value of
+each heuristic function, we assigned its weight to zero---effectively,
+disabling the function---and compared the result with \tname\ using
+default weights, obtained with the optimization procedure described on
 Section~\ref{sec:weight-optimization}.
@@
 ($\uparrow$) indicates that the performance of \tname{} on that
-configuration is worse when compared to \tname's default
-configuration. A down arrow ($\downarrow$), appearing highlighted in
-the table, indicates the contrary.
+configuration is worse when compared to ``base'' (as the ranking
+increased). A down arrow ($\downarrow$), appearing highlighted in the
+table, indicates the contrary.
 
@@
     Basebuilder LFI & 55  & 55 & 320$\uparrow$ & 112$\uparrow$ & 58$\uparrow$\\
-      Tomcat Snoop Access & 6 & 6 & 3$\cellcolor{lightgray}\downarrow$ & 6 & 2665$\uparrow$\\
+      Tomcat Snoop Access & 6 & 6 & 3$\cellcolor{lightgray}\downarrow$ & 6 & 2,665$\uparrow$\\
     \bottomrule
@@
   \footnotesize
-  \setlength{\tabcolsep}{5pt}
+  \setlength{\tabcolsep}{3pt}
   \renewcommand{\arraystretch}{-0.01}
@@
 the remaining attacks to evaluate soundness of the rule (at the Rules
-Creations step). It is expected that the number and variety of attacks
+Generation step). It is expected that the number and variety of attacks
 play a role in the performance of the technique.  We answered this
@@
 the position of the golden rule at the rank for an increasing number
-\begin{wrapfigure}[10]{r}{0.28\textwidth}  
+\begin{wrapfigure}[9]{r}{0.28\textwidth}  
   \vspace{-2ex}
@@
   \caption{\label{fig:impact-number-attacks}Positive examples.}
-  \vspace{-1ex}
+  \vspace{-2ex}
 \end{wrapfigure}
@@
 
-\begin{wrapfigure}[11]{r}{0.28\textwidth}  
-  \vspace{-2ex}
+\begin{wrapfigure}[8]{r}{0.28\textwidth}  
+  \vspace{-2.5ex}
   \centering
@@
   \caption{Negative examples.}
-  \vspace{-1ex}
+  \vspace{-2ex}
   \label{fig:impact-negative}

### 2020-06-02 paper/paper.tex
fix ref
pass IV.C5

@@
 weights.
+\Gui{We used Differential Evolution algorithm from SciPy to find the best weights for our set of attacks.
+The optmization for our weights consists of minimizing the function that returns the sum of every attack's golden rule ranking in our set given a combination of weights.
+\indent
+$\sum_{i=1}^{N}\mathit{R_i}$ ($\mathit{R_i}$, where $\mathit{R}$ is the ranking of a golden rule and $\mathit{N}$ is the number of attacks)
+The algorithm also receives as input the range which the weights can change (between 0 and 1) and the max number of generations, which we set to 10 (setting higher values could provide better results, but also would take more time).
+At the end, the optimization returned the weights [0.28058353, 0.05713404, 0.043285, 0.49022223]}
 
@@
 in the method used, results, and support for content
-options. Kao~\etal~\cite{kao2015AutomaticNR} proposed ARGS, a system
+options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed ARGS, a system
 to synthesize http rules that take malicious traffic as input. ARGS

### 2020-06-02 paper/paper.tex
minor changes

@@
 \begin{figure}[h!]
-  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
+  \lstinputlisting[columns=flexible, language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
   \vspace{-1ex}  
@@
 
-The rule creation procedure \tname{} uses iteratively discards options
-from rules, derived from the seed rule, using the negative traffic to
-guide the search for minimal rules.  Algorithm~\ref{algo1} shows the
-pseudocode for deriving these rules. The algorithm takes as input an
-over-constrained rule $sr$, denoting the seed rule produced at the
-Reverse Engineering step, and pcap files~\cite{pcap} to reconstruct
-the benign traffic ($\mathit{bTr}$) and the malicious traffic
-($\mathit{mTr}$). The algorithm produces on output a set of
-alternative rules to $sr$ that captures every instance of the attack
-(as per $\mathit{mTr}$), but does not capture any message from the
-benign traffic (as per $\mathit{bTr}$).
+\tname{} generates plausible rules while minimizing the seed
+rule. Algorithm~\ref{algo1} shows the pseudocode of that
+procedure. The algorithm takes as input an over-constrained rule $sr$,
+denoting the seed rule produced at the Reverse Engineering step, and
+pcap files~\cite{pcap} to reconstruct the benign traffic
+($\mathit{bTr}$) and the malicious traffic ($\mathit{mTr}$). The
+algorithm produces on output a set of plausible rules. Each of these
+rules should capture every instance of the attack (as per
+$\mathit{mTr}$) and not capture any message from the benign traffic
+(as per $\mathit{bTr}$).
 
@@
 candidate rule in $\mathcal R$. The algorithm mutates each of these
-rules $N$ times \rui{is N an input?
-How does one determine it in practice?} \Gui{Yes.
-Higher N values increases the chances of creating every possible rule for the attack (decreasing the negative effects of the non-determinism), at the same time that increases the computational cost.
-So the value can be increased in case the desired rule is not being created, or it can be reduced if the code is creating too many rules.
-In our code, we chose manually N=2, as it was already enough for showing the desired results.}, discarding one rule option each
-time. Line~\ref{copy} assigns a copy of the bitvector encoding
+rules $N$ times, discarding one rule option each
+time. Higher (resp., lower) values of $N$ produces more (resp.,
+less) rules. We found empirically that a value of 2 was sufficient to
+generate rules in all cases. Line~\ref{copy} assigns a copy of the bitvector encoding
 $\mathit{ro}$ to variable $r$. Function $\mathit{len}$ returns the
@@
 
+%\rui{is N an input?
+How does one determine it in practice?} \Gui{Yes.
+Higher N values increases the chances of creating every possible rule for the attack (decreasing the negative effects of the non-determinism), at the same time that increases the computational cost.
+So the value can be increased in case the desired rule is not being created, or it can be reduced if the code is creating too many rules.
+In our code, we chose manually N=2, as it was already enough for showing the desired results.}
+
 The do-while loop terminates when no rule in the current generation of
@@
 capture only the positive traffic. Every rule created during the
-search captures the attack associated with the input as every created
-rule contains a proper subset of the options from the original rule
+search captures the attack associated with the input as every rule
+created contains a proper subset of the options from the original rule
 and all those options are, by definition, satisfied by the traffic
@@
 function $\mathit{filter}$, only the rules in $\mathcal O$ that
-capture all the instances of the attack (\ie{}, variants) are reported
-on output.
-
-%% \begin{proposition}
-%%   The minimization algorithm may not produce minimal rules.
-%% \end{proposition}
-
-%% \begin{proof}
-%% The output of the minimization algorithm may contain rules that are
-%% not minimal given the non-determinism of the algorithm. In principle,
-%% it is possible that options that could be discarded are not selected
-%% in the loop from lines~\ref{start-for}--\ref{end-for}. However, we
-%% found empirically that a small value of $N$ suffices to discard those
-%% options from a rule.
-%% \end{proof}
-
-%% Public databases of negative traffic with
-%% \Fix{thousands} of messages exist  and can be used to
-%% increase the accuracy of the minimization.
-%\subsection{Limitations}
-%\Fix{...}
-
+capture all the instances of the attack are reported on output.
 \textbf{Implementation details.} We implemented function
@@
 
+%% \begin{proposition}
+%%   The minimization algorithm may not produce minimal rules.
+%% \end{proposition}
+
+%% \begin{proof}
+%% The output of the minimization algorithm may contain rules that are
+%% not minimal given the non-determinism of the algorithm. In principle,
+%% it is possible that options that could be discarded are not selected
+%% in the loop from lines~\ref{start-for}--\ref{end-for}. However, we
+%% found empirically that a small value of $N$ suffices to discard those
+%% options from a rule.
+%% \end{proof}
+
+%% Public databases of negative traffic with
+%% \Fix{thousands} of messages exist  and can be used to
+%% increase the accuracy of the minimization.
+%\subsection{Limitations}
+%\Fix{...}
+
 %% \textbf{Limitations.} It is worth
@@
 
-\rui{paragraph on complexity? I wonder whether that is a good ideia because
-this is already a long section.}
+%\rui{paragraph on complexity? I wonder whether that is a good ideia because
+%this is already a long section.}
 

### 2020-06-02 paper/paper.tex
minor changes
minor change to Reverse Engineering step

@@
 \begin{figure}[h!]
-  \lstinputlisting[columns=flexible, language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
+  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
   \vspace{-1ex}  
@@
 
-\begin{figure}[h!]
-  \lstinputlisting[language=C,numbers=none,keywords={content}]{adaptor-golden-rule.suricata}
-  \vspace{-1ex}  
-  \caption{Suricata rule for JBoss JMX attack.}
-  \vspace{-2ex}  
-  \label{fig:adaptor-golden-rule.suricata}
-\end{figure}
-
 For the field strategy, \tname{} parses the input messages looking for
@@
 \percRulesWithContent{} of these rules contain at least one content
-option. Figure~\ref{fig:adaptor-golden-rule.suricata} shows an example of a rule that
+option.
+
+\begin{figure}[h!]
+  \lstinputlisting[language=C,numbers=none,keywords={http,content}]{adaptor-golden-rule.suricata}
+  \vspace{-1ex}  
+  \caption{Suricata rule for JBoss JMX attack.}
+  \vspace{-2ex}  
+  \label{fig:adaptor-golden-rule.suricata}
+\end{figure}
+
+Figure~\ref{fig:adaptor-golden-rule.suricata} shows an example of a rule that
 captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the

### 2020-06-02 paper/paper.tex
minor change to Reverse Engineering step
saving space fig 3
merging

@@
 %\vspace{-1ex}
+
 \begin{figure}[h!]
-  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
+  % part removed...rev:4; metadata:created_at 2010_09_23, updated_at 2010_09_23;  
+  \lstinputlisting[language=C,numbers=none,keywords={icmp,dsize,itype}]{pingscan.suricata}
   \vspace{-1ex}  
@@
 bytes in the payload and the option \CodeIn{itype: 8} indicates that
-the packet to be captured are ICMP Echo requests.
+the packets to be captured are ICMP Echo requests.
 
@@
 this case, \tname{} sets the header to \CodeIn{icmp any any -> any
-  any}. It uses the most general flow description instructing the
-\nids\ to analyze the icmp traffic flowing from/to any address and
+  any}. It uses the most general flow description to command the
+\nids\ to monitor the icmp traffic flowing from/to any address and
 port.  The rationale is that the choice of these parameters are
-domain-specific and subject to change by the sys admins. Likewise,
+domain-specific and subject to change by sys admins. Likewise,
 \tname{} leaves the action of the corresponding rule constant, as
@@
 and \CodeIn{flow} are the most common. The options \CodeIn{dsize} and
-\CodeIn{itype}, the ones associated with the golden rule, appear as
-the 6th and 7th most popular, respectively. Restricted to the set of
+\CodeIn{itype}, used in the golden rule, appear as the 6th and 7th
+most popular, respectively. However, when restricted to the set of
 five options obtained at the first step, \CodeIn{dsize} and

### 2020-06-02 paper/paper.tex
pass in optimizatoin

@@
 \label{sec:weight-optimization}
-We optimized the weights of these heuristic functions as follows. We
-ran \tname{} using different combinations of weights on a sample of
-attacks and chose the combination that offered the best overall
-performance, which we computed with the sum of the rankings of the
-golden rule on each attack. The weight combination that led to the
-smaller sum of rankings was considered the best choice. We considered
-\emph{all} combinations of the following weights 0, 0.25, 0.5, 0.75,
-and 1. Given that there are four different functions and five possible
-weight values, we needed to run \tname{} on the sample of attacks for
-625 (=$5^4$) times. In the end of the process, the weight combination
-$\langle{}1,0.25,0.25,1\rangle$ was the best choice we found. Although
-results showed that this strategy results in good performance, we
-remain to evaluate other optimization strategies to find better
-weights.
-\Gui{We used Differential Evolution algorithm from SciPy to find the best weights for our set of attacks.
-The optmization for our weights consists of minimizing the function that returns the sum of every attack's golden rule ranking in our set given a combination of weights.
+A key strategy for obtaining the ideal ranking lie in finding optimal weights
+of these heuristic functions. 
+%We ran \tname{} using different combinations of weights on a sample of
+%attacks and chose the combination that offered the best overall
+%performance, which we computed with the sum of the rankings of the
+%golden rule on each attack. The weight combination that led to the
+%smaller sum of rankings was considered the best choice. We considered
+%\emph{all} combinations of the following weights 0, 0.25, 0.5, 0.75,
+%and 1. Given that there are four different functions and five possible
+%weight values, we needed to run \tname{} on the sample of attacks for
+%625 (=$5^4$) times. In the end of the process, the weight combination
+%$\langle{}1,0.25,0.25,1\rangle$ was the best choice we found. Although
+%results showed that this strategy results in good performance, we
+%remain to evaluate other optimization strategies to find better
+%weights.
+%
+In  order to compute the optimal weights, we used the Differential Evolution 
+algorithm from SciPy to find the optimal weights for our set of attacks.
+The optmization consists in minimizing the function that returns the sum of 
+every attack's golden rule ranking in our set given a combination of weights: 
 \indent
-$\sum_{i=1}^{N}\mathit{R_i}$ ($\mathit{R_i}$, where $\mathit{R}$ is the ranking of a golden rule and $\mathit{N}$ is the number of attacks)
-The algorithm also receives as input the range which the weights can change (between 0 and 1) and the max number of generations, which we set to 10 (setting higher values could provide better results, but also would take more time).
-At the end, the optimization returned the weights [0.28058353, 0.05713404, 0.043285, 0.49022223]}
-
-%% , such as evolutionary alternatives as
-%% AVM~\cite{10.1109/32.57624}.
-
-%\Gui{Fizemos um teste para encontrar melhores pesos para cada ataque.
-Calculamos o fitness de cada regra atribuindo diferentes pesos para cada função heurística.
-Os valores atribuídos foram 0, 0.25, 0.5, 0.75 e 1.
-Depois de calculado, ordenamos todas as listas criadas e pegamos a posição da Golden Rule.
-As melhores combinações de pesos (que conseguiram melhor posição da golden rule) estão na tabela 4, onde na coluna "Weights", há quatro números separados por vírgula, cada um deles representando o peso de uma função (na ordem que é representado no paper.
-O total de regras criadas é igual ao da Tabela 3.}
+$\sum_{i=1}^{N}\mathit{R_i}$ ($\mathit{R_i}$, where $\mathit{R}$ is the 
+ranking of a golden rule and $\mathit{N}$ is the number of attacks).
+
+The algorithm also takes as input the range of the weights (between 0 and 1) 
+and the max number of generations, which we set to 10 (higher values could 
+potentially better results, but also would take more time). The weights are 
+as follows $\langle{}0.280, 0.057, 0.043, 0.490\rangle$.
 

### 2020-06-02 paper/paper.tex
pass on abstract

@@
 
+% and examples of trusted rules
+
 \begin{abstract}
 Network Intrusion Detection Systems (\nids{}) are popular tools to
-defend against network attacks. These systems monitor the network
-traffic and flag suspicious network behavior. Rule-based \nids\ do
+defend local networks against attacks. These systems monitor the
+network traffic and flag suspicious behavior. Rule-based \nids\ do
 that by checking the network traffic against a set of rules, which
@@
 existing defenses. This paper proposes \tname{}, a novel approach to
-synthesize rules for \textit{rule-based} \nids. \tname\ leverages
-malicious (positive) and benign (negative) traffic 
-%\rui{positive and negative traffic are first introduced here, withuot being explaiend.} 
-and examples of trusted rules to create
-rules for new attacks. \tname{} is organized as a pipeline of three components that
+synthesize rules for rule-based \nids. \tname\ leverages malicious
+(positive) and benign (negative) traffic to create rules for new
+attacks. \tname{} is organized as a pipeline of three components to
 1)~create an over-specified seed rule, 2)~derive plausible rules from
-the seed, and 3)~rank plausible rules. Experimental results provide
-evidence that \tname\ is promising.
- \rui{the example is interesting, but this gives the 
-feeling that the work is too preliminary} 
-For example, it was
-able to synthesize the correct rule in all cases we analyzed and
-ranked the correct rule among the first ten rules of the reported
-ranking in \percTopFiveRanking\ of the cases.
+the seed, and 3)~rank plausible rules. We evaluated \tname\ against a
+set of 21 network attacks with varios
+characteristics. \tname\ performed very well, being capable of
+generating the correct rule among the top-3 and top-1 rules of the
+ranking, respectively, in \percTopThreeRankingPrefix\ and
+\percTopOneRankingPrefix\ of the cases.
 \end{abstract}

### 2020-06-02 paper/paper.tex
updated attack names

@@
   \renewcommand{\arraystretch}{0.35}
-  \begin{tabular}{lcccp{11cm}}
+  \begin{tabular}{lcccp{10.9cm}}
     \toprule
@@
     UDP Teardrop & \cite{udp-teardrop-source} & DoS & udp & \textbf{fragbits}:M; \textbf{id}:124; \\
-      2100384 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:8; \\
-      2100399 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:1; \textbf{itype}:3; \\
-      2100401 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:3; \\
-      2100402 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:3; \textbf{itype}:3; \\
-      2100408 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:0; \\
-      2100485 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:13; \textbf{itype}:3; \\
-      2100486 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:10; \textbf{itype}:3; \\
+      ICMP Info: Ping & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:8; \\
+      Dst/Host Unreachable & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:1; \textbf{itype}:3; \\
+      Dst/Net Unreachable & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:3; \\
+      Dst/Port Unreachable & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:3; \textbf{itype}:3; \\
+      ICMP Info: Echo Reply & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:0; \textbf{itype}:0; \\
+      ICMP Info: Code 13 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:13; \textbf{itype}:3; \\
+      ICMP Info: Code 10 & \cite{iscx} & ICMP Info & icmp & \textbf{icode}:10; \textbf{itype}:3; \\
     %%UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5; \\
@@
     UDP Teardrop & 1 & 4 & 8 & 4 \\
-    2100384 & 13 & 8 & 18 & 2 \\
-    2100399 & 132 & 6 & 3 & 1 \\
-    2100401 & 12 & 6 & 25 & 1 \\
-    2100402 & 3 & 6 & 12 & 1 \\
-    2100408 & 15 & 8 & 9 & 2 \\
-    2100485 & 4 & 6 & 3 & 1 \\
-    2100486 & 195 & 6 & 2 & 1 \\
+    ICMP Info: Ping & 13 & 8 & 18 & 2 \\
+    Dst/Host Unreachable & 132 & 6 & 3 & 1 \\
+    Dst/Net Unreachable & 12 & 6 & 25 & 1 \\
+    Dst/Port Unreachable & 3 & 6 & 12 & 1 \\
+    ICMP Info: Echo Reply & 15 & 8 & 9 & 2 \\
+    ICMP Info: Code 13 & 4 & 6 & 3 & 1 \\
+    ICMP Info: Code 10 & 195 & 6 & 2 & 1 \\
     %%UDP Flood & - & - & - & - \\
@@
     SYN Flood & 1 & 1 & 1 & 1 & 1 \\
-    2100384 & 2 & 18$\uparrow$& 2 & 2 & 2 \\
-    2100399 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
-    2100401 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
-    2100402 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
-    2100408 & 2 & 9$\uparrow$ & 2 & 2 & 2 \\
-    2100485 & 1 & 2$\uparrow$ & 1 & 1 & 1 \\
-    2100486 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
+    ICMP Info: Ping & 2 & 18$\uparrow$& 2 & 2 & 2 \\
+    Dst/Host Unreachable & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
+    Dst/Net Unreachable & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
+    Dst/Port Unreachable & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
+    ICMP Info: Echo Reply & 2 & 9$\uparrow$ & 2 & 2 & 2 \\
+    ICMP Info: Code 13 & 1 & 2$\uparrow$ & 1 & 1 & 1 \\
+    ICMP Info: Code 10 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
     %%UDP Flood & - & - & - & - & - \\

### 2020-06-02 paper/paper.tex
polishing

--- 2020-06-02 paper/paper.tex
@@
 RQ1 assesses \tname's performance. The metric we use is the position
-of the golden rule at the ranked list of plausible rules reported by
+of the golden rule in the ranked list of plausible rules reported by
 \tname\ (see Figure~\ref{fig:overview}). This metric is a proxy of
@@
 
-Table~\ref{table:results} shows results. Column ``Name'' shows the
-name of the attack, column ``\#~Vars.'' shows the number of variants
-of the attack, column ``\#~Option Cand. (step~1)'' shows the number of
-option candidates reported by \tname\ at step 1, column ``\#~Plausible
-Rules (step~2)'' shows the number of rules reported by \tname\ at step
-2, and column ``Rank (step~3)'' shows the position of the golden rule
+Table~\ref{table:results} shows the results, where ``Name'' is 
+name of the attack, ``\#~Vars.'' is the number of variants
+of the attack, ``\#~Option Cand. (step~1)'' is the number of
+option candidates reported by \tname\ at step 1, ``\#~Plausible
+Rules (step~2)'' is the number of rules reported by \tname\ at step
+2, and  ``Rank (step~3)'' is the position of the golden rule
 in the reported ranking. Observe that the number of option candidates

### 2020-06-02 paper/paper.tex
polishing abstract

@@
 set of 21 network attacks with varios
-characteristics. \tname\ performed very well, being capable of
-generating the correct rule among the top-3 and top-1 rules of the
-ranking, respectively, in \percTopThreeRankingPrefix\ and
-\percTopOneRankingPrefix\ of the cases.
+characteristics. \tname\ was capable of generating the correct rule 
+among the top-3 and top-1 rules of the ranking, respectively, in 
+\percTopThreeRankingPrefix\ and \percTopOneRankingPrefix\ of the cases.
 \end{abstract}
@@
 We evaluated \tname\ against \totalAttacks attacks of various kinds, such as 
-\Luc{This is already on section V, should it also be here?} Denial-of-Service (DoS)~\cite{denial-of-service},
+Denial-of-Service (DoS)~\cite{denial-of-service},
 Active Reconnaissance (AREC)~\cite{active-reconnaissance} and Privilege
@@
 example, \tname\ ranked the correct rule among the top ten rules in 
-\percTopFiveRanking\ of the cases. \rui{to discuss, I find 
-this a bit weak.}\Mar{agree. let us try to add more examples and data.}
+\percTopFiveRanking\ of the cases. \rui{should we quote the same numbers as in the abstract?}
 

### 2020-06-02 paper/paper.tex
polishing intro

@@
 
-\tname{} takes on input one or more examples \rui{We ought to explain how does one get hold on examples.
-How diffcult are they to obtain?} \Gui{Given a captured network traffic, an expert must identify which packets can be used to generate the rule.
-These packets are separated and given as input to Syrius.}of an attack and produces
-on output a ranked list of rules for that attack. It uses public data
+\tname{} takes on input one or more examples of an attack, identified by a domain expert, 
+and produces on output a ranked list of rules for that attack. It uses public data
 sets of benign messages~\cite{tcpreplay,stratosphere-normal} and

### 2020-06-02 paper/paper.tex
pass on intro

@@
 
-Open-source \nids{}, like Snort~\cite{snort} and
-Suricata~\cite{suricata}, provide rulesets for many existing
-attacks. Unfortunately, attackers actively look for new mechanisms to
-
-\begin{wrapfigure}[8]{r}{0.21\textwidth}
-    \vspace{-2ex}  
+\begin{wrapfigure}[7]{r}{0.21\textwidth}
+    \vspace{-4ex}  
     \centering
@@
 \end{wrapfigure}
+Open-source \nids{}, like Snort~\cite{snort} and
+Suricata~\cite{suricata}, provide rulesets for many existing
+attacks. Unfortunately, attackers actively look for new mechanisms to
 exploit systems through the
@@
 
-\tname{} takes on input one or more examples of an attack, identified by a domain expert, 
-and produces on output a ranked list of rules for that attack. It uses public data
-sets of benign messages~\cite{tcpreplay,stratosphere-normal} and
+%\rui{We ought to explain how does one get hold on examples.
+How diffcult are they to obtain?} \Gui{Given a captured network traffic, an expert must identify which packets can be used to generate the rule.
+These packets are separated and given as input to Syrius.}
+
+\tname{} takes on input one or more examples of an attack, identified by a domain expert, and produces
+on output a ranked list of rules for that attack. One assumption that
+we make is that system administrators are able to capture the traffic
+that perpetrated the attack. The capture can be obtained by other
+institutions, who shared that information, or the organization itself
+that suffered the attack and wants to avoid that
+repeated. \tname\ uses public data sets of benign
+messages~\cite{tcpreplay,stratosphere-normal} and
 rulesets~\cite{emerging-threats-open} to create those rules. The
@@
   %leave inconsistent data in the paper!}
-We evaluated \tname\ against \totalAttacks attacks of various kinds, such as 
-Denial-of-Service (DoS)~\cite{denial-of-service},
-Active Reconnaissance (AREC)~\cite{active-reconnaissance} and Privilege
-escalation (PESC)~\cite{privilege-escalation}. %\Fix{...}\rui{add few examples}. 
-The observed results are encouraging. For 
-example, \tname\ ranked the correct rule among the top ten rules in 
-\percTopFiveRanking\ of the cases. \rui{should we quote the same numbers as in the abstract?}
+
+We evaluated \tname\ against \totalAttacks attacks of various kinds
+(\eg{} Denial-of-Service (DoS)~\cite{denial-of-service}, Active
+Reconnaissance (AREC)~\cite{active-reconnaissance} and Privilege
+escalation (PESC)~\cite{privilege-escalation}) and from various
+sources. Preliminary results provide early evidence that \tname\ is
+promising. It reported the ground truth rule in \emph{all} cases and
+reported that rule within the top-1, top-3, top-5, and top-10 rules of
+the ranking, respectively, in \percTopOneRanking\,
+\percTopThreeRanking, \percTopFiveRanking, and \percTopTenRanking\ of
+the cases.
 

### 2020-06-02 paper/paper.tex
at --> in

@@
 most popular, respectively. However, when restricted to the set of
-five options obtained at the first step, \CodeIn{dsize} and
+five options obtained in the first step, \CodeIn{dsize} and
 \CodeIn{itype} are the most popular. At the end of this step, \tname{}
@@
 Unfortunately, for http attacks, the relevant data for the options is
-likely at the message payload; not the fields. By analyzing our data
+likely in the message payload; not the fields. By analyzing our data
 set of rules, we found that this occurs frequently in http attacks,
@@
 procedure. The algorithm takes as input an over-constrained rule $sr$,
-denoting the seed rule produced at the Reverse Engineering step, and
+denoting the seed rule produced in the Reverse Engineering step, and
 pcap files~\cite{pcap} to reconstruct the benign traffic
@@
 highest number of examples---idqaccess and scripttag---and measured
-the position of the golden rule at the rank for an increasing number
+the position of the golden rule in the ranking for an increasing number
 \begin{wrapfigure}[9]{r}{0.28\textwidth}  

### 2020-06-02 paper/paper.tex
polishing

@@
 
-Table~\ref{table:weights} shows results. Column ``Name'' shows the
-name of the attack and column ``base'' shows the result of
-\tname\ with all heuristic functions enabled, as in column ``Rank
-(step 3)'' from Table~\ref{table:results}. The remaining columns show
+Table~\ref{table:weights} shows the results, where ``Name'' is the
+name of the attack and ``base'' is the result of
+\tname\ with all heuristic functions enabled, as in ``Rank
+(step 3)'' in Table~\ref{table:results}. The remaining columns show
 results obtained by running the same experiment, as on
@@
 ($\uparrow$) indicates that the performance of \tname{} on that
-configuration is worse when compared to ``base'' (as the ranking
-increased). A down arrow ($\downarrow$), appearing highlighted in the
-table, indicates the contrary.
+configuration is worse when compared to ``base''. A down arrow ($\downarrow$), 
+highlighted in the table, indicates the contrary.
 

### 2020-06-02 paper/paper.tex
pass on related work

@@
 
-This section elaborates on the most similar work.
-
-%We discuss in the following techniques similar in goal to ours. Unfortunately
+This section elaborates on the most similar work to ours.
+
+%We discuss in the following techniques similar in goal to
+%ours. Unfortunately
+
+%% As detailed before,
+%% we could not access the tools implementing the techniques for
+%% comparison. Nevertheless, we found that t
 
 \noindent
-\textbf{Alternative methods for rule synthesis.} As detailed before,
-we could not access the tools implementing the techniques for
-comparison. Nevertheless, we found that there are important
+\textbf{Alternative methods for rule synthesis.} There are important
 differences between our work and prior work. Vollmer
@@
 negatives with respect to the benign and mallicious traffic
-considered. Guruprasad \etal{}\cite{guruprasad} also proposes an
-evolutionary framework, creating their own ICMP dataset as Vollmer,
-but also using DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx}
-datasets for non-ICMP attacks. Although Guruprasad supports more
-protocols than Vollmer, our approach still supports a larger set of
-protocols and the ability to ranks rules. Muthuregunathan
-\etal{}~\cite{muthuregunathan-etal2009} propose a clustering method to
-create rules. They characterize candidate solutions as a set of field
-values and use a clustering algorithm---hybrid k-means partitioned
-around menoids---to isolate malicious from benign
-traffic. Subsequently, they optimize the formed clusters using genetic
-algorithms and hill-climbing search. Ou approach differs from theirs
-in the method used, results, and support for content
-options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed ARGS, a system
-to synthesize http rules that take malicious traffic as input. ARGS
-looks for a common pattern across the different inputs to define
-rules. It does not take existing rules as input or benign
+considered. Guruprasad~\etal{}~\cite{guruprasad} also used an
+evolutionary approach to solve the problem. Although the solution of
+Guruprasad~\etal{} supports more protocols than Vollmer~\etal{}', our
+approach still supports a larger set of protocols and the ability to
+ranks rules. Muthuregunathan~\etal{}~\cite{muthuregunathan-etal2009}
+propose a clustering method to create rules. They characterize
+candidate solutions as a set of field values and use a clustering
+algorithm---hybrid k-means partitioned around menoids---to isolate
+malicious from benign traffic. Subsequently, they optimize the formed
+clusters using a genetic algorithm and hill-climbing search. Ou
+approach differs from theirs in the method used, results, and support
+for content options. Kao~\etal~\cite{Kao2015AutomaticNR} proposed
+ARGS, a system to synthesize http rules that take malicious traffic as
+input. ARGS looks for a common pattern across the different inputs to
+define rules. It does not take existing rules as input or benign
 traffic. Consequently, false positive can be produced and need to be

### 2020-06-02 paper/paper.tex
polishing
add scipy reference

@@
 In  order to compute the optimal weights, we used the Differential Evolution 
-algorithm from SciPy to find the optimal weights for our set of attacks.
+algorithm from SciPy\cite{diff-evo-scipy} to find the optimal weights for our set of attacks.
 The optmization consists in minimizing the function that returns the sum of 
@@
 \tname\ with all heuristic functions enabled, as in ``Rank
-(step 3)'' in Table~\ref{table:results}. The remaining columns show
-results obtained by running the same experiment, as on
-Section~\ref{sec:answer-rqone}, but suppressing one heuristic
+(step 3)'' in Table~\ref{table:results}. The remaining columns are the
+results when suppressing one heuristic
 function. The number corresponds to the order the heuristic function

### 2020-06-02 paper/paper.tex
polishing

--- 2020-06-02 paper/paper.tex
@@
 (LFI)~\cite{local-file-inclusion} and ICMP Traffic Alerts \cite{icmp-types}.
+
 \noindent\textbf{Malicious traffic.}~Recall that \tname\ takes
@@
 the SYN Flood attack (see Section~\ref{sec:dos}).
+
 \noindent\textbf{Benign traffic.}~Public data sets of benign traffic
@@
 traffic of a university network.
+
 \noindent\textbf{Ground Truth.}~To objectively evaluate \tname{}'s
@@
 
-Table~\ref{table:weights} shows the results, where ``Name'' is the
+The results are in Table~\ref{table:weights}, where ``Name'' is the
 name of the attack and ``base'' is the result of
 \tname\ with all heuristic functions enabled, as in ``Rank
-(step 3)'' in Table~\ref{table:results}. The remaining columns show
-results obtained by running the same experiment, as on
-Section~\ref{sec:answer-rqone}, but suppressing one heuristic
-function. The number corresponds to the order the heuristic function
-is introduced on Section~\ref{sec:ranking}. The symbol of an up arrow
-($\uparrow$) indicates that the performance of \tname{} on that
-configuration is worse when compared to ``base''. A down arrow ($\downarrow$), 
+(step 3)'' in Table~\ref{table:results}. The remaining columns are the
+results when suppressing one heuristic
+function. The number corresponds to the order of hte heuristic function
+as introduced on Section~\ref{sec:ranking}. The up arrow
+($\uparrow$) indicates that the performance is worse when compared to ``base'',
+whereas, a down arrow ($\downarrow$), 
 highlighted in the table, indicates the contrary.
@@
 %\subsubsection{Lack of comparison with other tools}
+
 \noindent\textbf{Lack of comparison with other tools}
@@
 only domain experts can identify that.
+
 \noindent\textbf{Running times.}~We ran all our tests using an
@@
 categorical and numerical options, the average running time was 6
-seconds per attack. For attacks with the content option, the average
-running time was 1,445 seconds (=24.1 minutes).  \textbf{Threats to
-  Validity.}~Considering external validity. Our results cannot be
-generalized beyond the subjects we selected. To mitigate that threat
-we selected a variety of attacks with different
-characteristics. Unfortunately, we could not access tools for related
-techniques, but there is strong evidence that \tname\ performs very
-well and is the only technique to support content options, which are
-very important in this domain. Considering internal validity, in
+seconds per attack. For attacks with the content option, it was 
+1,445 seconds (=24.1 minutes) on average.
+
+\noindent\textbf{Threats to Validity.}~Considering external validity, 
+our results cannot be generalized beyond the subjects we selected. To 
+mitigate that threat we selected a variety of attacks with different
+characteristics. Unfortunately, related techniques are unavailable, 
+but there is strong evidence that \tname\ performs very
+well and is the only technique to support content options, an important 
+aspect this domain. Considering internal validity, in
 principle, the implementation of \tname\ could contain errors. We
@@
 require detailed options, such as the HTTP PESC, RCE and LFI.
+
+\noindent
 \textbf{Rule repair.}~Hallahan~\etal{}~\cite{8102263} propose to
@@
 dataset. Their technique is complementary to \tname{}.
+
+\noindent
 \textbf{Anomaly-based intrusion detection.}

### 2020-06-02 paper/paper.tex
polishing

@@
 \noindent\textbf{Attacks.}~Table~\ref{table:attacks} shows the selection
-of attacks we analyzed.  Non-http attacks are listed at the top of the
-table and http attacks are listed at the bottom. Recall from
+of attacks considered in our analysis.  Non-http attacks are listed at 
+the top of the table and http attacks are listed at the bottom. Recall from
 Section~\ref{sec:reverse-engineering} that (1) http attacks are
 prevalent, (2) most http attacks use only content options, and (3)
-most non-http attacks only use field options. We selected
+most non-http attacks use field options only. We selected
 \numNonContentAttacks\ non-http attacks and \numContentAttacks\ http
 attacks. The higher number of http attacks reflects their
-popularity. Column ``Name'' shows the name of the attack, column
-``Source'' shows the source where we obtained data to reproduce the
-attack (see Section~\ref{subsec:malicious-traffic}), column ``Type''
-shows the kind of attack, column ``Protocol'' shows the name of the
-protocol explored in the attack, and column ``Ground Truth'' shows the
+popularity. ``Name'' is the name of the attack,
+``Source'' is the source where we obtained data to reproduce the
+attack (see Section~\ref{subsec:malicious-traffic}), ``Type''
+is the kind of attack, ``Protocol'' is the
+protocol explored in the attack, and ``Ground Truth'' is the
 rule options to capture the attack. The rationale for the selection

### 2020-06-03 paper/paper.tex
included keywords
fix fig 1
fix figs 6, 7 and 8
pass

@@
 
+\providecommand{\keywords}[1]
+{
+  \small	
+  \textbf{\textit{Keywords---}} #1
+}
+
 %% \title{Conference Paper Title*\\
@@
 
-%% \begin{IEEEkeywords}
+\keywords{Intrusion Detection, Suricata, Rule-based, Synthesis}
+
+% \begin{IEEEkeywords}
 %% component, formatting, style, styling, insert
@@
 
-\begin{wrapfigure}[7]{r}{0.21\textwidth}
+\begin{wrapfigure}[8]{r}{0.21\textwidth}
     \vspace{-4ex}  
@@
 Threats Open Ruleset''~\cite{emerging-threats-open}, a popular public
-data set of \suri\ rules. A closer observation on the changelogs of
+data set of Snort and \suri\ rules. A closer observation on the changelogs of
 this ruleset~\cite{emerging-threats-changelogs} shows that new rules
@@
 header, and \emph{initial} options expressed in the
-traffic. Considering the Ping Scan attack, \tname{} replays the
-attack---with NMAP, in this case---and monitors the traffic.  It
+traffic. Considering the Ping Scan attack, \tname{} 
 infers the header by checking the protocol of the input messages. In
@@
 problem. In this step, \tname{} searches for alternative rules that
-preserve the invariant to capture the positive (/malicious) traffic
-but not capture the negative benign traffic. It uses the rule obtained
+preserve the invariant to capture the positive (malicious) traffic
+but not capture the negative (benign) traffic. It uses the rule obtained
 in the previous step as seed to bootstrap the search for alternative
@@
 mentioned above. We limited the number of content options per rule to
-10 in the histogram for space. \MyComment{Rules without
+8 in the histogram for space. \MyComment{Rules without
   \CodeIn{content} options are an exception and half of the rules
@@
 rules $N$ times, discarding one rule option each
-time. Higher (resp., lower) values of $N$ produces more (resp.,
+time. Higher (respectively, lower) values of $N$ produces more (respectively,
 less) rules. We found empirically that a value of 2 was sufficient to
-generate rules in all cases. Line~\ref{copy} assigns a copy of the bitvector encoding
+generate the golden rule in all cases. Line~\ref{copy} assigns a copy of the bitvector encoding
 $\mathit{ro}$ to variable $r$. Function $\mathit{len}$ returns the
@@
 with a deployed server (\eg{}, web server) that \suri\ would monitor,
-our setup calls \suri\ directly through an API. \suri\ is invoked
+our setup calls \suri\ offiline by providing a \emph{pcap} file as input. \suri\ is invoked
 multiple times for creating rules.
@@
 compared to those containing unpopular
-rules. Figure~\ref{fig:distribution-options} shows the histogram of
+options. Figure~\ref{fig:distribution-options} shows the histogram of
 rule options for the ruleset we used~\cite{emerging-threats-open}. The
-x-axis shows the options used in rules (the 10 most common) and the
+x-axis shows the options used in rules (the 8 most common) and the
 y-axis shows the number of rules using the corresponding option. Note
@@
           %symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode, flags, tls.fingerprint},          
-          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype},          
+          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode},          
         ]
-        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155)};
+        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155) (icode,108)};
       \end{axis}
@@
 \caption{Distributions.}
-\vspace{-4ex}
+\vspace{-2ex}
 \end{figure*}
@@
 rule. Figure~\ref{fig:distribution-strings} shows the distribution of
-the ten most popular strings in our ruleset. To illustrate, consider a
-rule whose content options use the strings ``GET'' and ``ID''. The
+the seven most popular strings in our ruleset. To illustrate, consider a
+rule whose content options use the strings ``GET'' and ``id=''. The
 contribution of that rule for the computation of similarity score
 would be $0.601$ $(=(1$+$.203)$/$2)$. The relatively high weight
-reflects the fact that ``GET'' and ``ID'' are popular content strings.
-
-\begin{wrapfigure}[9]{r}{0.21\textwidth}
+reflects the fact that ``GET'' and ``id='' are popular content strings.
+
+\begin{wrapfigure}[10]{r}{0.21\textwidth}
   \vspace{-2ex}
@@
 through these two modifiers. The relatively high heuristic value
-reflects the fact that the string could be accessed by two popular
-content modifiers.
+reflects the fact that the string could be accessed by a popular
+content modifier.
 
@@
 the position of the golden rule in the ranking for an increasing number
-\begin{wrapfigure}[9]{r}{0.28\textwidth}  
+\begin{wrapfigure}[11]{r}{0.28\textwidth}  
   \vspace{-2ex}
@@
 
-\begin{wrapfigure}[8]{r}{0.28\textwidth}  
+\begin{wrapfigure}[10]{r}{0.28\textwidth}  
   \vspace{-2.5ex}

### 2020-06-03 paper/paper.tex
pass

@@
 prevalent, (2) most http attacks use only content options, and (3)
-most non-http attacks use field options only. We selected
-\numNonContentAttacks\ non-http attacks and \numContentAttacks\ http
-attacks. The higher number of http attacks reflects their
-popularity. ``Name'' is the name of the attack,
+most non-http attacks use field options only. % We selected
+%\numNonContentAttacks\ non-http attacks and \numContentAttacks\ http
+%attacks. The higher number of http attacks reflects their popularity.
+``Name'' is the name of the attack,
 ``Source'' is the source where we obtained data to reproduce the
@@
 exist and can be used to evaluate the rate of false positives of
-security tools. We used two popular public sources. The Bigflows.pcap
+security tools. We used two popular public sources. The bigFlows.pcap
 data set is part of the TcpReplay~\cite{tcpreplay} open source
@@
     SYN Flood & 4 & 2 & 2 & 1 \\
-    UDP Teardrop & 1 & 4 & 8 & 4 \\
+    UDP Teardrop & 1 & 4 & 8 & 3 \\
     ICMP Info: Ping & 13 & 8 & 18 & 2 \\
@@
     SYN Flood & 1 & 1 & 1 & 1 & 1 \\
+      UDP Teardrop & 3 & 5$\uparrow$ & 2 \cellcolor{lightgray}$\downarrow$ & 3 & 3 \\
     ICMP Info: Ping & 2 & 18$\uparrow$& 2 & 2 & 2 \\
@@
     ICMP Info: Code 10 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
-    %%UDP Flood & - & - & - & - & - \\
     \midrule

### 2020-06-03 paper/paper.tex
macro and typo fix

@@
 results when suppressing one heuristic
-function. The number corresponds to the order of hte heuristic function
+function. The number corresponds to the order of the heuristic function
 as introduced on Section~\ref{sec:ranking}. The up arrow
@@
 is important to recall that we did not report precision and recall as
-it is 100\% always. \tname captures, by construction, all positive
+it is 100\% always. \tname\ captures, by construction, all positive
 traffic and does not capture any negative traffic. We used a stronger

### 2020-06-03 paper/paper.tex
pass

@@
 promising. It reported the ground truth rule in \emph{all} cases and
-reported that rule within the top-1, top-3, top-5, and top-10 rules of
-the ranking, respectively, in \percTopOneRanking\,
-\percTopThreeRanking, \percTopFiveRanking, and \percTopTenRanking\ of
+reported that rule within the top-1, top-3 and top-10 rules of
+the ranking, respectively, in \percTopOneRanking\ ,
+\percTopThreeRanking\ and \percTopTenRanking\ of
 the cases.
@@
 cases. It reported the ground truth rule in \emph{all} cases and
-reported that rule within the top-1, top-3, top-5, and top-10 rules of
-the ranking, respectively, in \percTopOneRanking\,
-\percTopThreeRanking, \percTopFiveRanking, and \percTopTenRanking\ of
+reported that rule within the top-1, top-3 and top-10 rules of
+the ranking, respectively, in \percTopOneRanking\ ,
+\percTopThreeRanking\ and \percTopTenRanking\ of
 the cases.
@@
 \tname\ is promising. It was able to rank the correct rule among the
-first ten rules of the reported ranking in \percTopFiveRanking\ of
+first ten rules of the reported ranking in \percTopTenRanking\ of
 the cases we analyzed, for example. The data set containing our

### 2020-06-03 paper/paper.tex
rephrase optimization senttence

@@
 The algorithm also takes as input the range of the weights (between 0 and 1) 
-and the max number of generations, which we set to 10 (higher values could 
-potentially better results, but also would take more time). The weights are 
+and the max number of generations, which we set to 10 (confirmation whether 
+this number suffices remains to future work). The weights are 
 as follows $\langle{}0.280, 0.057, 0.043, 0.490\rangle$.

### 2020-06-03 paper/paper.tex
use template set of keywords

@@
 
-\providecommand{\keywords}[1]
-{
-  \small	
-  \textbf{\textit{Keywords---}} #1
-}
-
 %% \title{Conference Paper Title*\\
@@
 
-\keywords{Intrusion Detection, Suricata, Rule-based, Synthesis}
-
-% \begin{IEEEkeywords}
-%% component, formatting, style, styling, insert
-%% \end{IEEEkeywords}
+\begin{IEEEkeywords}
+Intrusion Detection, Suricata, Rule Synthesis.
+\end{IEEEkeywords}
 
@@
 Network Intrusion Detection Systems (\nids{}) are software systems
-that monitor the network traffic for malicious behavior and act
-accordingly by blocking messages or alerting humans about suspicious
+that monitor the network traffic for malicious behavior and block 
+messages or alerting humans about suspicious
 events~\cite{Mitchell:2014:SID:2597757.2542049}. \nids{} are typically
 placed behind a firewall, vetting the traffic that the firewall did
-not block. Various open-source (\eg{}, Snort~\cite{snort} and
+not block. Various open-source (\eg{}, Snort~\cite{snort},
 Suricata~\cite{suricata}) and commercial (\eg{},
-SolarWinds~\cite{solarwinds} and IBM QRadar~\cite{qradar})
+SolarWinds~\cite{solarwinds}, IBM QRadar~\cite{qradar})
 \nids\ implementations exist.
@@
 benign traffic to derive a set of plausible rules that isolate the
-benign traffic from the malicious traffic. Finally, it uses existing
-rules to learn heuristics to rank the plausible rules.
+benign traffic from the malicious traffic. Finally, it uses a set of 
+heuristics to rank the plausible rules.
 
@@
 uses benign traffic in the second step to create a set of plausible
-rules from the seed rule. Finally, in the third step, \tname{} uses
-existing rules to learn heuristics to rank plausible rules.
+rules from the seed rule. Finally, in the third step, \tname{} uses a 
+set of heuristics to rank plausible rules.
 

### 2020-06-03 paper/paper.tex
more polishing

@@
 uses benign traffic in the second step to create a set of plausible
-rules from the seed rule. Finally, in the third step, \tname{} uses a 
-set of heuristics to rank plausible rules.
+rules from the seed rule. Finally, in the third step, \tname{} ranks the 
+plausible rules.
 
@@
 %\vspace{0.5ex}
-\noindent\textbf{Attacks.}~Table~\ref{table:attacks} shows the selection
-of attacks considered in our analysis.  Non-http attacks are listed at 
-the top of the table and http attacks are listed at the bottom. Recall from
+\noindent\textbf{Attacks.}~Table~\ref{table:attacks} outlines the
+attacks considered in our analysis, where non-http attacks are at 
+the top and http attacks are at the bottom. Recall from
 Section~\ref{sec:reverse-engineering} that (1) http attacks are
-prevalent, (2) most http attacks use only content options, and (3)
-most non-http attacks use field options only. % We selected
-%\numNonContentAttacks\ non-http attacks and \numContentAttacks\ http
-%attacks. The higher number of http attacks reflects their popularity.
+prevalent, (2) most http attacks use content options only, and (3)
+most non-http attacks use field options only. 
 ``Name'' is the name of the attack,
-``Source'' is the source where we obtained data to reproduce the
+``Source'' is the source of the data to reproduce the
 attack (see Section~\ref{subsec:malicious-traffic}), ``Type''
@@
 rule options to capture the attack. The rationale for the selection
-were (1) to cover various kinds of attacks and (2) to cover attacks of
-various protocols. Considering (1), the selection includes attacks of
+are attack and protocol diversity. Considering (1), 
+the selection includes attacks of
 the following types: Denial-of-Service (DoS)~\cite{denial-of-service},
@@
 also used data sets from the Stratosphere IPS
-Project~\cite{stratosphere-normal} including, for example, the regular
+Project~\cite{stratosphere-normal} including, e.g., the regular
 traffic of a university network.
@@
 \noindent\textbf{Ground Truth.}~To objectively evaluate \tname{}'s
-performance we compared the rules it generates with \emph{trusted
-  rules}. We used the ``Emerging Threats Open
-Ruleset''~\cite{emerging-threats-open} for that. After finding a
-matching pair of attack and corresponding rule using the procedure
+performance, we compared the rules it generates with \emph{trusted
+  rules} using the ``Emerging Threats Open
+Ruleset''~\cite{emerging-threats-open}. After finding a
+matching pair of attack and corresponding rule as
 described in Section~\ref{subsec:malicious-traffic}, we reproduced the
@@
 \section{Related Work}
-
-%\Mar{mention something on ethical hacking.. ethical-hacking}
-
-This section elaborates on the most similar work to ours.
-
-%We discuss in the following techniques similar in goal to
-%ours. Unfortunately
-
-%% As detailed before,
-%% we could not access the tools implementing the techniques for
-%% comparison. Nevertheless, we found that t
 

### 2020-06-03 paper/paper.tex
minor changes

@@
 messages. Section~\ref{sec:rules-and-packets} shows the correspondence
-between packet fields and some of the rule options from \suri\ and
+between packet fields and some of the rule options from \suri\ while
 Section~\ref{sec:active-recon} shows a concrete example where \tname{}
-uses this strategy. Although we built the option-field mapping
-manually, from the \suri\ documentation, this is a one-time activity.
-Unfortunately, for http attacks, the relevant data for the options is
-likely in the message payload; not the fields. By analyzing our data
-set of rules, we found that this occurs frequently in http attacks,
-which are prevalent. For example, \percHttp\ of the \numrulessuri{}
-rules from the ruleset we analyzed~\cite{emerging-threats-open}
-prescribe defenses against http attacks. In addition,
-\percRulesWithContent{} of these rules contain at least one content
-option.
+uses this strategy to create the seed rule. Important to note that,
+although we built the option-field mapping manually (from the
+\suri\ documentation), this is a one-time activity.  Unfortunately,
+this strategy is insufficient for http attacks, which are
+prevalent. For example, \percHttp\ of the \numrulessuri{} rules from
+the ruleset we analyzed~\cite{emerging-threats-open} prescribe
+defenses against http attacks. By analyzing our data set of rules, we
+found that the relevant data for the options appears in the message
+payload, not the fields. For example, \percRulesWithContent{} of the
+rules for http attacks contain at least one content option.
 
@@
 
-Figure~\ref{fig:adaptor-golden-rule.suricata} shows an example of a rule that
-captures traffic containing the string \CodeIn{"/HtmlAdaptor"} in the
-payload.  Figure~\ref{fig:distribution-contents} shows the
-distribution of number of content options per rule for the data set
-mentioned above. We limited the number of content options per rule to
-8 in the histogram for space. \MyComment{Rules without
-  \CodeIn{content} options are an exception and half of the rules
-  contain at least two \CodeIn{content} options.}  To sum up, handling
-content options for http is very important.
+Figure~\ref{fig:adaptor-golden-rule.suricata} shows an example of an
+http rule that captures traffic containing the string
+\CodeIn{"/HtmlAdaptor"} in the payload.
+Figure~\ref{fig:distribution-contents} shows the distribution of
+number of content options per rule for the data set mentioned
+above. We limited the number of content options per rule to 7 in the
+histogram for space. \MyComment{Rules without \CodeIn{content} options
+  are an exception and half of the rules contain at least two
+  \CodeIn{content} options.}  To sum up, content options appear
+frequently in http rules and rules may contain several of these
+options. Handling content options is therefore very important.
 
@@
 with a deployed server (\eg{}, web server) that \suri\ would monitor,
-our setup calls \suri\ offiline by providing a \emph{pcap} file as input. \suri\ is invoked
-multiple times for creating rules.
+we calls \suri\ directly, through a public API, by providing a
+\emph{pcap} files on input. \suri\ is invoked multiple times for
+creating rules.
 
@@
   \setlength{\tabcolsep}{4.5pt}
-  \renewcommand{\arraystretch}{0.35}
+  \renewcommand{\arraystretch}{0.4}
   \begin{tabular}{lcccp{10.9cm}}
@@
   \setlength{\tabcolsep}{2pt}
-  \renewcommand{\arraystretch}{0.5}
+  \renewcommand{\arraystretch}{0.7}
   \caption{\label{table:results}Results.}
@@
   %\setlength{\tabcolsep}{5pt}
-  \renewcommand{\arraystretch}{-0.01}
+  \renewcommand{\arraystretch}{0.5}
   \caption{\label{table:weights}Contribution of heuristic functions.}
@@
 the position of the golden rule in the ranking for an increasing number
-\begin{wrapfigure}[11]{r}{0.28\textwidth}  
+\begin{wrapfigure}[9]{r}{0.28\textwidth}  
   \vspace{-2ex}
@@
   \caption{\label{fig:impact-number-attacks}Positive examples.}
-  \vspace{-2ex}
+  \vspace{-4ex}
 \end{wrapfigure}
@@
 
-\begin{wrapfigure}[10]{r}{0.28\textwidth}  
+\begin{wrapfigure}[8]{r}{0.28\textwidth}  
   \vspace{-2.5ex}

### 2020-06-03 paper/paper.tex
minor changes
including argmin

@@
 %
+% Math stuff
+%
+\newcommand{\argmin}{\mathop{\mathrm{{arg\,min}}}}
+
+%
 % Add comments in the text
@@
 
-\begin{wrapfigure}[8]{r}{0.21\textwidth}
+\begin{wrapfigure}[7]{r}{0.21\textwidth}
     \vspace{-4ex}  
@@
 
-\subsubsection{\label{sec:popularity-options}Popularity of options}
+\subsubsection{\label{sec:popularity-options}Popularity of options ($popt$)}
 
@@
 \indent
-$\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}/N$.
+$popt = \sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}/N$.
 
@@
 
-\subsubsection{Number of content options} We observed
+\subsubsection{Number of content options ($ncont$)} 
+We observed
 that \percContentOptions\ of the rules from the ruleset we analyzed
@@
 
-\subsubsection{Popularity of content strings} 
+\subsubsection{Popularity of content strings ($pcont$)} 
 This heuristic leverages content rarity to discriminate rules. It
@@
 
-\subsubsection{\label{sec:popularity-content-modifiers}Popularity of http content modifiers} As previously
-mentioned, \percRulesWithContent\ of \suri\ rules involve content
+\subsubsection{\label{sec:popularity-content-modifiers}Popularity of http content modifiers ($phttp$)} 
+As previously mentioned, \percRulesWithContent\ of \suri\ rules involve content
 options. Furthermore, nearly half of the rules that use content
@@
 A key strategy for obtaining the ideal ranking lie in finding optimal weights
-of these heuristic functions. 
-%We ran \tname{} using different combinations of weights on a sample of
-%attacks and chose the combination that offered the best overall
-%performance, which we computed with the sum of the rankings of the
-%golden rule on each attack. The weight combination that led to the
-%smaller sum of rankings was considered the best choice. We considered
-%\emph{all} combinations of the following weights 0, 0.25, 0.5, 0.75,
-%and 1. Given that there are four different functions and five possible
-%weight values, we needed to run \tname{} on the sample of attacks for
-%625 (=$5^4$) times. In the end of the process, the weight combination
-%$\langle{}1,0.25,0.25,1\rangle$ was the best choice we found. Although
-%results showed that this strategy results in good performance, we
-%remain to evaluate other optimization strategies to find better
-%weights.
+for the heuristic functions. The ranking of a golden rule, for a set of $N$ 
+attacks, is given by $R(W) = \sum_{i=1}^{N}\mathit{R_i}$, where 
+$R_i = (w_1 \cdot popt) \cdot (w_2 \cdot ncont) \cdot (w_3 \cdot pcont) \cdot (w_4 \cdot phttp)$
+and $W = \langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered for each heuristic.
 %
 In  order to compute the optimal weights, we used the Differential Evolution 
-algorithm from SciPy\cite{diff-evo-scipy} to find the optimal weights for our set of attacks.
-The optmization consists in minimizing the function that returns the sum of 
-every attack's golden rule ranking in our set given a combination of weights: 
-\indent
-$\sum_{i=1}^{N}\mathit{R_i}$ ($\mathit{R_i}$, where $\mathit{R}$ is the 
-ranking of a golden rule and $\mathit{N}$ is the number of attacks).
-
+algorithm from SciPy\cite{diff-evo-scipy}.
+The optmization consists in minimizing the function of every attack's golden rule 
+ranking in our set given a combination of weights. Formally, $W  = \displaystyle \argmin_W R$.
+%
 The algorithm also takes as input the range of the weights (between 0 and 1) 
@@
 this number suffices remains to future work). The weights are 
-as follows $\langle{}0.280, 0.057, 0.043, 0.490\rangle$.
+as follows $\langle{}0.280, 0.057, 0.043, 0.490\rangle{}$.
 

### 2020-06-03 paper/paper.tex
pass IV.C5

@@
 A key strategy for obtaining the ideal ranking lie in finding optimal weights
-for the heuristic functions. The ranking of a golden rule, for a set of $N$ 
-attacks, is given by $R(W) = \sum_{i=1}^{N}\mathit{R_i}$, where 
-$R_i = (w_1 \cdot popt) \cdot (w_2 \cdot ncont) \cdot (w_3 \cdot pcont) \cdot (w_4 \cdot phttp)$
-and $W = \langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered for each heuristic.
+for the heuristic functions. The sum of $R_i$ rankings of the golden rules, for a set of $N$ 
+attacks, is given by $R(W) = \sum_{i=1}^{N}\mathit{R_i}$, where $W = \langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered for each heuristic.
 %

### 2020-06-03 paper/paper.tex
pass IV.C5

@@
 A key strategy for obtaining the ideal ranking lie in finding optimal weights
-for the heuristic functions. The sum of $R_i$ rankings of the golden rules, for a set of $N$ 
-attacks, is given by $R(W) = \sum_{i=1}^{N}\mathit{R_i}$, where $W = \langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered for each heuristic.
+for the heuristic functions. The sum of rankings of the golden rules, for a set of $N$ 
+attacks, is given by $R(W) = \sum_{i=1}^{N}\mathit{R_i}$, where $R_i$ is the ranking based on the score provided by $(w_1 + popt) \cdot (w_2 + ncont) \cdot (w_3 + pcont) \cdot (w_4 + phttp)$ and $W = \langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered for each heuristic.
 %

### 2020-06-04 paper/paper.tex
changing title
polishing

@@
 
-\title{Rule Synthesis for Rule-based Intrusion Detectors}
+\title{Rule Synthesis for Network Intrusion Detectors}
 
@@
 \label{sec:weight-optimization}
-A key strategy for obtaining the ideal ranking lie in finding optimal weights
-for the heuristic functions. The sum of rankings of the golden rules, for a set of $N$ 
-attacks, is given by $R(W) = \sum_{i=1}^{N}\mathit{R_i}$, where $R_i$ is the ranking based on the score provided by $(w_1 + popt) \cdot (w_2 + ncont) \cdot (w_3 + pcont) \cdot (w_4 + phttp)$ and $W = \langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered for each heuristic.
+Obtaining the best possible ranking for the golden rule boils down to 
+finding optimal weights for the heuristic functions. The sum of rankings 
+of the golden rules, for a set of $N$ attacks, is given by 
+$R(W) = \sum_{i=1}^{N}\mathit{R_i}$, where $R_i$ is the ranking based on 
+the score provided by 
+$(w_1 + popt) \cdot (w_2 + ncont) \cdot (w_3 + pcont) \cdot (w_4 + phttp)$ 
+and $W = \langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered 
+for each heuristic.
 %
-In  order to compute the optimal weights, we used the Differential Evolution 
-algorithm from SciPy\cite{diff-evo-scipy}.
-The optmization consists in minimizing the function of every attack's golden rule 
-ranking in our set given a combination of weights. Formally, $W  = \displaystyle \argmin_W R$.
+We resorted to the Differential Evolution algorithm from SciPy\cite{diff-evo-scipy}
+to compute the optimal weights. The optmization consists in minimizing the function 
+of every attack's golden rule ranking in our set given a combination of weights. 
+Formally, $W  = \displaystyle \argmin_W R(W)$.
 %

### 2020-06-04 paper/paper.tex
fix score formula IV.C5

@@
 the score provided by 
-$(w_1 + popt) \cdot (w_2 + ncont) \cdot (w_3 + pcont) \cdot (w_4 + phttp)$ 
+$[(w_1 \cdot popt) + (w_2 \cdot ncont) + (w_3 \cdot pcont) + (w_4 \cdot phttp)]/4$ 
 and $W = \langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered 

### 2020-06-04 paper/paper.tex
table 3 and 4

@@
     scripttag & 23 & 22 & 2,221 & 1 \\
-    WeBid LFI & 1 & 23 & 35,444 & 29\\    
+    WeBid LFI & 1 & 23 & 35,444 & 21\\    
     %%Wordpress Login & 8 & 59 & - & - \\
@@
     /system32/ in URI & 4 & 20 & 1,881 & 1 \\
-    Basebuilder LFI & 1 & 22 & 60,222 & 55 \\
-    Tomcat Snoop Access & 1 & 24 & 58,122 & 6 \\
+    Basebuilder LFI & 1 & 22 & 60,222 & 51 \\
+    Tomcat Snoop Access & 1 & 24 & 58,122 & 5 \\
     \bottomrule
@@
     %%Apache Struts 2 & - & - & - & - & - & -\\
-    ColdFusion & 2 & 2 & 2 & 7$\uparrow$ & 5$\uparrow$ \\
-    iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
+    ColdFusion & 2 & 2 & 2 & 7$\uparrow$ & 6$\uparrow$ \\
+    iisadmin access & 1  & 1 & 1 & 1 & 13$\uparrow$\\        
     JBoss JMX & 36 & 36 & 14\cellcolor{lightgray}$\downarrow$ & 34\cellcolor{lightgray}$\downarrow$ & 269$\uparrow$ \\
-    scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
-    WeBid LFI & 29 & 29 & 72$\uparrow$ & 39$\uparrow$ & 41$\uparrow$ \\    
+    scripttag & 1 & 1 & 3$\uparrow$ & 1 & 16$\uparrow$ \\
+    WeBid LFI & 21 & 21 & 74$\uparrow$ & 39$\uparrow$ & 41$\uparrow$ \\    
     %%Wordpress Login & 8 & - & - & - & - \\
-    .htaccess access & 1  & 1 & 1 & 1 & 11$\uparrow$\\
-    .idq access & 2  & 2 & 2 & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ \\
-    /system32/ in URI & 1  & 1 & 4$\uparrow$ & 1 & 15$\uparrow$\\
-    Basebuilder LFI & 55  & 55 & 320$\uparrow$ & 112$\uparrow$ & 58$\uparrow$\\
-      Tomcat Snoop Access & 6 & 6 & 3$\cellcolor{lightgray}\downarrow$ & 6 & 2,665$\uparrow$\\
+    .htaccess access & 1  & 1 & 1 & 1 & 10$\uparrow$\\
+    .idq access & 2  & 2 & 2 & 1\cellcolor{lightgray}$\downarrow$ & 14$\uparrow$ \\
+    /system32/ in URI & 1  & 1 & 4$\uparrow$ & 1 & 14$\uparrow$\\
+    Basebuilder LFI & 51  & 51 & 370$\uparrow$ & 112$\uparrow$ & 58$\uparrow$\\
+      Tomcat Snoop Access & 5 & 5 & 2$\cellcolor{lightgray}\downarrow$ & 6 & 646$\uparrow$\\
     \bottomrule

### 2020-06-05 paper/paper.tex
fix captions fig 8 and 9
fig 7 and 8

@@
 
-\begin{wrapfigure}[10]{r}{0.21\textwidth}
+\begin{wrapfigure}[9]{r}{0.21\textwidth}
   \vspace{-2ex}
@@
 the position of the golden rule in the ranking for an increasing number
-\begin{wrapfigure}[9]{r}{0.28\textwidth}  
+\begin{wrapfigure}[8]{r}{0.28\textwidth}  
   \vspace{-2ex}
@@
     \centering
-    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
+    \includegraphics[scale=0.30,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
     \vspace{-3ex}
-    \caption{idqaccess}
+    \caption{.idq access}
   \end{subfigure}%
@@
     \centering
-    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
+    \includegraphics[scale=0.30,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
     \vspace{-3ex}        
@@
     \vspace{-3ex}    
-    \caption{inc}    
+    \caption{Basebuilder LFI}    
   \end{subfigure}%
@@
     \vspace{-3ex}    
-    \caption{jsp}    
+    \caption{Tomcat S. A.}    
   \end{subfigure}%

### 2020-06-05 paper/R/idqaccess/subject.r
fig 8

@@ -1,6 +1,6 @@
 #!/usr/bin/env Rscript
 
-pdf("idqaccess.pdf", width = 5, height = 2.9)
+pdf("idqaccess.pdf", width = 6.5, height = 2.9)
 
 df1 <- read.table("data1.txt", header = FALSE)
 df2 <- read.table("data2.txt", header = FALSE)

### 2020-06-05 paper/R/scripttag/subject.r

@@ -1,6 +1,6 @@
 #!/usr/bin/env Rscript
 
-pdf("scripttag.pdf", width = 5, height = 2.9)
+pdf("scripttag.pdf", width = 6.5, height = 2.9)
 
 df1 <- read.table("data1.txt", header = FALSE)
 df2 <- read.table("data2.txt", header = FALSE)

### 2020-06-05 paper/paper.tex
weights
pass IV.C5

@@
 The algorithm also takes as input the range of the weights (between 0 and 1) 
-and the max number of generations, which we set to 10 (confirmation whether 
-this number suffices remains to future work). The weights are 
-as follows $\langle{}0.280, 0.057, 0.043, 0.490\rangle{}$.
+and the max number of generations, which we set to the default value of 1000, but the algorithm stopped at the 45th generation as there was no further improvement. The weights are 
+as follows $\langle{}0.449, 0.103, 0.053, 0.796\rangle{}$.
 

### 2020-06-05 paper/paper.tex
minor changes

@@
 the score provided by 
-$[(w_1 \cdot popt) + (w_2 \cdot ncont) + (w_3 \cdot pcont) + (w_4 \cdot phttp)]/4$ 
-and $W = \langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered 
+$[(w_1 \cdot popt)$+$(w_2 \cdot ncont)$+$(w_3 \cdot pcont)$+$(w_4 \cdot phttp)]/4$ 
+and $W$=$\langle{}w_1, w_2, w_3, w_4\rangle{}$ are the weights considered 
 for each heuristic.
 %
-We resorted to the Differential Evolution algorithm from SciPy\cite{diff-evo-scipy}
-to compute the optimal weights. The optmization consists in minimizing the function 
-of every attack's golden rule ranking in our set given a combination of weights. 
-Formally, $W  = \displaystyle \argmin_W R(W)$.
+We resorted to the Differential Evolution algorithm from
+SciPy\cite{diff-evo-scipy} to compute the optimal weights. The
+optimization consists in minimizing the function of every attack's
+golden rule ranking in our set given a combination of weights.
+Formally, $W$=$\displaystyle \argmin_W R(W)$.
 %

### 2020-06-06 paper/paper.tex
minor changes
spell check

@@
 
+%\title{\tname{} for NIDS Rule Synthesis}
 \title{Rule Synthesis for Network Intrusion Detectors}
@@
 the seed, and 3)~rank plausible rules. We evaluated \tname\ against a
-set of 21 network attacks with varios
+set of 21 network attacks with various
 characteristics. \tname\ was capable of generating the correct rule 
@@
 The procedure we propose in this step builds on the observation that
-all options from the rule produced in the previous step are satisifed
+all options from the rule produced in the previous step are satisfied
 by the traffic produced by the attack. Consequently, any subset of
@@
 with a deployed server (\eg{}, web server) that \suri\ would monitor,
-we calls \suri\ directly, through a public API, by providing a
+we call \suri\ directly, through a public API, by providing a
 \emph{pcap} files on input. \suri\ is invoked multiple times for
@@
   performance of \tname. The function that builds on popularity of
-  content modifiers showed highest impact.}
+  content modifiers showed the highest impact.}
 
@@
 possible for two central reasons. First, we did not have access to the
-tools. We tried, unsuccssefully, making contact with authors to obtain
+tools. We tried, unsuccessfully, making contact with authors to obtain
 the tools. Second, the tools output different objects. For that
@@
 genetic search. The initial population is a set of candidate solutions
-that randomly combine options, modelling the packet fields. They
+that randomly combine options, modeling the packet fields. They
 showed that they were able to generate rules without false negatives
@@
 plausible options, which contain \emph{no} false positives or
-negatives with respect to the benign and mallicious traffic
+negatives with respect to the benign and malicious traffic
 considered. Guruprasad~\etal{}~\cite{guruprasad} also used an

### 2020-06-06 paper/paper.tex
minor changes
merging

@@
 %\title{\tname{} for NIDS Rule Synthesis}
-\title{Rule Synthesis for Network Intrusion Detectors}
+\title{\tname{}: Intrusion Detection Rule Synthesis}  
+%\title{Rule Synthesis for Network Intrusion Detectors}
 
@@
 
-\textbf{Correctness argument.}~The algorithm reports rules that
+\noindent\textbf{Correctness argument.}~The algorithm reports rules that
 capture only the positive traffic. Every rule created during the
@@
 capture all the instances of the attack are reported on output.
-\textbf{Implementation details.} We implemented function
+
+\noindent\textbf{Implementation details.} We implemented function
 \emph{isPlausible} by 1)~spawning Suricata to monitor a specified
@@
 
+% There are important differences between our work and prior work.
+
 \noindent
-\textbf{Alternative methods for rule synthesis.} There are important
-differences between our work and prior work. Vollmer
+\textbf{Alternative methods for rule synthesis.}~Vollmer
 \etal{}~\cite{vollmer-etal-cics2011} proposed the use of search to
@@
 genetic search. The initial population is a set of candidate solutions
-that randomly combine options, modeling the packet fields. They
-showed that they were able to generate rules without false negatives
-and only 3 false positives across 33K benign packets. The key
-differences between our work and Vollmer~\etal{}'s are 1) the support
-of different protocols and 2) the use of ranking to discriminate
-alternative plausible rules. As for 1, our approach supports a variety
-of protocols and options (\eg{}, options that refer to message
-payloads). The icmp rules they generate only refer to message fields.
-\MyComment{\Luc{eh possivel incluir contents em regras icmp tb. o q a
-    secao 2.3 diz q \textbf{geralmente} as regras nao-http possuem
-    apenas field options}. } As for 2, our solution additionally ranks
-plausible options, which contain \emph{no} false positives or
-negatives with respect to the benign and malicious traffic
+that randomly combine options, modelling the packet fields. They were
+able to generate rules without false negatives and only 3 false
+positives across 33K benign packets. The key differences between our
+work and Vollmer~\etal{}'s are 1) the support of different protocols
+and 2) the use of ranking to discriminate alternative plausible
+rules. As for 1, our approach supports a variety of protocols and
+options (\eg{}, options that refer to message payloads). The icmp
+rules they generate only refer to message fields.  \MyComment{\Luc{eh
+    possivel incluir contents em regras icmp tb. o q a secao 2.3 diz q
+    \textbf{geralmente} as regras nao-http possuem apenas field
+    options}. } As for 2, our solution additionally ranks plausible
+options, which contain \emph{no} false positives or negatives with
+respect to the benign and mallicious traffic
 considered. Guruprasad~\etal{}~\cite{guruprasad} also used an

### 2021-02-19 paper/old/paper.tex
organizing files

--- 2020-06-06 paper/paper.tex
@@
 %\title{\tname{} for NIDS Rule Synthesis}
-\title{\tname{}: Intrusion Detection Rule Synthesis}  
+\title{\tname{}: Synthesis of Rules for Intrusion Detectors}  
 %\title{Rule Synthesis for Network Intrusion Detectors}
@@
 http rule that captures traffic containing the string
-\CodeIn{"/HtmlAdaptor"} in the payload.
+\CodeIn{``/HtmlAdaptor''} in the payload.
 Figure~\ref{fig:distribution-contents} shows the distribution of
@@
     \midrule
-    %%Apache Struts 2 & \cite{wrccdc} & RCE & http & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
-    ColdFusion \MyComment{PESC} & \cite{nikto} & PESC & http \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
-    iisadmin access\MyComment{web-application-attack} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
-    JBoss JMX\MyComment{ Console Beanshell Exploit} & \cite{nikto} & RCE & http \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
-    scripttag\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
-    WeBid LFI \MyComment{Local File Inclusion} & \cite{nikto} & LFI & http & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
-    %%WordPress Login & \cite{wrccdc} & PVE & http & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
-    .htaccess access\MyComment{attempted-recon} & \cite{nikto} & AREC & http \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
-    .idq access\MyComment{EXPLOIT ISAPI} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
-    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & http \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
-    Basebuilder LFI & \cite{nikto} & LFI & http & \textbf{content}:"GET"; http\_method; \textbf{content}:"/main.inc.php?"; nocase; http\_uri; \textbf{content}:"mj\_config[src\_path]="; nocase; http\_uri; \\
-    Tomcat Snoop Access & \cite{nikto} & AREC & http & \textbf{content}:"/jsp/snp/"; http\_uri; \textbf{content}:".snp"; http\_uri;\\  
+    %%Apache Struts 2 & \cite{wrccdc} & RCE & http & \textbf{content}:``POST''; http\_method; \textbf{content}:``java.lang.ProcessBuilder''; nocase; http\_client\_body; fast\_pattern; \textbf{content}:``/struts2-rest-showcase/orders/3''; http\_uri; \\
+    ColdFusion \MyComment{PESC} & \cite{nikto} & PESC & http \MyComment{& Privilege Escalation} & \textbf{content}:``GET''; http\_method; nocase; \textbf{content}:``/CFIDE/administrator''; http\_uri; nocase; \\
+    iisadmin access\MyComment{web-application-attack} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:``/iisadmin''; nocase; http\_uri; \\
+    JBoss JMX\MyComment{ Console Beanshell Exploit} & \cite{nikto} & RCE & http \MyComment{& Remote Code Execution} & \textbf{content}:``/HtmlAdaptor''; nocase; http\_uri; \textbf{content}:``action=inspect''; nocase; http\_uri; \textbf{content}:``bean''; nocase; http\_uri; \textbf{content}:``name=''; http\_uri; \\
+    scripttag\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:``</script>''; nocase; http\_uri; \\
+    WeBid LFI \MyComment{Local File Inclusion} & \cite{nikto} & LFI & http & \textbf{content}:``GET''; depth:4; \textbf{content}:``/cron.php?''; http\_uri; nocase; \textbf{content}:``include\_path=''; http\_uri; nocase; \textbf{content}:``''../''; \\
+    %%WordPress Login & \cite{wrccdc} & PVE & http & \textbf{content}:``log=''; http\_client\_body; \textbf{content}:``''\&pwd=''; http\_client\_body; \textbf{content}:``''\&wp-submit=''; http\_client\_body; \\
+    .htaccess access\MyComment{attempted-recon} & \cite{nikto} & AREC & http \MyComment{& Adicionar} & \textbf{content}:``.htaccess''; nocase; http\_uri; \\
+    .idq access\MyComment{EXPLOIT ISAPI} & \cite{nikto} & RCE & http \MyComment{& Adicionar} & \textbf{content}:``.idq''; nocase; http\_uri; \\
+    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & http \MyComment{& Adicionar} & \textbf{content}:``/system32/''; nocase; http\_uri; \\
+    Basebuilder LFI & \cite{nikto} & LFI & http & \textbf{content}:``GET''; http\_method; \textbf{content}:``/main.inc.php?''; nocase; http\_uri; \textbf{content}:``mj\_config[src\_path]=''; nocase; http\_uri; \\
+    Tomcat Snoop Access & \cite{nikto} & AREC & http & \textbf{content}:``/jsp/snp/''; http\_uri; \textbf{content}:``.snp''; http\_uri;\\  
     \bottomrule

### 2021-02-19 paper/old/R/coldfusion/subject.r
deleting old material

@@ -1,21 +1,0 @@
-#!/usr/bin/env Rscript
-
-pdf("coldfusion.pdf", width = 6.5, height = 3)
-
-
-df1 <- read.table("data0.txt", header = FALSE)
-df2 <- read.table("data1.txt", header = FALSE)
-df3 <- read.table("data100.txt", header = FALSE)
-df4 <- read.table("data10000.txt", header = FALSE)
-df5 <- read.table("data100000.txt", header = FALSE)
-
-# adjusting margins
-par(mar = c(5, 4, 4, 0), 
-    mgp = c(1, 1, 0),
-    mai = c(0.4, 0.6, 0.1, 0),
-    mfrow = c(1,2))
-
-    bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,names=c("0","1","100","10k","100k"), boxwex=0.2, cex.axis=1, las=1)
-
-dev.off()
-

### 2021-02-19 paper/old/R/idqaccess/subject.r

@@ -1,22 +1,0 @@
-#!/usr/bin/env Rscript
-
-pdf("idqaccess.pdf", width = 6.5, height = 2.9)
-
-df1 <- read.table("data1.txt", header = FALSE)
-df2 <- read.table("data2.txt", header = FALSE)
-df3 <- read.table("data3.txt", header = FALSE)
-df4 <- read.table("data4.txt", header = FALSE)
-df5 <- read.table("data5.txt", header = FALSE)
-df6 <- read.table("data10.txt", header = FALSE)
-df7 <- read.table("data14.txt", header = FALSE)
-
-# adjusting margins
-par(mar = c(5, 4, 4, 0), 
-    mgp = c(1, 1, 0),
-    mai = c(0.4, 0.6, 0.1, 0),
-    mfrow = c(1,2))
-
-bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,df6$V1,df7$V1,names=c("1","2","3","4","5","10","14"), boxwex=0.2, cex.axis=1, las=1)
-
-dev.off()
-

### 2021-02-19 paper/old/R/inc/subject.r

@@ -1,21 +1,0 @@
-#!/usr/bin/env Rscript
-
-pdf("inc.pdf", width = 6.5, height = 3)
-
-
-df1 <- read.table("data0.txt", header = FALSE)
-df2 <- read.table("data1.txt", header = FALSE)
-df3 <- read.table("data100.txt", header = FALSE)
-df4 <- read.table("data10000.txt", header = FALSE)
-df5 <- read.table("data100000.txt", header = FALSE)
-
-# adjusting margins
-par(mar = c(5, 4, 4, 0), 
-    mgp = c(1, 1, 0),
-    mai = c(0.4, 0.6, 0.1, 0),
-    mfrow = c(1,2))
-
-bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,names=c("0","1","100","10k","100k"), boxwex=0.2, cex.axis=1, las=1)
-
-dev.off()
-

### 2021-02-19 paper/old/R/jsp/subject.r

@@ -1,21 +1,0 @@
-#!/usr/bin/env Rscript
-
-pdf("jsp.pdf", width = 6.5, height = 3)
-
-
-df1 <- read.table("data0.txt", header = FALSE)
-df2 <- read.table("data1.txt", header = FALSE)
-df3 <- read.table("data100.txt", header = FALSE)
-df4 <- read.table("data10000.txt", header = FALSE)
-df5 <- read.table("data100000.txt", header = FALSE)
-
-# adjusting margins
-par(mar = c(5, 4, 4, 0), 
-    mgp = c(1, 1, 0),
-    mai = c(0.4, 0.6, 0.1, 0),
-    mfrow = c(1,2))
-
-bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,names=c("0","1","100","10k","100k"), boxwex=0.2, cex.axis=1, las=1)
-
-dev.off()
-

### 2021-02-19 paper/old/R/scripttag/subject.r

@@ -1,22 +1,0 @@
-#!/usr/bin/env Rscript
-
-pdf("scripttag.pdf", width = 6.5, height = 2.9)
-
-df1 <- read.table("data1.txt", header = FALSE)
-df2 <- read.table("data2.txt", header = FALSE)
-df3 <- read.table("data3.txt", header = FALSE)
-df4 <- read.table("data4.txt", header = FALSE)
-df5 <- read.table("data5.txt", header = FALSE)
-df6 <- read.table("data10.txt", header = FALSE)
-df7 <- read.table("data23.txt", header = FALSE)
-
-# adjusting margins
-par(mar = c(5, 4, 4, 0), 
-    mgp = c(1, 1, 0),
-    mai = c(0.4, 0.6, 0.1, 0),
-    mfrow = c(1,2))
-
-bp <- boxplot(df1$V1,df2$V1,df3$V1,df4$V1,df5$V1,df6$V1,df7$V1,names=c("1","2","3","4","5","10","23"), boxwex=0.2, cex.axis=1, las=1)
-
-dev.off()
-
